### 2020-04-16 main.tex
init rpo

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber]{biblatex}
\usepackage{eurosym}
\usepackage{subcaption}
\usepackage{accents}
\usepackage[capitalise]{cleveref}

\addbibresource{main.bib}
\graphicspath{{figures/}}

\crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}

\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}

\newcommand{\generation}{g_{n,s,t}}
\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
\newcommand{\capacityGeneration}{G_{n,s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capexGeneration}{c_{n,s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}{o_{n,s}}
\newcommand{\demand}[1][n]{d_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}{K_{n,\ell}}
\newcommand{\mulowergeneration}{\ubar{\mu}_{n,s,t}}
\newcommand{\muuppergeneration}{\bar{\mu}_{n,s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}

\newcommand{\allocatePeer}{A_{n \rightarrow m,t}}
\newcommand{\allocateFlow}{A_{n \rightarrow m,\ell,t}}
\newcommand{\allocateCapexGeneration}{\mathcal{C}^{G}_{m,t}}
\newcommand{\allocateCapexFlow}{\mathcal{C}^{F}_{m,t}}
\newcommand{\allocateOpex}{\mathcal{O}_{m,t}}
\newcommand{\allocateEmissionCost}{\mathcal{E}_{m,t}}

\newcommand{\emission}{e_{n,s}}
\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{TC}}
\newcommand{\normedgeneration}{\tilde{g}_{n,s,t}}
\newcommand{\normeddemand}[1][n]{\tilde{d}_{#1,a,t}}
\newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
\newcommand{\impactcapexflow}{\Phi_{\ell,t}}
\newcommand{\resultsin}[1]{\hspace{20pt} \bot  \hspace{12pt} #1}
\newcommand{\Forall}[1]{\hspace{20pt} \forall \,\, #1}

\begin{document}

\title{From Flow Based Cost Allocation to Locational Marginal Prices}
\author{Fabian Hofmann}

\maketitle

% \begin{abstract}
% The abstract text goes here.
% \end{abstract}

\section*{Linear Energy Modelling and LMP}
We linearly optimize the capacity and dispatch of a simple power system.

\begin{align}
    \underset{\demand, \generation, \capacityGeneration}{\text{max}}
    \left(\utility(\demand) - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
\end{align}

where 
\begin{itemize}
\item[] $\utility$ denotes the utility function per bus $n$, demand type $a$ time step $t$ 
 \item[] $\demand$ denotes the eletric demand 
 \item[] $\capexGeneration$ denotes the capital expenditure (CAPEX) per node $n$ and generator type $s$
 \item[] $\capacityGeneration$ denotes the generation capacity
 \item[] $\opexGeneration$ denotes the operational cost (OPEX)
 \item[] $\generation$ denotes the net generation in MW
 \item[] $\capexFlow$ denotes the CAPEX per transmission line 
 \item[] $\capacityFlow$ denotes the transmission capacity.
\end{itemize}

\noindent
In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. The generation $\generation$ is constraint to its capacity
\begin{align}
 \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration}\Forall{n,s,t} 
 \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
 - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:lower_generation_capacity_constraint}
 \end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness are only non-zero if the corresponding constraint is binding.

\noindent
The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
\begin{align}
 \flow - \capacityFlow \le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
 \label[constraint]{eq:upper_flow_capacity_constraint} \\
 - \flow - \capacityFlow \le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
 \label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. \\

The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_s \generation - \sum_a \demand - \sum_l \incidence \, \flow = 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. \\

As the shown in \cite{brown_decreasing_2020}, the OPEX and CAPEX of generators follow the non-profit rule which states that their expenses equals their total revenue

\begin{align}
\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation = \sum_{t} \lmp \, \generation \Forall{n,s}\label{eq:non_profit_generator}
\end{align}

\noindent
The relation counts for transmission lines where the CAPEX amounts to the total congestion revenue
\begin{align}
\capexFlow \, \capacityFlow = \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
\end{align}
Note that we are neglecting the Kirchhoff Voltage Law (KVL) constraints. Finally, the total cost $\totalcost$, consisting of all CAPEX and OPEX have to be payed by the consumers

\begin{align}
\totalcost = \sum_{n,s} \capexGeneration \, \capacityGeneration + \sum_{n,s,t} \opexGeneration \, \generation + \sum_{\ell} \capexFlow \, \capacityFlow = \sum_{n,a,t} \lmp \, \demand 
\label{eq:total_system_cost}
\end{align}

\subsection*{Adding CO$_2$ Constraints}

Imposing an additional CO$_2$ constraint limiting the total emission to K,  
\begin{align}
 \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
 \label[constraint]{eq:co2_constraint}
\end{align}
with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to

\begin{align}
\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation = \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
\label{eq:non_profit_generator_emission}
\end{align}
This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs

\begin{align}
\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
\end{align}

% ----------------------------------------------------------------------------------------
% ----------------------------------------------------------------------------------------

\section*{Flow Based Cost Allocation}

Besides using LMP, there are others methods for distributing costs in a power system. A common approach is based on Flow Allocation (FA) methods which determine peer-to-peer (P2P) relations between producers (sources) and consumers (sinks) and thus enable to directly assign costs to the consumers. %Due to the non-determnistic tracing of power through the network the methods result in different allocation. 
Let $\allocatePeer$ denotes the power that is produced at bus $n$ at time $t$ and assigned to consumers at bus $m$. We will refer to it as a P2P relation between $n$ and $m$. Further let $\allocateFlow$ denote flow on line $\ell$ at time $t$ caused by the P2P connection $n\rightarrow m$.\\
\noindent
In the following we show how the LMP can be reproduced on the basis of FA.

\subsubsection*{Allocating Operational Expenditures}
Consumers at bus $m$ compensate the OPEX proportionally to the assigned production, thus 
\begin{align}
\allocateOpex = \sum_{n,s} \opexGeneration \, \normedgeneration \, \allocatePeer \Forall{m,t} \label{eq:allocate_opexGeneration}
\end{align}
where $\normedgeneration = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodal production. \\

\subsubsection*{Allocating CO$_2$ Cost}
 Similar to the OPEX (compare \cref{eq:shift_opex_by_emission_cost}) consumers at bus $m$ compensate the CO$_2$ cost, thus 
\begin{align}
 \allocateEmissionCost = \sum_{n,s} \emissionPrice \, \emission \, \normedgeneration \, \allocatePeer \Forall{m,t} \label{eq:allocate_emissionPrice}
\end{align}

\subsubsection*{Allocating Capital Expenditures for Generators}
According to the polluter pays principle, the CAPEX allocation should differentiate between consumers who are `responsible` for investments and those who are not. Thus, it is crucial to know whether a time step $t$ makes investments in $\capacityGeneration$ necessary or if it would be feasible without (further) investments. The KKT variable of the upper generator capacity constraint $\muuppergeneration$ gives us the needed estimator: If $\muuppergeneration > 0$, the upper capacity constraint \cref{eq:upper_generation_capacity_constraint} is binding and it can be fairly assumed that these times steps cause investments in $\capacityGeneration$. If in contrast $\muuppergeneration = 0$, the generation $\generation$ is not bound and does not push further investments in $\capacityGeneration$.
On the basis of the KKT variable $\muuppergeneration$ we are able to define a measure $\impactcapexgeneration$ for the impact on the investments in $\capacityGeneration$  
\begin{align}
    \impactcapexgeneration = \dfrac{\muuppergeneration}{\sum_t \muuppergeneration \, \generation}
\end{align}
which weights the contributions of consumers at bus $m$ at time $t$ to the CAPEX 
\begin{align}
    \allocateCapexGeneration = \sum_{n,s} \capexGeneration \, \capacityGeneration \, \impactcapexgeneration\, \allocatePeer \Forall{m,t} \label{eq:allocate_capexGeneration}
\end{align}

\noindent
If consumers at $m$ retrieve power from $n$ when $\impactcapexgeneration$ is large, their contribution to investments in $\capacityGeneration$ raises. On the other hand, if consumers at $m$ retrieve power from $n$ when $\impactcapexgeneration=0$ , their contribution to $\capacityGeneration$ remain unchanged.

\noindent
\textit{Side note}: A way more simple approach would neglect these thoughts and let  consumers at $m$ compensate CAPEX for generator $s$ at bus $n$ proportional to their retrieved power. This would lead us to $ \impactcapexgeneration = \left( \generation\right)^{-1}$.

\subsubsection*{Allocating Capital Expenditures for Transmission Lines}
The allocation of CAPEX per the transmission line $\capexFlow \, \capacityFlow $ follows a similar procedure as for the generators. Again, we build up on KKT variables $\muupperflow$ and $\mulowerflow$ which translate to the necessity of transmission expansion of $\ell$ at time $t$. The measure $\impactcapexflow$ respects the need of CAPEX of both flow directions when set to

\begin{align}
    \impactcapexflow = \dfrac{\muupperflow + \mulowerflow}{\sum_t \left( \muupperflow + \mulowerflow \right) \, \flow}
\end{align}

and weights the contributions of consumers at bus $m$ at time $t$ to the CAPEX in line $\ell$,
\begin{align}
    \allocateCapexFlow = \sum_{n,\ell} \capexFlow \, \capacityFlow\, \impactcapexflow, \allocateFlow \Forall{m,t} \label{eq:allocate_capexFlow}
\end{align}

\subsubsection*{Reproducing the LMP Based Consumption Costs}
The above presented quantities complete an allocation of the total cost $\totalcost$ which mirrors the cost allocation based on LMP, thus

\begin{align}
 \allocateOpex + \allocateEmissionCost + \allocateCapexGeneration + \allocateCapexFlow \simeq \lmp[m] \, \demand[m] \Forall{m,t} 
 \label{eq:allocated_cost_mirror_lmp}
\end{align}

\subsection*{play around}

\section*{Showcase}

\subsection*{Network with CO$_2$ constraint}
We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}. It consists of nine buses and ten time steps. The solver optimizes the capacity of two generators, wind and gas, per bus. ...

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{compare_allocation.png}
    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constaint. The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP. }
    \label{fig:cost_allocation}
\end{figure}

\begin{figure}[h]
\begin{subfigure}{.5\textwidth}
\centering
 \includegraphics[width=\textwidth]{network.png}
 \caption{}
 \label{fig:network}
\end{subfigure}
\begin{subfigure}{.5\textwidth}   
    \centering
    \includegraphics[width=\textwidth]{nodal_payments.png}
    \caption{}
    \label{fig:nodal_payments}
\end{subfigure}
\caption{Network used for showcasing. (a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$. (b) shows the total nodal payments according to the cost allocation.}
\end{figure}

\newpage
\subsubsection*{Relaxed CO$_2$ Constraint}

\begin{figure}[h]
\begin{subfigure}{.5\textwidth}
\centering
 \includegraphics[width=\textwidth]{network_relaxed_co2.png}
 \caption{}
 \label{fig:network_relaxed_co2}
\end{subfigure}
\begin{subfigure}{.5\textwidth}   
    \centering
    \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
    \caption{}
    \label{fig:nodal_payments_relaxed_co2}
\end{subfigure}
\caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}. Only one time-step $t=6$ determines the allocation of generator CAPEX $\allocateCapexGeneration$, as for all other time-steps \cref{eq:upper_generation_capacity_constraint} is not binding. Again note the cost scale difference between time step 6 and all others.}
    \label{fig:cost_allocation}
\end{figure}

% \subsection*{How does the cost flow through the network}
% 
% \begin{figure}[h]
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[width=\textwidth]{opex_flow.png}
%       \label{fig:opex_flow}
%     \end{subfigure}%
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[width=\textwidth]{capex_flow.png}
%       \label{fig:capex_flow}
%     \end{subfigure}
%     \caption{}
%     \label{fig:fig}
% \end{figure}

% \begin{appendix}
% Since the P2P comprises all net production and net consumption summing over all sources must result in the nodal demand  
% 
% \begin{align}
%  \sum_{n} \allocatePeer = \sum_a \demand[m]
% \end{align}
% as well as summing over all sinks results in the nodal generation
% \begin{align}
%  \sum_{m} \allocatePeer = \sum_s \generation
% \end{align}
% 
% The P2P relation can be further broke down to allocating generator specific $ \normedgeneration \, \allocatePeer $ where $\normedgeneration = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production. Similarly the allocation to consumers a are given by and consumer specific $ \normeddemand[m] \allocatePeer$ with the nodal comsumer share given by $\normeddemand = \demand/\sum_a \demand$.
% \end{appendix}

% \bibliographystyle{ieeetr}
% \printbibliography

\end{document}

### 2020-04-16 dist_cost_large.py

@@ -1,0 +1,139 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb 20 14:13:56 2020
+
+@author: fabian
+"""
+
+import pypsa
+import pandas as pd
+import numpy as np
+import netallocation as ntl
+from  netallocation.cost import (nodal_co2_cost, nodal_demand_cost,
+                                 nodal_production_revenue, congestion_revenue)
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+plt.rc('text', usetex=True)
+
+
+to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{m,t}$',
+                      branch_investment_cost = '$\mathcal{C}^{F}_{m,t}$',
+                      one_port_operational_cost = '$\mathcal{O}_{m,t}$',
+                      co2_cost = '$\mathcal{E}_{m,t}$',
+                      nodal_demand_cost = '$\sum_a \, \lambda_{m,t} \, d_{m,a,t}$')
+
+
+n = pypsa.Network('elec_s_50_ec_lvopt_Co2L-3H.nc')
+n.calculate_dependent_values()
+n.generators['p_nom_extendable'] = True
+n.lines['s_nom_min'] = 0
+n.links['p_nom_min'] = 0
+n.generators['p_nom_max'] = np.inf
+n.mremove('StorageUnit', n.storage_units.index)
+n.mremove('Line', n.lines.query('x_pu == inf').index)
+
+T = 10
+n.set_snapshots(n.snapshots[:T])
+n.global_constraints.loc['CO2Limit', 'constant'] *= T/2800
+n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
+tag = '_large'
+
+# %% Network plot
+
+fig, ax = plt.subplots(subplot_kw={'projection': ccrs.PlateCarree()}, figsize=(6,5))
+bus_sizes = n.generators.groupby(['bus', 'carrier']).p_nom_opt.sum()
+plot = n.plot(bus_sizes=bus_sizes/1e5, line_widths=n.lines.s_nom_opt/1e3,
+       link_widths=n.links.p_nom_opt/1e3, ax=ax)
+ax.legend(*ntl.plot.handles_labels_for(n.carriers.color), loc='center left',
+          bbox_to_anchor=(1, 0.5), ncol=1)
+fig.canvas.draw(); fig.tight_layout()
[…]

### 2020-04-16 dist_costs.py

@@ -1,0 +1,141 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb 20 14:13:56 2020
+
+@author: fabian
+"""
+
+import pypsa
+import pandas as pd
+import numpy as np
+import netallocation as ntl
+from  netallocation.cost import (nodal_co2_cost, nodal_demand_cost,
+                                 nodal_production_revenue, congestion_revenue)
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+plt.rc('text', usetex=True)
+
+
+to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{m,t}$',
+                      branch_investment_cost = '$\mathcal{C}^{F}_{m,t}$',
+                      one_port_operational_cost = '$\mathcal{O}_{m,t}$',
+                      co2_cost = '$\mathcal{E}_{m,t}$',
+                      nodal_demand_cost = '$\sum_a \, \lambda_{m,t} \, d_{m,a,t}$')
+
+
+n = ntl.test.get_network_ac_dc()
+n.carriers = n.carriers.drop('battery')
+n.generators['p_nom_min'] = 0
+
+tag = ''
+relax_co2 = False
+if relax_co2:
+    n.global_constraints = n.global_constraints.drop('co2_limit')
+    tag = '_relaxed_co2'
+
+sns = n.snapshots
+n.lopf(pyomo=False, keep_shadowprices=True)
+n.carriers.loc['wind', 'color'] = 'steelblue'
+n.buses.loc[n.buses.index.str.contains('DC'), 'y'] += 1
+n.buses.loc[n.buses.index.str.contains('DC'), 'x'] -= .5
+
+# %% Network plot
+
+fig, ax = plt.subplots(subplot_kw={'projection': ccrs.PlateCarree()}, figsize=(6,5))
+bus_sizes = n.generators.groupby(['bus', 'carrier']).p_nom_opt.sum()
+plot = n.plot(bus_sizes=bus_sizes/1e4, line_widths=n.lines.s_nom_opt/1e2,
+       link_widths=n.links.p_nom_opt/1e2, ax=ax)
+ntl.plot.annotate_bus_names(n, ax, shift=(0,0.3),
+                            transform=ccrs.PlateCarree(),
+                            bbox=dict(facecolor='white', alpha=0.5, edgecolor='None'))
[…]

### 2020-04-26 main.tex
add theory part

@@
 \usepackage{eurosym}
+\usepackage[dvipsnames]{xcolor} 
 \usepackage{subcaption}
@@
 \newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
+\newcommand{\note}[1]{\textcolor{Orange}{#1}}
 
@@
 \newcommand{\capexFlow}{c_{\ell}}
-\newcommand{\opexGeneration}{o_{n,s}}
+\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
 \newcommand{\demand}[1][n]{d_{#1,a,t}}
 \newcommand{\utility}{U_{n,a,t}}
-\newcommand{\incidence}{K_{n,\ell}}
-\newcommand{\mulowergeneration}{\ubar{\mu}_{n,s,t}}
-\newcommand{\muuppergeneration}{\bar{\mu}_{n,s,t}}
+\newcommand{\incidence}[1][n]{K_{#1,\ell}}
+\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
+\newcommand{\slackflow}{k_{\ell}}
+\newcommand{\slack}[1][n]{k_{#1}}
+\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
+\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
 \newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
@@
 \newcommand{\flow}{f_{\ell,t}}
+\newcommand{\cycle}{C_{\ell,c}}
+\newcommand{\impedance}{x_\ell}
+\newcommand{\cycleprice}{\lambda_{c,t}}
+\newcommand{\injection}{p_{n,t}}
+\newcommand{\netconsumption}{p^{-}_{n,t}}
+\newcommand{\netproduction}{p^{+}_{n,t}}
+\newcommand{\selfconsumption}{p^{\circ}_{n,t}}
+\newcommand{\lagrangian}{\mathcal{L}}
 
-\newcommand{\allocatePeer}{A_{n \rightarrow m,t}}
-\newcommand{\allocateFlow}{A_{n \rightarrow m,\ell,t}}
+\newcommand{\allocatePeer}[1][n \rightarrow m]{A_{#1,t}}
+\newcommand{\allocateTransaction}[1][n \rightarrow m]{A_{#1,\ell,t}}
+\newcommand{\allocatePeerNormed}[1][n \rightarrow m]{\tilde{A}_{#1,t}}
 \newcommand{\allocateCapexGeneration}{\mathcal{C}^{G}_{m,t}}
@@
 \newcommand{\totalcost}{\mathcal{TC}}
-\newcommand{\normedgeneration}{\tilde{g}_{n,s,t}}
+\newcommand{\generatorshare}{\tilde{g}_{n,s,t}}
 \newcommand{\normeddemand}[1][n]{\tilde{d}_{#1,a,t}}
@@
 \newcommand{\impactcapexflow}{\Phi_{\ell,t}}
-\newcommand{\resultsin}[1]{\hspace{20pt} \bot  \hspace{12pt} #1}
-\newcommand{\Forall}[1]{\hspace{20pt} \forall \,\, #1}
+
+%math 
+\newcommand{\resultsin}[1]{\hspace{12pt} \bot  \hspace{12pt} #1}
+\newcommand{\Forall}[1]{\hspace{20pt} \forall \,\, #1 }
+\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
 
@@
 
-\title{From Flow Based Cost Allocation to Locational Marginal Prices}
+\title{From Linear Optimization to Transmission Cost Allocation}
 \author{Fabian Hofmann}
@@
 \begin{align}
- \flow - \capacityFlow \le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
+ \flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
  \label[constraint]{eq:upper_flow_capacity_constraint} \\
- - \flow - \capacityFlow \le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
+ - \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
  \label[constraint]{eq:lower_flow_capacity_constraint}
@@
 \begin{align}
-    \sum_s \generation - \sum_a \demand - \sum_l \incidence \, \flow = 0 \resultsin{\lmp} \Forall{n,t}
+    \sum_l \incidence \, \flow - \sum_s \generation + \sum_a \demand &= 0 \resultsin{\lmp} \Forall{n,t}
     \label[constraint]{eq:nodal_balance_lin}
@@
 \begin{align}
-\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation = \sum_{t} \lmp \, \generation \Forall{n,s}\label{eq:non_profit_generator}
+\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \lmp \, \generation \Forall{n,s}\label{eq:non_profit_generator}
 \end{align}
@@
 \begin{align}
-\capexFlow \, \capacityFlow = \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
+\capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
 \end{align}
+Power flow from nodes with low locational prices to nodes with higher prices in the network, thus the nodal price difference in flow direction is negative.
+\note{Different to Tom's paper, there a minus is missing.}
 Note that we are neglecting the Kirchhoff Voltage Law (KVL) constraints. Finally, the total cost $\totalcost$, consisting of all CAPEX and OPEX have to be payed by the consumers
@@
 \begin{align}
-\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation = \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
+\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
 \label{eq:non_profit_generator_emission}
@@
 We will refer to it as a P2P relation between $n$ and $m$.
-Further let $\allocateFlow$ denote flow on line $\ell$ at time $t$ caused by the P2P connection $n\rightarrow m$.\\
+Further let $\allocateTransaction$ denote flow on line $\ell$ at time $t$ caused by the P2P connection $n\rightarrow m$.\\
 \noindent
@@
 \begin{align}
-\allocateOpex = \sum_{n,s} \opexGeneration \, \normedgeneration \, \allocatePeer \Forall{m,t} \label{eq:allocate_opexGeneration}
+\allocateOpex &= \sum_{n,s} \opexGeneration \, \generatorshare \, \allocatePeer \Forall{m,t} \label{eq:allocate_opexGeneration}
 \end{align}
-where $\normedgeneration = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodal production. \\
+where $\generatorshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodal production. \\
 
@@
 \begin{align}
- \allocateEmissionCost = \sum_{n,s} \emissionPrice \, \emission \, \normedgeneration \, \allocatePeer \Forall{m,t} \label{eq:allocate_emissionPrice}
+ \allocateEmissionCost &= \sum_{n,s} \emissionPrice \, \emission \, \generatorshare \, \allocatePeer \Forall{m,t} \label{eq:allocate_emissionPrice}
 \end{align}
@@
 \begin{align}
-    \allocateCapexGeneration = \sum_{n,s} \capexGeneration \, \capacityGeneration \, \impactcapexgeneration\, \allocatePeer \Forall{m,t} \label{eq:allocate_capexGeneration}
+    \allocateCapexGeneration &= \sum_{n,s} \capexGeneration \, \capacityGeneration \, \impactcapexgeneration\, \allocatePeer \Forall{m,t} \label{eq:allocate_capexGeneration}
 \end{align}
@@
 \begin{align}
-    \allocateCapexFlow = \sum_{n,\ell} \capexFlow \, \capacityFlow\, \impactcapexflow, \allocateFlow \Forall{m,t} \label{eq:allocate_capexFlow}
+    \allocateCapexFlow = \sum_{n,\ell} \capexFlow \, \capacityFlow\, \impactcapexflow, \allocateTransaction \Forall{m,t} \label{eq:allocate_capexFlow}
 \end{align}
@@
 
-\subsection*{play around}
+\newpage
 
-\section*{Showcase}
+\section*{Theory}
+\subsubsection*{Lagrange Mutliplier}
 
-\subsection*{Network with CO$_2$ constraint}
-We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
+Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
+
+\begin{align}
+ \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
+\end{align}
+...
+
+\subsection*{Full Lagrangian}
+\begin{align}
+ \lagrangian\left(\generation, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
+ &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
+%  &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
+ &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
+ &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
+\end{align}
+% 
+where $\boldsymbol{\lambda} = \left\lbrace \lmp \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
+
+In this formulation, we treat the flow $\flow$ an a passive quantity only.
+The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one net power production at bus $n$.
+Thus with a fix demand $\demand$, they links the generation $\generation$ to the flow on each line according to
+\begin{align}
+ \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  %= \sum_n \ptdf \injection 
+\end{align}
+The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
+It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
+The choice of slack modifies the PTDF accordingly
+\begin{align}
+ \ptdf  \rightarrow \ptdf - \sum_m \ptdf[m] \, \slack[m]
+ \label[relation]{eq:ptdf_slacked}
+\end{align}
+When bus $n$ injects excess power, is has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
+Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack
+\begin{align}
+\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
+\label{eq:slack}
+\end{align}
+where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
+% Sensitivity of flow $\flow$ against changes of the power production $\generation$  
+% \begin{align}
+%  \pdv{\flow}{\generation}  = \ptdf \label{eq:flow_sensitity} 
+% \end{align}
+The global maximum of the Lagrangian requires stationarity with respect to all variables. Using stationarity of the generation    
+\begin{align}
+ \pdv{\lagrangian}{\generation} = 0  \Forall{n,s} \\
+\end{align}
+and solving for $\lmp$, leads to
+
+% \begin{align}
+% - \opexGeneration + \lmp - \sum_{m,\ell} \lmp[m] \, \incidence[m] \, \ptdf  - \muuppergeneration + \mulowergeneration - \sum_{\ell} \muupperflow \ptdf + \sum_{\ell} \mulowerflow \, \ptdf = 0 \Forall{n,s}
+% \end{align}
+
+\begin{align}
+\lmp &= \opexGeneration  + \muuppergeneration - \mulowergeneration + \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf + \sum_{m,\ell}\lmp[m] \, \incidence[m] \, \ptdf   \Forall{n,s} \label{eq:lmp_generalized}
+% &= \opexGeneration  + \muuppergeneration - \mulowergeneration + \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf + \sum_{m}\lmp[m] \left( \delta_{m,n} - \slack[m] \right) &\Forall{n,s} \label{eq:lmp_generalized}
+\end{align}
+% 
+\Cref{eq:lmp_generalized} must hold for every choice of slack in the PTDF.
+According to \cref{eq:ptdf_slacked,eq:slack}, setting the slack to $\slack = \delta_{m,n}$ results in $\ptdf = \sum_\ell \incidence[m] \, \ptdf = 0$.
+This leads to our first represantion for Locational Market Price, which we will refer to as the ``Island Solution``.
+Accordingly, the LMP is directly determined by the local operational price and prices for the generation capacity constraint
+\begin{align}
+\lmp  =  \opexGeneration + \muuppergeneration - \mulowergeneration \Forall{n,s,t}
+\label{eq:lmp1}
+\end{align}
+However, using this Island Solution, we can derive a second representation of the LMP. Feeding the \cref{eq:lmp1} back into \cref{eq:lmp_generalized} and applying \cref{eq:slack}, leads to 
+\begin{align}
+\lmp =  - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf + \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \Forall{n,s,t} 
+\end{align}
+This equation generalizes \cref{eq:lmp1}.
+Note, that setting again $\slack = \delta_{m,n}$ will result in \cref{eq:lmp1}.
+It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints at each generator and each transmission line in the system.
+According to the slack $\slack$,  the possibility for allocating costs between buses and transmission components.
+
+\begin{align}
+ \sum_m \allocateTransaction[m \rightarrow n] = - \demand \, \ptdf  
+\end{align}
+\begin{align}
+ \allocatePeer[m \rightarrow n] = \demand[m] \, \slack[m] 
+\end{align}
+
+% The power injection per node and time denotes the nodal power imbalance between generation and demand
+% 
+% \begin{align}
+%  \injection = \sum_s \generation - \sum_a \demand
+% \end{align}
+% The net production is the amount of positive injection per node and time 
+% \begin{align}
+%  \netproduction = \text{min}\left( \injection, 0 \right) 
+% \end{align}
+% The nodal net consumption denotes the amount of negative injection 
+% \begin{align}
+%  \netconsumption = \text{min}\left( - \injection, 0 \right) 
+% \end{align}
+% The nodal self-consumption equals the minimum of the nodal production and demand. 
+% \begin{align}
+%  \selfconsumption = \text{min}\left( \netproduction, \netconsumption \right) 
+% \end{align}
+
+% \section*{Showcase}
+% 
+% \subsection*{Network with CO$_2$ constraint}
+% We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
 It consists of nine buses and ten time steps.
@@
 ...
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{compare_allocation.png}
-    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constaint.
+% 
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\textwidth]{compare_allocation.png}
+%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constraint.
 The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
 }
-    \label{fig:cost_allocation}
-\end{figure}
-
-\begin{figure}[h]
-\begin{subfigure}{.5\textwidth}
-\centering
- \includegraphics[width=\textwidth]{network.png}
- \caption{}
- \label{fig:network}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}   
-    \centering
-    \includegraphics[width=\textwidth]{nodal_payments.png}
-    \caption{}
-    \label{fig:nodal_payments}
-\end{subfigure}
-\caption{Network used for showcasing.
+%     \label{fig:cost_allocation}
+% \end{figure}
+% 
+% \begin{figure}[h]
+% \begin{subfigure}{.5\textwidth}
+% \centering
+%  \includegraphics[width=\textwidth]{network.png}
+%  \caption{}
+%  \label{fig:network}
+% \end{subfigure}
+% \begin{subfigure}{.5\textwidth}   
+%     \centering
+%     \includegraphics[width=\textwidth]{nodal_payments.png}
+%     \caption{}
+%     \label{fig:nodal_payments}
+% \end{subfigure}
+% \caption{Network used for showcasing.
 (a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$.
 (b) shows the total nodal payments according to the cost allocation.}
-\end{figure}
-
-\newpage
-\subsubsection*{Relaxed CO$_2$ Constraint}
-
-\begin{figure}[h]
-\begin{subfigure}{.5\textwidth}
-\centering
- \includegraphics[width=\textwidth]{network_relaxed_co2.png}
- \caption{}
- \label{fig:network_relaxed_co2}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}   
-    \centering
-    \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
-    \caption{}
-    \label{fig:nodal_payments_relaxed_co2}
-\end{subfigure}
-\caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
-    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}.
+% \end{figure}
+% 
+% 
+% 
+% 
+% 
+% \newpage
+% \subsubsection*{Relaxed CO$_2$ Constraint}
+% 
+% 
+% \begin{figure}[h]
+% \begin{subfigure}{.5\textwidth}
+% \centering
+%  \includegraphics[width=\textwidth]{network_relaxed_co2.png}
+%  \caption{}
+%  \label{fig:network_relaxed_co2}
+% \end{subfigure}
+% \begin{subfigure}{.5\textwidth}   
+%     \centering
+%     \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
+%     \caption{}
+%     \label{fig:nodal_payments_relaxed_co2}
+% \end{subfigure}
+% \caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
+% \end{figure}
+% 
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
+%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}.
 Only one time-step $t=6$ determines the allocation of generator CAPEX $\allocateCapexGeneration$, as for all other time-steps \cref{eq:upper_generation_capacity_constraint} is not binding.
 Again note the cost scale difference between time step 6 and all others.}
-    \label{fig:cost_allocation}
-\end{figure}
-
+%     \label{fig:cost_allocation_relaxed_co2}
+% \end{figure}
+% 
 % \subsection*{How does the cost flow through the network}
@@
 % 
-% The P2P relation can be further broke down to allocating generator specific $ \normedgeneration \, \allocatePeer $ where $\normedgeneration = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production.
+% The P2P relation can be further broke down to allocating generator specific $ \generatorshare \, \allocatePeer $ where $\generatorshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production.
 Similarly the allocation to consumers a are given by and consumer specific $ \normeddemand[m] \allocatePeer$ with the nodal comsumer share given by $\normeddemand = \demand/\sum_a \demand$.

### 2020-04-26 dist_cost_large.py

@@ -26,71 +26,132 @@ to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{m,t}$',
 
 n = pypsa.Network('elec_s_50_ec_lvopt_Co2L-3H.nc')
 n.calculate_dependent_values()
+n.snapshot_weightings[:] = 1
 n.generators['p_nom_extendable'] = True
 n.lines['s_nom_min'] = 0
-n.links['p_nom_min'] = 0
 n.generators['p_nom_max'] = np.inf
 n.mremove('StorageUnit', n.storage_units.index)
+n.mremove('Link', n.links.index)
 n.mremove('Line', n.lines.query('x_pu == inf').index)
+n.remove('Bus', 'DE0 0')
 
 T = 10
 n.set_snapshots(n.snapshots[:T])
-n.global_constraints.loc['CO2Limit', 'constant'] *= T/2800
+n.global_constraints.loc['CO2Limit', 'constant'] *= T/8700
 n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
 tag = '_large'
 
+# %% reproduce analytic
+
+from netallocation.cost import locational_market_price
+from netallocation.grid import PTDF, Incidence, power_demand, network_injection
+from netallocation.linalg import dedup_axis, dot
+from numpy import eye
+
+norm = lambda ds: ds/ds.sum()
+
+m = 'DE0 9'
+carrier = 'solar'
+s = m + ' ' + carrier
+t = n.snapshots[2]
+
+slack = norm(network_injection(n, t).clip(max=0))
+k = PTDF(n) @ slack
+H = PTDF(n) - k
+K = Incidence(n)
+y = locational_market_price(n, t)
+lhs = - (y @ K @ H.sel(bus=m))
+
+o = n.generators.marginal_cost[s] + ntl.cost.nodal_co2_price(n, t).loc[m, carrier]
+mu_up = n.generators_t.mu_upper.loc[t, s]
+mu_lo = n.generators_t.mu_lower.loc[t, s]
+line_mu_up = n.lines_t.mu_upper.loc[t]
+line_mu_lo = n.lines_t.mu_lower.loc[t]
+rhs  = ((line_mu_up - line_mu_lo) @ H.sel(bus=m))
+
+print(lhs.item(), rhs.item())
+
+lhs = - y.sel(bus=m) * K.sel(bus=m) @ H.sel(bus=m)
[…]

### 2020-04-29 main.tex
further work

@@
 
+\setlength\parindent{8pt}
+
 \newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
@@
 
-\newcommand{\generation}{g_{n,s,t}}
+\newcommand{\generation}[1][n]{g_{#1,s,t}}
 \newcommand{\generationpotential}{\bar{g}_{n,s,t}}
+\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
 \newcommand{\capacityGeneration}{G_{n,s}}
@@
 \newcommand{\demand}[1][n]{d_{#1,a,t}}
+\newcommand{\demandnodal}[1][n]{d_{#1,t}}
+\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
 \newcommand{\utility}{U_{n,a,t}}
@@
 \newcommand{\ptdf}[1][n]{H_{\ell,#1}}
+\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
 \newcommand{\slackflow}{k_{\ell}}
 \newcommand{\slack}[1][n]{k_{#1}}
+\newcommand{\Slack}{k_{m,n}}
 \newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
@@
 
-\newcommand{\allocatePeer}[1][n \rightarrow m]{A_{#1,t}}
-\newcommand{\allocateTransaction}[1][n \rightarrow m]{A_{#1,\ell,t}}
-\newcommand{\allocatePeerNormed}[1][n \rightarrow m]{\tilde{A}_{#1,t}}
-\newcommand{\allocateCapexGeneration}{\mathcal{C}^{G}_{m,t}}
-\newcommand{\allocateCapexFlow}{\mathcal{C}^{F}_{m,t}}
-\newcommand{\allocateOpex}{\mathcal{O}_{m,t}}
-\newcommand{\allocateEmissionCost}{\mathcal{E}_{m,t}}
+\newcommand{\allocatePeer}[1][m \rightarrow n]{A_{#1,t}}
+\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
+\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocatePeerNormed}[1][m \rightarrow n]{\tilde{A}_{#1,t}}
+\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
+\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
+\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
+\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}
 
-\newcommand{\emission}{e_{n,s}}
+\newcommand{\emission}[1][n]{e_{#1,s}}
 \newcommand{\emissionPrice}{\mu_{\text{CO2}}}
@@
 \newcommand{\totalcost}{\mathcal{TC}}
-\newcommand{\generatorshare}{\tilde{g}_{n,s,t}}
-\newcommand{\normeddemand}[1][n]{\tilde{d}_{#1,a,t}}
 \newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
@@
 
+\subsubsection*{Lagrange Mutliplier}
+
+Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
+
+\begin{align}
+ \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
+\end{align}
+...
+
 \section*{Linear Energy Modelling and LMP}
+
 We linearly optimize the capacity and dispatch of a simple power system.
@@
 \end{itemize}
+% 
+\subsubsection*{Constraints}
+In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. \\
 
-\noindent
-In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. The generation $\generation$ is constraint to its capacity
+The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
 \begin{align}
- \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration}\Forall{n,s,t} 
+    \sum_l \incidence \, \flow - \sum_s \generation + \sum_a \demand &= 0 \resultsin{\lmp} \Forall{n,t}
+    \label[constraint]{eq:nodal_balance_lin}
+\end{align}
+Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. \\
+
+The generation $\generation$ is constraint to its nominal capacity
+\begin{align}
+ \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
  \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
@@
 where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness are only non-zero if the corresponding constraint is binding.
+The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness,
+\begin{align}
+\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
+\label{eq:complementary_slackness_upper_generation} \\
+\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
+\label{eq:complementary_slackness_lower_generation} 
+\end{align}
+are only non-zero if the corresponding constraint is binding. \\
 
-\noindent
 The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
@@
 \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding.
+The complementary slackness
+\begin{align}
+ \muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
+ \label{eq:complementary_slackness_upper_flow} \\
+ \mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
+ \label{eq:complementary_slackness_lower_flow} 
+\end{align}
+set the respective KKT for flows staying below the thermal limit to zero. 
 \\
 
-The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
+\subsubsection*{Flow is a passive variable}
+We treat the flow $\flow$ an a passive quantity only that is it is only influenced by the nodal injection.
+
+The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit, commonly one MW, of net power production at bus $n$.
+Thus with a fix demand $\demand$, they direclty link the generation $\generation$ to the flow on each line according to
 \begin{align}
-    \sum_l \incidence \, \flow - \sum_s \generation + \sum_a \demand &= 0 \resultsin{\lmp} \Forall{n,t}
-    \label[constraint]{eq:nodal_balance_lin}
-\end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. \\
-
-As the shown in \cite{brown_decreasing_2020}, the OPEX and CAPEX of generators follow the non-profit rule which states that their expenses equals their total revenue
-
-\begin{align}
-\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \lmp \, \generation \Forall{n,s}\label{eq:non_profit_generator}
-\end{align}
-
-\noindent
-The relation counts for transmission lines where the CAPEX amounts to the total congestion revenue
-\begin{align}
-\capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
-\end{align}
-Power flow from nodes with low locational prices to nodes with higher prices in the network, thus the nodal price difference in flow direction is negative.
-\note{Different to Tom's paper, there a minus is missing.}
-Note that we are neglecting the Kirchhoff Voltage Law (KVL) constraints. Finally, the total cost $\totalcost$, consisting of all CAPEX and OPEX have to be payed by the consumers
-
-\begin{align}
-\totalcost = \sum_{n,s} \capexGeneration \, \capacityGeneration + \sum_{n,s,t} \opexGeneration \, \generation + \sum_{\ell} \capexFlow \, \capacityFlow = \sum_{n,a,t} \lmp \, \demand 
-\label{eq:total_system_cost}
-\end{align}
-
-\subsection*{Adding CO$_2$ Constraints}
-
-Imposing an additional CO$_2$ constraint limiting the total emission to K,  
-\begin{align}
- \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
- \label[constraint]{eq:co2_constraint}
-\end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$.
-The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
-
-\begin{align}
-\capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
-\label{eq:non_profit_generator_emission}
-\end{align}
-This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
-
-\begin{align}
-\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
-\end{align}
-
-% ----------------------------------------------------------------------------------------
-% ----------------------------------------------------------------------------------------
-
-\section*{Flow Based Cost Allocation}
-
-Besides using LMP, there are others methods for distributing costs in a power system.
-A common approach is based on Flow Allocation (FA) methods which determine peer-to-peer (P2P) relations between producers (sources) and consumers (sinks) and thus enable to directly assign costs to the consumers.
-%Due to the non-determnistic tracing of power through the network the methods result in different allocation.
-Let $\allocatePeer$ denotes the power that is produced at bus $n$ at time $t$ and assigned to consumers at bus $m$.
-We will refer to it as a P2P relation between $n$ and $m$.
-Further let $\allocateTransaction$ denote flow on line $\ell$ at time $t$ caused by the P2P connection $n\rightarrow m$.\\
-\noindent
-In the following we show how the LMP can be reproduced on the basis of FA.
-
-\subsubsection*{Allocating Operational Expenditures}
-Consumers at bus $m$ compensate the OPEX proportionally to the assigned production, thus 
-\begin{align}
-\allocateOpex &= \sum_{n,s} \opexGeneration \, \generatorshare \, \allocatePeer \Forall{m,t} \label{eq:allocate_opexGeneration}
-\end{align}
-where $\generatorshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodal production. \\
-
-\subsubsection*{Allocating CO$_2$ Cost}
- Similar to the OPEX (compare \cref{eq:shift_opex_by_emission_cost}) consumers at bus $m$ compensate the CO$_2$ cost, thus 
-\begin{align}
- \allocateEmissionCost &= \sum_{n,s} \emissionPrice \, \emission \, \generatorshare \, \allocatePeer \Forall{m,t} \label{eq:allocate_emissionPrice}
-\end{align}
-
-\subsubsection*{Allocating Capital Expenditures for Generators}
-According to the polluter pays principle, the CAPEX allocation should differentiate between consumers who are `responsible` for investments and those who are not.
-Thus, it is crucial to know whether a time step $t$ makes investments in $\capacityGeneration$ necessary or if it would be feasible without (further) investments.
-The KKT variable of the upper generator capacity constraint $\muuppergeneration$ gives us the needed estimator: If $\muuppergeneration > 0$, the upper capacity constraint \cref{eq:upper_generation_capacity_constraint} is binding and it can be fairly assumed that these times steps cause investments in $\capacityGeneration$.
-If in contrast $\muuppergeneration = 0$, the generation $\generation$ is not bound and does not push further investments in $\capacityGeneration$.
-On the basis of the KKT variable $\muuppergeneration$ we are able to define a measure $\impactcapexgeneration$ for the impact on the investments in $\capacityGeneration$  
-\begin{align}
-    \impactcapexgeneration = \dfrac{\muuppergeneration}{\sum_t \muuppergeneration \, \generation}
-\end{align}
-which weights the contributions of consumers at bus $m$ at time $t$ to the CAPEX 
-\begin{align}
-    \allocateCapexGeneration &= \sum_{n,s} \capexGeneration \, \capacityGeneration \, \impactcapexgeneration\, \allocatePeer \Forall{m,t} \label{eq:allocate_capexGeneration}
-\end{align}
-
-\noindent
-If consumers at $m$ retrieve power from $n$ when $\impactcapexgeneration$ is large, their contribution to investments in $\capacityGeneration$ raises.
-On the other hand, if consumers at $m$ retrieve power from $n$ when $\impactcapexgeneration=0$ , their contribution to $\capacityGeneration$ remain unchanged.
-
-\noindent
-\textit{Side note}: A way more simple approach would neglect these thoughts and let  consumers at $m$ compensate CAPEX for generator $s$ at bus $n$ proportional to their retrieved power.
-This would lead us to $ \impactcapexgeneration = \left( \generation\right)^{-1}$.
-
-\subsubsection*{Allocating Capital Expenditures for Transmission Lines}
-The allocation of CAPEX per the transmission line $\capexFlow \, \capacityFlow $ follows a similar procedure as for the generators.
-Again, we build up on KKT variables $\muupperflow$ and $\mulowerflow$ which translate to the necessity of transmission expansion of $\ell$ at time $t$.
-The measure $\impactcapexflow$ respects the need of CAPEX of both flow directions when set to
-
-\begin{align}
-    \impactcapexflow = \dfrac{\muupperflow + \mulowerflow}{\sum_t \left( \muupperflow + \mulowerflow \right) \, \flow}
-\end{align}
-
-and weights the contributions of consumers at bus $m$ at time $t$ to the CAPEX in line $\ell$,
-\begin{align}
-    \allocateCapexFlow = \sum_{n,\ell} \capexFlow \, \capacityFlow\, \impactcapexflow, \allocateTransaction \Forall{m,t} \label{eq:allocate_capexFlow}
-\end{align}
-
-\subsubsection*{Reproducing the LMP Based Consumption Costs}
-The above presented quantities complete an allocation of the total cost $\totalcost$ which mirrors the cost allocation based on LMP, thus
-
-\begin{align}
- \allocateOpex + \allocateEmissionCost + \allocateCapexGeneration + \allocateCapexFlow \simeq \lmp[m] \, \demand[m] \Forall{m,t} 
- \label{eq:allocated_cost_mirror_lmp}
-\end{align}
-
-\newpage
-
-\section*{Theory}
-\subsubsection*{Lagrange Mutliplier}
-
-Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
-
-\begin{align}
- \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
-\end{align}
-...
-
-\subsection*{Full Lagrangian}
-\begin{align}
- \lagrangian\left(\generation, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
- &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
-%  &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
- &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
- &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
-\end{align}
-% 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-
-In this formulation, we treat the flow $\flow$ an a passive quantity only.
-The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one net power production at bus $n$.
-Thus with a fix demand $\demand$, they links the generation $\generation$ to the flow on each line according to
-\begin{align}
- \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  %= \sum_n \ptdf \injection 
+ \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  
+ \label{eq:flow_from_ptdf}
 \end{align}
@@
 \begin{align}
- \ptdf  \rightarrow \ptdf - \sum_m \ptdf[m] \, \slack[m]
- \label[relation]{eq:ptdf_slacked}
+ \ptdf = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
+ \label{eq:ptdf_slacked}
 \end{align}
+where $\ptdfEqual$ denote the PTDF with equally distributed slack.
 When bus $n$ injects excess power, is has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
@@
 % \end{align}
-The global maximum of the Lagrangian requires stationarity with respect to all variables. Using stationarity of the generation    
+
+\subsubsection*{Breaking Down the Full Lagrangian}
 \begin{align}
- \pdv{\lagrangian}{\generation} = 0  \Forall{n,s} \\
-\end{align}
-and solving for $\lmp$, leads to
-
-% \begin{align}
-% - \opexGeneration + \lmp - \sum_{m,\ell} \lmp[m] \, \incidence[m] \, \ptdf  - \muuppergeneration + \mulowergeneration - \sum_{\ell} \muupperflow \ptdf + \sum_{\ell} \mulowerflow \, \ptdf = 0 \Forall{n,s}
-% \end{align}
-
-\begin{align}
-\lmp &= \opexGeneration  + \muuppergeneration - \mulowergeneration + \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf + \sum_{m,\ell}\lmp[m] \, \incidence[m] \, \ptdf   \Forall{n,s} \label{eq:lmp_generalized}
-% &= \opexGeneration  + \muuppergeneration - \mulowergeneration + \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf + \sum_{m}\lmp[m] \left( \delta_{m,n} - \slack[m] \right) &\Forall{n,s} \label{eq:lmp_generalized}
+ \lagrangian\left(\generation, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
+ &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
+%  &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
+ &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
+ &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
 \end{align}
 % 
-\Cref{eq:lmp_generalized} must hold for every choice of slack in the PTDF.
+where $\boldsymbol{\lambda} = \left\lbrace \lmp \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
+The global maximum of the Lagrangian requires stationarity with respect to all variables.
+The stationarity of the generation capacity variable leads to
+\begin{align}
+ \pdv{\lagrangian}{\capacityGeneration}  = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+ \label{eq:capexGeneration_duality}
+\end{align}
+
+the stationarity of the transmission capacity to
+\begin{align}
+ \pdv{\lagrangian}{\capacityFlow} = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
+ \label{eq:capexFlow_duality}
+\end{align}
+
+and the stationarity of the generation to 
+\begin{align}
+ \pdv{\lagrangian}{\generation} &= 0  &\Forall{n,s} \\
+ \rightarrow \hspace{10pt}  \opexGeneration &=  \lmp - \muuppergeneration + \mulowergeneration - \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf - \sum_{m,\ell}\lmp[m] \, \incidence[m] \, \ptdf   &\Forall{n,s} \label{eq:opex_duality}
+\end{align}
+% 
+For the latter, we used \cref{eq:flow_from_ptdf} which sets the derivative of the flow with respect to the generation to 
+\begin{align}
+\pdv{\flow}{\generation} = \ptdf                                                                                                                                                   \end{align}
+\Cref{eq:capexGeneration_duality,eq:capexFlow_duality,eq:opex_duality} show how the capital and operational prices translate into dual variables. \\
+
+Note that \cref{eq:opex_duality} must hold for every choice of slack in the PTDF.
 According to \cref{eq:ptdf_slacked,eq:slack}, setting the slack to $\slack = \delta_{m,n}$ results in $\ptdf = \sum_\ell \incidence[m] \, \ptdf = 0$.
-This leads to our first represantion for Locational Market Price, which we will refer to as the ``Island Solution``.
-Accordingly, the LMP is directly determined by the local operational price and prices for the generation capacity constraint
+This leads to our first represantion for Locational Market Price, which we will refer to as the ``Island Solution``,
 \begin{align}
@@
 \end{align}
-However, using this Island Solution, we can derive a second representation of the LMP. Feeding the \cref{eq:lmp1} back into \cref{eq:lmp_generalized} and applying \cref{eq:slack}, leads to 
+Accordingly, the LMP is directly determined by the local operational price and prices for the generation capacity constraint. 
+However, from the Island Solution, we can derive a second representation for the LMP. Feeding \cref{eq:lmp1} back into \cref{eq:opex_duality} and applying \cref{eq:slack}, finally leads to 
 \begin{align}
 \lmp =  - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf + \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \Forall{n,s,t} 
+\label{eq:lmp2}
 \end{align}
-This equation generalizes \cref{eq:lmp1}.
-Note, that setting again $\slack = \delta_{m,n}$ will result in \cref{eq:lmp1}.
-It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints at each generator and each transmission line in the system.
-According to the slack $\slack$,  the possibility for allocating costs between buses and transmission components.
+This equation generalizes the former Island Solution for the LMP which again can be reproduced by setting $\slack[m] = \delta_{m,n}$.
+It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints from all generators and transmission lines in the system.
+
+\subsubsection*{Cost Allocation}
+
+The choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to.
+As we will see, there are various valid options for this choice.
+Mutliplied by the nodal consumption $\demandnodal = \sum_a \demand$, the slack translates into the peer-to-peer relation
+\begin{align}
+ \allocatePeer = \demandnodal \, \slack[m]   \Forall{n,t}
+ \label{eq:allocate_peer}
+\end{align}
+that is the power which is produced at bus $m$ at time $t$ and assigned to consumers at bus $n$. Or the other words, the part of demand $\demandnodal$ that is covered by generators at bus $m$. \\
+
+Likewise, multipying the PTDF with the nodal demand will result in the flow induces by the net consumption at $n$ coming from the distributed slack 
+\begin{align}
+ \sum_m \allocateTransaction = - \demandnodal  \, \ptdf \Forall{n,\ell,t}  
+ \label{eq:allocate_transaction}
+\end{align}
+where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}. \\
+
+The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
+\begin{align}
+ \sum_n \allocatePeer = \sum_s \generation
+\end{align}
+Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\sum_s \generation[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
+\begin{align}
+  \sum_n \generationshare[m] \, \allocatePeer = \generation[m] \Forall{m,s,t}
+  \label{eq:recipients_sum}
+\end{align}
+and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ equals the power flow on line $\ell$,
+\begin{align}
+ \sum_{m,n} \allocateTransaction = \flow
+\end{align}
+
+These two quantities give a more intuitive insight on how the nodal payments can be decomposed.
+Multiplying both sides of \Cref{eq:lmp2}  with the nodal demand $\demandnodal$ will breakdown the costs payed by consumers at node $n$ to the different expeditures
 
 \begin{align}
- \sum_m \allocateTransaction[m \rightarrow n] = - \demand \, \ptdf  
+ \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration
 \end{align}
+
+with the allocated payments: \\
 \begin{align}
- \allocatePeer[m \rightarrow n] = \demand[m] \, \slack[m] 
+ \allocateOpex &= \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
+ &\text{OPEX for generators} 
+ \label{eq:allocate_opexGeneration}\\
+ \allocateCapexGeneration &= \sum_m \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
+ &\text{CAPEX for the generators} 
+  \label{eq:allocate_capexGeneration} \\
+ \allocateCapexFlow &=  \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
+ &\text{CAPEX for the transmission system} 
+  \label{eq:allocate_capexFlow}
 \end{align}
+
+Let's have a look at the allocated OPEX first.
+Consumers at bus $n$ retrieve power from different generators $(m,s)$ each producing a share of the nodal production.
+Hence consumers at $n$ compensate the OPEX proportionally to the assigned production.
+
+ In this way (mutliply both sides of \cref{eq:recipients_sum} by $\opexGeneration$ and sum over $t$), all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s). \\
+
+The CAPEX allocation for generators reveil a similar relation.
+According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
+If $\muuppergeneration > 0$, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
+Thus it is these times steps which push investments in $\capacityGeneration$.
+If in contrast $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
+When summing over all CAPEX payments to generator $(m,s)$ each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$,
+
+\begin{align}
+ \sum_{n,t} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer = \capexGeneration \, \capacityGeneration
+\end{align}
+where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
+Throughout all time steps, each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
+% The non-profit rule can also be seen when looking at the the total revenue per generator, 
+% \begin{align}
+%  \sum_t \lmp \, \generation &= \sum_t \opexGeneration \generation + \muuppergeneration \generation - \mulowergeneration \, \generation &\Forall{n,s} \\
+%  &= \sum_t \opexGeneration \, \generation + \capexGeneration \, \capacityGeneration &\Forall{n,s}
+% \end{align}
+%  where we multiplied both sides of \cref{eq:lmp1} by the generation and using the complementary slackness.  
+\\
+
+The allocation of CAPEX for the transmission system $\allocateCapexFlow$ builds up on the KKT variables $\muupperflow$ and $\mulowerflow$.
+Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$.
+Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compenstations for the resulting investments.
+Again the sum of all CAPEX payments to line $\ell$ equal the total CAPEX spent, thus
+
+\begin{align}
+ \sum_{m,n,t} \left( \muupperflow - \mulowerflow\right) \allocateTransaction = \capexFlow \, \capacityFlow  
+\end{align}
+where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow
+\begin{align}
+\sum_{n,m} \allocateTransaction = \flow
+\end{align}
+% A similar relation applies for the transmission CAPEX.
+Mulplying both side of \cref{eq:opex_duality} by the nodal injection $\sum_s \generation - \sum_a \demand$ and using the complementary slackness ...
+% shows that the CAPEX per line amounts to the total congestion revenue
+% \begin{align}
+% \capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
+% \end{align}
+% The transmission lines carry power from nodes with lower LMP to nodes with higher LMP, the price difference times the transported power is their congestion revenue.
+\note{Sign is different to Tom's paper (there a minus is missing).}
+
+\subsubsection*{Adding CO$_2$ Constraints}
+
+Imposing an additional CO$_2$ constraint limiting the total emission to K,  
+\begin{align}
+ \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
+ \label[constraint]{eq:co2_constraint}
+\end{align}
+with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
+% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
+% 
+% \begin{align}
+% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
+% \label{eq:non_profit_generator_emission}
+% \end{align}
+% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
+As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
+\begin{align}
+\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
+\end{align}
+This leads to allocated CO$_2$ cost compensation of node $n$ of
+ \begin{align}
+ \allocateEmissionCost &= \emissionPrice \, \sum_{m,s} \emission[m] \, \generationshare[m] \, \allocatePeer \Forall{n,t} \label{eq:allocate_emissionPrice}
+\end{align}
+
+% ----------------------------------------------------------------------------------------
+% ----------------------------------------------------------------------------------------
 
@@
 
-% \section*{Showcase}
-% 
-% \subsection*{Network with CO$_2$ constraint}
-% We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
+\section*{Showcase}
+
+\subsection*{Network with CO$_2$ constraint}
+We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
 It consists of nine buses and ten time steps.
@@
 ...
-% 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{compare_allocation.png}
-%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constraint.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{compare_allocation.png}
+    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constraint.
 The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
 }
-%     \label{fig:cost_allocation}
-% \end{figure}
-% 
-% \begin{figure}[h]
-% \begin{subfigure}{.5\textwidth}
-% \centering
-%  \includegraphics[width=\textwidth]{network.png}
-%  \caption{}
-%  \label{fig:network}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}   
-%     \centering
-%     \includegraphics[width=\textwidth]{nodal_payments.png}
-%     \caption{}
-%     \label{fig:nodal_payments}
-% \end{subfigure}
-% \caption{Network used for showcasing.
+    \label{fig:cost_allocation}
+\end{figure}
+
+\begin{figure}[h]
+\begin{subfigure}{.5\textwidth}
+\centering
+ \includegraphics[width=\textwidth]{network.png}
+ \caption{}
+ \label{fig:network}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}   
+    \centering
+    \includegraphics[width=\textwidth]{nodal_payments.png}
+    \caption{}
+    \label{fig:nodal_payments}
+\end{subfigure}
+\caption{Network used for showcasing.
 (a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$.
 (b) shows the total nodal payments according to the cost allocation.}
-% \end{figure}
-% 
-% 
+\end{figure}
+
 % 
@@
 % 
-% The P2P relation can be further broke down to allocating generator specific $ \generatorshare \, \allocatePeer $ where $\generatorshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production.
-Similarly the allocation to consumers a are given by and consumer specific $ \normeddemand[m] \allocatePeer$ with the nodal comsumer share given by $\normeddemand = \demand/\sum_a \demand$.
+% The P2P relation can be further broke down to allocating generator specific $ \generationshare \, \allocatePeer $ where $\generationshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production.
+Similarly the allocation to consumers a are given by and consumer specific $ \demandshare[m] \allocatePeer$ with the nodal comsumer share given by $\demandshare = \demand/\sum_a \demand$.
 % \end{appendix}

### 2020-04-29 dist_cost_large.py

@@ -47,6 +47,7 @@ from netallocation.cost import locational_market_price
 from netallocation.grid import PTDF, Incidence, power_demand, network_injection
 from netallocation.linalg import dedup_axis, dot
 from numpy import eye
+import xarray as xr
 
 norm = lambda ds: ds/ds.sum()
 
@@ -56,6 +57,8 @@ s = m + ' ' + carrier
 t = n.snapshots[2]
 
 slack = norm(network_injection(n, t).clip(max=0))
+slack = xr.zeros_like(slack)
+slack += norm(np.random.rand(*slack.shape))
 k = PTDF(n) @ slack
 H = PTDF(n) - k
 K = Incidence(n)
@@ -68,7 +71,6 @@ mu_lo = n.generators_t.mu_lower.loc[t, s]
 line_mu_up = n.lines_t.mu_upper.loc[t]
 line_mu_lo = n.lines_t.mu_lower.loc[t]
 rhs  = ((line_mu_up - line_mu_lo) @ H.sel(bus=m))
-
 print(lhs.item(), rhs.item())
 
 lhs = - y.sel(bus=m) * K.sel(bus=m) @ H.sel(bus=m)
@@ -78,10 +80,14 @@ rhs = ((line_mu_up - line_mu_lo) * H.sel(bus=m)).sum() +\
 print(lhs.item(), rhs.item())
 
 
+lhs = y.sel(bus=m)
+rhs = -(line_mu_up - line_mu_lo) @ H.sel(bus=m) + y @ slack
+print(lhs.item(), rhs.item())
+
+
+# %%
 N = len(n.buses)
 D = dedup_axis(dot(K, H), ('bus_0', 'bus_1'))
-print(D.to_pandas().round(3))
-# %%
 p = D.sel(bus_1=m).to_series()
 f = H.sel(bus=m).to_series()
 c = pd.Series('red', p.index).where(p<0, 'blue')

### 2020-05-05 main.tex
further writing

@@
 \newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
+\newcommand{\generationnodal}[1][n]{g_{#1,t}}
 \newcommand{\capacityGeneration}{G_{n,s}}
@@
 \newcommand{\injection}{p_{n,t}}
-\newcommand{\netconsumption}{p^{-}_{n,t}}
-\newcommand{\netproduction}{p^{+}_{n,t}}
-\newcommand{\selfconsumption}{p^{\circ}_{n,t}}
+\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
+\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
+\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}
+
+\newcommand{\totalnetconsumption}{p^{-}_{t}}
+\newcommand{\totalnetproduction}{p^{+}_{t}}
+\newcommand{\totalselfconsumption}{p^{\circ}_{t}}
+
 \newcommand{\lagrangian}{\mathcal{L}}
@@
 \newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
-\newcommand{\allocatePeerNormed}[1][m \rightarrow n]{\tilde{A}_{#1,t}}
 \newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
@@
 
-The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit, commonly one MW, of net power production at bus $n$.
+The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
 Thus with a fix demand $\demand$, they direclty link the generation $\generation$ to the flow on each line according to
@@
 where $\ptdfEqual$ denote the PTDF with equally distributed slack.
-When bus $n$ injects excess power, is has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
+When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
 Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack
@@
 
-The choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to.
-As we will see, there are various valid options for this choice.
-Mutliplied by the nodal consumption $\demandnodal = \sum_a \demand$, the slack translates into the peer-to-peer relation
+% As shown above, the choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to. 
+As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to.
+Established flow allocation schemes haved used this connection and define the slack $\slack[m]$ for allocating power flows.
+According to the Equivalent Bilateral Exchanges (EBE) for example, the net demand of bus $n$ is covered by all other buses proportional to their net power production.
+The Marginal Participation (MP) on the other hand allows to allocate power also to buses with net demand and net production.
+In the Appendix, we revise the mathematical expressions of the slack $\slack[m]$ for the different allocation schemes.
+\\
+
+For a generic slack, the effective power which is produced at bus $m$ and assigned to consumers at bus $n$ is given by
 \begin{align}
@@
 \end{align}
-that is the power which is produced at bus $m$ at time $t$ and assigned to consumers at bus $n$. Or the other words, the part of demand $\demandnodal$ that is covered by generators at bus $m$. \\
-
+which we will refer to as the peer-to-peer allocation.
+In other words, it indicates the part of demand $\demandnodal$ that is effectively covered by generators at bus $m$.
 Likewise, multipying the PTDF with the nodal demand will result in the flow induces by the net consumption at $n$ coming from the distributed slack
@@
 \end{align}
-where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}. \\
-
+where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}.
 The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
 \begin{align}
- \sum_n \allocatePeer = \sum_s \generation
+ \sum_n \allocatePeer = \generationnodal[m] \Forall{m,s,t}
+ \label{eq:generator_sum}
 \end{align}
-Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\sum_s \generation[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
+where $\generationnodal = \sum_s \generation$ is the nodal generation.
+Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\generationnodal[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
 \begin{align}
@@
 \end{align}
-and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ equals the power flow on line $\ell$,
+and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ results in the power flow on line $\ell$,
 \begin{align}
- \sum_{m,n} \allocateTransaction = \flow
+ \sum_{m,n} \allocateTransaction = \flow \Forall{\ell,t}
+ \label{eq:transaction_sum}
 \end{align}
+\\
 
-These two quantities give a more intuitive insight on how the nodal payments can be decomposed.
-Multiplying both sides of \Cref{eq:lmp2}  with the nodal demand $\demandnodal$ will breakdown the costs payed by consumers at node $n$ to the different expeditures
+Once defined the quantities $\allocatePeer$ and $\allocateTransaction$ on a basis of a valid choice of slack, an intuitive decomposition of the nodal payments becomes vivid: Multiplying both sides of \Cref{eq:lmp2}  with the nodal demand $\demandnodal$ will breakdown the costs payed by consumers at node $n$ to the different expeditures
 
 \begin{align}
- \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration
+ \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration \Forall{n,t}
+ \label{eq:lmp_allocation}
 \end{align}
@@
 \begin{align}
- \allocateOpex &= \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
+ \allocateOpex &= 
+ \sum_{m,s} \allocateOpex[n \rightarrow (m,s)]= 
+ \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
  &\text{OPEX for generators} 
 \label{eq:allocate_opexGeneration}\\
- \allocateCapexGeneration &= \sum_m \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
+ \allocateCapexGeneration &= 
+ \sum_{m,s} \allocateCapexGeneration[n \rightarrow (m,s)] = 
+ \sum_m \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
  &\text{CAPEX for the generators} 
 \label{eq:allocate_capexGeneration} \\
- \allocateCapexFlow &=  \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
+ \allocateCapexFlow &=  
+ \sum_{\ell} \allocateCapexFlow[n \rightarrow \ell] =  
+ \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
  &\text{CAPEX for the transmission system} 
@@
 
+\subsubsection*{Cost Terms}
+
 Let's have a look at the allocated OPEX first.
-Consumers at bus $n$ retrieve power from different generators $(m,s)$ each producing a share of the nodal production.
-Hence consumers at $n$ compensate the OPEX proportionally to the assigned production.
-
- In this way (mutliply both sides of \cref{eq:recipients_sum} by $\opexGeneration$ and sum over $t$), all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s). \\
+Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly pay for operational costs.
+The OPEX allocation can thus be seen as P2P tradings between producers and consumers with fixed production prices.
+In this way, the generator $(m,s)$ retrieves the exact amount of money from consumers that it spends on the operation.
+In other words, all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s), thus
+\begin{align}
+\sum_{n} \allocateOpex[n \rightarrow (m,s)] = \opexGeneration \, \generation
+\label{eq:no_profit_opex}
+\end{align}
 
@@
 \begin{align}
- \sum_{n,t} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer = \capexGeneration \, \capacityGeneration
+ \sum_{n,t} \allocateCapexGeneration[n \rightarrow (m,s)] = \capexGeneration \, \capacityGeneration
+\label{eq:no_profit_capex_generation}
 \end{align}
 where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
-Throughout all time steps, each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
+Hence, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
 % The non-profit rule can also be seen when looking at the the total revenue per generator, 
@@
 Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$.
-Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compenstations for the resulting investments.
+Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments.
 Again the sum of all CAPEX payments to line $\ell$ equal the total CAPEX spent, thus
@@
 \begin{align}
- \sum_{m,n,t} \left( \muupperflow - \mulowerflow\right) \allocateTransaction = \capexFlow \, \capacityFlow  
+ \sum_{n,t} \allocateCapexFlow[n \rightarrow \ell] = \capexFlow \, \capacityFlow  
 \end{align}
-where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow
-\begin{align}
-\sum_{n,m} \allocateTransaction = \flow
-\end{align}
+where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow as stated in \cref{eq:transaction_sum}.
 % A similar relation applies for the transmission CAPEX.
@@
 \end{align}
-
-% ----------------------------------------------------------------------------------------
-% ----------------------------------------------------------------------------------------
-
-% The power injection per node and time denotes the nodal power imbalance between generation and demand
-% 
-% \begin{align}
-%  \injection = \sum_s \generation - \sum_a \demand
-% \end{align}
-% The net production is the amount of positive injection per node and time 
-% \begin{align}
-%  \netproduction = \text{min}\left( \injection, 0 \right) 
-% \end{align}
-% The nodal net consumption denotes the amount of negative injection 
-% \begin{align}
-%  \netconsumption = \text{min}\left( - \injection, 0 \right) 
-% \end{align}
-% The nodal self-consumption equals the minimum of the nodal production and demand. 
-% \begin{align}
-%  \selfconsumption = \text{min}\left( \netproduction, \netconsumption \right) 
-% \end{align}
+which expands the allocation of the electricity cost in \cref{eq:lmp_allocation} to 
+\begin{align}
+ \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration  + \allocateEmissionCost \Forall{n,t}
+ \label{eq:lmp_allocation_with_emission}
+\end{align}
 
@@
     \includegraphics[width=\textwidth]{compare_allocation.png}
-    \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ constraint.
+    \caption{Comparison between the flow based cost allocation and the LMP based cost per consumer.
 The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
@@
 
+\newpage
+\subsection*{Appendix: Allocation and Choices of Slack}
+
+The presented cost allocation is sensitive to the choice of slack $\slack[m]$.
+The latter shift the shares of $\allocateOpex$, $\allocateCapexGeneration$, $\allocateCapexFlow$ and $\allocateEmissionCost$ for each node, whereas the total payment per node remains the same.
+In the following we show the connection of this cost allocation to established flow allocation methods.
+\\
+
+\subsection*{Marginal Participation}
+
+The peer-to-peer allocation for the MP scheme is given by 
+\begin{align}
+\allocatePeer &= \selfconsumption + \netconsumption  \, \dfrac{\netproduction[m]}{\totalnetproduction}
+\end{align}
+
+where 
+\begin{itemize}
+ \item $\injection = \left( \generationnodal - \demandnodal \right) $ denotes the nodal injection
+ \item $\netproduction = \text{min}\left( \injection, 0 \right) $ the nodal net production 
+ \item $\netconsumption = \text{min}\left( - \injection, 0 \right)$ the nodal net consumption and
+ \item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the nodal self-consumption. That is the power generated and at the same time consumed at node $n$. 
+\end{itemize}
+
+In order to allocate all gross demand $\demand$ to the slack, the latter is defined by 
+\begin{align}
+\slack[m,n] &= \dfrac{\selfconsumption + \netconsumption  \, \dfrac{\netproduction[m]}{\totalnetproduction}}{\demand}
+\end{align}
+
+...
+
+% 
+% % Direct Coupling:\\
+% % \begin{align}
+% %  \slack[m] = \dfrac{\generationnodal[m]}{\sum_{n} \generationnodal[n]} 
+% % \end{align}
+% %
+
 % \bibliographystyle{ieeetr}

### 2020-05-05 dist_costs.py

@@ -17,11 +17,11 @@ import cartopy.crs as ccrs
 plt.rc('text', usetex=True)
 
 
-to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{m,t}$',
-                      branch_investment_cost = '$\mathcal{C}^{F}_{m,t}$',
-                      one_port_operational_cost = '$\mathcal{O}_{m,t}$',
-                      co2_cost = '$\mathcal{E}_{m,t}$',
-                      nodal_demand_cost = '$\sum_a \, \lambda_{m,t} \, d_{m,a,t}$')
+to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n,t}$',
+                      branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',
+                      one_port_operational_cost = '$\mathcal{O}_{n,t}$',
+                      co2_cost = '$\mathcal{E}_{n,t}$',
+                      nodal_demand_cost = '$\lambda_{n,t} \, d_{n,t}$')
 
 
 n = ntl.test.get_network_ac_dc()
@@ -54,8 +54,7 @@ fig.canvas.draw(); fig.tight_layout()
 fig.savefig(f'figures/network{tag}.png')
 
 # %%
-ds = ntl.allocate_flow(n, method='ebe', q=0)
-pr = nodal_production_revenue(n).rename(bus='payer')
+ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
 dc = nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)
 ca = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
@@ -106,36 +105,43 @@ fig.savefig(f'figures/nodal_payments{tag}.png')
 
 
 #%% opex flow
+from netallocation.cost import (allocate_one_port_operational_cost,
+                                allocate_one_port_investment_cost)
+from netallocation.plot import pos_neg_buscolors as buscolors
 norm = lambda ds: ds/ ds.abs().sum()
-t = 0
-opex = ntl.cost.allocate_one_port_operational_cost(ds, n).sel(snapshot=sns[t])
-opex = opex.sum(['source_carrier'])
-opex_flow = opex.peer_on_branch_to_peer.sum(['source', 'sink']).to_series()
-opex_injection = opex.peer_to_peer.sum('sink').rename(source='bus') -\
[…]

### 2020-05-05 leftover.py

@@ -1,0 +1,62 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Apr 30 19:41:23 2020
+
+@author: fabian
+"""
+
+
+from netallocation.cost import locational_market_price
+from netallocation.grid import (PTDF, Incidence, power_demand, network_injection,
+                                power_production)
+from netallocation.linalg import dedup_axis, dot
+from numpy import eye
+import xarray as xr
+
+norm = lambda ds: ds/ds.sum()
+
+m = 'DE0 37'
+t = n.snapshots[1]
+sel = dict(bus=m, snapshot=t)
+
+k = norm(power_production(n, t))
+# k = xr.zeros_like(k) + norm(np.random.rand(*k.shape))
+H = PTDF(n) - PTDF(n) @ k
+K = Incidence(n)
+y = locational_market_price(n, t)
+d = ntl.power_demand(n, t).sel(bus=m)
+
+line_mu_up = n.lines_t.mu_upper.loc[t]
+line_mu_lo = n.lines_t.mu_lower.loc[t]
+lhs = y.sel(bus=m)
+rhs = -(line_mu_up - line_mu_lo) @ H.sel(bus=m) + y @ k
+print(lhs.item(), rhs.item())
+
+# test equations
+o = n.generators.marginal_cost[s] + ntl.cost.nodal_co2_price(n, t).loc[m, carrier]
+mu_up = n.generators_t.mu_upper.loc[t, s]
+mu_lo = n.generators_t.mu_lower.loc[t, s]
+line_mu_up = n.lines_t.mu_upper.loc[t]
+line_mu_lo = n.lines_t.mu_lower.loc[t]
+rhs  = ((line_mu_up - line_mu_lo) @ H.sel(bus=m))
+print(lhs.item(), rhs.item())
+
+lhs = - y.sel(bus=m) * K.sel(bus=m) @ H.sel(bus=m)
+wo_m = n.buses.drop(m).index
+rhs = ((line_mu_up - line_mu_lo) * H.sel(bus=m)).sum() +\
+       y.sel(bus=wo_m) @ K.sel(bus=wo_m) @ H.sel(bus=m)
+print(lhs.item(), rhs.item())
+
+
+# %% test standard equation
+
+from netallocation.cost import reindex_by_bus_carrier, nodal_co2_price
+c = 'Generator'
+(
[…]

### 2020-05-14 main.tex
wrapp up theory

@@
 
+\newcommand{\ie}{\textit{i.e.} }
+\newcommand{\eg}{\textit{e.g.} }
+
 \newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
 \newcommand{\note}[1]{\textcolor{Orange}{#1}}
+\newcommand{\vpad}{\vspace{1mm}}
 
@@
 
-\subsubsection*{Lagrange Mutliplier}
+In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
+We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA.
+This will lead us to a generalized form of peer-to-peer allocations and transmission usage allocations based on the choice of slack.
+In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion.
+By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.
 
-Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
+\subsubsection*{Nomenclature}
 
+\begin{table}[h]
+	\centering
+	\begin{tabular}{ll}
+        $\demand$ & Electric demand per bus $n$, demand type $a$, time step $t$ in MW  \vpad \\
+        $\generation$ & Electric generation per bus $n$, carrier $s$, time step $t$  in MW \vpad \\
+        $\flow$ & Active power flow on line $\ell$, time step $t$ in MW   \vpad \\
+        $\opexGeneration$ & Operational cost (OPEX) in \euro/MW \vpad \\
+        $\capexGeneration$ & Capital Expenditure (CAPEX) in \euro/MW \vpad \\
+        $\capexFlow$  & CAPEX per transmission line $\ell$ in \euro/MW  \vpad \\
+        $\capacityGeneration$ & Generation capacity in MW \vpad \\
+        $\capacityFlow$ & Transmission capacity in MW \vpad \\
+        $\incidence$ & Incidence matrix \vpad \\
+        $\injection$ & Net nodal injection in MW, \ie $\left( \generationnodal - \demandnodal \right)$  \vpad \\
+        $\netproduction$ & Net nodal production in MW, \ie $\text{min}\left( \injection, 0 \right)$ \vpad \\
+        $\netconsumption$ & Net nodal consumption in MW, \ie $\text{min}\left( - \injection, 0 \right)$ \vpad \\ 
+        $\selfconsumption$ & Nodal self-consumption in MW, \ie $\text{min}\left( \netproduction, \netconsumption \right)$
+	\end{tabular}
+\end{table}
+
+\subsubsection*{Power Transfer Distribition Factors}
+
+The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
+Thus with a fix demand $\demand$, they directly link the generation $\generation$ to the flow on each line according to
 \begin{align}
- \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
+ \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  
+ \label{eq:flow_from_ptdf}
 \end{align}
-...
+The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
+It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
+The choice of slack modifies the PTDF accordingly
+\begin{align}
+ \ptdf = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
+ \label{eq:ptdf_slacked}
+\end{align}
+where $\ptdfEqual$ denote the PTDF with equally distributed slack.
+When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
+Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack
+\begin{align}
+\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
+\label{eq:slack}
+\end{align}
+where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
 
-\section*{Linear Energy Modelling and LMP}
+\subsubsection*{Flow Allocation} \label{sec:flow_allocation}
 
-We linearly optimize the capacity and dispatch of a simple power system.
+As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to.
+Established flow allocation schemes haved used this relation and define the slack $\slack[m]$ in order to allocate power flows and exchanges.
+According to the Equivalent Bilateral Exchanges (EBE) for example, the net demand of bus $n$ is covered by all other buses proportional to their net power production.
+The Marginal Participation (MP) on the other hand allows to allocate power also to buses with net demand and net production at the same time.
+In the following we focus on the power that is allocated to net consumers in the network.
+\\
+
+For a generic slack, the effective power which is produced at bus $m$ and assigned to consumers at bus $n$ is given by
+\begin{align}
+%  \allocatePeer = \demandnodal \, \slack[m]   \Forall{n,t}
+ \allocatePeer = \selfconsumption + \netconsumption \, \slack[m]   \Forall{n,t}
+ \label{eq:allocate_peer}
+\end{align}
+which we will refer to as the peer-to-peer allocation.
+It consists of two terms.
+The first term is the power which is produced at bus $n$ and at the same time consumed at bus $n$, namely the self-consumption $\selfconsumption$.
+We set this quantity to the minimum of nodal generation and demand, \ie $\selfconsumption = \text{min}(\generation, \demand)$, that is as long as the nodal generation does not exceed the nodal demand, it cannot be exported and assigned to other buses.
+The second term reflects the net deficit power of bus $n$ which is provided by the slack $\slack$.
+Following this concept, $\allocatePeer$ splits the electricity demand $\demandnodal$ to different providing buses.
+
+Multipying the PTDF with the net demand $\netconsumption$ results in the flow induced by all imports to $n$
+\begin{align}
+%  \sum_m \allocateTransaction = - \demandnodal  \, \ptdf \Forall{n,\ell,t}
+ \sum_m \allocateTransaction = - \netconsumption  \, \ptdf \Forall{n,\ell,t}
+ \label{eq:allocate_transaction}
+\end{align}
+where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}.
+The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
+\begin{align}
+ \sum_n \allocatePeer = \generationnodal[m] \Forall{m,s,t}
+ \label{eq:generator_sum}
+\end{align}
+where $\generationnodal = \sum_s \generation$ is the nodal generation.
+Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\generationnodal[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
+\begin{align}
+  \sum_n \generationshare[m] \, \allocatePeer = \generation[m] \Forall{m,s,t}
+  \label{eq:recipients_sum}
+\end{align}
+and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ results in the power flow on line $\ell$,
+\begin{align}
+ \sum_{m,n} \allocateTransaction = \flow \Forall{\ell,t}
+ \label{eq:transaction_sum}
+\end{align}
+\\
+
+\subsubsection*{Network Optimisation}
+
+% Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
+% 
+% \begin{align}
+%  \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
+% \end{align}
+% ...
+%
+
+We linearly cost-optimize the capacity and dispatch of a simple power system.
 
@@
     \underset{\demand, \generation, \capacityGeneration}{\text{max}}
-    \left(\utility(\demand) - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
+    \left(\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
 \end{align}
-
-where 
-\begin{itemize}
-\item[] $\utility$ denotes the utility function per bus $n$, demand type $a$ time step $t$ 
- \item[] $\demand$ denotes the eletric demand 
- \item[] $\capexGeneration$ denotes the capital expenditure (CAPEX) per node $n$ and generator type $s$
- \item[] $\capacityGeneration$ denotes the generation capacity
- \item[] $\opexGeneration$ denotes the operational cost (OPEX)
- \item[] $\generation$ denotes the net generation in MW
- \item[] $\capexFlow$ denotes the CAPEX per transmission line 
- \item[] $\capacityFlow$ denotes the transmission capacity.
-\end{itemize}
-% 
-\subsubsection*{Constraints}
-In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. \\
+% where 
+% \begin{itemize}
+% \item[] $\utility$ denotes the utility function per bus $n$, demand type $a$ time step $t$ 
+%  \item[] $\demand$ denotes the eletric demand 
+%  \item[] $\capexGeneration$ denotes the capital expenditure (CAPEX) per node $n$ and generator type $s$
+%  \item[] $\capacityGeneration$ denotes the generation capacity
+%  \item[] $\opexGeneration$ denotes the operational cost (OPEX)
+%  \item[] $\generation$ denotes the net generation in MW
+%  \item[] $\flow$ denotes the active power flow on line $\ell$
+%  \item[] $\capexFlow$ denotes the CAPEX per transmission line 
+%  \item[] $\capacityFlow$ denotes the transmission capacity.
+subject to following physical constraints. 
+\\
+% In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. \\
 
@@
 \end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. \\
+Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
+In a power market this is the \euro/\megawatthour-price which a consumer has to pay.
+Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\
 
@@
 
-\subsubsection*{Flow is a passive variable}
-We treat the flow $\flow$ an a passive quantity only that is it is only influenced by the nodal injection.
-
-The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus with a fix demand $\demand$, they direclty link the generation $\generation$ to the flow on each line according to
-\begin{align}
- \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  
- \label{eq:flow_from_ptdf}
-\end{align}
-The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
-It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
-The choice of slack modifies the PTDF accordingly
-\begin{align}
- \ptdf = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
- \label{eq:ptdf_slacked}
-\end{align}
-where $\ptdfEqual$ denote the PTDF with equally distributed slack.
-When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
-Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack
-\begin{align}
-\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
-\label{eq:slack}
-\end{align}
-where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
-% Sensitivity of flow $\flow$ against changes of the power production $\generation$  
-% \begin{align}
-%  \pdv{\flow}{\generation}  = \ptdf \label{eq:flow_sensitity} 
-% \end{align}
-
-\subsubsection*{Breaking Down the Full Lagrangian}
+\subsubsection*{Decomposing the Locational Marginal Prices and Allocating the Electricity Costs}
 \begin{align}
@@
 It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints from all generators and transmission lines in the system.
-
-\subsubsection*{Cost Allocation}
-
 % As shown above, the choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to. 
-As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to.
-Established flow allocation schemes haved used this connection and define the slack $\slack[m]$ for allocating power flows.
-According to the Equivalent Bilateral Exchanges (EBE) for example, the net demand of bus $n$ is covered by all other buses proportional to their net power production.
-The Marginal Participation (MP) on the other hand allows to allocate power also to buses with net demand and net production.
-In the Appendix, we revise the mathematical expressions of the slack $\slack[m]$ for the different allocation schemes.
-\\
-
-For a generic slack, the effective power which is produced at bus $m$ and assigned to consumers at bus $n$ is given by
-\begin{align}
- \allocatePeer = \demandnodal \, \slack[m]   \Forall{n,t}
- \label{eq:allocate_peer}
-\end{align}
-which we will refer to as the peer-to-peer allocation.
-In other words, it indicates the part of demand $\demandnodal$ that is effectively covered by generators at bus $m$.
-Likewise, multipying the PTDF with the nodal demand will result in the flow induces by the net consumption at $n$ coming from the distributed slack
-\begin{align}
- \sum_m \allocateTransaction = - \demandnodal  \, \ptdf \Forall{n,\ell,t}  
- \label{eq:allocate_transaction}
-\end{align}
-where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}.
-The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
-\begin{align}
- \sum_n \allocatePeer = \generationnodal[m] \Forall{m,s,t}
- \label{eq:generator_sum}
-\end{align}
-where $\generationnodal = \sum_s \generation$ is the nodal generation.
-Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\generationnodal[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
-\begin{align}
-  \sum_n \generationshare[m] \, \allocatePeer = \generation[m] \Forall{m,s,t}
-  \label{eq:recipients_sum}
-\end{align}
-and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ results in the power flow on line $\ell$,
-\begin{align}
- \sum_{m,n} \allocateTransaction = \flow \Forall{\ell,t}
- \label{eq:transaction_sum}
-\end{align}
-\\
-
-Once defined the quantities $\allocatePeer$ and $\allocateTransaction$ on a basis of a valid choice of slack, an intuitive decomposition of the nodal payments becomes vivid: Multiplying both sides of \Cref{eq:lmp2}  with the nodal demand $\demandnodal$ will breakdown the costs payed by consumers at node $n$ to the different expeditures
+Further, \cref{eq:lmp2} reveals a crucial connection of the nodal electricity cost to the above presented flow allocation:
+Multiplying both sides of \cref{eq:lmp2}  with the net nodal demand $\netconsumption$
+breaks down the costs of power import to $n$. Then adding the price of the selfconsumption $\lmp \selfconsumption$ to both sides, finally breaks down the total gross demand cost at $n$,
 
@@
 \end{align}
-
-\subsubsection*{Cost Terms}
+Note that we used \cref{eq:allocate_peer,eq:allocate_transaction} which determine the peer-to-peer transactions between $m \rightarrow n$ and the corresponding usage of lines.
+\Cref{eq:lmp_allocation,eq:allocate_opexGeneration,eq:allocate_capexGeneration,eq:allocate_capexFlow} show how physically allocated flows on the basis of the slack directly tranlate into a cost allocation.\\
 
 Let's have a look at the allocated OPEX first.
-Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly pay for operational costs.
-The OPEX allocation can thus be seen as P2P tradings between producers and consumers with fixed production prices.
+Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly compensate their operational costs.
+The OPEX allocation behaves like P2P tradings between producers and consumers with fixed production prices.
 In this way, the generator $(m,s)$ retrieves the exact amount of money from consumers that it spends on the operation.
@@
 Thus it is these times steps which push investments in $\capacityGeneration$.
-If in contrast $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
+If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
 When summing over all CAPEX payments to generator $(m,s)$ each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$,
@@
 where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
-Hence, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
+So in total, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
 % The non-profit rule can also be seen when looking at the the total revenue per generator, 
@@
 
+\newpage
 \section*{Showcase}
@@
 
-\newpage
-\subsection*{Appendix: Allocation and Choices of Slack}
-
-The presented cost allocation is sensitive to the choice of slack $\slack[m]$.
-The latter shift the shares of $\allocateOpex$, $\allocateCapexGeneration$, $\allocateCapexFlow$ and $\allocateEmissionCost$ for each node, whereas the total payment per node remains the same.
-In the following we show the connection of this cost allocation to established flow allocation methods.
-\\
-
-\subsection*{Marginal Participation}
-
-The peer-to-peer allocation for the MP scheme is given by 
-\begin{align}
-\allocatePeer &= \selfconsumption + \netconsumption  \, \dfrac{\netproduction[m]}{\totalnetproduction}
-\end{align}
-
-where 
-\begin{itemize}
- \item $\injection = \left( \generationnodal - \demandnodal \right) $ denotes the nodal injection
- \item $\netproduction = \text{min}\left( \injection, 0 \right) $ the nodal net production 
- \item $\netconsumption = \text{min}\left( - \injection, 0 \right)$ the nodal net consumption and
- \item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the nodal self-consumption. That is the power generated and at the same time consumed at node $n$. 
-\end{itemize}
-
-In order to allocate all gross demand $\demand$ to the slack, the latter is defined by 
-\begin{align}
-\slack[m,n] &= \dfrac{\selfconsumption + \netconsumption  \, \dfrac{\netproduction[m]}{\totalnetproduction}}{\demand}
-\end{align}
-
-...
-
 % 

### 2020-05-14 main.tex
replace with cycle formulation

@@
 \begin{align}
- \lagrangian\left(\generation, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
+ \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
  &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
-%  &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
+ &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
  &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
@@
 % 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
+where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
 The global maximum of the Lagrangian requires stationarity with respect to all variables.
@@
 \begin{align}
- \pdv{\lagrangian}{\generation} &= 0  &\Forall{n,s} \\
- \rightarrow \hspace{10pt}  \opexGeneration &=  \lmp - \muuppergeneration + \mulowergeneration - \sum_{\ell} \left( \muupperflow - \mulowerflow\right)  \, \ptdf - \sum_{m,\ell}\lmp[m] \, \incidence[m] \, \ptdf   &\Forall{n,s} \label{eq:opex_duality}
+ \pdv{\lagrangian}{\generation} &= 0 
+ \rightarrow \hspace{10pt}  \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
 \end{align}
-% 
-For the latter, we used \cref{eq:flow_from_ptdf} which sets the derivative of the flow with respect to the generation to 
-\begin{align}
-\pdv{\flow}{\generation} = \ptdf                                                                                                                                                   \end{align}
-\Cref{eq:capexGeneration_duality,eq:capexFlow_duality,eq:opex_duality} show how the capital and operational prices translate into dual variables. \\
-
-Note that \cref{eq:opex_duality} must hold for every choice of slack in the PTDF.
-According to \cref{eq:ptdf_slacked,eq:slack}, setting the slack to $\slack = \delta_{m,n}$ results in $\ptdf = \sum_\ell \incidence[m] \, \ptdf = 0$.
-This leads to our first represantion for Locational Market Price, which we will refer to as the ``Island Solution``,
+Solving \cref{eq:opex_duality} for the $\lmp$, leads to our first representation for Locational Market Price, which we will refer to as the ``Island Solution``,
 \begin{align}
@@
 \end{align}
-Accordingly, the LMP is directly determined by the local operational price and prices for the generation capacity constraint. 
-However, from the Island Solution, we can derive a second representation for the LMP. Feeding \cref{eq:lmp1} back into \cref{eq:opex_duality} and applying \cref{eq:slack}, finally leads to 
+It connects the LMP directly with the local operational price and prices for the generation capacity constraint.
+However, we can derive a second representation for $\lmp$.
+Starting from the stationarity of the flow
 \begin{align}
-\lmp =  - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf + \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \Forall{n,s,t} 
+ 0 &= \pdv{\lagrangian}{\flow}  \\
+ &= - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  - \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance &= 0
+\end{align}
+and multipying each term with the Power Transfer Distribution Factor $\ptdf$ leaves us with
+
+\begin{align}
+  0 &= - \lmp + \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)
+  \label{eq:lmp_lmp1_reduced}
+\end{align}
+According to \cref{eq:slack}, the first term splits into the LMP at $n$ and the LMP weighted with the slack.
+The final term disappears as the $\cycle \, \impedance$ is the kernel of the PTDF $\ptdf$, so $\sum_l \cycle \, \impedance \ptdf = 0$.
+Solving \cref{eq:lmp_lmp1_reduced}, and replacing the LMP of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to
+
+\begin{align}
+\lmp =  \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
 \label{eq:lmp2}

### 2020-05-19 main.tex
restructure first section, add appendix

@@
 \newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
-\newcommand{\generationnodal}[1][n]{g_{#1,t}}
+\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
 \newcommand{\capacityGeneration}{G_{n,s}}
@@
 \newcommand{\demand}[1][n]{d_{#1,a,t}}
-\newcommand{\demandnodal}[1][n]{d_{#1,t}}
+\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
 \newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
@@
 \newcommand{\slack}[1][n]{k_{#1}}
+\newcommand{\slackk}[1][n]{k^*_{#1}}
 \newcommand{\Slack}{k_{m,n}}
+\newcommand{\Slackk}{k^*_{m,n}}
 \newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
@@
         $\capacityFlow$ & Transmission capacity in MW \vpad \\
-        $\incidence$ & Incidence matrix \vpad \\
-        $\injection$ & Net nodal injection in MW, \ie $\left( \generationnodal - \demandnodal \right)$  \vpad \\
-        $\netproduction$ & Net nodal production in MW, \ie $\text{min}\left( \injection, 0 \right)$ \vpad \\
-        $\netconsumption$ & Net nodal consumption in MW, \ie $\text{min}\left( - \injection, 0 \right)$ \vpad \\ 
-        $\selfconsumption$ & Nodal self-consumption in MW, \ie $\text{min}\left( \netproduction, \netconsumption \right)$
+        $\incidence$ & Incidence matrix \vpad 
+%         $\injection$ & Net nodal injection in MW, \ie $\left( \nodalgeneration - \nodaldemand \right)$  \vpad \\
+%         $\netproduction$ & Net nodal production in MW, \ie $\text{min}\left( \injection, 0 \right)$ \vpad \\
+%         $\netconsumption$ & Net nodal consumption in MW, \ie $\text{min}\left( - \injection, 0 \right)$ \vpad \\ 
+%         $\selfconsumption$ & Nodal self-consumption in MW, \ie $\text{min}\left( \netproduction, \netconsumption \right)$
 	\end{tabular}
@@
 
-\subsubsection*{Power Transfer Distribition Factors}
+\subsubsection*{Power Transfer Distribition Factors and Flow Allocation}
 
-The Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus with a fix demand $\demand$, they directly link the generation $\generation$ to the flow on each line according to
+In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
+Thus for a given gross production $\generation$ and gross demand $\demand$, they directly link to the the resulting flow on each line,
 \begin{align}
- \flow\left( \generation\right)  = \sum_n \ptdf \left( \sum_s \generation- \sum_a \demand \right)  
+ \flow  = \sum_n \ptdf \left( \nodalgeneration- \nodaldemand \right)  
  \label{eq:flow_from_ptdf}
 \end{align}
+where $\nodalgeneration = \sum_s \generation$ is the production of all generators $s$ attached to $n$ and $\nodaldemand = \sum_a \demand$ the demand of all consumers $a$ attached to $n$.
 The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
 It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
-The choice of slack modifies the PTDF accordingly
+The choice of slack modifies the PTDF according to
 \begin{align}
- \ptdf = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
+ \ptdf\left( \slack[m]\right)  = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
  \label{eq:ptdf_slacked}
@@
 where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
-
-\subsubsection*{Flow Allocation} \label{sec:flow_allocation}
-
-As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to.
-Established flow allocation schemes haved used this relation and define the slack $\slack[m]$ in order to allocate power flows and exchanges.
-According to the Equivalent Bilateral Exchanges (EBE) for example, the net demand of bus $n$ is covered by all other buses proportional to their net power production.
-The Marginal Participation (MP) on the other hand allows to allocate power also to buses with net demand and net production at the same time.
-In the following we focus on the power that is allocated to net consumers in the network.
+% As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
+Established flow allocation schemes [cite] haved used this degree of freedom in order to allocate power flows and exchanges to market participants.
+Under the assumption that consumers account for all power flows in the grid, the slack is set to $\slack^*$ such that
+\begin{align}
+ \flow  = - \sum_n \ptdf\left( \slackk[m]\right) \, \nodaldemand  
+ \label{eq:flow_from_demand}
+\end{align}
+Therefore, the flow can be reproduced from the demand-side of the system only.
+Each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
+In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add to $\flow$.
+The subflows, in turn, can be further broken down to contributions of each individual production-to-demand relation, $m \rightarrow n$, that is
+\begin{align}
+ \allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  
+\end{align}
+% 
+It indicates the flow on line $\ell$ induced by the power flowing from generators at $m$ to consumers at $n$.
+When summing over all sources $m$ it yields the subflows induced by $\nodaldemand$, the same terms as on the right hand side in \cref{eq:flow_from_demand};
+% \begin{align}
+% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
+% \end{align}
+when summing over all sources and sinks, it yields again the power flow, thus
+\begin{align}
+\flow = \sum_{m,n} \allocateTransaction
+\end{align}
+% 
+The consumed power at $n$ has to come from the slack $\slackk[m]$.
+As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
+\begin{align}
+ \allocatePeer &= \slackk[m] \, \nodaldemand 
+\label{eq:allocate_peer}
+\end{align}
+% 
+Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation 
+\begin{align}
+ \sum_n \allocatePeer = \nodalgeneration[m]
+ \label{eq:sum_n_allocate_peer}
+\end{align}
+which we proof in \cref{sec:proof_sum_n_allocate_peer}. Summing over all sources yields the nodal demand 
+\begin{align}
+ \sum_m \allocatePeer = \nodaldemand
+ \label{eq:sum_m_allocate_peer}
+\end{align}
+which straightforwardly follows from the fact that $\sum_n \slackk = 1$.
+Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$.
+The first comes through multiplying with the share that generator $s$ has in current nodal production $\generationshare = \generation/\sum_s \generation$, the second by multiplying with the nodal comsumer share $\demandshare = \demand/\sum_a \demand$.
 \\
 
-For a generic slack, the effective power which is produced at bus $m$ and assigned to consumers at bus $n$ is given by
-\begin{align}
-%  \allocatePeer = \demandnodal \, \slack[m]   \Forall{n,t}
- \allocatePeer = \selfconsumption + \netconsumption \, \slack[m]   \Forall{n,t}
- \label{eq:allocate_peer}
-\end{align}
-which we will refer to as the peer-to-peer allocation.
-It consists of two terms.
-The first term is the power which is produced at bus $n$ and at the same time consumed at bus $n$, namely the self-consumption $\selfconsumption$.
-We set this quantity to the minimum of nodal generation and demand, \ie $\selfconsumption = \text{min}(\generation, \demand)$, that is as long as the nodal generation does not exceed the nodal demand, it cannot be exported and assigned to other buses.
-The second term reflects the net deficit power of bus $n$ which is provided by the slack $\slack$.
-Following this concept, $\allocatePeer$ splits the electricity demand $\demandnodal$ to different providing buses.
-
-Multipying the PTDF with the net demand $\netconsumption$ results in the flow induced by all imports to $n$
-\begin{align}
-%  \sum_m \allocateTransaction = - \demandnodal  \, \ptdf \Forall{n,\ell,t}
- \sum_m \allocateTransaction = - \netconsumption  \, \ptdf \Forall{n,\ell,t}
- \label{eq:allocate_transaction}
-\end{align}
-where the PTDF include the slack $\slack[m]$ according to \cref{eq:ptdf_slacked}.
-The slack $\slack[m]$ in \cref{eq:allocate_peer,eq:allocate_transaction} should be chosen such that the sum over all recipients $n$ in $\allocatePeer$ results in the nodal production, thus
-\begin{align}
- \sum_n \allocatePeer = \generationnodal[m] \Forall{m,s,t}
- \label{eq:generator_sum}
-\end{align}
-where $\generationnodal = \sum_s \generation$ is the nodal generation.
-Then, weighted by the nodal production share $\generationshare[m] = \generation[m]/\generationnodal[m]$ the allocation distributes all power pruduced by generator $(m,s)$,
-\begin{align}
-  \sum_n \generationshare[m] \, \allocatePeer = \generation[m] \Forall{m,s,t}
-  \label{eq:recipients_sum}
-\end{align}
-and the sum over all producers $m$ and recipients $n$ in $\allocateTransaction$ results in the power flow on line $\ell$,
-\begin{align}
- \sum_{m,n} \allocateTransaction = \flow \Forall{\ell,t}
- \label{eq:transaction_sum}
-\end{align}
-\\
+One straightforward solution for $\slackk$ is the share in the total production $\slackk = c \, \generation$ with $c$ being defined as $c = 1/\sum_n \generation$.
+That leads to the nodal demand $\nodaldemand$ being covered by all nodal generators in the network proportional to their power production.
+In \cref{sec:solution_space_of_the_slack} we discuss the solution space of $\slackk$ and how this can be extended to an individual slack for each node $\Slackk$.
 
@@
 \end{align}
-% where 
-% \begin{itemize}
-% \item[] $\utility$ denotes the utility function per bus $n$, demand type $a$ time step $t$ 
-%  \item[] $\demand$ denotes the eletric demand 
-%  \item[] $\capexGeneration$ denotes the capital expenditure (CAPEX) per node $n$ and generator type $s$
-%  \item[] $\capacityGeneration$ denotes the generation capacity
-%  \item[] $\opexGeneration$ denotes the operational cost (OPEX)
-%  \item[] $\generation$ denotes the net generation in MW
-%  \item[] $\flow$ denotes the active power flow on line $\ell$
-%  \item[] $\capexFlow$ denotes the CAPEX per transmission line 
-%  \item[] $\capacityFlow$ denotes the transmission capacity.
 subject to following physical constraints. 
 \\
-% In the following we neglect the utility $\utility$ of the nodal demand while fixing the demand $\demand$ to a predefined time-series. \\
 
@@
 \begin{align}
-  0 &= - \lmp + \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)
+  0 &= - \lmp + \sum_m \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right) \ptdf
   \label{eq:lmp_lmp1_reduced}
@@
 \end{align}
-This equation generalizes the former Island Solution for the LMP which again can be reproduced by setting $\slack[m] = \delta_{m,n}$.
-It depicts the interdependence of the LMP, that is how $\lmp$ can be decomposed to operational prices and prices for capacity constraints from all generators and transmission lines in the system.
-% As shown above, the choice of slack $\slack[m]$ in \cref{eq:lmp2} decides on the generators and lines to which the nodal price at $n$ is allocated to. 
+This representation  decomposes $\lmp$ to operational prices and prices for capacity constraints from all generators and transmission lines in the system. 
 Further, \cref{eq:lmp2} reveals a crucial connection of the nodal electricity cost to the above presented flow allocation:
@@
 \begin{align}
- \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration \Forall{n,t}
+ \lmp \, \nodaldemand = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration \Forall{n,t}
  \label{eq:lmp_allocation}
@@
 \begin{align}
- \lmp \, \demandnodal = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration  + \allocateEmissionCost \Forall{n,t}
+ \lmp \, \nodaldemand = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration  + \allocateEmissionCost \Forall{n,t}
  \label{eq:lmp_allocation_with_emission}
@@
 
-\newpage
-\section*{Showcase}
-
-\subsection*{Network with CO$_2$ constraint}
-We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
+% \newpage
+% \section*{Showcase}
+% 
+% \subsection*{Network with CO$_2$ constraint}
+% We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
 It consists of nine buses and ten time steps.
@@
 ...
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{compare_allocation.png}
-    \caption{Comparison between the flow based cost allocation and the LMP based cost per consumer.
+% 
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\textwidth]{compare_allocation.png}
+%     \caption{Comparison between the flow based cost allocation and the LMP based cost per consumer.
 The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
 }
-    \label{fig:cost_allocation}
-\end{figure}
-
-\begin{figure}[h]
-\begin{subfigure}{.5\textwidth}
-\centering
- \includegraphics[width=\textwidth]{network.png}
- \caption{}
- \label{fig:network}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}   
-    \centering
-    \includegraphics[width=\textwidth]{nodal_payments.png}
-    \caption{}
-    \label{fig:nodal_payments}
-\end{subfigure}
-\caption{Network used for showcasing.
+%     \label{fig:cost_allocation}
+% \end{figure}
+% 
+% \begin{figure}[h]
+% \begin{subfigure}{.5\textwidth}
+% \centering
+%  \includegraphics[width=\textwidth]{network.png}
+%  \caption{}
+%  \label{fig:network}
+% \end{subfigure}
+% \begin{subfigure}{.5\textwidth}   
+%     \centering
+%     \includegraphics[width=\textwidth]{nodal_payments.png}
+%     \caption{}
+%     \label{fig:nodal_payments}
+% \end{subfigure}
+% \caption{Network used for showcasing.
 (a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$.
 (b) shows the total nodal payments according to the cost allocation.}
-\end{figure}
-
+% \end{figure}
+% 
+% 
 % 
@@
 
-% \begin{appendix}
-% Since the P2P comprises all net production and net consumption summing over all sources must result in the nodal demand  
-% 
-% \begin{align}
-%  \sum_{n} \allocatePeer = \sum_a \demand[m]
-% \end{align}
-% as well as summing over all sinks results in the nodal generation
-% \begin{align}
-%  \sum_{m} \allocatePeer = \sum_s \generation
-% \end{align}
-% 
-% The P2P relation can be further broke down to allocating generator specific $ \generationshare \, \allocatePeer $ where $\generationshare = \generation/\sum_s \generation$ denotes the share of generator $s$ of the nodel production.
-Similarly the allocation to consumers a are given by and consumer specific $ \demandshare[m] \allocatePeer$ with the nodal comsumer share given by $\demandshare = \demand/\sum_a \demand$.
-% \end{appendix}
+\appendix
 
-% 
-% % Direct Coupling:\\
-% % \begin{align}
-% %  \slack[m] = \dfrac{\generationnodal[m]}{\sum_{n} \generationnodal[n]} 
-% % \end{align}
-% %
+\section{Appendix}
+
+\subsection{Proof \Cref{eq:allocate_peer}}
+\label{sec:proof_allocate_peer}
+
+\Cref{eq:allocate_peer} follows from summing $\allocateTransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
+\begin{align}
+ \allocatePeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocateTransaction \\
+ &= \slackk \, \nodaldemand - \sum_\ell \incidence \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  \\
+ &= \slackk \, \nodaldemand - \left(  \delta_{n,m} - \dfrac{1}{N} - \delta_{n,n} + \dfrac{1}{N} \right) \slackk[m] \, \nodaldemand  \\
+ &= \slackk \, \nodaldemand - \left(  \delta_{n,m} - 1 \right) \slackk[m] \, \nodaldemand  \\
+ &= \slackk[m] \, \nodaldemand 
+\label{eq:proof_allocate_peer}
+\end{align}
+where we used \cref{eq:slack} and the fact that the equally distributed slack amounts to $1/N$ for all N nodes in the network.
+
+\subsection{Proof of \Cref{eq:sum_n_allocate_peer}}
+\label{sec:proof_sum_n_allocate_peer}
+The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatePeer$
+\begin{align}
+\sum_m \incidence[m] \, \flow &= - \sum_{m,n} \incidence[m] \, \ptdf \, \nodaldemand \\
+\nodalgeneration[m] - \nodaldemand[m] &= - \delta_{m,n} \, \nodaldemand + \slackk[m] \, \nodaldemand    \\
+\allocatePeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
+\sum_n \allocatePeer &= \nodalgeneration[m]
+\end{align}
+
+\subsection{Solution Space of $\boldsymbol{\slackk[m]}$}
+\label{sec:solution_space_of_the_slack}
+
+All choices of $\slackk[m]$ fulfilling \cref{eq:flow_from_demand} determine the solution space of $\slackk$.
+In the $N \times N$ nodal space this translates to the constraint given by \cref{eq:sum_n_allocate_peer} which denotes
+\begin{align}
+ \nodalgeneration[m] = \sum_n \slackk[m] \, \nodaldemand  
+ \label{eq:solution_space_slack}
+\end{align}
+Solving for $\slackk[m]$ directly leads to 
+\begin{align}
+ \slackk[m] = c \, \generation[m] 
+\end{align}
+where $c$ is the inverse of the total consumption or production $c = 1 / \sum_n \nodaldemand   = 1 / \sum_n \nodalgeneration$. \\
+However \cref{eq:solution_space_slack} has an inherent degree of freedom and can be reformulated as 
+\begin{align}
+ \nodalgeneration[m] = \sum_n \Slackk \, \nodaldemand  
+ \label{eq:solution_space_slack_extended}
+\end{align}
+where $\Slackk$ denote the individual choice of slack for each bus $n$. For example this
 

### 2020-05-19 config.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri May 15 15:20:25 2020
+
+@author: fabian
+"""
+
+import pandas as pd
+
+to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n,t}$',
+                      branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',
+                      one_port_operational_cost = '$\mathcal{O}_{n,t}$',
+                      co2_cost = '$\mathcal{E}_{n,t}$',
+                      nodal_demand_cost = '$\lambda_{n,t} \, d_{n,t}$')
+
+color = pd.Series({'one_port_operational_cost': 'darkkhaki',
+                   'co2_cost': 'tomato',
+                   'one_port_investment_cost': 'palevioletred',
+                   'branch_investment_cost': 'mediumaquamarine',
+                   'nodal_demand_cost': 'cadetblue'})
+
+latex_names = dict(p_nom_opt = '$G$', marginal_cost = '$o$', capital_cost='$c$',
+                   bus='$n$')

### 2020-05-19 dist_costs.py

@@ -48,7 +48,7 @@ plot = n.plot(bus_sizes=bus_sizes/1e4, line_widths=n.lines.s_nom_opt/1e2,
        link_widths=n.links.p_nom_opt/1e2, ax=ax)
 ntl.plot.annotate_bus_names(n, ax, shift=(0,0.3),
                             transform=ccrs.PlateCarree(),
-                            bbox=dict(facecolor='white', alpha=0.5, edgecolor='None'))
+                            bbox='fancy')
 ax.legend(*ntl.plot.handles_labels_for(n.carriers.color), loc='upper left')
 fig.canvas.draw(); fig.tight_layout()
 fig.savefig(f'figures/network{tag}.png')

### 2020-05-19 examples.py

@@ -1,0 +1,131 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri May 15 15:04:53 2020
+
+@author: fabian
+"""
+
+import netallocation as ntl
+from netallocation.plot import handles_labels_for
+import pypsa
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+import seabor as sns
+from config import to_symbol_dict, color
+plt.rc('text', usetex=False)
+
+tag = ''
+add_gen_args = {}
+
+# tag = '_constraint_capacity'
+# add_gen_args = {'p_nom_max': 80}
+
+
+n = pypsa.Network()
+n0 = '0'
+n1 = '1'
+d0 = 50
+d1 = 80
+o0 = 50
+o1 = 200
+c0 = 500
+c1 = 500
+c_l = 100
+
+n.madd('Bus', [n0, n1], x=[0, 1], y=[0, 0.5])
+n.madd('Load', [0, 1], bus=[n0, n1], p_set=[d0, d1])
+n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
+       marginal_cost=[o0, o1], capital_cost=[c0, c1], **add_gen_args)
+n.madd('Line', [0], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
+       capital_cost=c_l)
+n.lopf(pyomo=False, keep_shadowprices=True)
+
+sn = 'now'
+
+G0 = int(n.generators.p_nom_opt[0])
+G1 = int(n.generators.p_nom_opt[1])
+F = int(n.lines.s_nom_opt[0])
+
+g0 = int(n.generators_t.p['0']['now'])
+g1 = int(n.generators_t.p['1']['now'])
+f = int(n.lines_t.p0['0']['now'])
+
+ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
+dc = ntl.cost.nodal_demand_cost(n).rename(bus='payer')
+ca = ntl.allocate_cost(n, method=ds, q=0)
+ca_sum = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
+ca_df = ca_sum.sel(snapshot=sn, drop=True).to_dataframe()
+dc_df = dc.sel(snapshot=sn, drop=True).to_dataframe()
+
+# %%
+fig, ax = plt.subplots()
+
[…]

### 2020-05-24 main.tex
add price decomposition plot for intro

@@
-\documentclass[11pt]{article}
+\documentclass[11pt,twocolumn]{article}
 \usepackage{graphicx}
-\usepackage[left=2.00cm, right=2.00cm, top=2.00cm, bottom=2.00cm]{geometry}
+\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
 \usepackage{amsmath}
@@
 \newcommand{\totalcost}{\mathcal{TC}}
+\newcommand{\totalOpexGeneration}{\mathcal{O}}
+\newcommand{\totalOpexFlow}{\mathcal{O}^F}
+\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
+\newcommand{\totalCapexFlow}{\mathcal{C}^F}
+\newcommand{\totalEmissionCost}{\mathcal{E}}
+
 \newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
@@
 %math 
-\newcommand{\resultsin}[1]{\hspace{12pt} \bot  \hspace{12pt} #1}
-\newcommand{\Forall}[1]{\hspace{20pt} \forall \,\, #1 }
+\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
+\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
 \newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
@@
 \title{From Linear Optimization to Transmission Cost Allocation}
-\author{Fabian Hofmann}
+\author{F. Hofmann. T. Brown}
 
@@
 
-% \begin{abstract}
-% The abstract text goes here.
-% \end{abstract}
+\begin{abstract}
+Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and  
+\end{abstract}
 
-In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
+\section{Introduction}
+
+% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
 We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA.
@@
 By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.
+
+\section{Locational Market Price}
 
@@
 	\begin{tabular}{ll}
-        $\demand$ & Electric demand per bus $n$, demand type $a$, time step $t$ in MW  \vpad \\
-        $\generation$ & Electric generation per bus $n$, carrier $s$, time step $t$  in MW \vpad \\
-        $\flow$ & Active power flow on line $\ell$, time step $t$ in MW   \vpad \\
+        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
+        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
+        $\generation$ & Electric generation per bus $n$, carrier $s$, \\ & time step $t$  in MW \vpad \\
+        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
         $\opexGeneration$ & Operational cost (OPEX) in \euro/MW \vpad \\
@@
         $\incidence$ & Incidence matrix \vpad 
-%         $\injection$ & Net nodal injection in MW, \ie $\left( \nodalgeneration - \nodaldemand \right)$  \vpad \\
-%         $\netproduction$ & Net nodal production in MW, \ie $\text{min}\left( \injection, 0 \right)$ \vpad \\
-%         $\netconsumption$ & Net nodal consumption in MW, \ie $\text{min}\left( - \injection, 0 \right)$ \vpad \\ 
-%         $\selfconsumption$ & Nodal self-consumption in MW, \ie $\text{min}\left( \netproduction, \netconsumption \right)$
 	\end{tabular}
 \end{table}
+
+The total system cost $\totalcost$ has to be payed back by the consumers 
+\begin{align}
+ \totalcost =  \sum_{n,a,t} \lmp \, \demand
+\end{align}
+
+\begin{align}
+ \totalcost = \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
+\end{align}
+
+\begin{figure}[h]
+ \includegraphics[width=\linewidth]{figures/price_decomposition.png}
+ \label{fig:price_decomposition}
+ \caption{The Locational Market Price}
+\end{figure}
+
+\section{Mathematical Theory}
 
@@
 \end{align}
-where $\nodalgeneration = \sum_s \generation$ is the production of all generators $s$ attached to $n$ and $\nodaldemand = \sum_a \demand$ the demand of all consumers $a$ attached to $n$.
+where $\nodalgeneration = \sum_s \generation$ and $\nodaldemand = \sum_a \demand$ combine all generators $s$ and all comsumers $a$ attached to $n$.
 The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
@@
 When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
-Summing over all ingoing and outgoing flow changes resulting from an positive injection at $n$ yields again the slack
+Summing over all ingoing and outgoing flow changes resulting from a positive injection at $n$ yields again the slack
 \begin{align}
@@
 \end{align}
-where $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
+Note that $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
 % As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
@@
 Therefore, the flow can be reproduced from the demand-side of the system only.
-Each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
-In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add to $\flow$.
-The subflows, in turn, can be further broken down to contributions of each individual production-to-demand relation, $m \rightarrow n$, that is
+Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
+In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add up to $\flow$.
+These subflows, in turn, can be further broken down to contributions for each bus $m$ in the slack, such that we get the subflow of individual $m \rightarrow n$ relations, that is
 \begin{align}
  \allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
+ \label{eq:allocate_transaction}
 \end{align}
 % 
-It indicates the flow on line $\ell$ induced by the power flowing from generators at $m$ to consumers at $n$.
-When summing over all sources $m$ it yields the subflows induced by $\nodaldemand$, the same terms as on the right hand side in \cref{eq:flow_from_demand};
+It indicates the flow on line $\ell$ provided generators at $m$ and supplying the demand $\nodaldemand$.
+When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand};
 % \begin{align}
@@
 \flow = \sum_{m,n} \allocateTransaction
+\label{eq:transaction_sum}
 \end{align}
 % 
-The consumed power at $n$ has to come from the slack $\slackk[m]$.
+As mentioned, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
 As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
@@
 % 
-Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation 
+Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
 \begin{align}
@@
 \end{align}
-which we proof in \cref{sec:proof_sum_n_allocate_peer}. Summing over all sources yields the nodal demand 
+and summing over all sources yields the nodal demand 
 \begin{align}
@@
 which straightforwardly follows from the fact that $\sum_n \slackk = 1$.
-Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$.
-The first comes through multiplying with the share that generator $s$ has in current nodal production $\generationshare = \generation/\sum_s \generation$, the second by multiplying with the nodal comsumer share $\demandshare = \demand/\sum_a \demand$.
+Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
 \\
 
-One straightforward solution for $\slackk$ is the share in the total production $\slackk = c \, \generation$ with $c$ being defined as $c = 1/\sum_n \generation$.
-That leads to the nodal demand $\nodaldemand$ being covered by all nodal generators in the network proportional to their power production.
-In \cref{sec:solution_space_of_the_slack} we discuss the solution space of $\slackk$ and how this can be extended to an individual slack for each node $\Slackk$.
+The solution to $\slackk$ follows directly from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
+That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.
+However solution space can be extended to an individual slack for each node $\Slackk$, which we discuss in \cref{sec:solution_space_of_the_slack}.
 
@@
 \begin{align}
-    \sum_l \incidence \, \flow - \sum_s \generation + \sum_a \demand &= 0 \resultsin{\lmp} \Forall{n,t}
+    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
     \label[constraint]{eq:nodal_balance_lin}
@@
 \begin{align}
- \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   = &- \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
+\notag
+ & \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
+ \notag
+ & - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
+ \notag
  &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
- &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \label[constraint]{eq:langrange_cycle_constraint} \\ 
+ \notag
+ &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
+ \notag
  &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
  &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
+ \label{eq:full_lagrangian}
 \end{align}
@@
 \begin{align}
- \pdv{\lagrangian}{\capacityGeneration}  = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+ \pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
+ \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
  \label{eq:capexGeneration_duality}
@@
 \begin{align}
- \pdv{\lagrangian}{\capacityFlow} = 0 \hspace{10pt} \rightarrow \hspace{10pt} \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
+ \pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
+ \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
  \label{eq:capexFlow_duality}
@@
 \begin{align}
- \pdv{\lagrangian}{\generation} &= 0 
- \rightarrow \hspace{10pt}  \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
+ \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
+ \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
 \end{align}
@@
 \begin{align}
- 0 &= \pdv{\lagrangian}{\flow}  \\
- &= - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  - \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance &= 0
+ 0 =& \pdv{\lagrangian}{\flow}  \\
+ \notag
+ 0 =& - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  \\
+ & -\sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance
 \end{align}
@@
 The final term disappears as the $\cycle \, \impedance$ is the kernel of the PTDF $\ptdf$, so $\sum_l \cycle \, \impedance \ptdf = 0$.
-Solving \cref{eq:lmp_lmp1_reduced}, and replacing the LMP of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to
+Solving \cref{eq:lmp_lmp1_reduced} for $\lmp$ and replacing $\lmp[m]$ of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to
 
 \begin{align}
-\lmp =  \sum_m \left( \opexGeneration[m] + \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] - \sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
+\notag
+\lmp =&  \sum_m \opexGeneration[m]\, \slack[m]  + \sum_m \left( \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \\
+& -\sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
 \label{eq:lmp2}
 \end{align}
-This representation  decomposes $\lmp$ to operational prices and prices for capacity constraints from all generators and transmission lines in the system. 
-Further, \cref{eq:lmp2} reveals a crucial connection of the nodal electricity cost to the above presented flow allocation:
-Multiplying both sides of \cref{eq:lmp2}  with the net nodal demand $\netconsumption$
-breaks down the costs of power import to $n$. Then adding the price of the selfconsumption $\lmp \selfconsumption$ to both sides, finally breaks down the total gross demand cost at $n$,
-
+%  
+% NOTE: \cref{eq:lmp2} builds a starting point for breaking down the market value (MV) 
+% of generator $(m,s)$, defined as
+% \begin{align}
+%  MV_s = \dfrac{\sum_t \generation \, \lmp}{\sum_t \generation}
+% \end{align}
+In contrast to the Island Solution in \cref{eq:lmp1}, this representation connects $\lmp$ with all other prices in the system.
+Moreover, it must hold for any choice of $\slack[m]$ and correspondingly decomposes the LMP to operational prices $\opexGeneration[m]$ and prices for capacity bounds for generators and transmission lines.
+When setting the slack to $\slackk[m] = c \, \generation[m]$, the LMP decomposes to prices of generators proportional to their production plus an extra term for the transmission usage.
+This directly links to the flow allocation presented above.
+Multiplied with the nodal demand $\nodaldemand$ the left hand side of \cref{eq:lmp2} turns into the total payment of $n$ and the right hand side into the different payment allocations,
 \begin{align}
- \lmp \, \nodaldemand = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration \Forall{n,t}
+ \lmp \, \nodaldemand = \allocateOpex + \allocateCapexGeneration + \allocateCapexFlow \Forall{n,t}
  \label{eq:lmp_allocation}
 \end{align}
-
-with the allocated payments: \\
+which we define as  
 \begin{align}
  \allocateOpex &= 
- \sum_{m,s} \allocateOpex[n \rightarrow (m,s)]= 
  \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
- &\text{OPEX for generators} 
 \label{eq:allocate_opexGeneration}\\
  \allocateCapexGeneration &= 
- \sum_{m,s} \allocateCapexGeneration[n \rightarrow (m,s)] = 
- \sum_m \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
- &\text{CAPEX for the generators} 
+ \sum_{m,s} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
 \label{eq:allocate_capexGeneration} \\
  \allocateCapexFlow &=  
- \sum_{\ell} \allocateCapexFlow[n \rightarrow \ell] =  
  \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
- &\text{CAPEX for the transmission system} 
 \label{eq:allocate_capexFlow}
 \end{align}
-Note that we used \cref{eq:allocate_peer,eq:allocate_transaction} which determine the peer-to-peer transactions between $m \rightarrow n$ and the corresponding usage of lines.
-\Cref{eq:lmp_allocation,eq:allocate_opexGeneration,eq:allocate_capexGeneration,eq:allocate_capexFlow} show how physically allocated flows on the basis of the slack directly tranlate into a cost allocation.\\
+\Cref{eq:allocate_opexGeneration} denotes the payments for generators' OPEX, \cref{eq:allocate_capexGeneration} for generators' CAPEX and \cref{eq:allocate_capexFlow} for the transmission system's CAPEX respectively.
+All those payment allocation build up on the phycally allocated flows presented in \cref{eq:allocate_peer,eq:allocate_transaction}.
+Thus with the right choice of slack the power flow allocations directly tranlate into a cost allocation.
+Further, the allocated payments can be considered on a more detailed level as
+\begin{align}
+ \allocateOpex[n \rightarrow (m,s)] &= 
+ \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
+\label{eq:allocate_opexGeneration_detailed}\\
+ \allocateCapexGeneration[n \rightarrow (m,s)] &= 
+ \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
+\label{eq:allocate_capexGeneration_detailed} \\
+ \allocateCapexFlow[n \rightarrow \ell] &=  
+  \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
+\label{eq:allocate_capexFlow_detailed}
+\end{align}
+%
 
-Let's have a look at the allocated OPEX first.
+Let's have a look at the allocated OPEX in \cref{eq:allocate_opexGeneration,eq:allocate_opexGeneration_detailed} first.
 Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly compensate their operational costs.
@@
 
-The CAPEX allocation for generators reveil a similar relation.
+The CAPEX allocation for generators $\allocateCapexGeneration$ defined in \cref{eq:allocate_capexGeneration,eq:allocate_capexGeneration_detailed}, reveils a similar relation.
 According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
@@
 where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
-So in total, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation (non-profit rule).
+So in total, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation, reflecting the zero-profit rule.
 % The non-profit rule can also be seen when looking at the the total revenue per generator, 
@@
 
-The allocation of CAPEX for the transmission system $\allocateCapexFlow$ builds up on the KKT variables $\muupperflow$ and $\mulowerflow$.
+The allocation of CAPEX for the transmission system $\allocateCapexFlow$, defined in \cref{eq:allocate_capexFlow,eq:allocate_capexFlow_detailed}, builds up on the KKT variables $\muupperflow$ and $\mulowerflow$.
 Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$.
@@
 
+\newpage
 \appendix
@@
 \end{align}
-where $\Slackk$ denote the individual choice of slack for each bus $n$. For example this
+where $\Slackk$ denote the individual choice of slack for each bus $n$.
+For example the Marginal Participation cite[] algorithm takes only net injections into account.
+This sets the individual slack to
+\begin{align}
+\Slackk &= \dfrac{ \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]}{\nodaldemand}
+\end{align}
+where 
+\begin{itemize}
+%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
+ \item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
+ \item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
+ \item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
+ \item $\gamma_t = \dfrac{1}{\sum_n \netproduction} = \dfrac{1}{\sum_n \netconsumption}$ is the inverse of the total injected/extracted power at time $t$.
+\end{itemize}
 
 % \bibliographystyle{ieeetr}
-% \printbibliography
+\printbibliography
 

### 2020-05-24 examples.py

@@ -12,7 +12,7 @@ import pypsa
 import pandas as pd
 import matplotlib.pyplot as plt
 import numpy as np
-import seabor as sns
+import seaborn as sns
 from config import to_symbol_dict, color
 plt.rc('text', usetex=False)
 

### 2020-05-24 price_decomposition.py

@@ -1,0 +1,58 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Sat May 23 17:33:26 2020
+
+@author: fabian
+"""
+
+import matplotlib.pyplot as plt
+import pandas as pd
+from config import color
+from matplotlib.patches import FancyArrowPatch
+from mpl_toolkits.axisartist.axislines import SubplotZero
+
+plt.rc('patch', facecolor='k')
+
+
+cost_i = ['CAPEX Production', 'CAPEX Transmission System',
+          'OPEX Production', 'Other Cost (e.g. CO$_2$ Cost)']
+lmp = pd.DataFrame([[25, 8, 15, 5]], index=['lmp'], columns=cost_i)
+c = pd.Series(color[['one_port_investment_cost', 'branch_investment_cost',
+                     'one_port_operational_cost', 'co2_cost']].values,
+              index=cost_i)
+
+
+
+
+fig = plt.figure(figsize=(1.5,4))
+ax = SubplotZero(fig, 111)
+fig.add_subplot(ax)
+ax.axis['yzero'].set_axisline_style("-|>")
+ax.axis['yzero'].set_visible(True)
+for direction in ["left", "right", "bottom", "top"]:
+    # hides borders
+    ax.axis[direction].set_visible(False)
+
+# fig, ax = plt.subplots(figsize=(1.5,4))
+# for edge in ['bottom', 'right', 'top']:
+#     ax.spines[edge].set_visible(False)
+
+lmp.plot.bar(stacked=True, ax=ax, legend=False, color=c, position=-0.3)
+ax.set_xlim(left=0)
+
+bottom = 0
+for i, c in enumerate(lmp):
+    y = bottom + .5 * lmp.at['lmp', c]
+    # ax.bar(1, lmp.at['lmp', c], 0.2, bottom)
+    bottom += lmp.at['lmp', c]
+    ax.annotate(c, xy=(.67, y), xytext=(.9, y-.7), zorder=8,
+                arrowprops=dict(arrowstyle='-'))
+
+ax.set_ylabel('Electricity price  [€/MW]')
+ax.set_xticks([])
+ax.set_yticklabels([])
+
+
+# fig.tight_layout()
+fig.savefig('figures/price_decomposition.png', bbox_inches='tight')

### 2020-05-24 main.tex
add intro, restructure a bit

@@
 
-\section{Introduction}
-
 % In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
@@
 
-\section{Locational Market Price}
+% \section{Introduction}
 
@@
 
-The total system cost $\totalcost$ has to be payed back by the consumers 
+\subsubsection*{Cost Terms of the LMP}
+
+In a power system with minimized total cost $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of consumer $a$ at bus $n$.
+It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
 \begin{align}
- \totalcost =  \sum_{n,a,t} \lmp \, \demand
+ \lmp = \pdv{\,\totalcost}{\,\demand}
 \end{align}
-
+For long term operation and investment planning models, $\totalcost$ includes all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
 \begin{align}
- \totalcost = \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
+  \totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
+  \label{eq:total_cost_terms}
 \end{align}
-
+% 
 \begin{figure}[h]
- \includegraphics[width=\linewidth]{figures/price_decomposition.png}
+ \includegraphics[width=0.9\linewidth]{figures/price_decomposition.png}
+ \caption{Schematic decomposition of the Locational Market Price $\lmp$.
+In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \textit{i.e} OPEX and CAPEX for production and transmission and possibly other expenditures.
+As $\,\totalcost$ is fully payed by the consumers, each LMP $\lmp$ can be decomposed into cost terms compensating the different expenditures at timestep $t$.}
  \label{fig:price_decomposition}
- \caption{The Locational Market Price}
 \end{figure}
 
-\section{Mathematical Theory}
+Over the span of optimized timesteps $t$, the system costs are totally payed back by the consumers 
+\begin{align}
+ \totalcost &=  \sum_{n,a,t} \lmp \, \demand
+ \label{eq:total_cost_sum}
+ \end{align}
+% 
+As a direct consequence of \cref{eq:total_cost_terms,eq:total_cost_sum}, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
+We schematically show this behaviour in \cref{fig:price_decomposition}.
+Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection however leave the question open how the costs are allocated among the components.
+In the following we will show how the above presented cost decomposition is extended to full peer-to-peer cost allocation including all network components.
+
+% \section{Mathematical Theory}
 
@@
 In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus for a given gross production $\generation$ and gross demand $\demand$, they directly link to the the resulting flow on each line,
+Thus for a given production $\generation$ and demand $\demand$, they directly link to the the resulting flow on each line,
 \begin{align}
@@
 \end{align}
-Therefore, the flow can be reproduced from the demand-side of the system only.
+With such a choice the flow can be reproduced from the demand-side of the system only.
 Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
@@
 % 
-It indicates the flow on line $\ell$ provided generators at $m$ and supplying the demand $\nodaldemand$.
+It indicates the flow on line $\ell$ coming from generators at $m$ and supplying the demand $\nodaldemand$.
 When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand};
@@
 % 
-As mentioned, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
+As mentioned before, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
 As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
@@
 \end{align}
-which straightforwardly follows from the fact that $\sum_n \slackk = 1$.
-Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
+which follows from the fact that $\sum_n \slackk = 1$.
+Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
 \\
 
-The solution to $\slackk$ follows directly from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
+The solution to $\slackk$ follows from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
 That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.
-However solution space can be extended to an individual slack for each node $\Slackk$, which we discuss in \cref{sec:solution_space_of_the_slack}.
+However as we discuss in \cref{sec:solution_space_of_the_slack}, the solution space can be extended to an individual slack for each node $\Slackk$.
 
@@
 \begin{align}
-    \underset{\demand, \generation, \capacityGeneration}{\text{max}}
-    \left(\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
+    \underset{\generation, \capacityGeneration, \capacityFlow}{\text{max}}
+    \left(-\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
 \end{align}
@@
 According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration > 0$, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
+If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
 Thus it is these times steps which push investments in $\capacityGeneration$.

### 2020-05-24 examples.py

@@ -128,4 +128,5 @@ for i, v in enumerate(payoff):
     sns.heatmap(df, cmap='Blues', ax=ax, annot=True, fmt='10.0f', cbar=False)
     ax.set_title(to_symbol_dict[v])
 fig.tight_layout()
+fig.savefig(f'figures/example_payoff{tag}.png', bbox_inches='tight')
 

### 2020-05-24 price_decomposition.py

@@ -15,8 +15,9 @@ from mpl_toolkits.axisartist.axislines import SubplotZero
 plt.rc('patch', facecolor='k')
 
 
-cost_i = ['CAPEX Production', 'CAPEX Transmission System',
-          'OPEX Production', 'Other Cost (e.g. CO$_2$ Cost)']
+cost_i = ['CAPEX Production', 'CAPEX Transmission',
+          'OPEX Production', 'Other Cost Terms \n(CO$_2$, Maintenance etc.)']
+
 lmp = pd.DataFrame([[25, 8, 15, 5]], index=['lmp'], columns=cost_i)
 c = pd.Series(color[['one_port_investment_cost', 'branch_investment_cost',
                      'one_port_operational_cost', 'co2_cost']].values,
@@ -46,10 +47,11 @@ for i, c in enumerate(lmp):
     y = bottom + .5 * lmp.at['lmp', c]
     # ax.bar(1, lmp.at['lmp', c], 0.2, bottom)
     bottom += lmp.at['lmp', c]
-    ax.annotate(c, xy=(.67, y), xytext=(.9, y-.7), zorder=8,
-                arrowprops=dict(arrowstyle='-'))
+    ax.annotate(c, xy=(.67, y), xytext=(.9, y), zorder=8,
+                arrowprops=dict(arrowstyle='-'),
+                verticalalignment='center')
 
-ax.set_ylabel('Electricity price  [€/MW]')
+ax.set_ylabel('$\lambda_{n,t}$  [€/MWh]')
 ax.set_xticks([])
 ax.set_yticklabels([])
 

### 2020-05-25 main.tex
update figures

@@
 \newcommand{\vpad}{\vspace{1mm}}
+\newcommand{\hpad}{\hspace{15pt}}
 
@@
 \newcommand{\capacityGeneration}{G_{n,s}}
+\newcommand{\capacityGenerationUpper}{\bar{G}_{n,s}}
+\newcommand{\capacityGenerationLower}{\ubar{G}_{n,s}}
 \newcommand{\capacityFlow}{F_{\ell}}
+\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
+\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
 \newcommand{\capexGeneration}{c_{n,s}}
@@
 
-\subsubsection*{Cost Terms of the LMP}
+\section{Economic Context}
 
-In a power system with minimized total cost $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of consumer $a$ at bus $n$.
-It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
-\begin{align}
- \lmp = \pdv{\,\totalcost}{\,\demand}
-\end{align}
-For long term operation and investment planning models, $\totalcost$ includes all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
 \begin{align}
@@
 \end{align}
+In a network design with minimal $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of any consumer $a$ at bus $n$.
+It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
+\begin{align}
+ \lmp = \pdv{\,\totalcost}{\,\demand}
+ \label{eq:total_cost_derivative}
+\end{align}
+This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are totally payed back by the consumers 
+\begin{align}
+ \totalcost &=  \sum_{n,a,t} \lmp \, \demand
+ \label{eq:total_cost_sum}
+\end{align}
 % 
 \begin{figure}[h]
- \includegraphics[width=0.9\linewidth]{figures/price_decomposition.png}
+\centering
+ \includegraphics[width=.8\linewidth]{figures/price_decomposition.png}
  \caption{Schematic decomposition of the Locational Market Price $\lmp$.
@@
 
-Over the span of optimized timesteps $t$, the system costs are totally payed back by the consumers 
-\begin{align}
- \totalcost &=  \sum_{n,a,t} \lmp \, \demand
- \label{eq:total_cost_sum}
- \end{align}
 % 
-As a direct consequence of \cref{eq:total_cost_terms,eq:total_cost_sum}, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
+As a direct consequence of \cref{eq:total_cost_terms,eq:total_cost_derivative,eq:total_cost_sum}, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
 We schematically show this behaviour in \cref{fig:price_decomposition}.
@@
 
-% \section{Mathematical Theory}
+\section{Mathematical Theory}
 
-\subsubsection*{Power Transfer Distribition Factors and Flow Allocation}
+\subsection{Power Transfer Distribition Factors and Flow Allocation}
 
@@
 
-\subsubsection*{Network Optimisation}
+\subsection{Network Optimisation}
 
@@
 \begin{align}
-    \underset{\generation, \capacityGeneration, \capacityFlow}{\text{max}}
-    \left(-\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
+    \underset{\generation, \capacityGeneration, \capacityFlow}{\text{min}}
+    \left(\sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
 \end{align}
@@
 
-\subsubsection*{Decomposing the Locational Marginal Prices and Allocating the Electricity Costs}
+\subsection{Allocating Nodal Payments}
 \begin{align}
@@
 
-\subsubsection*{Adding CO$_2$ Constraints}
+\subsection{Adding Further Constraints}
+\subsubsection*{CO$_2$ Constraint}
 
@@
 
-% \newpage
+\subsubsection*{Lower and Upper Capacity Limits}
+
+Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of 
+\begin{align}
+ \capacityGeneration \ge \capacityGenerationLower \hpad \forall{n,s \in S} \\
+ \capacityGeneration \le \capacityGenerationUpper \hpad \forall{n,s \in S}
+\end{align}
+or doing likewise with a subset $L$ of transmission lines,
+\begin{align}
+ \capacityFlow \ge \capacityFlowLower \hpad \forall{\ell \in \textit{L}} \\
+ \capacityFlow \le \capacityFlowUpper \hpad \forall{\ell \in \textit{L}}
+\end{align}
+does not have any effect on the above presented allocation.
+The additional cost directly translate t<o the Quality of Supply $\muuppergeneration$ for production and $\muupperflow - \mulowerflow$ for transmission respectively.
+This also counts for limiting the overall production capacity $\sum_{n,s}\capacityGeneration$ or transmission capacity $\sum_\ell \capacityFlow$.
+
+\section{Numerical Example}
+
+\begin{figure*}[h]
+\centering
+ \includegraphics[width=\linewidth]{example_network_constraint_capacity.png}
+ \caption{Illustrative Example of a two bus optimized network with one time step. }
+ \label{fig:example_network}
+\end{figure*}
+
+\begin{figure}[h]
+\centering
+ \includegraphics[width=\linewidth]{example_payoff_constraint_capacity.png}
+ \caption{Full cost allocation for the example setup shown in \cref{fig:example_network}.
+The payments on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}}
+ \label{fig:example_payoff}
+\end{figure}
+
+\newpage
 % \section*{Showcase}

### 2020-05-25 examples.py

@@ -16,18 +16,18 @@ import seaborn as sns
 from config import to_symbol_dict, color
 plt.rc('text', usetex=False)
 
-tag = ''
-add_gen_args = {}
+# tag = ''
+# add_gen_args = {}
 
-# tag = '_constraint_capacity'
-# add_gen_args = {'p_nom_max': 80}
+tag = '_constraint_capacity'
+add_gen_args = {'p_nom_max': 100}
 
 
 n = pypsa.Network()
 n0 = '0'
 n1 = '1'
-d0 = 50
-d1 = 80
+d0 = 60
+d1 = 90
 o0 = 50
 o1 = 200
 c0 = 500
@@ -60,35 +60,44 @@ ca_df = ca_sum.sel(snapshot=sn, drop=True).to_dataframe()
 dc_df = dc.sel(snapshot=sn, drop=True).to_dataframe()
 
 # %%
-fig, ax = plt.subplots()
+fig, ax = plt.subplots(figsize=(12,5))
 
 n.plot(flow=ntl.network_flow(n, sn).to_series()/50, margin=0.3, ax=ax,
        geomap=False, bus_sizes = ca_df.stack().clip(lower=0)/5e6, bus_colors=color)
 bbox = dict(facecolor='w', alpha=0.3, edgecolor='grey', boxstyle='round')
-textkwargs = dict(size=12, color='darkslategray', ha="left", va="top",
+textkwargs = dict(size=12, color='darkslategray', va="top",
                   bbox=bbox, zorder=8)
+wpad = 0.04
+
 bus0_fix = (f'Fixed:  \n\no = {o0} €/MW\nc = {c0} €/MW\nd = {d0} MW')
 bus0_opt = (f'Optimized:  \n\ng = {g0} MW\nG = {G0} MW')
 bus0_pay = (f'Fix quantities:  \n\no = {o0} €/MW\nc = {c0} €/MW\nd = {d0} MW')
-ax.text(n.buses.x[0]-.22, n.buses.y[0]+.26, bus0_fix, **textkwargs)
-ax.text(n.buses.x[0], n.buses.y[0]+.26, bus0_opt, **textkwargs)
+
+dy = .3
+ax.text(n.buses.x[0]-wpad, n.buses.y[0]+dy, bus0_fix, ha='right', **textkwargs)
[…]

### 2020-06-16 code/dist_costs.py
tidy up

@@ -4,6 +4,10 @@
 Created on Thu Feb 20 14:13:56 2020
 
 @author: fabian
+We illustrate the flow based cost allocation under use of the fictive network .
+It consists of nine buses and ten time steps. The solver optimizes the capacity
+of two generators, wind and gas, per bus. ...
+
 """
 
 import pypsa
@@ -41,6 +45,12 @@ n.buses.loc[n.buses.index.str.contains('DC'), 'y'] += 1
 n.buses.loc[n.buses.index.str.contains('DC'), 'x'] -= .5
 
 # %% Network plot
+'''
+Network used for showcasing. (a) shows the distributing of generation capacities
+$\capacityGeneration$, the widths of the transmission lines are proportional to
+their thermal limit $\capacityFlow$. (b) shows the total nodal payments
+according to the cost allocation.
+'''
 
 fig, ax = plt.subplots(subplot_kw={'projection': ccrs.PlateCarree()}, figsize=(6,5))
 bus_sizes = n.generators.groupby(['bus', 'carrier']).p_nom_opt.sum()
@@ -60,6 +70,14 @@ ca = ntl.allocate_cost(n, method=ds, q=0)
 ca = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
 
 # %% Allocation plot
+'''
+Comparison between the flow based cost allocation and the LMP based cost
+per consumer. The left bars consist of the allocated OPEX $\allocateOpex$,
+the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator
+CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$,
+while the right bars show the of the nodal consumption times the LMP.
+'''
+
 
 color = pd.Series({'one_port_operational_cost': 'darkkhaki',
                    'co2_cost': 'tomato',

### 2020-06-16 code/examples.py

@@ -14,7 +14,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 from config import to_symbol_dict, color
-plt.rc('text', usetex=False)
+# plt.rc('text.latex', preamble=r'\Large')
 
 # tag = ''
 # add_gen_args = {}
@@ -24,8 +24,8 @@ add_gen_args = {'p_nom_max': 100}
 
 
 n = pypsa.Network()
-n0 = '0'
-n1 = '1'
+n0 = '1'
+n1 = '2'
 d0 = 60
 d1 = 90
 o0 = 50
@@ -38,19 +38,18 @@ n.madd('Bus', [n0, n1], x=[0, 1], y=[0, 0.5])
 n.madd('Load', [0, 1], bus=[n0, n1], p_set=[d0, d1])
 n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
        marginal_cost=[o0, o1], capital_cost=[c0, c1], **add_gen_args)
-n.madd('Line', [0], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
+n.madd('Line', [1], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
        capital_cost=c_l)
 n.lopf(pyomo=False, keep_shadowprices=True)
 
 sn = 'now'
 
-G0 = int(n.generators.p_nom_opt[0])
-G1 = int(n.generators.p_nom_opt[1])
-F = int(n.lines.s_nom_opt[0])
+G0, G1 = n.generators.p_nom_opt
+F = n.lines.s_nom_opt.item()
+Gupper = add_gen_args.get('p_nom_max', None)
 
-g0 = int(n.generators_t.p['0']['now'])
-g1 = int(n.generators_t.p['1']['now'])
-f = int(n.lines_t.p0['0']['now'])
+g0, g1 = n.generators_t.p.loc['now']
+f = n.lines_t.p0.loc['now'].item()
 
 ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
 dc = ntl.cost.nodal_demand_cost(n).rename(bus='payer')
@@ -65,32 +64,33 @@ fig, ax = plt.subplots(figsize=(12,5))
 n.plot(flow=ntl.network_flow(n, sn).to_series()/50, margin=0.3, ax=ax,
[…]

### 2020-06-18 code/examples.py
Further work on chapter numerical example

@@ -13,13 +13,14 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
-from config import to_symbol_dict, color
-# plt.rc('text.latex', preamble=r'\Large')
+from netallocation.plot import injection_plot_kwargs
+from config import to_static_symbol, color
+# plt.rc('tex', preamble=r'\Large')
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
 
-# tag = ''
-# add_gen_args = {}
 
-tag = '_constraint_capacity'
+tag = ''
 add_gen_args = {'p_nom_max': 100}
 
 
@@ -38,7 +39,7 @@ n.madd('Bus', [n0, n1], x=[0, 1], y=[0, 0.5])
 n.madd('Load', [0, 1], bus=[n0, n1], p_set=[d0, d1])
 n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
        marginal_cost=[o0, o1], capital_cost=[c0, c1], **add_gen_args)
-n.madd('Line', [1], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
+n.madd('Line', ['1'], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
        capital_cost=c_l)
 n.lopf(pyomo=False, keep_shadowprices=True)
 
@@ -50,8 +51,10 @@ Gupper = add_gen_args.get('p_nom_max', None)
 
 g0, g1 = n.generators_t.p.loc['now']
 f = n.lines_t.p0.loc['now'].item()
+l0, l1 = n.buses_t.marginal_price.loc['now']
 
 ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
+ds = ntl.convert.peer_to_peer(ds, n)
 dc = ntl.cost.nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)
 ca_sum = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
@@ -63,13 +66,12 @@ fig, ax = plt.subplots(figsize=(12,5))
 
 n.plot(flow=ntl.network_flow(n, sn).to_series()/50, margin=0.3, ax=ax,
        geomap=False, bus_sizes = ca_df.stack().clip(lower=0)/5e6, bus_colors=color)
[…]

### 2020-06-23 code/config.py
begin chapter localize allocation, restructure appendix

@@ -25,7 +25,7 @@ color = pd.Series({'one_port_operational_cost': 'darkkhaki',
                    'co2_cost': 'tomato',
                    'one_port_investment_cost': 'palevioletred',
                    'branch_investment_cost': 'mediumaquamarine',
-                   'nodal_demand_cost': 'cadetblue'})
+                   'nodal_demand_cost': 'cadetblue'}).sort_index()
 
 latex_names = dict(p_nom_opt = '$G$', marginal_cost = '$o$', capital_cost='$c$',
                    bus='$n$')

### 2020-06-23 code/price_decomposition.py

@@ -57,4 +57,4 @@ ax.set_yticklabels([])
 
 
 # fig.tight_layout()
-fig.savefig('figures/price_decomposition.png', bbox_inches='tight')
+fig.savefig('../figures/price_decomposition.png', bbox_inches='tight')

### 2020-06-27 paper/playaround.tex
play around with tom's approach

--- 2020-05-24 main.tex
@@
-\documentclass[11pt,twocolumn]{article}
+\documentclass[a4paper,10pt]{article}
 \usepackage{graphicx}
@@
 \usepackage[colorlinks]{hyperref}
-\usepackage[backend=biber]{biblatex}
+\usepackage[backend=biber,style=draft]{biblatex}
 \usepackage{eurosym}
@@
 \usepackage{subcaption}
+\usepackage{enumitem} % for alphabetical enumeration 
 \usepackage{accents}
@@
 
-\addbibresource{main.bib}
-\graphicspath{{figures/}}
+%opening
+\title{}
+\author{Fabian Hofmann}
 
-\crefname{relation}{Rel.}{Rels.}
-\creflabelformat{relation}{(#2#1#3)}
-\crefname{constraint}{Constr.}{Constrs.}
-\creflabelformat{constraint}{(#2#1#3)}
+\begin{document}
 
-\setlength\parindent{8pt}
+\maketitle
 
+\begin{abstract}
+
+\end{abstract}
+
+% style operators
 \newcommand{\ie}{\textit{i.e.} }
 \newcommand{\eg}{\textit{e.g.} }
-
 \newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
@@
 \newcommand{\vpad}{\vspace{1mm}}
+\newcommand{\hpad}{\hspace{15pt}}
 
+% general symbols
 \newcommand{\generation}[1][n]{g_{#1,s,t}}
@@
 
+% matrix notation
+\newcommand{\incidenceM}{K}
+\newcommand{\flowM}{f}
+\newcommand{\injectionM}{p}
+\newcommand{\slackM}{k}
+\newcommand{\DirectedIncidence}{\mathcal{K}}
+\newcommand{\InverseAPInjection}{\mathcal{J}}
+\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}
+
+% totals
 \newcommand{\totalnetconsumption}{p^{-}_{t}}
@@
 
-\newcommand{\allocatePeer}[1][m \rightarrow n]{A_{#1,t}}
+% allocation quantities
+\newcommand{\allocatePeer}[1][(m, s) \rightarrow n]{A_{#1,t}}
 \newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
@@
 
-\newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
-\newcommand{\impactcapexflow}{\Phi_{\ell,t}}
-
 %math 
@@
 \newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
-\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
+\newcommand{\pdv}[2]{\dfrac{\partial #1}{\partial #2}}
 
-\begin{document}
+The total cost $\totalcost$ are composed of
+\begin{align}
+\totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow  \\
+&= \sum_{n,t} \left(  \pdv{\totalOpexGeneration}{\nodaldemand} + \pdv{\totalCapexGeneration}{\nodaldemand} + \pdv{\totalCapexFlow}{\nodaldemand} \right) \nodaldemand 
+\label{eq:total_cost_terms}
+\end{align}
+% 
+The total OPEX is given by 
+\begin{align}
+ \totalOpexGeneration& = \sum_{m,s,t} \opexGeneration[m]\,  \generation[m] 
+\end{align}
+% 
+The OPEX payments assigned to node $n$ at time step $t$ is given by 
+% 
+\begin{align}
+\allocateOpex  = \pdv{\totalOpexGeneration}{\nodaldemand} \, \nodaldemand & =  \sum_{m,s,t}\opexGeneration[m] \,  \allocatePeer
+\end{align}
+% 
+with $\allocatePeer$ being defined as the power produced by generator $(m, s)$ and consumed at node $n$
+\begin{align}
+ \allocatePeer = \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
+\end{align}
+% 
+The derivative $\partial \generation[m] / \partial \nodaldemand$ is not well-defined but implies a degree of freedom.
+However it is constraint to allocating exactly the amount of power which is produced by generator $(m,s)$, thus
+\begin{align}
+\generation[m] = \sum_n \allocatePeer 
+\label{eq:solution_space_slack}
+\end{align}
+% 
+The direct payment of node $n$ to generator $(m,s)$ is then  
+% 
+\begin{align}
+ \allocateOpex[n \rightarrow (m,s)] &= 
+\opexGeneration[m] \,  \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
+\label{eq:allocate_opexGeneration_detailed}\\
+\end{align}
 
-\title{From Linear Optimization to Transmission Cost Allocation}
-\author{F. Hofmann. T. Brown}
+% \begin{align}
+% \allocateCapexGeneration[n \rightarrow (m,s)] &= 
+% \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
+% \label{eq:allocate_capexGeneration_detailed} \\
+% \allocateCapexFlow[n \rightarrow \ell] &=  
+% \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
+% \label{eq:allocate_capexFlow_detailed}
+% \end{align}
 
-\maketitle
+The CAPEX for the transmission system is given by
 
-\begin{abstract}
-Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and  
-\end{abstract}
+\begin{align}
+ \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow = \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow 
+\end{align}
 
-\section{Introduction}
-
-% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
-We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA.
-This will lead us to a generalized form of peer-to-peer allocations and transmission usage allocations based on the choice of slack.
-In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion.
-By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.
-
-\section{Locational Market Price}
-
-\subsubsection*{Nomenclature}
-
-\begin{table}[h]
-	\centering
-	\begin{tabular}{ll}
-        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
-        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
-        $\generation$ & Electric generation per bus $n$, carrier $s$, \\ & time step $t$  in MW \vpad \\
-        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
-        $\opexGeneration$ & Operational cost (OPEX) in \euro/MW \vpad \\
-        $\capexGeneration$ & Capital Expenditure (CAPEX) in \euro/MW \vpad \\
-        $\capexFlow$  & CAPEX per transmission line $\ell$ in \euro/MW  \vpad \\
-        $\capacityGeneration$ & Generation capacity in MW \vpad \\
-        $\capacityFlow$ & Transmission capacity in MW \vpad \\
-        $\incidence$ & Incidence matrix \vpad 
-	\end{tabular}
-\end{table}
-
-The total system cost $\totalcost$ has to be payed back by the consumers 
+The individual cost terms are 
 \begin{align}
- \totalcost =  \sum_{n,a,t} \lmp \, \demand
+\allocateCapexFlow = \pdv{\totalCapexFlow}{\nodaldemand}\nodaldemand & = \pdv{\left( \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow \right) }{\nodaldemand} \\
+&= \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf 
 \end{align}
@@
 \begin{align}
- \totalcost = \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
+ \pdv{\muupperflow}{\nodaldemand} = 0 \;\; ?
 \end{align}
 
-\begin{figure}[h]
- \includegraphics[width=\linewidth]{figures/price_decomposition.png}
- \label{fig:price_decomposition}
- \caption{The Locational Market Price}
-\end{figure}
-
-\section{Mathematical Theory}
-
-\subsubsection*{Power Transfer Distribition Factors and Flow Allocation}
-
-In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus for a given gross production $\generation$ and gross demand $\demand$, they directly link to the the resulting flow on each line,
 \begin{align}
- \flow  =   \sum_n \ptdf  \left( \nodalgeneration- \nodaldemand \right)    
- \label{eq:flow_from_ptdf}
+\pdv{\lagrangian}{\demand} & = \pdv{\lagrangian}{\muupperflow} \,  \pdv{\muupperflow} {\nodaldemand} \\
+&= \left( \flow - \capacityFlow \right)  \,  \pdv{\muupperflow} {\nodaldemand}
 \end{align}
-where $\nodalgeneration = \sum_s \generation$ and $\nodaldemand = \sum_a \demand$ combine all generators $s$ and all comsumers $a$ attached to $n$.
-The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
-It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
-The choice of slack modifies the PTDF according to
-\begin{align}
- \ptdf\left( \slack[m]\right)  = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
- \label{eq:ptdf_slacked}
-\end{align}
-where $\ptdfEqual$ denote the PTDF with equally distributed slack.
-When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
-Summing over all ingoing and outgoing flow changes resulting from a positive injection at $n$ yields again the slack
-\begin{align}
-\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
-\label{eq:slack}
-\end{align}
-Note that $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
-% As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
-Established flow allocation schemes [cite] haved used this degree of freedom in order to allocate power flows and exchanges to market participants.
-Under the assumption that consumers account for all power flows in the grid, the slack is set to $\slack^*$ such that
-\begin{align}
- \flow  = - \sum_n \ptdf\left( \slackk[m]\right) \, \nodaldemand  
- \label{eq:flow_from_demand}
-\end{align}
-Therefore, the flow can be reproduced from the demand-side of the system only.
-Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
-In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add up to $\flow$.
-These subflows, in turn, can be further broken down to contributions for each bus $m$ in the slack, such that we get the subflow of individual $m \rightarrow n$ relations, that is
-\begin{align}
- \allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
- \label{eq:allocate_transaction}
-\end{align}
-% 
-It indicates the flow on line $\ell$ provided generators at $m$ and supplying the demand $\nodaldemand$.
-When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand};
-% \begin{align}
-% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
-% \end{align}
-when summing over all sources and sinks, it yields again the power flow, thus
-\begin{align}
-\flow = \sum_{m,n} \allocateTransaction
-\label{eq:transaction_sum}
-\end{align}
-% 
-As mentioned, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
-As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
-\begin{align}
- \allocatePeer &= \slackk[m] \, \nodaldemand 
-\label{eq:allocate_peer}
-\end{align}
-% 
-Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
-\begin{align}
- \sum_n \allocatePeer = \nodalgeneration[m]
- \label{eq:sum_n_allocate_peer}
-\end{align}
-and summing over all sources yields the nodal demand 
-\begin{align}
- \sum_m \allocatePeer = \nodaldemand
- \label{eq:sum_m_allocate_peer}
-\end{align}
-which straightforwardly follows from the fact that $\sum_n \slackk = 1$.
-Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
-\\
-
-The solution to $\slackk$ follows directly from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
-That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.
-However solution space can be extended to an individual slack for each node $\Slackk$, which we discuss in \cref{sec:solution_space_of_the_slack}.
-
-\subsubsection*{Network Optimisation}
-
-% Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
-% 
-% \begin{align}
-%  \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
-% \end{align}
-% ...
-%
-
-We linearly cost-optimize the capacity and dispatch of a simple power system.
-
-\begin{align}
-    \underset{\demand, \generation, \capacityGeneration}{\text{max}}
-    \left(\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
-\end{align}
-subject to following physical constraints. 
-\\
-
-The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
-\begin{align}
-    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
-    \label[constraint]{eq:nodal_balance_lin}
-\end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
-In a power market this is the \euro/\megawatthour-price which a consumer has to pay.
-Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\
-
-The generation $\generation$ is constraint to its nominal capacity
-\begin{align}
- \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
- \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
- \label[constraint]{eq:lower_generation_capacity_constraint}
- \end{align}
-where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness,
-\begin{align}
-\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
-\label{eq:complementary_slackness_upper_generation} \\
-\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
-\label{eq:complementary_slackness_lower_generation} 
-\end{align}
-are only non-zero if the corresponding constraint is binding. \\
-
-The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
-\begin{align}
- \flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
- \label[constraint]{eq:upper_flow_capacity_constraint} \\
- - \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
- \label[constraint]{eq:lower_flow_capacity_constraint}
-\end{align}
-The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e.
-\cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding.
-The complementary slackness
-\begin{align}
- \muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
- \label{eq:complementary_slackness_upper_flow} \\
- \mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
- \label{eq:complementary_slackness_lower_flow} 
-\end{align}
-set the respective KKT for flows staying below the thermal limit to zero. 
-\\
-
-\subsubsection*{Decomposing the Locational Marginal Prices and Allocating the Electricity Costs}
-\begin{align}
-\notag
- & \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
- \notag
- & - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
- \notag
- &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
- \notag
- &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
- \notag
- &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
- &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
- \label{eq:full_lagrangian}
-\end{align}
-% 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-The global maximum of the Lagrangian requires stationarity with respect to all variables.
-The stationarity of the generation capacity variable leads to
-\begin{align}
- \pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
- \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
- \label{eq:capexGeneration_duality}
-\end{align}
-
-the stationarity of the transmission capacity to
-\begin{align}
- \pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
- \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
- \label{eq:capexFlow_duality}
-\end{align}
-
-and the stationarity of the generation to 
-\begin{align}
- \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
- \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
-\end{align}
-Solving \cref{eq:opex_duality} for the $\lmp$, leads to our first representation for Locational Market Price, which we will refer to as the ``Island Solution``,
-\begin{align}
-\lmp  =  \opexGeneration + \muuppergeneration - \mulowergeneration \Forall{n,s,t}
-\label{eq:lmp1}
-\end{align}
-It connects the LMP directly with the local operational price and prices for the generation capacity constraint.
-However, we can derive a second representation for $\lmp$.
-Starting from the stationarity of the flow
-\begin{align}
- 0 =& \pdv{\lagrangian}{\flow}  \\
- \notag
- 0 =& - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  \\
- & -\sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance
-\end{align}
-and multipying each term with the Power Transfer Distribution Factor $\ptdf$ leaves us with
-
-\begin{align}
-  0 &= - \lmp + \sum_m \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right) \ptdf
-  \label{eq:lmp_lmp1_reduced}
-\end{align}
-According to \cref{eq:slack}, the first term splits into the LMP at $n$ and the LMP weighted with the slack.
-The final term disappears as the $\cycle \, \impedance$ is the kernel of the PTDF $\ptdf$, so $\sum_l \cycle \, \impedance \ptdf = 0$.
-Solving \cref{eq:lmp_lmp1_reduced} for $\lmp$ and replacing $\lmp[m]$ of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to
 
@@
 \notag
-\lmp =&  \sum_m \opexGeneration[m]\, \slack[m]  + \sum_m \left( \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \\
-& -\sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
-\label{eq:lmp2}
-\end{align}
-%  
-% NOTE: \cref{eq:lmp2} builds a starting point for breaking down the market value (MV) 
-% of generator $(m,s)$, defined as
-% \begin{align}
-%  MV_s = \dfrac{\sum_t \generation \, \lmp}{\sum_t \generation}
-% \end{align}
-In contrast to the Island Solution in \cref{eq:lmp1}, this representation connects $\lmp$ with all other prices in the system.
-Moreover, it must hold for any choice of $\slack[m]$ and correspondingly decomposes the LMP to operational prices $\opexGeneration[m]$ and prices for capacity bounds for generators and transmission lines.
-When setting the slack to $\slackk[m] = c \, \generation[m]$, the LMP decomposes to prices of generators proportional to their production plus an extra term for the transmission usage.
-This directly links to the flow allocation presented above.
-Multiplied with the nodal demand $\nodaldemand$ the left hand side of \cref{eq:lmp2} turns into the total payment of $n$ and the right hand side into the different payment allocations,
-\begin{align}
- \lmp \, \nodaldemand = \allocateOpex + \allocateCapexGeneration + \allocateCapexFlow \Forall{n,t}
- \label{eq:lmp_allocation}
-\end{align}
-which we define as  
-\begin{align}
- \allocateOpex &= 
- \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
-\label{eq:allocate_opexGeneration}\\
- \allocateCapexGeneration &= 
- \sum_{m,s} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
-\label{eq:allocate_capexGeneration} \\
- \allocateCapexFlow &=  
- \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
-\label{eq:allocate_capexFlow}
-\end{align}
-\Cref{eq:allocate_opexGeneration} denotes the payments for generators' OPEX, \cref{eq:allocate_capexGeneration} for generators' CAPEX and \cref{eq:allocate_capexFlow} for the transmission system's CAPEX respectively.
-All those payment allocation build up on the phycally allocated flows presented in \cref{eq:allocate_peer,eq:allocate_transaction}.
-Thus with the right choice of slack the power flow allocations directly tranlate into a cost allocation.
-Further, the allocated payments can be considered on a more detailed level as
-\begin{align}
- \allocateOpex[n \rightarrow (m,s)] &= 
- \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
-\label{eq:allocate_opexGeneration_detailed}\\
- \allocateCapexGeneration[n \rightarrow (m,s)] &= 
- \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
-\label{eq:allocate_capexGeneration_detailed} \\
- \allocateCapexFlow[n \rightarrow \ell] &=  
-  \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
-\label{eq:allocate_capexFlow_detailed}
-\end{align}
-%
-
-Let's have a look at the allocated OPEX in \cref{eq:allocate_opexGeneration,eq:allocate_opexGeneration_detailed} first.
-Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly compensate their operational costs.
-The OPEX allocation behaves like P2P tradings between producers and consumers with fixed production prices.
-In this way, the generator $(m,s)$ retrieves the exact amount of money from consumers that it spends on the operation.
-In other words, all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s), thus
-\begin{align}
-\sum_{n} \allocateOpex[n \rightarrow (m,s)] = \opexGeneration \, \generation
-\label{eq:no_profit_opex}
+ \pdv{\lagrangian}{\demand}
+ =&    \sum_{m, s, t} \opexGeneration[m] \Slack  \\
+\notag
+&+ \sum_{m,t} \lmp[m] \left(\sum_\ell \incidence[m] \, \ptdf  - \Slack + \delta_{m,n}  \right)  \\ 
+\notag
+&+ \sum_{n,s,t} \muuppergeneration \generationshare \Slack %  - \sum_{n,s,t} \mulowergeneration \generation  
+\\
+&+ \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf      
+\label{eq:full_lagrangian}
 \end{align}
 
-The CAPEX allocation for generators $\allocateCapexGeneration$ defined in \cref{eq:allocate_capexGeneration,eq:allocate_capexGeneration_detailed}, reveils a similar relation.
-According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration > 0$, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
-Thus it is these times steps which push investments in $\capacityGeneration$.
-If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $(m,s)$ each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$,
+\subsubsection*{Production CAPEX}
+\begin{align}
+ \totalCapexGeneration& = \sum_{n,s} \capexGeneration \capacityGeneration\\
+ &= \sum_{n,t} \muuppergeneration  \generation 
+\end{align}
+
+and taking the derivative
 
 \begin{align}
- \sum_{n,t} \allocateCapexGeneration[n \rightarrow (m,s)] = \capexGeneration \, \capacityGeneration
-\label{eq:no_profit_capex_generation}
-\end{align}
-where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
-So in total, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation, reflecting the zero-profit rule.
-% The non-profit rule can also be seen when looking at the the total revenue per generator, 
-% \begin{align}
-%  \sum_t \lmp \, \generation &= \sum_t \opexGeneration \generation + \muuppergeneration \generation - \mulowergeneration \, \generation &\Forall{n,s} \\
-%  &= \sum_t \opexGeneration \, \generation + \capexGeneration \, \capacityGeneration &\Forall{n,s}
-% \end{align}
-%  where we multiplied both sides of \cref{eq:lmp1} by the generation and using the complementary slackness.  
-\\
-
-The allocation of CAPEX for the transmission system $\allocateCapexFlow$, defined in \cref{eq:allocate_capexFlow,eq:allocate_capexFlow_detailed}, builds up on the KKT variables $\muupperflow$ and $\mulowerflow$.
-Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$.
-Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments.
-Again the sum of all CAPEX payments to line $\ell$ equal the total CAPEX spent, thus
-
-\begin{align}
- \sum_{n,t} \allocateCapexFlow[n \rightarrow \ell] = \capexFlow \, \capacityFlow  
-\end{align}
-where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow as stated in \cref{eq:transaction_sum}.
-% A similar relation applies for the transmission CAPEX.
-Mulplying both side of \cref{eq:opex_duality} by the nodal injection $\sum_s \generation - \sum_a \demand$ and using the complementary slackness ...
-% shows that the CAPEX per line amounts to the total congestion revenue
-% \begin{align}
-% \capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
-% \end{align}
-% The transmission lines carry power from nodes with lower LMP to nodes with higher LMP, the price difference times the transported power is their congestion revenue.
-\note{Sign is different to Tom's paper (there a minus is missing).}
-
-\subsubsection*{Adding CO$_2$ Constraints}
-
-Imposing an additional CO$_2$ constraint limiting the total emission to K,  
-\begin{align}
- \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
- \label[constraint]{eq:co2_constraint}
-\end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
-% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
-% 
-% \begin{align}
-% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
-% \label{eq:non_profit_generator_emission}
-% \end{align}
-% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
-As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
-\begin{align}
-\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
-\end{align}
-This leads to allocated CO$_2$ cost compensation of node $n$ of
- \begin{align}
- \allocateEmissionCost &= \emissionPrice \, \sum_{m,s} \emission[m] \, \generationshare[m] \, \allocatePeer \Forall{n,t} \label{eq:allocate_emissionPrice}
-\end{align}
-which expands the allocation of the electricity cost in \cref{eq:lmp_allocation} to 
-\begin{align}
- \lmp \, \nodaldemand = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration  + \allocateEmissionCost \Forall{n,t}
- \label{eq:lmp_allocation_with_emission}
+\pdv{\totalCapexGeneration}{\nodaldemand} & = \pdv{ \left(  \sum_{m,t} \muuppergeneration[m]  \generation[m] \right) }{\nodaldemand} \\
+&= \sum_{m,t} \generationshare  \muuppergeneration \, \Slack
 \end{align}
 
-% \newpage
-% \section*{Showcase}
-% 
-% \subsection*{Network with CO$_2$ constraint}
-% We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
-It consists of nine buses and ten time steps.
-The solver optimizes the capacity of two generators, wind and gas, per bus.
-...
-% 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{compare_allocation.png}
-%     \caption{Comparison between the flow based cost allocation and the LMP based cost per consumer.
-The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
-}
-%     \label{fig:cost_allocation}
-% \end{figure}
-% 
-% \begin{figure}[h]
-% \begin{subfigure}{.5\textwidth}
-% \centering
-%  \includegraphics[width=\textwidth]{network.png}
-%  \caption{}
-%  \label{fig:network}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}   
-%     \centering
-%     \includegraphics[width=\textwidth]{nodal_payments.png}
-%     \caption{}
-%     \label{fig:nodal_payments}
-% \end{subfigure}
-% \caption{Network used for showcasing.
-(a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$.
-(b) shows the total nodal payments according to the cost allocation.}
-% \end{figure}
-% 
-% 
-% 
-% 
-% 
-% \newpage
-% \subsubsection*{Relaxed CO$_2$ Constraint}
-% 
-% 
-% \begin{figure}[h]
-% \begin{subfigure}{.5\textwidth}
-% \centering
-%  \includegraphics[width=\textwidth]{network_relaxed_co2.png}
-%  \caption{}
-%  \label{fig:network_relaxed_co2}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}   
-%     \centering
-%     \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
-%     \caption{}
-%     \label{fig:nodal_payments_relaxed_co2}
-% \end{subfigure}
-% \caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
-% \end{figure}
-% 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
-%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}.
-Only one time-step $t=6$ determines the allocation of generator CAPEX $\allocateCapexGeneration$, as for all other time-steps \cref{eq:upper_generation_capacity_constraint} is not binding.
-Again note the cost scale difference between time step 6 and all others.}
-%     \label{fig:cost_allocation_relaxed_co2}
-% \end{figure}
-% 
-% \subsection*{How does the cost flow through the network}
-% 
-% \begin{figure}[h]
-%     \begin{subfigure}{.5\textwidth}
-%       \centering
-%       \includegraphics[width=\textwidth]{opex_flow.png}
-%       \label{fig:opex_flow}
-%     \end{subfigure}%
-%     \begin{subfigure}{.5\textwidth}
-%       \centering
-%       \includegraphics[width=\textwidth]{capex_flow.png}
-%       \label{fig:capex_flow}
-%     \end{subfigure}
-%     \caption{}
-%     \label{fig:fig}
-% \end{figure}
-
-\newpage
-\appendix
-
-\section{Appendix}
-
-\subsection{Proof \Cref{eq:allocate_peer}}
-\label{sec:proof_allocate_peer}
-
-\Cref{eq:allocate_peer} follows from summing $\allocateTransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
-\begin{align}
- \allocatePeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocateTransaction \\
- &= \slackk \, \nodaldemand - \sum_\ell \incidence \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  \\
- &= \slackk \, \nodaldemand - \left(  \delta_{n,m} - \dfrac{1}{N} - \delta_{n,n} + \dfrac{1}{N} \right) \slackk[m] \, \nodaldemand  \\
- &= \slackk \, \nodaldemand - \left(  \delta_{n,m} - 1 \right) \slackk[m] \, \nodaldemand  \\
- &= \slackk[m] \, \nodaldemand 
-\label{eq:proof_allocate_peer}
-\end{align}
-where we used \cref{eq:slack} and the fact that the equally distributed slack amounts to $1/N$ for all N nodes in the network.
-
-\subsection{Proof of \Cref{eq:sum_n_allocate_peer}}
-\label{sec:proof_sum_n_allocate_peer}
-The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatePeer$
-\begin{align}
-\sum_m \incidence[m] \, \flow &= - \sum_{m,n} \incidence[m] \, \ptdf \, \nodaldemand \\
-\nodalgeneration[m] - \nodaldemand[m] &= - \delta_{m,n} \, \nodaldemand + \slackk[m] \, \nodaldemand    \\
-\allocatePeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
-\sum_n \allocatePeer &= \nodalgeneration[m]
-\end{align}
-
-\subsection{Solution Space of $\boldsymbol{\slackk[m]}$}
-\label{sec:solution_space_of_the_slack}
-
-All choices of $\slackk[m]$ fulfilling \cref{eq:flow_from_demand} determine the solution space of $\slackk$.
-In the $N \times N$ nodal space this translates to the constraint given by \cref{eq:sum_n_allocate_peer} which denotes
-\begin{align}
- \nodalgeneration[m] = \sum_n \slackk[m] \, \nodaldemand  
- \label{eq:solution_space_slack}
-\end{align}
-Solving for $\slackk[m]$ directly leads to 
-\begin{align}
- \slackk[m] = c \, \generation[m] 
-\end{align}
-where $c$ is the inverse of the total consumption or production $c = 1 / \sum_n \nodaldemand   = 1 / \sum_n \nodalgeneration$. \\
-However \cref{eq:solution_space_slack} has an inherent degree of freedom and can be reformulated as 
-\begin{align}
- \nodalgeneration[m] = \sum_n \Slackk \, \nodaldemand  
- \label{eq:solution_space_slack_extended}
-\end{align}
-where $\Slackk$ denote the individual choice of slack for each bus $n$.
-For example the Marginal Participation cite[] algorithm takes only net injections into account.
-This sets the individual slack to
-\begin{align}
-\Slackk &= \dfrac{ \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]}{\nodaldemand}
-\end{align}
-where 
-\begin{itemize}
-%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
- \item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
- \item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
- \item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
- \item $\gamma_t = \dfrac{1}{\sum_n \netproduction} = \dfrac{1}{\sum_n \netconsumption}$ is the inverse of the total injected/extracted power at time $t$.
-\end{itemize}
-
-% \bibliographystyle{ieeetr}
-\printbibliography
-
 \end{document}

### 2020-06-27 code/constraint_matrix.py

@@ -1,0 +1,210 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jun 24 11:00:12 2020
+
+@author: fabian
+"""
+
+import pypsa
+from pypsa.linopt import set_int_index
+import pandas as pd
+import netallocation as ntl
+from scipy.sparse.linalg import inv as spinv
+from scipy.sparse import diags
+from numpy.linalg import inv
+from xarray import DataArray
+import numpy as np
+from numpy.random import rand
+from numpy.testing import assert_allclose
+from pandas import DataFrame, Series
+from numpy import array
+from pypsa.linopt import get_var, get_con
+from pandas import concat
+
+# %% create example
+
+n0 = '1'
+d0 = 60
+d1 = 131
+o0 = 52
+o1 = 101
+o2 = 156
+c0 = 1060
+c1 = 1050
+c2 = 1020
+# c_l = 100
+
+n = pypsa.Network()
+
+# old example
+# n.madd('Bus', [n0, n1], x=[0, 1], y=[0, 0.4])
+# n.madd('Load', [0, 1], bus=[n0, n1], p_set=[d0, d1])
+# n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
+#         marginal_cost=[o0, o1], capital_cost=[c0, c1], p_nom_max=100)
+# n.madd('Line', ['1'], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
+#         capital_cost=c_l)
+
+
+# 1. example
+# n.set_snapshots(list(range(2)))
+# n.madd('Bus', [n0])
+# n.madd('Load', [0], bus=[n0], p_set=pd.DataFrame({0: [d0, d1]}))
+# n.madd('Generator', [0, 1, 2], bus=n0, p_nom_extendable=False,
+#         marginal_cost=[o0, o1, o2], capital_cost=[c0, c1, c2], p_nom=[80, 51, 40],
+#         p_min_pu=pd.DataFrame(rand(2,3)/100, range(2), range(3)),
+#         p_nom_min = rand(3)/100,
+#         carrier=list('abc'))
+
+# 2. example
+n.set_snapshots(list(range(2)))
+n.madd('Bus', [n0])
+n.madd('Load', [0], bus=[n0], p_set=pd.DataFrame({0: [d0, d1]}))
+n.madd('Generator', [0, 1, 2], bus=n0, p_nom_extendable=True,
[…]

### 2020-07-07 paper/cost_allocation_tom.tex
split paper in two, flow-based approach and constraint matrix approach

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% Title should be 75 characters
%
%Decreasing market value of variable renewables a result to policy and not variability
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

%\documentclass[sort&compress,preprint,review,3p]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%\documentclass[final,3p,times]{elsarticle}
\documentclass[final,3p,times]{elsarticle}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
\graphicspath{{graphics/}}

\DeclareGraphicsExtensions{.pdf,.jpeg,.png}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{float}

\usepackage[normalem]{ulem}

\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{threeparttable}
% \usepackage{siunitx}

\usepackage{url}
\usepackage[colorlinks=true, citecolor=blue, linkcolor=blue, filecolor=blue,urlcolor=blue]{hyperref}

\usepackage[gen]{eurosym}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
%\usepackage{lineno}

\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

%use this to add space between rows
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\def\co{CO${}_2$}
\def\el{${}_{\textrm{el}}$}
\def\th{${}_{\textrm{th}}$}

\newcommand{\ubar}[1]{\text{\b{$#1$}}}

\def\l{\lambda}
\def\K{\kappa}
\def\m{\mu}
\def\G{\Gamma}
\def\d{\partial}
\def\cL{\mathcal{L}}

\newcommand*\rot{\rotatebox{90}}
\newcommand*\OK{\ding{51}}

\usepackage{tikz}

\usepackage[europeanresistors,americaninductors]{circuitikz}
\usepackage{adjustbox}

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor under-represents over-representation}

\journal{}

\begin{document}
%\linenumbers
\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Cost allocation from KKT}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[kit]{T.~Brown\corref{cor1}}
\ead{tom.brown@kit.edu}

\cortext[cor1]{Corresponding author}
\address[kit]{Institute for Automation and Applied Informatics, Karlsruhe Institute of Technology, Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}

\begin{abstract}
  Blah.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
%Here are some suggestions:
 renewable energy policy \sep storage \sep large-scale integration of renewable power generation

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

\tableofcontents

\section{General optimisation problem}\label{sec:theory}

Start with a generic linear objective function over $N$ variables $x_i$
\begin{equation}
 \min_{x_i} f(x) =  \min_{x_i}  \sum_{i=1}^N c_i x_i
\end{equation}
such that they respect linear inequality constraints
\begin{equation}
  \sum_i A_{ji} x_i \leq d_j \hspace{1cm} \m_j \hspace{1cm} j=1,\dots M
\end{equation}
    [Linear equality constraints $\sum_i b_i x_i = c$ with $\lambda$ can be replaced by two inequalities $\leq c$, $\geq c$ with $\lambda = \bar{\m} - \ubar{\m}$.]

A subset $B$ of the inequality constraints will be binding at the
optimum point $x^*$, i.e. for $j\in B$, $\sum_i A_{ji} x^*_i = d_j$. We
write $A'_{ji}$ for the matrix that only runs over $j\in B$. For
non-degenerate solutions these will be enough binding constraints to
solve for a unique optimum $x^*$, i.e. $|B| = N \leq M$, so that $A'$ is an $N\times N$
square matrix.  Therefore we can invert the saturation equation to get
\begin{equation}
  x^*_i = \sum_{k\in B} A^{\prime -1}_{ik} d_k
\end{equation}
It might seem that we've now lost all information about the objective function in this expression for $x^*$, but remember that it is the objective function which determines which constraints are binding, i.e. which vertex of the feasible simplex is the optimum.

If we alter the problem to $d_j \to d_j + \varepsilon$ then the new optimum $x^{\varepsilon*}$ is related to the old by
\begin{equation}
  f(x^*) \to f(x^{\varepsilon*}) = f(x^*) + \mu_j^* \varepsilon
\end{equation}
If constraint $j$ was binding such that $\mu_j^* \geq 0$ then we have
\begin{equation}
  x^{\varepsilon*}_i = \sum_{k\in B} A^{\prime -1}_{ik} \left(d_k + \varepsilon \delta_{jk}\right) = x^*_i + \varepsilon A^{\prime -1}_{ij}
\end{equation}

Now we see that
\begin{equation}
  f(x^{\varepsilon*}) = f(x^*) + \mu_j^* \varepsilon = \sum_i c_i  x^{\varepsilon*}_i = \sum_i c_i \left(x^*_i + \varepsilon A^{\prime -1}_{ij}\right) = f(x^*) +   \varepsilon \sum_i c_i  A^{\prime -1}_{ij}
\end{equation}

So that
\begin{equation}
\mu_j^* = \sum_i c_i  A^{\prime -1}_{ij}
\end{equation}
This is nothing other than the solution of the KKT stationarity equations
\begin{equation}
 0 =    \frac{\d \cL}{\d x_i} = c_i - \sum_j \mu_j^* A_{ji}
\end{equation}
where we have discarded the non-binding $j$ to solve for the binding $j\in B$ with $\mu_j^*\geq 0$.

This is related to strong duality since
\begin{equation}
  \sum_{j=1}^M \mu_j^* d_j = \sum_{j\in B} \mu_j^* d_j = \sum_{i,j\in B} c_i  A^{\prime -1}_{ij} d_j = \sum_i c_i x^*_i
\end{equation}

[Normally strong duality is proved the other way around:]
\begin{equation}
 \sum_i c_i x^*_i = \sum_{i,j} x^*_i  \mu_j^* A_{ji} = \sum_{j} \mu_j^* d_j
\end{equation}
[First equality is stationarity, second is complementary slackness.]

\subsection{Example 1: Electricity market with 2 generators}

Consider an optimisation problem for a single period
\begin{equation}
  \min_{g_s} \sum_s o_s g_s
\end{equation}
such that
\begin{align}
  \sum_s g_s & = d  \hspace{1cm} \lambda \\
  g_s & \leq G_s  \hspace{1cm} \bar{\mu}_s \\
  -g_s & \leq 0  \hspace{1cm} \ubar{\mu}_s
\end{align}
For a 3-generator example, the constraint matrix is for $s\in\{1,2,3\}$
\begin{equation}
  A_{ji}=\left(\begin{matrix}
    1 & 1 & 1\\
    1 & 0 & 0 \\
    -1 & 0 & 0 \\
     0 & 1 & 0 \\
     0 & -1 & 0 \\
     0 & 0 & 1 \\
     0 & 0 & -1
  \end{matrix}\right)
\end{equation}
and the constant term is
\begin{equation}
d_{j}=\left(\begin{matrix}
d  & G_1 & 0 & G_2 &  0 & G_3 & 0
\end{matrix}\right)
\end{equation}

Suppose $G_1 < d < G_1+G_2$ so that generator 2 is setting the price. Then constraints $j\in \{1,2,7\}$ are binding and
\begin{equation}
A'_{ji}=\left(\begin{matrix}
 1 & 1 & 1\\
 1 & 0 & 0\\
 0 & 0 & -1
\end{matrix}\right)
\end{equation}
The inverse is
\begin{equation}
 A^{\prime -1}_{ij}
=\left(\begin{matrix}
 0 & 1 & 0\\
 1 & -1 & 1\\
 0 & 0 & -1
\end{matrix}\right)
\end{equation}
So we get
\begin{equation}
g^*_i = \sum_j A^{\prime -1}_{ij} d_j
=\left(\begin{matrix}
 0 & 1 & 0\\
 1 & -1 & 1\\
 0 & 0 & -1
\end{matrix}\right)\left(\begin{matrix}
 d\\
 G_1\\
 0
\end{matrix}\right) = \left(\begin{matrix}
 G_1\\
 d-G_1\\
 0
\end{matrix}\right)
\end{equation}
and
\begin{equation}
\m^*_j =\left(\begin{matrix}
 \m^*_1\\
 \m^*_2\\
 \m^*_7
\end{matrix}\right) =\left(\begin{matrix}
 \l^*\\
 \bar{\m}^*_1\\
 \ubar{\m}^*_3
\end{matrix}\right) = \sum_i c_i A^{\prime -1}_{ij}
= \left(\begin{matrix}
 o_1 &
 o_2 &
 o_3
\end{matrix}\right) \left(\begin{matrix}
 0 & 1 & 0\\
 1 & -1 & 1\\
 0 & 0 & -1
\end{matrix}\right) = \left(\begin{matrix}
 o_2\\
 o_1-o_2\\
 o_2-o_3
\end{matrix}\right)
\end{equation}

The market price is $\l^* = o_2$, set by generator 2.

Note that everything here is rather sparse.

\subsection{Example 2: Electricity market with investment in 1 generator, 2 periods}

Consider an optimisation problem with a single generator, investment, and 2 periods such that $d_1> d_2$
\begin{equation}
  \min_{g_1,g_2,G} \left[ cG + o(g_1 + g_2)\right]
\end{equation}
such that
\begin{align}
  g_1 & = d_1  \hspace{1cm} \lambda_1 \\
  g_2 & = d_2  \hspace{1cm} \lambda_2 \\
  g_t - G & \leq 0  \hspace{1cm} \bar{\mu}_t \\
  -g_t & \leq 0  \hspace{1cm} \ubar{\mu}_t
\end{align}
The constraint matrix is
\begin{equation}
  A_{ji}=\left(\begin{matrix}
    1 & 0 & 0\\
    0 & 1 & 0 \\
    1 & 0 & -1 \\
     -1 & 0 & 0 \\
     0 & 1 & -1 \\
     0 & -1 & 0
  \end{matrix}\right)
\end{equation}
and the constant term is
\begin{equation}
d_{j}=\left(\begin{matrix}
d_1  & d_2 & 0 & 0 &  0 & 0
\end{matrix}\right)
\end{equation}

Suppose $d_1 > d_2$ so that $g_1^* = d_1 = G^*$ and $g_2^* = d_2 < G^*$. Then constraints $j\in \{1,2,3\}$ are binding and
\begin{equation}
A'_{ji}=\left(\begin{matrix}
 1 & 0 & 0\\
 0 & 1 & 0\\
 1 & 0 & -1
\end{matrix}\right)
\end{equation}
The inverse is
\begin{equation}
 A^{\prime -1}_{ij}
=\left(\begin{matrix}
 1 & 0 & 0\\
 0 & 1 & 0\\
 1 & 0 & -1
\end{matrix}\right)
\end{equation}
So we get
\begin{equation}
x^*_i = \left(\begin{matrix}
 g^*_1\\
 g^*_2\\
 G^*
\end{matrix}\right) = \sum_j A^{\prime -1}_{ij} d_j
=\left(\begin{matrix}
 1 & 0 & 0\\
 0 & 1 & 0\\
 1 & 0 & -1
\end{matrix}\right)\left(\begin{matrix}
 d_1\\
 d_2\\
 0
\end{matrix}\right) = \left(\begin{matrix}
 d_1\\
 d_2\\
 d_1
\end{matrix}\right)
\end{equation}
and
\begin{equation}
\m^*_j =\left(\begin{matrix}
 \m^*_1\\
 \m^*_2\\
 \m^*_3
\end{matrix}\right) =\left(\begin{matrix}
 \l_1^*\\
 \l_2^*\\
 \bar{\m}^*_1
\end{matrix}\right) = \sum_i c_i A^{\prime -1}_{ij}
= \left(\begin{matrix}
 o &
 o &
 c
\end{matrix}\right) \left(\begin{matrix}
 1 & 0 & 0\\
 0 & 1 & 0\\
 1 & 0 & -1
\end{matrix}\right) = \left(\begin{matrix}
 o + c \\
 o\\
 -c
\end{matrix}\right)
\end{equation}

\section{Application to cost allocation}

Now suppose we have our big objective function over $x_i = (g_{st},G_s,f_{\ell t}, F_\ell)$
\begin{equation}
  \min_{ x_i} \sum_i c_i x_i
\end{equation}
such that
\begin{align}
  \sum_s g_{s t}K_{ns}-\sum_\ell f_{\ell t}K_{n\ell}  & =  d_{n,t}  \hspace{1cm} \lambda_{n,t} \\
  g_{s t} - G_s & \leq 0  \hspace{1cm} \bar{\mu}_{s,t} \\
  -g_{s t} & \leq 0  \hspace{1cm} \ubar{\mu}_{s,t} \\
  f_{\ell t} - F_\ell & \leq 0  \hspace{1cm} \bar{\mu}_{\ell,t} \\
  -f_{\ell t} - F_\ell & \leq 0  \hspace{1cm} \ubar{\mu}_{\ell,t} \\
  \sum_\ell C_{\ell c}x_\ell f_{\ell t} & = 0 \hspace{1cm} \lambda_{c,t}
\end{align}
$K_{ns}$ is 1 if generator $s$ is at node $n$, 0 otherwise.

From strong duality we have cost recovery from the demand payments
\begin{equation}
  \sum_i c_i x^*_i = \sum_j d_j \mu^*_j = \sum_{n,t} \lambda_{n,t}^* d_{n,t}
\end{equation}

We now want to disaggregate the sum $ \sum_{n,t} \lambda_{n,t}^* d_{n,t}$ and find the fractions $r^i_{n,t}$ of the dispatch/capacity $x^*_i$ apportioned to each demand in each hour such that the payments in each hour covers it
\begin{equation}
  \lambda_{n,t}^* d_{n,t} = \sum_i  c_i r^i_{n,t}
\end{equation}
This must satisfy for each $i$
\begin{equation}
  \sum_{n,t} r^i_{n,t}  = x^*_i
\end{equation}

But we have a very natural way of doing this from Section \ref{sec:theory}!

Find the $N\times N$ matrix $A'_{ji}$ of binding constraints $j\in B$. We have for the investment variables:
\begin{equation}
x^*_i = \sum_{j\in B} A^{\prime -1}_{ij} d_j
\end{equation}
But the only non-zero constants $d_j$ are the demands $d_{n,t}$ so we get
\begin{equation}
x^*_i = \sum_{n,t} A^{\prime -1}_{i(n,t)} d_{n,t}
\end{equation}
(NB: The energy balance constraints will always be binding.)

This gives us a very natural definition of the $r^i_{n,t}$
\begin{equation}
  r^i_{n,t} = A^{\prime -1}_{i(n,t)} d_{n,t}
\end{equation}

Now we have
\begin{equation}
   \sum_i c_i r^i_{n,t} = \sum_i A^{\prime -1}_{i(n,t)} d_{n,t} c_i
\end{equation}
This is none other than our market price since remember
\begin{equation}
\mu_j^* = \sum_i c_i  A^{\prime -1}_{ij}
\end{equation}
Extract the term for the market prices
\begin{equation}
\lambda_{nt}^* = \sum_i c_i  A^{\prime -1}_{i(nt)} \label{eq:pricetocost}
\end{equation}
Thus we get
\begin{equation}
   \sum_i c_i r^i_{n,t} = \sum_i A^{\prime -1}_{i(n,t)} d_{n,t} c_i = \lambda_{nt}^* d_{n,t}
\end{equation}

This gives a unique way of extracting the contribution of revenue at node $n$ and time $t$ to the costs of the system.

The difficulty in practical applications is computing $A^{\prime -1}_{ij}$ (see section below).

Hopefully it is so sparse that the solutions are often obvious.

\subsection{Interpretation}

The critical factor is $A^{\prime -1}_{i(nt)}$ which tells us the share of cost-causing asset $i$ that accrues to node $n$ at time $t$.

Equation \eqref{eq:pricetocost} shows us that the price is set by this factor, so that whatever the marginal asset is that is setting the price is where the cost allocation goes. This is natural.

If generator $s$ is setting the price $\lambda^*_{n,t}$ at node $n$ and time $t$, then we should have for $i$ for $g_{s,t}$ that $A^{\prime -1}_{(st)(nt)} = 1$ and other values are zero for this $n$ and $t$, though I'm wary of this right now, because there could be complex linear combinations of $o_s$ and $c_s$.

\subsection{Calculation}

Note that we're interested only in the $A^{\prime -1}_{i(nt)}$, not in $A^{\prime -1}_{ij}$ for other $j$.

Step 1: Find the binding constraints $j\in B$ (based on examining Gurobi solution), confirm that $|B| = N$ (you may need random numbers to force a unique solution to the problem) and build the sparse matrix $A^{\prime}_{ji}$.

Step 2: For each pair $(n,t)$ solve the following sparse linear system of equations to find $y_i$:
\begin{equation}
  \sum_i A^{\prime}_{ji} y_i = \delta_{j(n,t)}
\end{equation}
The solution will be
\begin{equation}
  y_i =  A^{\prime -1}_{i(n,t)}
\end{equation}

Done!

\subsection{Application away from long-term equilibrium}

The method is quite general, so should also apply even if we fix $G_s$, $F_\ell$?

We could introduce costs $o_{\ell t}$ for each direction of the link flows $f_{\ell t}$ to track who uses them.

If we only have operation variables being optimised $g_{st}$ and $f_{\ell t}$ then we still get a factor $A^{\prime -1}_{i(nt)}$ to use for the cost allocation of usage, which we can then use to divide up the fixed costs of $G_s$ and $F_\ell$? But is the interpretation of $A^{\prime -1}_{i(nt)}$ correct as marginal cost or full cost?

With these extra constants we move the constant to the RHS now, e.g.
\begin{align}
  g_{st}  & \leq G_s  \hspace{1cm} \bar{\mu}_{st}
\end{align}

Now we have for binding cases $g_{st}^* = G_s$ ($\bar{\mu}_{st}^* \leq 0$ by definition)
\begin{equation}
  \bar{\mu}_{st}^* = \sum_i c_i  A^{\prime -1}_{i(st)}
\end{equation}
So this just gives the combination of costs that reproduce the scarcity value of the limitation $G_s$.

And for the strong duality
\begin{equation}
  \sum_i c_i x_i^* = \sum_{n,t} \lambda^*_{nt} d_{nt} + \sum_{s,t} \bar{\mu}^*_{st} G_{s}
\end{equation}
Since the second term is negative, we get the result that demand payments $ \sum_{n,t} \lambda^*_{nt} d_{nt}$ cover both operational costs $\sum_i c_i x_i^*$ and the scarcity value of the generator capacities $-\sum_{s,t} \bar{\mu}^*_{st} G_{s}$ which help to pay down their capital costs.

\bibliographystyle{elsarticle-num}

\biboptions{sort&compress}
\bibliography{kkt_flow_tracing}

\end{document}

### 2020-07-07 paper/playaround.tex
split paper in two, flow-based approach and constraint matrix approach

@@
 % general symbols
-\newcommand{\generation}[1][n]{g_{#1,s,t}}
-\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
-\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
-\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
-\newcommand{\capacityGeneration}{G_{n,s}}
-\newcommand{\capacityGenerationUpper}{\bar{G}_{n,s}}
-\newcommand{\capacityGenerationLower}{\ubar{G}_{n,s}}
+\newcommand{\generation}{g_{s,t}}
+\newcommand{\generationpotential}{\bar{g}_{s,t}}
+\newcommand{\generationshare}{\omega_{s,t}}
+\newcommand{\incidenceGenerator}{K_{n,s}}
+\newcommand{\nodalgeneration}{\incidenceGenerator \, \generation}
+\newcommand{\capacityGeneration}{G_{s}}
+\newcommand{\capacityGenerationUpper}{\bar{G}_{s}}
+\newcommand{\capacityGenerationLower}{\ubar{G}_{s}}
 \newcommand{\capacityFlow}{F_{\ell}}
@@
 \newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
-\newcommand{\capexGeneration}{c_{n,s}}
+\newcommand{\capexGeneration}{c_{s}}
 \newcommand{\capexFlow}{c_{\ell}}
-\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
-\newcommand{\demand}[1][n]{d_{#1,a,t}}
-\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
-\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
+\newcommand{\opexGeneration}{o_{s}}
+\newcommand{\demand}{d_{n,t}}
+\newcommand{\demandshare}{\omega_{a,t}}
 \newcommand{\utility}{U_{n,a,t}}
@@
 \newcommand{\Slackk}{k^*_{m,n}}
-\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
-\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
+\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
+\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
 \newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
@@
 % allocation quantities
-\newcommand{\allocatePeer}[1][(m, s) \rightarrow n]{A_{#1,t}}
+\newcommand{\allocatePeer}[1][s \rightarrow n]{A_{#1,t}}
 \newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
-\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocateTransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
 \newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
@@
 \newcommand{\totalEmissionCost}{\mathcal{E}}
+\newcommand{\totalRest}{\mathcal{R}}
+\newcommand{\totalDemand}{\mathcal{D}}
 
@@
 
-The total cost $\totalcost$ are composed of
+\section{Optimization}
+
+\def\l{\lambda}
+\def\K{\kappa}
+\def\m{\mu}
+\def\G{\Gamma}
+\def\d{\partial}
+\def\cL{\mathcal{L}}
+
+Start with a generic linear objective function over $N$ variables $x_i$
+\begin{equation}
+ \min_{x_i} f(x) =  \min_{x_i}  \sum_{i=1}^N c_i x_i
+\end{equation}
+such that they respect linear inequality constraints
+\begin{equation}
+  \sum_i A_{ji} x_i \leq d_j \hspace{1cm} \m_j \hspace{1cm} j=1,\dots M
+\end{equation}
+    [Linear equality constraints $\sum_i b_i x_i = c$ with $\lambda$ can be replaced by two inequalities $\leq c$, $\geq c$ with $\lambda = \bar{\m} - \ubar{\m}$.]
+
+A subset $B$ of the inequality constraints will be binding at the
+optimum point $x^*$, i.e. for $j\in B$, $\sum_i A_{ji} x^*_i = d_j$. We
+write $A'_{ji}$ for the matrix that only runs over $j\in B$. For
+non-degenerate solutions these will be enough binding constraints to
+solve for a unique optimum $x^*$, i.e. $|B| = N \leq M$, so that $A'$ is an $N\times N$
+square matrix.  Therefore we can invert the saturation equation to get
+\begin{equation}
+  x^*_i = \sum_{k\in B} A^{\prime -1}_{ik} d_k
+\end{equation}
+It might seem that we've now lost all information about the objective function in this expression for $x^*$, but remember that it is the objective function which determines which constraints are binding, i.e.
+which vertex of the feasible simplex is the optimum.
+
+\section{Power system}
+
+We consider a power system with minimized total cost $\totalcost$ being subject to supplying demand $\demand$ at node $n$ at time $t$. 
+The change in total cost against variation of the nodal demand
 \begin{align}
-\totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow  \\
-&= \sum_{n,t} \left(  \pdv{\totalOpexGeneration}{\nodaldemand} + \pdv{\totalCapexGeneration}{\nodaldemand} + \pdv{\totalCapexFlow}{\nodaldemand} \right) \nodaldemand 
+\lmp  =  \pdv{\totalcost}{\demand}
+\end{align}
+defines the Locational Marginal Price $\lmp$ that consumers at node $n$ and time $t$. 
+The total cost $\totalcost$ splits into two cost terms
+\begin{align}
+\totalcost &= \totalDemand +  \totalRest
 \label{eq:total_cost_terms}
 \end{align}
-% 
+where $\totalDemand = \sum_i \totalDemand_i $ combines cost all terms which are sensitive to changes in the nodal demand 
+\begin{align}
+ \pdv{\totalDemand_i}{\demand} \ne 0
+\end{align}
+and thus determine the LMP,
+\begin{align}
+\lmp = \pdv{\totalDemand}{\demand} 
+\end{align}
+and $\totalRest = \sum_i \totalRest_i$ combines cost terms which is not, thus 
+\begin{align}
+ 0 = \pdv{\totalRest_i}{\demand} \Forall{i}
+\end{align}
+Due to strong duality the total cost $\totalcost$ minus the exogeous cost term $\totalRest$ is totally payed back by the consumers
+\begin{align}
+ \totalcost - \totalRest = \sum_{n,t} \lmp \demand
+\end{align}
+
+\subsubsection*{Allocate OPEX}
+
 The total OPEX is given by 
 \begin{align}
- \totalOpexGeneration& = \sum_{m,s,t} \opexGeneration[m]\,  \generation[m] 
+ \totalOpexGeneration& = \sum_{s,t} \opexGeneration\,  \generation 
+\label{eq:opex_generation}
+ \end{align}
+ We want to disaggregate the totel OPEX into contribution of bus n at time 
+ \begin{align}
+  \totalOpexGeneration = \sum_{n,t} \allocateOpex
+ \end{align}
+which are defined as 
+\begin{align}
+\allocateOpex  = \pdv{\totalOpexGeneration}{\demand} \, \demand & =  \sum_{s,t}\opexGeneration \, \pdv{\generation}{\demand} \, \demand
 \end{align}
 % 
-The OPEX payments assigned to node $n$ at time step $t$ is given by 
+From feeding this back into \cref{eq:opex_generation} we see that the derivative $\partial \generation / \partial \demand$ has to fulfill
+\begin{align}
+\generation = \sum_n \pdv{\generation}{\demand} \, \demand
+\end{align}
+Where the term at the right hand side denote the power produced by generator $s$ and consumed at node $n$, which we define as
+\begin{align}
+ \allocatePeer = \pdv{\generation}{\demand} \, \demand
+ \label{eq:allocate_peer}
+\end{align}
+% 
+The direct payment of node $n$ to generator $s$ is then  
 % 
 \begin{align}
-\allocateOpex  = \pdv{\totalOpexGeneration}{\nodaldemand} \, \nodaldemand & =  \sum_{m,s,t}\opexGeneration[m] \,  \allocatePeer
-\end{align}
-% 
-with $\allocatePeer$ being defined as the power produced by generator $(m, s)$ and consumed at node $n$
-\begin{align}
- \allocatePeer = \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
-\end{align}
-% 
-The derivative $\partial \generation[m] / \partial \nodaldemand$ is not well-defined but implies a degree of freedom.
-However it is constraint to allocating exactly the amount of power which is produced by generator $(m,s)$, thus
-\begin{align}
-\generation[m] = \sum_n \allocatePeer 
-\label{eq:solution_space_slack}
-\end{align}
-% 
-The direct payment of node $n$ to generator $(m,s)$ is then  
-% 
-\begin{align}
- \allocateOpex[n \rightarrow (m,s)] &= 
-\opexGeneration[m] \,  \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
+ \allocateOpex[n \rightarrow s] &= 
+\opexGeneration \,  \allocatePeer
 \label{eq:allocate_opexGeneration_detailed}\\
@@
 
-% \begin{align}
-% \allocateCapexGeneration[n \rightarrow (m,s)] &= 
-% \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
-% \label{eq:allocate_capexGeneration_detailed} \\
-% \allocateCapexFlow[n \rightarrow \ell] &=  
-% \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
-% \label{eq:allocate_capexFlow_detailed}
-% \end{align}
+\subsubsection*{Allocate Generation CAPEX}
+\begin{align}
+ \totalCapexGeneration& = \sum_{s} \capexGeneration \capacityGeneration 
+\end{align}
+with 
+\begin{align}
+ \capexGeneration = \muuppergeneration \, \generation
+\end{align}
+and \cref{eq:allocate_peer}, we define 
+\begin{align}
+\allocateCapexGeneration = \pdv{\totalCapexGeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatePeer
+\end{align}
+
+\subsubsection*{Allocate Transmission CAPEX}
 
@@
 \end{align}
-
+% 
 The individual cost terms are 
 \begin{align}
-\allocateCapexFlow = \pdv{\totalCapexFlow}{\nodaldemand}\nodaldemand & = \pdv{\left( \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow \right) }{\nodaldemand} \\
-&= \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf 
+\allocateCapexFlow = \pdv{\totalCapexFlow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
 \end{align}
+where we assume $\partial \muupperflow/\partial \demand \cong 0 $, as dual variables have a second order dependeny on variables.
+The marginal change of flow with repect to the nodal demand is again non-unique and requires assumption on where the power is resulting from.
 
 \begin{align}
- \pdv{\muupperflow}{\nodaldemand} = 0 \;\; ?
+ \pdv{\flow}{\demand}\, \demand = \pdv{\flow}{\generation} \pdv{\generation}{\demand} \, \demand  = \allocateTransaction
 \end{align}
-
+fulfilling 
 \begin{align}
-\pdv{\lagrangian}{\demand} & = \pdv{\lagrangian}{\muupperflow} \,  \pdv{\muupperflow} {\nodaldemand} \\
-&= \left( \flow - \capacityFlow \right)  \,  \pdv{\muupperflow} {\nodaldemand}
-\end{align}
-
-\begin{align}
-\notag
- \pdv{\lagrangian}{\demand}
- =&    \sum_{m, s, t} \opexGeneration[m] \Slack  \\
-\notag
-&+ \sum_{m,t} \lmp[m] \left(\sum_\ell \incidence[m] \, \ptdf  - \Slack + \delta_{m,n}  \right)  \\ 
-\notag
-&+ \sum_{n,s,t} \muuppergeneration \generationshare \Slack %  - \sum_{n,s,t} \mulowergeneration \generation  
-\\
-&+ \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf      
-\label{eq:full_lagrangian}
-\end{align}
-
-\subsubsection*{Production CAPEX}
-\begin{align}
- \totalCapexGeneration& = \sum_{n,s} \capexGeneration \capacityGeneration\\
- &= \sum_{n,t} \muuppergeneration  \generation 
-\end{align}
-
-and taking the derivative
-
-\begin{align}
-\pdv{\totalCapexGeneration}{\nodaldemand} & = \pdv{ \left(  \sum_{m,t} \muuppergeneration[m]  \generation[m] \right) }{\nodaldemand} \\
-&= \sum_{m,t} \generationshare  \muuppergeneration \, \Slack
+\sum_{s,n} \allocateTransaction = \flow 
 \end{align}

### 2020-07-07 code/helpers.py

@@ -21,13 +21,16 @@ from pandas import concat
 
 
 
-def add_noise(n, noise_order=1e-4):
+def add_noise(n, noise_order=1e-4, only_static=True):
     for c, attr in pypsa.descriptors.nominal_attrs.items():
         if n.df(c).empty:
             continue
 
         n.df(c)[attr + '_min'] += rand(len(n.df(c))) * noise_order
         n.df(c)[attr + '_max'] += rand(len(n.df(c))) * noise_order
+
+        if only_static:
+            continue
 
         min_pu = attr.replace('nom', 'min_pu')
         max_pu = attr.replace('nom', 'max_pu')
@@ -46,8 +49,8 @@ def add_noise(n, noise_order=1e-4):
             n.pnl(c)[min_pu] += rand(*n.pnl(c)[min_pu].shape) * noise_order
 
 
-def noisy_lopf(n, noise_order=1e-4):
-    add_noise(n, noise_order)
+def noisy_lopf(n, noise_order=1e-4, only_static=False):
+    add_noise(n, noise_order, only_static)
     n.lopf(pyomo=False, keep_shadowprices=True, keep_references=True,
        solver_name='gurobi')
 
@@ -119,10 +122,21 @@ def get_linear_system(n):
     s.d = Series(d[B], cons_index[B])
     s.m = Series(m[B], cons_index[B])
 
-    cost_effective_i = s.d[lambda ds: ds!=0].index
-    s.r = s.A_inv.loc[:, cost_effective_i] * s.d[cost_effective_i]
+    cost_eff_cons_i = s.d[lambda ds: ds!=0].index
+    cost_eff_vars_i = s.c[lambda ds: ds!=0].index
+
+    s.r = s.A_inv.loc[:, cost_eff_cons_i] * s.d[cost_eff_cons_i]
+    s.C = s.r.loc[cost_eff_vars_i].mul(s.c[cost_eff_vars_i], 0)
 
     assert_allclose(s.x, s.r.sum(1), **tolerance)
 
     return s
+
+
+def pinv(df):
+    return pd.DataFrame(np.linalg.pinv(df), *df.T.axes).round(3)
+
+def inv(df):
+    return pd.DataFrame(np.linalg.inv(df), *df.T.axes).round(3)
+
 

### 2020-07-07 code/incremental_change.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Jul  2 15:02:28 2020
+
+@author: fabian
+"""
+
+import networks
+
+n = networks.n_ac_dc()
+
+n.lopf()
+print(n.generators_t.p)
+print(n.objective)
+
+n.loads_t.p_set.iloc[0,0] += 1
+
+n.lopf()
+print(n.generators_t.p)
+print(n.objective)

### 2020-07-10 paper/flow_based.tex
flow based cost allocation: prepare application case

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{enumitem} % for alphabetical enumeration 
\usepackage{accents}
\usepackage[capitalise]{cleveref}

\addbibresource{main.bib}
\graphicspath{{../figures/}}

\crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}

\setlength\parindent{8pt}

% style operators
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}
\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}

% general symbols
\newcommand{\generation}{g_{s,t}}
\newcommand{\generationpotential}{\bar{g}_{s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacityGeneration}{G_{s}}
\newcommand{\capacityGenerationUpper}{\bar{G}_{s}}
\newcommand{\capacityGenerationLower}{\ubar{G}_{s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
\newcommand{\capexGeneration}{c_{s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}{o_{s}}
\newcommand{\demand}[1][n]{d_{#1,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\incidenceGenerators}[1][n]{K_{#1,s}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
\newcommand{\slackflow}{k_{\ell}}
\newcommand{\slack}[1][n]{k_{#1}}
\newcommand{\slackk}[1][n]{k^*_{#1}}
\newcommand{\Slack}{k_{m,n}}
\newcommand{\Slackk}{k^*_{m,n}}
\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\impedance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}
\newcommand{\injection}{p_{n,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}

\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}
\newcommand{\lagrangian}{\mathcal{L}}

% allocation quantities
\newcommand{\allocatePeer}[1][s \rightarrow n]{A_{#1,t}}
\newcommand{\allocateFlow}[1][n]{A_{\ell,#1,t}}
\newcommand{\allocateTransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}

\newcommand{\emission}{e_{s}}
\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{TC}}
\newcommand{\totalOpexGeneration}{\mathcal{O}}
\newcommand{\totalOpexFlow}{\mathcal{O}^F}
\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
\newcommand{\totalCapexFlow}{\mathcal{C}^F}
\newcommand{\totalEmissionCost}{\mathcal{E}}
\newcommand{\totalRest}{\mathcal{R}}

\begin{document}

\title{Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable}
\author{F. Hofmann}

\maketitle

\begin{abstract}
% Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and ... 
\end{abstract}

% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system. We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA. This will lead us to a generalized form of P2P allocations and transmission usage allocations based on the choice of slack. In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion. By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.

\subsection*{Highlights}
\begin{itemize}
 \item The LMP always breaks down into contributions of different cost terms (OPEX, CAPEX, etc.) of the system.
 \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments.
 \item Methods like Average Participation or Flow Based Market Coupling allow for localizing the P2P cost assignments and prioritize assignments within one market zone.
\end{itemize}

% \section{Introduction}

\subsubsection*{Nomenclature}

\begin{table}[h]
    \centering
    \begin{tabular}{ll}
        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
        $\generation$ & Electric generation of generator $s$, \\ & time step $t$  in MW \vpad \\
        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
        $\opexGeneration$ & Operational price in \euro/MW \vpad \\
        $\capexGeneration$ & Capital Price in \euro/MW \vpad \\
        $\capexFlow$  & Capital Price in \euro/MW for transmission \\ & capacity on line $\ell$  \vpad \\
        $\capacityGeneration$ & Generation capacity in MW \vpad \\
        $\capacityFlow$ & Transmission capacity in MW \vpad \\
        $\incidence$ & Incidence matrix \vpad 
    \end{tabular}
\end{table}

\section{Economic Context}

In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
\begin{align}
\totalcost &= \totalOpexGeneration + \totalEmissionCost +  \totalCapexGeneration +  \totalCapexFlow + ...
\label{eq:total_cost_terms}
\end{align}
In a network design with minimal $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of any consumer $a$ at bus $n$. It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
\begin{align}
\lmp = \pdv{\,\totalcost}{\,\demand}
\label{eq:total_cost_derivative}
\end{align}
% 
\begin{figure}[h]
\centering
\includegraphics[width=.8\linewidth]{price_decomposition.png}
\caption{Schematic decomposition of the Locational Market Price $\lmp$. In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \ie OPEX and CAPEX for production and transmission and possibly other expenditures. }
\label{fig:price_decomposition}
\end{figure}
%

This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are partially or totally payed back by the consumers 
\begin{align}
\totalcost - \totalRest &=  \sum_{n,a,t} \lmp \, \demand
\label{eq:total_cost_sum}
\end{align}
depending on the costs $\totalRest$ which are independent of the nodal demand  
\begin{align}
 \pdv{\totalRest}{\demand} = 0
\end{align}
% 
In most cases, where $\totalRest$ plays a minor role, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures. We schematically show this behaviour in \cref{fig:price_decomposition}. 
Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection however leave the question open how the costs are allocated among the components. In the following we will show how the above presented cost decomposition is extended to full peer-to-peer (P2P) cost allocation including all network components.

\section{Nodal Cost Allocation}
\label{sec:theory}

From inserting \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative}, we are able to disaggregate the cost terms $\totalOpexGeneration$, $\totalCapexGeneration$ and $\totalCapexFlow$ into constributions of demand $\demand$ at bus~$n$ and time~$t$. For the CAPEX allocation we will make use of the shadowprices resulting from the cost mimization.

\subsection{Allocate Operation Expenditures}

The total OPEX is given by the operational price $\opexGeneration$ times the generation $\generation$ of all generators $s$, 
\begin{align}
 \totalOpexGeneration& = \sum_{s,t} \opexGeneration\,  \generation 
\label{eq:opex_generation}
 \end{align}
The allocated OPEX $\allocateOpex$ which naturally sum up to $  \totalOpexGeneration = \sum_{n,t} \allocateOpex$, are given by 
\begin{align}
\allocateOpex  = \pdv{\totalOpexGeneration}{\demand} \, \demand & =  \sum_{s,t}\opexGeneration \, \pdv{\generation}{\demand} \, \demand
\end{align}
% 
The derivative on the right hand side time the nodal demand can be interpreted as power produced by generator $s$ and consumed at node $n$. Without further explaining how the derivative is calculated, we define this power allocation as 
\begin{align}
 \allocatePeer = \pdv{\generation}{\demand} \, \demand
 \label{eq:allocate_peer}
\end{align}
% 
which by feeding back into \cref{eq:opex_generation} has to fulfill
\begin{align}
\generation = \sum_n \allocatePeer
\label{eq:allocate_peer_constraint}
\end{align}
% 
% 
The direct payment of node $n$ to generator $s$ is then given by  
% 
\begin{align}
 \allocateOpex[n \rightarrow s] &= 
\opexGeneration \,  \allocatePeer
\label{eq:allocate_opexGeneration_detailed}
\end{align}
Accordingly, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs. The cost allocation behaves like P2P tradings between producers and consumers for fixed production prices. In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation, thus 
\begin{align}
\sum_{n} \allocateOpex[n \rightarrow s] = \opexGeneration \, \generation
\label{eq:no_profit_opex}
\end{align}

\subsection{Allocate Emission Expenditures}

Given a fix price for emissions in CO$_2$ equivalents $\emissionPrice$, the cost term for emission adds up to 
\begin{align}
 \totalEmissionCost = \emissionPrice \, \sum_s  \emission \, \generation
\end{align}
where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
Similar to the allocated OPEX, the emission cost for consumers at bus $n$ at time $t$ is allocated through
\begin{align}
 \allocateEmissionCost = \pdv{\totalEmissionCost}{\demand} \demand = \sum_s \emissionPrice \, \emission \, \allocatePeer  
\end{align}
with its peer-to-peer payments 
\begin{align}
 \allocateEmissionCost[n \rightarrow s] = \emissionPrice \, \emission \, \allocatePeer
\end{align}

\subsection{Allocate Capital Expenditures for Generators}

The optimization of the generation capacity $\capacityGeneration$ adds the cost term 
\begin{align}
 \totalCapexGeneration& = \sum_{s} \capexGeneration \capacityGeneration 
\end{align}
to $\totalcost$. It combines all capacities times their capital cost $\capexGeneration$. The nominal capacity constraints the generation $\generation$ in the form of 
\begin{align}
\generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
\label[constraint]{eq:lower_generation_capacity_constraint}
\end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. Using the result of \cite{brown_decreasing_2020} we show in that the capital cost  for generator $s$ translates is payed back by the generation times the shadow price for the upper capacity constraint, 
\begin{align}
 \capexGeneration \capacityGeneration = \sum_t \muuppergeneration \,  \generation 
 \label{eq:no_profit_capex_generation}
\end{align}
With our definition in \cref{eq:allocate_peer} the allocated CAPEX can then be defined as 
\begin{align}
\allocateCapexGeneration = \pdv{\totalCapexGeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatePeer
\end{align}
with the peer-to-peer allocation 
\begin{align}
 \allocateCapexGeneration[n \rightarrow s] = \muuppergeneration \, \allocatePeer
 \label{eq:allocate_capexGeneration_detailed}
\end{align}

So, how does this allocation behave? According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not. If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding. Thus it is these times steps which push investments in $\capacityGeneration$. If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary. 
When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_peer_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$.

\subsection{Allocate Capital Expenditures for Transmission Lines}

The CAPEX for the transmission system is given by 
\begin{align}
 \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow
 \label{eq:capex_flow}
\end{align}
The transmission capacity variable $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
\begin{align}
\flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
\label[constraint]{eq:upper_flow_capacity_constraint} \\
- \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
\label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
Again, we use the result of \cite{brown_decreasing_2020} which derives that the investment in line $\ell$ is payed back by the shadowprices times the flow 
\begin{align}
\capexFlow \capacityFlow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
\label{eq:no_profit_capex_flow}
\end{align}
We insert this into \cref{eq:capex_flow} and define the individual cost terms  as 
\begin{align}
\allocateCapexFlow = \pdv{\totalCapexFlow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
\end{align}
The marginal change of flow with repect to the nodal demand is again non-unique and requires assumption on where the power is resulting from.

\begin{align}
 \pdv{\flow}{\demand}\, \demand = \allocateFlow
\end{align}
fulfilling 
\begin{align}
\sum_{n} \allocateFlow = \flow 
\label{eq:allocate_transaction_constraint}
\end{align}
% 
The peer-to-peer cost allocation for the transmission CAPEX is then defined as 
\begin{align}
 \allocateCapexFlow[n \rightarrow \ell] &=  
\left( \muupperflow - \mulowerflow\right)  \allocateFlow
\label{eq:allocate_capexFlow_detailed}
\end{align}
% 
Again we can interpret the shadowprices $\muupperflow$ and $\mulowerflow$ as a measure for necessity of transmission investments at $\ell$ at time $t$. Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments. Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent. This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:allocate_transaction_constraint,eq:no_profit_capex_flow}

% \subsection{LMP independent Cost  Terms}
% 
% Example: \\
% Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of 
% \begin{align}
% \capacityGeneration \ge \capacityGenerationLower \hpad \forall{n,s \in S} \\
% \capacityGeneration \le \capacityGenerationUpper \hpad \forall{n,s \in S}
% \end{align}
% or doing likewise with a subset $L$ of transmission lines,
% \begin{align}
% \capacityFlow \ge \capacityFlowLower \hpad \forall{\ell \in \textit{L}} \\
% \capacityFlow \le \capacityFlowUpper \hpad \forall{\ell \in \textit{L}}
% \end{align}
% does not have any effect on the above presented allocation in \cref{eq:lmp_allocation}. The additional cost directly translate to the Quality of Supply $\muuppergeneration$ for production and $\muupperflow - \mulowerflow$ for transmission respectively. This also counts for limiting the overall production capacity $\sum_{n,s}\capacityGeneration$ or transmission capacity $\sum_\ell \capacityFlow$. However it will distort the no profit rule for generators \cref{eq:no_profit_capex_generation} and transmission \cref{eq:no_profit_capex_flow} respectively. As shown in \cite{brown_decreasing_2020}

\section{Assumptions on Power Allocations}
% \section{\texorpdfstring{Extended Solution Space of $\boldsymbol{\slackk[m]}$}{Solution Space of the Slack}}
\label{sec:localizing_allocations}

The P2P cost allocation suits for any type of topology and network setup. But so far, the question of how $\allocatePeer$ and $\allocateFlow$ are defined was left open. We recap that both rely on derivatives of $\generation$ and $\flow$  with respect to the nodal demand $\demand$ fulfilling \cref{eq:allocate_peer_constraint,eq:allocate_transaction_constraint}. \\
% 
In fact, it turns out that ones the peer-to-peer allocations $\allocatePeer$ are defined, the power flow allocations $\allocateFlow$ are determined by the linear power flow equations.
In order to show this, let $\ptdf$ denote Power Transfer Distribition Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.  Further let $\incidenceGenerators$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise. The linear power flow equals the 
\begin{align}
 \flow  = \sum_m \ptdf[m] \left( \nodalgeneration[m] - \demand[m] \right)  
\end{align}
where the nodal generation $\nodalgeneration[m] = \sum_s \incidenceGenerators[m] \, \generation $ combines the production of all generators attached to bus $m$. Taking the derivative with respect to the demand, leads to 
\begin{align}
 \allocateFlow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatePeer[m \rightarrow n]  - \delta_{n,m} \demand \right) 
 \label{eq:allocate_peer_to_allocate_flow}
\end{align}
where $\allocatePeer[m \rightarrow n] = \sum_s \incidenceGenerators \, \allocatePeer$ combines all power allocations of the source $m$. \Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocatePeer$ directly determines the flow allocation on line $\ell$.

This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocatePeer$. In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$. Unfortunately, the solution is non-unique and requires further assumptions. Established flow allocation schemes approach this problem from different directions. Principally two options exist \textit{what} is allocated 
% 
\begin{enumerate}
\item gross power injections \label{gross}
\item net power injections \label{net}
\end{enumerate}
% 
Further it is important \textit{what assumptions} define the allocation, \ie what method is used to define the pairs of sources and sinks. The three suitable approaches we present here are
% 
\begin{enumerate}[label=\alph*., ref=\alph*]
\item Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} which assumes
that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010-1} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power decifit or excess of the regions. \label{fbmc}
\end{enumerate}
% 
We show the mathematical formulation for all combinations \ref{ebe}\ref{gross} - \ref{fbmc}\ref{net} in \cref{sec:gross_ebe,sec:net_ebe,sec:gross_ap,sec:net_ap}.
% 
Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $n$ with $\nodalgeneration \le \nodaldemand$ is not assigned to other buses, only to $n$. 
Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocatePeer$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$. Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \generation \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.

\subsection{Numerical Example}
\label{sec:numerical_example}

In the following we showcase a two bus system with one optimized time step and its resulting allocated payments, illustrated in \cref{fig:example_network}. \\
% 
% 
\begin{figure*}[t]
\centering
\includegraphics[width=\linewidth]{example_network.png}
\caption{Illustrative example of a 2-bus network with one optimized time step. Fixed prices and constraining values are given in the left box for each bus and the transmission line. Optimized values are given in the right boxes. Bus~1 has a cheaper operational price $o$, capital prices are the same for both. As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2. The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments. \vspace{-10pt}}
\label{fig:example_network}
\end{figure*}
% 
% 
The two buses are connected via one transmission line, each has one generator. Whereas the generator at bus~1 has an operational price of 50 \euro/\megawatthour, the generator at bus~2 has a higher operational price of 200~\euro/\megawatthour. For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacityGenerationUpper$~=~100~MW. The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit. With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW. The 40~MW excess power not consumed at bus~1, flow to bus~2 where the generator is built with only 50~MW.  
% 
\begin{figure}[h!]
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1}
    \end{subfigure}
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2}
    \end{subfigure}
    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2. The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
    The sum of both net flows equals the resulting flow of $f=40$~MW. \vspace{-20pt}}
    \label{fig:example_allocation}
\end{figure}
% 
% 
% 
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{example_payoff.png}
\caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}. The payments are derived on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}. Consumers at bus $n$ have to pay each generator proportional to their consumption. As we only consider one time step the proportionality applies for OPEX $\mathcal{O}_n$ and CAPEX $\mathcal{C}_n$. As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further tranmission expansion, it is rewarded proportional to the relief.}
\label{fig:example_payoff}
\end{figure}
% 
% 
\subsubsection*{Allocating Gross Power using EBE (\ref{gross}\ref{ebe}) }

\Cref{fig:example_allocation} shows the allocated transactions on basis of equvalently traded gross power (type \ref{gross}\ref{ebe}) for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1,1}=40$~MW at bus~1 are self-sustained. Consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator at bus~1. The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{2 \rightarrow 1, 1} = -20$. As this flow is in contrary direction to the total flow, it is relieving the transmission system. This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{2 \rightarrow 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.

The lower graph \cref{fig:example_allocation_bus2} illustrates the impact of consumption at bus~2. As $d_2$ is higher than $d_1$ the retrievements from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators. But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion. Hence the payoff to the transmission system is positive and much higher than for bus~1.

The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$. These values match their overall spendings, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$. The sum of all columns yields the total payment of consumers at bus $n$. For example the sum of payments of consumers at bus~1 is 36k~\euro. This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$. \\

The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only. In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity contraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.

\subsubsection*{Allocating Net Power using EBE (\ref{net}\ref{ebe}) }

In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments. The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}. As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated. Neither has bus~1 to pay fee to the transmission system as it only exports power. So all its consumers pay to its local generators. Bus~2 in constrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
\caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when equivalently allocation net power injection (scheme \ref{ebe}\ref{net}).}
\label{fig:example_payoff_net_ebe}
\end{figure}

Again the cummulative payments per bus meet the nodal spendings $\lmp \, \demand$. The cummulative revenues per generator and transmission line meet the all CAPEX and OPEX.

As the other schemes \ref{ap}\ref{gross} - \ref{fbmc}\ref{net} apart from the EBE scheme don't distinguishably modify allocations within the presented example, we move on to a more realistic setup where systemic and local effects show up.

\section{Application Case}

\clearpage
\appendix

\section{Appendix}
\subsection{LMP from Optimization}
The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\

\subsection{Proof Generation Capacity Payback}

As stated in \cite{brown_decreasing_2020} the capical cost of a generator $s$ is equal to the shadowprice $\muuppergeneration$ times the upper availibility $\generationpotential$ of all time steps 
\begin{align}
 \capexGeneration  = \sum_t \muuppergeneration \, \generationpotential 
 \label{eq:capital_price_generation_sum}
\end{align}
\Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint} which yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ imply the complementary slackness,
\begin{align}
\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
\label{eq:complementary_slackness_upper_generation} \\
\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
\label{eq:complementary_slackness_lower_generation} 
\end{align}
Multiplying both sides of \cref{eq:complementary_slackness_upper_generation} with $\capacityGeneration$ and using \cref{eq:capital_price_generation_sum} leads to 
\begin{align}
 \capexGeneration \, \capacityGeneration  = \sum_t \muuppergeneration \, \generation 
 \label{eq:capital_price_generation_sum}
\end{align}

\subsection{Proof Flow Capacity Payback}

The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. The complementary slackness 
\begin{align}
\muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
\label{eq:complementary_slackness_upper_flow} \\
\mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
\label{eq:complementary_slackness_lower_flow} 
\end{align}
set the respective KKT for flows staying below the thermal limit to zero.

\subsection{Full Lagrangian}
\label{sec:full_lagrangian}
\begin{align}
\notag
& \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
\notag
& \;\;\; \sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow  \\
\notag
&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
\notag
&+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
\notag
&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
\label{eq:full_lagrangian}
\end{align}
% 
where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables. The global maximum of the Lagrangian requires stationarity with respect to all variables. The stationarity of the generation capacity variable leads to 
\begin{align}
\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
\label{eq:capexGeneration_duality}
\end{align}

the stationarity of the transmission capacity to
\begin{align}
\pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
\capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
\label{eq:capexFlow_duality}
\end{align}

and the stationarity of the generation to 
\begin{align}
\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
\opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
\end{align}

\subsection{CO$_2$ constraint}

Imposing an additional CO$_2$ constraint limiting the total emission to K,  
\begin{align}
\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
\label[constraint]{eq:co2_constraint}
\end{align}
with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
% 
% \begin{align}
% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
% \label{eq:non_profit_generator_emission}
% \end{align}
% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
\begin{align}
\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
\end{align}

\subsection{Power Transfer Distribition Factors and Flow Allocation}
\label{sec:ptdf_and_flow_allocation}

In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$. Thus for a given production $\generation$ and demand $\demand$, they directly link to the the resulting flow on each line, 
\begin{align}
\flow  =   \sum_n \ptdf  \left( \nodalgeneration- \nodaldemand \right)    
\label{eq:flow_from_ptdf}
\end{align}
where $\nodalgeneration = \sum_s \generation$ and $\nodaldemand = \sum_a \demand$ combine all generators $s$ and all comsumers $a$ attached to $n$.
The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system. It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses. The choice of slack modifies the PTDF according to 
\begin{align}
\ptdf\left( \slack[m]\right)  = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
\label{eq:ptdf_slacked}
\end{align}
where $\ptdfEqual$ denote the PTDF with equally distributed slack.
When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack. Summing over all ingoing and outgoing flow changes resulting from a positive injection at $n$ yields again the slack 
\begin{align}
\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
\label{eq:slack}
\end{align}
Note that $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
% As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
Established flow allocation schemes [cite] haved used this degree of freedom in order to allocate power flows and exchanges to market participants. Under the assumption that consumers account for all power flows in the grid, the slack is set to $\slack^*$ such that 
\begin{align}
\flow  = - \sum_n \ptdf\left( \slackk[m]\right) \, \nodaldemand  
\label{eq:flow_from_demand}
\end{align}
With such a choice the flow can be reproduced from the demand-side of the system only. Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$. In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add up to $\flow$. These subflows, in turn, can be further broken down to contributions for each bus $m$ in the slack, such that we get the subflow of individual $m \rightarrow n$ relations, that is 
\begin{align}
\allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
\label{eq:allocate_transaction}
\end{align}
% 
It indicates the flow on line $\ell$ coming from generators at $m$ and supplying the demand $\nodaldemand$. When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand}; 
% \begin{align}
% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
% \end{align}
when summing over all sources and sinks, it yields again the power flow, thus
\begin{align}
\flow = \sum_{m,n} \allocateTransaction
\label{eq:transaction_sum}
\end{align}
% 
As mentioned before, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$. As proofen in \cref{sec:proof_allocate_peer}, for each P2P relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
\begin{align}
\allocatePeer &= \slackk[m] \, \nodaldemand 
\label{eq:allocate_peer}
\end{align}
% 
Finally, when summing over all sinks the P2P trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
\begin{align}
\sum_n \allocatePeer = \nodalgeneration[m]
\label{eq:sum_n_allocate_peer}
\end{align}
and summing over all sources yields the nodal demand 
\begin{align}
\sum_m \allocatePeer = \nodaldemand
\label{eq:sum_m_allocate_peer}
\end{align}
which follows from the fact that $\sum_n \slackk = 1$.
Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively. \\

The solution to $\slackk$ follows from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$. That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.

\subsection{\texorpdfstring{Proof \Cref{eq:allocate_peer}}{First Proof}}
\label{sec:proof_allocate_peer}

\Cref{eq:allocate_peer} follows from summing $\allocateTransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
\begin{align*}
\allocatePeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocateTransaction \\
&= \slackk \, \nodaldemand - \sum_\ell \incidence \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  \\
&= \slackk \, \nodaldemand - \left(  \delta_{n,m} - \dfrac{1}{N} - \delta_{n,n} + \dfrac{1}{N} \right) \slackk[m] \, \nodaldemand  \\
&= \slackk \, \nodaldemand - \left(  \delta_{n,m} - 1 \right) \slackk[m] \, \nodaldemand  \\
&= \slackk[m] \, \nodaldemand 
\end{align*}
where we used \cref{eq:slack} and the fact that the equally distributed slack amounts to $1/N$ for all N nodes in the network.

\subsection{\texorpdfstring{Proof of \Cref{eq:sum_n_allocate_peer}}{Second proof}}
\label{sec:proof_sum_n_allocate_peer}
The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatePeer$
\begin{align*}
\sum_m \incidence[m] \, \flow &= - \sum_{m,n} \incidence[m] \, \ptdf \, \nodaldemand \\
\nodalgeneration[m] - \nodaldemand[m] &= - \delta_{m,n} \, \nodaldemand + \slackk[m] \, \nodaldemand    \\
\allocatePeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
\sum_n \allocatePeer &= \nodalgeneration[m]
\end{align*}

\subsection{Allocating Gross Injections with EBE}
\label{sec:gross_ebe}
Allocating gross injections using the Equivalent Bilateral Exchanges Principal is simplistic and straightforward scheme. Accordingly each generator supplies each bus proportional to its power demand. The allocation is given by 
\begin{align}
 \allocatePeer = \dfrac{\generation}{\sum_s \generation} \demand 
\end{align}

\subsection{Allocating Net Injections with EBE}
\label{sec:net_ebe}

Allocating net power injections using the EBE methods leads to the same result as the Marginal Participation (MP) \cite{rudnick_marginal_1995}  algorithm when allocating to consumers only, see \cite{hofmann_flow_2020-1} for further insight. We calculate it by setting 
\begin{align}
\allocatePeer[m \rightarrow n] &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
\label{eq:mp_slack}
\end{align}
where 
\begin{itemize}
%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
\item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
\item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
\item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
\item $\gamma_t = \left( \sum_n \netproduction\right) ^{-1} = \left( \sum_n \netconsumption\right) ^{-1}$ is the inverse of the total injected/extracted power at time $t$.
\end{itemize}

The allocation $\allocatePeer$ from generator $s$ to $n$, is given by multiplying $\allocatePeer[m \rightarrow n]$ with the nodal share $\generation / \nodalgeneration$.

\subsection{Allocating Net Power using AP}
\label{sec:net_ap}

\newcommand{\incidenceM}{K}
\newcommand{\flowM}{f}
\newcommand{\injectionM}{p}
\newcommand{\slackM}{k}
\newcommand{\DirectedIncidence}{\mathcal{K}}
\newcommand{\InverseAPInjection}{\mathcal{J}}
\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}

Allocating net injections using the AP method is derived from \cite{achayuthakan_electricity_2010-1}. In a lossless network the downstream and upstream formulations result in the same P2P allocation which is why we restrict ourselves to the downstream formulation only. In a first step we define a time-dependent auxiliary matrix $\InverseAPInjection_t$ which is the inverse of the $N\times N$ with directed power flow $m \rightarrow n$ at entry $(m, n)$ for $m \ne n$ and the total flow passing node $m$ at entry $\left( m, m\right)$ at time step $t$. Mathematically this translates to

\begin{align}
\InverseAPInjection_t = \left( \diag{\injectionM^+} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( \flow \right)  \incidence$. Then the distributed slack for time step $t$ is given by
\begin{align}
\allocatePeer[m \rightarrow n] = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
\end{align}

\subsection{Allocating Gross Power using AP}
\label{sec:gross_ap}

We use the same allocation as in \cref{sec:net_ap} but replace the net nodal production $\netproduction$ by the gross nodal production $\nodalgeneration$ which leads to  
\begin{align}
\InverseAPInjection_t = \left( \diag{g} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
The distributed slack is for time step $t$ is then given by
\begin{align}
\allocatePeer[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
\end{align}

\subsection{Example: Power Flow Allocations of different Types}
\label{sec:example_plots}

\begin{figure}[h!]
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1_net_ebe}
    \end{subfigure}
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_net_ebe}
    \end{subfigure}
    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allcated net power injections (scheme \ref{net}\ref{ebe}). Bus~1 retrieves 60~MW from itself and nothing from bus~2. The latter in turn retrieves 40~MW from bus~1 and self supplies 50~MW. The P2P trades are less in number and more intuitive then with allocating gross flow.}
    \label{fig:example_allocation_net_ebe}
\end{figure}

\clearpage
\printbibliography

\end{document}

### 2020-07-10 paper/playaround.tex
flow based cost allocation: prepare application case

@@
 \usepackage[dvipsnames]{xcolor}
+\usepackage{amssymb}
 \usepackage{subcaption}
@@
 \newcommand{\hpad}{\hspace{15pt}}
-
-% general symbols
-\newcommand{\generation}{g_{s,t}}
-\newcommand{\generationpotential}{\bar{g}_{s,t}}
-\newcommand{\generationshare}{\omega_{s,t}}
-\newcommand{\incidenceGenerator}{K_{n,s}}
-\newcommand{\nodalgeneration}{\incidenceGenerator \, \generation}
-\newcommand{\capacityGeneration}{G_{s}}
-\newcommand{\capacityGenerationUpper}{\bar{G}_{s}}
-\newcommand{\capacityGenerationLower}{\ubar{G}_{s}}
-\newcommand{\capacityFlow}{F_{\ell}}
-\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
-\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
-\newcommand{\capexGeneration}{c_{s}}
-\newcommand{\capexFlow}{c_{\ell}}
-\newcommand{\opexGeneration}{o_{s}}
-\newcommand{\demand}{d_{n,t}}
-\newcommand{\demandshare}{\omega_{a,t}}
-\newcommand{\utility}{U_{n,a,t}}
-\newcommand{\incidence}[1][n]{K_{#1,\ell}}
-\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
-\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
-\newcommand{\slackflow}{k_{\ell}}
-\newcommand{\slack}[1][n]{k_{#1}}
-\newcommand{\slackk}[1][n]{k^*_{#1}}
-\newcommand{\Slack}{k_{m,n}}
-\newcommand{\Slackk}{k^*_{m,n}}
-\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
-\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
-\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
-\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
-\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
-\newcommand{\flow}{f_{\ell,t}}
-\newcommand{\cycle}{C_{\ell,c}}
-\newcommand{\impedance}{x_\ell}
-\newcommand{\cycleprice}{\lambda_{c,t}}
-\newcommand{\injection}{p_{n,t}}
-\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
-\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
-\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}
-
-% matrix notation
-\newcommand{\incidenceM}{K}
-\newcommand{\flowM}{f}
-\newcommand{\injectionM}{p}
-\newcommand{\slackM}{k}
-\newcommand{\DirectedIncidence}{\mathcal{K}}
-\newcommand{\InverseAPInjection}{\mathcal{J}}
-\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}
-
-% totals
-\newcommand{\totalnetconsumption}{p^{-}_{t}}
-\newcommand{\totalnetproduction}{p^{+}_{t}}
-\newcommand{\totalselfconsumption}{p^{\circ}_{t}}
-
-\newcommand{\lagrangian}{\mathcal{L}}
-
-% allocation quantities
-\newcommand{\allocatePeer}[1][s \rightarrow n]{A_{#1,t}}
-\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
-\newcommand{\allocateTransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
-\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
-\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
-\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
-\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}
-
-\newcommand{\emission}[1][n]{e_{#1,s}}
-\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
-\newcommand{\megawatthour}{MWh$_\text{el}$}
-\newcommand{\totalcost}{\mathcal{TC}}
-\newcommand{\totalOpexGeneration}{\mathcal{O}}
-\newcommand{\totalOpexFlow}{\mathcal{O}^F}
-\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
-\newcommand{\totalCapexFlow}{\mathcal{C}^F}
-\newcommand{\totalEmissionCost}{\mathcal{E}}
-\newcommand{\totalRest}{\mathcal{R}}
-\newcommand{\totalDemand}{\mathcal{D}}
-
-%math 
 \newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
@@
 \newcommand{\pdv}[2]{\dfrac{\partial #1}{\partial #2}}
+\newcommand{\st}{\text{s.t.}\hpad}
+%math
+
+% We consider a power system with minimized total cost $\totalcost$ being subject to supplying demand $\demand$ at node $n$ at time $t$. 
+% The change in total cost against variation of the nodal demand
+% \begin{align}
+% \lmp  =  \pdv{\totalcost}{\demand}
+% \end{align}
+% defines the Locational Marginal Price $\lmp$ that consumers at node $n$ and time $t$. 
+% The total cost $\totalcost$ splits into two cost terms
+% \begin{align}
+% \totalcost &= \totalDemand +  \totalRest
+% \label{eq:total_cost_terms}
+% \end{align}
+% where $\totalDemand$ combines cost all terms which are sensitive to changes in the nodal demand 
+% % \begin{align}
+% %  \pdv{\totalDemand}{\demand} \ne 0
+% % \end{align}
+% and thus determine the LMP,
+% \begin{align}
+% \lmp = \pdv{\totalDemand}{\demand} 
+% \end{align}
+% and $\totalRest$ combines cost terms which are independent of $\lmp$, thus 
+% \begin{align}
+%  0 = \pdv{\totalRest}{\demand} \Forall{i}
+% \end{align}
+% Due to strong duality the total cost $\totalcost$ minus the exogeous cost term $\totalRest$ is totally payed back by the consumers
+% \begin{align}
+%  \totalcost - \totalRest = \sum_{n,t} \lmp \demand
+% \end{align}
+%
 
 \section{Optimization}
-
-\def\l{\lambda}
-\def\K{\kappa}
-\def\m{\mu}
-\def\G{\Gamma}
-\def\d{\partial}
-\def\cL{\mathcal{L}}
 
@@
 \begin{equation}
-  \sum_i A_{ji} x_i \leq d_j \hspace{1cm} \m_j \hspace{1cm} j=1,\dots M
+  \sum_i A_{ji} x_i \leq b_j \hspace{1cm} \mu_j \hspace{1cm} j=1,\dots M
 \end{equation}
-    [Linear equality constraints $\sum_i b_i x_i = c$ with $\lambda$ can be replaced by two inequalities $\leq c$, $\geq c$ with $\lambda = \bar{\m} - \ubar{\m}$.]
+%     [Linear equality constraints $\sum_i b_i x_i = c$ with $\lambda$ can be replaced by two inequalities $\leq c$, $\geq c$ with $\lambda = \bar{\mu} - \ubar{\mu}$.]\\
+The assymetric problem (a problem where $x_i \in \mathbb{R}$) translates to the dual problem of type 
+\begin{align}
+&\max_{\mu_j} \hpad \sum_{j} b_j \mu_j \\ 
+&\st  \sum_{j} A_{ij} \mu_j = c_i
+\end{align}
+
+\subsection{Reduction to Binding Constraints}
 
 A subset $B$ of the inequality constraints will be binding at the
-optimum point $x^*$, i.e. for $j\in B$, $\sum_i A_{ji} x^*_i = d_j$. We
+optimum point $x^*$, i.e. for $j\in B$, $\sum_i A_{ji} x^*_i = b_j$. We
 write $A'_{ji}$ for the matrix that only runs over $j\in B$. For
@@
 \begin{equation}
-  x^*_i = \sum_{k\in B} A^{\prime -1}_{ik} d_k
+  x^*_i = \sum_{j\in B} A^{\prime -1}_{ij} b_j
 \end{equation}
 It might seem that we've now lost all information about the objective function in this expression for $x^*$, but remember that it is the objective function which determines which constraints are binding, i.e.
-which vertex of the feasible simplex is the optimum.
-
-\section{Power system}
-
-We consider a power system with minimized total cost $\totalcost$ being subject to supplying demand $\demand$ at node $n$ at time $t$. 
-The change in total cost against variation of the nodal demand
+which vertex of the feasible simplex is the optimum.\\
+% 
+The matrix $A^{\prime}$ allows switching back and forth between primal and dual space
 \begin{align}
-\lmp  =  \pdv{\totalcost}{\demand}
+b_j = \sum_i x^*_i \, A'_{i,j} \hspace{10pt} \leftrightarrow \hspace{10pt} x^*_i = \sum_j {A'}_{i,j}^{-1} \, b_j \\
+c_i = \sum_j A'_{i,j} \, \mu^*_j  \hspace{10pt} \leftrightarrow \hspace{10pt} \mu^*_j = \sum_i c_i \, {A'}_{i,j}^{-1}
 \end{align}
-defines the Locational Marginal Price $\lmp$ that consumers at node $n$ and time $t$. 
-The total cost $\totalcost$ splits into two cost terms
+The solutions of both problems are connected through
 \begin{align}
-\totalcost &= \totalDemand +  \totalRest
-\label{eq:total_cost_terms}
-\end{align}
-where $\totalDemand = \sum_i \totalDemand_i $ combines cost all terms which are sensitive to changes in the nodal demand 
-\begin{align}
- \pdv{\totalDemand_i}{\demand} \ne 0
-\end{align}
-and thus determine the LMP,
-\begin{align}
-\lmp = \pdv{\totalDemand}{\demand} 
-\end{align}
-and $\totalRest = \sum_i \totalRest_i$ combines cost terms which is not, thus 
-\begin{align}
- 0 = \pdv{\totalRest_i}{\demand} \Forall{i}
-\end{align}
-Due to strong duality the total cost $\totalcost$ minus the exogeous cost term $\totalRest$ is totally payed back by the consumers
-\begin{align}
- \totalcost - \totalRest = \sum_{n,t} \lmp \demand
+\sum_i c_i \, x^*_i = \sum_{i,j}  c_i \, {A'}^{-1}_{ij} b_j = \sum_{i,j} x^*_i \, A'_{i,j} \, \mu^*_j = \sum_j \mu^*_j \, b_j 
 \end{align}
 
-\subsubsection*{Allocate OPEX}
+\subsection{Shifting Coefficients}
+Consider a subset $R$ with $| R | \le M- N$ of constraints which we want to exclude from the problem.
+Therefore we shift the cost coefficients such that they reflect the costs of the removed constraints,
+\begin{align}
+ \tilde{c}_i &= c_i - \sum_{j \in R} A_{ij} \mu^*_j \\
+ &= \sum_{j \notin R} A_{ij} \mu^*_j   \label{eq:coeff_shift}
+\end{align}
+Solving the shifted objective function $\tilde{f}(x)$ with reduced number of constraints,
+\begin{align}
+&\min_{x_i} \hpad \sum_{i=1}^N \tilde{c}_i  x_i \\
+&\st  \sum_i A_{ij} x_i \le b_j \Forall{j \notin R}
+\end{align}
+yields the solution $\tilde{x}^*_i$. Will $\tilde{x}^*_i$ still satisfy the removed constraints 
+\begin{align}
+ \sum_i A_{ij} \tilde{x}^*_i \le b_j \Forall{j \in R} 
+\end{align}
+even though they are not explicit constraints to the problem anymore? We write down the reduced dual problem,
+\begin{align}
+& \max_{\mu_j} \hpad \sum_{j \notin R} b_j \mu_j  \\
+& \st \sum_{j \notin R} A_{ij} \mu_j = \tilde{c}_i \label{eq:shifted_constraints}
+\end{align}
+which has the solution $\tilde{\mu}^*_j $. Inserting \cref{eq:coeff_shift} into \cref{eq:shifted_constraints} leads us to 
+\begin{align}
+ \sum_{j \notin R} A_{ij} \tilde{\mu}^*_j = \sum_{j \notin R} A_{ij} \mu^*_j 
+\end{align}
+which shows that the common dual variables of the reduced problem and the original problem are the same,
+\begin{align}
+ \tilde{\mu}^*_j =  \mu^*_j \Forall{j \notin R}
+\end{align}
+and thus $B - R \subseteq \tilde{B}$.
 
-The total OPEX is given by 
+Consequently the binding constraints matrix 
 \begin{align}
- \totalOpexGeneration& = \sum_{s,t} \opexGeneration\,  \generation 
-\label{eq:opex_generation}
- \end{align}
- We want to disaggregate the totel OPEX into contribution of bus n at time 
- \begin{align}
-  \totalOpexGeneration = \sum_{n,t} \allocateOpex
- \end{align}
-which are defined as 
-\begin{align}
-\allocateOpex  = \pdv{\totalOpexGeneration}{\demand} \, \demand & =  \sum_{s,t}\opexGeneration \, \pdv{\generation}{\demand} \, \demand
-\end{align}
-% 
-From feeding this back into \cref{eq:opex_generation} we see that the derivative $\partial \generation / \partial \demand$ has to fulfill
-\begin{align}
-\generation = \sum_n \pdv{\generation}{\demand} \, \demand
-\end{align}
-Where the term at the right hand side denote the power produced by generator $s$ and consumed at node $n$, which we define as
-\begin{align}
- \allocatePeer = \pdv{\generation}{\demand} \, \demand
- \label{eq:allocate_peer}
-\end{align}
-% 
-The direct payment of node $n$ to generator $s$ is then  
-% 
-\begin{align}
- \allocateOpex[n \rightarrow s] &= 
-\opexGeneration \,  \allocatePeer
-\label{eq:allocate_opexGeneration_detailed}\\
+ \tilde{A}^{\prime}_{ij} =  A^{\prime}_{ij} \Forall{i, j \in B - R}
 \end{align}
 
-\subsubsection*{Allocate Generation CAPEX}
-\begin{align}
- \totalCapexGeneration& = \sum_{s} \capexGeneration \capacityGeneration 
-\end{align}
-with 
-\begin{align}
- \capexGeneration = \muuppergeneration \, \generation
-\end{align}
-and \cref{eq:allocate_peer}, we define 
-\begin{align}
-\allocateCapexGeneration = \pdv{\totalCapexGeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatePeer
-\end{align}
-
-\subsubsection*{Allocate Transmission CAPEX}
-
-The CAPEX for the transmission system is given by
-
-\begin{align}
- \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow = \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow 
-\end{align}
-% 
-The individual cost terms are 
-\begin{align}
-\allocateCapexFlow = \pdv{\totalCapexFlow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
-\end{align}
-where we assume $\partial \muupperflow/\partial \demand \cong 0 $, as dual variables have a second order dependeny on variables.
-The marginal change of flow with repect to the nodal demand is again non-unique and requires assumption on where the power is resulting from.
-
-\begin{align}
- \pdv{\flow}{\demand}\, \demand = \pdv{\flow}{\generation} \pdv{\generation}{\demand} \, \demand  = \allocateTransaction
-\end{align}
-fulfilling 
-\begin{align}
-\sum_{s,n} \allocateTransaction = \flow 
-\end{align}
+% The shifted objective function is then
+% \begin{align}
+%  \min_{x_i} f^{\prime}(x) =  \min_{x_i}  \sum_{i=1}^N \left( c_i + c^{ex}_i \right)  x_i
+% \end{align}
+% where 
+% \begin{align}
+%  c_i^{ex} = \sum_{j \notin D} A^{\prime}_{ij} \mu_j 
+% \end{align}
+% yields a solution where all variable are bound by the subset D of constraints, 
+% \begin{align}
+%   \sum_i A^{\prime \prime}_{ji} x_i = b_j \hspace{1cm} \m_j \hspace{1cm} j \in D
+% \end{align}
+% with 
+% \begin{align}
+%  A^{\prime \prime}_{ji} = A^{\prime}_{ji} \Forall{i \in \{i\, |\, c^{ex}_{i} = 0\}}
+% \end{align}
 

### 2020-07-10 code/config.py

@@ -21,6 +21,12 @@ to_static_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n}$',
                       nodal_demand_cost = '$\lambda_{n} \, d_{n}$')
 
 
+to_total_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}$',
+                      branch_investment_cost = '$\mathcal{C}^{F}$',
+                      one_port_operational_cost = '$\mathcal{O}$',
+                      co2_cost = '$\mathcal{E}$',
+                      nodal_demand_cost = '$\mathcal{TC}$')
+
 color = pd.Series({'one_port_operational_cost': 'darkkhaki',
                    'co2_cost': 'tomato',
                    'one_port_investment_cost': 'palevioletred',

### 2020-07-10 code/constraint_matrix.py

@@ -13,7 +13,8 @@ from helpers import get_linear_system, noisy_lopf
 import networks
 
 n = networks.n_ac_dc()
-n.global_constraints['constant'] += 200
+em = n.generators.carrier.map(n.carriers.co2_emissions) / \
+        n.generators.efficiency
 noisy_lopf(n)
 s = get_linear_system(n)
 A_, A_inv, c, x, d, m, r, C = (s[k].round(5) for k in
@@ -27,20 +28,23 @@ bus = 'London'
 gen = n.generators_t.p.T.copy()
 TC = n.objective
 
-print(C)
+print(n.generators_t.p @ em)
+print(m)
 
 
 mu_co2 = n.global_constraints.mu.co2_limit
 n.global_constraints = n.global_constraints.drop('co2_limit')
 
-em = n.generators.carrier.map(n.carriers.co2_emissions)
-n.generators['marginal_cost'] += em * mu_co2 / n.generators.efficiency
+n.generators['marginal_cost'] += em * mu_co2
 
 n.lopf(pyomo=False, keep_shadowprices=True, keep_references=True,
         solver_name='gurobi')
 s = get_linear_system(n)
 A_, A_inv, c, x, d, m, r, C = (s[k].round(5) for k in
                             ['A_', 'A_inv', 'c', 'x', 'd', 'm', 'r', 'C'])
+
+
+# print(n.generators_t.p @ em)
 
 
 # n.loads_t.p_set.loc[sn,bus] += 1

### 2020-07-10 code/examples.py

@@ -19,6 +19,10 @@ from config import to_static_symbol, color
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
 
+net = False
+altype = 'net' if net else 'gross'
+method = 'ptpf'
+
 
 n = pypsa.Network()
 n0 = '1'
@@ -37,7 +41,7 @@ n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
        marginal_cost=[o0, o1], capital_cost=[c0, c1], p_nom_max=100)
 n.madd('Line', ['1'], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
        capital_cost=c_l)
-n.lopf(pyomo=False, keep_shadowprices=True)
+n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
 
 sn = 'now'
 
@@ -49,7 +53,7 @@ g0, g1 = n.generators_t.p.loc['now']
 f = n.lines_t.p0.loc['now'].item()
 l0, l1 = n.buses_t.marginal_price.loc['now']
 
-ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
+ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=net)
 ds = ntl.convert.peer_to_peer(ds, n)
 dc = ntl.cost.nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)
@@ -102,7 +106,7 @@ ax.legend(*handles_labels_for(color[ca_df.columns].rename(to_static_symbol)),
           ncol=3, loc='lower right', frameon=True, fontsize='large')
 ntl.plot.annotate_bus_names(n, ax, shift=0, bbox='fancy')
 fig.tight_layout()
-fig.savefig(f'../figures/example_network.png', bbox_inches='tight')
+fig.savefig(f'../figures/example_network_{altype}_{method}.png', bbox_inches='tight')
 
 
 # %% payoff matrix
@@ -126,7 +130,7 @@ for i, v in enumerate(payoff):
         ax.set_ylabel('')
     ax.set_title(to_static_symbol[v])
 fig.tight_layout(w_pad=0)
-fig.savefig(f'../figures/example_payoff.png', bbox_inches='tight')
+fig.savefig(f'../figures/example_payoff_{altype}_{method}.png', bbox_inches='tight')
 
 # %% sub flows
 
[…]

### 2020-07-10 code/tinylp.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jul  8 12:37:18 2020
+
+@author: fabian
+"""
+
+import gurobi
+from numpy import array
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+
+model = gurobi.read('tinylp.lp')
+model.optimize()
+
+O = model.getObjective()
+A = model.getA().T.todense() # first dimension {vars} second {cons}
+x = array([v.x for v in model.getVars()])
+d = array([c.RHS  for c in model.getConstrs()])
+c = np.zeros_like(x)
+for t in range(O.size()): c[O.getVar(t).index] = O.getCoeff(t)
+m = array([c.Pi for c in model.getConstrs()])
+
+
+model2 = gurobi.read('tinylp2.lp')
+model2.optimize()
+
+O2 = model2.getObjective()
+A2 = model2.getA().T.todense() # first dimension {vars} second {cons}
+x2 = array([v.x for v in model2.getVars()])
+d2 = array([c.RHS  for c in model2.getConstrs()])
+c2 = np.zeros_like(x)
+for t in range(O2.size()): c2[O2.getVar(t).index] = O2.getCoeff(t)
+m2 = array([c.Pi for c in model2.getConstrs()])
+
+print(x, x2)
+
+
+#%% plot
+N = 10
+fig, (ax1, ax2) = plt.subplots(1,2, sharey=True, figsize=(8,5))
+# ax1.axis('off')
+ax1.set_aspect('equal', adjustable='box')
+ax1.set_xlim(0,N-1)
+# ax2.axis('off')
+ax2.set_xlim(0,N-1)
+ax2.set_aspect('equal', adjustable='box')
+
+f = r'$2 x_0+4 x_1$'
+con1 = r'$ x_1 + x_2 = 7 $'
+con2 = r'$x_1 + 0.5 x_2 \le 4$'
+mu2 = r'$\perp \mu = -4 $'
+fmod = r'$5 x_0 + 5  x_1$'
+
+
+carray = pd.DataFrame([[c @ np.array([x1,x2]) for x1 in range(N)]
+                       for x2 in range(N)])
+xx = np.linspace(0,N,200)
[…]

### 2020-07-12 code/allocate_network.py
introduce snakemake workflow

@@ -1,0 +1,32 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul 10 20:56:10 2020
+
+@author: fabian
+"""
+
+import netallocation as ntl
+import pypsa
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_german_network', clusters=50,
+                                   method='ptpf', power='net')
+
+n = pypsa.Network(snakemake.input.network)
+# n.set_snapshots(n.snapshots[:10])
+method = snakemake.wildcards.method
+aggregated = snakemake.wildcards.power == 'net'
+
+ds = ntl.allocate_flow(n, method=method, aggregated=aggregated,)  # q=0
+ca = ntl.allocate_cost(n, method=ds)
+payments = ca.sum([d for d in ca.dims if d not in ["payer", "snapshot"]])
+
+ca.reset_index('receiver_transmission_cost').to_netcdf(snakemake.output.costs)
+payments.to_netcdf(snakemake.output.payments)
+
+
+# pr = nodal_production_revenue(n).rename(bus="payer")
+# dc = nodal_demand_cost(n).rename(bus="payer")

### 2020-07-12 code/dist_costs.py

@@ -39,7 +39,7 @@ if relax_co2:
     tag = '_relaxed_co2'
 
 sns = n.snapshots
-n.lopf(pyomo=False, keep_shadowprices=True)
+n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
 n.carriers.loc['wind', 'color'] = 'steelblue'
 n.buses.loc[n.buses.index.str.contains('DC'), 'y'] += 1
 n.buses.loc[n.buses.index.str.contains('DC'), 'x'] -= .5
@@ -67,7 +67,7 @@ fig.savefig(f'figures/network{tag}.png')
 ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
 dc = nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)
-ca = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
+# ca = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
 
 # %% Allocation plot
 '''

### 2020-07-12 code/plot_network.py

@@ -1,0 +1,46 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul 10 10:39:11 2020
+
+@author: fabian
+"""
+
+import pypsa
+import numpy as np
+import matplotlib.pyplot as plt
+import netallocation as ntl
+import cartopy.crs as ccrs
+import geopandas as gpd
+
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_german_network', clusters=50)
+
+
+n = pypsa.Network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.regions)
+
+fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                       figsize=(6, 5))
+bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_opt.sum()
+plot = n.plot(
+    bus_sizes=bus_sizes / 2e5,
+    line_widths=n.lines.s_nom_opt / 3e3,
+    link_widths=n.links.p_nom_opt / 3e3,
+    ax=ax,
+)
+ax.legend(
+    *ntl.plot.handles_labels_for(n.carriers.color),
+    loc="center left",
+    bbox_to_anchor=(1, 0.5),
+    ncol=1,
+)
+fig.canvas.draw()
+fig.tight_layout()
+fig.savefig(snakemake.output[0])
+
+
+

### 2020-07-12 code/plot_price_map.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul 10 23:09:45 2020
+
+@author: fabian
+"""
+
+import netallocation as ntl
+import pypsa
+import xarray as xr
+import geopandas as gpd
+import cartopy.crs as ccrs
+from config import to_explanation
+import matplotlib.pyplot as plt
+import os
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_price_maps', clusters=50,
+                                   method='ptpf', power='net')
+
+n = pypsa.Network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.regions)
+payments = xr.open_dataset(snakemake.input.payments)
+
+
+demand = ntl.power_demand(n).rename(bus="payer")
+bus_sizes = (payments).sum("snapshot").to_dataframe().stack().sort_index()
+prices = (payments/demand).mean('snapshot').to_dataframe()
+priceregions = regions.set_index('name').join(prices)
+
+if not os.path.isdir(snakemake.output.folder):
+    os.mkdir(snakemake.output.folder)
+
+
+for col, name in to_explanation.items():
+
+    fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                           figsize=(5, 4))
+    ax.outline_patch.set_visible(False)
+    priceregions.plot(column=col, legend=True, ax=ax,
+                      transform=ccrs.PlateCarree(),
+                      legend_kwds={'label': f'Average LMP for {name} [€/MWh]'})
+    fig.tight_layout()
+    fig.savefig(snakemake.output.folder + f'/{col}_average.png')

### 2020-07-12 code/plot_total_costs.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul 10 21:56:19 2020
+
+@author: fabian
+"""
+
+import xarray as xr
+import matplotlib.pyplot as plt
+from config import color, to_total_symbol
+
+plt.rc('font', family='serif')
+plt.rc("text", usetex=False)
+
+ca = xr.open_dataset(snakemake.input.costs)
+
+fig, ax = plt.subplots()
+ca.sum().to_array().to_series().reindex_like(color).dropna()\
+    .rename(to_total_symbol).rename('')\
+    .plot.pie(colors=color, explode=[.01]*len(ca), autopct='%1.0f%%')
+
+fig.savefig(snakemake.output[0])

### 2020-07-12 code/sanitize_network.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul 10 22:16:58 2020
+
+@author: fabian
+"""
+
+import pypsa
+import numpy as np
+
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('sanitize_network', clusters=50)
+
+
+
+
+n = pypsa.Network(snakemake.input.network)
+
+n.mremove('Line', n.lines.query('num_parallel == 0').index)
+n.add("GlobalConstraint", "CO2Limit_artificial",
+      carrier_attribute="co2_emissions", sense="<=",
+      constant=np.inf, mu=snakemake.config['costs']['emission_prices']['co2'])
+
+n.export_to_netcdf(snakemake.output.network)
+

### 2020-07-12 code/solve_network.py

@@ -9,28 +9,49 @@ Created on Fri Jul 10 10:39:11 2020
 import pypsa
 import numpy as np
 
-name = 'elec_s_50_ec_lvopt_Co2L-3H'
-n = pypsa.Network(f"../{name}.nc")
-mu_co2 = 100
+n = pypsa.Network(snakamake.input)
+
+fig, ax = plt.subplots(subplot_kw={"projection": ccrs.PlateCarree()},
+                       figsize=(6, 5))
+bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_opt.sum()
+plot = n.plot(
+    bus_sizes=bus_sizes / 1e5,
+    line_widths=n.lines.s_nom_opt / 1e3,
+    link_widths=n.links.p_nom_opt / 1e3,
+    ax=ax,
+)
+ax.legend(
+    *ntl.plot.handles_labels_for(n.carriers.color),
+    loc="center left",
+    bbox_to_anchor=(1, 0.5),
+    ncol=1,
+)
+fig.canvas.draw()
+fig.tight_layout()
+fig.savefig(snakemake.output)
 
 
-n.calculate_dependent_values()
-n.generators.p_nom_max.update(n.generators.query('not p_nom_extendable').p_nom)
-n.generators["p_nom_extendable"] = True
-n.lines["s_nom_min"] = 0
-n.mremove("StorageUnit", n.storage_units.index)
-n.mremove("Line", n.lines.query("x_pu == inf").index)
-n.remove("Bus", "DE0 0")
 
-em = n.generators.carrier.map(n.carriers.co2_emissions) / \
-        n.generators.efficiency
-n.global_constraints = n.global_constraints.drop('CO2Limit')
-n.generators['marginal_cost'] += em * mu_co2
+# mu_co2 = 100
 
 
-n.lopf(pyomo=False, keep_shadowprices=True, solver_name="gurobi",
-       solver_options={'crossover': 0, 'method': 2})
+# n.calculate_dependent_values()
+# n.generators.p_nom_max.update(n.generators.query('not p_nom_extendable').p_nom)
+# n.generators["p_nom_extendable"] = True
+# n.lines["s_nom_min"] = 0
+# n.mremove("StorageUnit", n.storage_units.index)
+# n.mremove("Line", n.lines.query("x_pu == inf").index)
+# n.remove("Bus", "DE0 0")
 
-# reinsert mu
-n.global_constraints.at['CO2Limit', 'mu'] = mu_co2
[…]

### 2020-07-15 paper/flow_based.tex
introduce snakemake workflow
include storages!

@@
 \addbibresource{main.bib}
-\graphicspath{{../figures/}}
+\graphicspath{{../figures/}{../figures/example/}}
 
@@
 \newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
+\newcommand{\muuppergenerationnom}{\bar{\mu}^\text{nom}_{s}}
+\newcommand{\mulowergenerationnom}{\ubar{\mu}^\text{nom}_{s}}
 \newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
 \newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
+\newcommand{\muupperflownom}{\bar{\mu}^\text{nom}_{\ell}}
+\newcommand{\mulowerflownom}{\ubar{\mu}^\text{nom}_{\ell}}
 \newcommand{\lmp}[1][n]{\lambda_{#1,t}}
@@
 
-% \subsection{LMP independent Cost  Terms}
-% 
-% Example: \\
-% Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of 
-% \begin{align}
-% \capacityGeneration \ge \capacityGenerationLower \hpad \forall{n,s \in S} \\
-% \capacityGeneration \le \capacityGenerationUpper \hpad \forall{n,s \in S}
-% \end{align}
-% or doing likewise with a subset $L$ of transmission lines,
-% \begin{align}
-% \capacityFlow \ge \capacityFlowLower \hpad \forall{\ell \in \textit{L}} \\
-% \capacityFlow \le \capacityFlowUpper \hpad \forall{\ell \in \textit{L}}
-% \end{align}
-% does not have any effect on the above presented allocation in \cref{eq:lmp_allocation}.
-The additional cost directly translate to the Quality of Supply $\muuppergeneration$ for production and $\muupperflow - \mulowerflow$ for transmission respectively.
-This also counts for limiting the overall production capacity $\sum_{n,s}\capacityGeneration$ or transmission capacity $\sum_\ell \capacityFlow$.
-However it will distort the no profit rule for generators \cref{eq:no_profit_capex_generation} and transmission \cref{eq:no_profit_capex_flow} respectively.
-As shown in \cite{brown_decreasing_2020}
-
 \section{Assumptions on Power Allocations}
@@
 \end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
-In a power market this is the \euro/\megawatthour-price which a consumer has to pay.
-Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\
+Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\
+
+\subsection{Full Lagrangian}
+\label{sec:full_lagrangian}
+\begin{align}
+\notag
+& \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
+\notag
+& \;\;\; \sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow  \\
+\notag
+&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidenceGenerators\, \generation +  \demand  \right)  \\ 
+\notag
+&+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
+\notag
+&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
+&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
+\label{eq:full_lagrangian}
+\end{align}
+% 
+where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
+The global maximum of the Lagrangian requires stationarity with respect to all variables.
+The stationarity of the generation capacity variable leads to
+\begin{align}
+\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
+\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+\label{eq:capex_generation_duality}
+\end{align}
+
+the stationarity of the transmission capacity to
+\begin{align}
+\pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
+\capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
+\label{eq:capexFlow_duality}
+\end{align}
+
+and the stationarity of the generation to 
+\begin{align}
+\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
+\opexGeneration =  \incidenceGenerators \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
+\end{align}
+
+\begin{align}
+0 &= \pdv{\lagrangian}{\flow}  \\ 
+0 &= \sum_n \incidence \lmp  + \cycleprice \cycle \impedance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
+\end{align}
 
@@
 \end{align}
-Multiplying both sides of \cref{eq:complementary_slackness_upper_generation} with $\capacityGeneration$ and using \cref{eq:capital_price_generation_sum} leads to 
+Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacityGeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to 
 \begin{align}
@@
 \end{align}
+The zero-profit rule for generators is obtained by multiplying \cref{eq:opex_duality} with $\generation$ and using \cref{eq:complementary_slackness_lower_generation,eq:capital_price_generation_sum} which results in
+\begin{align}
+  \capexGeneration \, \capacityGeneration + \sum_t \opexGeneration \generation = \sum_t \lmp \incidenceGenerators \generation
+\end{align}
+It states that over the whole time span, all OPEX and CAPEX for generator $s$ (left hand side) are payed back by its revenue (right hand side).
 
@@
 \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding.
-The complementary slackness
+For flows below the thermal limit, the complementary slackness
 \begin{align}
@@
 \end{align}
-set the respective KKT for flows staying below the thermal limit to zero.
-
-\subsection{Full Lagrangian}
-\label{sec:full_lagrangian}
+sets the respective KKT to zero. 
+When multipying \cref{eq:capexFlow_duality} with $\capacityFlow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
 \begin{align}
-\notag
-& \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
-\notag
-& \;\;\; \sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow  \\
-\notag
-&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
-\notag
-&+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
-\notag
-&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
-&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
-\label{eq:full_lagrangian}
+ \capexFlow \, \capacityFlow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow
+ \label{eq:capital_price_flow_sum}
 \end{align}
-% 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-The global maximum of the Lagrangian requires stationarity with respect to all variables.
-The stationarity of the generation capacity variable leads to
+Againg we can use this to formulate the zero-profit rule for transmission lines. We mulitply \cref{eq:opex_flow_duality} with $\flow$, which finally leads us to 
 \begin{align}
-\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
-\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
-\label{eq:capexGeneration_duality}
+\capexFlow \, \capacityFlow = \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
 \end{align}
-
-the stationarity of the transmission capacity to
-\begin{align}
-\pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
-\capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
-\label{eq:capexFlow_duality}
-\end{align}
-
-and the stationarity of the generation to 
-\begin{align}
-\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
-\opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
-\end{align}
+It states that the congestion revenue of a line (first term right hand side) substracted by the cost for cycle constraint exactly matches its CAPEX.
 
@@
 \label{sec:example_plots}
-
 \begin{figure}[h!]
@@
 
+\subsection{Brownfield Optimization and Capacity Restricitions}
+
+Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of
+\begin{align}
+\capacityGeneration \ge \capacityGenerationLower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \\
+\capacityGeneration \le \capacityGenerationUpper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
+\end{align}
+% or doing likewise with a subset $L$ of transmission lines,
+% \begin{align}
+% \capacityFlow \ge \capacityFlowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
+% \capacityFlow \le \capacityFlowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
+% \end{align}
+alters the objective value as soon as one of those become bounding. 
+The complementary slackness state is 
+\begin{align}
+ \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \\
+ \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \\
+\end{align}
+
+The CAPEX paybacks for generators %and transmission lines
+\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
+\begin{align}
+\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
+\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s}
+\label{eq:capex_generation_duality_bf}
+\end{align}
+for generators. % and 
+% \begin{align}
+% \pdv{\lagrangian}{\capacityFlow}  = 0 \,\, \rightarrow \,\, \\
+% \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
+% \label{eq:capex_flow_duality_bf}
+% \end{align}
+% for transmission lines.
+Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ 
+\begin{align}
+ \capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGenerationLower - \muuppergenerationnom \capacityGenerationUpper \Forall{n,s}
+\end{align}
+Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
+\begin{align}
+  \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
+\end{align}
+
+playaround
+
+\begin{align}
+\capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGeneration - \muuppergenerationnom \capacityGeneration \Forall{n,s}
+\end{align}
+
 \clearpage

### 2020-07-15 code/config.py

@@ -7,6 +7,13 @@ Created on Fri May 15 15:20:25 2020
 """
 
 import pandas as pd
+import matplotlib.pyplot as plt
+
+plt.rc('font', family='serif')
+plt.rcParams['savefig.dpi'] = 200
+
+source_dims = dict(bus='source', carrier='source_carrier')
+sink_dims = dict(bus='sink', carrier='sink_carrier')
 
 to_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n,t}$',
                       branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',

### 2020-07-15 code/dist_cost_large.py

@@ -1,65 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Feb 20 14:13:56 2020
-
-@author: fabian
-"""
-
-import pypsa
-import netallocation as ntl
-from netallocation.cost import nodal_demand_cost, nodal_production_revenue
-import matplotlib.pyplot as plt
-import cartopy.crs as ccrs
-from config import to_symbol, color, to_total_symbol, to_explanation
-import geopandas as gpd
-
-plt.rc("text", usetex=True)
-plt.rc('font', family='serif')
-
-
-n = pypsa.Network('../solved_germany_50.nc')
-n.set_snapshots(n.snapshots[:10])
-tag = "_large"
-
-
-# %% total cost
-plt.rc("text", usetex=False)
-ca.sum().to_array().to_series().reindex_like(color).dropna()\
-    .rename(to_total_symbol).rename('')\
-    .plot.pie(colors=color, explode=[.01]*len(ca), autopct='%1.0f%%')
-plt.rc("text", usetex=True)
-
-
-
-
-#%% LMP
-demand = ntl.power_demand(n).rename(bus="payer")
-bus_sizes = (payments).sum("snapshot").to_dataframe().stack().sort_index()
-prices = (payments/demand).mean('snapshot').to_dataframe()
-regions = gpd.read_file('../regions_onshore_elec_s_50.geojson')
-priceregions = regions.set_index('name').join(prices)
-
-
-for col, name in to_explanation.items():
-
-    fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
-                           figsize=(5, 4))
-    ax.outline_patch.set_visible(False)
-    priceregions.plot(column=col, legend=True, ax=ax,
-                      transform=ccrs.PlateCarree(),
-                      legend_kwds={'label': f'Average LMP for {name} [€/MWh]'})
-    fig.tight_layout()
-    fig.savefig(f'../figures/{col}_average.png')
-
-# %%
-
-fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
-                       figsize=(5, 4))
-n.plot(ax=ax, bus_sizes=bus_sizes / bus_sizes.sum() * 3, bus_colors=color)
-handles, labels = ntl.plot.handles_labels_for(color[bus_sizes.index.unique(1)])
[…]

### 2020-07-15 code/dist_costs.py

@@ -18,19 +18,16 @@ from  netallocation.cost import (nodal_co2_cost, nodal_demand_cost,
                                  nodal_production_revenue, congestion_revenue)
 import matplotlib.pyplot as plt
 import cartopy.crs as ccrs
+from config import to_symbol
+
 plt.rc('text', usetex=True)
 
-
-to_symbol_dict = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n,t}$',
-                      branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',
-                      one_port_operational_cost = '$\mathcal{O}_{n,t}$',
-                      co2_cost = '$\mathcal{E}_{n,t}$',
-                      nodal_demand_cost = '$\lambda_{n,t} \, d_{n,t}$')
 
 
 n = ntl.test.get_network_ac_dc()
 n.carriers = n.carriers.drop('battery')
-n.generators['p_nom_max'] = 3500
+n.generators['p_nom_min'] = 500
+
 
 tag = ''
 relax_co2 = False
@@ -68,6 +65,8 @@ ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=False)
 dc = nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)
 ca = ca.sum([d for d in ca.dims if d not in ['payer', 'snapshot']])
+# ca = ntl.breakdown.expand_by_sink_type(ca, n, dim='payer')
+        # .sel(sink_carrier='Load', drop=True).fillna(0)
 
 # %% Allocation plot
 '''
@@ -102,7 +101,7 @@ for sn, ax in zip(n.snapshots, axes.ravel()):
     ax.grid(axis='y', zorder=1, linestyle='dashed')
 
 handles, labels = ax.get_legend_handles_labels()
-labels = [to_symbol_dict[l] for l in labels]
+labels = [to_symbol[l] for l in labels]
 fig.legend(handles[:-1], labels[:-1], ncol=4, frameon=False, loc='lower center',
            bbox_to_anchor=(0.3, 1), fontsize='large', title='Flow Based (left bars)')
 fig.legend(handles[-1:], labels[-1:], ncol=1, frameon=False, loc='lower center',
@@ -112,54 +111,54 @@ fig.tight_layout()
 
 # %% Nodal payments
 
-fig, ax = plt.subplots(subplot_kw={'projection': ccrs.PlateCarree()}, figsize=(5,4))
[…]

### 2020-07-15 code/plot_bars.py

@@ -9,21 +9,22 @@ Created on Mon Jul 13 10:12:28 2020
 import xarray as xr
 import matplotlib.pyplot as plt
 import pypsa
-from config import color, to_symbol
+from config import color, to_symbol, sink_dims
 import netallocation as ntl
 
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_bars', nname='de50',
+        snakemake = mock_snakemake('plot_bars', nname='de10gf',
                                    method='ptpf', power='net')
 
 
 n = pypsa.Network(snakemake.input.network)
 payments = xr.open_dataset(snakemake.input.payments)
-demand = ntl.power_demand(n).rename(bus="payer")
-dc = demand * ntl.cost.locational_market_price(n).rename(bus="payer")
+demand = ntl.power_demand(n, per_carrier=True).rename(sink_dims)
+dc = demand * ntl.cost.locational_market_price(n).rename(bus='sink')
+dc = dc.sel(sink_carrier='Load', drop=True)
 
 fig, axes = plt.subplots(3, 1, figsize=(12, 7), sharex=True)
 

### 2020-07-15 code/plot_cost_map.py

@@ -1,0 +1,58 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Jul 14 10:11:03 2020
+
+@author: fabian
+"""
+
+import netallocation as ntl
+import pypsa
+import xarray as xr
+import geopandas as gpd
+import cartopy.crs as ccrs
+from config import to_explanation
+import matplotlib.pyplot as plt
+import os
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_cost_maps', nname='de10gf',
+                                   method='ptpf', power='net')
+
+n = pypsa.Network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.regions)
+
+payments = xr.open_dataset(snakemake.input.payments)
+costs = (payments).sum('snapshot').to_dataframe()
+
+costs = costs.assign(lmp = (n.buses_t.marginal_price * n.loads_t.p).sum())
+regions = regions.set_index('name').join(costs)
+
+if not os.path.isdir(snakemake.output.folder):
+    os.mkdir(snakemake.output.folder)
+
+
+for col, name in to_explanation.items():
+
+    fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                           figsize=(5.5, 4))
+    ax.outline_patch.set_visible(False)
+    regions.plot(column=col, legend=True, ax=ax,
+                      transform=ccrs.PlateCarree(),
+                      legend_kwds={'label': f'Total Payments for {name} [€]'})
+    fig.canvas.draw()
+    fig.tight_layout()
+    fig.savefig(snakemake.output.folder + f'/{col}_total.png')
+
+
+fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                        figsize=(5.5, 4))
+ax.outline_patch.set_visible(False)
+regions.plot(column='lmp', legend=True, ax=ax,
+                  transform=ccrs.PlateCarree(),
+                  legend_kwds={'label': f'Total Payments [€]'})
+fig.canvas.draw()
+fig.tight_layout()
+fig.savefig(snakemake.output.folder + f'/electricity_total.png')

### 2020-07-15 code/plot_expenditure_maps.py

@@ -1,0 +1,84 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Jul 14 10:36:46 2020
+
+@author: fabian
+"""
+
+import netallocation as ntl
+import pypsa
+import xarray as xr
+import geopandas as gpd
+import cartopy.crs as ccrs
+from config import to_explanation
+import matplotlib.pyplot as plt
+import os
+
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_expenditure_maps', nname='de10gf',
+                                   method='ptpf', power='net')
+
+nice_names = snakemake.config['plotting']['nice_names']
+
+n = pypsa.Network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.regions)
+
+expenditures = xr.open_dataset(snakemake.input.costs).sum(['sink', 'snapshot'])\
+                 .sel(sink_carrier='Load', drop=True) \
+                 .rename(source='bus', source_carrier='carrier')\
+                 .set_index(branch=['component', 'branch_i'])
+
+by_bus_carrier = (expenditures.drop('branch_investment_cost')
+                  .drop_dims('branch').to_dataframe().unstack('carrier'))
+by_bus = by_bus_carrier.sum(level=0, axis=1)
+by_branch = expenditures.branch_investment_cost.to_series()
+
+# regions = regions.assign(production = production.mean())
+
+if not os.path.isdir(snakemake.output.folder):
+    os.mkdir(snakemake.output.folder)
+
+for col in by_bus:
+    fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                            figsize=(5, 4))
+    ax.outline_patch.set_visible(False)
+    regions.plot(column=by_bus[col], legend=True, ax=ax,
+                      transform=ccrs.PlateCarree(),
+                      legend_kwds={'label':
+                                   f'Total {to_explanation[col]} [€]'})
+    fig.canvas.draw()
+    fig.tight_layout()
+    fig.savefig(snakemake.output.folder + f'/{col}_total.png')
+
+
+if not os.path.isdir(snakemake.output.folder + '/by_carrier'):
+    os.mkdir(snakemake.output.folder + '/by_carrier')
+
+for col in by_bus_carrier:
+    carrier = nice_names[col[1]]
[…]

### 2020-07-15 code/plot_network.py

@@ -12,35 +12,128 @@ import matplotlib.pyplot as plt
 import netallocation as ntl
 import cartopy.crs as ccrs
 import geopandas as gpd
-
+import config
+from matplotlib.pyplot import Line2D
+from netallocation.plot_helpers import (make_handler_map_to_scale_circles_as_in,
+                                        make_legend_circles_for)
+from pypsa.plot import projected_area_factor
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_german_network', clusters=50)
+        snakemake = mock_snakemake('plot_network', nname='de10bf')
 
+is_brownfield = 'bf' in snakemake.wildcards.nname
 
 n = pypsa.Network(snakemake.input.network)
 regions = gpd.read_file(snakemake.input.regions)
 
-fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
-                       figsize=(6, 5))
-bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_opt.sum()
-plot = n.plot(
-    bus_sizes=bus_sizes / 2e5,
-    line_widths=n.lines.s_nom_opt / 3e3,
-    link_widths=n.links.p_nom_opt / 3e3,
-    ax=ax,
+bus_scale = 1e-5
+branch_scale = 1e-3
+
+
+if is_brownfield:
+    fig, axes = plt.subplots(1, 2, subplot_kw={"projection": ccrs.EqualEarth()},
+                             figsize=(10, 5))
+    # existent
+    bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_min.sum()
+    plot = n.plot(
+        bus_sizes=bus_sizes * bus_scale,
+        line_widths=n.lines.s_nom_min * branch_scale,
+        link_widths=n.links.p_nom_min * branch_scale,
+        ax=axes[0],
+        geomap='10m',
+        title='Existent Capacities',
+        boundaries=regions.total_bounds[[0,2,1,3]]
+    )
+    regions.plot(ax=axes[0], transform=ccrs.PlateCarree())
+
+    # expanded
+    bus_sizes = n.generators.eval('p_nom_ext = p_nom_opt - p_nom_min')\
+                 .groupby(["bus", "carrier"]).p_nom_ext.sum()\
+                     [lambda ds: ds>0.001]
+    plot = n.plot(
+        bus_sizes=bus_sizes * bus_scale,
[…]

### 2020-07-15 code/plot_price_map.py

@@ -18,19 +18,20 @@ import os
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_price_maps', clusters=50,
+        snakemake = mock_snakemake('plot_price_maps', nname='de10gf',
                                    method='ptpf', power='net')
 
 n = pypsa.Network(snakemake.input.network)
 regions = gpd.read_file(snakemake.input.regions)
+
 payments = xr.open_dataset(snakemake.input.payments)
+demand = ntl.power_demand(n, per_carrier=True).rename(bus="sink")\
+            .sel(carrier='Load', drop=True)
 
-
-demand = ntl.power_demand(n).rename(bus="payer")
-bus_sizes = (payments).sum("snapshot").to_dataframe().stack().sort_index()
 prices = (payments/demand).mean('snapshot').to_dataframe()
 prices = prices.assign(lmp = n.buses_t.marginal_price.mean())
-priceregions = regions.set_index('name').join(prices)
+regions = regions.set_index('name').join(prices)
+regions = regions.assign(demand = n.loads_t.p.mean())
 
 if not os.path.isdir(snakemake.output.folder):
     os.mkdir(snakemake.output.folder)
@@ -41,7 +42,7 @@ for col, name in to_explanation.items():
     fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
                            figsize=(5, 4))
     ax.outline_patch.set_visible(False)
-    priceregions.plot(column=col, legend=True, ax=ax,
+    regions.plot(column=col, legend=True, ax=ax,
                       transform=ccrs.PlateCarree(),
                       legend_kwds={'label': f'Average LMP for \n {name} [€/MWh]'})
     fig.canvas.draw()
@@ -52,9 +53,20 @@ for col, name in to_explanation.items():
 fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
                         figsize=(5, 4))
 ax.outline_patch.set_visible(False)
-priceregions.plot(column='lmp', legend=True, ax=ax,
+regions.plot(column='lmp', legend=True, ax=ax,
                   transform=ccrs.PlateCarree(),
                   legend_kwds={'label': f'Average LMP [€/MWh]'})
 fig.canvas.draw()
 fig.tight_layout()
[…]

### 2020-07-15 code/solve_and_sanitize_network.py

@@ -14,7 +14,8 @@ import shutil
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('solve_and_sanitize_german_network', clusters=50)
+        snakemake = mock_snakemake('solve_and_sanitize_german_network',
+                                   clusters=10, field='bf')
 
 
 solver_opts = snakemake.config['solving']['solver']
@@ -25,10 +26,22 @@ n = pypsa.Network(snakemake.input.network)
 n.generators.p_nom_max.update(n.generators.query('not p_nom_extendable').p_nom)
 n.generators['p_nom_extendable'] = True
 
+
+if snakemake.wildcards.field == 'bf':
+    for c, attr in pypsa.descriptors.nominal_attrs.items():
+        n.df(c)[attr + '_min'] = n.df(c)[attr]
+else:
+    for c, attr in pypsa.descriptors.nominal_attrs.items():
+        n.df(c)[attr + '_min'] = 0
+        n.df(c)[attr] = 0
+
+
 n.lopf(pyomo=False, solver_name=solver_name, solver_options=solver_opts,
         keep_shadowprices=True)
 
 n.mremove('Line', n.lines.query('num_parallel == 0').index)
+
+n.name = snakemake.output.network.split('/')[-1].replace('.nc', '')
 
 n.export_to_netcdf(snakemake.output.network)
 

### 2020-07-15 code/solve_network.py

@@ -1,57 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Fri Jul 10 10:39:11 2020
-
-@author: fabian
-"""
-
-import pypsa
-import numpy as np
-
-n = pypsa.Network(snakamake.input)
-
-fig, ax = plt.subplots(subplot_kw={"projection": ccrs.PlateCarree()},
-                       figsize=(6, 5))
-bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_opt.sum()
-plot = n.plot(
-    bus_sizes=bus_sizes / 1e5,
-    line_widths=n.lines.s_nom_opt / 1e3,
-    link_widths=n.links.p_nom_opt / 1e3,
-    ax=ax,
-)
-ax.legend(
-    *ntl.plot.handles_labels_for(n.carriers.color),
-    loc="center left",
-    bbox_to_anchor=(1, 0.5),
-    ncol=1,
-)
-fig.canvas.draw()
-fig.tight_layout()
-fig.savefig(snakemake.output)
-
-
-
-# mu_co2 = 100
-
-
-# n.calculate_dependent_values()
-# n.generators.p_nom_max.update(n.generators.query('not p_nom_extendable').p_nom)
-# n.generators["p_nom_extendable"] = True
-# n.lines["s_nom_min"] = 0
-# n.mremove("StorageUnit", n.storage_units.index)
-# n.mremove("Line", n.lines.query("x_pu == inf").index)
-# n.remove("Bus", "DE0 0")
-
-# em = n.generators.carrier.map(n.carriers.co2_emissions) / \
-#         n.generators.efficiency
-# n.global_constraints = n.global_constraints.drop('CO2Limit')
-# n.generators['marginal_cost'] += em * mu_co2
-
-
-# n.lopf(pyomo=False, keep_shadowprices=True, solver_name="gurobi",
-#        solver_options={'crossover': 0, 'method': 2})
-
-# # reinsert mu
-# n.global_constraints.at['CO2Limit', 'mu'] = mu_co2
-# n.export_to_netcdf(snakamake.output)

### 2020-07-21 paper/flow_based.tex
adjust brownfield appendix

@@
 \section{Appendix}
+
+\renewcommand\theequation{\thesection.\arabic{equation}}
+\setcounter{equation}{0}
+
+\renewcommand\thefigure{\thesection.\arabic{figure}}    
+\setcounter{figure}{0}
+
 \subsection{LMP from Optimization}
@@
     \end{subfigure}
-    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allcated net power injections (scheme \ref{net}\ref{ebe}).
+    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allocated net power injections (scheme \ref{net}\ref{ebe}).
 Bus~1 retrieves 60~MW from itself and nothing from bus~2.
@@
 \end{align}
+alters the objective value as soon as one of those become bounding. 
+The complementary slackness for both are 
+\begin{align}
+    \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \Forall{n,s \in S} \\
+    \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \Forall{n,s \in S}
+\end{align}
+
+The CAPEX paybacks for generators and transmission lines
+\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
+\begin{align}
+\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
+\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s \in S}
+\label{eq:capex_generation_duality_bf}
+\end{align}
+for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ leads us to 
+\begin{align}
+ \capexGeneration \, \capacityGeneration - \totalRest_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
+ \label{eq:capex_generation_duality_bf2}
+\end{align}
+where we define the cost resulting from the capacity expansion limits as
+\begin{align}
+    \totalRest_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacityGeneration
+    \Forall{s}
+\end{align}
+As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \le 0$ the latter represents a net positive cost term.
+
+Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
+\begin{align}
+    \capexGeneration \capacityGeneration + \sum_t \opexGeneration \generation  - \totalRest_s = \sum_t \lmp \incidenceGenerators \generation \Forall{s \in S}
+\end{align}
+
 % or doing likewise with a subset $L$ of transmission lines,
@@
 % \end{align}
-alters the objective value as soon as one of those become bounding. 
-The complementary slackness state is 
-\begin{align}
- \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \\
- \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \\
-\end{align}
 
-The CAPEX paybacks for generators %and transmission lines
-\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
-\begin{align}
-\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
-\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s}
-\label{eq:capex_generation_duality_bf}
-\end{align}
-for generators. % and 
+% and 
+% \begin{align}
+%     \muupperflownom \left( \capacityFlow - \capacityFlowUpper \right) =0 \\
+%     \mulowerflownom \left( \capacityFlow - \capacityFlowLower \right) =0 \\
+% \end{align}
+
 % \begin{align}
@@
 %     for transmission lines.
-Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ 
-\begin{align}
- \capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGenerationLower - \muuppergenerationnom \capacityGenerationUpper \Forall{n,s}
-\end{align}
-Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
-\begin{align}
-  \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
-\end{align}
 
-playaround
-
-\begin{align}
-\capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGeneration - \muuppergenerationnom \capacityGeneration \Forall{n,s}
-\end{align}
+% Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
+% \begin{align}
+%   \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
+% \end{align}
 

### 2020-07-21 paper/flow_based.tex
paper rename newcommands, finish brownfield section

@@
 \newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
-\newcommand{\capacityGeneration}{G_{s}}
-\newcommand{\capacityGenerationUpper}{\bar{G}_{s}}
-\newcommand{\capacityGenerationLower}{\ubar{G}_{s}}
-\newcommand{\capacityFlow}{F_{\ell}}
-\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
-\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
-\newcommand{\capexGeneration}{c_{s}}
-\newcommand{\capexFlow}{c_{\ell}}
-\newcommand{\opexGeneration}{o_{s}}
+\newcommand{\capacitygeneration}{G_{s}}
+\newcommand{\capacitygenerationupper}{\bar{G}_{s}}
+\newcommand{\capacitygenerationlower}{\ubar{G}_{s}}
+\newcommand{\capacityflow}{F_{\ell}}
+\newcommand{\capacityflowUpper}{\bar{F}_{\ell}}
+\newcommand{\capacityflowLower}{\ubar{F}_{\ell}}
+\newcommand{\capitalpricegeneration}{c_{s}}
+\newcommand{\capitalpriceflow}{c_{\ell}}
+\newcommand{\operationalpricegeneration}{o_{s}}
 \newcommand{\demand}[1][n]{d_{#1,t}}
@@
 \newcommand{\incidence}[1][n]{K_{#1,\ell}}
-\newcommand{\incidenceGenerators}[1][n]{K_{#1,s}}
+\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
 \newcommand{\ptdf}[1][n]{H_{\ell,#1}}
-\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
-\newcommand{\slackflow}{k_{\ell}}
-\newcommand{\slack}[1][n]{k_{#1}}
-\newcommand{\slackk}[1][n]{k^*_{#1}}
-\newcommand{\Slack}{k_{m,n}}
-\newcommand{\Slackk}{k^*_{m,n}}
+
 \newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
@@
 % allocation quantities
-\newcommand{\allocatePeer}[1][s \rightarrow n]{A_{#1,t}}
-\newcommand{\allocateFlow}[1][n]{A_{\ell,#1,t}}
-\newcommand{\allocateTransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
-\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
-\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
-\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
-\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}
+\newcommand{\allocatepeer}[1][s \rightarrow n]{A_{#1,t}}
+\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
+\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
+\newcommand{\allocatecapexflow}[1][n]{\mathcal{C}^{F}_{#1,t}}
+\newcommand{\allocateopex}[1][n]{\mathcal{O}_{#1,t}}
+\newcommand{\allocateemissioncost}[1][n]{\mathcal{E}_{#1,t}}
 
 \newcommand{\emission}{e_{s}}
-\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
+\newcommand{\emissionprice}{\mu_{\text{CO2}}}
 \newcommand{\megawatthour}{MWh$_\text{el}$}
 \newcommand{\totalcost}{\mathcal{TC}}
-\newcommand{\totalOpexGeneration}{\mathcal{O}}
-\newcommand{\totalOpexFlow}{\mathcal{O}^F}
-\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
-\newcommand{\totalCapexFlow}{\mathcal{C}^F}
-\newcommand{\totalEmissionCost}{\mathcal{E}}
-\newcommand{\totalRest}{\mathcal{R}}
+\newcommand{\opexgeneration}{\mathcal{O}}
+\newcommand{\opexflow}{\mathcal{O}^F}
+\newcommand{\capexgeneration}{\mathcal{C}^G}
+\newcommand{\capexflow}{\mathcal{C}^F}
+\newcommand{\emissioncost}{\mathcal{E}}
+\newcommand{\remainingcost}{\mathcal{R}}
 
@@
 \end{abstract}
-
-% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
-We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA.
-This will lead us to a generalized form of P2P allocations and transmission usage allocations based on the choice of slack.
-In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion.
-By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.
 
@@
         $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
-        $\opexGeneration$ & Operational price in \euro/MW \vpad \\
-        $\capexGeneration$ & Capital Price in \euro/MW \vpad \\
-        $\capexFlow$  & Capital Price in \euro/MW for transmission \\ & capacity on line $\ell$  \vpad \\
-        $\capacityGeneration$ & Generation capacity in MW \vpad \\
-        $\capacityFlow$ & Transmission capacity in MW \vpad \\
+        $\operationalpricegeneration$ & Operational price in \euro/MW \vpad \\
+        $\capitalpricegeneration$ & Capital Price in \euro/MW \vpad \\
+        $\capitalpriceflow$  & Capital Price in \euro/MW for transmission \\ & capacity on line $\ell$  \vpad \\
+        $\capacitygeneration$ & Generation capacity in MW \vpad \\
+        $\capacityflow$ & Transmission capacity in MW \vpad \\
         $\incidence$ & Incidence matrix \vpad 
@@
 
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\opexgeneration$ and capital expenditures (CAPEX) $\capexgeneration$ and $\capexflow$ for production  and transmission respectively, thus
 \begin{align}
-\totalcost &= \totalOpexGeneration + \totalEmissionCost +  \totalCapexGeneration +  \totalCapexFlow + ...
+\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration +  \capexflow + ...
 \label{eq:total_cost_terms}
@@
 \begin{align}
-\totalcost - \totalRest &=  \sum_{n,a,t} \lmp \, \demand
+\totalcost - \remainingcost &=  \sum_{n,a,t} \lmp \, \demand
 \label{eq:total_cost_sum}
 \end{align}
-depending on the costs $\totalRest$ which are independent of the nodal demand  
+depending on the costs $\remainingcost$ which are independent of the nodal demand  
 \begin{align}
- \pdv{\totalRest}{\demand} = 0
+ \pdv{\remainingcost}{\demand} = 0
 \end{align}
 % 
-In most cases, where $\totalRest$ plays a minor role, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
+In most cases, where $\remainingcost$ plays a minor role, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
 We schematically show this behaviour in \cref{fig:price_decomposition}.
@@
 
-From inserting \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative}, we are able to disaggregate the cost terms $\totalOpexGeneration$, $\totalCapexGeneration$ and $\totalCapexFlow$ into constributions of demand $\demand$ at bus~$n$ and time~$t$.
+From inserting \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative}, we are able to disaggregate the cost terms $\opexgeneration$, $\capexgeneration$ and $\capexflow$ into constributions of demand $\demand$ at bus~$n$ and time~$t$.
 For the CAPEX allocation we will make use of the shadowprices resulting from the cost mimization.
@@
 
-The total OPEX is given by the operational price $\opexGeneration$ times the generation $\generation$ of all generators $s$, 
+The total OPEX is given by the operational price $\operationalpricegeneration$ times the generation $\generation$ of all generators $s$, 
 \begin{align}
- \totalOpexGeneration& = \sum_{s,t} \opexGeneration\,  \generation 
+ \opexgeneration& = \sum_{s,t} \operationalpricegeneration\,  \generation 
 \label{eq:opex_generation}
  \end{align}
-The allocated OPEX $\allocateOpex$ which naturally sum up to $  \totalOpexGeneration = \sum_{n,t} \allocateOpex$, are given by 
+The allocated OPEX $\allocateopex$ which naturally sum up to $  \opexgeneration = \sum_{n,t} \allocateopex$, are given by 
 \begin{align}
-\allocateOpex  = \pdv{\totalOpexGeneration}{\demand} \, \demand & =  \sum_{s,t}\opexGeneration \, \pdv{\generation}{\demand} \, \demand
+\allocateopex  = \pdv{\opexgeneration}{\demand} \, \demand & =  \sum_{s,t}\operationalpricegeneration \, \pdv{\generation}{\demand} \, \demand
 \end{align}
@@
 \begin{align}
- \allocatePeer = \pdv{\generation}{\demand} \, \demand
+ \allocatepeer = \pdv{\generation}{\demand} \, \demand
  \label{eq:allocate_peer}
@@
 \begin{align}
-\generation = \sum_n \allocatePeer
+\generation = \sum_n \allocatepeer
 \label{eq:allocate_peer_constraint}
@@
 \begin{align}
- \allocateOpex[n \rightarrow s] &= 
-\opexGeneration \,  \allocatePeer
+ \allocateopex[n \rightarrow s] &= 
+\operationalpricegeneration \,  \allocatepeer
 \label{eq:allocate_opexGeneration_detailed}
@@
 \begin{align}
-\sum_{n} \allocateOpex[n \rightarrow s] = \opexGeneration \, \generation
+\sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
 \label{eq:no_profit_opex}
@@
 
-Given a fix price for emissions in CO$_2$ equivalents $\emissionPrice$, the cost term for emission adds up to 
+Given a fix price for emissions in CO$_2$ equivalents $\emissionprice$, the cost term for emission adds up to 
 \begin{align}
- \totalEmissionCost = \emissionPrice \, \sum_s  \emission \, \generation
+ \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
 \end{align}
@@
 \begin{align}
- \allocateEmissionCost = \pdv{\totalEmissionCost}{\demand} \demand = \sum_s \emissionPrice \, \emission \, \allocatePeer  
+ \allocateemissioncost = \pdv{\emissioncost}{\demand} \demand = \sum_s \emissionprice \, \emission \, \allocatepeer  
 \end{align}
@@
 \begin{align}
- \allocateEmissionCost[n \rightarrow s] = \emissionPrice \, \emission \, \allocatePeer
+ \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocatepeer
 \end{align}
@@
 
-The optimization of the generation capacity $\capacityGeneration$ adds the cost term 
+The optimization of the generation capacity $\capacitygeneration$ adds the cost term 
 \begin{align}
- \totalCapexGeneration& = \sum_{s} \capexGeneration \capacityGeneration 
+ \capexgeneration& = \sum_{s} \capitalpricegeneration \capacitygeneration 
 \end{align}
-to $\totalcost$. It combines all capacities times their capital cost $\capexGeneration$. The nominal capacity constraints the generation $\generation$ in the form of 
+to $\totalcost$. It combines all capacities times their capital cost $\capitalpricegeneration$. The nominal capacity constraints the generation $\generation$ in the form of 
 \begin{align}
-\generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
+\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
@@
 \begin{align}
- \capexGeneration \capacityGeneration = \sum_t \muuppergeneration \,  \generation 
+ \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
  \label{eq:no_profit_capex_generation}
@@
 \begin{align}
-\allocateCapexGeneration = \pdv{\totalCapexGeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatePeer
+\allocatecapexgeneration = \pdv{\capexgeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatepeer
 \end{align}
@@
 \begin{align}
- \allocateCapexGeneration[n \rightarrow s] = \muuppergeneration \, \allocatePeer
+ \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocatepeer
  \label{eq:allocate_capexGeneration_detailed}
@@
 If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
-Thus it is these times steps which push investments in $\capacityGeneration$.
+Thus it is these times steps which push investments in $\capacitygeneration$.
 If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_peer_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$.
+When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_peer_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
 
@@
 \begin{align}
- \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow
+ \capexflow & = \sum_\ell \capitalpriceflow \capacityflow
  \label{eq:capex_flow}
 \end{align}
-The transmission capacity variable $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
+The transmission capacity variable $\capacityflow$ limits the flow $\flow$ in both directions, such that 
 \begin{align}
-\flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
+\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
 \label[constraint]{eq:upper_flow_capacity_constraint} \\
-- \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
+- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
 \label[constraint]{eq:lower_flow_capacity_constraint}
@@
 \begin{align}
-\capexFlow \capacityFlow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
+\capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
 \label{eq:no_profit_capex_flow}
@@
 \begin{align}
-\allocateCapexFlow = \pdv{\totalCapexFlow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
+\allocatecapexflow = \pdv{\capexflow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
 \end{align}
@@
 \begin{align}
- \pdv{\flow}{\demand}\, \demand = \allocateFlow
+ \pdv{\flow}{\demand}\, \demand = \allocateflow
 \end{align}
@@
 \begin{align}
-\sum_{n} \allocateFlow = \flow 
+\sum_{n} \allocateflow = \flow 
 \label{eq:allocate_transaction_constraint}
@@
 \begin{align}
- \allocateCapexFlow[n \rightarrow \ell] &=  
-\left( \muupperflow - \mulowerflow\right)  \allocateFlow
+ \allocatecapexflow[n \rightarrow \ell] &=  
+\left( \muupperflow - \mulowerflow\right)  \allocateflow
 \label{eq:allocate_capexFlow_detailed}
@@
 The P2P cost allocation suits for any type of topology and network setup.
-But so far, the question of how $\allocatePeer$ and $\allocateFlow$ are defined was left open.
+But so far, the question of how $\allocatepeer$ and $\allocateflow$ are defined was left open.
 We recap that both rely on derivatives of $\generation$ and $\flow$  with respect to the nodal demand $\demand$ fulfilling \cref{eq:allocate_peer_constraint,eq:allocate_transaction_constraint}.
@@
 % 
-In fact, it turns out that ones the peer-to-peer allocations $\allocatePeer$ are defined, the power flow allocations $\allocateFlow$ are determined by the linear power flow equations.
+In fact, it turns out that ones the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
 In order to show this, let $\ptdf$ denote Power Transfer Distribition Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Further let $\incidenceGenerators$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise.
+Further let $\incidencegenerator$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise.
 The linear power flow equals the
@@
 \end{align}
-where the nodal generation $\nodalgeneration[m] = \sum_s \incidenceGenerators[m] \, \generation $ combines the production of all generators attached to bus $m$.
+where the nodal generation $\nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation $ combines the production of all generators attached to bus $m$.
 Taking the derivative with respect to the demand, leads to
 \begin{align}
- \allocateFlow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatePeer[m \rightarrow n]  - \delta_{n,m} \demand \right) 
+ \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer[m \rightarrow n]  - \delta_{n,m} \demand \right) 
  \label{eq:allocate_peer_to_allocate_flow}
 \end{align}
-where $\allocatePeer[m \rightarrow n] = \sum_s \incidenceGenerators \, \allocatePeer$ combines all power allocations of the source $m$.
-\Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocatePeer$ directly determines the flow allocation on line $\ell$.
+where $\allocatepeer[m \rightarrow n] = \sum_s \incidencegenerator \, \allocatepeer$ combines all power allocations of the source $m$.
+\Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocatepeer$ directly determines the flow allocation on line $\ell$.
 
-This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocatePeer$.
+This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocatepeer$.
 In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$.
@@
 The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal.
-Therefore $\allocatePeer$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
+Therefore $\allocatepeer$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
 Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation.
@@
 Whereas the generator at bus~1 has an operational price of 50 \euro/\megawatthour, the generator at bus~2 has a higher operational price of 200~\euro/\megawatthour.
-For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacityGenerationUpper$~=~100~MW.
+For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacitygenerationupper$~=~100~MW.
 The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit.
@@
 \notag
-& \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
+& \lagrangian\left(\generation, \flow, \capacitygeneration, \capacityflow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
 \notag
-& \;\;\; \sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow  \\
+& \;\;\; \sum_{n,s} \capitalpricegeneration \capacitygeneration + \sum_{n, s, t} \operationalpricegeneration \generation + \sum_{\ell} \capitalpriceflow \, \capacityflow  \\
 \notag
-&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidenceGenerators\, \generation +  \demand  \right)  \\ 
+&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidencegenerator\, \generation +  \demand  \right)  \\ 
 \notag
@@
 \notag
-&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
-&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
+&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacitygeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
+&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityflow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityflow \right)     
 \label{eq:full_lagrangian}
@@
 \begin{align}
-\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
-\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, 
+\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
 \label{eq:capex_generation_duality}
@@
 \begin{align}
-\pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
-\capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
+\pdv{\lagrangian}{\capacityflow} = 0 \,\, \rightarrow \,\, 
+\capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
 \label{eq:capexFlow_duality}
@@
 \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
-\opexGeneration =  \incidenceGenerators \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
+\operationalpricegeneration =  \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
 \end{align}
@@
 \begin{align}
- \capexGeneration  = \sum_t \muuppergeneration \, \generationpotential 
+ \capitalpricegeneration  = \sum_t \muuppergeneration \, \generationpotential 
  \label{eq:capital_price_generation_sum}
@@
 \begin{align}
-\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
+\muuppergeneration \left( \generation - \generationpotential \, \capacitygeneration \right)  &= 0  \Forall{n,s,t} 
 \label{eq:complementary_slackness_upper_generation} \\
@@
 \end{align}
-Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacityGeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to 
+Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacitygeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to 
 \begin{align}
- \capexGeneration \, \capacityGeneration  = \sum_t \muuppergeneration \, \generation 
+ \capitalpricegeneration \, \capacitygeneration  = \sum_t \muuppergeneration \, \generation 
  \label{eq:capital_price_generation_sum}
@@
 \begin{align}
-  \capexGeneration \, \capacityGeneration + \sum_t \opexGeneration \generation = \sum_t \lmp \incidenceGenerators \generation
+  \capitalpricegeneration \, \capacitygeneration + \sum_t \operationalpricegeneration \generation = \sum_t \lmp \incidencegenerator \generation
 \end{align}
@@
 \begin{align}
-\muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
+\muupperflow \left( \flow - \capacityflow \right)  &= 0 \Forall{\ell,t}
 \label{eq:complementary_slackness_upper_flow} \\
-\mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
+\mulowerflow \left( \flow - \capacityflow \right) &=  0 \Forall{\ell,t}
 \label{eq:complementary_slackness_lower_flow} 
@@
 sets the respective KKT to zero. 
-When multipying \cref{eq:capexFlow_duality} with $\capacityFlow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
+When multipying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
 \begin{align}
- \capexFlow \, \capacityFlow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow
+ \capitalpriceflow \, \capacityflow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow
  \label{eq:capital_price_flow_sum}
@@
 \begin{align}
-\capexFlow \, \capacityFlow = \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
+\capitalpriceflow \, \capacityflow = \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
 \end{align}
@@
 \begin{align}
-\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
+\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionprice} 
 \label[constraint]{eq:co2_constraint}
 \end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
+with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionprice$ in \euro/tonne-CO$_2$. 
 % The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
@@
 % \begin{align}
-% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
+% \capitalpricegeneration \, \capacitygeneration + \sum_{t} \operationalpricegeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionprice \right)  \, \generation \Forall{n,s} 
 % \label{eq:non_profit_generator_emission}
@@
 \begin{align}
-\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
+\operationalpricegeneration \rightarrow \operationalpricegeneration + \emission \, \emissionprice \label[relation]{eq:shift_opex_by_emission_cost}
 \end{align}
 
-\subsection{Power Transfer Distribition Factors and Flow Allocation}
-\label{sec:ptdf_and_flow_allocation}
+\subsection{Brownfield Optimization and Capacity Restricitions}
 
-In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus for a given production $\generation$ and demand $\demand$, they directly link to the the resulting flow on each line,
+Constraining the capacities $\capacitygeneration$  for a subset $S$ of generators to lower or upper limits in the form of
 \begin{align}
-\flow  =   \sum_n \ptdf  \left( \nodalgeneration- \nodaldemand \right)    
-\label{eq:flow_from_ptdf}
+\capacitygeneration \ge \capacitygenerationlower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \label{eq:capacityexpansionminimum}\\
+\capacitygeneration \le \capacitygenerationupper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
+\label{eq:capacityexpansionmaximum}
 \end{align}
-where $\nodalgeneration = \sum_s \generation$ and $\nodaldemand = \sum_a \demand$ combine all generators $s$ and all comsumers $a$ attached to $n$.
-The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system.
-It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses.
-The choice of slack modifies the PTDF according to
+alters the objective value as soon as one of those become bounding. 
+The complementary slackness for both are 
 \begin{align}
-\ptdf\left( \slack[m]\right)  = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
-\label{eq:ptdf_slacked}
+    \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{n,s \in S} \\
+    \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{n,s \in S}
 \end{align}
-where $\ptdfEqual$ denote the PTDF with equally distributed slack.
-When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack.
-Summing over all ingoing and outgoing flow changes resulting from a positive injection at $n$ yields again the slack
+
+The CAPEX paybacks for generators and transmission lines
+\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
 \begin{align}
-\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
-\label{eq:slack}
+\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
+\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s \in S}
+\label{eq:capex_generation_duality_bf}
 \end{align}
-Note that $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
-% As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
-Established flow allocation schemes [cite] haved used this degree of freedom in order to allocate power flows and exchanges to market participants.
-Under the assumption that consumers account for all power flows in the grid, the slack is set to $\slack^*$ such that
+for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacitygeneration$ leads us to 
 \begin{align}
-\flow  = - \sum_n \ptdf\left( \slackk[m]\right) \, \nodaldemand  
-\label{eq:flow_from_demand}
+ \capitalpricegeneration \, \capacitygeneration - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
+ \label{eq:capex_generation_duality_bf2}
 \end{align}
-With such a choice the flow can be reproduced from the demand-side of the system only.
-Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$.
-In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add up to $\flow$.
-These subflows, in turn, can be further broken down to contributions for each bus $m$ in the slack, such that we get the subflow of individual $m \rightarrow n$ relations, that is
+where we define the cost resulting from the capacity expansion limits as
 \begin{align}
-\allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
-\label{eq:allocate_transaction}
+    \remainingcost_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacitygeneration
+    \Forall{s}
 \end{align}
-% 
-It indicates the flow on line $\ell$ coming from generators at $m$ and supplying the demand $\nodaldemand$.
-When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand};
+As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \le 0$ the latter represents a net positive cost term.
+
+Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
+\begin{align}
+    \capitalpricegeneration \capacitygeneration + \sum_t \operationalpricegeneration \generation  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
+\end{align}
+We see that the revenue of generator $s$ will not fully pay back its full CAPEX and OPEX.
+The exogenous constraint shifts the zero-profit equations such that some of the expenditures for $s \in S$ cannot directly be allocated to consumers.
+The collected payments of all consumers will not cover the costs induced by \cref{eq:capacityexpansionminimum,eq:capacityexpansionmaximum},
+
+\begin{align}
+    \capitalpricegeneration \capacitygeneration - \remainingcost_s = \sum_{n,t} \allocatecapexgeneration[n \rightarrow s] \Forall{s \in S}
+\end{align}
+
+Doing likewise with a subset $L$ of transmission lines,
+\begin{align}
+\capacityflow \ge \capacityflowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
+\capacityflow \le \capacityflowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
+\end{align}
+results in a shifted zero-profit rule for transmission line in the form 
+\begin{align}
+    \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow[n \rightarrow \ell] \Forall{\ell \in L}
+\end{align}
+where $\remainingcost_\ell$ is given by 
+\begin{align}
+    \remainingcost_\ell = \left(\mulowerflownom - \muupperflownom \right) \, \capacityflow
+    \Forall{\ell \in L}
+\end{align}
+
+% and 
 % \begin{align}
-% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
+%     \muupperflownom \left( \capacityflow - \capacityflowUpper \right) =0 \\
+%     \mulowerflownom \left( \capacityflow - \capacityflowLower \right) =0 \\
 % \end{align}
-when summing over all sources and sinks, it yields again the power flow, thus
-\begin{align}
-\flow = \sum_{m,n} \allocateTransaction
-\label{eq:transaction_sum}
-\end{align}
-% 
-As mentioned before, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
-As proofen in \cref{sec:proof_allocate_peer}, for each P2P relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
-\begin{align}
-\allocatePeer &= \slackk[m] \, \nodaldemand 
-\label{eq:allocate_peer}
-\end{align}
-% 
-Finally, when summing over all sinks the P2P trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
-\begin{align}
-\sum_n \allocatePeer = \nodalgeneration[m]
-\label{eq:sum_n_allocate_peer}
-\end{align}
-and summing over all sources yields the nodal demand 
-\begin{align}
-\sum_m \allocatePeer = \nodaldemand
-\label{eq:sum_m_allocate_peer}
-\end{align}
-which follows from the fact that $\sum_n \slackk = 1$.
-Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
-\\
 
-The solution to $\slackk$ follows from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
-That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.
+% \begin{align}
+%     \pdv{\lagrangian}{\capacityflow}  = 0 \,\, \rightarrow \,\, \\
+%     \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
+%     \label{eq:capex_flow_duality_bf}
+%     \end{align}
+%     for transmission lines.
 
-\subsection{\texorpdfstring{Proof \Cref{eq:allocate_peer}}{First Proof}}
-\label{sec:proof_allocate_peer}
-
-\Cref{eq:allocate_peer} follows from summing $\allocateTransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
-\begin{align*}
-\allocatePeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocateTransaction \\
-&= \slackk \, \nodaldemand - \sum_\ell \incidence \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  \\
-&= \slackk \, \nodaldemand - \left(  \delta_{n,m} - \dfrac{1}{N} - \delta_{n,n} + \dfrac{1}{N} \right) \slackk[m] \, \nodaldemand  \\
-&= \slackk \, \nodaldemand - \left(  \delta_{n,m} - 1 \right) \slackk[m] \, \nodaldemand  \\
-&= \slackk[m] \, \nodaldemand 
-\end{align*}
-where we used \cref{eq:slack} and the fact that the equally distributed slack amounts to $1/N$ for all N nodes in the network.
-
-\subsection{\texorpdfstring{Proof of \Cref{eq:sum_n_allocate_peer}}{Second proof}}
-\label{sec:proof_sum_n_allocate_peer}
-The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatePeer$
-\begin{align*}
-\sum_m \incidence[m] \, \flow &= - \sum_{m,n} \incidence[m] \, \ptdf \, \nodaldemand \\
-\nodalgeneration[m] - \nodaldemand[m] &= - \delta_{m,n} \, \nodaldemand + \slackk[m] \, \nodaldemand    \\
-\allocatePeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
-\sum_n \allocatePeer &= \nodalgeneration[m]
-\end{align*}
-
-\subsection{Allocating Gross Injections with EBE}
-\label{sec:gross_ebe}
-Allocating gross injections using the Equivalent Bilateral Exchanges Principal is simplistic and straightforward scheme.
-Accordingly each generator supplies each bus proportional to its power demand.
-The allocation is given by
-\begin{align}
- \allocatePeer = \dfrac{\generation}{\sum_s \generation} \demand 
-\end{align}
+% Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
+% \begin{align}
+%   \capitalpricegeneration \, \capacitygeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacitygenerationlower \right) + \muuppergenerationnom \left( \capacitygenerationupper - \capacitygenerationlower \right) \Forall{n,s}
+% \end{align}
 
@@
 \begin{align}
-\allocatePeer[m \rightarrow n] &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
+\allocatepeer[m \rightarrow n] &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
 \label{eq:mp_slack}
@@
 
-The allocation $\allocatePeer$ from generator $s$ to $n$, is given by multiplying $\allocatePeer[m \rightarrow n]$ with the nodal share $\generation / \nodalgeneration$.
+The allocation $\allocatepeer$ from generator $s$ to $n$, is given by multiplying $\allocatepeer[m \rightarrow n]$ with the nodal share $\generation / \nodalgeneration$.
 
@@
 \begin{align}
-\allocatePeer[m \rightarrow n] = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
+\allocatepeer[m \rightarrow n] = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
 \end{align}
@@
 \begin{align}
-\allocatePeer[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
+\allocatepeer[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
 \end{align}
@@
 
-\subsection{Brownfield Optimization and Capacity Restricitions}
-
-Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of
-\begin{align}
-\capacityGeneration \ge \capacityGenerationLower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \\
-\capacityGeneration \le \capacityGenerationUpper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
-\end{align}
-alters the objective value as soon as one of those become bounding. 
-The complementary slackness for both are 
-\begin{align}
-    \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \Forall{n,s \in S} \\
-    \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \Forall{n,s \in S}
-\end{align}
-
-The CAPEX paybacks for generators and transmission lines
-\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
-\begin{align}
-\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
-\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s \in S}
-\label{eq:capex_generation_duality_bf}
-\end{align}
-for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ leads us to 
-\begin{align}
- \capexGeneration \, \capacityGeneration - \totalRest_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
- \label{eq:capex_generation_duality_bf2}
-\end{align}
-where we define the cost resulting from the capacity expansion limits as
-\begin{align}
-    \totalRest_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacityGeneration
-    \Forall{s}
-\end{align}
-As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \le 0$ the latter represents a net positive cost term.
-
-Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
-\begin{align}
-    \capexGeneration \capacityGeneration + \sum_t \opexGeneration \generation  - \totalRest_s = \sum_t \lmp \incidenceGenerators \generation \Forall{s \in S}
-\end{align}
-
-% or doing likewise with a subset $L$ of transmission lines,
-% \begin{align}
-% \capacityFlow \ge \capacityFlowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
-% \capacityFlow \le \capacityFlowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
-% \end{align}
-
-% and 
-% \begin{align}
-%     \muupperflownom \left( \capacityFlow - \capacityFlowUpper \right) =0 \\
-%     \mulowerflownom \left( \capacityFlow - \capacityFlowLower \right) =0 \\
-% \end{align}
-
-% \begin{align}
-%     \pdv{\lagrangian}{\capacityFlow}  = 0 \,\, \rightarrow \,\, \\
-%     \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
-%     \label{eq:capex_flow_duality_bf}
-%     \end{align}
-%     for transmission lines.
-
-% Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
-% \begin{align}
-%   \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
-% \end{align}
-
 \clearpage

### 2020-07-21 paper/ptdf_detailed.tex
paper rename newcommands, finish brownfield section

--- 2020-05-24 main.tex
@@
 \addbibresource{main.bib}
-\graphicspath{{figures/}}
+\graphicspath{{../figures/}{../figures/example/}}
 
@@
 
+% style operators
 \newcommand{\ie}{\textit{i.e.} }
 \newcommand{\eg}{\textit{e.g.} }
-
 \newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
@@
 \newcommand{\vpad}{\vspace{1mm}}
+\newcommand{\hpad}{\hspace{15pt}}
+\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
+\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
+\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
 
-\newcommand{\generation}[1][n]{g_{#1,s,t}}
-\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
+% general symbols
+\newcommand{\generation}{g_{s,t}}
+\newcommand{\generationpotential}{\bar{g}_{s,t}}
 \newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
 \newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
-\newcommand{\capacityGeneration}{G_{n,s}}
+\newcommand{\capacitygeneration}{G_{s}}
+\newcommand{\capacitygenerationupper}{\bar{G}_{s}}
+\newcommand{\capacitygenerationlower}{\ubar{G}_{s}}
 \newcommand{\capacityFlow}{F_{\ell}}
-\newcommand{\capexGeneration}{c_{n,s}}
-\newcommand{\capexFlow}{c_{\ell}}
-\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
-\newcommand{\demand}[1][n]{d_{#1,a,t}}
+\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
+\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
+\newcommand{\capitalpricegeneration}{c_{s}}
+\newcommand{\capitalpriceflow}{c_{\ell}}
+\newcommand{\operationalpricegeneration}{o_{s}}
+\newcommand{\demand}[1][n]{d_{#1,t}}
 \newcommand{\nodaldemand}[1][n]{d_{#1,t}}
@@
 \newcommand{\incidence}[1][n]{K_{#1,\ell}}
+\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
 \newcommand{\ptdf}[1][n]{H_{\ell,#1}}
@@
 \newcommand{\Slackk}{k^*_{m,n}}
-\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
-\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
+\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
+\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
+\newcommand{\muuppergenerationnom}{\bar{\mu}^\text{nom}_{s}}
+\newcommand{\mulowergenerationnom}{\ubar{\mu}^\text{nom}_{s}}
 \newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
 \newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
+\newcommand{\muupperflownom}{\bar{\mu}^\text{nom}_{\ell}}
+\newcommand{\mulowerflownom}{\ubar{\mu}^\text{nom}_{\ell}}
 \newcommand{\lmp}[1][n]{\lambda_{#1,t}}
@@
 \newcommand{\totalselfconsumption}{p^{\circ}_{t}}
-
 \newcommand{\lagrangian}{\mathcal{L}}
 
-\newcommand{\allocatePeer}[1][m \rightarrow n]{A_{#1,t}}
-\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
-\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
-\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
-\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
-\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
-\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}
+% allocation quantities
+\newcommand{\allocatepeer}[1][s \rightarrow n]{A_{#1,t}}
+\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
+\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
+\newcommand{\allocatecapexflow}[1][n]{\mathcal{C}^{F}_{#1,t}}
+\newcommand{\allocateopex}[1][n]{\mathcal{O}_{#1,t}}
+\newcommand{\allocateemissioncost}[1][n]{\mathcal{E}_{#1,t}}
 
-\newcommand{\emission}[1][n]{e_{#1,s}}
-\newcommand{\emissionPrice}{\mu_{\text{CO2}}}
+\newcommand{\emission}{e_{s}}
+\newcommand{\emissionprice}{\mu_{\text{CO2}}}
 \newcommand{\megawatthour}{MWh$_\text{el}$}
 \newcommand{\totalcost}{\mathcal{TC}}
-\newcommand{\totalOpexGeneration}{\mathcal{O}}
-\newcommand{\totalOpexFlow}{\mathcal{O}^F}
-\newcommand{\totalCapexGeneration}{\mathcal{C}^G}
-\newcommand{\totalCapexFlow}{\mathcal{C}^F}
-\newcommand{\totalEmissionCost}{\mathcal{E}}
-
-\newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
-\newcommand{\impactcapexflow}{\Phi_{\ell,t}}
-
-%math 
-\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
-\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
-\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
+\newcommand{\opexgeneration}{\mathcal{O}}
+\newcommand{\opexflow}{\mathcal{O}^F}
+\newcommand{\capexgeneration}{\mathcal{C}^G}
+\newcommand{\capexflow}{\mathcal{C}^F}
+\newcommand{\emissioncost}{\mathcal{E}}
+\newcommand{\remainingcost}{\mathcal{R}}
 
@@
 
-\title{From Linear Optimization to Transmission Cost Allocation}
-\author{F. Hofmann. T. Brown}
-
-\maketitle
-
-\begin{abstract}
-Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and  
-\end{abstract}
-
-\section{Introduction}
-
-% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system.
-We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA.
-This will lead us to a generalized form of peer-to-peer allocations and transmission usage allocations based on the choice of slack.
-In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion.
-By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.
-
-\section{Locational Market Price}
-
-\subsubsection*{Nomenclature}
-
-\begin{table}[h]
-	\centering
-	\begin{tabular}{ll}
-        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
-        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
-        $\generation$ & Electric generation per bus $n$, carrier $s$, \\ & time step $t$  in MW \vpad \\
-        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
-        $\opexGeneration$ & Operational cost (OPEX) in \euro/MW \vpad \\
-        $\capexGeneration$ & Capital Expenditure (CAPEX) in \euro/MW \vpad \\
-        $\capexFlow$  & CAPEX per transmission line $\ell$ in \euro/MW  \vpad \\
-        $\capacityGeneration$ & Generation capacity in MW \vpad \\
-        $\capacityFlow$ & Transmission capacity in MW \vpad \\
-        $\incidence$ & Incidence matrix \vpad 
-	\end{tabular}
-\end{table}
-
-The total system cost $\totalcost$ has to be payed back by the consumers 
-\begin{align}
- \totalcost =  \sum_{n,a,t} \lmp \, \demand
-\end{align}
-
-\begin{align}
- \totalcost = \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
-\end{align}
-
-\begin{figure}[h]
- \includegraphics[width=\linewidth]{figures/price_decomposition.png}
- \label{fig:price_decomposition}
- \caption{The Locational Market Price}
-\end{figure}
-
-\section{Mathematical Theory}
-
-\subsubsection*{Power Transfer Distribition Factors and Flow Allocation}
+\subsection{Power Transfer Distribition Factors and Flow Allocation}
+\label{sec:ptdf_and_flow_allocation}
 
 In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Thus for a given gross production $\generation$ and gross demand $\demand$, they directly link to the the resulting flow on each line,
+Thus for a given production $\generation$ and demand $\demand$, they directly link to the the resulting flow on each line,
 \begin{align}
@@
 \begin{align}
- \allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
+\allocatetransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
 \label{eq:allocate_transaction}
@@
 % 
-It indicates the flow on line $\ell$ provided generators at $m$ and supplying the demand $\nodaldemand$.
+It indicates the flow on line $\ell$ coming from generators at $m$ and supplying the demand $\nodaldemand$.
 When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand};
 % \begin{align}
-% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
+% \sum_m \allocatetransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
 % \end{align}
@@
 \begin{align}
-\flow = \sum_{m,n} \allocateTransaction
+\flow = \sum_{m,n} \allocatetransaction
 \label{eq:transaction_sum}
@@
 % 
-As mentioned, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
-As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
+As mentioned before, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$.
+As proofen in \cref{sec:proof_allocate_peer}, for each P2P relation $m \rightarrow n$, the ``traded`` power $\allocatepeer$  amounts to
 \begin{align}
- \allocatePeer &= \slackk[m] \, \nodaldemand 
+\allocatepeer &= \slackk[m] \, \nodaldemand 
 \label{eq:allocate_peer}
@@
 % 
-Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
+Finally, when summing over all sinks the P2P trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
 \begin{align}
- \sum_n \allocatePeer = \nodalgeneration[m]
+\sum_n \allocatepeer = \nodalgeneration[m]
 \label{eq:sum_n_allocate_peer}
@@
 \begin{align}
- \sum_m \allocatePeer = \nodaldemand
+\sum_m \allocatepeer = \nodaldemand
 \label{eq:sum_m_allocate_peer}
 \end{align}
-which straightforwardly follows from the fact that $\sum_n \slackk = 1$.
-Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be further broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
+which follows from the fact that $\sum_n \slackk = 1$.
+Both allocation quantities $\allocatepeer$ and $\allocatetransaction$ can be broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively.
 \\
 
-The solution to $\slackk$ follows directly from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
+The solution to $\slackk$ follows from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$.
 That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$.
-However solution space can be extended to an individual slack for each node $\Slackk$, which we discuss in \cref{sec:solution_space_of_the_slack}.
 
-\subsubsection*{Network Optimisation}
-
-% Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
-% 
-% \begin{align}
-%  \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
-% \end{align}
-% ...
-%
-
-We linearly cost-optimize the capacity and dispatch of a simple power system.
-
-\begin{align}
-    \underset{\demand, \generation, \capacityGeneration}{\text{max}}
-    \left(\sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
-\end{align}
-subject to following physical constraints. 
-\\
-
-The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
-\begin{align}
-    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
-    \label[constraint]{eq:nodal_balance_lin}
-\end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
-In a power market this is the \euro/\megawatthour-price which a consumer has to pay.
-Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\
-
-The generation $\generation$ is constraint to its nominal capacity
-\begin{align}
- \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
- \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
- \label[constraint]{eq:lower_generation_capacity_constraint}
- \end{align}
-where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness,
-\begin{align}
-\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
-\label{eq:complementary_slackness_upper_generation} \\
-\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
-\label{eq:complementary_slackness_lower_generation} 
-\end{align}
-are only non-zero if the corresponding constraint is binding. \\
-
-The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
-\begin{align}
- \flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
- \label[constraint]{eq:upper_flow_capacity_constraint} \\
- - \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
- \label[constraint]{eq:lower_flow_capacity_constraint}
-\end{align}
-The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e.
-\cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding.
-The complementary slackness
-\begin{align}
- \muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
- \label{eq:complementary_slackness_upper_flow} \\
- \mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
- \label{eq:complementary_slackness_lower_flow} 
-\end{align}
-set the respective KKT for flows staying below the thermal limit to zero. 
-\\
-
-\subsubsection*{Decomposing the Locational Marginal Prices and Allocating the Electricity Costs}
-\begin{align}
-\notag
- & \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
- \notag
- & - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
- \notag
- &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
- \notag
- &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
- \notag
- &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
- &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
- \label{eq:full_lagrangian}
-\end{align}
-% 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-The global maximum of the Lagrangian requires stationarity with respect to all variables.
-The stationarity of the generation capacity variable leads to
-\begin{align}
- \pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
- \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
- \label{eq:capexGeneration_duality}
-\end{align}
-
-the stationarity of the transmission capacity to
-\begin{align}
- \pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
- \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
- \label{eq:capexFlow_duality}
-\end{align}
-
-and the stationarity of the generation to 
-\begin{align}
- \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
- \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
-\end{align}
-Solving \cref{eq:opex_duality} for the $\lmp$, leads to our first representation for Locational Market Price, which we will refer to as the ``Island Solution``,
-\begin{align}
-\lmp  =  \opexGeneration + \muuppergeneration - \mulowergeneration \Forall{n,s,t}
-\label{eq:lmp1}
-\end{align}
-It connects the LMP directly with the local operational price and prices for the generation capacity constraint.
-However, we can derive a second representation for $\lmp$.
-Starting from the stationarity of the flow
-\begin{align}
- 0 =& \pdv{\lagrangian}{\flow}  \\
- \notag
- 0 =& - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  \\
- & -\sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance
-\end{align}
-and multipying each term with the Power Transfer Distribution Factor $\ptdf$ leaves us with
-
-\begin{align}
-  0 &= - \lmp + \sum_m \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right) \ptdf
-  \label{eq:lmp_lmp1_reduced}
-\end{align}
-According to \cref{eq:slack}, the first term splits into the LMP at $n$ and the LMP weighted with the slack.
-The final term disappears as the $\cycle \, \impedance$ is the kernel of the PTDF $\ptdf$, so $\sum_l \cycle \, \impedance \ptdf = 0$.
-Solving \cref{eq:lmp_lmp1_reduced} for $\lmp$ and replacing $\lmp[m]$ of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to
-
-\begin{align}
-\notag
-\lmp =&  \sum_m \opexGeneration[m]\, \slack[m]  + \sum_m \left( \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \\
-& -\sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
-\label{eq:lmp2}
-\end{align}
-%  
-% NOTE: \cref{eq:lmp2} builds a starting point for breaking down the market value (MV) 
-% of generator $(m,s)$, defined as
-% \begin{align}
-%  MV_s = \dfrac{\sum_t \generation \, \lmp}{\sum_t \generation}
-% \end{align}
-In contrast to the Island Solution in \cref{eq:lmp1}, this representation connects $\lmp$ with all other prices in the system.
-Moreover, it must hold for any choice of $\slack[m]$ and correspondingly decomposes the LMP to operational prices $\opexGeneration[m]$ and prices for capacity bounds for generators and transmission lines.
-When setting the slack to $\slackk[m] = c \, \generation[m]$, the LMP decomposes to prices of generators proportional to their production plus an extra term for the transmission usage.
-This directly links to the flow allocation presented above.
-Multiplied with the nodal demand $\nodaldemand$ the left hand side of \cref{eq:lmp2} turns into the total payment of $n$ and the right hand side into the different payment allocations,
-\begin{align}
- \lmp \, \nodaldemand = \allocateOpex + \allocateCapexGeneration + \allocateCapexFlow \Forall{n,t}
- \label{eq:lmp_allocation}
-\end{align}
-which we define as  
-\begin{align}
- \allocateOpex &= 
- \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
-\label{eq:allocate_opexGeneration}\\
- \allocateCapexGeneration &= 
- \sum_{m,s} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
-\label{eq:allocate_capexGeneration} \\
- \allocateCapexFlow &=  
- \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
-\label{eq:allocate_capexFlow}
-\end{align}
-\Cref{eq:allocate_opexGeneration} denotes the payments for generators' OPEX, \cref{eq:allocate_capexGeneration} for generators' CAPEX and \cref{eq:allocate_capexFlow} for the transmission system's CAPEX respectively.
-All those payment allocation build up on the phycally allocated flows presented in \cref{eq:allocate_peer,eq:allocate_transaction}.
-Thus with the right choice of slack the power flow allocations directly tranlate into a cost allocation.
-Further, the allocated payments can be considered on a more detailed level as
-\begin{align}
- \allocateOpex[n \rightarrow (m,s)] &= 
- \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
-\label{eq:allocate_opexGeneration_detailed}\\
- \allocateCapexGeneration[n \rightarrow (m,s)] &= 
- \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
-\label{eq:allocate_capexGeneration_detailed} \\
- \allocateCapexFlow[n \rightarrow \ell] &=  
-  \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
-\label{eq:allocate_capexFlow_detailed}
-\end{align}
-%
-
-Let's have a look at the allocated OPEX in \cref{eq:allocate_opexGeneration,eq:allocate_opexGeneration_detailed} first.
-Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly compensate their operational costs.
-The OPEX allocation behaves like P2P tradings between producers and consumers with fixed production prices.
-In this way, the generator $(m,s)$ retrieves the exact amount of money from consumers that it spends on the operation.
-In other words, all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s), thus
-\begin{align}
-\sum_{n} \allocateOpex[n \rightarrow (m,s)] = \opexGeneration \, \generation
-\label{eq:no_profit_opex}
-\end{align}
-
-The CAPEX allocation for generators $\allocateCapexGeneration$ defined in \cref{eq:allocate_capexGeneration,eq:allocate_capexGeneration_detailed}, reveils a similar relation.
-According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration > 0$, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
-Thus it is these times steps which push investments in $\capacityGeneration$.
-If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $(m,s)$ each generator retrieves exactly the cost that were spent to build the capacity $\capacityGeneration$,
-
-\begin{align}
- \sum_{n,t} \allocateCapexGeneration[n \rightarrow (m,s)] = \capexGeneration \, \capacityGeneration
-\label{eq:no_profit_capex_generation}
-\end{align}
-where we used \cref{eq:complementary_slackness_upper_generation,eq:capexGeneration_duality}.
-So in total, throughout all time steps each generator $(m,s)$ receives the money it spends for invesments and operation, reflecting the zero-profit rule.
-% The non-profit rule can also be seen when looking at the the total revenue per generator, 
-% \begin{align}
-%  \sum_t \lmp \, \generation &= \sum_t \opexGeneration \generation + \muuppergeneration \generation - \mulowergeneration \, \generation &\Forall{n,s} \\
-%  &= \sum_t \opexGeneration \, \generation + \capexGeneration \, \capacityGeneration &\Forall{n,s}
-% \end{align}
-%  where we multiplied both sides of \cref{eq:lmp1} by the generation and using the complementary slackness.  
-\\
-
-The allocation of CAPEX for the transmission system $\allocateCapexFlow$, defined in \cref{eq:allocate_capexFlow,eq:allocate_capexFlow_detailed}, builds up on the KKT variables $\muupperflow$ and $\mulowerflow$.
-Again the latter translate to the necessity of transmission invesments at $\ell$ at time $t$.
-Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments.
-Again the sum of all CAPEX payments to line $\ell$ equal the total CAPEX spent, thus
-
-\begin{align}
- \sum_{n,t} \allocateCapexFlow[n \rightarrow \ell] = \capexFlow \, \capacityFlow  
-\end{align}
-where we used the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} and the fact that summing over all sources $m$ and sinks $n$ the allocation equals the actual power flow as stated in \cref{eq:transaction_sum}.
-% A similar relation applies for the transmission CAPEX.
-Mulplying both side of \cref{eq:opex_duality} by the nodal injection $\sum_s \generation - \sum_a \demand$ and using the complementary slackness ...
-% shows that the CAPEX per line amounts to the total congestion revenue
-% \begin{align}
-% \capexFlow \, \capacityFlow &= - \sum_{n,t} \lmp \, \incidence \, \flow \Forall{\ell} \label{eq:non_profit_branch}
-% \end{align}
-% The transmission lines carry power from nodes with lower LMP to nodes with higher LMP, the price difference times the transported power is their congestion revenue.
-\note{Sign is different to Tom's paper (there a minus is missing).}
-
-\subsubsection*{Adding CO$_2$ Constraints}
-
-Imposing an additional CO$_2$ constraint limiting the total emission to K,  
-\begin{align}
- \sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionPrice} 
- \label[constraint]{eq:co2_constraint}
-\end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionPrice$ in \euro/tonne-CO$_2$. 
-% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
-% 
-% \begin{align}
-% \capexGeneration \, \capacityGeneration + \sum_{t} \opexGeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionPrice \right)  \, \generation \Forall{n,s} 
-% \label{eq:non_profit_generator_emission}
-% \end{align}
-% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
-As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
-\begin{align}
-\opexGeneration \rightarrow \opexGeneration + \emission \, \emissionPrice \label[relation]{eq:shift_opex_by_emission_cost}
-\end{align}
-This leads to allocated CO$_2$ cost compensation of node $n$ of
- \begin{align}
- \allocateEmissionCost &= \emissionPrice \, \sum_{m,s} \emission[m] \, \generationshare[m] \, \allocatePeer \Forall{n,t} \label{eq:allocate_emissionPrice}
-\end{align}
-which expands the allocation of the electricity cost in \cref{eq:lmp_allocation} to 
-\begin{align}
- \lmp \, \nodaldemand = \allocateCapexFlow + \allocateOpex + \allocateCapexGeneration  + \allocateEmissionCost \Forall{n,t}
- \label{eq:lmp_allocation_with_emission}
-\end{align}
-
-% \newpage
-% \section*{Showcase}
-% 
-% \subsection*{Network with CO$_2$ constraint}
-% We illustrate the flow based cost allocation under use of the fictive network shown in \cref{fig:network}.
-It consists of nine buses and ten time steps.
-The solver optimizes the capacity of two generators, wind and gas, per bus.
-...
-% 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{compare_allocation.png}
-%     \caption{Comparison between the flow based cost allocation and the LMP based cost per consumer.
-The left bars consist of the allocated OPEX $\allocateOpex$, the allocated CO$_2$ cost $\allocateEmissionCost$, the allocated generator CAPEX $\allocateCapexGeneration$ and transmission CAPEX $\allocateCapexFlow$, while the right bars show the of the nodal consumption times the LMP.
-}
-%     \label{fig:cost_allocation}
-% \end{figure}
-% 
-% \begin{figure}[h]
-% \begin{subfigure}{.5\textwidth}
-% \centering
-%  \includegraphics[width=\textwidth]{network.png}
-%  \caption{}
-%  \label{fig:network}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}   
-%     \centering
-%     \includegraphics[width=\textwidth]{nodal_payments.png}
-%     \caption{}
-%     \label{fig:nodal_payments}
-% \end{subfigure}
-% \caption{Network used for showcasing.
-(a) shows the distributing of generation capacities $\capacityGeneration$, the widths of the transmission lines are proportional to their thermal limit $\capacityFlow$.
-(b) shows the total nodal payments according to the cost allocation.}
-% \end{figure}
-% 
-% 
-% 
-% 
-% 
-% \newpage
-% \subsubsection*{Relaxed CO$_2$ Constraint}
-% 
-% 
-% \begin{figure}[h]
-% \begin{subfigure}{.5\textwidth}
-% \centering
-%  \includegraphics[width=\textwidth]{network_relaxed_co2.png}
-%  \caption{}
-%  \label{fig:network_relaxed_co2}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}   
-%     \centering
-%     \includegraphics[width=\textwidth]{nodal_payments_relaxed_co2.png}
-%     \caption{}
-%     \label{fig:nodal_payments_relaxed_co2}
-% \end{subfigure}
-% \caption{Similar to \cref{fig:network} and \cref{fig:nodal_payments} but without CO$_2$ constraint.}
-% \end{figure}
-% 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{compare_allocation_relaxed_co2.png}
-%     \caption{Comparison of the stacked flow based cost allocation with the LMP based cost per consumer for each time step $t$ without CO$_2$ \cref{eq:co2_constraint}.
-Only one time-step $t=6$ determines the allocation of generator CAPEX $\allocateCapexGeneration$, as for all other time-steps \cref{eq:upper_generation_capacity_constraint} is not binding.
-Again note the cost scale difference between time step 6 and all others.}
-%     \label{fig:cost_allocation_relaxed_co2}
-% \end{figure}
-% 
-% \subsection*{How does the cost flow through the network}
-% 
-% \begin{figure}[h]
-%     \begin{subfigure}{.5\textwidth}
-%       \centering
-%       \includegraphics[width=\textwidth]{opex_flow.png}
-%       \label{fig:opex_flow}
-%     \end{subfigure}%
-%     \begin{subfigure}{.5\textwidth}
-%       \centering
-%       \includegraphics[width=\textwidth]{capex_flow.png}
-%       \label{fig:capex_flow}
-%     \end{subfigure}
-%     \caption{}
-%     \label{fig:fig}
-% \end{figure}
-
-\newpage
-\appendix
-
-\section{Appendix}
-
-\subsection{Proof \Cref{eq:allocate_peer}}
+\subsection{\texorpdfstring{Proof \Cref{eq:allocate_peer}}{First Proof}}
 \label{sec:proof_allocate_peer}
 
-\Cref{eq:allocate_peer} follows from summing $\allocateTransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
-\begin{align}
- \allocatePeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocateTransaction \\
+\Cref{eq:allocate_peer} follows from summing $\allocatetransaction$ over all incoming flows to $n$ and taking into account the power that $n$ provides by itsself, $\slackk \, \nodaldemand$, which leads us to
+\begin{align*}
+\allocatepeer &= \slackk \, \nodaldemand - \sum_{\ell} \incidence \, \allocatetransaction \\
 &= \slackk \, \nodaldemand - \sum_\ell \incidence \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand  \\
@@
 &= \slackk[m] \, \nodaldemand 
-\label{eq:proof_allocate_peer}
-\end{align}
+\end{align*}
 where we used \cref{eq:slack} and the fact that the equally distributed slack amounts to $1/N$ for all N nodes in the network.
 
-\subsection{Proof of \Cref{eq:sum_n_allocate_peer}}
+\subsection{\texorpdfstring{Proof of \Cref{eq:sum_n_allocate_peer}}{Second proof}}
 \label{sec:proof_sum_n_allocate_peer}
-The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatePeer$
-\begin{align}
+The relation follows from multiplying \cref{eq:flow_from_demand} with $\sum_m \incidence[m]$, and solving for $\allocatepeer$
+\begin{align*}
 \sum_m \incidence[m] \, \flow &= - \sum_{m,n} \incidence[m] \, \ptdf \, \nodaldemand \\
 \nodalgeneration[m] - \nodaldemand[m] &= - \delta_{m,n} \, \nodaldemand + \slackk[m] \, \nodaldemand    \\
-\allocatePeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
-\sum_n \allocatePeer &= \nodalgeneration[m]
+\allocatepeer &= \nodalgeneration[m] - \nodaldemand[m] + \delta_{m,n} \nodaldemand \\
+\sum_n \allocatepeer &= \nodalgeneration[m]
+\end{align*}
+
+\subsection{Allocating Gross Injections with EBE}
+\label{sec:gross_ebe}
+Allocating gross injections using the Equivalent Bilateral Exchanges Principal is simplistic and straightforward scheme.
+Accordingly each generator supplies each bus proportional to its power demand.
+The allocation is given by
+\begin{align}
+ \allocatepeer = \dfrac{\generation}{\sum_s \generation} \demand 
 \end{align}
 
-\subsection{Solution Space of $\boldsymbol{\slackk[m]}$}
-\label{sec:solution_space_of_the_slack}
-
-All choices of $\slackk[m]$ fulfilling \cref{eq:flow_from_demand} determine the solution space of $\slackk$.
-In the $N \times N$ nodal space this translates to the constraint given by \cref{eq:sum_n_allocate_peer} which denotes
-\begin{align}
- \nodalgeneration[m] = \sum_n \slackk[m] \, \nodaldemand  
- \label{eq:solution_space_slack}
-\end{align}
-Solving for $\slackk[m]$ directly leads to 
-\begin{align}
- \slackk[m] = c \, \generation[m] 
-\end{align}
-where $c$ is the inverse of the total consumption or production $c = 1 / \sum_n \nodaldemand   = 1 / \sum_n \nodalgeneration$. \\
-However \cref{eq:solution_space_slack} has an inherent degree of freedom and can be reformulated as 
-\begin{align}
- \nodalgeneration[m] = \sum_n \Slackk \, \nodaldemand  
- \label{eq:solution_space_slack_extended}
-\end{align}
-where $\Slackk$ denote the individual choice of slack for each bus $n$.
-For example the Marginal Participation cite[] algorithm takes only net injections into account.
-This sets the individual slack to
-\begin{align}
-\Slackk &= \dfrac{ \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]}{\nodaldemand}
-\end{align}
-where 
-\begin{itemize}
-%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
- \item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
- \item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
- \item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
- \item $\gamma_t = \dfrac{1}{\sum_n \netproduction} = \dfrac{1}{\sum_n \netconsumption}$ is the inverse of the total injected/extracted power at time $t$.
-\end{itemize}
-
-% \bibliographystyle{ieeetr}
-\printbibliography
-
 \end{document}

### 2020-07-23 paper/flow_based.tex
clean up paper

@@
 \usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
-\usepackage{amsmath}
+\usepackage{amsmath,bm}
 \usepackage[colorlinks]{hyperref}
@@
 \newcommand{\totalcost}{\mathcal{TC}}
+\newcommand{\cost}{\mathcal{C}}
 \newcommand{\opexgeneration}{\mathcal{O}}
@@
 
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\opexgeneration$ and capital expenditures (CAPEX) $\capexgeneration$ and $\capexflow$ for production  and transmission respectively, thus
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of mutliple cost terms.
+Typically, these include operational expenditures (OPEX) $\opexgeneration$ for generators, expenditures for emissions $\emissioncost$, capital expenditures (CAPEX) for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and possible other terms, \ie
 \begin{align}
@@
 \end{align}
-In a network design with minimal $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of any consumer $a$ at bus $n$.
+In turn, each single cost term on the right side $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
+\begin{align}
+    \cost = \sum_i \cost_i
+    \label{eq:asset_costs}
+\end{align} 
+For example the total OPEX, $\opexgeneration = \sum_s \opexgeneration_s$, is the combined OPEX of all generators $s$.
+
+Minimizing the total cost $\totalcost$  of a network, which translates to maximizing the total welfare, will determine the single expenditure $\cost_i$ dedicated to asset $i$.
+with maximized welfare.
+Then, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$.
 It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
@@
 % 
-
 This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are partially or totally payed back by the consumers 
@@
 % 
-In most cases, where $\remainingcost$ plays a minor role, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures.
-We schematically show this behaviour in \cref{fig:price_decomposition}.
-Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection however leave the question open how the costs are allocated among the components.
-In the following we will show how the above presented cost decomposition is extended to full peer-to-peer (P2P) cost allocation including all network components.
+Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits in certain regions of the network, or minimum share of one technology in the power mix.
+However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
+Feeding \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative} illustrates that the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission (and other possible expenditures), schematically shown in \cref{fig:price_decomposition}.
+Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection.
+However the question of how the LMP translates to the single cost terms $\cost_i$ associated to asset $i$ in the network remains unanswered.
+This work aims at presenting and illustrating a an intuitive scheme how to cost breakdown the cost decomposition to a full peer-to-peer (P2P) cost allocation including all network assets.
 
@@
 
-From inserting \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative}, we are able to disaggregate the cost terms $\opexgeneration$, $\capexgeneration$ and $\capexflow$ into constributions of demand $\demand$ at bus~$n$ and time~$t$.
-For the CAPEX allocation we will make use of the shadowprices resulting from the cost mimization.
+In following we show how to disaggregate the cost terms $\opexgeneration$, $\emissioncost$, $\capexgeneration$ and $\capexflow$ for asset $i$ into contributions of demand $\demand$ at bus~$n$ and time~$t$.
+That is what amount consumers at $\left(n, t\right)$ pay for the operation, investment and other cost-intensive, but welfare contributing, actions at asset $i$.
+We follow the general procedure: For each cost term $\cost$ we break down the contributions of asset $i$, \ie define \cref{eq:asset_costs}.
+According to \cref{eq:total_cost_derivative,eq:total_cost_sum}, we define its derivative with respect to $\demand$ and multiply the result with $\demand$.
+This returns $\cost_{n\rightarrow i,t}$, the contribution of $\demand$ to the cost term $\cost_i$.
+Follows naturally that when summing up all contributions the cost term $\cost_i$ must be covered, \ie
+\begin{align}
+    \cost_i = \sum_{n,t} \cost_{n \rightarrow i, t}
+\end{align}
+and the total money spent at $\left(n,t\right)$ for the cost term $\cost$ amounts 
+\begin{align}
+    \cost_{n,t} = \sum_i \cost_{n \rightarrow i, t}
+\end{align}
+Note that for allocating CAPEX, we will make use of the shadowprices associated to $\cost_i$, given by the dual values at cost optimum $\bm{\mu}$.
 
@@
  \end{align}
-The allocated OPEX $\allocateopex$ which naturally sum up to $  \opexgeneration = \sum_{n,t} \allocateopex$, are given by 
+
+Hence, the contribution of $\demand$ to the OPEX at generator $s$ is straightforward given by
 \begin{align}
-\allocateopex  = \pdv{\opexgeneration}{\demand} \, \demand & =  \sum_{s,t}\operationalpricegeneration \, \pdv{\generation}{\demand} \, \demand
+    \allocateopex[n \rightarrow s] &= 
+   \operationalpricegeneration \,  \pdv{\generation}{\demand} \, \demand
+\label{eq:allocate_opexGeneration_detailed}
 \end{align}
-% 
-The derivative on the right hand side time the nodal demand can be interpreted as power produced by generator $s$ and consumed at node $n$.
-Without further explaining how the derivative is calculated, we define this power allocation as
+
+We define the derivative on the right hand side times the nodal demand as 
 \begin{align}
@@
 % 
-which by feeding back into \cref{eq:opex_generation} has to fulfill
+The quantity $\allocatepeer$ can be considered as a power allocation, that is the power produced by generator $s$ and consumed at node $n$ and time $t$.
+By feeding back \cref{eq:allocate_opexGeneration_detailed} into \cref{eq:opex_generation}, we see that the power allocation has to fulfill
 \begin{align}
@@
 % 
-The direct payment of node $n$ to generator $s$ is then given by  
-% 
-\begin{align}
- \allocateopex[n \rightarrow s] &= 
-\operationalpricegeneration \,  \allocatepeer
-\label{eq:allocate_opexGeneration_detailed}
-\end{align}
-Accordingly, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs.
+The definition of $\allocateopex[n \rightarrow s]$ in \cref{eq:allocate_opexGeneration_detailed} can be interpreted as the following.
+Consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs.
 The cost allocation behaves like P2P tradings between producers and consumers for fixed production prices.
-In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation, thus
+In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$, thus
 \begin{align}
@@
 where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
-Similar to the allocated OPEX, the emission cost for consumers at bus $n$ at time $t$ is allocated through
-\begin{align}
- \allocateemissioncost = \pdv{\emissioncost}{\demand} \demand = \sum_s \emissionprice \, \emission \, \allocatepeer  
-\end{align}
-with its peer-to-peer payments 
+Similar to the allocated OPEX, the emission cost for consumers at bus $n$ at time $t$ for operation at generator $s$ is allocated through
 \begin{align}
@@
 where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-Using the result of \cite{brown_decreasing_2020} we show in that the capital cost  for generator $s$ translates is payed back by the generation times the shadow price for the upper capacity constraint,
+As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation},  the capital cost  for generator $s$ is payed back by the generation times the shadow price for the upper capacity constraint,
 \begin{align}
@@
 With our definition in \cref{eq:allocate_peer} the allocated CAPEX can then be defined as 
-\begin{align}
-\allocatecapexgeneration = \pdv{\capexgeneration}{\demand} \, \demand & = \sum_{s} \muuppergeneration \, \allocatepeer
-\end{align}
-with the peer-to-peer allocation 
 \begin{align}
@@
 \end{align}
-Again, we use the result of \cite{brown_decreasing_2020} which derives that the investment in line $\ell$ is payed back by the shadowprices times the flow 
+We use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that the investment in line $\ell$ is payed back by the shadowprices times the flow 
 \begin{align}
@@
 \begin{align}
-\allocatecapexflow = \pdv{\capexflow}{\demand}\demand & =  \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \pdv{\flow}{\demand} \, \demand
+    \allocatecapexflow[n \rightarrow \ell] &=  
+   \left( \muupperflow - \mulowerflow\right) \,  \pdv{\flow}{\demand}\, \demand
+   \label{eq:allocate_capexFlow_detailed}
 \end{align}
-The marginal change of flow with repect to the nodal demand is again non-unique and requires assumption on where the power is resulting from.
-
+Again, we define the derivative on the right side times the demand as 
 \begin{align}
- \pdv{\flow}{\demand}\, \demand = \allocateflow
+ \allocateflow =  \pdv{\flow}{\demand}\, \demand
 \end{align}
-fulfilling 
+which from inserting into \cref{eq:no_profit_capex_flow} has to fulfill 
 \begin{align}
-\sum_{n} \allocateflow = \flow 
+ \flow = \sum_{n} \allocateflow
 \label{eq:allocate_transaction_constraint}
 \end{align}
-% 
-The peer-to-peer cost allocation for the transmission CAPEX is then defined as 
-\begin{align}
- \allocatecapexflow[n \rightarrow \ell] &=  
-\left( \muupperflow - \mulowerflow\right)  \allocateflow
-\label{eq:allocate_capexFlow_detailed}
-\end{align}
-% 
-Again we can interpret the shadowprices $\muupperflow$ and $\mulowerflow$ as a measure for necessity of transmission investments at $\ell$ at time $t$.
-Consumers which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments.
+$\allocateflow$ can be interpreted as the flow that the demand at $\left(n,t\right)$ induces on line $\ell$.
+The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
+Hence, the definition of $\allocatecapexflow[n \rightarrow \ell]$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
 Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
@@
 
-The P2P cost allocation suits for any type of topology and network setup.
+The presented cost allocation suits for any type of topology and network setup.
 But so far, the question of how $\allocatepeer$ and $\allocateflow$ are defined was left open.
@@
 % 
-In fact, it turns out that ones the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
+In fact, it turns out that ones, the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
 In order to show this, let $\ptdf$ denote Power Transfer Distribition Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
@@
 \subsection{Proof Generation Capacity Payback}
+\label{sec:zero_profit_generation}
 
@@
 \subsection{Proof Flow Capacity Payback}
+\label{sec:zero_profit_flow}
 
@@
 \end{align}
+
+\subsection{Storage Units}
 
@@
 
-% and 
 % \begin{align}
@@
 %     for transmission lines.
-
 % Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
@@
 % \end{align}
+
+\subsection{Allocating Gross Injections with EBE}
+\label{sec:gross_ebe}
+
+The allocation of gross generation to demands $\demand$ is straigthforwardly obained by a proportional distribution of the generation, \ie
+
+\begin{align}
+    \allocatepeer[s \rightarrow n] = \dfrac{\generation}{\sum_s \generation} \demand 
+\end{align}
 

### 2020-08-09 paper/flow_based.tex
further cleanup/fixes

@@
 \usepackage{subcaption}
+\usepackage{booktabs} % for table (toprule, bottomrule)
 \usepackage{enumitem} % for alphabetical enumeration 
@@
 \creflabelformat{constraint}{(#2#1#3)}
+\crefname{subequations}{Eqs.}{Eqs.}
+\creflabelformat{subequations}{(#2#1#3)}
 
@@
 % general symbols
+\newcommand{\state}{s_{i,t}}
+\newcommand{\costfactor}{c_{i,t}}
+
+%generation
 \newcommand{\generation}{g_{s,t}}
@@
 \newcommand{\capacitygenerationlower}{\ubar{G}_{s}}
-\newcommand{\capacityflow}{F_{\ell}}
-\newcommand{\capacityflowUpper}{\bar{F}_{\ell}}
-\newcommand{\capacityflowLower}{\ubar{F}_{\ell}}
+\newcommand{\operationalpricegeneration}{o_{s}}
 \newcommand{\capitalpricegeneration}{c_{s}}
-\newcommand{\capitalpriceflow}{c_{\ell}}
-\newcommand{\operationalpricegeneration}{o_{s}}
-\newcommand{\demand}[1][n]{d_{#1,t}}
-\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
-\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
-\newcommand{\utility}{U_{n,a,t}}
-\newcommand{\incidence}[1][n]{K_{#1,\ell}}
-\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
-\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
-
 \newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
@@
 \newcommand{\mulowergenerationnom}{\ubar{\mu}^\text{nom}_{s}}
+
+% flow
+\newcommand{\flow}{f_{\ell,t}}
+\newcommand{\capacityflow}{F_{\ell}}
+\newcommand{\capacityflowUpper}{\bar{F}_{\ell}}
+\newcommand{\capacityflowLower}{\ubar{F}_{\ell}}
+\newcommand{\operationalpriceflow}{o_\ell}
+\newcommand{\capitalpriceflow}{c_{\ell}}
 \newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
@@
 \newcommand{\mulowerflownom}{\ubar{\mu}^\text{nom}_{\ell}}
+
+% storage
+\newcommand{\storage}{g_{r,t}}
+\newcommand{\storagedispatch}{\storage^\text{dis}}
+\newcommand{\storagecharge}{\storage^\text{sto}}
+\newcommand{\storagesoc}{\storage^\text{ene}}
+
+\newcommand{\efficiency}{\eta_{r}}
+\newcommand{\efficiencydispatch}{\efficiency^\text{dis}}
+\newcommand{\efficiencycharge}{\efficiency^\text{sto}}
+\newcommand{\efficiencysoc}{\efficiency^\text{ene}}
+
+\newcommand{\operationalpricestorage}{o_r}
+\newcommand{\capitalpricestorage}{c_r}
+\newcommand{\capacitystorage}{G_r}
+\newcommand{\capacitystorageupper}{\bar{G}_r}
+\newcommand{\capacitystoragelower}{\ubar{G}_r}
+\newcommand{\mulowerstorage}{\ubar{\mu}_{r,t}}
+\newcommand{\muupperstorage}{\bar{\mu}_{r,t}}
+\newcommand{\mulowerstoragedispatch}{\ubar{\mu}_{r,t}^\text{dis}}
+\newcommand{\muupperstoragedispatch}{\bar{\mu}_{r,t}^\text{dis}}
+\newcommand{\mulowerstoragecharge}{\ubar{\mu}_{r,t}^\text{sto}}
+\newcommand{\muupperstoragecharge}{\bar{\mu}_{r,t}^\text{sto}}
+\newcommand{\mulowerstoragesoc}{\ubar{\mu}_{r,t}^\text{ene}}
+\newcommand{\muupperstoragesoc}{\bar{\mu}_{r,t}^\text{ene}}
+\newcommand{\muupperstoragenom}{\bar{\mu}^\text{nom}_r}
+\newcommand{\mulowerstoragenom}{\ubar{\mu}^\text{nom}_r}
+\newcommand{\mustateofcharge}{\lambda^\text{ene}_{r,t}}
+
+% other
+\newcommand{\lagrangian}{\mathcal{L}}
 \newcommand{\lmp}[1][n]{\lambda_{#1,t}}
-\newcommand{\flow}{f_{\ell,t}}
-\newcommand{\cycle}{C_{\ell,c}}
-\newcommand{\impedance}{x_\ell}
-\newcommand{\cycleprice}{\lambda_{c,t}}
+\newcommand{\demand}[1][n]{d_{#1,t}}
+\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
+\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
 \newcommand{\injection}{p_{n,t}}
@@
 \newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}
-
 \newcommand{\totalnetconsumption}{p^{-}_{t}}
@@
 \newcommand{\totalselfconsumption}{p^{\circ}_{t}}
-\newcommand{\lagrangian}{\mathcal{L}}
+
+% network
+\newcommand{\incidence}[1][n]{K_{#1,\ell}}
+\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
+\newcommand{\incidencestorage}[1][n]{K_{#1,r}}
+\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
+\newcommand{\cycle}{C_{\ell,c}}
+\newcommand{\impedance}{x_\ell}
+\newcommand{\cycleprice}{\lambda_{c,t}}
 
@@
 \newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
 \newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
@@
 \newcommand{\cost}{\mathcal{C}}
-\newcommand{\opexgeneration}{\mathcal{O}}
+\newcommand{\opex}{\mathcal{O}}
+\newcommand{\opexgeneration}{\mathcal{O}^G}
 \newcommand{\opexflow}{\mathcal{O}^F}
+\newcommand{\opexstorage}{\mathcal{O}^E}
 \newcommand{\capexgeneration}{\mathcal{C}^G}
 \newcommand{\capexflow}{\mathcal{C}^F}
+\newcommand{\capexstorage}{\mathcal{C}^E}
 \newcommand{\emissioncost}{\mathcal{E}}
@@
 \title{Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable}
-\author{F. Hofmann}
+\author{F. Hofmann, Brown T.}
 
@@
 \totalcost &= \opexgeneration + \emissioncost +  \capexgeneration +  \capexflow + ...
-\label{eq:total_cost_terms}
+\label{eq:total_cost}
 \end{align}
@@
 \begin{align}
-    \cost = \sum_i \cost_i
-    \label{eq:asset_costs}
+    \cost = \sum_{i} \cost_{i}
+    \label{eq:asset_cost}
 \end{align} 
-For example the total OPEX, $\opexgeneration = \sum_s \opexgeneration_s$, is the combined OPEX of all generators $s$.
+For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.
 
-Minimizing the total cost $\totalcost$  of a network, which translates to maximizing the total welfare, will determine the single expenditure $\cost_i$ dedicated to asset $i$.
-with maximized welfare.
-Then, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$.
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$.
 It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
@@
 \lmp = \pdv{\,\totalcost}{\,\demand}
-\label{eq:total_cost_derivative}
+\label{eq:lmp}
 \end{align}
@@
 \begin{align}
-\totalcost - \remainingcost &=  \sum_{n,a,t} \lmp \, \demand
-\label{eq:total_cost_sum}
+\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
+\label{eq:total_cost_reduced}
 \end{align}
@@
 % 
-Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits in certain regions of the network, or minimum share of one technology in the power mix.
+Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
 However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
-Feeding \cref{eq:total_cost_terms} into \cref{eq:total_cost_derivative} illustrates that the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission (and other possible expenditures), schematically shown in \cref{fig:price_decomposition}.
-Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection.
-However the question of how the LMP translates to the single cost terms $\cost_i$ associated to asset $i$ in the network remains unanswered.
-This work aims at presenting and illustrating a an intuitive scheme how to cost breakdown the cost decomposition to a full peer-to-peer (P2P) cost allocation including all network assets.
+From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission (and other possible expenditures).
+This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
 
-\section{Nodal Cost Allocation}
+However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
+This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.
+
+\section{Dispatch-Based Cost Allocation}
 \label{sec:theory}
+\begin{subequations}\label[subequations]{eq:general_scheme}
 
-In following we show how to disaggregate the cost terms $\opexgeneration$, $\emissioncost$, $\capexgeneration$ and $\capexflow$ for asset $i$ into contributions of demand $\demand$ at bus~$n$ and time~$t$.
-That is what amount consumers at $\left(n, t\right)$ pay for the operation, investment and other cost-intensive, but welfare contributing, actions at asset $i$.
-We follow the general procedure: For each cost term $\cost$ we break down the contributions of asset $i$, \ie define \cref{eq:asset_costs}.
-According to \cref{eq:total_cost_derivative,eq:total_cost_sum}, we define its derivative with respect to $\demand$ and multiply the result with $\demand$.
-This returns $\cost_{n\rightarrow i,t}$, the contribution of $\demand$ to the cost term $\cost_i$.
-Follows naturally that when summing up all contributions the cost term $\cost_i$ must be covered, \ie
+Let $\cost_{i}$ denote a general cost term associated with asset $i$.
+Consider a long-term equilibrium in a power system with perfect competition, then, according to the zero-profit condition, each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$, \ie
+\begin{align}
+    \cost_{i} = \sum_t  \costfactor \, \state
+    \label{eq:cost_decomposition}
+\end{align}
+where $c_{i,t}$ denotes a general cost factor in \euro/MW. 
+For OPEX the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$ for asset $i$.
+However for CAPEX, $c_{i,t}$ is a composition of shadow prices $\mu_i$, given by the KKT variables of the capacity limit constraints.
+As we will show later the composition of $\mu_i$ must be determined for each cost term individually.
+
+% where $c_{i,t}$ is either a fixed operational price or a composition of shadowprices
+Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\cost_{n\rightarrow i,t}$ that consumers at bus $n$ have pay to asset $i$ at time $t$ as  
+\begin{align}
+    \cost_{n \rightarrow i,t} = \costfactor \,\pdv{\state}{\demand} \demand
+\end{align}
+Therefore, $\cost_{n\rightarrow i,t}$ indicates the contribution of $\demand$ to the cost term $\cost_{i}$.
+It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$.
+Without further specifying, we define the derivative multiplied with the nodal demand as
+\begin{align}
+    \allocatestate = \pdv{\state}{\demand} \demand
+    \label{eq:allocate_state}
+\end{align}    
+This quantity may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with.
+From the natural fact that the sum of all contributions must return the cost term, 
 \begin{align}
@@
 \end{align}
-and the total money spent at $\left(n,t\right)$ for the cost term $\cost$ amounts 
+it follows that $\allocatestate$ must fulfill
+\begin{align}
+    \state = \sum_n \allocatestate
+    \label{eq:allocate_state_constraint}
+\end{align}
+Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
 \begin{align}
@@
 \end{align}
-Note that for allocating CAPEX, we will make use of the shadowprices associated to $\cost_i$, given by the dual values at cost optimum $\bm{\mu}$.
+\end{subequations}
 
-\subsection{Allocate Operation Expenditures}
+\begin{table*}[t]
+    \begin{center}
+        \begin{tabular}{l|c|c|c|c|c}
+        & $i$ & $\cost$ & $\cost_i$  & $\costfactor$ & $\state$  \\
+        \toprule 
+        OPEX Production & $s$ & $\opexgeneration$ & $\sum_{t} \operationalpricegeneration \, \generation$   & $\operationalpricegeneration$ & $\generation$ \\  
+        OPEX Transmission  & $\ell$ & $\opexflow$ & $\sum_{t} \operationalpriceflow \, |\flow|  $ & $\operationalpriceflow$ & $|\flow|$ \\  
+        OPEX Storage & $r$  & $\opexstorage$ & $\sum_{t} \operationalpricestorage \, \storagedispatch$ &  $\operationalpricestorage$ & $\storage$ \\
+        \midrule   
+        CAPEX Production & $s$ & $\capexgeneration$ & $ \capitalpricegeneration \capacitygeneration$ & $\muuppergeneration$ & $\generation$ \\
+        CAPEX Transmission & $\ell$ & $\capexflow$ & $ \capitalpriceflow \capacityflow$ & $\left(\muupperflow - \mulowerflow \right)$ & $\flow$ \\
+        CAPEX Storage & $r$ & $\capexstorage$ & $ \capitalpricestorage \capacitystorage$ & $ (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch$ & $\storage$ \\
+        \midrule
+        Emission Cost & $s$ & $\emissioncost$ & $ \emissionprice \, \emission \, \generation$ & $\emissionprice \,\emission$ & $\generation$ \\
+        % \bottomrule   
+    \end{tabular}
+    \end{center}
+    \caption{Mapping of different cost terms to the cost allocation scheme given in \cref{eq:general_scheme}.
+These include OPEX \& CAPEX for production, transmission and storage assets in the network, as well as a cost term for the total Green House Gas (GHG) emissions.}
+    \label{tab:cost_allocation_map}
+\end{table*}
 
-The total OPEX is given by the operational price $\operationalpricegeneration$ times the generation $\generation$ of all generators $s$, 
+Now, assume a network with generators $s$, transmission lines $\ell$ and storage units $r$.
+Each asset $i = \{s, \ell, r\}$ adds an term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
+The operational price for an asset $i$ is given by $o_i$, the capital price for one unit capacity expansion by $c_i$.
+Then, for example, the OPEX occasioned by generator $s$ is given by
 \begin{align}
- \opexgeneration& = \sum_{s,t} \operationalpricegeneration\,  \generation 
-\label{eq:opex_generation}
+    \opexgeneration_s = \sum_t \operationalpricegeneration \, \generation 
+    \label{eq:opexgeneration}
 \end{align}
-
-Hence, the contribution of $\demand$ to the OPEX at generator $s$ is straightforward given by
+where $\generation$ denotes its production at time $t$.
+It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition}.
+This allows us to use the above presented scheme in \cref{eq:general_scheme} and formulate a full cost allocation for the OPEX on production.
+As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by
 \begin{align}
     \allocateopex[n \rightarrow s] &= 
-   \operationalpricegeneration \,  \pdv{\generation}{\demand} \, \demand
+   \operationalpricegeneration \,  \allocatepeer
 \label{eq:allocate_opexGeneration_detailed}
 \end{align}
-
-We define the derivative on the right hand side times the nodal demand as 
+The quantity  
 \begin{align}
@@
 \end{align}
-% 
-The quantity $\allocatepeer$ can be considered as a power allocation, that is the power produced by generator $s$ and consumed at node $n$ and time $t$.
-By feeding back \cref{eq:allocate_opexGeneration_detailed} into \cref{eq:opex_generation}, we see that the power allocation has to fulfill
+can be considered as a power allocation which indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
+It follows naturally that, in the same manner, the scheme can be applied to the OPEX on the flow $\flow$ and storage unit dispatch $\storagedispatch$.
+As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow.
+The upper half in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
+Note, in all cases, the allocation $\allocatestate$ remains undefined which we will resolve later.
+
+For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and the capacity limit.
+All quantities for the CAPEX allocation, which we discuss in detail in the following, are summarized in the lower half of \cref{tab:cost_allocation_map}.
+
+\subsubsection*{Allocate CAPEX on Generators}
+
+The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
 \begin{align}
-\generation = \sum_n \allocatepeer
-\label{eq:allocate_peer_constraint}
+\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
+\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
+- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
+\label[constraint]{eq:lower_generation_capacity_constraint}
 \end{align}
-% 
-% 
-The definition of $\allocateopex[n \rightarrow s]$ in \cref{eq:allocate_opexGeneration_detailed} can be interpreted as the following.
-Consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs.
-The cost allocation behaves like P2P tradings between producers and consumers for fixed production prices.
-In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$, thus
+where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
+At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.
+As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
 \begin{align}
-\sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
-\label{eq:no_profit_opex}
+ \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
+ \label{eq:no_profit_capex_generation}
+\end{align}
+This representation connects the CAPEX with the operational state of generator $s$ and matches the needed form in \cref{eq:cost_decomposition} for using the allocation scheme.
+The resulting the CAPEX allocation is given by
+\begin{align}
+ \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocatepeer
+ \label{eq:allocate_capexGeneration_detailed}
+\end{align}
+How does this allocation behave?
+According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
+If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
+Thus it is these times steps which push investments in $\capacitygeneration$.
+If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
+When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
+
+\subsubsection*{Allocate CAPEX on Transmission Lines}
+
+The transmission capacity $\capacityflow$ limits the flow $\flow$ in both directions,
+\begin{align}
+\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
+\label[constraint]{eq:upper_flow_capacity_constraint} \\
+- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
+\label[constraint]{eq:lower_flow_capacity_constraint}
+\end{align}
+which yield the shadow prices $\muupperflow$ and $\mulowerflow$.
+Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow
+\begin{align}
+\capexflow_\ell = \capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
+\label{eq:no_profit_capex_flow}
+\end{align}
+Again, we follow the scheme in \cref{eq:general_scheme} which  defines the CAPEX allocation as 
+\begin{align}
+    \allocatecapexflow[n \rightarrow \ell] &=  
+   \left( \muupperflow - \mulowerflow\right) \, \allocateflow
+   \label{eq:allocate_capexFlow_detailed}
+\end{align}
+The quantity 
+\begin{align}
+ \allocateflow =  \pdv{\flow}{\demand}\, \demand
+\end{align}
+can be interpreted as the flow that the demand at $\left(n,t\right)$ causes on line $\ell$.
+% According to \cref{eq:allocate_state_constraint} it has to fulfill 
+% \begin{align}
+%  \flow = \sum_{n} \allocateflow
+% \label{eq:allocate_flow_constraint}
+% \end{align}
+The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
+Hence, the definition of $\allocatecapexflow[n \rightarrow \ell]$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
+Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
+This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_flow}
+
+\subsubsection*{Allocate CAPEX on Storages}
+
+% The complementary slackness of the above constraints,  
+% \begin{align}
+%     \capexstorage_r = \capitalpricestorage \, \capacitystorage = \sum_t \muupperstoragedispatch \, \storagedispatch + \muupperstoragecharge \, \storagecharge + \muupperstoragesoc \, \storagesoc
+% \end{align}
+
+We us the result of Appendix B.3 in \cite{brown_decreasing_2020} which states that a storage recovers its capital (here we added the operational) costs from aligning dispatch and charging to the LMP, thus
+\begin{align}
+    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
+\end{align}
+where $\incidencestorage$ is one if storage $s$ is at node $n$ and zero otherwise. Using \cref{eq:lagrangian_storage_dispatch}, we can reformulate to 
+\begin{align}
+    \capitalpricestorage \, \capacitystorage = \sum_t \incidencestorage \left( (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch \right) \storagedispatch 
 \end{align}
 
-\subsection{Allocate Emission Expenditures}
+% \subsection{Allocate Operational Expenditures for Generators}
+
+% Assume the cost-minimal power system includes multiple generators whose dispatch $\generation$ is optimized.
+The cost term added to $\totalcost$ combines the OPEX for all generators $s$ over the whole time span
+% \begin{align}
+%     \opexgeneration =  \sum_{s,t} \operationalpricegeneration \, \generation
+% \end{align}   
+% where $\operationalpricegeneration$ denotes the fixed operational price for generator $s$ to produce 1~MW of electricity.
+The full cost allocation for the operation of generator $s$, arises from replacing
+% \begin{subequations}
+% \label[relation]{eq:opex_allocation}
+% \begin{align}
+%     i &\Rightarrow s \\
+%     \costfactor &\Rightarrow \operationalpricegeneration \\
+%     \state &\Rightarrow \generation 
+% \end{align}
+% \end{subequations}
+% % $i$ by $s$, $\costfactor$ by $\operationalpricegeneration$, $\cost$ by $\opexgeneration$ and $\state$ by $\generation$
+% in \cref{eq:general_scheme}. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
+% \begin{align}
+%     \allocateopex[n \rightarrow s] &= 
+%    \operationalpricegeneration \,  \allocatepeer
+% \label{eq:allocate_opexGeneration_detailed}
+% \end{align}
+% where the quantity  
+% \begin{align}
+%  \allocatepeer = \pdv{\generation}{\demand} \, \demand
+%  \label{eq:allocate_peer}
+% \end{align}
+% obtained through \cref{eq:allocate_state}, can be considered as a power allocation. It indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
+% From \cref{eq:allocate_state_constraint} it follows that all the power allocations must sum up to the power produced by generator $s$, \ie
+% \begin{align}
+% \generation = \sum_n \allocatepeer
+% \label{eq:allocate_peer_constraint}
+% \end{align}
+% % 
+% % 
+% The cost allocation $\allocateopex[n \rightarrow s]$ assumes peer-to-peer tradings between producers and consumers for fixed production prices $\operationalpricegeneration$.
+% Hence, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs.
+In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$,
+% \begin{align}
+% \sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
+% \label{eq:no_profit_opex}
+% \end{align}
+
+\subsubsection*{Allocate Emission Expenditures}
 
@@
 where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
-Similar to the allocated OPEX, the emission cost for consumers at bus $n$ at time $t$ for operation at generator $s$ is allocated through
+We use the same set of equation as given through \cref{eq:general_scheme,eq:opex_allocation} but update the cost factor to $\costfactor \Rightarrow \emissionprice \, \emission$.
+For consumers at bus $n$ at time $t$ this leads to the emission cost of
 \begin{align}
@@
 \end{align}
-
-\subsection{Allocate Capital Expenditures for Generators}
-
-The optimization of the generation capacity $\capacitygeneration$ adds the cost term 
-\begin{align}
- \capexgeneration& = \sum_{s} \capitalpricegeneration \capacitygeneration 
-\end{align}
-to $\totalcost$. It combines all capacities times their capital cost $\capitalpricegeneration$. The nominal capacity constraints the generation $\generation$ in the form of 
-\begin{align}
-\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
-\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
-- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
-\label[constraint]{eq:lower_generation_capacity_constraint}
-\end{align}
-where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation},  the capital cost  for generator $s$ is payed back by the generation times the shadow price for the upper capacity constraint,
-\begin{align}
- \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
- \label{eq:no_profit_capex_generation}
-\end{align}
-With our definition in \cref{eq:allocate_peer} the allocated CAPEX can then be defined as 
-\begin{align}
- \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocatepeer
- \label{eq:allocate_capexGeneration_detailed}
-\end{align}
-
-So, how does this allocation behave?
-According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
-Thus it is these times steps which push investments in $\capacitygeneration$.
-If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_peer_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
-
-\subsection{Allocate Capital Expenditures for Transmission Lines}
-
-The CAPEX for the transmission system is given by 
-\begin{align}
- \capexflow & = \sum_\ell \capitalpriceflow \capacityflow
- \label{eq:capex_flow}
-\end{align}
-The transmission capacity variable $\capacityflow$ limits the flow $\flow$ in both directions, such that 
-\begin{align}
-\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
-\label[constraint]{eq:upper_flow_capacity_constraint} \\
-- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
-\label[constraint]{eq:lower_flow_capacity_constraint}
-\end{align}
-We use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that the investment in line $\ell$ is payed back by the shadowprices times the flow 
-\begin{align}
-\capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
-\label{eq:no_profit_capex_flow}
-\end{align}
-We insert this into \cref{eq:capex_flow} and define the individual cost terms  as 
-\begin{align}
-    \allocatecapexflow[n \rightarrow \ell] &=  
-   \left( \muupperflow - \mulowerflow\right) \,  \pdv{\flow}{\demand}\, \demand
-   \label{eq:allocate_capexFlow_detailed}
-\end{align}
-Again, we define the derivative on the right side times the demand as 
-\begin{align}
- \allocateflow =  \pdv{\flow}{\demand}\, \demand
-\end{align}
-which from inserting into \cref{eq:no_profit_capex_flow} has to fulfill 
-\begin{align}
- \flow = \sum_{n} \allocateflow
-\label{eq:allocate_transaction_constraint}
-\end{align}
-$\allocateflow$ can be interpreted as the flow that the demand at $\left(n,t\right)$ induces on line $\ell$.
-The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
-Hence, the definition of $\allocatecapexflow[n \rightarrow \ell]$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
-Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
-This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:allocate_transaction_constraint,eq:no_profit_capex_flow}
+associated with generator $s$.
 
@@
 But so far, the question of how $\allocatepeer$ and $\allocateflow$ are defined was left open.
-We recap that both rely on derivatives of $\generation$ and $\flow$  with respect to the nodal demand $\demand$ fulfilling \cref{eq:allocate_peer_constraint,eq:allocate_transaction_constraint}.
+We recap that both rely on derivatives of $\generation$ and $\flow$  with respect to the nodal demand $\demand$ fulfilling \cref{eq:allocate_state_constraint}.
 \\
@@
 In fact, it turns out that ones, the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
-In order to show this, let $\ptdf$ denote Power Transfer Distribition Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
+In order to show this, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
 Further let $\incidencegenerator$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise.
-The linear power flow equals the
+The linear power flow equation can be written as
 \begin{align}
@@
 
+For an simplified storage model, the upper capacity $\capacitystorage$ limits the discharging dispatch $\storagedispatch$, the storing power $\storagecharge$ and state of charge $\storagesoc$ of a storage unit $r$ by
+\begin{align}
+    \storagedispatch - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragedispatch} \\
+    \storagecharge - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragecharge} \\
+    \storagesoc - h_r \, \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragesoc}
+\end{align}
+where we assume a fixed ratio between dispatch and storage capacity of $h_r$. 
+From stationarity we obtain 
+\begin{align}
+    \pdv{\lagrangian}{\capacitystorage} &= 0 \\
+    \capitalpricestorage &= \sum_t \left( \muupperstoragedispatch + \muupperstoragecharge + h_r \muupperstoragesoc  \right)
+\end{align}
+\begin{align}
+    \pdv{\lagrangian}{\storagedispatch} &= 0 \\
+    \operationalpricestorage +& \lmp + \mulowerstoragedispatch - \muupperstoragedispatch - (\efficiencydispatch )^{-1} \mustateofcharge = 0  
+    \label{eq:lagrangian_storage_dispatch}
+\end{align}
+\begin{align}
+    \pdv{\lagrangian}{\storagecharge} &= 0 \\
+    - \lmp +& \mulowerstoragecharge - \muupperstoragecharge + \efficiencycharge \mustateofcharge = 0 
+    \label{eq:lagrangian_storage_charge}
+\end{align}
+
 \subsection{Brownfield Optimization and Capacity Restricitions}

### 2020-08-09 code/allocate_network.py

@@ -9,6 +9,7 @@ Created on Fri Jul 10 20:56:10 2020
 import netallocation as ntl
 import pypsa
 import xarray as xr
+import pandas as pd
 
 from netallocation.utils import (reindex_by_bus_carrier as by_bus_carrier,
                                  get_as_dense_by_bus_carrier)
@@ -19,22 +20,25 @@ from netallocation.convert import peer_to_peer, virtual_patterns
 from netallocation.cost import (nodal_co2_price, snapshot_weightings,
                                 allocate_one_port_operational_cost,
                                 allocate_branch_investment_cost)
-from config import source_dims
+from config import source_dims, sink_dims
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('allocate_network', nname='de10gf',
+        snakemake = mock_snakemake('allocate_network', nname='acdc',
                                    method='ptpf', power='net')
 
 n = pypsa.Network(snakemake.input.network)
 
 method = snakemake.wildcards.method
 aggregated = snakemake.wildcards.power == 'net'
-co2_price = snakemake.config['costs']['emission_prices']['co2']
+if 'co2_limit' in n.global_constraints.index:
+    co2_price = 0
+else:
+    co2_price = snakemake.config['costs']['emission_prices']['co2']
 
 
-ds = ntl.allocate_flow(n, method=method, aggregated=aggregated, dask=False)  # q=0
+ds = ntl.allocate_flow(n, method=method, aggregated=aggregated)  # q=0
 ds = ds.chunk(dict(snapshot=5))
 
 A = expand_by_source_type(peer_to_peer(ds, n).peer_to_peer, n, chunksize=5)
@@ -50,7 +54,7 @@ A_emission = A * ep
 # opex
 comps = ['Generator', 'StorageUnit', 'Store']
 o = get_as_dense_by_bus_carrier(n, 'marginal_cost', comps).rename(source_dims)
-o = o - ep.reindex_like(o, fill_value=0)
+o = o * snapshot_weightings(n) - ep.reindex_like(o, fill_value=0)
 A_opex = A * o
 
 
@@ -59,15 +63,28 @@ c = 'Generator'
[…]

### 2020-08-09 code/plot_expenditure_maps.py

@@ -18,7 +18,7 @@ import os
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_expenditure_maps', nname='de10gf',
+        snakemake = mock_snakemake('plot_expenditure_maps', nname='test-de10gf',
                                    method='ptpf', power='net')
 
 nice_names = snakemake.config['plotting']['nice_names']

### 2020-08-09 code/plot_price_map.py

@@ -18,15 +18,15 @@ import os
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_price_maps', nname='de10gf',
+        snakemake = mock_snakemake('plot_price_maps', nname='test-de10gf',
                                    method='ptpf', power='net')
+
 
 n = pypsa.Network(snakemake.input.network)
 regions = gpd.read_file(snakemake.input.regions)
 
-payments = xr.open_dataset(snakemake.input.payments)
-demand = ntl.power_demand(n, per_carrier=True).rename(bus="sink")\
-            .sel(carrier='Load', drop=True)
+payments = xr.open_dataset(snakemake.input.payments).sum('sink_carrier')
+demand = ntl.power_demand(n).rename(bus="sink")\
 
 prices = (payments/demand).mean('snapshot').to_dataframe()
 prices = prices.assign(lmp = n.buses_t.marginal_price.mean())

### 2020-08-09 code/solve_tiny_network.py

@@ -1,0 +1,22 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jul 20 17:50:06 2020
+
+@author: fabian
+"""
+
+import pypsa
+import netallocation as ntl
+
+
+n = ntl.test.get_network_ac_dc()
+n.generators['p_nom_min'] = 500
+
+# n.carriers = n.carriers.drop('battery')
+n.add('StorageUnit', 'battery', bus='Frankfurt', capital_cost=50,
+      p_nom_extendable=True, marginal_cost=10, carrier='battery')
+
+n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
+
+n.export_to_netcdf(snakemake.output[0])

### 2020-08-10 paper/flow_based.tex
update paper gitignore test figures

@@
 % allocation quantities
-\newcommand{\allocatepeer}[1][s \rightarrow n]{A_{#1,t}}
+\newcommand{\allocatepeer}[1][s, n]{A_{#1,t}}
 \newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
@@
 The operational price for an asset $i$ is given by $o_i$, the capital price for one unit capacity expansion by $c_i$.
-Then, for example, the OPEX occasioned by generator $s$ is given by
+For example the OPEX occasioned by generator $s$ is given by
 \begin{align}
@@
 where $\generation$ denotes its production at time $t$.
-It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition}.
-This allows us to use the above presented scheme in \cref{eq:general_scheme} and formulate a full cost allocation for the OPEX on production.
+It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme} for a full cost allocation of OPEX on production.
 As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by
@@
 As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow.
-The upper half in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
+The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
 Note, in all cases, the allocation $\allocatestate$ remains undefined which we will resolve later.
@@
 For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and the capacity limit.
-All quantities for the CAPEX allocation, which we discuss in detail in the following, are summarized in the lower half of \cref{tab:cost_allocation_map}.
+All quantities for the CAPEX allocation, which we discuss in detail in the following, are summarized in the middle section of \cref{tab:cost_allocation_map}.
 
@@
 The presented cost allocation suits for any type of topology and network setup.
-But so far, the question of how $\allocatepeer$ and $\allocateflow$ are defined was left open.
-We recap that both rely on derivatives of $\generation$ and $\flow$  with respect to the nodal demand $\demand$ fulfilling \cref{eq:allocate_state_constraint}.
+But so far, the question of how $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.
+We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:allocate_state_constraint}).
 \\

### 2020-08-11 paper/flow_based.tex
fix non extendable SUs to extendable

@@
 % allocation quantities
-\newcommand{\allocatepeer}[1][s, n]{A_{#1,t}}
+\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
+\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
 \newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
@@
     \allocateopex[n \rightarrow s] &= 
-   \operationalpricegeneration \,  \allocatepeer
+   \operationalpricegeneration \,  \allocategeneration
 \label{eq:allocate_opexGeneration_detailed}
@@
 \begin{align}
- \allocatepeer = \pdv{\generation}{\demand} \, \demand
+ \allocategeneration = \pdv{\generation}{\demand} \, \demand
  \label{eq:allocate_peer}
@@
 \begin{align}
- \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocatepeer
+ \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocategeneration
  \label{eq:allocate_capexGeneration_detailed}
@@
 
-We us the result of Appendix B.3 in \cite{brown_decreasing_2020} which states that a storage recovers its capital (here we added the operational) costs from aligning dispatch and charging to the LMP, thus
+We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus 
 \begin{align}
@@
 %     \allocateopex[n \rightarrow s] &= 
-%    \operationalpricegeneration \,  \allocatepeer
+%    \operationalpricegeneration \,  \allocategeneration
 % \label{eq:allocate_opexGeneration_detailed}
@@
 % \begin{align}
-%  \allocatepeer = \pdv{\generation}{\demand} \, \demand
+%  \allocategeneration = \pdv{\generation}{\demand} \, \demand
 %  \label{eq:allocate_peer}
@@
 % \begin{align}
-% \generation = \sum_n \allocatepeer
+% \generation = \sum_n \allocategeneration
 % \label{eq:allocate_peer_constraint}
@@
 \begin{align}
- \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocatepeer
+ \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocategeneration
 \end{align}
@@
 \begin{align}
- \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer[m \rightarrow n]  - \delta_{n,m} \demand \right) 
+ \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
  \label{eq:allocate_peer_to_allocate_flow}
 \end{align}
-where $\allocatepeer[m \rightarrow n] = \sum_s \incidencegenerator \, \allocatepeer$ combines all power allocations of the source $m$.
-\Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocatepeer$ directly determines the flow allocation on line $\ell$.
+where $\allocatepeer = \sum_s \incidencegenerator \, \allocategeneration$ combines all power allocations of the source $m$.
+\Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocategeneration$ directly determines the flow allocation on line $\ell$.
 
-This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocatepeer$.
+This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocategeneration$.
 In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$.
@@
 The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal.
-Therefore $\allocatepeer$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
+Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
 Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation.
@@
 \begin{align}
-    \allocatepeer[s \rightarrow n] = \dfrac{\generation}{\sum_s \generation} \demand 
+    \allocategeneration = \dfrac{\generation}{\sum_s \generation} \demand 
 \end{align}
@@
 \begin{align}
-\allocatepeer[m \rightarrow n] &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
+\allocatepeer &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
 \label{eq:mp_slack}
@@
 
-The allocation $\allocatepeer$ from generator $s$ to $n$, is given by multiplying $\allocatepeer[m \rightarrow n]$ with the nodal share $\generation / \nodalgeneration$.
+The allocation $\allocategeneration$ from generator $s$ to $n$, is given by multiplying $\allocatepeer$ with the nodal share $\generation / \nodalgeneration$.
 
@@
 \begin{align}
-\allocatepeer[m \rightarrow n] = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
+\allocatepeer = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
 \end{align}
@@
 \begin{align}
-\allocatepeer[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
+\allocategeneration[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
 \end{align}

### 2020-08-11 code/plot_network.py

@@ -37,7 +37,6 @@ if is_brownfield:
                              figsize=(10, 5))
     # existent
     bus_sizes = n.generators.groupby(["bus", "carrier"]).p_nom_min.sum()
-    bus_sizes = bus_sizes.reindex(n.buses.index, level=0, fill_value=0)
     plot = n.plot(
         bus_sizes=bus_sizes * bus_scale,
         line_widths=n.lines.s_nom_min * branch_scale,

### 2020-08-11 code/plot_total_costs.py

@@ -5,7 +5,8 @@ Created on Fri Jul 10 21:56:19 2020
 
 @author: fabian
 """
-
+import pypsa
+import pandas as pd
 import xarray as xr
 import matplotlib.pyplot as plt
 from config import color, to_total_symbol
@@ -13,11 +14,37 @@ from config import color, to_total_symbol
 plt.rc('font', family='serif')
 plt.rc("text", usetex=False)
 
-ca = xr.open_dataset(snakemake.input.costs)
 
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('plot_total_costs', nname_wo_field='de10')
+
+ngf = pypsa.Network(snakemake.input.network_gf)
+nbf = pypsa.Network(snakemake.input.network_bf)
+cagf = xr.open_dataset(snakemake.input.costs_gf)
+cabf = xr.open_dataset(snakemake.input.costs_bf)
+
+# %%
 fig, ax = plt.subplots()
-ca.sum().to_array().to_series().reindex_like(color).dropna()\
-    .rename(to_total_symbol).rename('')\
-    .plot.pie(colors=color, explode=[.01]*len(ca), autopct='%1.0f%%')
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+
+TC_gf = ngf.objective_constant + ngf.objective
+TC_bf = nbf.objective_constant + nbf.objective
+TC = [TC_gf, TC_bf]
+
+stacked_gf = (cagf.sum().to_array().to_series().reindex_like(color).dropna()
+              .rename(to_total_symbol).rename('').to_frame().T)
+stacked_bf = (cabf.sum().to_array().to_series().reindex_like(color).dropna()
+              .rename(to_total_symbol).rename('').to_frame().T)
+stacked_cost = pd.concat([stacked_gf, stacked_bf], keys=['Greenfield', 'Brownfield'])
+
+R = TC - stacked_cost.sum(1)
+stacked_cost[r'$\mathcal{R}$'] = R
+
+stacked_cost.plot.bar(color=color, stacked=True, ax=ax, zorder=4)
+ax.legend(loc='lower center', bbox_to_anchor=(0.5, 1), ncol=5, frameon=False)
+ax.grid(axis='y', linestyle='dashed', color='gray', zorder=1, alpha=.5)
 
 fig.savefig(snakemake.output[0])

### 2020-08-11 paper/flow_based.tex
fix .gitignore
update numerical example

@@
 \newcommand{\incidencestorage}[1][n]{K_{#1,r}}
+\newcommand{\incidenceasset}[1][n]{K_{#1,i}}
 \newcommand{\ptdf}[1][n]{H_{\ell,#1}}
@@
 \newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
+\newcommand{\allocatestoragedispatch}[1][r, n]{A_{#1,t}}
 \newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
@@
 \newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
-\newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
-\newcommand{\allocatecapexflow}[1][n]{\mathcal{C}^{F}_{#1,t}}
-\newcommand{\allocateopex}[1][n]{\mathcal{O}_{#1,t}}
-\newcommand{\allocateemissioncost}[1][n]{\mathcal{E}_{#1,t}}
+\newcommand{\allocatecost}[1][n \rightarrow i]{\mathcal{C}_{#1, t}}
+\newcommand{\allocatecapexgeneration}[1][n \rightarrow s]{\mathcal{C}^{G}_{#1,t}}
+\newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\mathcal{C}^{F}_{#1,t}}
+\newcommand{\allocateopex}[1][n \rightarrow s]{\mathcal{O}_{#1,t}}
+\newcommand{\allocateemissioncost}[1][n \rightarrow s]{\mathcal{E}_{#1,t}}
 
@@
 % where $c_{i,t}$ is either a fixed operational price or a composition of shadowprices
-Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\cost_{n\rightarrow i,t}$ that consumers at bus $n$ have pay to asset $i$ at time $t$ as  
+Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$ as  
 \begin{align}
-    \cost_{n \rightarrow i,t} = \costfactor \,\pdv{\state}{\demand} \demand
+    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
 \end{align}
-Therefore, $\cost_{n\rightarrow i,t}$ indicates the contribution of $\demand$ to the cost term $\cost_{i}$.
+Therefore, $\allocatecost$ indicates the contribution of $\demand$ to the cost term $\cost_{i}$.
 It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$.
@@
 \begin{align}
-    \cost_i = \sum_{n,t} \cost_{n \rightarrow i, t}
+    \cost_i = \sum_{n,t} \allocatecost
 \end{align}
@@
 \begin{align}
-    \cost_{n,t} = \sum_i \cost_{n \rightarrow i, t}
+    \cost_{n,t} = \sum_i \allocatecost
 \end{align}
@@
 \begin{align}
- \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocategeneration
+ \allocatecapexgeneration = \muuppergeneration \, \allocategeneration
  \label{eq:allocate_capexGeneration_detailed}
@@
 \begin{align}
-    \allocatecapexflow[n \rightarrow \ell] &=  
+    \allocatecapexflow &=  
    \left( \muupperflow - \mulowerflow\right) \, \allocateflow
@@
 The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
-Hence, the definition of $\allocatecapexflow[n \rightarrow \ell]$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
+Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
 Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
@@
 \begin{align}
- \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocategeneration
+ \allocateemissioncost = \emissionprice \, \emission \, \allocategeneration
 \end{align}
@@
 \\
-% 
-In fact, it turns out that ones, the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
-In order to show this, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-Further let $\incidencegenerator$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise.
+
+Let $\nodalgeneration[m]$ denote the nodal power generation which combines the power production of all producing assets, in this case generators $S$ and storages $R$, at node $n$ and time $t$.
+It is given by
+\begin{align}
+    \nodalgeneration[m] = \sum_{i \in \{S, R\}} \incidenceasset[m] \, \state
+\end{align}
+with $\incidenceasset[m]$ being 1 if asset $i$ is attached to bus $m$ and zero otherwise. Further, let $\allocatepeer$ collect the power produced by assets at node $m$ and consumed at $n$, given by 
+\begin{align}
+    \allocatepeer = \sum_{i \in \{S,R\}} \incidenceasset[m] \, \allocatestate  
+    % \nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation + \sum_r \incidencestorage[m] \storagedispatch \Forall{n}
+\end{align}
+Now, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
 The linear power flow equation can be written as
@@
 \end{align}
-where the nodal generation $\nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation $ combines the production of all generators attached to bus $m$.
-Taking the derivative with respect to the demand, leads to
+Note that for transport models or mixed AC-DC networks, $\ptdf$ can be artificially calculated using the formulation presented in \cite{hofmann_flow_2020-1}.
+Taking the derivative with respect to the demand, 
 \begin{align}
  \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
- \label{eq:allocate_peer_to_allocate_flow}
+ \label{eq:allocate_peer_to_allocate_flow},
 \end{align}
-where $\allocatepeer = \sum_s \incidencegenerator \, \allocategeneration$ combines all power allocations of the source $m$.
-\Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocategeneration$ directly determines the flow allocation on line $\ell$.
-
-This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocategeneration$.
-In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$.
-Unfortunately, the solution is non-unique and requires further assumptions.
+shows that $\allocateflow$ is fully determined through the peer-to-peer allocation $\allocatepeer$.
+In other words, we only need to know how much power produced at node $m$ is consumed at node $n$ in order to derive the allocated flow $\allocateflow$.
+Further we can breakdown $\allocatepeer$ to $\allocategeneration$ for generators and $\allocatestoragedispatch$ for storages proportionally to their contribution to the nodal generation $\nodalgeneration[m]$.
+Unfortunately, the solution for $\allocatepeer$ is non-unique and requires further assumptions.
 Established flow allocation schemes approach this problem from different directions.
@@
 % 
-Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $n$ with $\nodalgeneration \le \nodaldemand$ is not assigned to other buses, only to $n$. 
+Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $m$ with $\nodalgeneration[m] \le \nodaldemand[m]$ is not assigned to other buses, only to $m$. 
 Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
@@
 Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
-Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation.
-If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \generation \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.
+Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first step.
+If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.
 
@@
 
-In the following we showcase a two bus system with one optimized time step and its resulting allocated payments, illustrated in \cref{fig:example_network}. \\
+In the following, we illustrate the cost allocation and explain its functionality by taking the example of a two bus system with one optimized time step, shown in \cref{fig:example_network}. \\
 % 
@@
 The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
-    The sum of both net flows equals the resulting flow of $f=40$~MW. \vspace{-20pt}}
+    The sum of both net flows equals the resulting flow of $f_1=40$~MW. \vspace{-20pt}}
     \label{fig:example_allocation}
@@
 Consumers at bus $n$ have to pay each generator proportional to their consumption.
-As we only consider one time step the proportionality applies for OPEX $\mathcal{O}_n$ and CAPEX $\mathcal{C}_n$.
+As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_n$.
 As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further tranmission expansion, it is rewarded proportional to the relief.}
@@
 Consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator at bus~1.
-The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{2 \rightarrow 1, 1} = -20$.
+The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$.
 As this flow is in contrary direction to the total flow, it is relieving the transmission system.
-This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{2 \rightarrow 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.
+This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{\ell = 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.
 
@@
 \begin{align}
-    \capitalpricegeneration \capacitygeneration - \remainingcost_s = \sum_{n,t} \allocatecapexgeneration[n \rightarrow s] \Forall{s \in S}
+    \capitalpricegeneration \capacitygeneration - \remainingcost_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
 \end{align}
@@
 \begin{align}
-    \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow[n \rightarrow \ell] \Forall{\ell \in L}
+    \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow \Forall{\ell \in L}
 \end{align}

### 2020-08-11 code/examples.py

@@ -14,13 +14,15 @@ import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 from netallocation.plot import injection_plot_kwargs
-from config import to_static_symbol, color
+from config import to_static_symbol, color, to_total_symbol
 # plt.rc('tex', preamble=r'\Large')
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
 
-net = False
-altype = 'net' if net else 'gross'
+path = '../figures/example'
+
+net = True
+power = 'net' if net else 'gross'
 method = 'ptpf'
 
 
@@ -106,7 +108,7 @@ ax.legend(*handles_labels_for(color[ca_df.columns].rename(to_static_symbol)),
           ncol=3, loc='lower right', frameon=True, fontsize='large')
 ntl.plot.annotate_bus_names(n, ax, shift=0, bbox='fancy')
 fig.tight_layout()
-fig.savefig(f'../figures/example_network_{altype}_{method}.png', bbox_inches='tight')
+fig.savefig(f'{path}/example_network_{power}_{method}.png', bbox_inches='tight')
 
 
 # %% payoff matrix
@@ -128,9 +130,9 @@ for i, v in enumerate(payoff):
                 annot_kws={'horizontalalignment': 'left'})
     if i:
         ax.set_ylabel('')
-    ax.set_title(to_static_symbol[v])
+    ax.set_title(to_total_symbol[v])
 fig.tight_layout(w_pad=0)
-fig.savefig(f'../figures/example_payoff_{altype}_{method}.png', bbox_inches='tight')
+fig.savefig(f'{path}/example_payoff_{power}_{method}.png', bbox_inches='tight')
 
 # %% sub flows
 
@@ -151,24 +153,24 @@ for bus in n.buses.index:
 
     for b in n.buses.index:
         bbox.update({'facecolor': bcolor, 'edgecolor': 'None', 'pad':.5, 'alpha':1})
-        A = r'$A_{%s, %s} = %i$'%(b, bus, p2p[b])
+        A = r'$A_{%s \rightarrow %s} = %i$'%(b, bus, p2p[b])
         ax.text(*n.buses.loc[b, ['x', 'y']] + [0, 0.2], A, zorder=8, color='white',
-                ha='center', va='center', bbox=bbox)
+                ha='center', va='bottom', bbox=bbox)
 
     source = p2p.drop(bus).index.item()
 
[…]

### 2020-08-11 paper/flow_based.tex
further update on numerical example section

@@
 As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2.
-The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.
-\vspace{-10pt}}
+The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.}
 \label{fig:example_network}
@@
 The two buses are connected via one transmission line, each has one generator.
-Whereas the generator at bus~1 has an operational price of 50 \euro/\megawatthour, the generator at bus~2 has a higher operational price of 200~\euro/\megawatthour.
+Whereas generator~1 (at bus~1) has an operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a higher operational price of 200~\euro/\megawatthour.
 For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacitygenerationupper$~=~100~MW.
@@
 With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW.
-The 40~MW excess power not consumed at bus~1, flow to bus~2 where the generator is built with only 50~MW.
+The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW.
 % 
-\begin{figure}[h!]
-    \begin{subfigure}[c]{\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus1.png}
+% 
+% \begin{figure}[h]
+%     \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
+%     \label{fig:example_allocation_bus1}
+%     \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
+Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.}
+% \end{figure}
+% % 
+% \begin{figure}[b]
+%     \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
+%     \label{fig:example_allocation_bus2_gross_ebe}
+%     \caption{Type \ref{gross}\ref{ebe} power allocation for bus~2 of the example network in \cref{fig:example_network}.
+Bus~2 retrieves 60~MW from bus~1 and self supplies 30~MW.
+The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
+% \end{figure}
+% 
+\begin{figure*}[h!]
+    \begin{subfigure}[c]{.495\linewidth}
+    \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
     \vspace{-40pt}
@@
     \end{subfigure}
-    \begin{subfigure}[c]{\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus2.png}
+    \begin{subfigure}[c]{.495\linewidth}
+    \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
     \vspace{-40pt}
     \subcaption{}
-    \label{fig:example_allocation_bus2}
+    \label{fig:example_allocation_bus2_gross_ebe}
     \end{subfigure}
-    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
+    \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
 Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.
 The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
-    The sum of both net flows equals the resulting flow of $f_1=40$~MW. \vspace{-20pt}}
+    The sum of both net flows equals the resulting flow of $f_1=40$~MW. \vspace{-0pt}}
     \label{fig:example_allocation}
-\end{figure}
-% 
+\end{figure*}
 % 
@@
     \centering
-\includegraphics[width=\linewidth]{example_payoff.png}
+    \includegraphics[width=\linewidth]{example_payoff_gross_ebe.png}
     \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}.
@@
 Consumers at bus $n$ have to pay each generator proportional to their consumption.
-As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_n$.
+As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$.
 As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further tranmission expansion, it is rewarded proportional to the relief.}
@@
 \end{figure}    
+% 
 % 
@@
 The resulting P2P payments are given in \cref{fig:example_payoff}.
-The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1,1}=40$~MW at bus~1 are self-sustained.
-Consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator at bus~1.
+The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained.
+With only one generator at bus 1, consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator~1.
 The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$.
@@
 
-The lower graph \cref{fig:example_allocation_bus2} illustrates the impact of consumption at bus~2.
+The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2.
 As $d_2$ is higher than $d_1$ the retrievements from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
+    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when equivalently allocation net power injection (scheme \ref{ebe}\ref{net}).}
+    \label{fig:example_payoff_net_ebe}
+    \end{figure}    
+% 
 In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments.
@@
 Bus~2 in constrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
-\begin{figure}[h]
-\centering
-\includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
-\caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when equivalently allocation net power injection (scheme \ref{ebe}\ref{net}).}
-\label{fig:example_payoff_net_ebe}
-\end{figure}
-
+% 
 Again the cummulative payments per bus meet the nodal spendings $\lmp \, \demand$. The cummulative revenues per generator and transmission line meet the all CAPEX and OPEX.

### 2020-08-11 code/examples.py

@@ -23,7 +23,7 @@ path = '../figures/example'
 
 net = True
 power = 'net' if net else 'gross'
-method = 'ptpf'
+method = 'ebe'
 
 
 n = pypsa.Network()
@@ -108,12 +108,12 @@ ax.legend(*handles_labels_for(color[ca_df.columns].rename(to_static_symbol)),
           ncol=3, loc='lower right', frameon=True, fontsize='large')
 ntl.plot.annotate_bus_names(n, ax, shift=0, bbox='fancy')
 fig.tight_layout()
-fig.savefig(f'{path}/example_network_{power}_{method}.png', bbox_inches='tight')
+fig.savefig(f'{path}/example_network.png', bbox_inches='tight')
 
 
 # %% payoff matrix
 
-to_indices = pd.Series(dict(payer='n', receiver_nodal_cost='m',
+to_indices = pd.Series(dict(payer='n', receiver_nodal_cost='s',
                             receiver_transmission_cost='$\ell$'))
 fig, axes = plt.subplots(1, 3, figsize=(5,2.5), sharey=True,
                           gridspec_kw={'width_ratios':[0.4,.4,.2]})

### 2020-08-13 paper/flow_based.tex
further work on chapter storage, seems good now

@@
 \newcommand{\state}{s_{i,t}}
-\newcommand{\costfactor}{c_{i,t}}
+\newcommand{\capacity}{S_{i}}
+\newcommand{\costfactor}{\gamma_{i,t}}
 
@@
 \newcommand{\storagesoc}{\storage^\text{ene}}
+\newcommand{\storageprevioussoc}{g_{r,t-1}^\text{ene}}
 
@@
 \newcommand{\mustateofcharge}{\lambda^\text{ene}_{r,t}}
+\newcommand{\munextstateofcharge}{\lambda^\text{ene}_{r,t+1}}
 
@@
 \newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\mathcal{C}^{F}_{#1,t}}
+\newcommand{\allocatecapexstorage}[1][n \rightarrow r]{\mathcal{C}^{E}_{#1,t}}
 \newcommand{\allocateopex}[1][n \rightarrow s]{\mathcal{O}_{#1,t}}
@@
 \end{align}
-In turn, each single cost term on the right side $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
+In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
 \begin{align}
@@
 \end{align} 
-For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.
+where an asset describes any component of the network, such as a generator, line, energy storage etc.
+%For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.
 
@@
 However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
-From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission (and other possible expenditures).
+
+From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms.
 This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
-
 However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
@@
 \end{align}
-where $c_{i,t}$ denotes a general cost factor in \euro/MW. 
-For OPEX the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$ for asset $i$.
-However for CAPEX, $c_{i,t}$ is a composition of shadow prices $\mu_i$, given by the KKT variables of the capacity limit constraints.
-As we will show later the composition of $\mu_i$ must be determined for each cost term individually.
-
-% where $c_{i,t}$ is either a fixed operational price or a composition of shadowprices
-Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$ as  
+where $\costfactor$ denotes a cost factor in \euro/MW. 
+If $\cost_i$ describes the OPEX occasioned by asset $i$, the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$.
+However, as we will show later, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given at the optimum.
+%As we will show later the composition of $\mu_{i,t}$ must be determined for each type of asset individually.
+
+% where $\costfactor$ is either a fixed operational price or a composition of shadowprices
+Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
 \begin{align}
     \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
-\end{align}
-Therefore, $\allocatecost$ indicates the contribution of $\demand$ to the cost term $\cost_{i}$.
+    \label{eq:cost_allocation}
+\end{align}
+The product on the right hand side 
+\begin{align}
+    \allocatestate = \pdv{\state}{\demand} \demand
+    \label{eq:state_allocation}
+\end{align}    
+may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with.
 It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$.
-Without further specifying, we define the derivative multiplied with the nodal demand as
-\begin{align}
-    \allocatestate = \pdv{\state}{\demand} \demand
-    \label{eq:allocate_state}
-\end{align}    
-This quantity may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with.
 From the natural fact that the sum of all contributions must return the cost term, 
@@
     \cost_i = \sum_{n,t} \allocatecost
+    \label{eq:cost_payback}
 \end{align}
@@
     \state = \sum_n \allocatestate
-    \label{eq:allocate_state_constraint}
-\end{align}
+    \label{eq:state_allocation_constraint}
+\end{align}
+The last equation states that all power produced or processed by asset $i$ must be totally consumed by the network demand $\demand$.  
 Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
@@
         CAPEX Transmission & $\ell$ & $\capexflow$ & $ \capitalpriceflow \capacityflow$ & $\left(\muupperflow - \mulowerflow \right)$ & $\flow$ \\
-        CAPEX Storage & $r$ & $\capexstorage$ & $ \capitalpricestorage \capacitystorage$ & $ (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch$ & $\storage$ \\
+        CAPEX Storage & $r$ & $\capexstorage$ & $ \capitalpricestorage \capacitystorage$ & $ \muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge $ & $\storage$ \\
         \midrule
@@
 \end{align}
-where $\generation$ denotes its production at time $t$.
-It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme} for a full cost allocation of OPEX on production.
-As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by
+where $\generation$ denotes its power generation at time $t$.
+As \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme}.
+As a result we obtain
 \begin{align}
@@
 \end{align}
+which is the contribution of $\demand$ to the OPEX at generator $s$.
 The quantity  
@@
 \end{align}
-can be considered as a power allocation which indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
-It follows naturally that, in the same manner, the scheme can be applied to the OPEX on the flow $\flow$ and storage unit dispatch $\storagedispatch$.
+can be considered as the power that is produced by generator $s$ and consumed at node $n$ at time $t$. \\
+In the same manner, we can follow the scheme allocate OPEX for flow $\flow$ and storage unit dispatch $\storagedispatch$.
 As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow.
 The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
-Note, in all cases, the allocation $\allocatestate$ remains undefined which we will resolve later.
 
 For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and the capacity limit.
-All quantities for the CAPEX allocation, which we discuss in detail in the following, are summarized in the middle section of \cref{tab:cost_allocation_map}.
+All quantities for the CAPEX allocation, which we now discuss in detail in the following, are summarized in the middle section of \cref{tab:cost_allocation_map}.
 
@@
 \end{align}
-This representation connects the CAPEX with the operational state of generator $s$ and matches the needed form in \cref{eq:cost_decomposition} for using the allocation scheme.
+This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition} allows for using the cost allocation scheme.
 The resulting the CAPEX allocation is given by
@@
 If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
+When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} , we can use \cref{eq:state_allocation_constraint,eq:no_profit_capex_generation} to see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
 
@@
 \end{align}
-Again, we follow the scheme in \cref{eq:general_scheme} which  defines the CAPEX allocation as 
+From here, we follow the scheme in \cref{eq:general_scheme} which finally defines the CAPEX allocation as 
 \begin{align}
@@
 \end{align}
-can be interpreted as the flow that the demand at $\left(n,t\right)$ causes on line $\ell$.
-% According to \cref{eq:allocate_state_constraint} it has to fulfill 
+can be interpreted as the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
+% According to \cref{eq:state_allocation_constraint} it has to fulfill 
 % \begin{align}
@@
 Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
-This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_flow}
+This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:state_allocation_constraint,eq:no_profit_capex_flow}
 
@@
 
-We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus 
-\begin{align}
-    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
-\end{align}
-where $\incidencestorage$ is one if storage $s$ is at node $n$ and zero otherwise. Using \cref{eq:lagrangian_storage_dispatch}, we can reformulate to 
-\begin{align}
-    \capitalpricestorage \, \capacitystorage = \sum_t \incidencestorage \left( (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch \right) \storagedispatch 
-\end{align}
-
-% \subsection{Allocate Operational Expenditures for Generators}
-
-% Assume the cost-minimal power system includes multiple generators whose dispatch $\generation$ is optimized.
-The cost term added to $\totalcost$ combines the OPEX for all generators $s$ over the whole time span
-% \begin{align}
-%     \opexgeneration =  \sum_{s,t} \operationalpricegeneration \, \generation
-% \end{align}   
-% where $\operationalpricegeneration$ denotes the fixed operational price for generator $s$ to produce 1~MW of electricity.
-The full cost allocation for the operation of generator $s$, arises from replacing
-% \begin{subequations}
-% \label[relation]{eq:opex_allocation}
-% \begin{align}
-%     i &\Rightarrow s \\
-%     \costfactor &\Rightarrow \operationalpricegeneration \\
-%     \state &\Rightarrow \generation 
-% \end{align}
-% \end{subequations}
-% % $i$ by $s$, $\costfactor$ by $\operationalpricegeneration$, $\cost$ by $\opexgeneration$ and $\state$ by $\generation$
-% in \cref{eq:general_scheme}. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
-% \begin{align}
-%     \allocateopex[n \rightarrow s] &= 
-%    \operationalpricegeneration \,  \allocategeneration
-% \label{eq:allocate_opexGeneration_detailed}
-% \end{align}
-% where the quantity  
-% \begin{align}
-%  \allocategeneration = \pdv{\generation}{\demand} \, \demand
-%  \label{eq:allocate_peer}
-% \end{align}
-% obtained through \cref{eq:allocate_state}, can be considered as a power allocation. It indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
-% From \cref{eq:allocate_state_constraint} it follows that all the power allocations must sum up to the power produced by generator $s$, \ie
-% \begin{align}
-% \generation = \sum_n \allocategeneration
-% \label{eq:allocate_peer_constraint}
-% \end{align}
-% % 
-% % 
-% The cost allocation $\allocateopex[n \rightarrow s]$ assumes peer-to-peer tradings between producers and consumers for fixed production prices $\operationalpricegeneration$.
-% Hence, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs.
-In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$,
-% \begin{align}
-% \sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
-% \label{eq:no_profit_opex}
-% \end{align}
+In a simplified storage model, $\capacitystorage$ limits the storage dispatch $\storagedispatch$ and charging $\storagecharge$.
+Further it limits the maximal storage capacity $\storagesoc$ by a fix ratio $h_r$, denoting the maximum hours at full discharge.
+The storage $r$ dispatches power with efficiency $\efficiencydispatch$, charges power with efficiency $\efficiencycharge$ and preserves power from one time step $t$ to the next, $t+1$, with an efficiency of $\efficiencysoc$.
+In \cref{sec:zero_profit_storage_units} we formulate the mathematical details.
+As already shown in \cite{brown_decreasing_2020}, the total expenditures at $r$ are fully paid back by differences of the LMP at which the storage ``buys`` and ``sells`` power.
+Taking only the CAPEX into account the zero profit condition reduces to
+\begin{align}
+    \notag
+    \capexstorage =& \capitalpricestorage \, \capacitystorage \\
+    \notag
+    =& \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
+    &- \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
+    \label{eq:no_profit_capex_storage}
+\end{align}
+where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower capacity bound of the power dispatch of $r$ and $\mustateofcharge$ is the shadow price of the energy balance constraint.
+\Cref{eq:no_profit_capex_storage} matches the form of \cref{eq:cost_decomposition}.
+However when applying \cref{eq:cost_allocation} it follows naturally that to assume $\partial \storagecharge / \partial \demand = 0$,
+% that 
+% \begin{align}
+%     \pdv{\storagecharge}{\demand} \demand = 0 
+% \end{align}
+implying that the when a storage charges power, it does not supply any demand. Rather it stands with the demand on the consumer side, retrieving power from producing assets. 
+This leaves us with 
+\begin{align}
+     \allocatecapexstorage = \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \allocatestoragedispatch
+\end{align}
+and the power allocation 
+\begin{align}
+    \allocatestoragedispatch = \pdv{\storagedispatch}{\demand} \demand
+\end{align}
+The latter only allocates dispatched power of storage $r$. Note that this will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
+% \begin{align}
+%     \capexstorage_r + \remainingcost^E_r = \sum_n \allocatecapexstorage
+% \end{align} 
+However the fact that storages themselves ``buy`` power from other the assets, when charging, the cost allocation in \cref{eq:general_scheme} may be applied to the consumption of storages themselves, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
+This derives payments that a storage unit $r$ has to pay to asset $i$, namely $\cost_{r \rightarrow i}$.
+The sum of those payments due to $r$ will the add up to $\remainingcost^E_r$.
 
@@
 But so far, the question of how $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.
-We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:allocate_state_constraint}).
+We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:state_allocation_constraint}).
 \\
@@
 \subsection{LMP from Optimization}
-The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
-\begin{align}
-    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
+The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL).
+Alternativaly, we can the demand $\demand$ has to be supplied by the attached assets,
+\begin{align}
+    \sum_i \incidenceasset \state  &=  \demand 
+    % \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0
+     \resultsin{\lmp} \Forall{n,t}
     \label[constraint]{eq:nodal_balance_lin}
 \end{align}
-Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\
+where $\incidenceasset$ is +1 if $i$ is attached to $n$ and a positive operation $\state$ delivers power to $n$, -1 if is attached to $n$ and a positive operation retrieves power from $n$ and zero else (note that for lines this results in the negative of the conventional Incidence Matrix).
+The shadow price of the nodal balance constraint mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
+In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\
 
@@
 \label{sec:full_lagrangian}
+The Lagrangian for the investment model can be condensed to the following expression
+
 \begin{align}
 \notag
-& \lagrangian\left(\generation, \flow, \capacitygeneration, \capacityflow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
+ \lagrangian\left(\state, \capacity, \lmp, \mu_j \right) =& 
+ \;\;\; \sum_{i,t} o_i \state + \sum_{i} c_i \capacity  \\
 \notag
-& \;\;\; \sum_{n,s} \capitalpricegeneration \capacitygeneration + \sum_{n, s, t} \operationalpricegeneration \generation + \sum_{\ell} \capitalpriceflow \, \capacityflow  \\
-\notag
-&+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidencegenerator\, \generation +  \demand  \right)  \\ 
-\notag
-&+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
-\notag
-&+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacitygeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
-&+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityflow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityflow \right)     
+&+  \sum_{n,t} \lmp \left( \demand - \sum_i \incidenceasset \, \state \right) \\
+&+ \sum_j \mu_j \, h_j \left(\state, \capacity \right)
 \label{eq:full_lagrangian}
 \end{align}
-% 
-where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-The global maximum of the Lagrangian requires stationarity with respect to all variables.
-The stationarity of the generation capacity variable leads to
-\begin{align}
-\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, 
-\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
-\label{eq:capex_generation_duality}
-\end{align}
-
-the stationarity of the transmission capacity to
-\begin{align}
-\pdv{\lagrangian}{\capacityflow} = 0 \,\, \rightarrow \,\, 
-\capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
-\label{eq:capexFlow_duality}
-\end{align}
-
-and the stationarity of the generation to 
-\begin{align}
-\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
-\operationalpricegeneration =  \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
-\end{align}
-
-\begin{align}
-0 &= \pdv{\lagrangian}{\flow}  \\ 
-0 &= \sum_n \incidence \lmp  + \cycleprice \cycle \impedance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
-\end{align}
-
-\subsection{Proof Generation Capacity Payback}
+where $h_j \left(\state, \capacity \right)$ denotes all inequality constraints attached to $\state$ and $\capacity$.
+% \begin{align}
+% \notag
+% & \lagrangian\left(\generation, \flow, \capacitygeneration, \capacityflow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
+% \notag
+% & \;\;\; \sum_{n,s} \capitalpricegeneration \capacitygeneration + \sum_{n, s, t} \operationalpricegeneration \generation + \sum_{\ell} \capitalpriceflow \, \capacityflow  \\
+% \notag
+% &+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidencegenerator\, \generation +  \demand  \right)  \\ 
+% \notag
+% &+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
+% \notag
+% &+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacitygeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
+% &+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityflow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityflow \right)     
+% \label{eq:full_lagrangian}
+% \end{align}
+% 
+% where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
+% 
+In order to impose the Kirchhoff Voltage Law (KVL) for the linearized AC flow, the term 
+\begin{align}
+    \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow 
+\end{align}
+can be added to $\lagrangian$, with $\impedance$ denoting the line's impedance and $\cycle$ being 1 if $\ell$ is part of the cycle $c$ and zero otherwise.
+
+The global maximum of the Lagrangian requires stationarity with respect to all variables:
+\begin{align}
+    \pdv{\lagrangian}{\state} = \pdv{\lagrangian}{\capacity} = 0    
+\end{align}
+
+\subsection{Zero Profit Generation}
 \label{sec:zero_profit_generation}
-
-As stated in \cite{brown_decreasing_2020} the capical cost of a generator $s$ is equal to the shadowprice $\muuppergeneration$ times the upper availibility $\generationpotential$ of all time steps 
-\begin{align}
- \capitalpricegeneration  = \sum_t \muuppergeneration \, \generationpotential 
- \label{eq:capital_price_generation_sum}
-\end{align}
-\Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint} which yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ imply the complementary slackness,
+\Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint}, which yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$, imply the complementary slackness,
 \begin{align}
@@
 \end{align}
+
+The stationarity of the generation capacity variable leads to 
+\begin{align}
+\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, 
+\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+\label{eq:capex_generation_duality}
+\end{align}
+and the stationarity of the generation to 
+\begin{align}
+\pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
+\operationalpricegeneration =  \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
+\end{align}
+
 Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacitygeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to 
@@
 
-\subsection{Proof Flow Capacity Payback}
+\subsection{Zero Profit Transmission System}
 \label{sec:zero_profit_flow}
@@
 sets the respective KKT to zero.
+
+The stationarity of the transmission capacity to
+\begin{align}
+\pdv{\lagrangian}{\capacityflow} = 0 \,\, \rightarrow \,\, 
+\capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
+\label{eq:capexFlow_duality}
+\end{align}
+and the stationarity with respect to the flow to
+\begin{align}
+    0 &= \pdv{\lagrangian}{\flow}  \\ 
+    0 &= - \sum_n \incidence \lmp  + \cycleprice \cycle \impedance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
+\end{align}
+
 When multipying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
@@
 \begin{align}
-\capitalpriceflow \, \capacityflow = \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
-\end{align}
+\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
+\end{align}
+% Watch out Incidence = - ordinary Incidence !!
 It states that the congestion revenue of a line (first term right hand side) substracted by the cost for cycle constraint exactly matches its CAPEX.
 
-\subsection{CO$_2$ constraint}
-
-Imposing an additional CO$_2$ constraint limiting the total emission to K,  
-\begin{align}
-\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionprice} 
-\label[constraint]{eq:co2_constraint}
-\end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionprice$ in \euro/tonne-CO$_2$. 
-% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
-% 
-% \begin{align}
-% \capitalpricegeneration \, \capacitygeneration + \sum_{t} \operationalpricegeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionprice \right)  \, \generation \Forall{n,s} 
-% \label{eq:non_profit_generator_emission}
-% \end{align}
-% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
-As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
-\begin{align}
-\operationalpricegeneration \rightarrow \operationalpricegeneration + \emission \, \emissionprice \label[relation]{eq:shift_opex_by_emission_cost}
-\end{align}
-
-\subsection{Storage Units}
+\subsection{Zero Profit Storage Units}
+\label{sec:zero_profit_storage_units}
 
@@
 where we assume a fixed ratio between dispatch and storage capacity of $h_r$. 
-From stationarity we obtain 
-\begin{align}
-    \pdv{\lagrangian}{\capacitystorage} &= 0 \\
-    \capitalpricestorage &= \sum_t \left( \muupperstoragedispatch + \muupperstoragecharge + h_r \muupperstoragesoc  \right)
-\end{align}
-\begin{align}
+% From stationarity we obtain 
+% \begin{align}
+%     \notag
+%     \pdv{\lagrangian}{\capacitystorage} &= 0 \\
+%     \capitalpricestorage &= \sum_t \left( \muupperstoragedispatch + \muupperstoragecharge + h_r \muupperstoragesoc  \right)
+% \end{align}
+The state of charge must be consistent throughout every time step according to what is dispatched and stored, 
+\begin{align}
+    \notag
+    \storagesoc - \efficiencysoc \storageprevioussoc - \efficiencycharge \storagecharge &+ (\efficiencydispatch)^{-1} \storagedispatch = 0 \\
+    &\resultsin{\mustateofcharge} \Forall{r,t}
+\end{align}
+
+We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus 
+\begin{align}
+    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
+\end{align}
+The stationarity of the dispatched power leads us to  
+\begin{align}
+    \notag
     \pdv{\lagrangian}{\storagedispatch} &= 0 \\
-    \operationalpricestorage +& \lmp + \mulowerstoragedispatch - \muupperstoragedispatch - (\efficiencydispatch )^{-1} \mustateofcharge = 0  
-    \label{eq:lagrangian_storage_dispatch}
-\end{align}
-\begin{align}
-    \pdv{\lagrangian}{\storagecharge} &= 0 \\
-    - \lmp +& \mulowerstoragecharge - \muupperstoragecharge + \efficiencycharge \mustateofcharge = 0 
-    \label{eq:lagrangian_storage_charge}
+    \operationalpricestorage -  \lmp \, \incidencestorage - \mulowerstoragedispatch + \muupperstoragedispatch + (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
+    \label{eq:stationarity_storagedispatch}
+\end{align}
+which we can use to define the revenue which compensates the CAPEX at $r$, 
+\begin{align}
+    \notag
+    \capitalpricestorage \, \capacitystorage = \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
+    - \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
+\end{align}
+
+% With this constraint we can derive the following stationarities:
+% \begin{align}
+%     \notag
+%     \pdv{\lagrangian}{\storagedispatch} &= 0 \\
+%     \operationalpricestorage + \lmp + \mulowerstoragedispatch - \muupperstoragedispatch - (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
+%     \label{eq:stationarity_storagedispatch}
+% \end{align}
+% \begin{align}
+%     \notag
+%     \pdv{\lagrangian}{\storagecharge} &= 0 \\
+%     - \lmp + \mulowerstoragecharge - \muupperstoragecharge + \efficiencycharge \mustateofcharge &= 0 
+%     \label{eq:stationarity_storagecharge}
+% \end{align}
+% \begin{align}
+%     \notag
+%     \pdv{\lagrangian}{\storagesoc} &= 0 \\
+%     \mulowerstoragesoc - \muupperstoragesoc - \mustateofcharge + \efficiencysoc \munextstateofcharge &= 0 
+%     \label{eq:stationarity_storagesoc}
+% \end{align}
+
+\subsection{CO$_2$ constraint}
+
+Imposing an additional CO$_2$ constraint limiting the total emission to K,  
+\begin{align}
+\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionprice} 
+\label[constraint]{eq:co2_constraint}
+\end{align}
+with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionprice$ in \euro/tonne-CO$_2$. 
+% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
+% 
+% \begin{align}
+% \capitalpricegeneration \, \capacitygeneration + \sum_{t} \operationalpricegeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionprice \right)  \, \generation \Forall{n,s} 
+% \label{eq:non_profit_generator_emission}
+% \end{align}
+% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
+As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
+\begin{align}
+\operationalpricegeneration \rightarrow \operationalpricegeneration + \emission \, \emissionprice \label[relation]{eq:shift_opex_by_emission_cost}
 \end{align}
@@
 \begin{align}
-\capacitygeneration \ge \capacitygenerationlower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \label{eq:capacityexpansionminimum}\\
-\capacitygeneration \le \capacitygenerationupper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
+    \capacitygenerationlower - \capacitygeneration \le 0 \resultsin{\mulowergenerationnom} \hpad \forall{s \in S} \label{eq:capacityexpansionminimum}\\
+    \capacitygeneration - \capacitygenerationupper \le 0 \resultsin{\muuppergenerationnom} \hpad \forall{s \in S}
 \label{eq:capacityexpansionmaximum}
@@
 \begin{align}
-    \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{n,s \in S} \\
-    \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{n,s \in S}
+    \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{s \in S} \\
+    \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{s \in S}
 \end{align}
@@
 \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
-\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s \in S}
+\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{s \in S}
 \label{eq:capex_generation_duality_bf}
@@
 \begin{align}
- \capitalpricegeneration \, \capacitygeneration - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
+ \capexgeneration_s - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
  \label{eq:capex_generation_duality_bf2}
@@
 \end{align}
-As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \le 0$ the latter represents a net positive cost term.
+As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \ge 0$, the latter can be a net positive or net negative cost term.
 
@@
 \begin{align}
-    \capitalpricegeneration \capacitygeneration + \sum_t \operationalpricegeneration \generation  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
-\end{align}
-We see that the revenue of generator $s$ will not fully pay back its full CAPEX and OPEX.
+    \capexgeneration_s + \opexgeneration_s  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
+\end{align}
+We see that the revenue of generator $s$ will not exactly pay back its full CAPEX and OPEX.
 The exogenous constraint shifts the zero-profit equations such that some of the expenditures for $s \in S$ cannot directly be allocated to consumers.
-The collected payments of all consumers will not cover the costs induced by \cref{eq:capacityexpansionminimum,eq:capacityexpansionmaximum},
-
-\begin{align}
-    \capitalpricegeneration \capacitygeneration - \remainingcost_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
+In order to take this effect into account for the cost allocation, we update the shadow price $\muuppergeneration$ in \cref{tab:cost_allocation_map}, to
+
+\begin{align}
+    \muuppergeneration \rightarrow \dfrac{\capitalpricegeneration}{\capitalpricegeneration - \mulowergenerationnom + \muuppergenerationnom}\, \muuppergeneration
+\end{align}
+By doing so we ensure that the allocated CAPEX sum up to the actual CAPEX at generator $s$, thus 
+\begin{align}
+    \capexgeneration_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
 \end{align}

### 2020-08-13 code/allocate_network.py

@@ -58,6 +58,14 @@ o = o * snapshot_weightings(n) - ep.reindex_like(o, fill_value=0)
 A_opex = A * o
 
 
+"""
+WATCH OUT! In the paper the constraints always have a '<=' sign which lead
+to positive shadowprices.
+This is not the case for the pypsa.linopf implementation -> we have
+to adjust the mu's in some cases in order to take into account that they're
+not positive.
+"""
+
 # capex one port
 c = 'Generator'
 mu_gen = by_bus_carrier(n.pnl(c).mu_upper, c, n)

### 2020-08-13 code/check_sus_shadowprices.py

@@ -1,0 +1,45 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Aug 12 11:06:12 2020
+
+@author: fabian
+"""
+
+import pypsa
+
+
+n = pypsa.Network('/home/fabian/papers/costallocation/resources/de10gf.nc')
+
+close = lambda df1, df2: ((df1/df2).fillna(1).round(2) == 1).all()
+
+# Generators
+
+g = n.generators
+gt = n.generators_t
+
+assert close(g.capital_cost, (gt.mu_upper.mul(gt.p_max_pu, fill_value=1)).sum()
+                              + g.mu_upper_p_nom + g.mu_lower_p_nom)
+
+g.eval('p_nom_opt * capital_cost')
+
+
+# Lines/Links
+
+
+
+
+# Storage Units
+
+s = n.storage_units
+st = n.storage_units_t
+
+
+# This must be equal
+assert close(s.capital_cost, (st.mu_upper_p_dispatch + st.mu_upper_p_store +
+                              st.mu_upper_state_of_charge * s.max_hours).sum())
+
+
+assert close((st.mu_lower_p_dispatch + st.mu_upper_p_dispatch + st.mu_state_of_charge / s.efficiency_dispatch)['DE0 0 battery'], n.buses_t.marginal_price['DE0 0'])
+
+

### 2020-08-14 paper/flow_based.tex
allocation fix storage shadowprice
correct shadow price correction

@@
 \newcommand{\costfactor}{\gamma_{i,t}}
+\newcommand{\capacityupper}{\bar{S}}
+\newcommand{\capacitylower}{\ubar{S}}
+\newcommand{\muuppernom}{\bar{\mu}^\text{nom}_{i}}
+\newcommand{\mulowernom}{\ubar{\mu}^\text{nom}_{i}}
 
@@
 Each asset $i = \{s, \ell, r\}$ adds an term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
-The operational price for an asset $i$ is given by $o_i$, the capital price for one unit capacity expansion by $c_i$.
-For example the OPEX occasioned by generator $s$ is given by
+
+\subsection{Allocating OPEX}
+
+Let the operational price for an asset $i$ be given by $o_i$. Then, for example the OPEX occasioned by generator $s$ is given by 
 \begin{align}
@@
 
+\subsection{CAPEX Allocation}
 For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and the capacity limit.
-All quantities for the CAPEX allocation, which we now discuss in detail in the following, are summarized in the middle section of \cref{tab:cost_allocation_map}.
-
-\subsubsection*{Allocate CAPEX on Generators}
+For all assets, let the capital price for one unit capacity expansion be denoted by $c_i$.
+All quantities for the CAPEX allocation, which we now discuss in detail, are summarized in the middle section of \cref{tab:cost_allocation_map}.
+
+\subsubsection{Generators}
 
@@
 
-\subsubsection*{Allocate CAPEX on Transmission Lines}
+\subsubsection{Transmission Lines}
 
@@
 
-\subsubsection*{Allocate CAPEX on Storages}
+\subsubsection{Storages}
 
@@
 \end{align}
-where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower capacity bound of the power dispatch of $r$ and $\mustateofcharge$ is the shadow price of the energy balance constraint.
-\Cref{eq:no_profit_capex_storage} matches the form of \cref{eq:cost_decomposition}.
-However when applying \cref{eq:cost_allocation} it follows naturally that to assume $\partial \storagecharge / \partial \demand = 0$,
-% that 
-% \begin{align}
-%     \pdv{\storagecharge}{\demand} \demand = 0 
-% \end{align}
-implying that the when a storage charges power, it does not supply any demand. Rather it stands with the demand on the consumer side, retrieving power from producing assets. 
+where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint.
+When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand = 0$, implying that the when a storage charges power, it does not supply any demand.
+Rather it stands with the demand on the consumer side, retrieving power from producing assets.
 This leaves us with 
@@
 % \end{align} 
-However the fact that storages themselves ``buy`` power from other the assets, when charging, the cost allocation in \cref{eq:general_scheme} may be applied to the consumption of storages themselves, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
-This derives payments that a storage unit $r$ has to pay to asset $i$, namely $\cost_{r \rightarrow i}$.
-The sum of those payments due to $r$ will the add up to $\remainingcost^E_r$.
-
-\subsubsection*{Allocate Emission Expenditures}
+It is crucial to note that like this, storage units perform a redistribution of money, and therefore distort the cost allocation.
+So, certain share of what is allocated to the CAPEX of a storage is in another time step spent by the storage in order to buy power from other assets.
+This effect scales with the amount of installed capacity.
+
+It is possible to incorporate this redistribution effect into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
+Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$.
+The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.
+
+% In a final step we can still allocate all cost to the demand $\demand$.
+However, this would require knowledge about when the power $\storagedispatch$ was initially charged: Given the amount of power $A_{r, t_{in} \rightarrow t_{out}}$ that is charged by storage $r$ at time step $t_{in}$ and discharged at $t_{out}$ ...
+
+\subsection{Allocate Emission Expenditures}
 
@@
 associated with generator $s$.
+
+\subsection{Capacity Expansion Limit}
+
+In more realistic setups, generators, lines or other assets can only be built up to a certain limit.
+This might caused by land use restrictions or social acceptance problems.
+%In the following give a general solution how the cost allocation may handle this.
+However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit,
+\begin{align}
+    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \hpad \Forall{i \in I}
+\label{eq:capacityexpansionmaximum}
+\end{align}
+the zero profit condition for asset $i$ alters and the revenues don't match the expenditures anymore.
+More precisely, the revenue from the shadow prices given in \cref{tab:cost_allocation_map} exceed the CAPEX $\cost_i$ by $\remainingcost_i$, thus
+\begin{align}
+ \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
+ \label{eq:capex_generation_duality_bf2}
+\end{align}
+where the difference is given by 
+\begin{align}
+    \remainingcost_i = - \muuppernom \capacity
+    \Forall{i}
+\end{align}
+In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to
+
+\begin{align}
+    \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppergenerationnom}\, \costfactor
+\end{align}
+This ensures that the allocated CAPEX sum up to the actual CAPEX at asset $i$ (and \cref{eq:cost_payback} holds).
+
+\subsection{Brownfield Constraints}
+
+In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. Mathematically this translates to 
+\begin{align}
+    \capacitylower - \capacity  \le 0 \resultsin{\mulowernom} \hpad \forall{i \in I}
+\label{eq:capacityexpansionminimum}
+\end{align}
+Again, such a setup alters the zero profit condition of asset $i$ as soon as the constraint becomes binding. 
+In that case, asset $i$ does not collect enough revenue in order to match the CAPEX by a difference is given by 
+\begin{align}
+    \remainingcost_i = \mulowernom \capacity
+    \Forall{i}
+\end{align}
+% In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to
+
+% \begin{align}
+%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppergenerationnom}\, \costfactor
+% \end{align}
+% This ensures that the allocated CAPEX sum up to the actual CAPEX at asset $i$ (and \cref{eq:cost_payback} holds).
 
@@
 
-In the following, we illustrate the cost allocation and explain its functionality by taking the example of a two bus system with one optimized time step, shown in \cref{fig:example_network}. \\
-% 
-% 
 \begin{figure*}[t]
@@
     \end{figure*}
+    % 
+In the following, we illustrate the cost allocation and explain its functionality by taking the example of a two bus system with one optimized time step, shown in \cref{fig:example_network}. \\
+% 
+% 
 % 

### 2020-08-14 code/allocate_network.py

@@ -9,23 +9,21 @@ Created on Fri Jul 10 20:56:10 2020
 import netallocation as ntl
 import pypsa
 import xarray as xr
+import numpy as np
 import pandas as pd
 
 from netallocation.utils import (reindex_by_bus_carrier as by_bus_carrier,
                                  get_as_dense_by_bus_carrier)
-from netallocation.breakdown import (by_carriers as breakdown_carriers,
-                                     expand_by_sink_type,
+from netallocation.breakdown import (expand_by_sink_type,
                                      expand_by_source_type)
 from netallocation.convert import peer_to_peer, virtual_patterns
-from netallocation.cost import (nodal_co2_price, snapshot_weightings,
-                                allocate_one_port_operational_cost,
-                                allocate_branch_investment_cost)
-from config import source_dims, sink_dims
+from netallocation.cost import (nodal_co2_price, snapshot_weightings)
+from config import source_dims
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('allocate_network', nname='test-de50gf',
+        snakemake = mock_snakemake('allocate_network', nname='acdc',
                                    method='ptpf', power='net')
 
 n = pypsa.Network(snakemake.input.network)
@@ -66,16 +64,20 @@ A_opex = A * o
 # not positive.
 # =============================================================================
 
+
 # capex one port
 c = 'Generator'
-mu_gen = by_bus_carrier(n.pnl(c).mu_upper, c, n)
+# calculate corrections for capacity restrictions
+correction_upper = n.df(c).capital_cost / (n.df(c).capital_cost + n.df(c).mu_upper_p_nom)
+correction_lower = n.df(c).mu_lower_p_nom * n.df(c).p_nom_opt / n.pnl(c).p.sum()
+correction_lower = correction_lower.replace(np.inf, 0)
+mu_gen = by_bus_carrier(n.pnl(c).mu_upper * correction_upper + correction_lower, c, n)
+
 
 c = 'StorageUnit'
 if not n.df(c).empty:
     mu_sus = (n.pnl(c).mu_state_of_charge / n.df(c).efficiency_dispatch
               + n.pnl(c).mu_upper_p_dispatch + n.pnl(c).mu_lower_p_dispatch)
[…]

### 2020-08-14 code/check_shadowprices.py

@@ -9,6 +9,12 @@ Created on Wed Aug 12 11:06:12 2020
 import pypsa
 
 
+if __name__ == "__main__":
+    if 'snakemake' not in globals():
+        from _helpers import mock_snakemake
+        snakemake = mock_snakemake('check_shadowprices', nname='acdc')
+
+
 n = pypsa.Network(snakemake.input[0])
 
 close = lambda df1, df2: ((df1/df2).fillna(1).round(2) == 1).all()
@@ -21,7 +27,16 @@ gt = n.generators_t
 assert close(g.capital_cost, (gt.mu_upper.mul(gt.p_max_pu, fill_value=1)).sum()
                               + g.mu_upper_p_nom + g.mu_lower_p_nom)
 
-g.eval('p_nom_opt * capital_cost')
+investment = g.eval('p_nom_opt * capital_cost')
+revenue = (gt.mu_upper * gt.p).sum()
+
+correction_factor_upper = g.capital_cost / (g.capital_cost + g.mu_upper_p_nom)
+correction_term_lower = g.mu_lower_p_nom * g.p_nom_opt / gt.p.sum()
+
+revenue_corrected = ((gt.mu_upper * correction_factor_upper
+                      + correction_term_lower) * gt.p).sum()
+
+assert close(investment.where(gt.p.sum()>0, 0), revenue_corrected)
 
 
 # Lines/Links
@@ -39,7 +54,7 @@ st = n.storage_units_t
 assert close(s.capital_cost, (st.mu_upper_p_dispatch + st.mu_upper_p_store +
                               st.mu_upper_state_of_charge * s.max_hours).sum())
 
-
-assert close((st.mu_lower_p_dispatch + st.mu_upper_p_dispatch + st.mu_state_of_charge / s.efficiency_dispatch)['DE0 0 battery'], n.buses_t.marginal_price['DE0 0'])
+store, bus = s.index[0], s.iloc[0].bus
+assert close((s.marginal_cost + st.mu_lower_p_dispatch + st.mu_upper_p_dispatch + st.mu_state_of_charge / s.efficiency_dispatch)[store], n.buses_t.marginal_price[bus])
 
 

### 2020-08-17 code/examples.py
adjust examples plot

@@ -14,6 +14,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 from netallocation.plot import injection_plot_kwargs
+from helpers import adjust_shadowprice
 from config import to_static_symbol, color, to_total_symbol
 # plt.rc('tex', preamble=r'\Large')
 plt.rc('text', usetex=True)
@@ -21,7 +22,7 @@ plt.rc('font', family='serif')
 
 path = '../figures/example'
 
-net = True
+net = False
 power = 'net' if net else 'gross'
 method = 'ebe'
 
@@ -44,6 +45,7 @@ n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
 n.madd('Line', ['1'], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
        capital_cost=c_l)
 n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
+n.generators_t.mu_upper = adjust_shadowprice(n.generators_t.mu_upper, n, 'Generator')
 
 sn = 'now'
 

### 2020-08-17 paper/flow_based.tex
adjust examples plot
adjust shadowprice for branches properly

@@
 Again, such a setup alters the zero profit condition of asset $i$ as soon as the constraint becomes binding. 
-In that case, asset $i$ does not collect enough revenue in order to match the CAPEX by a difference is given by 
+In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference is given by 
 \begin{align}
@@
 \end{align}
+Again, we can balance this effect out by updating the cost factor to 
+\begin{align}
+    \costfactor \rightarrow \costfactor + 
+\end{align}
+
 % In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to
@@
 \section{Application Case}
+
+We apply the cost allocation on a German power system model
 

### 2020-08-17 code/check_shadowprices.py

@@ -19,7 +19,12 @@ n = pypsa.Network(snakemake.input[0])
 
 tol = (n.objective + n.objective_constant)*1e-6
 
-
+if 'lv_limit' in n.global_constraints.index:
+    for c in n.branch_components:
+        if n.df(c).empty: continue
+        mu_upper = (n.global_constraints.at['lv_limit', 'mu'] * n.df(c).length)
+        nom = pypsa.descriptors.nominal_attrs[c]
+        n.df(c)['mu_upper_'+nom] += mu_upper
 
 close = lambda df1, df2: ((df1 - df2).abs() < tol ).all()
 
@@ -51,8 +56,6 @@ revenue = ((lt.mu_upper + lt.mu_lower) * lt.p0).sum()
 
 gamma = lt.mu_upper + lt.mu_lower
 gamma_adj = adjust_shadowprice(gamma, 'Line', n)
-if 'lv_limit' in n.global_constraints.index:
-    gamma_adj += - n.global_constraints.at['lv_limit', 'mu'] * l.length
 revenue = (gamma_adj * lt.p0).sum()
 assert close(investment, revenue)
 
@@ -66,8 +69,6 @@ revenue = ((lt.mu_upper + lt.mu_lower) * lt.p0).sum()
 lv_mu = - n.global_constraints.at['lv_limit', 'mu']
 gamma = lt.mu_upper + lt.mu_lower
 gamma_adj = adjust_shadowprice(gamma, 'Link', n)
-if 'lv_limit' in n.global_constraints.index:
-    gamma_adj += - n.global_constraints.at['lv_limit', 'mu'] * l.length
 revenue = (gamma_adj * lt.p0).sum()
 assert close(investment, revenue)
 

### 2020-08-17 code/helpers.py

@@ -32,7 +32,7 @@ def adjust_shadowprice(gamma, c, n, s=None):
     lower = (n.df(c)['mu_lower_'+nom] * n.df(c)[nom +'_opt'] / n.pnl(c)[p].sum())
     lower = lower.replace(np.inf, 0)
 
-    return gamma * upper + lower
+    return (gamma + lower) * upper
 
 
 

### 2020-08-17 paper/flow_based.tex
make price adjustment flexible

@@
 \end{align}
-the zero profit condition for asset $i$ alters and the revenues don't match the expenditures anymore.
-More precisely, the revenue from the shadow prices given in \cref{tab:cost_allocation_map} exceed the CAPEX $\cost_i$ by $\remainingcost_i$, thus
-\begin{align}
- \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
- \label{eq:capex_generation_duality_bf2}
-\end{align}
-where the difference is given by 
+the zero profit condition for asset $i$ alters.
+Then the revenues don't match the expenditures as soon as the constraint becomes binding.
+More precisely, the revenue from the shadow prices given in \cref{tab:cost_allocation_map} exceed the CAPEX $\cost_i$ by $\remainingcost_i$,  where  $\remainingcost_i$ is given by
 \begin{align}
@@
 \end{align}
-In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to
-
-\begin{align}
-    \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppergenerationnom}\, \costfactor
-\end{align}
-This ensures that the allocated CAPEX sum up to the actual CAPEX at asset $i$ (and \cref{eq:cost_payback} holds).
+% \begin{align}
+%  \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
+%  \label{eq:capex_generation_duality_bf2}
+% \end{align}
+For the cost allocation, these means the prices in \cref{tab:cost_allocation_map} have to be adjusted, as otherwise consumers would pay too much. As we show in ... updating the prices by a fix ratio 
+\begin{align}
+    \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppernom}\, \costfactor
+    \label{eq:adjustment_lowernom}
+\end{align}
+will ensure that the allocated CAPEX sum up to the actual CAPEX at asset $i$, meaning that \cref{eq:cost_payback} holds.
+Note that $\muuppernom$ is a positive number, so the adjustment cannot increase the paid price.
 
@@
 \end{align}
-Again, such a setup alters the zero profit condition of asset $i$ as soon as the constraint becomes binding. 
+Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding. 
 In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference is given by 
@@
 \end{align}
+In constrast to 
 Again, we can balance this effect out by updating the cost factor to 
 \begin{align}
-    \costfactor \rightarrow \costfactor + 
-\end{align}
+    \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
+\end{align}
+As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.
 

### 2020-08-17 code/allocate_network.py

@@ -30,7 +30,8 @@ if __name__ == "__main__":
 
 n = pypsa.Network(snakemake.input.network)
 
-method = snakemake.wildcards.method
+method = snakemake.wildcards.method.split('-')[0]
+adjust_mu = '-adj' in snakemake.wildcards.method
 aggregated = snakemake.wildcards.power == 'net'
 if 'co2_limit' in n.global_constraints.index:
     co2_price = 0
@@ -76,7 +77,9 @@ A_opex = A * o
 # capex one port
 c = 'Generator'
 # calculate corrections for capacity restrictions
-mu_gen = by_bus_carrier(adjust_shadowprice(n.pnl(c).mu_upper, c, n), c, n)
+mu_gen = n.pnl(c).mu_upper
+mu_gen = adjust_shadowprice(mu_gen, c, n) if adjust_mu else mu_gen
+mu_gen = by_bus_carrier(mu_gen, c, n)
 
 
 c = 'StorageUnit'
@@ -84,7 +87,8 @@ if not n.df(c).empty:
     # calculate corrections for capacity restrictions
     mu_sus = (n.pnl(c).mu_state_of_charge / n.df(c).efficiency_dispatch
               + n.pnl(c).mu_upper_p_dispatch + n.pnl(c).mu_lower_p_dispatch)
-    mu_sus = by_bus_carrier(adjust_shadowprice(mu_sus, c, n), c, n)
+    mu_sus = adjust_shadowprice(mu_sus, c, n) if adjust_mu else mu_sus
+    mu_sus = by_bus_carrier(mu_sus, c, n)
     mu = xr.concat([mu_gen, mu_sus], dim='carrier').rename(source_dims).fillna(0)
 else:
     mu = mu_gen.rename(source_dims).fillna(0)
@@ -94,8 +98,11 @@ A_capex = mu * A
 # capex branch
 names = ['component', 'branch_i']
 comps = set(A_f.component.data)
-mu = pd.concat({c: n.pnl(c).mu_upper + n.pnl(c).mu_lower for c in comps},
-               axis=1, names=names)
+mu = []
+for c in comps:
+    mu_b = n.pnl(c).mu_upper + n.pnl(c).mu_lower
+    mu.append(adjust_shadowprice(mu_b, c, n) if adjust_mu else mu_b)
+mu = pd.concat(mu, axis=1, names=names, keys=comps)
 mu = xr.DataArray(mu, dims=['snapshot', 'branch'])
 A_capex_branch = (A_f * mu).rename(bus='sink')
 

### 2020-08-17 code/check_shadowprices.py

@@ -14,88 +14,88 @@ if __name__ == "__main__":
         from _helpers import mock_snakemake
         snakemake = mock_snakemake('check_shadowprices', nname='de10bf')
 
+if not 'test' in snakemake.input[0]:
+    n = pypsa.Network(snakemake.input[0])
 
-n = pypsa.Network(snakemake.input[0])
+    tol = (n.objective + n.objective_constant)*1e-6
 
-tol = (n.objective + n.objective_constant)*1e-6
+    if 'lv_limit' in n.global_constraints.index:
+        for c in n.branch_components:
+            if n.df(c).empty: continue
+            mu_upper = (n.global_constraints.at['lv_limit', 'mu'] * n.df(c).length)
+            nom = pypsa.descriptors.nominal_attrs[c]
+            n.df(c)['mu_upper_'+nom] += mu_upper
 
-if 'lv_limit' in n.global_constraints.index:
-    for c in n.branch_components:
-        if n.df(c).empty: continue
-        mu_upper = (n.global_constraints.at['lv_limit', 'mu'] * n.df(c).length)
-        nom = pypsa.descriptors.nominal_attrs[c]
-        n.df(c)['mu_upper_'+nom] += mu_upper
+    close = lambda df1, df2: ((df1 - df2).abs() < tol ).all()
 
-close = lambda df1, df2: ((df1 - df2).abs() < tol ).all()
+    # Generators
 
-# Generators
+    g = n.generators
+    gt = n.generators_t
 
-g = n.generators
-gt = n.generators_t
+    assert close(g.capital_cost, (gt.mu_upper.mul(gt.p_max_pu, fill_value=1)).sum()
+                                  + g.mu_upper_p_nom + g.mu_lower_p_nom)
 
-assert close(g.capital_cost, (gt.mu_upper.mul(gt.p_max_pu, fill_value=1)).sum()
-                              + g.mu_upper_p_nom + g.mu_lower_p_nom)
+    investment = g.eval('p_nom_opt * capital_cost')
+    revenue = (gt.mu_upper * gt.p).sum()
 
-investment = g.eval('p_nom_opt * capital_cost')
-revenue = (gt.mu_upper * gt.p).sum()
-
-gamma = adjust_shadowprice(gt.mu_upper, 'Generator', n)
-revenue = (gamma * gt.p).sum()
[…]

### 2020-08-18 paper/flow_based.tex
reintegrate sparsity into capex allocation

@@
 \usepackage[colorlinks]{hyperref}
-\usepackage[backend=biber]{biblatex}
+\usepackage[backend=biber, bibencoding=utf8]{biblatex}
 \usepackage{eurosym}
@@
 
-% allocation quantities
-\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
-\newcommand{\allocatestoragedispatch}[1][r, n]{A_{#1,t}}
-\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
-\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
-\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
-\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
-\newcommand{\allocatecost}[1][n \rightarrow i]{\mathcal{C}_{#1, t}}
-\newcommand{\allocatecapexgeneration}[1][n \rightarrow s]{\mathcal{C}^{G}_{#1,t}}
-\newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\mathcal{C}^{F}_{#1,t}}
-\newcommand{\allocatecapexstorage}[1][n \rightarrow r]{\mathcal{C}^{E}_{#1,t}}
-\newcommand{\allocateopex}[1][n \rightarrow s]{\mathcal{O}_{#1,t}}
-\newcommand{\allocateemissioncost}[1][n \rightarrow s]{\mathcal{E}_{#1,t}}
-
+% Cost Terms
 \newcommand{\emission}{e_{s}}
@@
 \newcommand{\remainingcost}{\mathcal{R}}
+\newcommand{\scarcitycost}{\remainingcost^\text{scarcity}}
+\newcommand{\subsidycost}{\remainingcost^\text{subsidy}}
+
+% allocation quantities
+\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
+\newcommand{\allocatestoragedispatch}[1][r, n]{A_{#1,t}}
+\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
+\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
+\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
+\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
+
+\newcommand{\allocatecost}[1][n \rightarrow i]{\cost_{#1, t}}
+\newcommand{\allocatecapexgeneration}[1][n \rightarrow s]{\capexgeneration_{#1,t}}
+\newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\capexflow_{#1,t}}
+\newcommand{\allocatecapexstorage}[1][n \rightarrow r]{\capexstorage_{#1,t}}
+\newcommand{\allocateopex}[1][n \rightarrow s]{\opex_{#1,t}}
+\newcommand{\allocateemissioncost}[1][n \rightarrow s]{\emissioncost_{#1,t}}
+\newcommand{\allocatescarcitycost}[1][n \rightarrow i]{\scarcitycost_{#1,t}}
+\newcommand{\allocatesubsidycost}[1][n \rightarrow i]{\subsidycost_{#1,t}}
 
@@
 \totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
-\label{eq:total_cost_reduced}
+\label{eq:total_revenue}
 \end{align}
@@
 % where $\costfactor$ is either a fixed operational price or a composition of shadowprices
-Following the implications of \cref{eq:lmp,eq:total_cost_reduced},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
+Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
 \begin{align}
@@
 The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
+\\
+
+The scheme works for all other cost attached to the operational state of an asset $i$.
+Given for example a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, the cost term for emission adds up to
+\begin{align}
+ \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
+\end{align}
+where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$.
+The allocated payment for consumers at bus $n$ at time $t$ assigned to generator $s$ is then given by 
+\begin{align}
+ \allocateemissioncost = \emissionprice \, \emission \, \allocategeneration
+\end{align}
 
@@
 
-\subsection{Allocate Emission Expenditures}
-
-Given a fix price for emissions in CO$_2$ equivalents $\emissionprice$, the cost term for emission adds up to 
-\begin{align}
- \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
-\end{align}
-where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
-We use the same set of equation as given through \cref{eq:general_scheme,eq:opex_allocation} but update the cost factor to $\costfactor \Rightarrow \emissionprice \, \emission$.
-For consumers at bus $n$ at time $t$ this leads to the emission cost of
-\begin{align}
- \allocateemissioncost = \emissionprice \, \emission \, \allocategeneration
-\end{align}
-associated with generator $s$.
-
-\subsection{Capacity Expansion Limit}
+\subsection{Design Constraints}
+
+Power system modelling does rarely follow a pure Greenfield approach with unlimited capacity expansion.
+Rather, today's models are setting various constraints defining socio-political or  technical requirements.
+As mentioned before this will alter the equality of total cost and total revenue, \ie leads to $\remainingcost \ne 0$ in \cref{eq:total_revenue}.
+More precisely, each constraint $h_j$ (other then the nodal balance constraint) of the form
+\begin{align}
+    h_j \left(\state, \capacity \right) - K < 0
+\end{align}
+where $K$ is any non-zero constant, will result in a cost term contributing to $\remainingcost$ and in some cases alter \cref{eq:cost_decomposition} to 
+\begin{align}
+    \cost_i - \remainingcost_i = \sum_t \costfactor \, \state
+\end{align}
+In the following we highlight two often used classes of constraints and show how to incorporate them into the cost allocation.
+
+\subsubsection{Capacity Expansion Limit}
 
 In more realistic setups, generators, lines or other assets can only be built up to a certain limit.
-This might caused by land use restrictions or social acceptance problems.
+This might be due to land use restrictions or social acceptance problems.
 %In the following give a general solution how the cost allocation may handle this.
-However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit,
-\begin{align}
-    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \hpad \Forall{i \in I}
-\label{eq:capacityexpansionmaximum}
-\end{align}
-the zero profit condition for asset $i$ alters.
-Then the revenues don't match the expenditures as soon as the constraint becomes binding.
-More precisely, the revenue from the shadow prices given in \cref{tab:cost_allocation_map} exceed the CAPEX $\cost_i$ by $\remainingcost_i$,  where  $\remainingcost_i$ is given by
-\begin{align}
-    \remainingcost_i = - \muuppernom \capacity
-    \Forall{i}
+However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit $\capacityupper$, in the form of 
+\begin{align}
+    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \Forall{i \in I}
+\label{eq:capacityexpansionmaximum},
+\end{align}
+the zero profit condition alters as soon as the constraint becomes binding.
+Then, the revenue of asset $i$ exceeds its total expenditures (OPEX + CAPEX).
+More precisely, the allocated CAPEX in \cref{tab:cost_allocation_map} will surpass the actual CAPEX of asset $i$ by the cost it has to pay for the scarcity, given by the absolute value of
+\begin{align}
+    \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
 \end{align}
@@
 % \end{align}
-For the cost allocation, these means the prices in \cref{tab:cost_allocation_map} have to be adjusted, as otherwise consumers would pay too much. As we show in ... updating the prices by a fix ratio 
-\begin{align}
-    \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppernom}\, \costfactor
-    \label{eq:adjustment_lowernom}
-\end{align}
-will ensure that the allocated CAPEX sum up to the actual CAPEX at asset $i$, meaning that \cref{eq:cost_payback} holds.
-Note that $\muuppernom$ is a positive number, so the adjustment cannot increase the paid price.
-
-\subsection{Brownfield Constraints}
+
+% In order to ensure that the allocated CAPEX sum up to the actual CAPEX at asset $i$, meaning that \cref{eq:cost_payback} holds, we adjust the CAPEX allocations to 
+% \begin{align}
+%     \allocatecost = \left(\dfrac{c_i}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
+% \end{align}
+% with the cost factors $\costfactor$ for CAPEX given in \cref{tab:cost_allocation_map}. 
+% The costs which consumers at $n$ have to pay for the scarcity impacting asset $i$ are given   
+% \begin{align}
+%     \allocatescarcitycost = \left(\dfrac{\muuppernom}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
+% \end{align}
+
+The costs which consumers at $n$ have to pay for the scarcity impacting asset $i$ are given by 
+\begin{align}
+    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost^I \Forall{i \in I}
+\end{align}
+where $\allocatecost^I$ denotes the CAPEX allocation presented above.
+
+% have to be adjusted, as otherwise consumers would pay too much. As we show in ... updating the prices by a fix ratio 
+% \begin{align}
+%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppernom}\, \costfactor
+%     \label{eq:adjustment_lowernom}
+% \end{align}
+% will . Note that $\muuppernom$ is a positive number, so the adjustment cannot increase the paid price.
+
+\subsubsection{Brownfield Constraints}
 
@@
 Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding. 
-In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference is given by 
-\begin{align}
-    \remainingcost_i = \mulowernom \capacity
+In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference, given by 
+\begin{align}
+    \subsidycost_i = \mulowernom \capacity
     \Forall{i}
 \end{align}
-In constrast to 
-Again, we can balance this effect out by updating the cost factor to 
-\begin{align}
-    \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
-\end{align}
-As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.
+has to be subsidized by governments or communities.
+It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost^I = \subsidycost_i $.
+
+% Again, we can balance this effect out by updating the cost factor to 
+% \begin{align}
+%     \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
+% \end{align}
+% In constrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
+% As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.
 
@@
 
-We apply the cost allocation on a German power system model
+In order to showcase the cost allocation at a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
+The model builds on the PyPSA-EUR workflow \cite{horsch_pypsa-eur_2020} with technigal details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
+We follow a browfield approach where transmission lines start from today's capacity values originally retreived from the ENTSO-E Transmission System Map \cite{entsoe-e_entso-e_nodate}.
+Pre-installed generation capacities for wind and solar are distributed proportionally to the average power potential at each site excluding those with an average capacity factor of 10\%.
+Further, wind and solar capacity expansion are limited by land use restriction.
+These consider agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{corine2012,natura2000}.
+Pumped Hydro Storages (PHS) are fixed to today's capacities with no more expansion allowed.
+Additionally, unlimited batteries and H$_{2}$-storages are allowed at each node.
+We impose a effective carbon price of 120 \euro per tonne-CO$_{2}$.
+
+\Cref{fig:network} shows the estimate of existent capacities on the left hand side and the capacity expansions on the right. 
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{de50bf}
+    \label{fig:network}
+\end{figure*}
 

### 2020-08-18 code/config.py

@@ -16,28 +16,38 @@ source_dims = dict(bus='source', carrier='source_carrier')
 sink_dims = dict(bus='sink', carrier='sink_carrier')
 
 to_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n,t}$',
-                      branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',
-                      one_port_operational_cost = '$\mathcal{O}_{n,t}$',
-                      co2_cost = '$\mathcal{E}_{n,t}$',
-                      nodal_demand_cost = '$\lambda_{n,t} \, d_{n,t}$')
+                 branch_investment_cost = '$\mathcal{C}^{F}_{n,t}$',
+                 one_port_operational_cost = '$\mathcal{O}_{n,t}$',
+                 branch_sparcity_cost = '$\mathcal{R}^{F sparcity}_{n,t}$',
+                 one_port_sparcity_cost = '$\mathcal{G sparcity}_{n,t}$',
+                 co2_cost = '$\mathcal{E}_{n,t}$',
+                 nodal_demand_cost = '$\lambda_{n,t} \, d_{n,t}$'
+                 )
 
 to_static_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}_{n}$',
-                      branch_investment_cost = '$\mathcal{C}^{F}_{n}$',
-                      one_port_operational_cost = '$\mathcal{O}_{n}$',
-                      co2_cost = '$\mathcal{E}_{n}$',
-                      nodal_demand_cost = '$\lambda_{n} \, d_{n}$')
+                        branch_investment_cost = '$\mathcal{C}^{F}_{n}$',
+                        one_port_operational_cost = '$\mathcal{O}_{n}$',
+                        branch_sparcity_cost = '$\mathcal{R}^{F sparcity}_{n}$',
+                        one_port_sparcity_cost = '$\mathcal{G sparcity}_{n}$',
+                        co2_cost = '$\mathcal{E}_{n}$',
+                        nodal_demand_cost = '$\lambda_{n} \, d_{n}$')
 
 
 to_total_symbol = dict(one_port_investment_cost = '$\mathcal{C}^{G}$',
                       branch_investment_cost = '$\mathcal{C}^{F}$',
                       one_port_operational_cost = '$\mathcal{O}$',
+                      branch_sparcity_cost = '$\mathcal{R}^{F sparcity}$',
+                      one_port_sparcity_cost = '$\mathcal{G sparcity}$',
[…]

### 2020-08-18 code/examples.py

@@ -22,7 +22,7 @@ plt.rc('font', family='serif')
 
 path = '../figures/example'
 
-net = False
+net = True
 power = 'net' if net else 'gross'
 method = 'ebe'
 
@@ -45,7 +45,7 @@ n.madd('Generator', [0, 1], bus=[n0, n1], p_nom_extendable=True,
 n.madd('Line', ['1'], s_nom_extendable=True, x=0.01, bus0=[n0], bus1=[n1],
        capital_cost=c_l)
 n.lopf(pyomo=False, keep_shadowprices=True, solver_name='gurobi')
-n.generators_t.mu_upper = adjust_shadowprice(n.generators_t.mu_upper, n, 'Generator')
+# n.generators_t.mu_upper = adjust_shadowprice(n.generators_t.mu_upper, n, 'Generator')
 
 sn = 'now'
 

### 2020-08-21 code/allocate_network.py
fix 'sparcity' -> 'scarcity' add scarcity and subsidy cost in total cost plot

@@ -78,8 +78,8 @@ O = A * o
 # not positive.
 # =============================================================================
 
-def sparcity_share(c):
-    """Get the ratio of payments allocated to the sparcity constraint."""
+def scarcity_share(c):
+    """Get the ratio of payments allocated to the scarcity constraint."""
     mu_upper = 'mu_upper_' + nominal_attrs[c]
     return (- n.df(c)[mu_upper] /
             (n.df(c).capital_cost - n.df(c)[mu_upper]))
@@ -89,16 +89,16 @@ def sparcity_share(c):
 c = 'Generator'
 mu_gen = by_bus_carrier(n.pnl(c).mu_upper, c, n).rename(source_dims)
 C_gen = mu_gen * A.sel(source_carrier=n.df(c).carrier.unique())
-C_spa_gen = C_gen * by_bus_carrier(sparcity_share(c), c, n).rename(source_dims)
+C_sca_gen = C_gen * by_bus_carrier(scarcity_share(c), c, n).rename(source_dims)
 # rename for not conflicting later
 C_gen = C_gen.rename(source_carrier='source_carrier_gen')
 
 if not 'test' in snakemake.input.network:
-    sparcity = n.df(c).p_nom_max.replace(np.inf, 0) @ n.df(c).mu_upper_p_nom
+    scarcity = n.df(c).p_nom_max.replace(np.inf, 0) @ n.df(c).mu_upper_p_nom
     subsidy = n.df(c).p_nom_min @ n.df(c).mu_lower_p_nom
 
     investment = n.df(c).p_nom_opt @ n.df(c).capital_cost
-    revenue = (C_gen.sum() + subsidy + sparcity)
+    revenue = (C_gen.sum() + subsidy + scarcity)
     assert_allclose(investment, revenue, rtol=1e-2, atol=np.inf)
 
 
@@ -121,10 +121,10 @@ mu = pd.concat([n.pnl(c).mu_upper + n.pnl(c).mu_lower for c in comps],
 mu = xr.DataArray(mu, dims=['snapshot', 'branch'])
 C_branch = (A_f * mu).rename(bus='sink')
 
-share_sparcity = pd.concat([sparcity_share(c) for c in comps], keys=comps,
+share_scarcity = pd.concat([scarcity_share(c) for c in comps], keys=comps,
                            names=names)
-share_sparcity = xr.DataArray(share_sparcity, dims='branch')
-C_spa_branch = C_branch * share_sparcity
[…]

### 2020-08-21 code/check_shadowprices.py

@@ -8,11 +8,12 @@ Created on Wed Aug 12 11:06:12 2020
 
 import pypsa
 from helpers import adjust_shadowprice
+import numpy as np
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('check_shadowprices', nname='de10bf')
+        snakemake = mock_snakemake('check_shadowprices', nname='acdc')
 
 if not 'test' in snakemake.input[0]:
     n = pypsa.Network(snakemake.input[0])
@@ -38,10 +39,9 @@ if not 'test' in snakemake.input[0]:
 
     investment = g.eval('p_nom_opt * capital_cost')
     revenue = (gt.mu_upper * gt.p).sum()
-
-    gamma = adjust_shadowprice(gt.mu_upper, 'Generator', n)
-    revenue = (gamma * gt.p).sum()
-    assert close(investment.where(gt.p.sum()>0, 0), revenue)
+    subsidy = g.p_nom_min * g.mu_lower_p_nom
+    scarcity = g.p_nom_max.replace(np.inf, 0) * g.mu_upper_p_nom
+    assert close(investment, revenue + subsidy + scarcity)
 
 
 
@@ -52,12 +52,9 @@ if not 'test' in snakemake.input[0]:
 
     investment = l.eval('s_nom_opt * capital_cost')
     revenue = ((lt.mu_upper + lt.mu_lower) * lt.p0).sum()
-
-
-    gamma = lt.mu_upper + lt.mu_lower
-    gamma_adj = adjust_shadowprice(gamma, 'Line', n)
-    revenue = (gamma_adj * lt.p0).sum()
-    assert close(investment, revenue)
+    subsidy = l.s_nom_min * l.mu_lower_s_nom
+    scarcity = l.s_nom_max.replace(np.inf, 0) * l.mu_upper_s_nom
+    assert close(investment, revenue + subsidy + scarcity)
 
 
     l = n.links
@@ -65,12 +62,10 @@ if not 'test' in snakemake.input[0]:
 
     investment = l.eval('p_nom_opt * capital_cost')
     revenue = ((lt.mu_upper + lt.mu_lower) * lt.p0).sum()
+    subsidy = l.p_nom_min * l.mu_lower_p_nom
+    scarcity = l.p_nom_max.replace(np.inf, 0) * l.mu_upper_p_nom
+    assert close(investment, revenue + subsidy + scarcity)
[…]

### 2020-08-21 code/config.py

@@ -9,6 +9,8 @@ Created on Fri May 15 15:20:25 2020
 import pandas as pd
 import matplotlib.pyplot as plt
 
+# plt.rc('text', usetex=True)
+# plt.rc('text.latex', preamble=r'\usepackage{amsmath}')
 plt.rc('font', family='serif')
 plt.rcParams['savefig.dpi'] = 200
 
@@ -25,10 +27,12 @@ symbol = dict(generator_investment_cost = '\mathcal{C}^{G}',
               branch_investment_cost = '\mathcal{C}^{F}',
               one_port_operational_cost = '\mathcal{O}',
               one_port_investment_cost = '\mathcal{C}',
-              branch_sparcity_cost = '\mathcal{R}^{F sparcity}',
-              one_port_sparcity_cost = '\mathcal{G sparcity}',
+              branch_scarcity_cost = '\mathcal{R}^{F scarcity}',
+              one_port_scarcity_cost = '\mathcal{G scarcity}',
               co2_cost = '\mathcal{E}',
               remaining_cost = '\mathcal{R}',
+              scarcity_cost = '\mathcal{R}^{scarcity}',
+              subsidy_cost = '\mathcal{R}^{subsidy}',
               lmp = '\lambda',
               demand = 'd')
 
@@ -47,18 +51,23 @@ to_explanation = {'one_port_operational_cost': 'OPEX',
                   'generator_investment_cost': 'Production CAPEX',
                   'storage_investment_cost': 'Storage CAPEX',
                   'branch_investment_cost': 'Transmission CAPEX',
-                  'branch_sparcity_cost': 'Sparcity Cost',
-                  'generator_sparcity_cost': 'Sparcity Cost',
+                  'branch_scarcity_cost': 'Scarcity Cost',
+                  'generator_scarcity_cost': 'Scarcity Cost',
+                  'scarcity_cost': 'Scarcity Cost',
+                  'subsidy_cost': 'Subsidy Cost',
                   'nodal_demand_cost': 'Nodal Payment'}
 
 color = pd.Series({'one_port_operational_cost': 'darkkhaki',
                    'one_port_investment_cost': 'palevioletred',
                    'co2_cost': 'tomato',
                    'generator_investment_cost': 'palevioletred',
-                   'storage_investment_cost': 'royalblue',
+                   'storage_investment_cost': 'skyblue',
                    'branch_investment_cost': 'mediumaquamarine',
[…]

### 2020-08-21 code/helpers.py

@@ -21,7 +21,7 @@ def load(n):
               .rename(bus="sink")
 
 def combine_oneports(ds):
-    if 'generator_sparcity_cost' in ds:
+    if 'generator_scarcity_cost' in ds:
         return ds
     gen = ds.generator_investment_cost.rename(source_carrier_gen='source_carrier')
     sus = ds.storage_investment_cost.rename(source_carrier_sus='source_carrier')

### 2020-08-21 code/plot_maps_transfer.py

@@ -18,7 +18,7 @@ from helpers import combine_oneports, scfmt, load
 
 if 'snakemake' not in globals():
     from _helpers import mock_snakemake
-    snakemake = mock_snakemake('plot_sparcity_maps', nname='test-de10bf',
+    snakemake = mock_snakemake('plot_scarcity_maps', nname='test-de10bf',
                                method='ptpf', power='net')
 
 

### 2020-08-21 code/plot_network.py

@@ -7,6 +7,7 @@ Created on Fri Jul 10 10:39:11 2020
 """
 
 import pypsa
+import config
 import matplotlib.pyplot as plt
 import netallocation as ntl
 import cartopy.crs as ccrs

### 2020-08-21 code/plot_total_costs.py

@@ -10,6 +10,8 @@ import pandas as pd
 import xarray as xr
 import matplotlib.pyplot as plt
 from config import color, to_total_symbol
+import numpy as np
+from pypsa.descriptors import nominal_attrs
 
 plt.rc('font', family='serif')
 plt.rc("text", usetex=False)
@@ -18,15 +20,33 @@ plt.rc("text", usetex=False)
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('plot_total_costs', nname='de10bf')
+        snakemake = mock_snakemake('plot_total_costs', nname='acdc')
 
 n = pypsa.Network(snakemake.input.network)
 ca = xr.open_dataset(snakemake.input.costs)
 
-TC = n.objective_constant + n.objective
+
 
 stacked_cost = ca.sum().to_array().to_series()
-stacked_cost['remaining_cost'] =  TC - stacked_cost.sum()
+
+
+if 'lv_limit' in n.global_constraints.index:
+    for c in n.branch_components:
+        if n.df(c).empty: continue
+        mu_upper = (n.global_constraints.at['lv_limit', 'mu'] * n.df(c).length)
+        nom = nominal_attrs[c]
+        n.df(c)['mu_upper_'+nom] += mu_upper
+
+stacked_cost['subsidy_cost'] = 0
+stacked_cost['scarcity_cost'] = 0
+for c in ['Generator', 'StorageUnit', 'Line', 'Link']:
+    nom = nominal_attrs[c]
+    stacked_cost['subsidy_cost'] += n.df(c)[nom + '_min'] @ n.df(c)['mu_lower_' + nom]
+    stacked_cost['scarcity_cost'] += (n.df(c)[nom + '_max'].replace(np.inf, 0) @
+                                      n.df(c)['mu_upper_' + nom])
+
+
+assert ((n.objective_constant + n.objective)/stacked_cost.sum()).round(2) == 1
 
 # %%
 fig, ax = plt.subplots(figsize=[3,5])

### 2020-08-21 code/price_decomposition.py

@@ -56,5 +56,5 @@ ax.set_xticks([])
 ax.set_yticklabels([])
 
 
-# fig.tight_layout()
+fig.tight_layout()
 fig.savefig('../figures/price_decomposition.png', bbox_inches='tight')

### 2020-08-26 paper/flow_based.tex
fix 'sparcity' -> 'scarcity' add scarcity and subsidy cost in total cost plot

@@
 
-\subsection{Allocating OPEX}
+\subsection{OPEX Allocation}
 
@@
 In order to showcase the cost allocation at a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
-The model builds on the PyPSA-EUR workflow \cite{horsch_pypsa-eur_2020} with technigal details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
-We follow a browfield approach where transmission lines start from today's capacity values originally retreived from the ENTSO-E Transmission System Map \cite{entsoe-e_entso-e_nodate}.
+The model builds on the PyPSA-EUR workflow \cite{horsch_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
+We follow a browfield approach where transmission lines can be expanded starting from today's capacity values, originally retreived from the ENTSO-E Transmission System Map \cite{entsoe-e_entso-e_nodate}.
 Pre-installed generation capacities for wind and solar are distributed proportionally to the average power potential at each site excluding those with an average capacity factor of 10\%.
@@
 
-\Cref{fig:network} shows the estimate of existent capacities on the left hand side and the capacity expansions on the right. 
 \begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{de50bf}
+    \includegraphics[width=\linewidth]{de50bf/network}
     \label{fig:network}
+    \caption{Brownfield optimization of the German power system.
+The left side shows existent renewable capacities which serve as lower capacity limits for the optimization.
+The right side shows the capacity expansion of renewables as well as installation of backup gas power plants.
+The effective CO$_2$ price is set to 120 \euro per tonne CO$_2$ emission.}
 \end{figure*}
+The left side of \cref{fig:network} shows the lower capacity bounds for renewable generators and transmission infrastructure, the right side the capacity expansion for generation, storage and transmission.
+The optimization expands the solar capacity in the south, onshore and offshore wind in the upper north and most west.
+OCGT generators are build all over the network except for most south and north.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
+    \label{fig:total_cost}
+    \caption{Total allocated payments of the system. }
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{de50bf/average_price}
+    \label{fig:average_price}
+    \caption{Average electricity price $\lmp$ in the network.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{de50bf/locality_ptpf_net}
+    \label{fig:locality}
+    \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
+\end{figure}
 

### 2020-08-26 code/plot_network.py
update tables, small things

@@ -42,7 +42,7 @@ if is_brownfield:
         link_widths=n.links.p_nom_min * branch_scale,
         ax=axes[0],
         geomap='10m',
-        title='Existent Capacities',
+        title='Lower Capacity Bounds',
         boundaries=regions.total_bounds[[0,2,1,3]]
     )
     regions.plot(ax=axes[0], transform=ccrs.PlateCarree(), aspect='equal')

### 2020-08-26 paper/flow_based.tex
add write table rule
paper, last chapter update install.sh for custom power plants

@@
 In order to showcase the cost allocation at a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
-The model builds on the PyPSA-EUR workflow \cite{horsch_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
-We follow a browfield approach where transmission lines can be expanded starting from today's capacity values, originally retreived from the ENTSO-E Transmission System Map \cite{entsoe-e_entso-e_nodate}.
-Pre-installed generation capacities for wind and solar are distributed proportionally to the average power potential at each site excluding those with an average capacity factor of 10\%.
+The model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
+We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
+Pre-installed generation capacities totals for wind and solar are distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
 Further, wind and solar capacity expansion are limited by land use restriction.
@@
 Additionally, unlimited batteries and H$_{2}$-storages are allowed at each node.
-We impose a effective carbon price of 120 \euro per tonne-CO$_{2}$.
+We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$.
+All effective cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are reported in detail in \cref{sec:cost_assumptions}.
 
@@
     \includegraphics[width=\linewidth]{de50bf/network}
-    \label{fig:network}
     \caption{Brownfield optimization of the German power system.
@@
 The effective CO$_2$ price is set to 120 \euro per tonne CO$_2$ emission.}
+    \label{fig:network}
 \end{figure*}
-The left side of \cref{fig:network} shows the lower capacity bounds for renewable generators and transmission infrastructure, the right side the capacity expansion for generation, storage and transmission.
+The resulting network is shown in  \cref{fig:network} with its lower capacity bounds for renewable generators and transmission infrastructure on the left and capacity expansion for generation, storage and transmission on the right.
 The optimization expands the solar capacity in the south, onshore and offshore wind in the upper north and most west.
-OCGT generators are build all over the network except for most south and north.
+Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network.
+Transmission lines are amplified in along the North-South axis, including one large DC link, leading from the coastal region to the South-West.
+
+\Cref{fig:average_price} displays the average electricity price $\lmp$ per region. We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).  
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{de50bf/average_price}
+    \caption{Average electricity price $\lmp$ in the network.}
+    \label{fig:average_price}
+\end{figure}
 
@@
     \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
+    \caption{Total allocated payments of the system. }
     \label{fig:total_cost}
-    \caption{Total allocated payments of the system. }
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{de50bf/average_price}
-    \label{fig:average_price}
-    \caption{Average electricity price $\lmp$ in the network.}
 \end{figure}
@@
     \includegraphics[width=\linewidth]{de50bf/locality_ptpf_net}
+    \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
     \label{fig:locality}
-    \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
 \end{figure}
@@
 
-\subsection{Example: Power Flow Allocations of different Types}
+\subsection{Numerical Example: Power Flow Allocation of type \ref{net}\ref{ebe}}
 \label{sec:example_plots}
@@
 
+\subsection{Working Example}
+
+\subsubsection{Cost Assumptions}
+\label{sec:cost_assumptions}
+
+\begin{table*}
+    \centering
+    \input{../tables/de50bf_ptpf_net/prices}    
+    \caption{Operational and capital price assumptions for all type of assets used in the working example.
+The capital price for transmission lines are given in [k\,\euro/MW/km].
+The cost assumptions are retrieved fromt the PyPSA-EUR model \cite{horsch_jonas_pypsa-eur_2020}.}
+\end{table*}
+
 \clearpage

### 2020-08-27 paper/flow_based.tex
update paper (insert more figures) allocate_payment adjust figsize

@@
     \caption{Brownfield optimization of the German power system.
-The left side shows existent renewable capacities which serve as lower capacity limits for the optimization.
+The left side shows existent renewable capacities, matching the total capacity for the year 2017, which serve as lower capacity limits for the optimization.
 The right side shows the capacity expansion of renewables as well as installation of backup gas power plants.
@@
 Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network.
-Transmission lines are amplified in along the North-South axis, including one large DC link, leading from the coastal region to the South-West.
-
-\Cref{fig:average_price} displays the average electricity price $\lmp$ per region. We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).  
+Transmission lines are amplified in along the North-South axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the South-West.
+
+\Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
+We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).
+Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices.
+The node with the lowest LMP in the upper North-West, stands out through high pre-installed offshore capacities.
 \begin{figure}
@@
 
+\begin{figure*}
+    \centering
+    \begin{subfigure}[c]{.6\linewidth}
+    \includegraphics[width=\linewidth]{de50bf/ptpf_net_to_lowest-lmp}
+    \end{subfigure}
+    \hspace{-.2151\linewidth}
+    \begin{subfigure}[c]{.6\linewidth}
+    \includegraphics[width=\linewidth]{de50bf/ptpf_net_to_highest-lmp}
+    \end{subfigure}
+    \caption{Comparison of payments of the node with the \textbf{lowest LMP (left)} and the node with the \textbf{highest LMP (right)}.
+The region of the paying bus is colored in dark blue.
+The circles indicate where to which bus and technology combined OPEX and CAPEX payments.
+Further the thickness of the lines indicates the dedicated amount of payments.
+The cheap prices in the North...
+}
+    \label{fig:direct-allocation}
+\end{figure*}
+
 \begin{figure}
@@
 
-\section{Appendix}
+\section{Network Optimization}
 
@@
 
-\subsection{Brownfield Optimization and Capacity Restricitions}
-
-Constraining the capacities $\capacitygeneration$  for a subset $S$ of generators to lower or upper limits in the form of
-\begin{align}
-    \capacitygenerationlower - \capacitygeneration \le 0 \resultsin{\mulowergenerationnom} \hpad \forall{s \in S} \label{eq:capacityexpansionminimum}\\
-    \capacitygeneration - \capacitygenerationupper \le 0 \resultsin{\muuppergenerationnom} \hpad \forall{s \in S}
-\label{eq:capacityexpansionmaximum}
-\end{align}
-alters the objective value as soon as one of those become bounding. 
-The complementary slackness for both are 
-\begin{align}
-    \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{s \in S} \\
-    \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{s \in S}
-\end{align}
-
-The CAPEX paybacks for generators and transmission lines
-\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
-\begin{align}
-\pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
-\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{s \in S}
-\label{eq:capex_generation_duality_bf}
-\end{align}
-for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacitygeneration$ leads us to 
-\begin{align}
- \capexgeneration_s - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
- \label{eq:capex_generation_duality_bf2}
-\end{align}
-where we define the cost resulting from the capacity expansion limits as
-\begin{align}
-    \remainingcost_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacitygeneration
-    \Forall{s}
-\end{align}
-As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \ge 0$, the latter can be a net positive or net negative cost term.
-
-Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
-\begin{align}
-    \capexgeneration_s + \opexgeneration_s  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
-\end{align}
-We see that the revenue of generator $s$ will not exactly pay back its full CAPEX and OPEX.
+% \subsection{Brownfield Optimization and Capacity Restricitions}
+
+% Constraining the capacities $\capacitygeneration$  for a subset $S$ of generators to lower or upper limits in the form of
+% \begin{align}
+%     \capacitygenerationlower - \capacitygeneration \le 0 \resultsin{\mulowergenerationnom} \hpad \forall{s \in S} \label{eq:capacityexpansionminimum}\\
+%     \capacitygeneration - \capacitygenerationupper \le 0 \resultsin{\muuppergenerationnom} \hpad \forall{s \in S}
+% \label{eq:capacityexpansionmaximum}
+% \end{align}
+% alters the objective value as soon as one of those become bounding. 
+% The complementary slackness for both are 
+% \begin{align}
+%     \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{s \in S} \\
+%     \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{s \in S}
+% \end{align}
+
+% The CAPEX paybacks for generators and transmission lines
+% \cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
+% \begin{align}
+% \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
+% \capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{s \in S}
+% \label{eq:capex_generation_duality_bf}
+% \end{align}
+% for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacitygeneration$ leads us to 
+% \begin{align}
+%  \capexgeneration_s - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
+%  \label{eq:capex_generation_duality_bf2}
+% \end{align}
+% where we define the cost resulting from the capacity expansion limits as
+% \begin{align}
+%     \remainingcost_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacitygeneration
+%     \Forall{s}
+% \end{align}
+% As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \ge 0$, the latter can be a net positive or net negative cost term.
+
+% Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
+% \begin{align}
+%     \capexgeneration_s + \opexgeneration_s  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
+% \end{align}
+% We see that the revenue of generator $s$ will not exactly pay back its full CAPEX and OPEX.
 The exogenous constraint shifts the zero-profit equations such that some of the expenditures for $s \in S$ cannot directly be allocated to consumers.
@@
 
-\begin{align}
-    \muuppergeneration \rightarrow \dfrac{\capitalpricegeneration}{\capitalpricegeneration - \mulowergenerationnom + \muuppergenerationnom}\, \muuppergeneration
-\end{align}
-By doing so we ensure that the allocated CAPEX sum up to the actual CAPEX at generator $s$, thus 
-\begin{align}
-    \capexgeneration_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
-\end{align}
-
-Doing likewise with a subset $L$ of transmission lines,
-\begin{align}
-\capacityflow \ge \capacityflowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
-\capacityflow \le \capacityflowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
-\end{align}
-results in a shifted zero-profit rule for transmission line in the form 
-\begin{align}
-    \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow \Forall{\ell \in L}
-\end{align}
-where $\remainingcost_\ell$ is given by 
-\begin{align}
-    \remainingcost_\ell = \left(\mulowerflownom - \muupperflownom \right) \, \capacityflow
-    \Forall{\ell \in L}
-\end{align}
-
-% \begin{align}
-%     \muupperflownom \left( \capacityflow - \capacityflowUpper \right) =0 \\
-%     \mulowerflownom \left( \capacityflow - \capacityflowLower \right) =0 \\
-% \end{align}
-
-% \begin{align}
-%     \pdv{\lagrangian}{\capacityflow}  = 0 \,\, \rightarrow \,\, \\
-%     \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
-%     \label{eq:capex_flow_duality_bf}
-%     \end{align}
-%     for transmission lines.
-% Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
-% \begin{align}
-%   \capitalpricegeneration \, \capacitygeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacitygenerationlower \right) + \muuppergenerationnom \left( \capacitygenerationupper - \capacitygenerationlower \right) \Forall{n,s}
-% \end{align}
-
+% \begin{align}
+%     \muuppergeneration \rightarrow \dfrac{\capitalpricegeneration}{\capitalpricegeneration - \mulowergenerationnom + \muuppergenerationnom}\, \muuppergeneration
+% \end{align}
+% By doing so we ensure that the allocated CAPEX sum up to the actual CAPEX at generator $s$, thus 
+% \begin{align}
+%     \capexgeneration_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
+% \end{align}
+
+% Doing likewise with a subset $L$ of transmission lines,
+% \begin{align}
+% \capacityflow \ge \capacityflowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
+% \capacityflow \le \capacityflowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
+% \end{align}
+% results in a shifted zero-profit rule for transmission line in the form 
+% \begin{align}
+%     \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow \Forall{\ell \in L}
+% \end{align}
+% where $\remainingcost_\ell$ is given by 
+% \begin{align}
+%     \remainingcost_\ell = \left(\mulowerflownom - \muupperflownom \right) \, \capacityflow
+%     \Forall{\ell \in L}
+% \end{align}
+
+% % \begin{align}
+% %     \muupperflownom \left( \capacityflow - \capacityflowUpper \right) =0 \\
+% %     \mulowerflownom \left( \capacityflow - \capacityflowLower \right) =0 \\
+% % \end{align}
+
+% % \begin{align}
+% %     \pdv{\lagrangian}{\capacityflow}  = 0 \,\, \rightarrow \,\, \\
+% %     \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
+% %     \label{eq:capex_flow_duality_bf}
+% %     \end{align}
+% %     for transmission lines.
+% % Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
+% % \begin{align}
+% %   \capitalpricegeneration \, \capacitygeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacitygenerationlower \right) + \muuppergenerationnom \left( \capacitygenerationupper - \capacitygenerationlower \right) \Forall{n,s}
+% % \end{align}
+
+\section{Allocation Schemes}
 \subsection{Allocating Gross Injections with EBE}
@@
 
-Allocating net injections using the AP method is derived from \cite{achayuthakan_electricity_2010-1}.
+Allocating net injections using the AP method is derived from \cite{achayuthakan_electricity_2010}.
 In a lossless network the downstream and upstream formulations result in the same P2P allocation which is why we restrict ourselves to the downstream formulation only.
@@
 \begin{figure}[h!]
-    \begin{subfigure}[c]{\linewidth}
+    \begin{subfigure}[c]{.99\linewidth}
     \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
@@
     \end{subfigure}
-    \begin{subfigure}[c]{\linewidth}
+    \begin{subfigure}[c]{.99\linewidth}
     \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
@@
 \end{figure}
-
-\subsection{Working Example}
-
-\subsubsection{Cost Assumptions}
-\label{sec:cost_assumptions}
-
-\begin{table*}
+% 
+\section{Working Example}
+The following figures contain more detailed information about the peer-to-peer cost allocation.
+The cost or prices payed by consumers are indicated by the region color.
+The dedicated revenue is displayed in proportion to the size of cycles (for assets attached to buses) or to the thickness of transmission branches.
+\begin{figure*}
+    \centering
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/OCGT_one_port_investment_cost}
+        \subcaption{OCGT}
+        \label{fig:ocgt_capex}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/onwind_one_port_investment_cost}
+        \subcaption{Onshore Wind}
+        \label{fig:onshore_capex}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/solar_one_port_investment_cost}
+        \subcaption{Solar}
+        \label{fig:solar_capex}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/H2_one_port_investment_cost}
+        \subcaption{H$_{2}$-Storage}
+        \label{fig:h2_capex}
+    \end{subfigure}
+    \caption{Average electricity price due to \textbf{CAPEX allocation}, $\sum_t \allocatecapexgeneration$ and $\sum_t \allocatecapexstorage$, for OCGT, Onshore Wind, batteries and Hydrogen Storage.
+Average Allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
+    \label{fig:scarcity-price}
+\end{figure*}
+
+\begin{table*}[h]
     \centering
@@
 
+\begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
+    \caption{Average electricity \textbf{price for emission cost}, $\sum_t \allocateemissioncost$.}
+    \label{fig:emission_cost}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
+    \caption{Average electricity \textbf{price for the scarcity}, $\sum_t \allocatescarcitycost$ for the transmission system, due to the upper transmission expansion limit of 25\%.}
+    \label{fig:branch_scarcity_cost}
+\end{figure}
+
+\begin{figure*}
+    \centering
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/offwind-dc_generator_scarcity_cost}
+        \subcaption{Offshore Wind}
+        \label{fig:offwind-dc_generator_scarcity_cost}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/onwind_generator_scarcity_cost}
+        \subcaption{Onshore Wind}
+        \label{fig:onwind_generator_scarcity_cost}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/solar_generator_scarcity_cost}
+        \subcaption{Solar}
+        \label{fig:solar_generator_scarcity_cost}
+    \end{subfigure}
+    \begin{subfigure}[c]{.49\linewidth}
+        \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/ror_generator_scarcity_cost}
+        \subcaption{Run-of-River}
+        \label{fig:ror_generator_scarcity_cost}
+    \end{subfigure}
+    \caption{Average electricity \textbf{price for the scarcity}, $\sum_t \allocatescarcitycost$, due to technology resource limits, for offshore wind, onshore wind,  solar, Run-of-River.
+Average Allocated Scarcity Cost per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
+    \label{fig:scarcity-price}
+\end{figure*}
+
 \clearpage

### 2020-08-27 code/plot_allocated_payment.py

@@ -74,7 +74,7 @@ bus_colors = pd.concat([capex_colors, opex_colors, emission_color])
 bus_scale = 1e-6
 branch_scale = 3e-3
 
-fig, ax = plt.subplots(figsize=(5,5), subplot_kw={'projection': ccrs.EqualEarth()})
+fig, ax = plt.subplots(figsize=(4.5,4.5), subplot_kw={'projection': ccrs.EqualEarth()})
 n.plot(line_widths=branch_widths['Line'] * branch_scale,
        link_widths=branch_widths['Link'] * branch_scale,
        line_colors=branch_colors['Line'],

### 2020-08-27 install.sh

@@ -12,5 +12,5 @@ git clone https://github.com/PyPSA/technology-data.git
 # python 
 # '''
 # import powerplantmatching as pm
-# pm.data.OPSD_VRE_country("DE").to_csv("pypsa-de/data/custom_powerplants.csv")
+# pm.data.OPSD_VRE_country("DE").powerplant.convert_country_to_alpha2().to_csv("pypsa-de/data/custom_powerplants.csv")
 # '''

### 2020-08-28 paper/flow_based.tex
adjust numerical example section, but (!) change it now to only ap net

@@
 
-\title{Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable}
+\title{A dispatch-based approach to fully allocate costs in optimal power systems}
 \author{F. Hofmann, Brown T.}
@@
  \item The LMP always breaks down into contributions of different cost terms (OPEX, CAPEX, etc.) of the system.
- \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments.
- \item Methods like Average Participation or Flow Based Market Coupling allow for localizing the P2P cost assignments and prioritize assignments within one market zone.
+ \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments, where prominent methods as Average Participation or Flow Based Market Coupling can be applied.
+ \item ...
+ \item ...
 \end{itemize}
@@
 
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of mutliple cost terms.
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms.
 Typically, these include operational expenditures (OPEX) $\opexgeneration$ for generators, expenditures for emissions $\emissioncost$, capital expenditures (CAPEX) for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and possible other terms, \ie
@@
 % 
-This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are partially or totally payed back by the consumers 
+This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
 \begin{align}
@@
 
-% where $\costfactor$ is either a fixed operational price or a composition of shadowprices
-Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
+% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
+Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$.
+This leads us to
 \begin{align}
@@
 \end{align}
-The product on the right hand side 
-\begin{align}
-    \allocatestate = \pdv{\state}{\demand} \demand
+The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. 
+\begin{align}
+    \allocatestate \rightarrow\pdv{\state}{\demand} \demand
     \label{eq:state_allocation}
 \end{align}    
+...
 may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with.
@@
 % \end{align}
-The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
+The  shadow prices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
 Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
@@
     \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
+    \label{eq:scarcitycost}
 \end{align}
@@
 % \end{align}
-% In constrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
+% In contrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
 % As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.
@@
 that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
-\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010-1} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
+\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
 \item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions.
-The interregional exchange is only allocating net power decifit or excess of the regions.
+The interregional exchange is only allocating net power deficit or excess of the regions.
 \label{fbmc}
@@
 If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.
+However, it relies on further assumptions such as the Generation Shift Keys determining the production which are deployed for the inter-regional exchange.
+From this point of view, we set the main focus of this research to the AP based scheme, as locality and little de-pendency on external decisions are strong arguments for a transparent cost allocation.
+However, we will use the EBE based cost allocation for the following example, in order to depict the functionality of the cost allocation and the fundamental difference between the allocation of gross and net power.
 
@@
     % 
-In the following, we illustrate the cost allocation and explain its functionality by taking the example of a two bus system with one optimized time step, shown in \cref{fig:example_network}. \\
-% 
-% 
-% 
-% 
-The two buses are connected via one transmission line, each has one generator.
+% 
+% 
+Consider a two bus system, as shown in \cref{fig:example_network}, with one transmission line and one generator per bus.
 Whereas generator~1 (at bus~1) has an operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a higher operational price of 200~\euro/\megawatthour.
-For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacitygenerationupper$~=~100~MW.
+For both the CAPEX rate is set to 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW.
 The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit.
@@
     \end{subfigure}
-    \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
+    \caption{Power allocations for gross power injection using the EBE scheme, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
 Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.
 The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
-    The sum of both net flows equals the resulting flow of $f_1=40$~MW. \vspace{-0pt}}
+    The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
     \label{fig:example_allocation}
@@
 As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$.
-As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further tranmission expansion, it is rewarded proportional to the relief.}
+As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further transmission expansion, it is rewarded proportional to the relief.}
     \label{fig:example_payoff}
@@
 % 
-\subsubsection*{Allocating Gross Power using EBE (\ref{gross}\ref{ebe}) }
-
-\Cref{fig:example_allocation} shows the allocated transactions on basis of equvalently traded gross power (type \ref{gross}\ref{ebe}) for both buses 1~\&~2 separately.
-The resulting P2P payments are given in \cref{fig:example_payoff}.
+\subsubsection*{Allocating Gross Power Production}
+
+\Cref{fig:example_allocation} shows the allocated transactions on basis of gross power injection for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
 The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained.
@@
 The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2.
-As $d_2$ is higher than $d_1$ the retrievements from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators.
+As $d_2$ is higher than $d_1$, the reception from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators.
 But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion.
@@
 The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$.
-These values match their overall spendings, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$.
+These values match their overall spending, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$.
 The sum of all columns yields the total payment of consumers at bus $n$.
@@
 The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only.
-In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity contraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
-
-\subsubsection*{Allocating Net Power using EBE (\ref{net}\ref{ebe}) }
+In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
+
+\subsubsection*{Allocating Net Power Production}
 
@@
     \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
-    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when equivalently allocation net power injection (scheme \ref{ebe}\ref{net}).}
+    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when allocating net power injection using the EBE scheme (type \ref{net}\ref{ebe}).
+This leads to less and more intuitive payments.}
     \label{fig:example_payoff_net_ebe}
@@
 Neither has bus~1 to pay fee to the transmission system as it only exports power.
-So all its consumers pay to its local generators.
-Bus~2 in constrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
-% 
-Again the cummulative payments per bus meet the nodal spendings $\lmp \, \demand$. The cummulative revenues per generator and transmission line meet the all CAPEX and OPEX.
-
-As the other schemes \ref{ap}\ref{gross} - \ref{fbmc}\ref{net} apart from the EBE scheme don't distinguishably modify allocations within the presented example, we move on to a more realistic setup where systemic and local effects show up.
+So, consumers at bus~1 pay to its local generator.
+Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
+% 
+Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$.
+The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX.
+Note this gives the same result as when allocating net injection with the Average Participation \ref{net}\ref{ap}.
+
+The example shows that allocating net power injection only, reduces the number of peer-to-peer payments significantly, which leads to a much clearer pictures.
+The same counts for the AP scheme which we will use in the following application case.
+
+%We emphasize that the AP based allocation does not imply traceability of the resulting flows, as only the \textit{source}-\textit{sink} relations $\allocatepeer$ are determined by it, the flow allocation still follow the power plow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
 
@@
 
-In order to showcase the cost allocation at a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
+For showcasing the behavior of the cost allocation in a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
 The model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
 We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
-Pre-installed generation capacities totals for wind and solar are distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
+Pre-installed generation capacity totals for the year 2017 for wind and solar were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
 Further, wind and solar capacity expansion are limited by land use restriction.
 These consider agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{corine2012,natura2000}.
-Pumped Hydro Storages (PHS) are fixed to today's capacities with no more expansion allowed.
-Additionally, unlimited batteries and H$_{2}$-storages are allowed at each node.
-We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$.
-All effective cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are reported in detail in \cref{sec:cost_assumptions}.
-
-\begin{figure*}
+Pumped Hydro Storages (PHS) and Run-of-River power plants are fixed to today's capacities with no more expansion allowed.
+Additionally, unlimited expansion of batteries and H$_{2}$-storages and Open-Cycle Gas Turbines (OCGT) are allowed at each node.
+% 
+\begin{figure*}[t]
     \centering
@@
 \end{figure*}
-The resulting network is shown in  \cref{fig:network} with its lower capacity bounds for renewable generators and transmission infrastructure on the left and capacity expansion for generation, storage and transmission on the right.
+We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, with an gross emission of 180~kg/MWh and an efficiency of 39\% for OCGT, adds an price of 55 \euro/\megawatthour.
+All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.
+The optimized network is shown in  \cref{fig:network} with lower capacity bounds for renewable generators and transmission infrastructure on the left and capacity expansion for generation, storage and transmission on the right.
 The optimization expands the solar capacity in the south, onshore and offshore wind in the upper north and most west.
@@
 
-\Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
-We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).
-Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices.
-The node with the lowest LMP in the upper North-West, stands out through high pre-installed offshore capacities.
 \begin{figure}
@@
 \end{figure}
+
+The total annualized cost of the power system roughly sum up to 42 billion \euro.
+\Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
+We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).
+Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices.
+The node with the lowest LMP in the upper North-West, stands out through high pre-installed offshore capacities.
+\\
+
+We allocate all costs under the assumtion that the \textit{net} dispatch is assigned according to the AP method (type \ref{net}\ref{ap}). This allocation suits the best for 
+From the perspective of consumers, the largest proportion of the system cost is covered by payments associated with CAPEX for generators, transmission system and storage units in decreasing order, see \cref{fig:total_cost}.
+In \cref{fig:capex_price} we depict the regional distribution of the average price which consumers pay for a selection of different technologies.
+It turns out that  mostly areas with high CAPEX allocation are the ones with low average LMP.
+In particular, this counts for the coastal regions where strong offshore and onshore investments are taken.
+Inside the country, some regions also pay a fairly high shares dedicated to CAPEX, which are however allocated to non-local assets.
+The most even spatial distribution of CAPEX allocations accounts for OCGT.
+These are however primarily built nearby strong wind capacities.
+
+Together with the emission cost $\emissioncost$, the total OPEX $\opex$ amount around 16 billion \euro.
+As to expect, XX\% are dedicated to OCGT alone.
+In \cref{fig:opex_price,fig:emission_cost} we show the regional distribution of payments per MWh and the resulting revenues for generator...
+
+The scarcity cost $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit, sum up to roughly 7.5 b\euro.
+Note that according to \cref{eq:scarcitycost} this term is negative.
+It translates to  depicting the cost that consumers pay ``too much'' for assets limited in their capacity expansion.
+In the real world this money would be used to pay for augmented land costs in the dedicated areas.
+...
+
+The cost caused by lower capacity constraints for pre-existing assets which violate the optimal design are collected in $\subsidycost$. This ...
 
@@
 The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL).
-Alternativaly, we can the demand $\demand$ has to be supplied by the attached assets,
+Alternatively, we can the demand $\demand$ has to be supplied by the attached assets,
 \begin{align}
@@
 
-The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e.
+The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the transmission capacity in positive or negative direction, i.e.
 \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding.
@@
 
-When multipying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
+When multiplying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
 \begin{align}
@@
 \end{align}
-Againg we can use this to formulate the zero-profit rule for transmission lines. We mulitply \cref{eq:opex_flow_duality} with $\flow$, which finally leads us to 
+Again we can use this to formulate the zero-profit rule for transmission lines. We multiply \cref{eq:opex_flow_duality} with $\flow$, which finally leads us to 
 \begin{align}
@@
 % Watch out Incidence = - ordinary Incidence !!
-It states that the congestion revenue of a line (first term right hand side) substracted by the cost for cycle constraint exactly matches its CAPEX.
+It states that the congestion revenue of a line (first term right hand side) reduced by the cost for cycle constraint exactly matches its CAPEX.
 
@@
 
-\subsection{CO$_2$ constraint}
+\subsection{Emission Constraint}
 
@@
 
-% \subsection{Brownfield Optimization and Capacity Restricitions}
+% \subsection{Brownfield Optimization and Capacity Restrictions}
 
@@
 
-The allocation of gross generation to demands $\demand$ is straigthforwardly obained by a proportional distribution of the generation, \ie
+The allocation of gross generation to demands $\demand$ is straightforwardly obtained by a proportional distribution of the generation, \ie
 
@@
     \begin{subfigure}[c]{.49\linewidth}
-        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/OCGT_one_port_investment_cost}
-        \subcaption{OCGT}
-        \label{fig:ocgt_capex}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_investment_cost_total}
+        \subcaption{All production and storage technologies}
+        \label{fig:total_capex}
     \end{subfigure}
@@
     \begin{subfigure}[c]{.49\linewidth}
-        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/H2_one_port_investment_cost}
-        \subcaption{H$_{2}$-Storage}
-        \label{fig:h2_capex}
+        \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/by_carrier/OCGT_one_port_investment_cost}
+        \subcaption{OCGT}
+        \label{fig:ocgt_capex}
     \end{subfigure}
-    \caption{Average electricity price due to \textbf{CAPEX allocation}, $\sum_t \allocatecapexgeneration$ and $\sum_t \allocatecapexstorage$, for OCGT, Onshore Wind, batteries and Hydrogen Storage.
+    \caption{Average electricity price due to \textbf{CAPEX allocation}, $\sum_t \allocatecapexgeneration$ and $\sum_t \allocatecapexstorage$, for all assets, onshore wind, solar and OCGT.
 Average Allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
-    \label{fig:scarcity-price}
+    \label{fig:capex_price}
 \end{figure*}
@@
 The capital price for transmission lines are given in [k\,\euro/MW/km].
-The cost assumptions are retrieved fromt the PyPSA-EUR model \cite{horsch_jonas_pypsa-eur_2020}.}
+The cost assumptions are retrieved from the PyPSA-EUR model \cite{horsch_jonas_pypsa-eur_2020}.}
+    \label{tab:cost_assumptions}
 \end{table*}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_operational_cost_total}
+    \caption{Average electricity price \textbf{OPEX allocation}, $\sum_t \allocateopex$.}
+    \label{fig:opex_price}
+\end{figure}
 
@@
     \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
-    \caption{Average electricity \textbf{price for emission cost}, $\sum_t \allocateemissioncost$.}
+    \caption{Average electricity price for \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$.}
     \label{fig:emission_cost}
@@
     \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
-    \caption{Average electricity \textbf{price for the scarcity}, $\sum_t \allocatescarcitycost$ for the transmission system, due to the upper transmission expansion limit of 25\%.}
+    \caption{Average electricity price for the \textbf{allocated scarcity cost}, $\sum_t \allocatescarcitycost$ for the transmission system, due to the upper transmission expansion limit of 25\%.}
     \label{fig:branch_scarcity_cost}
@@
     \end{subfigure}
-    \caption{Average electricity \textbf{price for the scarcity}, $\sum_t \allocatescarcitycost$, due to technology resource limits, for offshore wind, onshore wind,  solar, Run-of-River.
+    \caption{Average electricity price for \textbf{allocated scarcity cost}, $\sum_t \allocatescarcitycost$, due to technology resource limits, for offshore wind, onshore wind,  solar, run-of-river.
 Average Allocated Scarcity Cost per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
-    \label{fig:scarcity-price}
+    \label{fig:scarcity_price}
 \end{figure*}

### 2020-08-28 code/examples.py

@@ -22,9 +22,9 @@ plt.rc('font', family='serif')
 
 path = '../figures/example'
 
-net = True
+net = False
 power = 'net' if net else 'gross'
-method = 'ebe'
+method = 'ptpf'
 
 
 n = pypsa.Network()
@@ -57,7 +57,8 @@ g0, g1 = n.generators_t.p.loc['now']
 f = n.lines_t.p0.loc['now'].item()
 l0, l1 = n.buses_t.marginal_price.loc['now']
 
-ds = ntl.allocate_flow(n, method='ebe', q=0, aggregated=net)
+kwargs = {q: 0} if method == 'ebe' else {}
+ds = ntl.allocate_flow(n, method=method, aggregated=net, **kwargs)
 ds = ntl.convert.peer_to_peer(ds, n)
 dc = ntl.cost.nodal_demand_cost(n).rename(bus='payer')
 ca = ntl.allocate_cost(n, method=ds, q=0)

### 2020-08-28 paper/flow_based.tex
last chapter

@@
 \newcommand{\mulowernom}{\ubar{\mu}^\text{nom}_{i}}
+
+\newcommand{\kk}{k~\euro\,}
 
@@
 \newcommand{\lmp}[1][n]{\lambda_{#1,t}}
+\newcommand{\averagelmp}{\left<\lmp\right>_t}
 \newcommand{\demand}[1][n]{d_{#1,t}}
@@
 However, it relies on further assumptions such as the Generation Shift Keys determining the production which are deployed for the inter-regional exchange.
-From this point of view, we set the main focus of this research to the AP based scheme, as locality and little de-pendency on external decisions are strong arguments for a transparent cost allocation.
-However, we will use the EBE based cost allocation for the following example, in order to depict the functionality of the cost allocation and the fundamental difference between the allocation of gross and net power.
+
+From this point of view, we restrict the focus of this research to \textbf{net} injection allocation on the basis of \textbf{AP}, type \ref{net}\ref{ap}, as locality and no further dependency on external decisions are strong arguments for a transparent cost allocation.
+We emphasize that the AP based allocation does not imply traceability of the resulting flows, as only the \textit{source}-\textit{sink} relations $\allocatepeer$ are determined by it, the flow allocation still follows the power plow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
+In the following example, we illustrate the functionality of the cost allocation by an example of a small artificial network.
 
@@
 Consider a two bus system, as shown in \cref{fig:example_network}, with one transmission line and one generator per bus.
-Whereas generator~1 (at bus~1) has an operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a higher operational price of 200~\euro/\megawatthour.
-For both the CAPEX rate is set to 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW.
-The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit.
+Whereas generator~1 (at bus~1) has a cheap operational price of 50 \euro/\megawatthour, an expensive operational price of 200~\euro/\megawatthour is set to  generator~2 (at bus~2).
+For both the capital investments amount 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW.
+The transmission line has a capital price of 100~\euro/MW and no upper capacity limit.
 With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW.
 The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW.
-% 
-% 
-% \begin{figure}[h]
-%     \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
-%     \label{fig:example_allocation_bus1}
-%     \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
-Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.}
-% \end{figure}
-% % 
-% \begin{figure}[b]
-%     \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
-%     \label{fig:example_allocation_bus2_gross_ebe}
-%     \caption{Type \ref{gross}\ref{ebe} power allocation for bus~2 of the example network in \cref{fig:example_network}.
-Bus~2 retrieves 60~MW from bus~1 and self supplies 30~MW.
-The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
-% \end{figure}
+\\
 %  
@@
     \begin{subfigure}[c]{.495\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
+    % NOTE: Net EBE and Net AP lead to the same result here. Due to time limits just take the ebe figures
+    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
     \vspace{-40pt}
@@
     \begin{subfigure}[c]{.495\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
+    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
     \vspace{-40pt}
     \subcaption{}
-    \label{fig:example_allocation_bus2_gross_ebe}
+    \label{fig:example_allocation_bus2}
     \end{subfigure}
-    \caption{Power allocations for gross power injection using the EBE scheme, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}.
-Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.
-The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
-    The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
+    \caption{Power allocations of net power injections using the AP scheme, type \ref{net}\ref{ap}, in the example network, \cref{fig:example_network}.
+Consumers at for bus~1, figure (a), are, with 60~MW, totally supplied by the local generator and do not consume any imported power from bus 2.
+In contrast consumer at bus~2, figure (b), retrieve 40~MW from generator 1, induce a flow of 40~MW on the transmission line and consume 50~MW from the local generator.}
     \label{fig:example_allocation}
@@
 % 
-% 
 \begin{figure}[h]
     \centering
-    \includegraphics[width=\linewidth]{example_payoff_gross_ebe.png}
-    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}.
-The payments are derived on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}.
-Consumers at bus $n$ have to pay each generator proportional to their consumption.
-As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$.
-As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further transmission expansion, it is rewarded proportional to the relief.}
+    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
+    \caption{Full P2P cost allocation on the basis of power allocations shown in \cref{fig:example_allocation}.
+Consumers compensate OPEX and CAPEX of the generators they retrieve from, see \cref{fig:example_allocation}.
+As bus~1 is totally self-supplying, all it payment is assigned to the local generator.
+As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX at bus~1 and CAPEX for the transmission.}
     \label{fig:example_payoff}
@@
 % 
-\subsubsection*{Allocating Gross Power Production}
-
-\Cref{fig:example_allocation} shows the allocated transactions on basis of gross power injection for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
-The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained.
-With only one generator at bus 1, consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator~1.
-The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$.
-As this flow is in contrary direction to the total flow, it is relieving the transmission system.
-This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{\ell = 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.
-
-The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2.
-As $d_2$ is higher than $d_1$, the reception from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators.
-But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion.
-Hence the payoff to the transmission system is positive and much higher than for bus~1.
-
-The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$.
-These values match their overall spending, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$.
-The sum of all columns yields the total payment of consumers at bus $n$.
+\Cref{fig:example_allocation} shows the allocated transactions on the basis of allocation \ref{net}\ref{ap} for both buses 1~\&~2 separately.
+Note, the ``sum'' of the two figures give to the actual dispatch.
+The corresponding P2P payments are given in \cref{fig:example_payoff}.
+
+The left graph \cref{fig:example_allocation_bus1} shows that $d_1$ is with 60~MW totally supplied by the local production.
+Consequently consumers at bus~1 pay 3k~\euro~OPEX, which is the operational price of 50~\euro\, times the 60~MW.
+Further they pay 33k~\euro~ for the CAPEX at generator~1.
+Note that 3k~\euro~ of these account for the scarcity imposed buy the upper expansion limit $\capacitygenerationupper$.
+The rest makes up 60\% of the CAPEX spent at generator~1, exactly the share of power allocated to $d_1$.
+Consumers at bus~1 don't pay any transmission CAPEX as no flow is associated with there demand.
+\\
+The right graph \cref{fig:example_allocation_bus2} shows the power allocations to $d_2$.
+We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from bus~1.
+Thus, consumers have the pay for the local OPEX and CAPEX as well as the expenditures at bus~1 and the transmission system.
+As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it.
+The allocated CAPEX exactly matches the actual CAPEX of generator~2.
+However in the case of generator~1, 2\kk of the 22\kk CAPEX allocation are associated to the scarcity cost.
+The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line (zero-profit condition).
+% 
+The sum of all values in the payoff matrix in \cref{fig:example_payoff} yield $\totalcost - \scarcitycost$, the total system cost minus the scarcity cost (which in turn is negative).
+The sum of a column  yields the total revenue per the asset $i$.
+These values match their overall spending plus the cost of scarcity.
+The sum a row returns the total payment of consumers at bus $n$.
 For example the sum of payments of consumers at bus~1 is 36k~\euro.
@@
 \\
-
-The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only.
-In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
-
-\subsubsection*{Allocating Net Power Production}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
-    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when allocating net power injection using the EBE scheme (type \ref{net}\ref{ebe}).
-This leads to less and more intuitive payments.}
-    \label{fig:example_payoff_net_ebe}
-    \end{figure}    
-% 
-In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments.
+%
+
+The fact that OPEX and CAPEX allocations are proportional to each other results from the fact of optimizing one time step only.
+This coherence breaks for for larger optimization problems with multiple time steps.
+Then CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
+\\
+The example shows that allocating net power injection leads to intuitive and transparent  cost allocations.
+
+% In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments.
 The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}.
@@
 Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
-% 
-Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$.
+% % 
+% Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$.
 The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX.
@@
 
-The example shows that allocating net power injection only, reduces the number of peer-to-peer payments significantly, which leads to a much clearer pictures.
-The same counts for the AP scheme which we will use in the following application case.
-
-%We emphasize that the AP based allocation does not imply traceability of the resulting flows, as only the \textit{source}-\textit{sink} relations $\allocatepeer$ are determined by it, the flow allocation still follow the power plow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
-
 \section{Application Case}
@@
 The model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
+
 We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
@@
 \end{figure*}
-We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, with an gross emission of 180~kg/MWh and an efficiency of 39\% for OCGT, adds an price of 55 \euro/\megawatthour.
+% 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{de50bf/average_price}
+    \caption{Average electricity price $\averagelmp$ per region as a result from the optimization of the German power system.
+Regions in the middle and south of Germany have high prices whereas electricity in the North with a strong wind, transmission and OCGT infrastructure is cheaper.}
+    \label{fig:average_price}
+\end{figure}
+We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, with an gross emission of 180~kg/MWh and an efficiency of 39\% for OCGT, adds an effective price of 55 \euro/\megawatthour.
 All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.
+
 The optimized network is shown in  \cref{fig:network} with lower capacity bounds for renewable generators and transmission infrastructure on the left and capacity expansion for generation, storage and transmission on the right.
@@
 Transmission lines are amplified in along the North-South axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the South-West.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{de50bf/average_price}
-    \caption{Average electricity price $\lmp$ in the network.}
-    \label{fig:average_price}
-\end{figure}
-
-The total annualized cost of the power system roughly sum up to 42 billion \euro.
+% 
+% 
+The total annualized cost of the power system roughly sums up to 42 billion \euro.
 \Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
@@
 
-We allocate all costs under the assumtion that the \textit{net} dispatch is assigned according to the AP method (type \ref{net}\ref{ap}). This allocation suits the best for 
+We allocate all costs under the assumption that the \textit{net} dispatch is assigned according to the AP method (type \ref{net}\ref{ap}). 
 From the perspective of consumers, the largest proportion of the system cost is covered by payments associated with CAPEX for generators, transmission system and storage units in decreasing order, see \cref{fig:total_cost}.
-In \cref{fig:capex_price} we depict the regional distribution of the average price which consumers pay for a selection of different technologies.
+In \cref{fig:capex_price} we depict the share of the average nodal price $\left<\lmp\right>_t$ which is dedicated to the CAPEX of different technologies.
 It turns out that  mostly areas with high CAPEX allocation are the ones with low average LMP.
@@
 The most even spatial distribution of CAPEX allocations accounts for OCGT.
-These are however primarily built nearby strong wind capacities.
+These are however primarily built nearby strong wind capacities in the North.
 
 Together with the emission cost $\emissioncost$, the total OPEX $\opex$ amount around 16 billion \euro.
-As to expect, XX\% are dedicated to OCGT alone.
-In \cref{fig:opex_price,fig:emission_cost} we show the regional distribution of payments per MWh and the resulting revenues for generator...
-
-The scarcity cost $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit, sum up to roughly 7.5 b\euro.
-Note that according to \cref{eq:scarcitycost} this term is negative.
-It translates to  depicting the cost that consumers pay ``too much'' for assets limited in their capacity expansion.
-In the real world this money would be used to pay for augmented land costs in the dedicated areas.
-...
-
-The cost caused by lower capacity constraints for pre-existing assets which violate the optimal design are collected in $\subsidycost$. This ...
+As to expect, 99.97\% are dedicated to OCGT alone.
+For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price,fig:emission_cost}.
+Compared to the CAPEX allocation for OCGT, the OPEX distributes in a more local manner.
+This results from the fact that during ``standard'' demand peaks, it is rather local OCGT generators which serve as backup generators.
+In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power.
+The OPEX allocations for renewables play with 0.2 \euro/MWh in average an inferior role.
+
+The negative segment in \cref{fig:total_cost} is associated with the scarcity costs $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit.
+These sum up approximately to 7.5 b\euro.
+Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments.
+It translates to the cost that consumers pay ``too much'' for assets limited in their capacity expansion.
+In the real world this money would be used to pay for augmented land costs or civic participation in the dedicated areas.
+In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average price $\left<\lmp\right>_t$.
+The scarcity for wind and solar are relatively low and impact the average price by around 2\euro/MWh.
+They mostly affect regions with strong investments and those in close vicinity.
+Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16\euro.
+This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
+However, as these power plants in particular are already amortized and scarcity cost should be reconsidered and removed from a final cost allocation.
+The scarcity cost for the transmission system, resulting from the limited transmission volume expansion of 25\% mostly influence regions in the middle of Germany, where in average 8\euro/MWh account for transmission scarcity, see \cref{fig:branch_scarcity_price}.
+This illustrates the strong impact that the line volume constraint has on the optimization.
+Regions in the North are not touched by the constraint.
+
+The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
+These violate the optimal design and are not compensated by the nodal payments $\demand \, \lmp$.
+This ...
 
@@
 
-\subsection{Numerical Example: Power Flow Allocation of type \ref{net}\ref{ebe}}
-\label{sec:example_plots}
-\begin{figure}[h!]
-    \begin{subfigure}[c]{.99\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
-    \vspace{-40pt}
-    \subcaption{}
-    \label{fig:example_allocation_bus1_net_ebe}
-    \end{subfigure}
-    \begin{subfigure}[c]{.99\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
-    \vspace{-40pt}
-    \subcaption{}
-    \label{fig:example_allocation_bus2_net_ebe}
-    \end{subfigure}
-    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allocated net power injections (scheme \ref{net}\ref{ebe}).
-Bus~1 retrieves 60~MW from itself and nothing from bus~2.
-The latter in turn retrieves 40~MW from bus~1 and self supplies 50~MW.
-The P2P trades are less in number and more intuitive then with allocating gross flow.}
-    \label{fig:example_allocation_net_ebe}
-\end{figure}
 % 
@@
     \end{subfigure}
-    \caption{Average electricity price due to \textbf{CAPEX allocation}, $\sum_t \allocatecapexgeneration$ and $\sum_t \allocatecapexstorage$, for all assets, onshore wind, solar and OCGT.
+    \caption{Average \textbf{CAPEX allocation} per MWh, $\sum_t \allocatecapexgeneration / \sum_t \demand$ and $\sum_t \allocatecapexstorage / \sum_t \demand$, for all production and storage assets (a), onshore wind (b), solar (c) and OCGT (d).
 Average Allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
@@
     \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_operational_cost_total}
-    \caption{Average electricity price \textbf{OPEX allocation}, $\sum_t \allocateopex$.}
+    \caption{Average \textbf{OPEX allocation} per consumed MWh, $\sum_t \allocateopex/\sum_t \demand$.}
     \label{fig:opex_price}
@@
     \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
-    \caption{Average electricity price for \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$.}
+    \caption{Average \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$, per consumed MWh.}
     \label{fig:emission_cost}
@@
     \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
-    \caption{Average electricity price for the \textbf{allocated scarcity cost}, $\sum_t \allocatescarcitycost$ for the transmission system, due to the upper transmission expansion limit of 25\%.}
-    \label{fig:branch_scarcity_cost}
+    \caption{Average \textbf{allocated transmission scarcity cost} per consumed MWh. This scarcity cost results from the upper transmission expansion limit of 25\%.}
+    \label{fig:branch_scarcity_price}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
+    \caption{Average power mix per region.}
+    \label{fig:power_mix}
 \end{figure}

### 2020-08-28 paper/leftovers/example_gross_ebe_description.tex
last chapter

\subsubsection*{Allocating Gross Power Production with EBE}

\begin{figure*}[h!]
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1}
    \end{subfigure}
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_gross_ebe}
    \end{subfigure}
    \caption{Power allocations for gross power injection using the EBE scheme, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2. The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
    The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
    \label{fig:example_allocation}
\end{figure*}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_gross_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}. The payments are derived on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}. Consumers at bus $n$ have to pay each generator proportional to their consumption. As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$. As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further transmission expansion, it is rewarded proportional to the relief.}
    \label{fig:example_payoff}
\end{figure}

\Cref{fig:example_allocation} shows the allocated transactions on basis of gross power injection for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained. With only one generator at bus 1, consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator~1. The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$. As this flow is in contrary direction to the total flow, it is relieving the transmission system. This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{\ell = 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.

The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2. As $d_2$ is higher than $d_1$, the reception from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators. But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion. Hence the payoff to the transmission system is positive and much higher than for bus~1.

The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$. These values match their overall spending, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$. The sum of all columns yields the total payment of consumers at bus $n$. For example the sum of payments of consumers at bus~1 is 36k~\euro. This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$. \\

The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only. In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.

\subsubsection*{Allocating Net Power Production}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when allocating net power injection using the EBE scheme (type \ref{net}\ref{ebe}). This leads to less and more intuitive payments.}
    \label{fig:example_payoff_net_ebe}
    \end{figure}    
% 
In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments. The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}. As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated. Neither has bus~1 to pay fee to the transmission system as it only exports power. So, consumers at bus~1 pay to its local generator. Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1. 
% 
Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$. The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX. Note this gives the same result as when allocating net injection with the Average Participation \ref{net}\ref{ap}.

The example shows that allocating net power injection only, reduces the number of peer-to-peer payments significantly, which leads to a much clearer pictures. The same counts for the AP scheme which we will use in the following application case.

\subsection{Numerical Example: Power Flow Allocation of type \ref{net}\ref{ebe}}
\label{sec:example_plots}
\begin{figure}[h!]
    \begin{subfigure}[c]{.99\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1_net_ebe}
    \end{subfigure}
    \begin{subfigure}[c]{.99\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_net_ebe}
    \end{subfigure}
    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allocated net power injections (scheme \ref{net}\ref{ebe}). Bus~1 retrieves 60~MW from itself and nothing from bus~2. The latter in turn retrieves 40~MW from bus~1 and self supplies 50~MW. The P2P trades are less in number and more intuitive then with allocating gross flow.}
    \label{fig:example_allocation_net_ebe}
\end{figure}

### 2020-09-07 paper/flow_based.tex
paper: revise application section, add last section code: add branch width to plot_map_transfer, correct scale in subsidy plot

@@
 \end{align} 
-where an asset describes any component of the network, such as a generator, line, energy storage etc.
+where an asset describes any operating part of the network, such as a generator, line, energy storage etc.
 %For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.
@@
 \subsubsection{Storages}
+\label{sec:storages}
 
@@
 \subsection{Design Constraints}
-
+\label{sec:design_constraints}
 Power system modelling does rarely follow a pure Greenfield approach with unlimited capacity expansion.
@@
     h_j \left(\state, \capacity \right) - K < 0
+    \label{eq:design_constraint}
 \end{align}
@@
 \end{align}
-In the following we highlight two often used classes of constraints and show how to incorporate them into the cost allocation.
+The effective revenue per asset $i$ associated with cost term $\cost$ can then be defined as 
+\begin{align}
+    \cost' = \sum_i  \cost_i - \remainingcost_i
+    \label{altered_cost}
+\end{align}
+Note that $\cost'$ is the amount which is allocated to consumers which, according to the quality of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, can either be larger, equal or lower than $\cost$.
+In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to incorporate them into the cost allocation.
 
@@
  \flow  = \sum_m \ptdf[m] \left( \nodalgeneration[m] - \demand[m] \right)
+ \label{eq:power_flow_equation}  
 \end{align}
@@
 
-For showcasing the behavior of the cost allocation in a more complex system, we apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
+We showcase the behavior of the cost allocation in a more complex system, by applying it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
 The model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
@@
 We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
-Pre-installed generation capacity totals for the year 2017 for wind and solar were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
+Pre-installed wind and solar generation capacity totals of the year 2017 were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
 Further, wind and solar capacity expansion are limited by land use restriction.
@@
 \end{figure}
-We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, with an gross emission of 180~kg/MWh and an efficiency of 39\% for OCGT, adds an effective price of 55 \euro/\megawatthour.
+We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, for OCGT, adds an effective price of 55 \euro/\megawatthour (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%).
 All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.
 
-The optimized network is shown in  \cref{fig:network} with lower capacity bounds for renewable generators and transmission infrastructure on the left and capacity expansion for generation, storage and transmission on the right.
-The optimization expands the solar capacity in the south, onshore and offshore wind in the upper north and most west.
+The optimized network is shown in  \cref{fig:network}.
+On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission.
+The optimization expands solar capacities in the south, onshore and offshore wind in the upper north and most west.
 Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network.
 Transmission lines are amplified in along the North-South axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the South-West.
-% 
-% 
 The total annualized cost of the power system roughly sums up to 42 billion \euro.
+%
+
 \Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
@@
 
-We allocate all costs under the assumption that the \textit{net} dispatch is assigned according to the AP method (type \ref{net}\ref{ap}). 
-From the perspective of consumers, the largest proportion of the system cost is covered by payments associated with CAPEX for generators, transmission system and storage units in decreasing order, see \cref{fig:total_cost}.
-In \cref{fig:capex_price} we depict the share of the average nodal price $\left<\lmp\right>_t$ which is dedicated to the CAPEX of different technologies.
+The costs are allocated according to type \ref{net}\ref{ap}, DC-lines are incorporated into the PTDF in \cref{eq:power_flow_equation} using the method described in \cite{hofmann_flow_2020-1}.
+In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost'$.
+Due to the design constraints, \cref{sec:design_constraints}, and the role of storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$.
+The figure also includes costs for scarcity $\scarcitycost$ and subsidy  $\subsidycost$.
+Note that the sum of all contributions in \cref{fig:total_cost} equals the total cost $\totalcost$.
+
+The largest proportion of the payments is associated with CAPEX for generators, transmission system and storage units in decreasing order. 
 It turns out that  mostly areas with high CAPEX allocation are the ones with low average LMP.
-In particular, this counts for the coastal regions where strong offshore and onshore investments are taken.
-Inside the country, some regions also pay a fairly high shares dedicated to CAPEX, which are however allocated to non-local assets.
-The most even spatial distribution of CAPEX allocations accounts for OCGT.
-These are however primarily built nearby strong wind capacities in the North.
+In particular, this counts for the coastal regions where strong offshore and onshore investments are taken, see \cref{fig:capex_price}.
+Particular regions inside the country also pay comparably high share to CAPEX, however these are allocated to non-local assets.
+We further observe that CAPEX for OCGT is allocated very evenly among all consumers.
+Again, these payments benefit primarily regions nearby strong wind capacities in the North.
 
@@
 The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
-These violate the optimal design and are not compensated by the nodal payments $\demand \, \lmp$.
-This ...
+These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$.
 
@@
 \end{figure}
+
+\section{Conclusion}
+
+In this work we presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from asset to consumer.
+Starting from an optimized long-term investment model, we were able to decompose separate cost terms into contributions of single consumers.
+For three typical classes of assets, namely generators, transmission lines and storage units, we showed how operational prices and shadow prices must be weighted with the dispatch allocation in order to allocate all system costs.
+Further we highlighted the impact of constraints with a non-zero constant on the right hand side, such as lower and upper capacity expansion limits, which alter the locational marginal price and therefore distort the cost allocations.
+In the case of upper capacity expansion limits, this leads to an additional charge for consumers which have to compensate for \eg land use restrictions.
+For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated to consumers.
+.
 

### 2020-09-07 code/plot_maps_transfer.py

@@ -64,7 +64,7 @@ for var in cost:
         if 'branch' in var: continue
         if (p[carrier].sum() <= 0.001): continue
 
-        expend = varname.replace('Production & Storage ', '')
+        expand = varname.replace('Production & Storage ', '')
         ncarrier = n.carriers.nice_name[carrier]
 
         fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
@@ -73,7 +73,7 @@ for var in cost:
 
         n.plot(bus_sizes=r[carrier]/r[carrier].sum()*0.7, ax=ax, **nplot_kwargs)
         regions.plot(column=p[carrier], ax=ax, **rplot_kwargs,
-                     legend_kwds={'label': f'{payment_type} {ncarrier} {expend} [{unit}]',
+                     legend_kwds={'label': f'{payment_type} {ncarrier} {expand} [{unit}]',
                                   'format': fmt})
 
         fig.canvas.draw(), fig.tight_layout()
@@ -84,7 +84,12 @@ for var in cost:
                            figsize=(5, 4))
     ax.spines['geo'].set_visible(False)
 
-    if 'branch' not in var:
+    if 'branch' in var:
+        normed = r/r.sum()
+        n.plot(bus_sizes=0, line_widths=normed['Line'],
+               link_widths=normed['Link'], ax=ax,
+               boundaries=regions.total_bounds[[0,2,1,3]], geomap=res)
+    else:
         n.plot(bus_sizes=r.sum(1)/p.sum().sum(), ax=ax, **nplot_kwargs)
     regions.plot(column=p if 'branch' in var else p.sum(1),
                  ax=ax, **rplot_kwargs,

### 2020-09-07 code/plot_subsidy.py

@@ -25,11 +25,12 @@ n = pypsa.Network(snakemake.input.network)
 regions = gpd.read_file(snakemake.input.regions).set_index('name')
 
 
-oneports = pd.concat([n.df(c).groupby(['bus', 'carrier']).mu_lower_p_nom.sum()
+oneports = pd.concat([n.df(c).eval('cost = mu_lower_p_nom * p_nom_opt')
+                      .groupby(['bus', 'carrier']).cost.sum()
                       for c in ['Generator', 'StorageUnit']]).sort_index()
 
-line_widths  = n.lines.mu_lower_s_nom
-link_widths = n.links.mu_lower_p_nom
+line_widths  = n.lines.eval('mu_lower_s_nom * s_nom_opt')
+link_widths = n.links.eval('mu_lower_p_nom * p_nom_opt')
 branch_sum = line_widths.sum() + link_widths.sum()
 
 bus_scale = 2
@@ -60,12 +61,12 @@ fig.legend(
 
 
 # legend generator capacities
-reference_caps = [50e3, 10e3]
+reference_caps = [50e6, 10e6]
 scale = oneports.sum() / bus_scale / projected_area_factor(ax)**2
 handles = make_legend_circles_for(reference_caps, scale=scale,
                                   facecolor="w", edgecolor='grey',
                                   alpha=.5)
-labels = ["%ik €"%(s / 1e3) for s in reference_caps]
+labels = ["%iM €"%(s / 1e6) for s in reference_caps]
 handler_map = make_handler_map_to_scale_circles_as_in(ax)
 legend = fig.legend(handles, labels,
                 loc="upper left", bbox_to_anchor=(1., 0.4),
@@ -75,10 +76,10 @@ fig.add_artist(legend)
 
 # legend transmission capacitues
 handles, labels = [], []
-reference_caps = [10e3, 5e3]
+reference_caps = [10e6, 5e6]
 handles = [Line2D([0], [0], color='grey', linewidth=s * branch_scale / branch_sum)
            for s in reference_caps]
-labels = ['%ik €'%(s / 1e3) for s in reference_caps]
+labels = ['%iM €'%(s / 1e6) for s in reference_caps]
 
 legend = fig.legend(handles, labels,
                     loc="lower left", bbox_to_anchor=(1, .1),

### 2020-09-07 code/plot_subsidy.py
update legend sizes in allocated payment, maps_transfer, and subsidy

@@ -48,7 +48,7 @@ n.plot(bus_sizes=oneports/oneports.sum() * bus_scale,
        boundaries=regions.total_bounds[[0,2,1,3]],
        )
 regions.plot(ax=ax, transform=ccrs.PlateCarree(), aspect='equal',
-             color='white', lw=0.3, edgecolor='grey')
+             color='white', lw=0.2, edgecolor='grey')
 
 
 fig.legend(
@@ -75,8 +75,8 @@ legend = fig.legend(handles, labels,
 fig.add_artist(legend)
 
 # legend transmission capacitues
+reference_caps = [100e6, 50e6]
 handles, labels = [], []
-reference_caps = [10e6, 5e6]
 handles = [Line2D([0], [0], color='grey', linewidth=s * branch_scale / branch_sum)
            for s in reference_caps]
 labels = ['%iM €'%(s / 1e6) for s in reference_caps]

### 2020-09-08 paper/flow_based.tex
add time_series plot

@@
 In particular, this counts for the coastal regions where strong offshore and onshore investments are taken, see \cref{fig:capex_price}.
-Particular regions inside the country also pay comparably high share to CAPEX, however these are allocated to non-local assets.
+Several regions inside the country also pay comparably high share to CAPEX, however these are allocated to non-local assets.
 We further observe that CAPEX for OCGT is allocated very evenly among all consumers.
@@
 In this work we presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from asset to consumer.
-Starting from an optimized long-term investment model, we were able to decompose separate cost terms into contributions of single consumers.
+Starting from an optimized long-term investment model, we were able to decompose single cost terms into contributions of single consumers.
 For three typical classes of assets, namely generators, transmission lines and storage units, we showed how operational prices and shadow prices must be weighted with the dispatch allocation in order to allocate all system costs.
-Further we highlighted the impact of constraints with a non-zero constant on the right hand side, such as lower and upper capacity expansion limits, which alter the locational marginal price and therefore distort the cost allocations.
+Further we highlighted the impact of additional design constraints, that is constraints with a non-zero constant on the right hand side, such as lower and upper capacity expansion limits.
+These alter the locational marginal price and therefore distort the cost allocations.
 In the case of upper capacity expansion limits, this leads to an additional charge for consumers which have to compensate for \eg land use restrictions.
-For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated to consumers.
-.
+For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
 
@@
 \begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
+    \caption{Average power mix per region.}
+    \label{fig:power_mix}
+\end{figure}
+
+\begin{figure}
     \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
-    \caption{Average power mix per region.}
-    \label{fig:power_mix}
+    \includegraphics[width=\linewidth]{de50bf/subsidy}
+    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).}
+    \label{fig:subsidy}
 \end{figure}

### 2020-09-08 code/plot_time_series.py

@@ -1,0 +1,41 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Sep  8 10:34:01 2020
+
+@author: fabian
+"""
+import pypsa
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+
+
+if 'snakemake' not in globals():
+    from _helpers import mock_snakemake
+    snakemake = mock_snakemake('plot_capex_timeseries', nname='acdc')
+
+
+n = pypsa.Network(snakemake.input.network)
+
+months = mdates.MonthLocator()  # every month
+
+
+gen = ((n.generators_t.p * n.generators_t.mu_upper)
+       .groupby(n.generators.carrier, axis=1).sum())
+
+sto = ((n.storage_units_t.p * n.storage_units_t.mu_upper)
+       .groupby(n.storage_units.carrier, axis=1).sum())
+
+capex = pd.concat([gen, sto], axis=1)
+
+fig, ax = plt.subplots(figsize = (8, 3))
+
+capex.div(1e6).plot(ax=ax)
+
+ax.set_ylabel('Million €')
+ax.set_xlabel('')
+ax.xaxis.set_major_formatter(mdates.DateFormatter('%m-%d'))
+fig.autofmt_xdate()
+
+

### 2020-09-08 paper/flow_based.tex
add plot of snapshot with highest expenditure

@@
 \begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/subsidy}
+    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).}
+    \label{fig:subsidy}
+\end{figure}
+
+\begin{figure*}
+    \includegraphics[width=\linewidth]{de50bf/capex_duration_curve}
+    \caption{Duration curve of the CAPEX allocation for production and storage technologies.
+Hours are sorted such that the total amount of allocated expenditure per hour decreases.
+With 2.7 bn~\euro\, the first value pushes investments extraordinary high.
+It is dominated by CAPEX for OCGT to which 92\% of the payments are allocated.
+This hour alone occasions about half of all OCGT CAPEX.
+The remaining time-steps are dominated by revenues for onshore wind.}
+    \label{fig:capex_duration}
+\end{figure*}
+
+\begin{figure}
     \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
@@
     \label{fig:branch_scarcity_price}
-\end{figure}
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{de50bf/subsidy}
-    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).}
-    \label{fig:subsidy}
 \end{figure}

### 2020-09-08 code/plot_duration_curve.py

@@ -8,24 +8,23 @@ Created on Tue Sep  8 10:34:01 2020
 import pypsa
 import pandas as pd
 import matplotlib.pyplot as plt
-import matplotlib.dates as mdates
 
 
 if 'snakemake' not in globals():
     from _helpers import mock_snakemake
-    snakemake = mock_snakemake('plot_capex_timeseries', nname='de50bf')
+    snakemake = mock_snakemake('plot_capex_duration_curve', nname='de50bf')
 
 
 n = pypsa.Network(snakemake.input.network)
 
-months = mdates.MonthLocator()  # every month
 
+c = 'Generator'
+gen = (n.pnl(c).p * n.pnl(c).mu_upper).groupby(n.df(c).carrier, axis=1).sum()
 
-gen = ((n.generators_t.p * n.generators_t.mu_upper)
-       .groupby(n.generators.carrier, axis=1).sum())
-
-sto = ((n.storage_units_t.p * n.storage_units_t.mu_upper)
-       .groupby(n.storage_units.carrier, axis=1).sum())
+c = 'StorageUnit'
+mu = (n.pnl(c).mu_state_of_charge / n.df(c).efficiency_dispatch
+          + n.pnl(c).mu_upper_p_dispatch + n.pnl(c).mu_lower_p_dispatch)
+sto = (n.pnl(c).p_dispatch * mu).groupby(n.df(c).carrier, axis=1).sum().clip(lower=0)
 
 capex = pd.concat([gen, sto], axis=1)
 # %%

### 2020-09-08 code/plot_operation.py

@@ -1,0 +1,35 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Sep  8 15:43:11 2020
+
+@author: fabian
+"""
+import netallocation as ntl
+import pypsa
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+import geopandas as gpd
+
+if 'snakemake' not in globals():
+    from _helpers import mock_snakemake
+    snakemake = mock_snakemake('plot_operation_high_expenditure', nname='test-de10bf')
+
+
+n = pypsa.Network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.regions).set_index('name')
+
+sn = (n.loads_t.p * n.buses_t.marginal_price).sum(1).idxmax()
+
+
+fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
+                        figsize=(5, 4))
+
+prod = ntl.power_production(n, sn, per_carrier=True).to_pandas().stack()
+
+n.plot(flow=sn, line_widths=0.01, link_widths=.01, bus_sizes=prod/1e5, ax=ax,
+       boundaries=regions.total_bounds[[0,2,1,3]], geomap='10m')
+
+fig.canvas.draw()
+fig.tight_layout()
+fig.savefig(snakemake.output[0], bbox_inches='tight')

### 2020-09-08 code/plot_subsidy.py

@@ -51,16 +51,6 @@ regions.plot(ax=ax, transform=ccrs.PlateCarree(), aspect='equal',
              color='white', lw=0.2, edgecolor='grey')
 
 
-# legend = ax.legend(
-#     *ntl.plot.handles_labels_for(n.carriers.set_index('nice_name').color),
-#     loc="upper center",
-#     ncol=2,
-#     bbox_to_anchor=(.5, 0),
-#     title='Carrier',
-#     frameon=False
-# )
-# ax.add_artist(legend)
-
 # legend generator capacities
 reference_caps = [50e6, 10e6]
 scale = oneports.sum() / bus_scale / projected_area_factor(ax)**2

### 2020-09-08 paper/flow_based.tex
follow up

@@
 It is dominated by CAPEX for OCGT to which 92\% of the payments are allocated.
-This hour alone occasions about half of all OCGT CAPEX.
-The remaining time-steps are dominated by revenues for onshore wind.}
+This hour alone occasions about the half of all OCGT CAPEX.
+The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution.
+In hours of low CAPEX allocation (after the second drop) spendings for OCGT start to increase again.
+These  }
     \label{fig:capex_duration}

### 2020-09-08 code/plot_operation.py

@@ -29,9 +29,9 @@ fig, ax = plt.subplots(subplot_kw={"projection": ccrs.EqualEarth()},
 prod = ntl.power_production(n, sn, per_carrier=True).to_pandas().stack()
 
 n.plot(flow=sn, line_widths=0.01, link_widths=.01, bus_sizes=prod/1e5, ax=ax,
-       boundaries=regions.total_bounds[[0,2,1,3]], geomap='10m')
+       boundaries=regions.total_bounds[[0,2,1,3]], geomap='10m', link_colors='rosybrown')
 regions.plot(column=n.loads_t.p.loc[sn], ax=ax, legend=True,
-             transform=ccrs.PlateCarree(), aspect='equal', cmap='Purples',
+             transform=ccrs.PlateCarree(), aspect='equal', cmap='Greens',
              alpha=0.5, legend_kwds={'alpha':.5, 'label': 'Load [MW]'}, vmin=0)
 
 fig.canvas.draw()

### 2020-09-08 paper/flow_based.tex
add new figure to paper, adjust legend in subsidy plot

@@
 With 2.7 bn~\euro\, the first value pushes investments extraordinary high.
-It is dominated by CAPEX for OCGT to which 92\% of the payments are allocated.
+Due to low renewable potentials, it is dominated by CAPEX for OCGT, to which 92\% of the payments are allocated.
 This hour alone occasions about the half of all OCGT CAPEX.
+\Cref{capex_duration} gives a detailed picture of the operational state at this time-step.
 The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution.
 In hours of low CAPEX allocation (after the second drop) spendings for OCGT start to increase again.
-These  }
+These time-steps however play a minor role.}
     \label{fig:capex_duration}
 \end{figure*}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{de50bf/operation_high_expenditure}
+    \caption{Production, flow and consumption in the system at the hour with the highest expenditures.
+The size of the circles are proportional to the power production at a node.
+Size of arrows are proportional to the flow on the transmission line.
+The depicted hour corresponds to the first value in the duration curve in \cref{fig:capex_duration}.}
+    \label{fig:operation_high_expenditure}
+\end{figure}
 

### 2020-09-08 code/plot_subsidy.py

@@ -60,7 +60,7 @@ handles = make_legend_circles_for(reference_caps, scale=scale,
 labels = ["%iM €"%(s / 1e6) for s in reference_caps]
 handler_map = make_handler_map_to_scale_circles_as_in(ax)
 legend = ax.legend(handles, labels,
-                loc="upper center", bbox_to_anchor=(0.3, 0),
+                loc="lower center", bbox_to_anchor=(0.3, 1),
                 frameon=False,
                 handler_map=handler_map)
 ax.add_artist(legend)
@@ -73,7 +73,7 @@ handles = [Line2D([0], [0], color='grey', linewidth=s * branch_scale / branch_su
 labels = ['%iM €'%(s / 1e6) for s in reference_caps]
 
 legend = ax.legend(handles, labels,
-                    loc="upper center", bbox_to_anchor=(.7, 0),
+                    loc="lower center", bbox_to_anchor=(.7, 1),
                     frameon=False,
                     )
 ax.artists.append(legend)

### 2020-09-09 paper/flow_based.tex
continue writing in application section

@@
 
-The costs are allocated according to type \ref{net}\ref{ap}, DC-lines are incorporated into the PTDF in \cref{eq:power_flow_equation} using the method described in \cite{hofmann_flow_2020-1}.
+The costs are allocated according to type \ref{net}\ref{ap}. For allocating flows on  DC-lines we use the method described in \cite{hofmann_flow_2020-1}.
 In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost'$.
-Due to the design constraints, \cref{sec:design_constraints}, and the role of storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$.
+Due to the design constraints, \cref{sec:design_constraints}, and storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$.
 The figure also includes costs for scarcity $\scarcitycost$ and subsidy  $\subsidycost$.
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
+    \caption{Total allocated payments of the system. }
+    \label{fig:total_cost}
+\end{figure}
+
 The largest proportion of the payments is associated with CAPEX for generators, transmission system and storage units in decreasing order. 
-It turns out that  mostly areas with high CAPEX allocation are the ones with low average LMP.
-In particular, this counts for the coastal regions where strong offshore and onshore investments are taken, see \cref{fig:capex_price}.
-Several regions inside the country also pay comparably high share to CAPEX, however these are allocated to non-local assets.
-We further observe that CAPEX for OCGT is allocated very evenly among all consumers.
-Again, these payments benefit primarily regions nearby strong wind capacities in the North.
+Taking the different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shows in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
+At this time (morning, end of February), the system hits the highest mismatch between low renewable power potential and high demand.
+The necessity for backup generators manifests in high allocation of CAPEX and consequently high LMP.
+With a few exceptions in the South West and East, all consumers receive power from OCGT at this time, thus all pay high amounts for the needed backup infrastructure (operational state of the system is in detailed shown in \cref{fig:operation_high_expenditure}).
+Note that this is the most extreme event, which ensures backup infrastructure for other inferior extreme events where backup power is needed.
+The total CAPEX allocation for OCGT infrastructure is depicted in \cref{fig:ocgt_capex} which correlates with out findings from the extreme event.
+\\
+Contrary to this, CAPEX for renewable infrastructure are allocated evenly throughout several thousands of hours.
+As for onshore and offshore wind farms, the produced power deeply penetrates the network, see \cref{fig:power_mix}, thus it is not only local, but also remote consumers which cover the CAPEX, see \cref{fig:capex_price}.
+This in turn benefits local consumers, which profit from the cheap power which explains why these regions end up with a low average LMP.
 
@@
 These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
-    \caption{Total allocated payments of the system. }
-    \label{fig:total_cost}
-\end{figure}
 
@@
     \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/one_port_operational_cost_total}
-    \caption{Average \textbf{OPEX allocation} per consumed MWh, $\sum_t \allocateopex/\sum_t \demand$.}
+    \caption{Average \textbf{OPEX allocation} per consumed MWh, $\sum_t \allocateopex/\sum_t \demand$.
+The effective prices for OPEX are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators and storages.
+As OCGT is the only allowed fossil based technology, the drawn allocation is proportional to OPEX allocation of OCGT generators.}
     \label{fig:opex_price}
@@
     \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
-    \caption{Average \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$, per consumed MWh.}
+    \caption{Average \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$, per consumed MWh.
+The effective prices are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators.}
     \label{fig:emission_cost}
@@
     \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
-    \caption{Average power mix per region.}
+    \caption{Average power mix per region calculated by Average Participation \ref{ap}.
+Coastal regions are mainly supplied by local offshore and onshore wind farms.
+Their strong power injections additionally penetrate the network up to the southern border.
+In the middle and South, the supply is dominated by a combination of OCGT and solar power.}
     \label{fig:power_mix}
@@
 \begin{figure}
+    \vspace{2cm}
     \includegraphics[width=\linewidth]{de50bf/subsidy}
-    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).}
+    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).
+The figure shows the built infrastructure that does not gain back its CAPEX from its market revenue, but is only build due to lower capacity limits.}
     \label{fig:subsidy}
@@
     \caption{Duration curve of the CAPEX allocation for production and storage technologies.
-Hours are sorted such that the total amount of allocated expenditure per hour decreases.
-With 2.7 bn~\euro\, the first value pushes investments extraordinary high.
-Due to low renewable potentials, it is dominated by CAPEX for OCGT, to which 92\% of the payments are allocated.
+Hours are sorted by the total amount of allocated expenditure.
+With 2.7 bn~\euro\, the first value pushes investments extraordinarily high.
+Due to low renewable potentials, it is dominated by CAPEX for OCGT which receives 92\% of the payments.
 This hour alone occasions about the half of all OCGT CAPEX.
-\Cref{capex_duration} gives a detailed picture of the operational state at this time-step.
+\Cref{fig:capex_duration} gives a detailed picture of the operational state at this time-step.
 The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution.
@@
 These time-steps however play a minor role.}
-    \label{fig:capex_duration}
+    \label{fig:capex_duration_curve}
 \end{figure*}
@@
     \includegraphics[width=\linewidth]{de50bf/operation_high_expenditure}
-    \caption{Production, flow and consumption in the system at the hour with the highest expenditures.
+    \caption{Production, flow and consumption in the system at the hour with the highest allocated expenditures.
 The size of the circles are proportional to the power production at a node.
@@
     \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/branch_scarcity_cost_total}
-    \caption{Average \textbf{allocated transmission scarcity cost} per consumed MWh. This scarcity cost results from the upper transmission expansion limit of 25\%.}
+    \caption{Average \textbf{allocated transmission scarcity cost} per consumed MWh, $\sum_t \scarcitycost_{n \rightarrow \ell, t} / \sum_t \demand $.
+This scarcity cost results from the upper transmission expansion limit of 25\%.
+The costs are indicated by the regional color.
+The lines are drawn in proportion to revenue dedicated to scarcity cost.
+}
     \label{fig:branch_scarcity_price}
@@
     \end{subfigure}
-    \caption{Average electricity price for \textbf{allocated scarcity cost}, $\sum_t \allocatescarcitycost$, due to technology resource limits, for offshore wind, onshore wind,  solar, run-of-river.
-Average Allocated Scarcity Cost per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
+    \caption{Average \textbf{allocated scarcity cost} per consumed MWh, $\sum_t \allocatescarcitycost / \sum_t \demand$.
+These cost result from land use restricitions for offshore wind, onshore wind,  solar, run-of-river.
+The cost per MWh are indicated by the color of a region.
+The revenue per production asset is given by the size of the circle at the corresponding bus.}
     \label{fig:scarcity_price}

### 2020-09-09 code/plot_duration_curve.py

@@ -8,6 +8,7 @@ Created on Tue Sep  8 10:34:01 2020
 import pypsa
 import pandas as pd
 import matplotlib.pyplot as plt
+import config
 
 
 if 'snakemake' not in globals():
@@ -53,7 +54,7 @@ ax2.legend(handles[::-1], n.carriers.nice_name[duration.columns][::-1],
            ncol=3, frameon=False)
 
 
-ax1.set_ylabel('CAPEX [€]')
+ax1.set_ylabel('Allocated CAPEX [€]')
 ax2.set_xlabel('time [h]')
 ax1.set_yscale('log')
 ax1.set_ylim(bottom=10)

### 2020-09-09 paper/flow_based.tex
follow up

@@
 Together with the emission cost $\emissioncost$, the total OPEX $\opex$ amount around 16 billion \euro.
-As to expect, 99.97\% are dedicated to OCGT alone.
+As to expect, 99.97\% are dedicated to OCGT alone, as these have by far the highest operational price.
 For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price,fig:emission_cost}.
-Compared to the CAPEX allocation for OCGT, the OPEX distributes in a more local manner.
-This results from the fact that during ``standard'' demand peaks, it is rather local OCGT generators which serve as backup generators.
+For OCGT, the spatial allocation of the OPEX clearly differs from  the CAPEX allocation.
+This results from the fact that during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators.
 In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power.

### 2020-10-13 paper/flow_based.tex
write intro text (without citations); now change text for a better story

@@
 \newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
-\newcommand{\injection}{p_{n,t}}
+\newcommand{\injection}[1][n]{p_{#1,t}}
 \newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
@@
 \newcommand{\cycle}{C_{\ell,c}}
-\newcommand{\impedance}{x_\ell}
+\newcommand{\reactance}{x_\ell}
 \newcommand{\cycleprice}{\lambda_{c,t}}
@@
 \newcommand{\megawatthour}{MWh$_\text{el}$}
-\newcommand{\totalcost}{\mathcal{TC}}
+\newcommand{\totalcost}{\mathcal{T}}
 \newcommand{\cost}{\mathcal{C}}
@@
 
-\title{A dispatch-based approach to fully allocate costs in optimal power systems}
-\author{F. Hofmann, Brown T.}
+\title{Tracing prices: A flow-based cost allocation for optimized power systems}
+\author{F. Hofmann}
 
@@
 \begin{itemize}
- \item The LMP always breaks down into contributions of different cost terms (OPEX, CAPEX, etc.) of the system.
- \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments, where prominent methods as Average Participation or Flow Based Market Coupling can be applied.
- \item ...
- \item ...
+ \item Flow Allocation schemes determine peer-to-peer (P2P) allocation between sources and sinks 
+ \item P2P allocation can be used to assign payments of consumers to single network assets
+ \item Using shadow prices, Capital Expenditures are accurately allolcated to users which profit from investments
+ \item The P2P cost assignments are in alignment with the nodal pricing scheme of a welfare-maximized system   
+ \item This opens the possibility to transparently integrate transmission constraints and investment costs in national and international development plans 
+ \item Applying to a German future scenario shows how higher electricity costs arise for buses which rely on investments in transmission and local backup generation  
 \end{itemize}
-
-% \section{Introduction}
 
@@
 
+\section{Introduction}
+
+Today's power systems are subject to a deep and ongoing transformation.
+The needed shift to from controllable to weather-driven power generation, the constant improvement and innovation of technology require rigorous system planning and international cooperation.
+The core of the problem manifests in the costs.
+Firstly, these should be as small as possible while meeting ecological and sociological standards.
+Secondly, they must be distributed in a fair and transparent manner.
+Cost-driving aspects or market players should be detected and addressed appropriately.
+In this context, power system modelling has increasingly gained attention during the last years.
+While many studies show how cost-efficiency and renewable energy are brought together, the question of how and, more importantly, upon what grounds costs are distributed was often left open.
+\\
+This study fills this gap.
+On the basis of an optimized network the operational state of each time step and corresponding characteristics are considered to allocate all system costs.
+Therefore, two basic concepts are used: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
+Secondly, \textit{flow tracing}, following Bialek's Average Participation, assigns the operation of single network assets to consumers in an plausible manner.
+The following shows how combining these allow for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the end consumers.
+
 \section{Economic Context}
@@
 In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms.
-Typically, these include operational expenditures (OPEX) $\opexgeneration$ for generators, expenditures for emissions $\emissioncost$, capital expenditures (CAPEX) for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and possible other terms, \ie
-\begin{align}
-\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration +  \capexflow + ...
+Typically, these include OPEX for generators $\opexgeneration$, expenditures for emissions $\emissioncost$, CAPEX for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and so on, \ie
+\begin{align}
+\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration + ...
 \label{eq:total_cost}
 \end{align}
-In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
+In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, ...\}$ consists of cost associated to the asset $i$ in the system, 
 \begin{align}
@@
 
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$.
-It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
+Hence, it is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
 \begin{align}
@@
 \end{figure}
+From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms.
+This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
 % 
@@
 % 
-Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
+Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
 However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
 
-From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms.
-This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
 However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
@@
 
-Let $\cost_{i}$ denote a general cost term associated with asset $i$.
+Let $\cost_{i}$ denote an arbitrary cost term associated with asset $i$.
 Consider a long-term equilibrium in a power system with perfect competition, then, according to the zero-profit condition, each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$, \ie
@@
 % where $\costfactor$ is either a fixed operational price or a composition of shadow prices
-Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$.
-This leads us to
+Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
 \begin{align}
@@
 \end{align}
-The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. 
-\begin{align}
-    \allocatestate \rightarrow\pdv{\state}{\demand} \demand
+The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. However, we replace this quantity by 
+\begin{align}
+    % \pdv{\state}{\demand} \demand \rightarrow \int \pdv{\state}{\demand}  \,\mathrm{d}\demand  \rightarrow \allocatestate
+    \pdv{\state}{\demand} \demand \Leftrightarrow  \allocatestate
     \label{eq:state_allocation}
 \end{align}    
-...
-may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with.
-It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$.
-From the natural fact that the sum of all contributions must return the cost term, 
+where $\allocatestate$ represents the amount of power that  asset $i$ supplies demand $\demand$ with.
+As we will show later, it is calculated from established flow allocation schemes.
+By following this approach, the differential quantity on the left side, is replaced by an integral quantity.
+The latter does not only rely on a marginal sensitivity of the nodal demand $\demand$, but makes further assumptions on how the operational state $\state$, all in all, serves consumers in the network.
+The dispatch allocation $\allocatestate$ ensures that all power produced or processed by asset $i$ at time $t$  is assigned to consumers,
+\begin{align}
+    \state = \sum_n \allocatestate,
+    \label{eq:state_allocation_constraint}
+\end{align}
+from which follows that the sum of all cost contributions return the total cost term, 
 \begin{align}
@@
 \end{align}
-it follows that $\allocatestate$ must fulfill
-\begin{align}
-    \state = \sum_n \allocatestate
-    \label{eq:state_allocation_constraint}
-\end{align}
-The last equation states that all power produced or processed by asset $i$ must be totally consumed by the network demand $\demand$.  
-Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
-\begin{align}
-    \cost_{n,t} = \sum_i \allocatecost
-\end{align}
+% Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
+% \begin{align}
+%     \cost_{n,t} = \sum_i \allocatecost
+% \end{align}
+
 \end{subequations}
+
+By combining \cref{eq:cost_allocation,eq:state_allocation}, we are able to formulate a full cost allocation from demand to network assets for every time-step $t$.
+In particular, the definition of cost factors $\costfactor$ must be derived for each cost term individually.
+In the following we depict the presented scheme for typical classes of assets.
+Therefore, we assume a network with generators $s$, transmission lines $\ell$ and storage units $r$, each asset $i = \{s, \ell, r\}$ adds a term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
 
@@
         \midrule
+        Emission Cost & $s$ & $\emissioncost$ & $ \emissionprice \, \emission \, \generation$ & $\emissionprice \,\emission$ & $\generation$ \\
+        \midrule   
         CAPEX Production & $s$ & $\capexgeneration$ & $ \capitalpricegeneration \capacitygeneration$ & $\muuppergeneration$ & $\generation$ \\
@@
         CAPEX Storage & $r$ & $\capexstorage$ & $ \capitalpricestorage \capacitystorage$ & $ \muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge $ & $\storage$ \\
-        \midrule
-        Emission Cost & $s$ & $\emissioncost$ & $ \emissionprice \, \emission \, \generation$ & $\emissionprice \,\emission$ & $\generation$ \\
         % \bottomrule   
@@
 
-Now, assume a network with generators $s$, transmission lines $\ell$ and storage units $r$.
-Each asset $i = \{s, \ell, r\}$ adds an term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
-
 \subsection{OPEX Allocation}
 
-Let the operational price for an asset $i$ be given by $o_i$. Then, for example the OPEX occasioned by generator $s$ is given by 
+Let the operational price for an asset $i$ be given by $o_i$. Then, the OPEX occasioned, for example, by generator $s$ is given by 
 \begin{align}
@@
 where $\generation$ denotes its power generation at time $t$.
-As \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme}.
+\Cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme}.
 As a result we obtain
@@
 \begin{align}
- \allocategeneration = \pdv{\generation}{\demand} \, \demand
+ \allocategeneration \Leftrightarrow \pdv{\generation}{\demand} \, \demand
  \label{eq:allocate_peer}
 \end{align}
-can be considered as the power that is produced by generator $s$ and consumed at node $n$ at time $t$. \\
-In the same manner, we can follow the scheme allocate OPEX for flow $\flow$ and storage unit dispatch $\storagedispatch$.
+superseding the derivative on the right side, describes the power that is produced by generator $s$ and consumed at node $n$ at time $t$. \\
+In the same manner, we can follow the scheme to allocate OPEX for flow $\flow$ and storage unit dispatch $\storagedispatch$.
 As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow.
@@
 \subsection{CAPEX Allocation}
-For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and the capacity limit.
-For all assets, let the capital price for one unit capacity expansion be denoted by $c_i$.
-All quantities for the CAPEX allocation, which we now discuss in detail, are summarized in the middle section of \cref{tab:cost_allocation_map}.
+For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and its binding constraints.
+For all assets, let the capital price for one MW capacity expansion be denoted by $c_i$.
+All quantities for the CAPEX allocation, which we now discuss in detail, are summarized in the lower section of \cref{tab:cost_allocation_map}.
 
@@
 \end{align}
-This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition} allows for using the cost allocation scheme.
-The resulting the CAPEX allocation is given by
+This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}. The resulting the CAPEX allocation is given by
 \begin{align}
@@
 According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding.
-Thus it is these times steps which push investments in $\capacitygeneration$.
+If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding, which pushes investments in $\capacitygeneration$.
 If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
@@
 which yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow
+Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which shows that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow
 \begin{align}
@@
 \begin{align}
- \allocateflow =  \pdv{\flow}{\demand}\, \demand
-\end{align}
-can be interpreted as the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
+ \allocateflow \Leftrightarrow  \pdv{\flow}{\demand}\, \demand
+\end{align}
+represents the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
 % According to \cref{eq:state_allocation_constraint} it has to fulfill 
@@
 where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint.
-When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand = 0$, implying that the when a storage charges power, it does not supply any demand.
-Rather it stands with the demand on the consumer side, retrieving power from producing assets.
+When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand \Leftrightarrow 0$, implying that the when a storage charges power, it does not supply any demand.
+At these times, it the storage is a effective consumer retrieving power from producing assets.
 This leaves us with 
@@
 \end{align}
-and the power allocation 
-\begin{align}
-    \allocatestoragedispatch = \pdv{\storagedispatch}{\demand} \demand
-\end{align}
-The latter only allocates dispatched power of storage $r$. Note that this will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
+where the dispatch allocation only assigns discharge power to demand $\demand$, thus
+\begin{align}
+    \allocatestoragedispatch \Leftrightarrow \pdv{\storagedispatch}{\demand} \demand
+\end{align}
+Note that this relation will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
 % \begin{align}
@@
 % \end{align} 
-It is crucial to note that like this, storage units perform a redistribution of money, and therefore distort the cost allocation.
-So, certain share of what is allocated to the CAPEX of a storage is in another time step spent by the storage in order to buy power from other assets.
-This effect scales with the amount of installed capacity.
-
-It is possible to incorporate this redistribution effect into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
+It is crucial to note that storage units perform a redistribution of money, and therefore distort the cost allocation.
+So, a certain share of what is allocated to the CAPEX of a storage is in another time step spent by the same storage in order to buy power from other assets.
+This effect scales with the amount of installed storage capacity.
+
+We shortly annotate that it is possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
 Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$.
@@
 \end{align}
-where $K$ is any non-zero constant, will result in a cost term contributing to $\remainingcost$ and in some cases alter \cref{eq:cost_decomposition} to 
+where $K$ is any non-zero constant and $\capacity$ is collects all capacities of asset $i$, will result in a cost term contributing to $\remainingcost$ and alter \cref{eq:cost_decomposition} to 
 \begin{align}
@@
 \end{align}
-The effective revenue per asset $i$ associated with cost term $\cost$ can then be defined as 
+The total allocatable cost associated to cost term $\cost$ is then given by 
 \begin{align}
@@
 \end{align}
-Note that $\cost'$ is the amount which is allocated to consumers which, according to the quality of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, can either be larger, equal or lower than $\cost$.
+According to the nature of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, it is either larger, equal or lower than $\cost$.  
 In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to incorporate them into the cost allocation.
@@
 
-The costs which consumers at $n$ have to pay for the scarcity impacting asset $i$ are given by 
+The share of the CAPEX allocation, which results from the scarcity, is given by 
 \begin{align}
@@
 The presented cost allocation suits for any type of topology and network setup.
-But so far, the question of how $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.
-We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:state_allocation_constraint}).
+It may be extended by other classes of network assets or design constraints.
+But so far, the question of how the allocation $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.
+
+In a first step, we show that the needed information narrows down to peer-to-peer allocations between producing assets and consumers.
+%We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:state_allocation_constraint}).
 \\
-
-Let $\nodalgeneration[m]$ denote the nodal power generation which combines the power production of all producing assets, in this case generators $S$ and storages $R$, at node $n$ and time $t$.
-It is given by
-\begin{align}
-    \nodalgeneration[m] = \sum_{i \in \{S, R\}} \incidenceasset[m] \, \state
-\end{align}
-with $\incidenceasset[m]$ being 1 if asset $i$ is attached to bus $m$ and zero otherwise. Further, let $\allocatepeer$ collect the power produced by assets at node $m$ and consumed at $n$, given by 
+Let $\injection[m]$ denote the net power injection at node $m$ and time $t$. It is given by 
+\begin{align}
+    \injection[m] = \sum_{i \in \{s, r\}} \incidenceasset[m] \, \state - \demand[m]
+\end{align}
+where $\incidenceasset[m]$ is 1 if asset $i$ is attached to bus $m$ and zero otherwise. Now, let $\ptdf$ denote a linear mapping between the injection $\injection[m]$ and the flow $\flow$, such that  
+\begin{align}
+ \flow  = \sum_m \ptdf[m]\, \injection[m]
+ \label{eq:power_flow_equation}  
+\end{align}
+Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
+For transport models or mixed AC-DC networks, these can calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
+
+Using the two expression the derivative of the flow $\flow$ with respect to the demand can be expressed through 
+\begin{align}
+    \pdv{\flow}{\demand} \demand \Leftrightarrow  \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
+ \label{eq:allocate_peer_to_allocate_flow},
+\end{align}
+where we collected all power produced at bus $m$ and consumed at bus $n$ in $\allocatepeer$, defined as
 \begin{align}
@@
 \end{align}
-Now, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.
-The linear power flow equation can be written as
-\begin{align}
- \flow  = \sum_m \ptdf[m] \left( \nodalgeneration[m] - \demand[m] \right)
- \label{eq:power_flow_equation}  
-\end{align}
-Note that for transport models or mixed AC-DC networks, $\ptdf$ can be artificially calculated using the formulation presented in \cite{hofmann_flow_2020-1}.
-Taking the derivative with respect to the demand, 
-\begin{align}
- \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
- \label{eq:allocate_peer_to_allocate_flow},
-\end{align}
-shows that $\allocateflow$ is fully determined through the peer-to-peer allocation $\allocatepeer$.
-In other words, we only need to know how much power produced at node $m$ is consumed at node $n$ in order to derive the allocated flow $\allocateflow$.
-Further we can breakdown $\allocatepeer$ to $\allocategeneration$ for generators and $\allocatestoragedispatch$ for storages proportionally to their contribution to the nodal generation $\nodalgeneration[m]$.
+\Cref{eq:allocate_peer_to_allocate_flow} shows that $\allocateflow$ is fully determined through the peer-to-peer allocation $\allocatepeer$.
+In other words, we only need to know how much of the power produced at node $m$ is consumed at node $n$.
+With this information we may also derive the cost allocation for the transmission system.
 Unfortunately, the solution for $\allocatepeer$ is non-unique and requires further assumptions.
@@
 % 
-Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $m$ with $\nodalgeneration[m] \le \nodaldemand[m]$ is not assigned to other buses, only to $m$. 
-Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
-The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal.
-Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$.
-Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first step.
+Principally, allocating net power instead of gross power leads to less P2P trades, as power from a bus $m$ with $\nodalgeneration[m] \le \nodaldemand[m]$ is not assigned to other buses, only to $m$.
+
+When it comes to the methods, as literature has often pointed out, the EBE principal does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
+The AP allocation tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal.
+Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when following in the direction of $\flow$, ends up at bus $n$.
+In contrast, the FBMC allows to control the regions or market zones which are netted out in a first step.
 If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.
-However, it relies on further assumptions such as the Generation Shift Keys determining the production which are deployed for the inter-regional exchange.
-
-From this point of view, we restrict the focus of this research to \textbf{net} injection allocation on the basis of \textbf{AP}, type \ref{net}\ref{ap}, as locality and no further dependency on external decisions are strong arguments for a transparent cost allocation.
-We emphasize that the AP based allocation does not imply traceability of the resulting flows, as only the \textit{source}-\textit{sink} relations $\allocatepeer$ are determined by it, the flow allocation still follows the power plow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
-In the following example, we illustrate the functionality of the cost allocation by an example of a small artificial network.
-
-\subsection{Numerical Example}
-\label{sec:numerical_example}
-
+However, it relies on further assumptions such as the Generation Shift Keys determining the producing assets which are deployed for the inter-regional exchange.
+% 
 \begin{figure*}[t]
@@
 %
-% 
-% 
-Consider a two bus system, as shown in \cref{fig:example_network}, with one transmission line and one generator per bus.
-Whereas generator~1 (at bus~1) has a cheap operational price of 50 \euro/\megawatthour, an expensive operational price of 200~\euro/\megawatthour is set to  generator~2 (at bus~2).
-For both the capital investments amount 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW.
+
+From this point of view, we restrict the focus of this research to \textbf{net} injection allocation on the basis of \textbf{AP}, type \ref{net}\ref{ap}, as locality and no further dependency on external decisions are strong arguments for a transparent cost allocation.
+We emphasize that the AP based allocation does not imply traceability of the resulting flows, as it only determines the \textit{source}-\textit{sink} relations $\allocatepeer$.
+The flow allocation still follows the power flow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
+In the following example, we illustrate the functionality of the cost allocation by an example of a small artificial network.
+% % 
+% 
+\subsection{Numerical Example}
+\label{sec:numerical_example}
+% 
+Consider a two bus system, depicted in \cref{fig:example_network}, with one transmission line and one generator per bus.
+Generator~1 (at bus~1) has a cheap operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a expensive operational price of 200~\euro/\megawatthour.
+For both, capital investments amount 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW.
 The transmission line has a capital price of 100~\euro/MW and no upper capacity limit.
@@
 Note, the ``sum'' of the two figures give to the actual dispatch.
-The corresponding P2P payments are given in \cref{fig:example_payoff}.
+The resulting P2P payments are given in \cref{fig:example_payoff}.
 
@@
 Note that 3k~\euro~ of these account for the scarcity imposed buy the upper expansion limit $\capacitygenerationupper$.
-The rest makes up 60\% of the CAPEX spent at generator~1, exactly the share of power allocated to $d_1$.
+The rest makes up 60\% of the total CAPEX spent at generator~1, exactly the share of power allocated to $d_1$.
 Consumers at bus~1 don't pay any transmission CAPEX as no flow is associated with there demand.
@@
 We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from bus~1.
-Thus, consumers have the pay for the local OPEX and CAPEX as well as the expenditures at bus~1 and the transmission system.
+Thus, consumers have the pay for the local OPEX and CAPEX as well as the proportional share of expenditures at bus~1 and the transmission system.
 As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it.
-The allocated CAPEX exactly matches the actual CAPEX of generator~2.
-However in the case of generator~1, 2\kk of the 22\kk CAPEX allocation are associated to the scarcity cost.
-The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line (zero-profit condition).
+The allocated CAPEX $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the actual CAPEX of generator~2.
+Again.
+in the case of generator~1, 2\kk of the 22\kk CAPEX allocation are associated with the scarcity cost for generator~1.
+The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line, which directly reflects the zero-profit condition.
 % 
@@
 
-The fact that OPEX and CAPEX allocations are proportional to each other results from the fact of optimizing one time step only.
-This coherence breaks for for larger optimization problems with multiple time steps.
+The fact that OPEX and CAPEX allocations are proportional to each other results from optimizing one time step only.
+This coherence breaks for larger optimization problems with multiple time steps.
 Then CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
 \\
-The example shows that allocating net power injection leads to intuitive and transparent  cost allocations.
+% The example shows that allocating net power injection leads to intuitive and transparent  cost allocations.
 
@@
 \section{Application Case}
-
-We showcase the behavior of the cost allocation in a more complex system, by applying it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
-The model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
-
-We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
-Pre-installed wind and solar generation capacity totals of the year 2017 were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
-Further, wind and solar capacity expansion are limited by land use restriction.
-These consider agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{corine2012,natura2000}.
-Pumped Hydro Storages (PHS) and Run-of-River power plants are fixed to today's capacities with no more expansion allowed.
-Additionally, unlimited expansion of batteries and H$_{2}$-storages and Open-Cycle Gas Turbines (OCGT) are allowed at each node.
 % 
@@
 % 
+We showcase the behavior of the cost allocation in a more complex system, by applying it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
+The model builds up on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
+
+We follow a brownfield approach where transmission lines can be expanded starting from today's capacity values, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}.
+Pre-installed wind and solar generation capacity totals of the year 2017 were distributed in proportion to the average power potential at each site excluding those with an average capacity factor of 10\%.
+Further, wind and solar capacity expansion are limited by land use restriction.
+These consider agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{corine2012,natura2000}.
+Pumped Hydro Storages (PHS) and Run-of-River power plants are fixed to today's capacities with no more expansion allowed.
+Additionally, unlimited expansion of batteries and H$_{2}$-storages and Open-Cycle Gas Turbines (OCGT) are allowed at each node.
+% 
+% 
 \begin{figure}
@@
 The optimized network is shown in  \cref{fig:network}.
-On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission.
+On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission are displayed.
 The optimization expands solar capacities in the south, onshore and offshore wind in the upper north and most west.
@@
 The largest proportion of the payments is associated with CAPEX for generators, transmission system and storage units in decreasing order. 
-Taking the different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shows in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
+Taking different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shown in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
 At this time (morning, end of February), the system hits the highest mismatch between low renewable power potential and high demand.
@@
 With a few exceptions in the South West and East, all consumers receive power from OCGT at this time, thus all pay high amounts for the needed backup infrastructure (operational state of the system is in detailed shown in \cref{fig:operation_high_expenditure}).
-Note that this is the most extreme event, which ensures backup infrastructure for other inferior extreme events where backup power is needed.
-The total CAPEX allocation for OCGT infrastructure is depicted in \cref{fig:ocgt_capex} which correlates with out findings from the extreme event.
+Note that this is the most extreme event, which ensures backup infrastructure for other inferior extreme events.
+The total CAPEX allocation for OCGT infrastructure is depicted in \cref{fig:ocgt_capex}.
+This  correlates with our findings of the extreme event.
 \\
@@
 As for onshore and offshore wind farms, the produced power deeply penetrates the network, see \cref{fig:power_mix}, thus it is not only local, but also remote consumers which cover the CAPEX, see \cref{fig:capex_price}.
-This in turn benefits local consumers, which profit from the cheap power which explains why these regions end up with a low average LMP.
+This in turn benefits local consumers, which profit from the cheap power which explains why these regions end up with a low average LMP.\\
 
@@
 The OPEX allocations for renewables play with 0.2 \euro/MWh in average an inferior role.
-
-The negative segment in \cref{fig:total_cost} is associated with the scarcity costs $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit.
-These sum up approximately to 7.5 b\euro.
-Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments.
-It translates to the cost that consumers pay ``too much'' for assets limited in their capacity expansion.
-In the real world this money would be used to pay for augmented land costs or civic participation in the dedicated areas.
-In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average price $\left<\lmp\right>_t$.
-The scarcity for wind and solar are relatively low and impact the average price by around 2\euro/MWh.
-They mostly affect regions with strong investments and those in close vicinity.
-Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16\euro.
-This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
-However, as these power plants in particular are already amortized and scarcity cost should be reconsidered and removed from a final cost allocation.
-The scarcity cost for the transmission system, resulting from the limited transmission volume expansion of 25\% mostly influence regions in the middle of Germany, where in average 8\euro/MWh account for transmission scarcity, see \cref{fig:branch_scarcity_price}.
-This illustrates the strong impact that the line volume constraint has on the optimization.
-Regions in the North are not touched by the constraint.
-
-The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
-These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$.
+\\
 
@@
     \centering
-    \includegraphics[width=\linewidth]{de50bf/locality_ptpf_net}
+    \includegraphics[width=\linewidth]{de50bf/locality_all_ptpf_net}
     \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
@@
 \end{figure}
+
+The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit.
+These sum up approximately to 7.5 b\euro.
+Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments.
+It translates to the cost that consumers pay ``too much'' for assets limited in their capacity expansion.
+In the real world this money would be spent for augmented land costs or civic participation in the dedicated areas.
+In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average cost per consumed MWh.
+The scarcity for wind and solar is relatively low and impact the average price at roughly 2\euro/MWh.
+It primarily affects, different to the pure CAPEX allocation, regions in close vicinity.
+Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16\euro.
+This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
+However, as these power plants in particular are already amortized, the scarcity cost should be reconsidered and removed from a final cost allocation.
+
+A high influence on the price comes from the scarcity of transmission expansion.
+Right beside regions with high wind infrastructure, it occasions costs higher than MWh 4\euro/MWh (maximal 8\euro/MWh), see \cref{fig:branch_scarcity_price}.
+As to expect, the constraint mainly suppresses transmission expansion along the north-south axis.
+
+The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
+These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$.
+Most of these ``non-allocatable'' costs account for Pumped Hydro Storage, onshore wind, offshore wind and the transmission system, see \cref{fig:subsidy} for further details.
+Again, as most of the named assets are already amortized (PHS, transmission system), the subsidy should be reconsidered in a final cost allocation.
+
+In \cref{fig:direct-allocation} we compare the total cost allocation of the region with the lowest average LMP (left side) against the one of the region with the highest LMP (right side).
+The region with the cheapest electricity is in the Northwest.
+We see that this bus is fairly independent of investments in the transmission system, as it is supplied by steady, cheap offshore power.
+Moreover, the CAPEX of the local offshore wind farms are partly payed by subsidies, see \cref{fig:subsidy}.
+Only a small share of the payments is allocated to remote OCGTs, resulting from the extreme event mentioned earlier.
+
+In contrast the regions with the highest LMP, spends most its money to local OCGT operations, emission cost and the transmission system.
+Its payments to  onshore and offshore wind infrastructure are low despite a third of its supply comes from wind power.
+Instead, a lot of its payments are assigned to the transmission infrastructure.
+Hence, the wind power supply at this region is not restricted by exhausted wind power resources but by bottlenecks in the transmission system.
 
@@
 For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
-
+\dots
 \clearpage
@@
 % \notag
-% &+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
+% &+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \reactance \, \flow  \\
 % \notag
@@
 \begin{align}
-    \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow 
-\end{align}
-can be added to $\lagrangian$, with $\impedance$ denoting the line's impedance and $\cycle$ being 1 if $\ell$ is part of the cycle $c$ and zero otherwise.
+    \sum_{\ell,c,t} \cycleprice \, \cycle \, \reactance \, \flow 
+\end{align}
+can be added to $\lagrangian$, with $\reactance$ denoting the line's impedance and $\cycle$ being 1 if $\ell$ is part of the cycle $c$ and zero otherwise.
 
@@
     0 &= \pdv{\lagrangian}{\flow}  \\ 
-    0 &= - \sum_n \incidence \lmp  + \cycleprice \cycle \impedance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
+    0 &= - \sum_n \incidence \lmp  + \cycleprice \cycle \reactance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
 \end{align}
@@
 \begin{align}
-\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \impedance\, \flow 
+\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \reactance\, \flow 
 \end{align}
@@
 This hour alone occasions about the half of all OCGT CAPEX.
-\Cref{fig:capex_duration} gives a detailed picture of the operational state at this time-step.
+\Cref{fig:capex_duration_curve} gives a detailed picture of the operational state at this time-step.
 The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution.
@@
 Size of arrows are proportional to the flow on the transmission line.
-The depicted hour corresponds to the first value in the duration curve in \cref{fig:capex_duration}.}
+The depicted hour corresponds to the first value in the duration curve in \cref{fig:capex_duration_curve}.}
     \label{fig:operation_high_expenditure}

### 2020-10-13 lp-based/constraint_matrix.py

@@ -1,61 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Jun 24 11:00:12 2020
-
-@author: fabian
-"""
-
-import pypsa
-import numpy as np
-import pandas as pd
-from helpers import get_linear_system, noisy_lopf
-import networks
-
-n = networks.n_ac_dc()
-em = n.generators.carrier.map(n.carriers.co2_emissions) / \
-        n.generators.efficiency
-noisy_lopf(n)
-s = get_linear_system(n)
-A_, A_inv, c, x, d, m, r, C = (s[k].round(5) for k in
-                            ['A_', 'A_inv', 'c', 'x', 'd', 'm', 'r', 'C'])
-# C = C.loc[:, d.round(1)!=0]
-
-
-sn = n.snapshots[0]
-bus = 'London'
-
-gen = n.generators_t.p.T.copy()
-TC = n.objective
-
-print(n.generators_t.p @ em)
-print(m)
-
-
-mu_co2 = n.global_constraints.mu.co2_limit
-n.global_constraints = n.global_constraints.drop('co2_limit')
-
-n.generators['marginal_cost'] += em * mu_co2
-
-n.lopf(pyomo=False, keep_shadowprices=True, keep_references=True,
-        solver_name='gurobi')
-s = get_linear_system(n)
-A_, A_inv, c, x, d, m, r, C = (s[k].round(5) for k in
-                            ['A_', 'A_inv', 'c', 'x', 'd', 'm', 'r', 'C'])
-
-
-# print(n.generators_t.p @ em)
-
-
-# n.loads_t.p_set.loc[sn,bus] += 1
-
-# n.lopf(pyomo=False, keep_shadowprices=True, keep_references=True,
-#         solver_name='gurobi')
-# print(n.generators_t.p.T - gen)
-# print(n.objective - TC)
-
-
-# n.global_constraints = n.global_constraints.drop('co2_limit')
-
-# print('Objective diff:', objective - n.objective)
-# print('CO2 cost diff:', total_co2_cost)

### 2020-10-13 lp-based/incremental_change.py

@@ -1,21 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Jul  2 15:02:28 2020
-
-@author: fabian
-"""
-
-import networks
-
-n = networks.n_ac_dc()
-
-n.lopf()
-print(n.generators_t.p)
-print(n.objective)
-
-n.loads_t.p_set.iloc[0,0] += 1
-
-n.lopf()
-print(n.generators_t.p)
-print(n.objective)

### 2020-10-13 lp-based/networks.py

@@ -1,98 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jun 29 14:24:39 2020
-
-@author: fabian
-"""
-
-import pypsa
-import pandas as pd
-
-def n1_t1_g2_wo(d = 70, marginal_cost=[5, 10], p_nom=40):
-    """
-    $G_1 < d < G_1 + G_2$
-
-    Note, for such a system the total system cost $TC$ are less then the nodal
-    payments as soon as one generator is at its limit:
-
-    $\lambda = o_s + \bar{\mu_s} \;\;\;\;\; \forall s$
-
-    $d \, \lambda = d \, (o_s + \bar{\mu_s} ) \ge \sum_s g_s \, o_s = TC$
-    """
-    n = pypsa.Network()
-    n.add('Bus', 'Bus1')
-    n.add('Load', 0, bus='Bus1', p_set=d)
-    n.madd('Generator', ['Gen0', 'Gen1'], bus='Bus1',
-           marginal_cost=marginal_cost, p_nom=p_nom)
-    return n
-
-
-
-def n1_t1_g2_w(d = 70, marginal_cost=[5, 10], capital_cost=[100, 30]):
-    """
-    Here the total cost are payed by the consumers
-
-    $TC = \lambda d$
-    """
-    n = pypsa.Network()
-    n.add('Bus', 'Bus1')
-    n.add('Load', 0, bus='Bus1', p_set=d)
-    n.madd('Generator', ['Gen0', 'Gen1'], bus='Bus1', p_nom_extendable=True,
-           marginal_cost=marginal_cost, capital_cost=capital_cost)
-    return n
-
-
-def n1_t9_g2_w(d = 70, marginal_cost=[2, 4.], capital_cost=[50, 30]):
-    """Initialize network with base and one peak demand at one bus."""
-    load = pd.DataFrame({0: [50, 49, 48, 47, 46, 90, 46.5, 47.5, 48.5]})
-
-    n = pypsa.Network()
-    n.set_snapshots(list(range(len(load))))
-    n.add('Bus', 'Bus1')
-    n.madd('Load', [0], bus='Bus1', p_set=load)
-    n.madd('Generator', ['Gen0', 'Gen1'], bus='Bus1', p_nom_extendable=True,
-           marginal_cost=marginal_cost, capital_cost=capital_cost)
-    return n
-
-
[…]

### 2020-10-13 lp-based/tinylp.py

@@ -1,96 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Jul  8 12:37:18 2020
-
-@author: fabian
-"""
-
-import gurobi
-from numpy import array
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-plt.rc('text', usetex=True)
-plt.rc('font', family='serif')
-
-
-model = gurobi.read('tinylp.lp')
-model.optimize()
-
-O = model.getObjective()
-A = model.getA().T.todense() # first dimension {vars} second {cons}
-x = array([v.x for v in model.getVars()])
-d = array([c.RHS  for c in model.getConstrs()])
-c = np.zeros_like(x)
-for t in range(O.size()): c[O.getVar(t).index] = O.getCoeff(t)
-m = array([c.Pi for c in model.getConstrs()])
-
-
-model2 = gurobi.read('tinylp2.lp')
-model2.optimize()
-
-O2 = model2.getObjective()
-A2 = model2.getA().T.todense() # first dimension {vars} second {cons}
-x2 = array([v.x for v in model2.getVars()])
-d2 = array([c.RHS  for c in model2.getConstrs()])
-c2 = np.zeros_like(x)
-for t in range(O2.size()): c2[O2.getVar(t).index] = O2.getCoeff(t)
-m2 = array([c.Pi for c in model2.getConstrs()])
-
-print(x, x2)
-
-
-#%% plot
-N = 10
-fig, (ax1, ax2) = plt.subplots(1,2, sharey=True, figsize=(8,5))
-# ax1.axis('off')
-ax1.set_aspect('equal', adjustable='box')
-ax1.set_xlim(0,N-1)
-# ax2.axis('off')
-ax2.set_xlim(0,N-1)
-ax2.set_aspect('equal', adjustable='box')
-
-f = r'$2 x_0+4 x_1$'
-con1 = r'$ x_1 + x_2 = 7 $'
-con2 = r'$x_1 + 0.5 x_2 \le 4$'
-mu2 = r'$\perp \mu = -4 $'
-fmod = r'$5 x_0 + 5  x_1$'
-
-
-carray = pd.DataFrame([[c @ np.array([x1,x2]) for x1 in range(N)]
-                       for x2 in range(N)])
-xx = np.linspace(0,N,200)
[…]

### 2020-10-14 paper/flow_based.tex
strong rearrangement! but look okay (next cost allocation)

@@
 \newcommand{\capacity}{S_{i}}
-\newcommand{\costfactor}{\gamma_{i,t}}
+\newcommand{\costfactor}{\gamma^\circ_{i,t}}
 \newcommand{\capacityupper}{\bar{S}}
@@
 \newcommand{\totalcost}{\mathcal{T}}
-\newcommand{\cost}{\mathcal{C}}
+\newcommand{\cost}[1][\circ]{\mathcal{C}^{#1}}
 \newcommand{\opex}{\mathcal{O}}
@@
 \newcommand{\opexstorage}{\mathcal{O}^E}
-\newcommand{\capexgeneration}{\mathcal{C}^G}
-\newcommand{\capexflow}{\mathcal{C}^F}
-\newcommand{\capexstorage}{\mathcal{C}^E}
+\newcommand{\capexgeneration}{\mathcal{I}^G}
+\newcommand{\capexflow}{\mathcal{I}^F} 
+\newcommand{\capexstorage}{\mathcal{I}^E}
 \newcommand{\emissioncost}{\mathcal{E}}
@@
         $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
-        $\generation$ & Electric generation of generator $s$, \\ & time step $t$  in MW \vpad \\
-        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
-        $\operationalpricegeneration$ & Operational price in \euro/MW \vpad \\
-        $\capitalpricegeneration$ & Capital Price in \euro/MW \vpad \\
-        $\capitalpriceflow$  & Capital Price in \euro/MW for transmission \\ & capacity on line $\ell$  \vpad \\
-        $\capacitygeneration$ & Generation capacity in MW \vpad \\
-        $\capacityflow$ & Transmission capacity in MW \vpad \\
-        $\incidence$ & Incidence matrix \vpad 
+        $\state$ & Operational state of asset $i$, \\ & at time step $t$  in MW \vpad \\
+        % $\generation$ & Electric generation of generator $s$, \\ & time step $t$  in MW \vpad \\
+        % $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
+        $o_i$ & Operational price of asset $i$ in \euro/MW \vpad \\
+        $c_i$ & Capital Price of asset $i$ in \euro/MW \vpad \\
+        % $\capacitygeneration$ & Generation capacity in MW \vpad \\
+        % $\capacityflow$ & Transmission capacity in MW \vpad \\
+        $\incidenceasset$ & Incidence values ($\ne$0  if $i$ is attached to $n$)  \vpad 
     \end{tabular}
@@
 In this context, power system modelling has increasingly gained attention during the last years.
-While many studies show how cost-efficiency and renewable energy are brought together, the question of how and, more importantly, upon what grounds costs are distributed was often left open.
+Many studies show how cost-efficiency and renewable energy are brought together.
+However, the question of how and, more importantly, upon what grounds costs are distributed was often left open.
 \\
 This study fills this gap.
-On the basis of an optimized network the operational state of each time step and corresponding characteristics are considered to allocate all system costs.
-Therefore, two basic concepts are used: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
+On the basis of an optimized network the operational state of each time step and corresponding characteristics are are in detailed considered to allocate all system costs.
+The presented approach builds upon two basic concepts: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
 Secondly, \textit{flow tracing}, following Bialek's Average Participation, assigns the operation of single network assets to consumers in an plausible manner.
@@
 
-\section{Economic Context}
-
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms.
-Typically, these include OPEX for generators $\opexgeneration$, expenditures for emissions $\emissioncost$, CAPEX for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and so on, \ie
-\begin{align}
-\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration + ...
+\section{Flow-based Cost Allocation}
+
+\subsection{Zero-Profit Rule}
+
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms $\cost$.
+Typically, these include operational expenditures for generators $\opexgeneration$, expenditures for emissions $\emissioncost$,  capital investments for the transmission system $\capexflow$ and so on, \ie
+\begin{align}
+\totalcost &= \sum_\circ \cost = \opexgeneration + \emissioncost +  \capexflow + ...
 \label{eq:total_cost}
 \end{align}
-In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, ...\}$ consists of cost associated to the asset $i$ in the system, 
+In turn, each of the terms $\cost$ consists of the costs associated to an asset $i$ in the system, 
 \begin{align}
@@
 \end{align} 
-where an asset describes any operating part of the network, such as a generator, line, energy storage etc.
-%For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.
-
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
-Hence, it is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
-\begin{align}
-\lmp = \pdv{\,\totalcost}{\,\demand}
-\label{eq:lmp}
-\end{align}
-% 
-\begin{figure}[h]
-\centering
-\includegraphics[width=.8\linewidth]{price_decomposition.png}
-\caption{Schematic decomposition of the Locational Market Price $\lmp$.
-In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \ie OPEX and CAPEX for production and transmission and possibly other expenditures.
-}
-\label{fig:price_decomposition}
-\end{figure}
-From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms.
-This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
-% 
-This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
-\begin{align}
-\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
-\label{eq:total_revenue}
-\end{align}
-depending on the costs $\remainingcost$ which are independent of the nodal demand  
-\begin{align}
- \pdv{\remainingcost}{\demand} = 0
-\end{align}
-% 
-Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
-However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
-
-However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
-This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.
-
-\section{Dispatch-Based Cost Allocation}
-\label{sec:theory}
-\begin{subequations}\label[subequations]{eq:general_scheme}
-
-Let $\cost_{i}$ denote an arbitrary cost term associated with asset $i$.
-Consider a long-term equilibrium in a power system with perfect competition, then, according to the zero-profit condition, each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$, \ie
+where an ``asset'' describes any operating components of the network, such as a generator, line, energy storage etc. We refer to the set of assets as $I$. 
+In a long-term equilibrium of a power system with perfect competition, the zero-profit condition states that each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
 \begin{align}
@@
 \end{align}
-where $\costfactor$ denotes a cost factor in \euro/MW. 
+where $\costfactor$ denotes a corresponding cost factor in \euro/MW. 
 If $\cost_i$ describes the OPEX occasioned by asset $i$, the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$.
-However, as we will show later, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given at the optimum.
-%As we will show later the composition of $\mu_{i,t}$ must be determined for each type of asset individually.
-
-% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
-Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
-\begin{align}
-    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
-    \label{eq:cost_allocation}
-\end{align}
-The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. However, we replace this quantity by 
-\begin{align}
-    % \pdv{\state}{\demand} \demand \rightarrow \int \pdv{\state}{\demand}  \,\mathrm{d}\demand  \rightarrow \allocatestate
-    \pdv{\state}{\demand} \demand \Leftrightarrow  \allocatestate
-    \label{eq:state_allocation}
-\end{align}    
-where $\allocatestate$ represents the amount of power that  asset $i$ supplies demand $\demand$ with.
-As we will show later, it is calculated from established flow allocation schemes.
-By following this approach, the differential quantity on the left side, is replaced by an integral quantity.
-The latter does not only rely on a marginal sensitivity of the nodal demand $\demand$, but makes further assumptions on how the operational state $\state$, all in all, serves consumers in the network.
-The dispatch allocation $\allocatestate$ ensures that all power produced or processed by asset $i$ at time $t$  is assigned to consumers,
-\begin{align}
-    \state = \sum_n \allocatestate,
-    \label{eq:state_allocation_constraint}
-\end{align}
-from which follows that the sum of all cost contributions return the total cost term, 
-\begin{align}
-    \cost_i = \sum_{n,t} \allocatecost
-    \label{eq:cost_payback}
-\end{align}
-% Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
-% \begin{align}
-%     \cost_{n,t} = \sum_i \allocatecost
-% \end{align}
-
-\end{subequations}
-
-By combining \cref{eq:cost_allocation,eq:state_allocation}, we are able to formulate a full cost allocation from demand to network assets for every time-step $t$.
-In particular, the definition of cost factors $\costfactor$ must be derived for each cost term individually.
-In the following we depict the presented scheme for typical classes of assets.
-Therefore, we assume a network with generators $s$, transmission lines $\ell$ and storage units $r$, each asset $i = \{s, \ell, r\}$ adds a term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
+However, as we will show in the following, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given at the cost-optimum.
+
+For a detailed demonstration, we derive  \cref{eq:cost_decomposition} for each cost term separately while considering generators, transmission lines and storages in the network.
+Therefore, let $o_i$ denote the operational price per MWh for asset $i$ and $c_i$ the capital price for one MW capacity expansion.
+\Cref{tab:cost_allocation_map} summarizes all derived relations, these can be inserted into \cref{eq:asset_cost,eq:cost_decomposition} for each single cost term.
 
@@
 
-\subsection{OPEX Allocation}
-
-Let the operational price for an asset $i$ be given by $o_i$. Then, the OPEX occasioned, for example, by generator $s$ is given by 
+\subsubsection*{Generators}
+\begin{subequations}
+Let $S \subseteq I$ be the set of generators in the network, such that $\generation = s_{s,t}$ describes the power production of generator $s \in S$.
+The OPEX occasioned by generator $s$ is given by a cost-weighted sum of the production, thus
 \begin{align}
@@
 \end{align}
-where $\generation$ denotes its power generation at time $t$.
-\Cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme}.
-As a result we obtain
-\begin{align}
-    \allocateopex[n \rightarrow s] &= 
-   \operationalpricegeneration \,  \allocategeneration
-\label{eq:allocate_opexGeneration_detailed}
-\end{align}
-which is the contribution of $\demand$ to the OPEX at generator $s$.
-The quantity  
-\begin{align}
- \allocategeneration \Leftrightarrow \pdv{\generation}{\demand} \, \demand
- \label{eq:allocate_peer}
-\end{align}
-superseding the derivative on the right side, describes the power that is produced by generator $s$ and consumed at node $n$ at time $t$. \\
-In the same manner, we can follow the scheme to allocate OPEX for flow $\flow$ and storage unit dispatch $\storagedispatch$.
-As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow.
-The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation.
-\\
-
-The scheme works for all other cost attached to the operational state of an asset $i$.
-Given for example a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, the cost term for emission adds up to
-\begin{align}
- \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
-\end{align}
-where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$.
-The allocated payment for consumers at bus $n$ at time $t$ assigned to generator $s$ is then given by 
-\begin{align}
- \allocateemissioncost = \emissionprice \, \emission \, \allocategeneration
-\end{align}
-
-\subsection{CAPEX Allocation}
-For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and its binding constraints.
-For all assets, let the capital price for one MW capacity expansion be denoted by $c_i$.
-All quantities for the CAPEX allocation, which we now discuss in detail, are summarized in the lower section of \cref{tab:cost_allocation_map}.
-
-\subsubsection{Generators}
-
-The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
-\begin{align}
-\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
-\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
-- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
-\label[constraint]{eq:lower_generation_capacity_constraint}
-\end{align}
-where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
+In case a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, is assumed, a further the cost term per generator $s$     
+\begin{align}
+ \emissioncost_s = \emissionprice \, \sum_t  \emission \, \generation
+\end{align}
+adds to $\totalcost$.
+Here, $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$.
+In contrast to OPEX and emission costs, the CAPEX of $S$ are not a function of the production $\generation$, but of the actual installed capacity $\capacitygeneration$ of generator $s$.
+In the optimization it limits the generation $\generation$ in the form of
+\begin{align}
+\generation - \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
+\label[constraint]{eq:upper_generation_capacity_constraint}
+% - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
+% \label[constraint]{eq:lower_generation_capacity_constraint}
+\end{align}
+The constraint yields a shadow-price of $\muuppergeneration$ in literature often denoted as the Quality of Supply.
+It can be interpreted as the the price per MW that \cref{eq:upper_generation_capacity_constraint} imposes to the system.
+If $\muuppergeneration$  is bigger than zero, the constraint is binding, which pushes investments in $\capacitygeneration$.
+As shown in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
+\begin{align}
+ \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
+ \label{eq:no_profit_capex_generation}
+\end{align}
+This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}.
+%If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
+
+% \subsection{CAPEX Allocation}
+% For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and its binding constraints.
+
+% \subsubsection{Generators}
+
+% The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
+% \begin{align}
+% \generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
+% \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
+% - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
+% \label[constraint]{eq:lower_generation_capacity_constraint}
+% \end{align}
+% where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
 At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.
 As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
-\begin{align}
- \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
- \label{eq:no_profit_capex_generation}
-\end{align}
-This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}. The resulting the CAPEX allocation is given by
-\begin{align}
- \allocatecapexgeneration = \muuppergeneration \, \allocategeneration
- \label{eq:allocate_capexGeneration_detailed}
-\end{align}
-How does this allocation behave?
+% \begin{align}
+%  \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
+%  \label{eq:no_profit_capex_generation}
+% \end{align}
+% This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}. The resulting the CAPEX allocation is given by
+% \begin{align}
+%  \allocatecapexgeneration = \muuppergeneration \, \allocategeneration
+%  \label{eq:allocate_capexGeneration_detailed}
+% \end{align}
+% How does this allocation behave?
 According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
@@
 If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} , we can use \cref{eq:state_allocation_constraint,eq:no_profit_capex_generation} to see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
-
-\subsubsection{Transmission Lines}
-
-The transmission capacity $\capacityflow$ limits the flow $\flow$ in both directions,
+% When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} , we can use \cref{eq:state_allocation_constraint,eq:no_profit_capex_generation} to see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
+
+\end{subequations}
+
+\subsubsection*{Transmission Lines}
+\begin{subequations}
+
+Let $L \subset I$ be the set of transmission lines in the system, these may include Alternating Current (AC) as well as Directed Current (DC) lines.
+Further let $\flow = s_{\ell,t}$ represent the power flow on line $\ell \in L$.
+If the OPEX of the transmission system is taken into account in $\totalcost$ (these are often neglected in power system models), these may be approximated by $\opexflow_\ell = \sum_t o_\ell |\flow|$, that is, a cost weighted sum of the net flow on line $\ell$.
+Again this stands in contrast to the CAPEX which not a function of $\flow$ but of the transmission capacity $\capacityflow$.
+It limits the flow $\flow$ in both directions,
 \begin{align}
@@
 \end{align}
-which yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which shows that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow
+At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$.
+Again, we use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
 \begin{align}
@@
 \end{align}
-From here, we follow the scheme in \cref{eq:general_scheme} which finally defines the CAPEX allocation as 
-\begin{align}
-    \allocatecapexflow &=  
-   \left( \muupperflow - \mulowerflow\right) \, \allocateflow
-   \label{eq:allocate_capexFlow_detailed}
-\end{align}
-The quantity 
-\begin{align}
- \allocateflow \Leftrightarrow  \pdv{\flow}{\demand}\, \demand
-\end{align}
-represents the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
-% According to \cref{eq:state_allocation_constraint} it has to fulfill 
-% \begin{align}
-%  \flow = \sum_{n} \allocateflow
-% \label{eq:allocate_flow_constraint}
-% \end{align}
-The  shadow prices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
-Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$.
-Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent.
-This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:state_allocation_constraint,eq:no_profit_capex_flow}
-
-\subsubsection{Storages}
+Again, the shadow prices $\muupperflow$ and $\mulowerflow$ can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
+Hence, a non-zero values indicate that \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint} are bound and therefore that the congestion on line $\ell$ at time $t$ is imposing costs to the system.
+
+\end{subequations}
+
+\subsubsection*{Storages}
 \label{sec:storages}
+\begin{subequations}
 
@@
 
+Let $R \in S$ denote all storages in the system.
 In a simplified storage model, $\capacitystorage$ limits the storage dispatch $\storagedispatch$ and charging $\storagecharge$.
@@
 In \cref{sec:zero_profit_storage_units} we formulate the mathematical details.
-As already shown in \cite{brown_decreasing_2020}, the total expenditures at $r$ are fully paid back by differences of the LMP at which the storage ``buys`` and ``sells`` power.
-Taking only the CAPEX into account the zero profit condition reduces to
+The OPEX term which adds to the objective function is given by
+\begin{align}
+    \opexstorage = \sum_r o_r \storagedispatch 
+\end{align}
+Using the result of \cite{brown_decreasing_2020} the CAPEX can be related to the operation of a storage unit $r$ through
 \begin{align}
@@
 where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint.
-When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand \Leftrightarrow 0$, implying that the when a storage charges power, it does not supply any demand.
+% When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand \Leftrightarrow 0$, implying that the when a storage charges power, it does not supply any demand.
 At these times, it the storage is a effective consumer retrieving power from producing assets.
-This leaves us with 
-\begin{align}
-     \allocatecapexstorage = \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \allocatestoragedispatch
-\end{align}
-where the dispatch allocation only assigns discharge power to demand $\demand$, thus
-\begin{align}
-    \allocatestoragedispatch \Leftrightarrow \pdv{\storagedispatch}{\demand} \demand
-\end{align}
-Note that this relation will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
-% \begin{align}
-%     \capexstorage_r + \remainingcost^E_r = \sum_n \allocatecapexstorage
-% \end{align} 
-It is crucial to note that storage units perform a redistribution of money, and therefore distort the cost allocation.
+% This leaves us with 
+% \begin{align}
+    %      \allocatecapexstorage = \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \allocatestoragedispatch
+% \end{align}
+% where the dispatch allocation only assigns discharge power to demand $\demand$, thus
+% \begin{align}
+%     \allocatestoragedispatch \Leftrightarrow \pdv{\storagedispatch}{\demand} \demand
+% \end{align}
+% Note that this relation will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
+% % \begin{align}
+    % %     \capexstorage_r + \remainingcost^E_r = \sum_n \allocatecapexstorage
+% % \end{align} 
+% It is crucial to note that storage units perform a redistribution of money, and therefore distort the cost allocation.
 So, a certain share of what is allocated to the CAPEX of a storage is in another time step spent by the same storage in order to buy power from other assets.
@@
 
-We shortly annotate that it is possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
+% We shortly annotate that it is possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
 Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$.
 The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.
-
-% In a final step we can still allocate all cost to the demand $\demand$.
-However, this would require knowledge about when the power $\storagedispatch$ was initially charged: Given the amount of power $A_{r, t_{in} \rightarrow t_{out}}$ that is charged by storage $r$ at time step $t_{in}$ and discharged at $t_{out}$ ...
+\end{subequations}
+
+\subsection{Allocation of Power Dispatch}
+\label{sec:localizing_allocations}
+\begin{subequations}
+
+After showing that cost terms $\cost$ for OPEX and CAPEX can be represented by cost-weighted sum of the operational state $\state$, we now demonstrate how the operational state in turn is allocated to consumers.
+This builds the second crucial ingredient of the cost allocation.
+
+Dispatch and flow in a system can be considered as a superposition of individual contributions of nodes and assets.
+In order to artificially quantify these contribution, the literature provides various methods, named flow allocation schemes.
+Each of these follow a specific set of assumptions which result in peer-to-peer allocations $A_{m \rightarrow n}$.
+That is a measure for the power supplied at node $m$ and consumed at node $n$.
+
+In the following, we resort to one specific flow allocation scheme Average Participation (AP), also known as Flow Tracing.
+The basic idea is to trace the power injection at bus $m$ through the network while following the real power flow on transmission lines and applying the principal of proportional sharing.
+At each bus, including the starting bus $m$, the traced flow might be mixed with incoming power flows from other buses.
+As soon as the power is absorbed or flowing out of a bus, the traced flow originating from $m$ splits in the same proportion as the total flow.
+This assumption leads to regionally confined allocation $A_{m \rightarrow n}$ based on a straightforward principal.
+A mathematical formulation of the AP scheme is documented in \cref{sec:net_ap}.
+For a detailed comparison with other scheme we refer to [].
+
+% The Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} for example assumes that every producer supplies every consumer proportional to its share in the total consumption.
+Despite the straightforward principal, this scheme does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
+The Flow Based Market Coupling (FBMC) is based on predefined zonal sensitivities for allocating intra-regional and inter-regional power flow.
+However, with its reliance on the mentioned sensitivity, the Generation Shift Keys, the scheme is strongly dependent on exogenous assumptions.
+
+The peer-to-peer allocations $\allocatepeer$ fulfill some basic properties.
+On the one hand it allocates the all power productions at time $t$, \ie when summing over all receiving nodes $n$ the allocations yield the gross power generation of producing assets $i \in S \cup R$ (generators and storages) attached to $m$, denoted by $\nodalgeneration[m]$.
+Mathematically this translates to
+\begin{align}         
+    \nodalgeneration[m] = 
+    \sum_{i \in S \cup R} \incidenceasset[m] \, \state = \sum_n \allocatepeer
+    \label{eq:nodalgeneration}
+\end{align}
+where $\incidenceasset[m]$ is 1 if asset $i$ is attached to bus $m$ and zero otherwise.
+On the other hand, when summing over all supplying nodes, the allocation  $\allocatepeer$ yields the gross nodal demand $\demand$ at node $n$ and time $t$, which leaves us with
+\begin{align}
+    \demand = \sum_m \allocatepeer 
+\end{align}
+As the standard formulation implies, we assume that only net power production of $m$ is allocated to other buses.
+That is, if the nodal generation $\nodalgeneration[m]$ does not exceed the nodal demand $\demand[m]$, all of it is allocated to local consumers, \ie $\allocatepeer[m \rightarrow m] = \min\left(\nodalgeneration[m], \demand[m]\right)$.
+The principal of proportional sharing sets the allocation of a single producing asset $i \in S \cup R$ proportional to the nodal allocation $\allocatepeer$ and weights it by the share $w_{i,t} = \incidenceasset[m] \state / \nodalgeneration[m]$ that asset $i$ contributes to the nodal generation $\nodalgeneration[m]$, leading to
+\begin{align}
+    \allocatestate = w_{i,t} \allocatepeer \Forall{i \in S \cup R, n , t}
+    \label{eq:allocate_production}
+\end{align}
+
+This relation builds one crucial key to the cost allocation. It states how much of the power produced by asset $i \in S \cup R$ is finally consumed by consumers at $n$.
+
+Yet, we didn't touch the allocation from transporting assets $i \in L$ to consumers.
+To tackle this, let $\ptdf$ denote a linear mapping between the injection $\left(\nodalgeneration[m] - \demand[m]\right)$ and the flow $\flow$, such that
+\begin{align}
+ \flow  = \sum_m \ptdf[m]\, \left(\nodalgeneration[m] - \demand[m]\right) \Forall{\ell \in L, t}
+ \label{eq:power_flow_equation}  
+\end{align}
+Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
+For transport models or networks with High Voltage Directed Current (HVDC) lines, these can calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
+Inserting \cref{eq:nodalgeneration} into \cref{eq:power_flow_equation} and expanding the sum yields 
+\begin{align}
+    \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) \Forall{\ell \in L, n, t}
+ \label{eq:allocate_flow}
+\end{align}
+Complementary to \cref{eq:allocate_production}, this allocation indicates the power flow on line $\ell$ and time $t$ which is ending up at node $n$ and therefore consumed by $\demand$.  
+Note that the assignment of line flows to consumers is different from the traced flow returned by the AP scheme, as we're only taking the peer-to-peer relation $\allocatepeer$ and, for reasons to be yet to come, from there calculate the power flow .
+
+With \cref{eq:allocate_production,eq:allocate_flow} a full set of allocations between the operational state $\state$ of asset $i \in I$ is derived.
+Naturally, the sum over all receiving nodes reproduces the operational state $\state$ of asset $i$,
+\begin{align}
+    \state = \sum_n \allocatestate \Forall{i, t}
+\end{align}
+
+\end{subequations}
+
+\subsection{Cost Allocation}
+\label{sec:general_scheme}
+\begin{subequations}
+\label[subequations]{eq:general_scheme}
+
+% Let $\cost_{i}$ denote an arbitrary cost term associated with asset $i$.  %As we will show later the composition of $\mu_{i,t}$ must be determined for each type of asset individually.
+
+% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
+Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
+\begin{align}
+    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
+    \label{eq:cost_allocation}
+\end{align}
+The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. However, we replace this quantity by 
+\begin{align}
+    % \pdv{\state}{\demand} \demand \rightarrow \int \pdv{\state}{\demand}  \,\mathrm{d}\demand  \rightarrow \allocatestate
+    \pdv{\state}{\demand} \demand \Leftrightarrow  \allocatestate
+    \label{eq:state_allocation}
+\end{align}    
+where $\allocatestate$ represents the amount of power that  asset $i$ supplies demand $\demand$ with.
+As we will show later, it is calculated from established flow allocation schemes.
+By following this approach, the differential quantity on the left side, is replaced by an integral quantity.
+The latter does not only rely on a marginal sensitivity of the nodal demand $\demand$, but makes further assumptions on how the operational state $\state$, all in all, serves consumers in the network.
+The dispatch allocation $\allocatestate$ ensures that all power produced or processed by asset $i$ at time $t$  is assigned to consumers,
+\begin{align}
+    \state = \sum_n \allocatestate,
+    \label{eq:state_allocation_constraint}
+\end{align}
+from which follows that the sum of all cost contributions return the total cost term, 
+\begin{align}
+    \cost_i = \sum_{n,t} \allocatecost
+    \label{eq:cost_payback}
+\end{align}
+% Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
+% \begin{align}
+%     \cost_{n,t} = \sum_i \allocatecost
+% \end{align}
+
+\end{subequations}
+
+By combining \cref{eq:cost_allocation,eq:state_allocation}, we are able to formulate a full cost allocation from demand to network assets for every time-step $t$.
+In particular, the definition of cost factors $\costfactor$ must be derived for each cost term individually.
+In the following we depict the presented scheme for typical classes of assets.
+Therefore, we assume a network with generators $s$, transmission lines $\ell$ and storage units $r$, each asset $i = \{s, \ell, r\}$ adds a term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
+
+\subsection{Connection to Locational Marginal Prices}
+
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
+Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
+\begin{align}
+\lmp = \pdv{\,\totalcost}{\,\demand}
+\label{eq:lmp}
+\end{align}
+% 
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\linewidth]{price_decomposition.png}
+\caption{Schematic decomposition of the Locational Market Price $\lmp$.
+As the cost function of the power system model different cost terms as indicated by \cref{eq:total_cost}, so does each $\lmp$ at node $n$ and time $t$.}
+\label{fig:price_decomposition}
+\end{figure}
+
+From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms. 
+This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
+
+% 
+This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
+\begin{align}
+\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
+\label{eq:total_revenue}
+\end{align}
+depending on the costs $\remainingcost$ which are independent of the nodal demand  
+\begin{align}
+ \pdv{\remainingcost}{\demand} = 0
+\end{align}
+% 
+Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
+However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
+
+However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
+This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.
 
@@
 \begin{align}
-    \cost' = \sum_i  \cost_i - \remainingcost_i
+    \cost[\circ '] = \sum_i  \cost_i - \remainingcost_i
     \label{altered_cost}
@@
 \begin{align}
-    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost^I \Forall{i \in I}
-\end{align}
-where $\allocatecost^I$ denotes the CAPEX allocation presented above.
+    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost \Forall{i \in I}
+\end{align}
+where $\allocatecost$ denotes the CAPEX allocation presented above.
 
@@
 has to be subsidized by governments or communities.
-It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost^I = \subsidycost_i $.
+It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost = \subsidycost_i $.
 
@@
 
-\section{Assumptions on Power Allocations}
-% \section{\texorpdfstring{Extended Solution Space of $\boldsymbol{\slackk[m]}$}{Solution Space of the Slack}}
-\label{sec:localizing_allocations}
-
-The presented cost allocation suits for any type of topology and network setup.
-It may be extended by other classes of network assets or design constraints.
-But so far, the question of how the allocation $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open.
-
-In a first step, we show that the needed information narrows down to peer-to-peer allocations between producing assets and consumers.
-%We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:state_allocation_constraint}).
-\\
-Let $\injection[m]$ denote the net power injection at node $m$ and time $t$. It is given by 
-\begin{align}
-    \injection[m] = \sum_{i \in \{s, r\}} \incidenceasset[m] \, \state - \demand[m]
-\end{align}
-where $\incidenceasset[m]$ is 1 if asset $i$ is attached to bus $m$ and zero otherwise. Now, let $\ptdf$ denote a linear mapping between the injection $\injection[m]$ and the flow $\flow$, such that  
-\begin{align}
- \flow  = \sum_m \ptdf[m]\, \injection[m]
- \label{eq:power_flow_equation}  
-\end{align}
-Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
-For transport models or mixed AC-DC networks, these can calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
-
-Using the two expression the derivative of the flow $\flow$ with respect to the demand can be expressed through 
-\begin{align}
-    \pdv{\flow}{\demand} \demand \Leftrightarrow  \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
- \label{eq:allocate_peer_to_allocate_flow},
-\end{align}
-where we collected all power produced at bus $m$ and consumed at bus $n$ in $\allocatepeer$, defined as
-\begin{align}
-    \allocatepeer = \sum_{i \in \{S,R\}} \incidenceasset[m] \, \allocatestate  
-    % \nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation + \sum_r \incidencestorage[m] \storagedispatch \Forall{n}
-\end{align}
-\Cref{eq:allocate_peer_to_allocate_flow} shows that $\allocateflow$ is fully determined through the peer-to-peer allocation $\allocatepeer$.
-In other words, we only need to know how much of the power produced at node $m$ is consumed at node $n$.
-With this information we may also derive the cost allocation for the transmission system.
-Unfortunately, the solution for $\allocatepeer$ is non-unique and requires further assumptions.
-Established flow allocation schemes approach this problem from different directions.
-Principally two options exist \textit{what} is allocated
-% 
-\begin{enumerate}
-\item gross power injections \label{gross}
-\item net power injections \label{net}
-\end{enumerate}
-% 
-Further it is important \textit{what assumptions} define the allocation, \ie what method is used to define the pairs of sources and sinks. The three suitable approaches we present here are
-% 
-\begin{enumerate}[label=\alph*., ref=\alph*]
-\item Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} which assumes
-that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
-\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
-\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions.
-The interregional exchange is only allocating net power deficit or excess of the regions.
-\label{fbmc}
-\end{enumerate}
-% 
-We show the mathematical formulation for all combinations \ref{ebe}\ref{gross} - \ref{fbmc}\ref{net} in \cref{sec:gross_ebe,sec:net_ebe,sec:gross_ap,sec:net_ap}.
-% 
-Principally, allocating net power instead of gross power leads to less P2P trades, as power from a bus $m$ with $\nodalgeneration[m] \le \nodaldemand[m]$ is not assigned to other buses, only to $m$.
-
-When it comes to the methods, as literature has often pointed out, the EBE principal does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
-The AP allocation tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal.
-Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when following in the direction of $\flow$, ends up at bus $n$.
-In contrast, the FBMC allows to control the regions or market zones which are netted out in a first step.
-If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.
-However, it relies on further assumptions such as the Generation Shift Keys determining the producing assets which are deployed for the inter-regional exchange.
-% 
-\begin{figure*}[t]
-    \centering
-    \includegraphics[width=\linewidth]{example_network.png}
-    \caption{Illustrative example of a 2-bus network with one optimized time step.
-Fixed prices and constraining values are given in the left box for each bus and the transmission line.
-Optimized values are given in the right boxes.
-Bus~1 has a cheaper operational price $o$, capital prices are the same for both.
-As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2.
-The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.}
-    \label{fig:example_network}
-\end{figure*}
-%
-
-From this point of view, we restrict the focus of this research to \textbf{net} injection allocation on the basis of \textbf{AP}, type \ref{net}\ref{ap}, as locality and no further dependency on external decisions are strong arguments for a transparent cost allocation.
-We emphasize that the AP based allocation does not imply traceability of the resulting flows, as it only determines the \textit{source}-\textit{sink} relations $\allocatepeer$.
-The flow allocation still follows the power flow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.
-In the following example, we illustrate the functionality of the cost allocation by an example of a small artificial network.
-% % 
 % 
@@
 The costs are allocated according to type \ref{net}\ref{ap}. For allocating flows on  DC-lines we use the method described in \cite{hofmann_flow_2020-1}.
-In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost'$.
+In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost[\circ ']$.
 Due to the design constraints, \cref{sec:design_constraints}, and storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$.

### 2020-10-15 paper/flow_based.tex
coming to chapter "numerical example" now

@@
 \newcommand{\cost}[1][\circ]{\mathcal{C}^{#1}}
+\newcommand{\payment}[1][n]{\mathcal{C}_{#1,t}}
 \newcommand{\opex}{\mathcal{O}}
@@
     \state = \sum_n \allocatestate \Forall{i, t}
+    \label{eq:decompose_state}
 \end{align}
@@
 
-% Let $\cost_{i}$ denote an arbitrary cost term associated with asset $i$.  %As we will show later the composition of $\mu_{i,t}$ must be determined for each type of asset individually.
-
-% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
-Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as
-\begin{align}
-    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
+Using the presented relations, we are able to straightforwardly define the full cost allocation.
+Therefore, the expression in \cref{eq:decompose_state} is inserted in \cref{eq:cost_decomposition} which leads to
+\begin{align}
+    \allocatecost = \costfactor \, \allocatestate
     \label{eq:cost_allocation}
 \end{align}
-The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. However, we replace this quantity by 
-\begin{align}
-    % \pdv{\state}{\demand} \demand \rightarrow \int \pdv{\state}{\demand}  \,\mathrm{d}\demand  \rightarrow \allocatestate
-    \pdv{\state}{\demand} \demand \Leftrightarrow  \allocatestate
-    \label{eq:state_allocation}
-\end{align}    
-where $\allocatestate$ represents the amount of power that  asset $i$ supplies demand $\demand$ with.
-As we will show later, it is calculated from established flow allocation schemes.
-By following this approach, the differential quantity on the left side, is replaced by an integral quantity.
-The latter does not only rely on a marginal sensitivity of the nodal demand $\demand$, but makes further assumptions on how the operational state $\state$, all in all, serves consumers in the network.
-The dispatch allocation $\allocatestate$ ensures that all power produced or processed by asset $i$ at time $t$  is assigned to consumers,
-\begin{align}
-    \state = \sum_n \allocatestate,
-    \label{eq:state_allocation_constraint}
-\end{align}
-from which follows that the sum of all cost contributions return the total cost term, 
-\begin{align}
-    \cost_i = \sum_{n,t} \allocatecost
-    \label{eq:cost_payback}
-\end{align}
-% Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
-% \begin{align}
-%     \cost_{n,t} = \sum_i \allocatecost
-% \end{align}
+By default, all cost of a cost term $\cost_i$ associated with asset $i$ are allocated, thus $\cost_i = \sum_n \allocatecost$; likewise all system costs are distributed, $\totalcost = \sum_{\circ, i, n, t} \allocatecost$.
+
+The cost allocation entails a further important property.
+When summing over all assets $i$ and cost terms $\circ$, the cost allocation yields $\payment$, which we refer to as the nodal payment or payment.
+This exactly matches the payment determined by the Locational Marginal Price (LMP) $\lmp$, thus
+\begin{align}
+    \payment = \sum_{\circ, i} \allocatecost  = \lmp\, \demand  \Forall{n,t}
+    \label{eq:payment}
+\end{align}
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
+Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
+At the optimum the LMP is given by a shadow price of the nodal balance constraint (see \cref{sec:lmp} for details).
+\Cref{eq:payment} which is in detailed proved in ... connects the cost allocation to a nodal pricing scheme implied by the LMP.
 
@@
 
-By combining \cref{eq:cost_allocation,eq:state_allocation}, we are able to formulate a full cost allocation from demand to network assets for every time-step $t$.
-In particular, the definition of cost factors $\costfactor$ must be derived for each cost term individually.
-In the following we depict the presented scheme for typical classes of assets.
-Therefore, we assume a network with generators $s$, transmission lines $\ell$ and storage units $r$, each asset $i = \{s, \ell, r\}$ adds a term for OPEX and a term for CAPEX to the total system cost $\totalcost$.
-
-\subsection{Connection to Locational Marginal Prices}
-
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
-Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
-\begin{align}
-\lmp = \pdv{\,\totalcost}{\,\demand}
-\label{eq:lmp}
-\end{align}
-% 
-\begin{figure}[h]
-\centering
-\includegraphics[width=.8\linewidth]{price_decomposition.png}
-\caption{Schematic decomposition of the Locational Market Price $\lmp$.
+% \subsection{Connection to Locational Marginal Prices}
+
+% % 
+% \begin{figure}[h]
+% \centering
+% \includegraphics[width=.8\linewidth]{price_decomposition.png}
+% \caption{Schematic decomposition of the Locational Market Price $\lmp$.
 As the cost function of the power system model different cost terms as indicated by \cref{eq:total_cost}, so does each $\lmp$ at node $n$ and time $t$.}
-\label{fig:price_decomposition}
-\end{figure}
-
-From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms. 
-This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
-
-% 
-This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
-\begin{align}
-\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
-\label{eq:total_revenue}
-\end{align}
-depending on the costs $\remainingcost$ which are independent of the nodal demand  
-\begin{align}
- \pdv{\remainingcost}{\demand} = 0
-\end{align}
-% 
-Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
+% \label{fig:price_decomposition}
+% \end{figure}
+
+% From feeding \cref{eq:total_cost} into \cref{eq:lmp} it follows naturally that the LMP splits into contribution to the above mentioned cost terms. 
+% This relation, which we schematically show in \cref{fig:price_decomposition}, was already shown in extensive investigations of the LMP \cite{schweppe_spot_1988}.
+
+% % 
+% This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
+% \begin{align}
+% \totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
+% \label{eq:total_revenue}
+% \end{align}
+% depending on the costs $\remainingcost$ which are independent of the nodal demand  
+% \begin{align}
+%  \pdv{\remainingcost}{\demand} = 0
+% \end{align}
+% % 
+% Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
 However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
 
-However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
+% However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
 This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.
@@
 \label{sec:design_constraints}
+\begin{subequations}
+
 Power system modelling does rarely follow a pure Greenfield approach with unlimited capacity expansion.
 Rather, today's models are setting various constraints defining socio-political or  technical requirements.
-As mentioned before this will alter the equality of total cost and total revenue, \ie leads to $\remainingcost \ne 0$ in \cref{eq:total_revenue}.
+However, this will alter the equality of total cost and total revenue.
 More precisely, each constraint $h_j$ (other then the nodal balance constraint) of the form
 \begin{align}
-    h_j \left(\state, \capacity \right) - K < 0
+    h_j \left(\state, \capacity \right) < K
     \label{eq:design_constraint}
 \end{align}
-where $K$ is any non-zero constant and $\capacity$ is collects all capacities of asset $i$, will result in a cost term contributing to $\remainingcost$ and alter \cref{eq:cost_decomposition} to 
-\begin{align}
-    \cost_i - \remainingcost_i = \sum_t \costfactor \, \state
-\end{align}
-The total allocatable cost associated to cost term $\cost$ is then given by 
-\begin{align}
-    \cost[\circ '] = \sum_i  \cost_i - \remainingcost_i
-    \label{altered_cost}
-\end{align}
+where $K$ is any non-zero constant and $\capacity$ denotes the nominal capacities of asset $i$, will  alter \cref{eq:cost_decomposition} to 
+\begin{align}
+    \cost_i - \remainingcost^\circ_i &= \sum_t \costfactor \, \state \\
+    &= \sum_{n,t} \allocatecost
+\end{align}
+% The total allocatable cost associated to cost term $\cost$ is then given by 
+% \begin{align}
+%     \cost[\circ '] = \sum_i  \cost_i - \remainingcost_i
+%     \label{altered_cost}
+% \end{align}
 According to the nature of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, it is either larger, equal or lower than $\cost$.
+Whereas \cref{eq:payment} still holds true, the total of payments do not return the total system cost $\totalcost$ anymore.
+The total mismatch $\remainingcost$ is given by
+\begin{align}
+    \remainingcost = \totalcost - \sum_{n,t} \payment
+\end{align}
 In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to incorporate them into the cost allocation.
 
+\end{subequations}
+
 \subsubsection{Capacity Expansion Limit}
@@
 In more realistic setups, generators, lines or other assets can only be built up to a certain limit.
-This might be due to land use restrictions or social acceptance problems.
+This might be due to land use restrictions or social acceptance considerations.
 %In the following give a general solution how the cost allocation may handle this.
-However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit $\capacityupper$, in the form of 
+When constraining the capacity $\capacity$ to an upper limit $\capacityupper$, in the form of 
 \begin{align}
@@
 \end{align}
-the zero profit condition alters as soon as the constraint becomes binding.
-Then, the revenue of asset $i$ exceeds its total expenditures (OPEX + CAPEX).
-More precisely, the allocated CAPEX in \cref{tab:cost_allocation_map} will surpass the actual CAPEX of asset $i$ by the cost it has to pay for the scarcity, given by the absolute value of
+the zero profit condition alters as soon as the constraint becomes binding. Then, asset $i$ is payed an additional scarcity rent 
 \begin{align}
@@
 \end{align}
-% \begin{align}
-%  \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
-%  \label{eq:capex_generation_duality_bf2}
-% \end{align}
-
-% In order to ensure that the allocated CAPEX sum up to the actual CAPEX at asset $i$, meaning that \cref{eq:cost_payback} holds, we adjust the CAPEX allocations to 
-% \begin{align}
-%     \allocatecost = \left(\dfrac{c_i}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
-% \end{align}
-% with the cost factors $\costfactor$ for CAPEX given in \cref{tab:cost_allocation_map}. 
-% The costs which consumers at $n$ have to pay for the scarcity impacting asset $i$ are given   
-% \begin{align}
-%     \allocatescarcitycost = \left(\dfrac{\muuppernom}{c_i + \muuppernom}\right) \, \costfactor \, \allocatestate \Forall{i \in I}
-% \end{align}
-
-The share of the CAPEX allocation, which results from the scarcity, is given by 
-\begin{align}
-    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost \Forall{i \in I}
-\end{align}
-where $\allocatecost$ denotes the CAPEX allocation presented above.
-
-% have to be adjusted, as otherwise consumers would pay too much. As we show in ... updating the prices by a fix ratio 
-% \begin{align}
-%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppernom}\, \costfactor
-%     \label{eq:adjustment_lowernom}
-% \end{align}
-% will . Note that $\muuppernom$ is a positive number, so the adjustment cannot increase the paid price.
+This rent may account for different possible realms, as for example the increased market price due in higher competed areas or additional costs for social or environmental compensation.
+To end this, the share in $\allocatecost$ which consumers pay for the scarcity rent can be recalculated by a correct weighting the shadow price $\muuppergenerationnom$ with the capital price $c_i$, leading to
+\begin{align}
+    \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost \Forall{i}
+\end{align}
 
@@
 
-In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. Mathematically this translates to 
+In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. This introduces a constraint of the form 
 \begin{align}
@@
 Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding. 
-In that case, asset $i$ does not collect enough revenue in order to match the CAPEX. The difference, given by 
+In that case, asset $i$ does not collect enough revenue from $\allocatecost$ in order to match the CAPEX. The difference, given by 
 \begin{align}
@@
 
-% Again, we can balance this effect out by updating the cost factor to 
-% \begin{align}
-%     \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
-% \end{align}
-% In contrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
-% As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.
-
-% In order to take this effect into account for the cost allocation, we update the cost factors in \cref{tab:cost_allocation_map} to
-
-% \begin{align}
-%     \costfactor \rightarrow \dfrac{c_i}{c_i + \muuppergenerationnom}\, \costfactor
-% \end{align}
-% This ensures that the allocated CAPEX sum up to the actual CAPEX at asset $i$ (and \cref{eq:cost_payback} holds).
-
-% 
 \subsection{Numerical Example}
@@
 \subsection{LMP from Optimization}
+\label{sec:lmp}
 The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL).
@@
 \begin{align}
-    \sum_i \incidenceasset \state  &=  \demand 
+    \nodalgeneration - \sum_\ell \incidence \flow  &=  \demand 
     % \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0
@@
 \end{align}
-where $\incidenceasset$ is +1 if $i$ is attached to $n$ and a positive operation $\state$ delivers power to $n$, -1 if is attached to $n$ and a positive operation retrieves power from $n$ and zero else (note that for lines this results in the negative of the conventional Incidence Matrix).
+where $\incidence$ is +1 if line $\ell$ starts at bus $n$, -1 if it ends at $n$, 0 otherwise.
+The nodal generation $\nodalgeneration$ collects all nodal production, see \cref{eq:nodalgeneration}.
 The shadow price of the nodal balance constraint mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.

### 2020-10-15 paper/flow_based.tex
follow up

@@
 % general symbols
-\newcommand{\state}{s_{i,t}}
+\newcommand{\state}[1][i]{s_{#1,t}}
 \newcommand{\capacity}{S_{i}}
@@
 
-\newcommand{\kk}{k~\euro\,}
+\newcommand{\kk}{k\euro~}
 
@@
 \newcommand{\lmp}[1][n]{\lambda_{#1,t}}
-\newcommand{\averagelmp}{\left<\lmp\right>_t}
+\newcommand{\averagelmp}{\bar{\lambda}_n}
 \newcommand{\demand}[1][n]{d_{#1,t}}
@@
 \newcommand{\allocatecost}[1][n \rightarrow i]{\cost_{#1, t}}
-\newcommand{\allocatecapexgeneration}[1][n \rightarrow s]{\capexgeneration_{#1,t}}
-\newcommand{\allocatecapexflow}[1][n \rightarrow \ell]{\capexflow_{#1,t}}
-\newcommand{\allocatecapexstorage}[1][n \rightarrow r]{\capexstorage_{#1,t}}
+\newcommand{\allocatecapexgeneration}[1][n \rightarrow s,t]{\capexgeneration_{#1}}
+\newcommand{\allocatecapexflow}[1][n \rightarrow \ell,t]{\capexflow_{#1}}
+\newcommand{\allocatecapexstorage}[1][n \rightarrow r, t]{\capexstorage_{#1}}
 \newcommand{\allocateopex}[1][n \rightarrow s]{\opex_{#1,t}}
@@
 Secondly, \textit{flow tracing}, following Bialek's Average Participation, assigns the operation of single network assets to consumers in an plausible manner.
-The following shows how combining these allow for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the end consumers.
+The following shows how combining these allow for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network.
 
@@
 \label[constraint]{eq:upper_generation_capacity_constraint}
-% - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
-% \label[constraint]{eq:lower_generation_capacity_constraint}
 \end{align}
@@
 \end{align}
-% The total allocatable cost associated to cost term $\cost$ is then given by 
-% \begin{align}
-%     \cost[\circ '] = \sum_i  \cost_i - \remainingcost_i
+% We define the allocated costs associated to cost term $\cost$ is then given by 
+% \begin{align}
+%     \bar{\cost} = \sum_i  \cost_i - \remainingcost_i
 %     \label{altered_cost}
@@
 
-\subsubsection{Capacity Expansion Limit}
-
+\subsubsection*{Capacity Expansion Limit}
+
+\begin{subequations}
 In more realistic setups, generators, lines or other assets can only be built up to a certain limit.
@@
 \end{align}
-
-\subsubsection{Brownfield Constraints}
-
+\end{subequations}
+
+\subsubsection*{Brownfield Constraints}
+
+\begin{subequations}
 In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. This introduces a constraint of the form 
@@
 It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost = \subsidycost_i $.
+\end{subequations}
 
@@
 % 
+\begin{figure*}[t]
+    \centering
+    \includegraphics[width=\linewidth]{example_network.png}
+    \caption{Illustrative example of a 2~bus network with one optimized time step.
+Fixed prices and constraining values are given in the left box for each bus and the transmission line.
+Optimized values are given in the right boxes.
+Generator~1 at bus~1 has a cheaper operational price $o$, capital prices are the same for both.
+As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2.
+}
+    \label{fig:example_network}
+\end{figure*}
+\begin{figure*}[h!]
+    \begin{subfigure}[c]{.495\linewidth}
+    % NOTE: Net EBE and Net AP lead to the same result here. Due to time limits just take the ebe figures
+    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ptpf.png}
+    \vspace{-40pt}
+    \subcaption{}
+    \label{fig:example_allocation_bus1}
+    \end{subfigure}
+    \begin{subfigure}[c]{.495\linewidth}
+    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ptpf.png}
+    \vspace{-40pt}
+    \subcaption{}
+    \label{fig:example_allocation_bus2}
+    \end{subfigure}
+    \caption{Power allocations $\allocatestate$ in the example network \cref{fig:example_network} using Average Participation.
+Consumers at for bus~1, figure (a), are, with 60~MW, totally supplied by the local generator and do not consume any imported power from bus 2.
+In contrast consumer at bus~2, figure (b), retrieve 40~MW from generator 1, induce a flow of 40~MW on the transmission line and consume 50~MW from the local generator.}
+    \label{fig:example_allocation}
+\end{figure*}
+
 Consider a two bus system, depicted in \cref{fig:example_network}, with one transmission line and one generator per bus.
@@
 %  
-\begin{figure*}[h!]
-    \begin{subfigure}[c]{.495\linewidth}
-    % NOTE: Net EBE and Net AP lead to the same result here. Due to time limits just take the ebe figures
-    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
-    \vspace{-40pt}
-    \subcaption{}
-    \label{fig:example_allocation_bus1}
-    \end{subfigure}
-    \begin{subfigure}[c]{.495\linewidth}
-    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
-    \vspace{-40pt}
-    \subcaption{}
-    \label{fig:example_allocation_bus2}
-    \end{subfigure}
-    \caption{Power allocations of net power injections using the AP scheme, type \ref{net}\ref{ap}, in the example network, \cref{fig:example_network}.
-Consumers at for bus~1, figure (a), are, with 60~MW, totally supplied by the local generator and do not consume any imported power from bus 2.
-In contrast consumer at bus~2, figure (b), retrieve 40~MW from generator 1, induce a flow of 40~MW on the transmission line and consume 50~MW from the local generator.}
-    \label{fig:example_allocation}
-\end{figure*}
-% 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
-    \caption{Full P2P cost allocation on the basis of power allocations shown in \cref{fig:example_allocation}.
-Consumers compensate OPEX and CAPEX of the generators they retrieve from, see \cref{fig:example_allocation}.
-As bus~1 is totally self-supplying, all it payment is assigned to the local generator.
-As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX at bus~1 and CAPEX for the transmission.}
-    \label{fig:example_payoff}
-\end{figure}    
-% 
-% 
-% 
-\Cref{fig:example_allocation} shows the allocated transactions on the basis of allocation \ref{net}\ref{ap} for both buses 1~\&~2 separately.
-Note, the ``sum'' of the two figures give to the actual dispatch.
+% 
+% 
+% 
+\Cref{fig:example_allocation} shows the allocation $\allocatestate$ for $n$~=~1 and $n$~=~2 separately.
+The ``sum'' of the two figures give to the actual dispatch and flow.
 The resulting P2P payments are given in \cref{fig:example_payoff}.
@@
 The right graph \cref{fig:example_allocation_bus2} shows the power allocations to $d_2$.
-We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from bus~1.
-Thus, consumers have the pay for the local OPEX and CAPEX as well as the proportional share of expenditures at bus~1 and the transmission system.
+We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from generator~1.
+Thus, consumers pay for the local OPEX and CAPEX as well as the proportional share of expenditures at bus~1 and the transmission system.
 As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it.
-The allocated CAPEX $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the actual CAPEX of generator~2.
-Again.
-in the case of generator~1, 2\kk of the 22\kk CAPEX allocation are associated with the scarcity cost for generator~1.
-The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line, which directly reflects the zero-profit condition.
+The allocated $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the full investment of generator~2.
+Again, 2\kk of the 22\kk which are allocated to investment in generator~1 are associated with the scarcity cost for generator~1.
+The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line.
+% 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{example_payoff_net_ptpf.png}
+    \caption{Full P2P cost allocation $\allocatecost$ on the optimized example network in \cref{fig:example_network}.
+Consumers compensate OPEX and CAPEX of the generators they retrieve from (compare with \cref{fig:example_allocation}).
+As bus~1 is totally self-supplying, all it payment is assigned to the local generator.
+As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX at bus~1 and CAPEX for the transmission.}
+    \label{fig:example_payoff}
+\end{figure}    
 % 
@@
 These values match their overall spending plus the cost of scarcity.
-The sum a row returns the total payment of consumers at bus $n$.
-For example the sum of payments of consumers at bus~1 is 36k~\euro.
+The sum a row returns the nodal payments $C_n = \lambda_n \, d_n$.
+For example the sum of payments of consumers at bus~1 is 36\kk.
 This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$.
@@
 \\
-% The example shows that allocating net power injection leads to intuitive and transparent  cost allocations.
-
-% In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments.
-The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}.
-As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated.
-Neither has bus~1 to pay fee to the transmission system as it only exports power.
-So, consumers at bus~1 pay to its local generator.
-Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
-% % 
-% Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$.
-The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX.
-Note this gives the same result as when allocating net injection with the Average Participation \ref{net}\ref{ap}.
 
@@
 Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network.
-Transmission lines are amplified in along the North-South axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the South-West.
+Transmission lines are amplified in along the north-south axis, including one large DC link, associated with the German S\"ud-Link, leading from the coastal region to the southwest.
 The total annualized cost of the power system roughly sums up to 42 billion \euro.
@@
 
-\Cref{fig:average_price} displays the average electricity price $\lmp$ per region.
+\Cref{fig:average_price} displays the load-weighted average electricity price $\averagelmp$ per region, defined by 
+\begin{align}
+    \averagelmp = \dfrac{\sum_{t} \lmp \demand}{\sum_t \demand}    
+\end{align}
 We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).
 Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices.
-The node with the lowest LMP in the upper North-West, stands out through high pre-installed offshore capacities.
+The node with the lowest LMP in the upper northwest, stands out through high pre-installed offshore capacities.
 \\
 
-The costs are allocated according to type \ref{net}\ref{ap}. For allocating flows on  DC-lines we use the method described in \cite{hofmann_flow_2020-1}.
-In \cref{fig:total_cost} we show the total of all ``allocatable'' costs $\cost[\circ ']$.
-Due to the design constraints, \cref{sec:design_constraints}, and storage units, \cref{sec:storages}, these are different from the original cost terms $\cost$.
-The figure also includes costs for scarcity $\scarcitycost$ and subsidy  $\subsidycost$.
+% The costs are allocated according to type \ref{net}\ref{ap}. For allocating flows on  DC-lines we use the method described in \cite{hofmann_flow_2020-1}.
+In \cref{fig:total_cost} we show the total of all allocated costs.
+The allocated cost to capital expenditures is given by $\cost - \remainingcost^\circ$.
+The difference $\remainingcost$ consists of scarcity rents $\scarcitycost$ and subsidies $\subsidycost$.
 Note that the sum of all contributions in \cref{fig:total_cost} equals the total cost $\totalcost$.
+In the following, we address each of the displayed cost terms and its corresponding allocations separately.
 
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{de50bf/total_costs}
+    \includegraphics[width=0.85\linewidth]{de50bf/total_costs}
     \caption{Total allocated payments of the system. }
@@
 Taking different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shown in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
-At this time (morning, end of February), the system hits the highest mismatch between low renewable power potential and high demand.
-The necessity for backup generators manifests in high allocation of CAPEX and consequently high LMP.
+At this time (morning, end of February), the system hits the highest mismatch between low renewable power potentials and high demands.
+The necessity for backup generators manifests in high allocation of CAPEX for OCGT and consequently high LMP.
 With a few exceptions in the South West and East, all consumers receive power from OCGT at this time, thus all pay high amounts for the needed backup infrastructure (operational state of the system is in detailed shown in \cref{fig:operation_high_expenditure}).
@@
 As for onshore and offshore wind farms, the produced power deeply penetrates the network, see \cref{fig:power_mix}, thus it is not only local, but also remote consumers which cover the CAPEX, see \cref{fig:capex_price}.
-This in turn benefits local consumers, which profit from the cheap power which explains why these regions end up with a low average LMP.\\
+This in turn benefits local consumers, which profit from the cheap operational prices of local wind farms.
+This explains why these regions end up with a low average LMP.\\
 
@@
 For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price,fig:emission_cost}.
-For OCGT, the spatial allocation of the OPEX clearly differs from  the CAPEX allocation.
-This results from the fact that during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators.
+Since during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators, thus the OPEX allocation of OCGT clearly differs from the CAPEX allocation.
 In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power.
-The OPEX allocations for renewables play with 0.2 \euro/MWh in average an inferior role.
+The average operational price for renewables is extremely low (0.2 \euro/MWh), thus they play an inferior role in the OPEX allocation.
 \\
@@
 \dots
+
+% ----------------------- APPENDIX -------------------------------------
+
 \clearpage
@@
 \notag
- \lagrangian\left(\state, \capacity, \lmp, \mu_j \right) =& 
- \;\;\; \sum_{i,t} o_i \state + \sum_{i} c_i \capacity  \\
+& \lagrangian\left(\state, \capacity, \lmp, \mu_j \right) = \\ 
 \notag
-&+  \sum_{n,t} \lmp \left( \demand - \sum_i \incidenceasset \, \state \right) \\
-&+ \sum_j \mu_j \, h_j \left(\state, \capacity \right)
+& \qquad + \sum_{i,t} o_i \state + \sum_{i} c_i \capacity  \\
+\notag
+&\qquad +  \sum_{n,t} \lmp \left( \demand - \nodalgeneration + \sum_\ell \incidence \, \flow \right) \\
+&\qquad + \sum_j \mu_j \, h_j \left(\state, \capacity \right)
 \label{eq:full_lagrangian}
@@
 where $h_j \left(\state, \capacity \right)$ denotes all inequality constraints attached to $\state$ and $\capacity$.
-% \begin{align}
-% \notag
-% & \lagrangian\left(\generation, \flow, \capacitygeneration, \capacityflow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
-% \notag
-% & \;\;\; \sum_{n,s} \capitalpricegeneration \capacitygeneration + \sum_{n, s, t} \operationalpricegeneration \generation + \sum_{\ell} \capitalpriceflow \, \capacityflow  \\
-% \notag
-% &+ \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \incidencegenerator\, \generation +  \demand  \right)  \\ 
-% \notag
-% &+ \sum_{\ell,c,t} \cycleprice \, \cycle \, \reactance \, \flow  \\
-% \notag
-% &+ \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacitygeneration \right)  - \sum_{n,s,t} \mulowergeneration \generation  \\
-% &+ \sum_{\ell,t} \muupperflow \left( \flow - \capacityflow \right) - \sum_{\ell,t} \mulowerflow \left( \flow + \capacityflow \right)     
-% \label{eq:full_lagrangian}
-% \end{align}
-% 
-% where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables.
-% 
+Note that according to \cref{eq:nodalgeneration} $\nodalgeneration$ is a composition of $\state$ and $\flow = \state[\ell] \forall \ell \in L$.
+
 In order to impose the Kirchhoff Voltage Law (KVL) for the linearized AC flow, the term 
@@
 \label{sec:zero_profit_generation}
+For each generator the optimization defines a lower capacity constraint, given by 
+\begin{align}
+    - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
+    \label[constraint]{eq:lower_generation_capacity_constraint}
+\end{align} 
 \Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint}, which yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$, imply the complementary slackness,
@@
 
-% \subsection{Brownfield Optimization and Capacity Restrictions}
-
-% Constraining the capacities $\capacitygeneration$  for a subset $S$ of generators to lower or upper limits in the form of
-% \begin{align}
-%     \capacitygenerationlower - \capacitygeneration \le 0 \resultsin{\mulowergenerationnom} \hpad \forall{s \in S} \label{eq:capacityexpansionminimum}\\
-%     \capacitygeneration - \capacitygenerationupper \le 0 \resultsin{\muuppergenerationnom} \hpad \forall{s \in S}
-% \label{eq:capacityexpansionmaximum}
-% \end{align}
-% alters the objective value as soon as one of those become bounding. 
-% The complementary slackness for both are 
-% \begin{align}
-%     \muuppergenerationnom \left( \capacitygeneration - \capacitygenerationupper \right) =0 \Forall{s \in S} \\
-%     \mulowergenerationnom \left( \capacitygeneration - \capacitygenerationlower \right) =0 \Forall{s \in S}
-% \end{align}
-
-% The CAPEX paybacks for generators and transmission lines
-% \cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
-% \begin{align}
-% \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, \\
-% \capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{s \in S}
-% \label{eq:capex_generation_duality_bf}
-% \end{align}
-% for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacitygeneration$ leads us to 
-% \begin{align}
-%  \capexgeneration_s - \remainingcost_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
-%  \label{eq:capex_generation_duality_bf2}
-% \end{align}
-% where we define the cost resulting from the capacity expansion limits as
-% \begin{align}
-%     \remainingcost_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacitygeneration
-%     \Forall{s}
-% \end{align}
-% As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \ge 0$, the latter can be a net positive or net negative cost term.
-
-% Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits,
-% \begin{align}
-%     \capexgeneration_s + \opexgeneration_s  - \remainingcost_s = \sum_t \lmp \incidencegenerator \generation \Forall{s \in S}
-% \end{align}
-% We see that the revenue of generator $s$ will not exactly pay back its full CAPEX and OPEX.
-The exogenous constraint shifts the zero-profit equations such that some of the expenditures for $s \in S$ cannot directly be allocated to consumers.
-In order to take this effect into account for the cost allocation, we update the shadow price $\muuppergeneration$ in \cref{tab:cost_allocation_map}, to
-
-% \begin{align}
-%     \muuppergeneration \rightarrow \dfrac{\capitalpricegeneration}{\capitalpricegeneration - \mulowergenerationnom + \muuppergenerationnom}\, \muuppergeneration
-% \end{align}
-% By doing so we ensure that the allocated CAPEX sum up to the actual CAPEX at generator $s$, thus 
-% \begin{align}
-%     \capexgeneration_s = \sum_{n,t} \allocatecapexgeneration \Forall{s \in S}
-% \end{align}
-
-% Doing likewise with a subset $L$ of transmission lines,
-% \begin{align}
-% \capacityflow \ge \capacityflowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
-% \capacityflow \le \capacityflowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
-% \end{align}
-% results in a shifted zero-profit rule for transmission line in the form 
-% \begin{align}
-%     \capitalpriceflow \, \capacityflow - \remainingcost_\ell = \sum_{n,t}\allocatecapexflow \Forall{\ell \in L}
-% \end{align}
-% where $\remainingcost_\ell$ is given by 
-% \begin{align}
-%     \remainingcost_\ell = \left(\mulowerflownom - \muupperflownom \right) \, \capacityflow
-%     \Forall{\ell \in L}
-% \end{align}
-
-% % \begin{align}
-% %     \muupperflownom \left( \capacityflow - \capacityflowUpper \right) =0 \\
-% %     \mulowerflownom \left( \capacityflow - \capacityflowLower \right) =0 \\
-% % \end{align}
-
-% % \begin{align}
-% %     \pdv{\lagrangian}{\capacityflow}  = 0 \,\, \rightarrow \,\, \\
-% %     \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
-% %     \label{eq:capex_flow_duality_bf}
-% %     \end{align}
-% %     for transmission lines.
-% % Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacitygeneration - \capacitygenerationlower \right) $
-% % \begin{align}
-% %   \capitalpricegeneration \, \capacitygeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacitygenerationlower \right) + \muuppergenerationnom \left( \capacitygenerationupper - \capacitygenerationlower \right) \Forall{n,s}
-% % \end{align}
-
-\section{Allocation Schemes}
-\subsection{Allocating Gross Injections with EBE}
-\label{sec:gross_ebe}
-
-The allocation of gross generation to demands $\demand$ is straightforwardly obtained by a proportional distribution of the generation, \ie
-
-\begin{align}
-    \allocategeneration = \dfrac{\generation}{\sum_s \generation} \demand 
-\end{align}
-
-\subsection{Allocating Net Injections with EBE}
-\label{sec:net_ebe}
-
-Allocating net power injections using the EBE methods leads to the same result as the Marginal Participation (MP) \cite{rudnick_marginal_1995}  algorithm when allocating to consumers only, see \cite{hofmann_flow_2020-1} for further insight.
-We calculate it by setting
-\begin{align}
-\allocatepeer &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
-\label{eq:mp_slack}
-\end{align}
-where 
-\begin{itemize}
-%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
-\item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
-\item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
-\item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
-\item $\gamma_t = \left( \sum_n \netproduction\right) ^{-1} = \left( \sum_n \netconsumption\right) ^{-1}$ is the inverse of the total injected/extracted power at time $t$.
-\end{itemize}
-
-The allocation $\allocategeneration$ from generator $s$ to $n$, is given by multiplying $\allocatepeer$ with the nodal share $\generation / \nodalgeneration$.
-
-\subsection{Allocating Net Power using AP}
+\section{Power Allocation}
 \label{sec:net_ap}
@@
 
-\subsection{Allocating Gross Power using AP}
-\label{sec:gross_ap}
-
-We use the same allocation as in \cref{sec:net_ap} but replace the net nodal production $\netproduction$ by the gross nodal production $\nodalgeneration$ which leads to  
-\begin{align}
-\InverseAPInjection_t = \left( \diag{g} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
-\end{align}
-The distributed slack is for time step $t$ is then given by
-\begin{align}
-\allocategeneration[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
-\end{align}
-
-% 
 \section{Working Example}
@@
     \includegraphics[width=\linewidth]{de50bf/power_mix_ptpf_net}
-    \caption{Average power mix per region calculated by Average Participation \ref{ap}.
+    \caption{Average power mix per region calculated by Average Participation.
 Coastal regions are mainly supplied by local offshore and onshore wind farms.

### 2020-10-19 paper/flow_based.tex
Further literature review; add limitation section

@@
 Today's power systems are subject to a deep and ongoing transformation.
-The needed shift to from controllable to weather-driven power generation, the constant improvement and innovation of technology require rigorous system planning and international cooperation.
+The needed shift to from controllable to weather-driven power generation, the constant improvement and innovation of technology require rigorous system planning and international cooperation \cite{pfenninger_energy_2014,schlachtberger_benefits_2017}.
 The core of the problem manifests in the costs.
@@
 Cost-driving aspects or market players should be detected and addressed appropriately.
-In this context, power system modelling has increasingly gained attention during the last years.
+In this context, power system modelling has increasingly gained attention during the last years \cite{bazmi_sustainable_2011,pfenninger_energy_2014,pereira_generation_2017}.
 Many studies show how cost-efficiency and renewable energy are brought together.
@@
 The presented approach builds upon two basic concepts: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
-Secondly, \textit{flow tracing}, following Bialek's Average Participation, assigns the operation of single network assets to consumers in an plausible manner.
-The following shows how combining these allow for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network.
+Secondly, \textit{flow tracing}, following Bialek's Average Participation (AP), assigns the operation of single network assets to consumers in an plausible manner.
+Combining these two relation allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
+Cost allocation based on flow allocation was already discussed and performed in various works.
+Shahidehpour et al.
+give a deep insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} based on different allocation techniques.
+Based on Generation Shift Factors, \ie the marginal contribution of generators to a flow on line, the Locational Marginal Price (LMP) can be represented as a superposition of LMP at the reference bus, the price for congestion and a price for losses.
+\cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however does not reflect the exact LMP of an optimized network.
+A similar approach was taken in \cite{tranberg_flow-based_2018} where electricity prices of a non-optimal power dispatch were allocated using flow tracing.
 
@@
 where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint.
-% When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that $\partial \storagecharge / \partial \demand \cdot \demand \Leftrightarrow 0$, implying that the when a storage charges power, it does not supply any demand.
-At these times, it the storage is a effective consumer retrieving power from producing assets.
-% This leaves us with 
-% \begin{align}
-    %      \allocatecapexstorage = \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \allocatestoragedispatch
-% \end{align}
-% where the dispatch allocation only assigns discharge power to demand $\demand$, thus
-% \begin{align}
-%     \allocatestoragedispatch \Leftrightarrow \pdv{\storagedispatch}{\demand} \demand
-% \end{align}
-% Note that this relation will break \cref{eq:cost_payback} as the payments to $r$ surpass the CAPEX by an amount $\remainingcost^E_r$. 
-% % \begin{align}
-    % %     \capexstorage_r + \remainingcost^E_r = \sum_n \allocatecapexstorage
-% % \end{align} 
-% It is crucial to note that storage units perform a redistribution of money, and therefore distort the cost allocation.
-So, a certain share of what is allocated to the CAPEX of a storage is in another time step spent by the same storage in order to buy power from other assets.
-This effect scales with the amount of installed storage capacity.
-
-% We shortly annotate that it is possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
-Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$.
-The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.
+Following the considerations in \cref{sec:zero_profit_storage_units} we restrict to the revenue from dispatched power, \ie the first term in \cref{eq:no_profit_capex_storage}, for the cost allocation.
+
 \end{subequations}
@@
 
-After showing that cost terms $\cost$ for OPEX and CAPEX can be represented by cost-weighted sum of the operational state $\state$, we now demonstrate how the operational state in turn is allocated to consumers.
+After representing the cost terms $\cost$ for OPEX and CAPEX by a cost-weighted sum of the operational state $\state$, we now show how the operational state in turn is allocated to consumers.
 This builds the second crucial ingredient of the cost allocation.
@@
 \begin{align}
-    \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) \Forall{\ell \in L, n, t}
+    \allocateflow = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n m} \demand \right) \Forall{\ell \in L, n, t}
  \label{eq:allocate_flow}
@@
 
+\section{Limitations and possible Extensions}
+
+The presented cost allocation is based on the linear power flow approximation.
+Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
+Here the AP scheme traces active power flows and their losses as done in \cite{meng_investigation_2007}.
+This introduces an additional cost term $\remainingcost^\text{Loss}$ compensated by the consumers and mirroring in the LMP.
+
+The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
+
+In the presented work, we restricted the application to long term investment models.
+However, the cost allocation can be applied to short term planning models with fixed capacities as well.
+The revenue from the capacity limits then builds basis for future investments.
+
 \section{Conclusion}
 
-In this work we presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from asset to consumer.
-Starting from an optimized long-term investment model, we were able to decompose single cost terms into contributions of single consumers.
+We presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from assets to consumers.
+Within a ling-term equilibrium OPEX and CAPEX of each asset are payed back by the operation based revenue.
+Using flow allocation, we are able to allocate the operation and therefore the assigned costs of assets to consumers in the network.
 For three typical classes of assets, namely generators, transmission lines and storage units, we showed how operational prices and shadow prices must be weighted with the dispatch allocation in order to allocate all system costs.
-Further we highlighted the impact of additional design constraints, that is constraints with a non-zero constant on the right hand side, such as lower and upper capacity expansion limits.
-These alter the locational marginal price and therefore distort the cost allocations.
-In the case of upper capacity expansion limits, this leads to an additional charge for consumers which have to compensate for \eg land use restrictions.
-For lower capacity investment bounds, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
-\dots
+Further we highlighted the impact of minimum capacity requirements and maximum installation potentials.
+These alter the revenues per asset and therefore the cost allocations.
+For lower capacity requirements, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
+Contrary, upper capacity expansion limits lead to an additional charge for consumers which have to compensate for \eg land use restrictions.
 
@@
 \begin{align}
-\muuppergeneration \left( \generation - \generationpotential \, \capacitygeneration \right)  &= 0  \Forall{n,s,t} 
+\muuppergeneration \left( \generation - \generationpotential \, \capacitygeneration \right)  &= 0  \Forall{s,t} 
 \label{eq:complementary_slackness_upper_generation} \\
-\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
+\mulowergeneration  \, \generation &= 0 \Forall{s,t}
 \label{eq:complementary_slackness_lower_generation} 
@@
 \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\, 
-\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
+\capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{s}
 \label{eq:capex_generation_duality}
@@
 \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
-\operationalpricegeneration =  \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
+\operationalpricegeneration =  \sum_n \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{s} \label{eq:opex_duality}
 \end{align}
@@
 \begin{align}
- \capitalpricegeneration \, \capacitygeneration  = \sum_t \muuppergeneration \, \generation 
+ \capitalpricegeneration \, \capacitygeneration  = \sum_t \muuppergeneration \, \generation \Forall{s} 
  \label{eq:capital_price_generation_sum}
@@
 \begin{align}
-  \capitalpricegeneration \, \capacitygeneration + \sum_t \operationalpricegeneration \generation = \sum_t \lmp \incidencegenerator \generation
+  \capitalpricegeneration \, \capacitygeneration + \sum_t \operationalpricegeneration \generation = \sum_{n,t} \lmp \incidencegenerator \generation \Forall{s}
 \end{align}
@@
 \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
-\label{eq:capexFlow_duality}
+\label{eq:capacity_flow_duality}
 \end{align}
@@
     0 &= \pdv{\lagrangian}{\flow}  \\ 
-    0 &= - \sum_n \incidence \lmp  + \cycleprice \cycle \reactance  - \muupperflow + \mulowerflow \Forall{n,s} \label{eq:opex_flow_duality}
-\end{align}
-
-When multiplying \cref{eq:capexFlow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain 
-\begin{align}
- \capitalpriceflow \, \capacityflow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow
+    0 &= - \sum_n \incidence \lmp  + \sum_c \cycleprice \cycle \reactance  - \muupperflow + \mulowerflow \Forall{\ell, t} \label{eq:flow_duality}
+\end{align}
+
+When multiplying \cref{eq:capacity_flow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain
+\begin{align}
+ \capitalpriceflow \, \capacityflow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow \Forall{\ell} 
  \label{eq:capital_price_flow_sum}
 \end{align}
-Again we can use this to formulate the zero-profit rule for transmission lines. We multiply \cref{eq:opex_flow_duality} with $\flow$, which finally leads us to 
-\begin{align}
-\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \cycleprice\, \cycle\, \reactance\, \flow 
-\end{align}
+Again we can use this to formulate the zero-profit rule for transmission lines. We multiply \cref{eq:flow_duality} with $\flow$, which finally leads us to 
+\begin{align}
+\capitalpriceflow \, \capacityflow = - \sum_n \incidence\, \lmp\, \flow + \sum_c \cycleprice\, \cycle\, \reactance\, \flow 
+\Forall{\ell} \end{align}
 % Watch out Incidence = - ordinary Incidence !!
@@
 \begin{align}
-    \storagedispatch - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragedispatch} \\
-    \storagecharge - \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragecharge} \\
-    \storagesoc - h_r \, \capacitystorage &\le 0 \Forall{r,t} \resultsin{\muupperstoragesoc}
+    \storagedispatch - \capacitystorage &\le 0  \resultsin{\muupperstoragedispatch} \Forall{r,t} \\
+    \storagecharge - \capacitystorage &\le 0  \resultsin{\muupperstoragecharge} \Forall{r,t} \\
+    \storagesoc - h_r \, \capacitystorage &\le 0  \resultsin{\muupperstoragesoc} \Forall{r,t}
 \end{align}
@@
 \begin{align}
-    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
+    \notag
+    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage &\left(\storagedispatch - \storagecharge \right) \Forall{r,t}
 \end{align}
@@
     \pdv{\lagrangian}{\storagedispatch} &= 0 \\
-    \operationalpricestorage -  \lmp \, \incidencestorage - \mulowerstoragedispatch + \muupperstoragedispatch + (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
+    \operationalpricestorage - \sum_n \lmp \, \incidencestorage - \mulowerstoragedispatch + \muupperstoragedispatch + (\efficiencydispatch )^{-1} \mustateofcharge &= 0 \;  \forall r,t
     \label{eq:stationarity_storagedispatch}
@@
     - \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
+    \label{eq:no_profit_capex_storage2}
 \end{align}
@@
 
-\subsection{Emission Constraint}
-
-Imposing an additional CO$_2$ constraint limiting the total emission to K,  
-\begin{align}
-\sum_{n,s,t} \emission \, \generation \le \text{K} \resultsin{\emissionprice} 
-\label[constraint]{eq:co2_constraint}
-\end{align}
-with $\emission$ being the emission factor in tonne-CO$_2$ per \megawatthour, returns an effective CO$_2$ price $\emissionprice$ in \euro/tonne-CO$_2$. 
-% The CO$_2$ price shifts the right hand side of the non-profit relation for generators \cref{eq:non_profit_generator} to
-% 
-% \begin{align}
-% \capitalpricegeneration \, \capacitygeneration + \sum_{t} \operationalpricegeneration \, \generation &= \sum_{t} \left( \lmp - \emission \, \emissionprice \right)  \, \generation \Forall{n,s} 
-% \label{eq:non_profit_generator_emission}
-% \end{align}
-% This shows nicely the duality for exchanging the CO$_2$ \cref{eq:co2_constraint} for a shifted OPEX which includes the CO$_2$ costs
-As shown in ... the constraint can be translated in a dual price which shift the operational price per generator
-\begin{align}
-\operationalpricegeneration \rightarrow \operationalpricegeneration + \emission \, \emissionprice \label[relation]{eq:shift_opex_by_emission_cost}
+When applying the cost allocation scheme \cref{eq:general_scheme}, it stands to reason to assume that when a storage charges power, it does not supply any demand.
+Thus consumers only pay storage units in times the storage dispatches power.
+Hence, we restrict the allocatable revenue per storage unit to the first term in \cref{eq:no_profit_capex_storage,eq:no_profit_capex_storage2}.
+This allocates then the CAPEX of $r$ plus the costs $\remainingcost^E_r$ it needs to buy the charging power,
+\begin{align}
+        \capexstorage_r + \remainingcost^E_r = \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch 
+\end{align} 
+In charging times the total of remaining costs $\remainingcost^E_r$ is spent to power from other assets. These costs scale with the amount of installed storage capacity.
+Note that it would be possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}.
+Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$.
+The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.
+
+\subsection{Proof: Equivalence of local and imported prices}
+
+We start with \cref{eq:flow_duality} which we recall here, 
+\begin{align}
+    \notag
+    0 = &- \sum_m \incidence[m] \lmp[m]  + \sum_c  \cycleprice \cycle \reactance  \\
+    &- \muupperflow + \mulowerflow \Forall{\ell,t} 
+\end{align}
+It states that the price difference between two adjacent buses minus the price for the KVL, is the revenue per line $\ell$, $\left(- \muupperflow + \mulowerflow\right)$.
+We multiply the equation by the flow allocation $\allocateflow$ and obtain
+\begin{align}
+    \notag
+    0 =& - \allocateflow \sum_m \incidence[m] \lmp[m]  \\
+    \notag
+    &+ \allocateflow \sum_c  \cycleprice \cycle \reactance  \\
+    &- \allocateflow \left(\muupperflow - \mulowerflow\right) \Forall{\ell,t} 
+    \label{eq:flow_duality2}
+\end{align}  
+The allocation $\allocateflow$ defined in \cref{eq:allocate_flow}follows the linear power flow laws. We slightly reformulate the expression to
+\begin{align}
+    \allocateflow = \sum_{m'} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \Forall{\ell, n, t}
+    \label{eq:allocate_flow2}
+\end{align}
+and insert it into \cref{eq:flow_duality2}. When taking the sum over all lines $L$, the first term yields  
+\begin{align}
+    \notag
+    &- \sum_{\ell, m'} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \sum_m \incidence[m] \lmp[m] \\
+    \notag
+    =& - \sum_{\ell, m', m} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \incidence[m] \lmp[m] \\
+    \notag
+    =& - \sum_{m', m} \delta_{m m'} \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \lmp[m] \\
+    \notag
+    =& - \sum_{m} \left( \allocatepeer - \delta_{n m} \demand \right) \lmp[m] \\
+    =& - \sum_{m} \allocatepeer \lmp[m] + \demand \lmp 
+\end{align}
+where in the third step we used the relation $\sum_\ell \ptdf \incidence[m] = \delta_{n m}$.
+The second term in \cref{eq:flow_duality2} vanishes as the basis cycles $\cycle$ are the kernel of the PTDF, $\sum_{\ell} \cycle \ptdf  = 0 \; \forall c,n$. Thus, we end up with 
+\begin{align}
+    \notag
+    \demand \lmp =& \sum_{m} \allocatepeer \lmp[m] \\
+    +& \sum_\ell \allocateflow \left(\muupperflow - \mulowerflow\right) \Forall{n,t}
+\end{align}
+This relation shows that for any P2P allocations $\allocatepeer$ the combined price of the imported power is always be the same as the locational price.
+This holds only true if the flow allocation $\allocateflow$ follow the power flow laws, \ie are defined as in \cref{eq:allocate_flow,eq:allocate_flow2}.
+
+Naturally, the power production of the supplying node $m$ decomposes into contributions of assets, following the definition in \cref{eq:allocate_production}.
+At the same time, the LMP at $m$ decomposes into asset related prices, using \cref{eq:opex_duality,eq:stationarity_storagedispatch}.
+This finally reproduces the allocations of \cref{tab:cost_allocation_map} and results in
+\begin{align}
+    \demand \, \lmp =  \sum_{\circ, i} \allocatecost 
 \end{align}
@@
 \end{align}
-where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( \flow \right)  \incidence$.
-Then the distributed slack for time step $t$ is given by
+where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( f_\ell \right)  \incidence$.
+Then the P2P allocation for time step $t$ is given by
 \begin{align}

### 2020-10-19 paper/flow_based.tex
finish intro

@@
 This study fills this gap.
-On the basis of an optimized network the operational state of each time step and corresponding characteristics are are in detailed considered to allocate all system costs.
+On the basis of an optimized network the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
 The presented approach builds upon two basic concepts: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
@@
 Combining these two relation allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
-Cost allocation based on flow allocation was already discussed and performed in various works.
+Flow-based cost allocation was already discussed and performed in various works \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,tranberg_flow-based_2018,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
 Shahidehpour et al.
-give a deep insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} based on different allocation techniques.
+give a deep insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} with different allocation techniques.
 Based on Generation Shift Factors, \ie the marginal contribution of generators to a flow on line, the Locational Marginal Price (LMP) can be represented as a superposition of LMP at the reference bus, the price for congestion and a price for losses.
-\cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however does not reflect the exact LMP of an optimized network.
+The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however does not reflect the exact LMP of an optimized network.
 A similar approach was taken in \cite{tranberg_flow-based_2018} where electricity prices of a non-optimal power dispatch were allocated using flow tracing.
 
+The following work brings advantages of all the mentioned studies together.
+It reassures localized cost allocations while fully aligning payments to the nodal pricing scheme based on the LMP.
+The work serves to facilitate transparency and cost-benefit analysis in network plannings such as the Ten Year Network Development Plan \cite{entso-e_completing_2020} or the German Netzentwicklungsplan (NEP) \cite{bundesnetzagentur_netzentwicklungsplan_2020}.
+Further, it builds a basis of a usage based transmission cost allocation.
+The first section which presents the full method splits into: deriving the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  presenting the AP allocation scheme and resulting allocation from assets to consumers (\cref{sec:dispatch_allocations}); presenting the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
+The second part in \cref{sec:application_case} applies the cost allocation to a optimized German power system with high shares of renewable power sources and evaluates the allocated costs.
+
 \section{Flow-based Cost Allocation}
@@
 \subsection{Zero-Profit Rule}
+\label{sec:zero_profit_rule}
 
@@
 \subsection{Allocation of Power Dispatch}
-\label{sec:localizing_allocations}
+\label{sec:dispatch_allocations}
 \begin{subequations}
@@
 Complementary to \cref{eq:allocate_production}, this allocation indicates the power flow on line $\ell$ and time $t$ which is ending up at node $n$ and therefore consumed by $\demand$.  
-Note that the assignment of line flows to consumers is different from the traced flow returned by the AP scheme, as we're only taking the peer-to-peer relation $\allocatepeer$ and, for reasons to be yet to come, from there calculate the power flow .
+Note that the assignment of line flows to consumers is different from the traced flow returned by the AP scheme.
+As pointed out in \cite{galiana_transmission_2003} the AP scheme, does not obey the Kirchhoff Voltage Law but only the Kirchhoff Current Law.
+As we show later and in \cref{sec:proof_equivalence} a consideration of both is needed in order to align the allocated costs with the Locational Marginal Prices.
 
@@
 \end{align}
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$.
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
 Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
 At the optimum the LMP is given by a shadow price of the nodal balance constraint (see \cref{sec:lmp} for details).
-\Cref{eq:payment} which is in detailed proved in ... connects the cost allocation to a nodal pricing scheme implied by the LMP.
+\Cref{eq:payment} which is in detailed proved in \cref{sec:proof_equivalence} connects the cost allocation to a nodal pricing scheme implied by the LMP.
 
@@
 \section{Application Case}
+\label{sec:application_case}
 % 
@@
 \subsection{Proof: Equivalence of local and imported prices}
+\label{sec:proof_equivalence}
 
@@
 This relation shows that for any P2P allocations $\allocatepeer$ the combined price of the imported power is always be the same as the locational price.
-This holds only true if the flow allocation $\allocateflow$ follow the power flow laws, \ie are defined as in \cref{eq:allocate_flow,eq:allocate_flow2}.
+The representation matches the findings in \cite{wu_locational_2005}.
+However, the latter builds its formulation on a evenly distributed slack, which translated to a peer-to-peer allocation $\allocatepeer$ corresponding to the non-local Equivalent Bilateral Exchanges \cite{galiana_transmission_2003}.
+However, the representation here holds only true any $\allocatepeer$ if the corresponding flow allocation $\allocateflow$ follow the power flow laws, \ie are defined as in \cref{eq:allocate_flow,eq:allocate_flow2}.
 
 Naturally, the power production of the supplying node $m$ decomposes into contributions of assets, following the definition in \cref{eq:allocate_production}.
-At the same time, the LMP at $m$ decomposes into asset related prices, using \cref{eq:opex_duality,eq:stationarity_storagedispatch}.
+At the same time, the LMP at $m$ decomposes into asset related prices (\cref{eq:opex_duality,eq:stationarity_storagedispatch}).
 This finally reproduces the allocations of \cref{tab:cost_allocation_map} and results in

### 2020-10-19 paper/flow_based.tex
review some parts, still missing: abstract, graphical abstract, acknowlegdement, review

@@
 \usepackage[colorlinks]{hyperref}
-\usepackage[backend=biber, bibencoding=utf8]{biblatex}
+\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8]{biblatex}
 \usepackage{eurosym}
@@
 
+\twocolumn[
+  \begin{@twocolumnfalse}
+
     \maketitle
@@
     \begin{abstract}
-% Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and ... 
+        Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling.
+It leads to perfectly scheduled operations and investments while securing all power demands.
+\dots
+        This work introduces a peer-to-peer cost allocation for welfare-maximized systems   \dots
+        This opens the possibility to transparently integrate transmission constraints and investment costs in national and international development plans.
     \end{abstract}
+
+\end{@twocolumnfalse}
+]
 
@@
 \begin{itemize}
- \item Flow Allocation schemes determine peer-to-peer (P2P) allocation between sources and sinks 
- \item P2P allocation can be used to assign payments of consumers to single network assets
- \item Using shadow prices, Capital Expenditures are accurately allolcated to users which profit from investments
- \item The P2P cost assignments are in alignment with the nodal pricing scheme of a welfare-maximized system   
- \item This opens the possibility to transparently integrate transmission constraints and investment costs in national and international development plans 
- \item Applying to a German future scenario shows how higher electricity costs arise for buses which rely on investments in transmission and local backup generation  
+    \item In long-term investment models every network asset gains back its variable and investment cost from its operation based revenue
+    \item Flow tracing is used as a basis to allocate the operation of assets to consumers. 
+    \item Using operational and shadow prices, the allocations assign all system costs in an P2P manner.
+    \item The P2P cost assignments are locally constraint while being in alignment with the nodal pricing scheme of a welfare-maximized system.    
+    \item Applying to a German future scenario shows how higher electricity costs arise for buses relying on investments in transmission and local backup generation  
 \end{itemize}
@@
 Cost-driving aspects or market players should be detected and addressed appropriately.
-In this context, power system modelling has increasingly gained attention during the last years \cite{bazmi_sustainable_2011,pfenninger_energy_2014,pereira_generation_2017}.
+In this context, power system modelling has increasingly gained attention during the last years \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017}.
 Many studies show how cost-efficiency and renewable energy are brought together.
@@
 \\
-This study fills this gap.
+This study bridges this gap.
 On the basis of an optimized network the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
 The presented approach builds upon two basic concepts: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
-Secondly, \textit{flow tracing}, following Bialek's Average Participation (AP), assigns the operation of single network assets to consumers in an plausible manner.
-Combining these two relation allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
+Secondly, flow tracing, following Bialek's Average Participation (AP) \cite{bialek_tracing_1996}, can be used to assigns the operation of single network assets to consumers in an plausible and locally constraint manner.
+Combining these two relations allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
 Flow-based cost allocation was already discussed and performed in various works \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,tranberg_flow-based_2018,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
@@
 Further, it builds a basis of a usage based transmission cost allocation.
-The first section which presents the full method splits into: deriving the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  presenting the AP allocation scheme and resulting allocation from assets to consumers (\cref{sec:dispatch_allocations}); presenting the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
+
+The first section which presents the full method splits into the presentation of: the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  the AP allocation scheme and resulting allocation from assets to consumers (\cref{sec:dispatch_allocations}); the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
 The second part in \cref{sec:application_case} applies the cost allocation to a optimized German power system with high shares of renewable power sources and evaluates the allocated costs.
@@
 \end{align}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\linewidth]{cost_decomposition.png}
+\caption{Schematic decomposition of the total system cost $\totalcost$ in an long-term investment model.}
+\label{fig:cost_decomposition}
+\end{figure}
 In turn, each of the terms $\cost$ consists of the costs associated to an asset $i$ in the system, 
@@
 If $\cost_i$ describes the OPEX occasioned by asset $i$, the cost factor $\costfactor$ is simply given by the marginal operational price $o_i$.
-However, as we will show in the following, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given at the cost-optimum.
-
-For a detailed demonstration, we derive  \cref{eq:cost_decomposition} for each cost term separately while considering generators, transmission lines and storages in the network.
-Therefore, let $o_i$ denote the operational price per MWh for asset $i$ and $c_i$ the capital price for one MW capacity expansion.
+However, as we will show in the following, if it describes the CAPEX of asset $i$, $\costfactor$ is a composition of shadow prices $\mu_{i,t}$ given from the corresponding constraint at the cost-optimum.
+
+For a detailed demonstration, we derive  \cref{eq:cost_decomposition} for generators, transmission lines and storages separately.
+Therefore, let $o_i$ denote the operational price per MWh of asset $i$ and $c_i$ the capital price for one MW capacity expansion.
 \Cref{tab:cost_allocation_map} summarizes all derived relations, these can be inserted into \cref{eq:asset_cost,eq:cost_decomposition} for each single cost term.
@@
 \end{align}
-The constraint yields a shadow-price of $\muuppergeneration$ in literature often denoted as the Quality of Supply.
-It can be interpreted as the the price per MW that \cref{eq:upper_generation_capacity_constraint} imposes to the system.
+The constraint yields a shadow-price of $\muuppergeneration$ in literature often denoted as the Quality of Supply \cite{schweppe_spot_1988}.
+It can be interpreted as the price per MW that \cref{eq:upper_generation_capacity_constraint} imposes to the system.
 If $\muuppergeneration$  is bigger than zero, the constraint is binding, which pushes investments in $\capacitygeneration$.
-As shown in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
+As shown in \cite{brown_decreasing_2020} and in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
 \begin{align}
@@
 This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}.   
-%If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-
-% \subsection{CAPEX Allocation}
-% For the CAPEX allocation, it becomes crucial to look at the individual relations between operational state $\state$ and its binding constraints.
-
-% \subsubsection{Generators}
-
-% The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
-% \begin{align}
-% \generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
-% \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
-% - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
-% \label[constraint]{eq:lower_generation_capacity_constraint}
-% \end{align}
-% where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators.
-At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.
-As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
-% \begin{align}
-%  \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
-%  \label{eq:no_profit_capex_generation}
-% \end{align}
-% This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}. The resulting the CAPEX allocation is given by
-% \begin{align}
-%  \allocatecapexgeneration = \muuppergeneration \, \allocategeneration
-%  \label{eq:allocate_capexGeneration_detailed}
-% \end{align}
-% How does this allocation behave?
-According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not.
-If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding, which pushes investments in $\capacitygeneration$.
-If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary.
-% When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} , we can use \cref{eq:state_allocation_constraint,eq:no_profit_capex_generation} to see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.
-
 \end{subequations}
@@
 At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-Again, we use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
+We use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
 \begin{align}
@@
 \begin{subequations}
-
-% The complementary slackness of the above constraints,  
-% \begin{align}
-%     \capexstorage_r = \capitalpricestorage \, \capacitystorage = \sum_t \muupperstoragedispatch \, \storagedispatch + \muupperstoragecharge \, \storagecharge + \muupperstoragesoc \, \storagesoc
-% \end{align}
 
@@
 
-After representing the cost terms $\cost$ for OPEX and CAPEX by a cost-weighted sum of the operational state $\state$, we now show how the operational state in turn is allocated to consumers.
+The fact that all asset related costs $\cost_i$ can be represented as a cost-weighted sum of the operational state $\state$, prompts the question how the operational state $\state$, in turn, is allocated to consumers.
 This builds the second crucial ingredient of the cost allocation.
@@
 
-% The Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} for example assumes that every producer supplies every consumer proportional to its share in the total consumption.
-Despite the straightforward principal, this scheme does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}.
-The Flow Based Market Coupling (FBMC) is based on predefined zonal sensitivities for allocating intra-regional and inter-regional power flow.
-However, with its reliance on the mentioned sensitivity, the Generation Shift Keys, the scheme is strongly dependent on exogenous assumptions.
-
 The peer-to-peer allocations $\allocatepeer$ fulfill some basic properties.
@@
 
-Yet, we didn't touch the allocation from transporting assets $i \in L$ to consumers.
+Yet, we didn't touch the allocation of transporting assets $i \in L$ to consumers.
+Note that the traced flow based in the AP scheme does obey the Kirchhoff Current Law but not the Kirchhoff Voltage Law, as already pointed out in \cite{galiana_transmission_2003}.
+However, as we show later a consideration of both laws is needed in order to align the allocated costs with the optimized Locational Marginal Prices.
 To tackle this, let $\ptdf$ denote a linear mapping between the injection $\left(\nodalgeneration[m] - \demand[m]\right)$ and the flow $\flow$, such that
@@
 Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
-For transport models or networks with High Voltage Directed Current (HVDC) lines, these can calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
+For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
 Inserting \cref{eq:nodalgeneration} into \cref{eq:power_flow_equation} and expanding the sum yields 
@@
 \end{align}
-Complementary to \cref{eq:allocate_production}, this allocation indicates the power flow on line $\ell$ and time $t$ which is ending up at node $n$ and therefore consumed by $\demand$.  
-Note that the assignment of line flows to consumers is different from the traced flow returned by the AP scheme.
-As pointed out in \cite{galiana_transmission_2003} the AP scheme, does not obey the Kirchhoff Voltage Law but only the Kirchhoff Current Law.
-As we show later and in \cref{sec:proof_equivalence} a consideration of both is needed in order to align the allocated costs with the Locational Marginal Prices.
+Complementary to \cref{eq:allocate_production}, this allocation indicates the power flow on line $\ell$ and time $t$ which is finally consumed by $\demand$.
 
@@
 Using the presented relations, we are able to straightforwardly define the full cost allocation.
-Therefore, the expression in \cref{eq:decompose_state} is inserted in \cref{eq:cost_decomposition} which leads to
+Therefore, the expanded sum in \cref{eq:decompose_state} is inserted in \cref{eq:cost_decomposition} which leads to
 \begin{align}
@@
 The cost allocation entails a further important property.
-When summing over all assets $i$ and cost terms $\circ$, the cost allocation yields $\payment$, which we refer to as the nodal payment or payment.
-This exactly matches the payment determined by the Locational Marginal Price (LMP) $\lmp$, thus
+In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
+Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
+At the optimum this quantity is given by a shadow price of the nodal balance constraint (see \cref{sec:lmp} for details).
+Now, when summing over all assets $i$ and cost terms $\circ$, the cost allocation yields
 \begin{align}
@@
 \end{align} 
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
-Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
-At the optimum the LMP is given by a shadow price of the nodal balance constraint (see \cref{sec:lmp} for details).
-\Cref{eq:payment} which is in detailed proved in \cref{sec:proof_equivalence} connects the cost allocation to a nodal pricing scheme implied by the LMP.
+which we refer to as the nodal payment or payment.
+It exactly matches the payment determined by the LMP $\lmp$.
+Hence, \Cref{eq:payment} which is in detailed proved in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the nodal pricing scheme implied by the LMP.
 
@@
 
-% \subsection{Connection to Locational Marginal Prices}
-
-% % 
 % \begin{figure}[h]
@@
 
-% % 
-% This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
-% \begin{align}
-% \totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
-% \label{eq:total_revenue}
-% \end{align}
-% depending on the costs $\remainingcost$ which are independent of the nodal demand  
-% \begin{align}
-%  \pdv{\remainingcost}{\demand} = 0
-% \end{align}
-% % 
-% Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional request to the network design, such as capacity expansion limits or minimum share of one technology in the power mix.
-However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.
-
-% However the question of how the LMP can be decomposed into contributions of single cost terms $\cost_{i}$ associated with asset $i$ remains unanswered.
-This work aims at presenting and illustrating a an intuitive, peer-to-peer cost allocation including all network assets.
-
 \subsection{Design Constraints}
@@
 \end{align}
-% We define the allocated costs associated to cost term $\cost$ is then given by 
-% \begin{align}
-%     \bar{\cost} = \sum_i  \cost_i - \remainingcost_i
-%     \label{altered_cost}
-% \end{align}
+
 According to the nature of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, it is either larger, equal or lower than $\cost$.
@@
 \end{align}
-In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to incorporate them into the cost allocation.
+In the following we highlight two often used classes of designs constraints in the form of \cref{eq:design_constraint} and show how to consider them into the cost allocation.
 
@@
 \begin{subequations}
-In more realistic setups, generators, lines or other assets can only be built up to a certain limit.
-This might be due to land use restrictions or social acceptance considerations.
-%In the following give a general solution how the cost allocation may handle this.
+In real-world setups, generators, lines or other assets can often only be built up to a certain limit. This might be due to land use restrictions or social acceptance considerations. 
 When constraining the capacity $\capacity$ to an upper limit $\capacityupper$, in the form of 
@@
 \end{align}
-has to be subsidized by governments or communities.
-It is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost = \subsidycost_i $.
+has to be subsidized by governments or communities or is simply ignored when investments are amortized.
+Note, it is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost = \subsidycost_i $.
 \end{subequations}
@@
 % 
-% 
-% 
-% 
 \Cref{fig:example_allocation} shows the allocation $\allocatestate$ for $n$~=~1 and $n$~=~2 separately.
@@
 The left side shows existent renewable capacities, matching the total capacity for the year 2017, which serve as lower capacity limits for the optimization.
-The right side shows the capacity expansion of renewables as well as installation of backup gas power plants.
+The right side shows the capacity expansion of renewable resources as well as installation of backup gas power plants.
 The effective CO$_2$ price is set to 120 \euro per tonne CO$_2$ emission.}
@@
 We observe a relatively strong gradient from south (at roughly 92 \euro/MWh) to north (80 \euro/MWh).
-Regions with with little pre-installed capacity and capacity expansion, especially for renewables, tend to have higher prices.
+Regions with little pre-installed capacity and capacity expansion, especially with respect to renewable generation, tend to have higher prices.
 The node with the lowest LMP in the upper northwest, stands out through high pre-installed offshore capacities.
@@
 In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power.
-The average operational price for renewables is extremely low (0.2 \euro/MWh), thus they play an inferior role in the OPEX allocation.
+The average operational price for renewable generators is extremely low (0.2 \euro/MWh), thus they play an inferior role in the OPEX allocation.
 \\
@@
 
-The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewables and the transmission expansion limit.
+The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewable resources and the transmission expansion limit.
 These sum up approximately to 7.5 b\euro.
@@
 
-\section{Limitations and possible Extensions}
-
-The presented cost allocation is based on the linear power flow approximation.
-Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
-Here the AP scheme traces active power flows and their losses as done in \cite{meng_investigation_2007}.
-This introduces an additional cost term $\remainingcost^\text{Loss}$ compensated by the consumers and mirroring in the LMP.
-
-The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
-
-In the presented work, we restricted the application to long term investment models.
-However, the cost allocation can be applied to short term planning models with fixed capacities as well.
-The revenue from the capacity limits then builds basis for future investments.
-
-\section{Conclusion}
+\section{Conclusion and Outlook}
 
@@
 Contrary, upper capacity expansion limits lead to an additional charge for consumers which have to compensate for \eg land use restrictions.
+Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent, the cost allocation shows how
+
+The presented cost allocation is based on the linear power flow approximation.
+Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
+However, the AP scheme might not the right choice as the Kirchhoff Voltage Law cannot not be considered subsequently as done here in the linear case.
+Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it considers by default the two circuit laws.
+The full power flow introduces an additional cost term $\remainingcost^\text{Loss}$ accounting for the transmission loss which is compensated by the consumers and mirrors in the LMP.
+
+% The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
+
+In the presented work, we restricted the application to long term investment models with perfect foresight.
+However, the cost allocation can as well be applied to short term planning models with fixed capacities.
+The revenue from the capacity limits then builds basis for future investments.
 
@@
 The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution.
-In hours of low CAPEX allocation (after the second drop) spendings for OCGT start to increase again.
+In hours of low CAPEX allocation (after the second drop) spending for OCGT start to increase again.
 These time-steps however play a minor role.}
@@
     \caption{Average \textbf{allocated scarcity cost} per consumed MWh, $\sum_t \allocatescarcitycost / \sum_t \demand$.
-These cost result from land use restricitions for offshore wind, onshore wind,  solar, run-of-river.
+These cost result from land use restrictions for offshore wind, onshore wind,  solar, run-of-river.
 The cost per MWh are indicated by the color of a region.

### 2020-10-19 code/examples.py

@@ -169,6 +169,7 @@ for bus in n.buses.index:
 
     bbox.update({'facecolor': fcolor})
     A = r'$A_{1,%s} = %+i$ MW'%(bus, round(subflow.item()))
+    A = A.replace('+', '')
     ax.text(*n.buses[['x', 'y']].mean() + [0, 0.2], A, zorder=8,
             ha='center', va='center', color='white',
             bbox=bbox)

### 2020-10-19 code/price_decomposition.py

@@ -58,3 +58,6 @@ ax.set_yticklabels([])
 
 fig.tight_layout()
 fig.savefig('../figures/price_decomposition.png', bbox_inches='tight')
+
+ax.set_ylabel('Total System Cost')
+fig.savefig('../figures/cost_decomposition.png', bbox_inches='tight')

### 2020-10-20 paper/flow_based.tex
add graphical abstract; abstract; ackowledgement etc.

@@
     \begin{abstract}
-        Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling.
-It leads to perfectly scheduled operations and investments while securing all power demands.
-\dots
-        This work introduces a peer-to-peer cost allocation for welfare-maximized systems   \dots
-        This opens the possibility to transparently integrate transmission constraints and investment costs in national and international development plans.
+        Cost minimization is a common approach in energy system modelling.
+In long-term studies, it leads to perfectly scheduled operations and investments while securing all power demands.
+System needs reports from regional, national and international agencies increasingly rely on cost minization techniques.
+However, cost impact to single regions and thefore the question of who profits to what extent from an investment is often neglected.
+        This work introduces a peer-to-peer cost allocation for cost-minimized system models.
+It is based on two basic principal.
+The first being that in a long-term equilibrium the expenses occasioned by a network asset can be expressed as a function of its operation.
+The second is that using flow allocation techniques, the operationel state can be allocated to consumers.
+        The cost allocation reflects the behaviour of the optimized Locational Marginal Prices while allocating costs in a plausible and spatially constraint manner.
+It is showcased by means of an optimized German power system with an carbon tax of 120 \euro\, per tonne CO$_2$.
+        % Applying it to an optimized German with a carbon tax of 120 \euro\, per tonne CO$_2$, network shows that transmission and backup investments capacity drive the prices for regions remote from
+        % This opens the possibility to transparently integrate transmission constraints and investment costs in system development plans and possibly into the power market. 
+        \vtop{%
+         \vskip0pt
+          \hbox{%
+          \includegraphics[width=\linewidth]{graphical_abstract.png}
+          }%
+        }
+
     \end{abstract}
@@
 \begin{itemize}
-    \item In long-term investment models every network asset gains back its variable and investment cost from its operation based revenue
+    \item In long-term equilibria a network asset gains back its variable and investment cost from its operation based revenue
     \item Flow tracing is used as a basis to allocate the operation of assets to consumers. 
-    \item Using operational and shadow prices, the allocations assign all system costs in an P2P manner.
+    \item Allocated flows have to be reshaped such that the Kirchhoff Current Law and the Kirchhoff Voltage Law are respected. 
+    \item Using operational and shadow prices from constraints, all costs are assigned in an P2P manner.
     \item The P2P cost assignments are locally constraint while being in alignment with the nodal pricing scheme of a welfare-maximized system.    
-    \item Applying to a German future scenario shows how higher electricity costs arise for buses relying on investments in transmission and local backup generation  
+    % \item Upper and lower capacity expansion limits modify the revenue per asset, which however correctly reflects in the cost allocation
+    \item In a low-carbon scenario for Germany regions with high renewable potentials profit from investment compensated by remote buses.   
 \end{itemize}
@@
 
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms $\cost$.
-Typically, these include operational expenditures for generators $\opexgeneration$, expenditures for emissions $\emissioncost$,  capital investments for the transmission system $\capexflow$ and so on, \ie
-\begin{align}
-\totalcost &= \sum_\circ \cost = \opexgeneration + \emissioncost +  \capexflow + ...
-\label{eq:total_cost}
-\end{align}
 \begin{figure}[h]
@@
 \end{figure}
+
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms $\cost$.
+Typically, these include operational expenditures for generators $\opexgeneration$, expenditures for emissions $\emissioncost$,  capital investments for the transmission system $\capexflow$ and so on, \ie
+\begin{align}
+\totalcost &= \sum_\circ \cost = \opexgeneration + \emissioncost +  \capexflow + ...
+\label{eq:total_cost}
+\end{align}
+
 In turn, each of the terms $\cost$ consists of the costs associated to an asset $i$ in the system, 
@@
 
+\section{Limitations}
+The presented cost allocation is based on the linear power flow approximation.
+Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
+However, the AP scheme might not the right choice as the Kirchhoff Voltage Law cannot not be considered subsequently as done here.
+Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it considers by default the two circuit laws.
+Allocating on the basis of the full power flow introduces an additional cost term $\remainingcost^\text{Loss}$ accounting for the transmission loss which is compensated by the consumers and mirrors in the LMP.
+
+The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
+
+In the presented work, we restricted the application to long term investment models with perfect foresight.
+However, the cost allocation can as well be applied to short term planning models with fixed capacities.
+The revenue from the capacity limits then builds basis for amortization and future investments.
+
+The optimization assumed a fix demand time series.
+As shown in \cref{fig:capex_duration_curve} this leads to high if not unrealistic LMP.
+Introducing a value of loss load as proposed in \cite{schroder_value_2015}  would screen away these and lead to more evenly distributed allocations.
+
 \section{Conclusion and Outlook}
 
-We presented a new cost-allocation scheme based on peer-to-peer dispatch allocations from assets to consumers.
-Within a ling-term equilibrium OPEX and CAPEX of each asset are payed back by the operation based revenue.
+A new cost-allocation scheme based on peer-to-peer dispatch allocations from assets to consumers was presented.
+Within a long-term equilibrium OPEX and CAPEX of each asset are payed back by the operation based revenue.
 Using flow allocation, we are able to allocate the operation and therefore the assigned costs of assets to consumers in the network.
@@
 Contrary, upper capacity expansion limits lead to an additional charge for consumers which have to compensate for \eg land use restrictions.
-Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent, the cost allocation shows how
-
-The presented cost allocation is based on the linear power flow approximation.
-Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
-However, the AP scheme might not the right choice as the Kirchhoff Voltage Law cannot not be considered subsequently as done here in the linear case.
-Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it considers by default the two circuit laws.
-The full power flow introduces an additional cost term $\remainingcost^\text{Loss}$ accounting for the transmission loss which is compensated by the consumers and mirrors in the LMP.
-
-% The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
-
-In the presented work, we restricted the application to long term investment models with perfect foresight.
-However, the cost allocation can as well be applied to short term planning models with fixed capacities.
-The revenue from the capacity limits then builds basis for future investments.
+Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent, the cost allocation shows how buses remote from wind farms pay higher prices due to increased reliance in transmission and backup capacity.
+On the other hand, buses with high renewable installation spend most payments to local assets.
+
+\subsubsection*{Reproducibility}
+
+All figures and number of the plot can be reproduced by using the \textit{snakemake} workflow in [].
+
+\subsection*{Funding}
+This research was funded by the by the German Federal Ministry for Economics Affairs and Energy in
+the frame of the NetAllok project \cite{bundesministerium_fur_wirtschaft_und_energie_verbundvorhaben_nodate}.
+
+\subsubsection*{Acknowledgement}
+
+Special thanks to Tom Brown for fruitful discussions and well-placed inputs.
 

### 2020-10-21 paper/flow_based.tex
update graphical abstract

@@
 \usepackage[colorlinks]{hyperref}
-\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8]{biblatex}
+\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8, style=numeric-comp]{biblatex}
 \usepackage{eurosym}
@@
     \begin{abstract}
-        Cost minimization is a common approach in energy system modelling.
-In long-term studies, it leads to perfectly scheduled operations and investments while securing all power demands.
-System needs reports from regional, national and international agencies increasingly rely on cost minization techniques.
+        Cost minimization of power systems leads to perfectly scheduled operations and investments while securing all power demands. 
+        System needs reports from regional, national and international agencies increasingly rely on cost minimization techniques.
 However, cost impact to single regions and thefore the question of who profits to what extent from an investment is often neglected.
-        This work introduces a peer-to-peer cost allocation for cost-minimized system models.
-It is based on two basic principal.
-The first being that in a long-term equilibrium the expenses occasioned by a network asset can be expressed as a function of its operation.
-The second is that using flow allocation techniques, the operationel state can be allocated to consumers.
-        The cost allocation reflects the behaviour of the optimized Locational Marginal Prices while allocating costs in a plausible and spatially constraint manner.
-It is showcased by means of an optimized German power system with an carbon tax of 120 \euro\, per tonne CO$_2$.
-        % Applying it to an optimized German with a carbon tax of 120 \euro\, per tonne CO$_2$, network shows that transmission and backup investments capacity drive the prices for regions remote from
-        % This opens the possibility to transparently integrate transmission constraints and investment costs in system development plans and possibly into the power market. 
+The literature provides multiple cost allocation methods dedicated to this problem, yet they hardly suit for large networks and an integration of all costs.
+        This work introduces a peer-to-peer allocation of all system costs in optimized models.
+Combining basic relations from economic optimization with flow allocation, its cost assignments locally constraint while still being aligned with the Locational Marginal Prices at the optimum.
+It is showcased by means of an optimized near-future scenario of the German power system.
         \vtop{%
@@
     \item Using operational and shadow prices from constraints, all costs are assigned in an P2P manner.
-    \item The P2P cost assignments are locally constraint while being in alignment with the nodal pricing scheme of a welfare-maximized system.    
+    \item The cost assignments are locally constraint and in alignment with the nodal pricing scheme.    
     % \item Upper and lower capacity expansion limits modify the revenue per asset, which however correctly reflects in the cost allocation
@@
 Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
-For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated or extended using the formulation presented in \cite{hofmann_flow_2020-1}.
+For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated or extended using the formulation presented in \cite{hofmann_flow_2020}.
 Inserting \cref{eq:nodalgeneration} into \cref{eq:power_flow_equation} and expanding the sum yields 

### 2020-10-21 code/examples.py

@@ -153,11 +153,15 @@ for bus in n.buses.index:
     fig, ax = plt.subplots(1, 1, figsize=(5, 2.5))
     subflow = ds.virtual_flow_pattern.sel(snapshot='now', bus=bus).to_series()
     p2p = ds.peer_to_peer.sel(sink=bus, snapshot='now').to_series()
+    if subflow.item() == 0:
+        n.plot(bus_sizes=p2p*3e-4, ax=ax, geomap=False, line_widths=1,
+               bus_colors=bcolor, line_colors=fcolor)
+    else:
+        n.plot(flow=subflow*5e-2, bus_sizes=p2p*3e-4, ax=ax, geomap=False,
+               bus_colors=bcolor, line_colors=fcolor)
 
-    n.plot(flow=subflow*5e-2, bus_sizes=p2p*3e-4, ax=ax, geomap=False,
-           bus_colors=bcolor, line_colors=fcolor)
-    bbox2 = dict(facecolor='lightgrey', alpha=0.5, edgecolor='None', boxstyle='circle')
-    ntl.plot.annotate_bus_names(n, ax, shift=0, bbox=bbox2)
+    bbox2 = dict(facecolor='lightgrey', alpha=1, edgecolor='None', boxstyle='circle')
+    ntl.plot.annotate_bus_names(n, ax, shift=0, bbox=bbox2,)
 
     for b in n.buses.index:
         bbox.update({'facecolor': bcolor, 'edgecolor': 'None', 'pad':.5, 'alpha':1})
@@ -173,7 +177,7 @@ for bus in n.buses.index:
     ax.text(*n.buses[['x', 'y']].mean() + [0, 0.2], A, zorder=8,
             ha='center', va='center', color='white',
             bbox=bbox)
-
+    ax.autoscale()
     fig.tight_layout(h_pad=0, w_pad=0)
     fig.savefig(f'{path}/example_allocation_bus{bus}_{power}_{method}.png',
                 bbox_inches='tight')

### 2020-10-22 paper/flow_based.tex
follow up

@@
 \newcommand{\allocatecost}[1][n \rightarrow i]{\cost_{#1, t}}
+\newcommand{\allocatecapex}[1][n \rightarrow s,t]{\mathcal{I}_{#1}}
 \newcommand{\allocatecapexgeneration}[1][n \rightarrow s,t]{\capexgeneration_{#1}}
@@
         This work introduces a peer-to-peer allocation of all system costs in optimized models.
-Combining basic relations from economic optimization with flow allocation, its cost assignments locally constraint while still being aligned with the Locational Marginal Prices at the optimum.
+Combining basic relations from economic optimization with flow allocation, its cost assignments are locally constraint while still being aligned with the Locational Marginal Prices at the optimum.
 It is showcased by means of an optimized near-future scenario of the German power system.
@@
 \begin{itemize}
-    \item In long-term equilibria a network asset gains back its variable and investment cost from its operation based revenue
+    \item In long-term equilibria a network asset recovers its variable and investment cost from its operation based revenue
     \item Flow tracing is used as a basis to allocate the operation of assets to consumers. 
@@
     \item Using operational and shadow prices from constraints, all costs are assigned in an P2P manner.
-    \item The cost assignments are locally constraint and in alignment with the nodal pricing scheme.    
+    \item The cost assignments are locally constraint and in alignment with the nodal pricing scheme of the optimum.    
     % \item Upper and lower capacity expansion limits modify the revenue per asset, which however correctly reflects in the cost allocation
-    \item In a low-carbon scenario for Germany regions with high renewable potentials profit from investment compensated by remote buses.   
+    \item In a low-carbon scenario for Germany, regions with high renewable potentials profit from investments compensated by remote buses.   
 \end{itemize}
@@
 Today's power systems are subject to a deep and ongoing transformation.
-The needed shift to from controllable to weather-driven power generation, the constant improvement and innovation of technology require rigorous system planning and international cooperation \cite{pfenninger_energy_2014,schlachtberger_benefits_2017}.
-The core of the problem manifests in the costs.
-Firstly, these should be as small as possible while meeting ecological and sociological standards.
-Secondly, they must be distributed in a fair and transparent manner.
-Cost-driving aspects or market players should be detected and addressed appropriately.
-In this context, power system modelling has increasingly gained attention during the last years \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017}.
-Many studies show how cost-efficiency and renewable energy are brought together.
-However, the question of how and, more importantly, upon what grounds costs are distributed was often left open.
+The shift from controllable to variable, weather-driven power generation as well as the constant improvement and innovation of technology require rigorous system planning and international cooperation \cite{pfenninger_energy_2014,schlachtberger_benefits_2017}.
+The core of the challenge manifests in the system costs.
+Firstly, these should be as small as possible while meeting ecological and techno-economic standards.
+Secondly, they must be distributed in a fair and transparent manner among all agents in the power system.
+In this respect, it is central to identify the drivers of costs and address them appropriately.
+To this end, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019}.
+However, the question of how and, more importantly, upon what grounds costs are distributed is often left open.
 \\
@@
 On the basis of an optimized network the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
-The presented approach builds upon two basic concepts: Firstly, the revenue of an network asset at the cost-optimum exactly matches its operational and capital expenditures, also known as the zero-profit condition.
-Secondly, flow tracing, following Bialek's Average Participation (AP) \cite{bialek_tracing_1996}, can be used to assigns the operation of single network assets to consumers in an plausible and locally constraint manner.
+The approach builds upon two concepts: Firstly, according to the zero-profit condition at the cost-optimum, the revenue of an network asset exactly matches its operational and capital expenditures.
+Secondly, flow tracing, following Bialek's Average Participation (AP) \cite{bialek_tracing_1996}, is used to assign the operation of single network assets to consumers in an plausible and locally constraint manner.
 Combining these two relations allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
@@
 Based on Generation Shift Factors, \ie the marginal contribution of generators to a flow on line, the Locational Marginal Price (LMP) can be represented as a superposition of LMP at the reference bus, the price for congestion and a price for losses.
-The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however does not reflect the exact LMP of an optimized network.
+The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however this does not reflect the exact LMP of an optimized network.
 A similar approach was taken in \cite{tranberg_flow-based_2018} where electricity prices of a non-optimal power dispatch were allocated using flow tracing.
@@
 
-The first section which presents the full method splits into the presentation of: the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  the AP allocation scheme and resulting allocation from assets to consumers (\cref{sec:dispatch_allocations}); the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
-The second part in \cref{sec:application_case} applies the cost allocation to a optimized German power system with high shares of renewable power sources and evaluates the allocated costs.
+The first section which presents the full method splits into the presentation of: the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  the AP allocation scheme and derived allocations from asset to consumer (\cref{sec:dispatch_allocations}); the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
+The second part in \cref{sec:application_case} applies the cost allocation to a optimized German power system with a high share of renewable power generation and evaluates the allocated costs.
 
@@
 \end{align}
-In case a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, is assumed, a further the cost term per generator $s$     
+In case a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, is assumed, a further the cost term per generator $s$,   
 \begin{align}
@@
 At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-We use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
+Again, we use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
 \begin{align}
@@
 \end{align}
-Again, the shadow prices $\muupperflow$ and $\mulowerflow$ can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
+The shadow prices $\muupperflow$ and $\mulowerflow$ can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$.
 Hence, a non-zero values indicate that \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint} are bound and therefore that the congestion on line $\ell$ at time $t$ is imposing costs to the system.
@@
 
-Let $R \in S$ denote all storages in the system.
+Let $R \subset I$ denote all storages in the system.
 In a simplified storage model, $\capacitystorage$ limits the storage dispatch $\storagedispatch$ and charging $\storagecharge$.
@@
 In \cref{sec:zero_profit_storage_units} we formulate the mathematical details.
-The OPEX term which adds to the objective function is given by
+The OPEX which adds to $\totalcost$ is given by
 \begin{align}
@@
 Each of these follow a specific set of assumptions which result in peer-to-peer allocations $A_{m \rightarrow n}$.
-That is a measure for the power supplied at node $m$ and consumed at node $n$.
+That is a measure for the power produced at node $m$ and consumed at node $n$.
 
@@
 A mathematical formulation of the AP scheme is documented in \cref{sec:net_ap}.
-For a detailed comparison with other scheme we refer to [].
+For a detailed comparison with other scheme we refer to \cite{hofmann_techno-economic_2020}.
 
 The peer-to-peer allocations $\allocatepeer$ fulfill some basic properties.
-On the one hand it allocates the all power productions at time $t$, \ie when summing over all receiving nodes $n$ the allocations yield the gross power generation of producing assets $i \in S \cup R$ (generators and storages) attached to $m$, denoted by $\nodalgeneration[m]$.
+On the one hand it allocates the all power productions at time $t$, \ie when summing over all receiving nodes $n$ the allocations yield the gross power generation $\nodalgeneration[m]$ of producing assets $i \in S \cup R$ (generators and storages) attached to $m$.
 Mathematically this translates to
@@
 \end{align}
-As the standard formulation implies, we assume that only net power production of $m$ is allocated to other buses.
-That is, if the nodal generation $\nodalgeneration[m]$ does not exceed the nodal demand $\demand[m]$, all of it is allocated to local consumers, \ie $\allocatepeer[m \rightarrow m] = \min\left(\nodalgeneration[m], \demand[m]\right)$.
+As the standard formulation implies, we assume that only net power production of $m$ is allocated to other buses, \ie $\allocatepeer[m \rightarrow m] = \min\left(\nodalgeneration[m], \demand[m]\right)$.
+Like this nodal generations which do not exceed the nodal demand are completely allocated to local consumers.
 The principal of proportional sharing sets the allocation of a single producing asset $i \in S \cup R$ proportional to the nodal allocation $\allocatepeer$ and weights it by the share $w_{i,t} = \incidenceasset[m] \state / \nodalgeneration[m]$ that asset $i$ contributes to the nodal generation $\nodalgeneration[m]$, leading to
@@
 \end{align}
-
-This relation builds one crucial key to the cost allocation. It states how much of the power produced by asset $i \in S \cup R$ is finally consumed by consumers at $n$.
+% 
+This relation states how much of the power produced by asset $i \in S \cup R$ is finally consumed by consumers at $n$.
 
@@
 Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
-For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated or extended using the formulation presented in \cite{hofmann_flow_2020}.
+For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated using the formulation presented in \cite{hofmann_flow_2020}.
 Inserting \cref{eq:nodalgeneration} into \cref{eq:power_flow_equation} and expanding the sum yields 
@@
 
-Using the presented relations, we are able to straightforwardly define the full cost allocation.
-Therefore, the expanded sum in \cref{eq:decompose_state} is inserted in \cref{eq:cost_decomposition} which leads to
+Using the presented relations, we are able to straightforwardly define the cost allocation.
+Therefore, we insert \cref{eq:decompose_state} in \cref{eq:cost_decomposition} and decompose the sum, which leads to
 \begin{align}
@@
 The cost allocation entails a further important property.
-In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
+In a cost-optimal setup with minimized $\totalcost$, the LMP describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
 Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
@@
 which we refer to as the nodal payment or payment.  
-It exactly matches the payment determined by the LMP $\lmp$.
-Hence, \Cref{eq:payment} which is in detailed proved in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the nodal pricing scheme implied by the LMP.
+This relation which is in detailed proven in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the optimized nodal pricing scheme, \ie the nodal payment exactly match the payment determined by the LMP $\lmp$.
 
@@
 \end{align}
-This rent may account for different possible realms, as for example the increased market price due in higher competed areas or additional costs for social or environmental compensation.
-To end this, the share in $\allocatecost$ which consumers pay for the scarcity rent can be recalculated by a correct weighting the shadow price $\muuppergenerationnom$ with the capital price $c_i$, leading to
+This rent may account for different possible realms, as for example the increased market price in higher competed areas or additional costs for social or environmental compensation.
+To end this, the share in $\allocatecost$ which consumers pay for the scarcity rent can be recalculated by a correct weighting of the shadow price $\muuppergenerationnom$ with the capital price $c_i$, leading to
 \begin{align}
@@
 Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding. 
-In that case, asset $i$ does not collect enough revenue from $\allocatecost$ in order to match the CAPEX. The difference, given by 
+In that case, asset $i$ does not collect enough revenue from $\allocatecost$ in order to compensate the CAPEX. The difference, given by 
 \begin{align}
@@
 % 
-\Cref{fig:example_allocation} shows the allocation $\allocatestate$ for $n$~=~1 and $n$~=~2 separately.
+\Cref{fig:example_allocation} shows the allocation $A_{i,n}$ for bus~1 and bus~2 separately.
 The ``sum'' of the two figures give to the actual dispatch and flow.
@@
 The rest makes up 60\% of the total CAPEX spent at generator~1, exactly the share of power allocated to $d_1$.
-Consumers at bus~1 don't pay any transmission CAPEX as no flow is associated with there demand.
+Consumers at bus~1 don't pay any transmission CAPEX as no flow is assigned.
 \\
@@
 We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from generator~1.
-Thus, consumers pay for the local OPEX and CAPEX as well as the proportional share of expenditures at bus~1 and the transmission system.
+Thus, consumers pay for the local OPEX and CAPEX as well as the corresponding proportion of expenditures at bus~1 and the transmission system.
 As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it.
 The allocated $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the full investment of generator~2.
-Again, 2\kk of the 22\kk which are allocated to investment in generator~1 are associated with the scarcity cost for generator~1.
+In contrast, 2\kk of the 22\kk which are allocated to investment in generator~1 are associated with the scarcity cost for generator~1.
 The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line.
@@
 These values match their overall spending plus the cost of scarcity.
-The sum a row returns the nodal payments $C_n = \lambda_n \, d_n$.
+The sum a row returns the nodal payments $\mathcal{C}_n = \lambda_n \, d_n$.
 For example the sum of payments of consumers at bus~1 is 36\kk.
-This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$.
+This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW.
 \\
@@
 The right side shows the capacity expansion of renewable resources as well as installation of backup gas power plants.
-The effective CO$_2$ price is set to 120 \euro per tonne CO$_2$ emission.}
+The effective CO$_2$ price is set to 120 \euro\, per tonne CO$_2$ emission.}
     \label{fig:network}
@@
 % 
-We showcase the behavior of the cost allocation in a more complex system, by applying it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
+We now showcase the behavior of the cost allocation in a more complex system and apply it to an cost-optimized German power system model with 50 nodes and one year time span with hourly resolution.
 The model builds up on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} with technical details and assumptions reported in \cite{horsch_pypsa-eur_2018}.
@@
     \includegraphics[width=\linewidth]{de50bf/average_price}
-    \caption{Average electricity price $\averagelmp$ per region as a result from the optimization of the German power system.
+    \caption{Load-weighted average electricity price $\averagelmp$ per region in the optimized German power system.
 Regions in the middle and south of Germany have high prices whereas electricity in the North with a strong wind, transmission and OCGT infrastructure is cheaper.}
@@
 \end{figure}
-We impose a effective carbon price of 120 \euro\, per tonne-CO$_{2}$ which, for OCGT, adds an effective price of 55 \euro/\megawatthour (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%).
+We impose an carbon price of 120 \euro\, per tonne-CO$_{2}$ which, for OCGT, adds an effective price of 55 \euro/\megawatthour (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%).
 All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.
@@
 The optimized network is shown in  \cref{fig:network}.
-On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission are displayed.
+On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission.
 The optimization expands solar capacities in the south, onshore and offshore wind in the upper north and most west.
@@
 As to expect, 99.97\% are dedicated to OCGT alone, as these have by far the highest operational price.
-For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price,fig:emission_cost}.
+For a detailed regional distribution of payments per MWh and the resulting revenues for generator, see \cref{fig:opex_price}.
 Since during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators, thus the OPEX allocation of OCGT clearly differs from the CAPEX allocation.
@@
 The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewable resources and the transmission expansion limit.
-These sum up approximately to 7.5 b\euro.
+These sum up approximately to 7.5 bn\euro.
 Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments.
@@
 In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average cost per consumed MWh.
-The scarcity for wind and solar is relatively low and impact the average price at roughly 2\euro/MWh.
+The scarcity for wind and solar is relatively low and impact the average price at roughly 2~\euro/MWh.
 It primarily affects, different to the pure CAPEX allocation, regions in close vicinity.
-Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16\euro.
+Note however that the scarcity rent only apportions to a few areas with the highest renewable power potential.
+Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16~\euro.
 This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
@@
 The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
-These violate the optimal design and are not compensated through the nodal payments $\demand \, \lmp$.
+These violate the optimal design and are not recovered by the revenue.
 Most of these ``non-allocatable'' costs account for Pumped Hydro Storage, onshore wind, offshore wind and the transmission system, see \cref{fig:subsidy} for further details.
@@
 
-In \cref{fig:direct-allocation} we compare the total cost allocation of the region with the lowest average LMP (left side) against the one of the region with the highest LMP (right side).
+In \cref{fig:direct-allocation} we compare the P2P cost assignments of the region with the lowest average LMP (left side) against the one with the highest LMP (right side).
 The region with the cheapest electricity is in the Northwest.
@@
 Hence, the wind power supply at this region is not restricted by exhausted wind power resources but by bottlenecks in the transmission system.
+
+Finally, \cref{fig:locality} draws the cumulative share of P2P cost assignments of a function of the distance between producer and consumer.
+The data is shown for all technologies separately.
+The later the curve reaches 100\% the deeper the price of a technologies penetrates into the network.
+Offshore wind has the strongest price influence to remote buses.
+Only 10-30\% of its expenses are compensated by local consumers and the rest is assigned to remote buses especially those with high demand (compare \cref{fig:average_demand}).
+Interestingly, the contrast between Hydrogen Storage and Battery sticks out.
+Whereas Battery costs are hardly assigned to other buses, almost 50\% of the Hydrogen Storage costs are payed by remote buses.
+It underlines the fundamental functionality of the two technologies.
+Hydrogen storage are located at buses with high wind generation and balance out their long-term excess and deficit energy.
+The dispatched power follows a similar way through the network as the wind power.
+The battery on the other hand pairs with local solar production and its locally constraint dispatch and flow.
 
@@
 
-In the presented work, we restricted the application to long term investment models with perfect foresight.
+We restricted the application to long term investment models with perfect foresight.
 However, the cost allocation can as well be applied to short term planning models with fixed capacities.
@@
 
-All figures and number of the plot can be reproduced by using the \textit{snakemake} workflow in [].
+All figures and data points can be reproduced by using the \textit{snakemake} workflow in \cite{hofmann_pypsa-costallocation_2020}.
 
@@
 
-Special thanks to Tom Brown for fruitful discussions and well-placed inputs.
+Thanks to Tom Brown for fruitful discussions and well-placed inputs. Thanks to Alexander Zerrahn and Alexander Kies for helpful comments.
 
@@
 \end{align}
-which we  can use to define the revenue which compensates the CAPEX at $r$, 
+which we  can use to define the revenue which recovers the CAPEX at $r$, 
 \begin{align}
@@
     \end{subfigure}
-    \caption{Average \textbf{CAPEX allocation} per MWh, $\sum_t \allocatecapexgeneration / \sum_t \demand$ and $\sum_t \allocatecapexstorage / \sum_t \demand$, for all production and storage assets (a), onshore wind (b), solar (c) and OCGT (d).
-Average Allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
+    \caption{Average \textbf{CAPEX allocation} per MWh, $\sum_t \allocatecapex / \sum_t \demand$  for all production and storage assets (a), onshore wind (b), solar (c) and OCGT (d).
+Average allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
     \label{fig:capex_price}
@@
     \caption{Average \textbf{OPEX allocation} per consumed MWh, $\sum_t \allocateopex/\sum_t \demand$.
-The effective prices for OPEX are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators and storages.
+The effective prices for OPEX are indicated by the color of the region, the circles are drawn in proportion to the revenue per regional generators and storages.
 As OCGT is the only allowed fossil based technology, the drawn allocation is proportional to OPEX allocation of OCGT generators.}
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{de50bf/maps_price_ptpf_net/co2_cost_total}
-    \caption{Average \textbf{allocated emission cost}, $\sum_t \allocateemissioncost$, per consumed MWh.
-The effective prices are indicated by the color of the region, the size of the circles are set proportional to the revenue per regional generators.}
-    \label{fig:emission_cost}
+    \includegraphics[width=\linewidth]{de50bf/average_demand}
+    \caption{Average demand, $\sum_t \demand/T$ per regions. The regions with high population densities and larger areas reveal a higher demand.}
+    \label{fig:average_demand}
 \end{figure}
@@
     \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints).
-The figure shows the built infrastructure that does not gain back its CAPEX from its market revenue, but is only build due to lower capacity limits.}
+The figure shows the built infrastructure that does not gain back its CAPEX from its market revenue, but is only built due to lower capacity limits.}
     \label{fig:subsidy}

### 2020-10-22 paper/flow_based.tex
incorporate alex Z. review

@@
     \begin{abstract}
-        Cost minimization of power systems leads to perfectly scheduled operations and investments while securing all power demands. 
-        System needs reports from regional, national and international agencies increasingly rely on cost minimization techniques.
+        % Cost minimization of power systems leads to perfectly scheduled operations and investments while securing all power demands. 
+        % System needs reports from regional, national and international agencies increasingly rely on cost minimization techniques.
 However, cost impact to single regions and thefore the question of who profits to what extent from an investment is often neglected.
 The literature provides multiple cost allocation methods dedicated to this problem, yet they hardly suit for large networks and an integration of all costs.
-        This work introduces a peer-to-peer allocation of all system costs in optimized models.
+        % This work introduces a peer-to-peer allocation of all system costs in optimized models.
 Combining basic relations from economic optimization with flow allocation, its cost assignments are locally constraint while still being aligned with the Locational Marginal Prices at the optimum.
 It is showcased by means of an optimized near-future scenario of the German power system.
+        Power system models are a valuable and widely used tool to determine cost-minimal future operation and investment under political or ecological boundary conditions.
+Yet they are silent about the allocation of costs of single assets, \ie generators, transmission lines etc., to consumers in the network.
+Existing cost-allocation methods hardly suit large networks and do not take all relevant costs into account.
+This paper bridges this gap.
+Based on flow tracing, it introduces a peer-to-peer allocation of all costs in an optimized power system to consumers.
+The resulting cost allocation is both locally constrained and aligned with locational marginal prices in the optimum.
+We apply and discuss our approach using a future German scenario.
         \vtop{%
@@
 The shift from controllable to variable, weather-driven power generation as well as the constant improvement and innovation of technology require rigorous system planning and international cooperation \cite{pfenninger_energy_2014,schlachtberger_benefits_2017}.
-The core of the challenge manifests in the system costs.
-Firstly, these should be as small as possible while meeting ecological and techno-economic standards.
+The core of the challenge manifests in the total costs of the system.
+Firstly, these should be as low as possible while meeting ecological and techno-economic standards.
 Secondly, they must be distributed in a fair and transparent manner among all agents in the power system.
-In this respect, it is central to identify the drivers of costs and address them appropriately.
-To this end, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019}.
-However, the question of how and, more importantly, upon what grounds costs are distributed is often left open.
-\\
-This study bridges this gap.
-On the basis of an optimized network the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
-The approach builds upon two concepts: Firstly, according to the zero-profit condition at the cost-optimum, the revenue of an network asset exactly matches its operational and capital expenditures.
-Secondly, flow tracing, following Bialek's Average Participation (AP) \cite{bialek_tracing_1996}, is used to assign the operation of single network assets to consumers in an plausible and locally constraint manner.
-Combining these two relations allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network. \\
-Flow-based cost allocation was already discussed and performed in various works \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,tranberg_flow-based_2018,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
+It is central to identify the drivers of costs and address them appropriately.
+In this respect, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019}.
+Many studies for countries and regions throughout the world exist that lay out how renewable energy penetration can be expanded at minimum costs.
+Yet they largely remain silent how and on which grounds these costs are allocated among consumers.
+
+This paper fills this gap.
+In an optimized network, the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
+We build on two fundamental concepts: first, the zero profit condition that states that, in the optimum, the revenue of each network asset, i.e., generator, transmission line etc., matches its operational and capital expenditures.
+Second, flow tracing, following Bialek’s Average Participation (AP) approach \cite{bialek_tracing_1996}, that allocates the use of network assets to consumers in a locally constrained fashion.
+Combining these two concepts allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network.
+
+The literature discussed and applied the concept of flow-based cost allocation in a range of papers \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,tranberg_flow-based_2018,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
 Shahidehpour et al.
-give a deep insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} with different allocation techniques.
-Based on Generation Shift Factors, \ie the marginal contribution of generators to a flow on line, the Locational Marginal Price (LMP) can be represented as a superposition of LMP at the reference bus, the price for congestion and a price for losses.
-The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however this does not reflect the exact LMP of an optimized network.
-A similar approach was taken in \cite{tranberg_flow-based_2018} where electricity prices of a non-optimal power dispatch were allocated using flow tracing.
-
-The following work brings advantages of all the mentioned studies together.
+provide a profound insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} using different allocation techniques.
+Specifically, they set out that Generation Shift Factors, i.e.
+the marginal contribution of generators to a flow on line, allow to represent locational marginal prices (LMP) as a superposition of the LMP at the reference bus, the price for congestion, and a price for losses.
+The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however it does not reflect the exact LMP of an optimized network.
+A similar approach was used in \cite{tranberg_flow-based_2018} that allocates electricity prices of a non-optimal power dispatch using flow tracing.
+
+In this paper, we bring together the advantages of the studies discussed above.
 It reassures localized cost allocations while fully aligning payments to the nodal pricing scheme based on the LMP.
-The work serves to facilitate transparency and cost-benefit analysis in network plannings such as the Ten Year Network Development Plan \cite{entso-e_completing_2020} or the German Netzentwicklungsplan (NEP) \cite{bundesnetzagentur_netzentwicklungsplan_2020}.
-Further, it builds a basis of a usage based transmission cost allocation.
-
-The first section which presents the full method splits into the presentation of: the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule});  the AP allocation scheme and derived allocations from asset to consumer (\cref{sec:dispatch_allocations}); the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}); a numerical example (\cref{sec:numerical_example}).
-The second part in \cref{sec:application_case} applies the cost allocation to a optimized German power system with a high share of renewable power generation and evaluates the allocated costs.
+It serves to facilitate transparency and cost-benefit analysis in network plannings such as the Ten Year Network Development Plan \cite{entso-e_completing_2020} or the German Netzentwicklungsplan (NEP) \cite{bundesnetzagentur_netzentwicklungsplan_2020}.
+Further, it provides a point of departure for usage-based transmission cost allocation.
+
+The first section presents the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule}),  the AP allocation scheme and derived allocations from asset to consumer (\cref{sec:dispatch_allocations}), the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}), a numerical example (\cref{sec:numerical_example}).
+\Cref{sec:application_case} applies the cost allocation to a optimized German power system with a high share of renewable power generation and evaluates the allocated costs.
 
@@
 Further the thickness of the lines indicates the dedicated amount of payments.
-The cheap prices in the North...
 }
@@
 
-All figures and data points can be reproduced by using the \textit{snakemake} workflow in \cite{hofmann_pypsa-costallocation_2020}.
+All figures and data points can be reproduced by using the \textit{python} workflow in \cite{hofmann_pypsa-costallocation_2020}.
 
@@
 
-Thanks to Tom Brown for fruitful discussions and well-placed inputs. Thanks to Alexander Zerrahn and Alexander Kies for helpful comments.
+In particular, I thank Tom Brown for very fruitful discussions.
+I am very grateful to Alexander Zerrahn for reviewing and helping out with important parts.
+Further, I want to thank Alexander Kies and Markus Schlott who steadily helped with creative thoughts.
 

### 2020-10-23 paper/flow_based.tex
submitted version

@@
 \usepackage[colorlinks]{hyperref}
-\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8, style=numeric-comp]{biblatex}
+\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8, style=ieee]{biblatex}
 \usepackage{eurosym}
@@
 The resulting cost allocation is both locally constrained and aligned with locational marginal prices in the optimum.
-We apply and discuss our approach using a future German scenario.
+The approach is applied and discussed using a future German scenario.
         \vtop{%
@@
         $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
-        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
+        $\demand$ & Electric demand per bus $n$ \\ & time step $t$ in MW  \vpad \\
         $\state$ & Operational state of asset $i$, \\ & at time step $t$  in MW \vpad \\
@@
         % $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
-        $o_i$ & Operational price of asset $i$ in \euro/MW \vpad \\
+        $o_i$ & Operational price of asset $i$ in \euro/MWh \vpad \\
         $c_i$ & Capital Price of asset $i$ in \euro/MW \vpad \\
@@
 This paper fills this gap.
-In an optimized network, the operational state of each time step and corresponding characteristics are in detailed considered to allocate all system costs.
+In an optimized network, the characteristics each time step are in detailed considered to allocate all system costs.
 We build on two fundamental concepts: first, the zero profit condition that states that, in the optimum, the revenue of each network asset, i.e., generator, transmission line etc., matches its operational and capital expenditures.
-Second, flow tracing, following Bialek’s Average Participation (AP) approach \cite{bialek_tracing_1996}, that allocates the use of network assets to consumers in a locally constrained fashion.
+Second, the flow tracing method, following Bialek’s Average Participation (AP) approach \cite{bialek_tracing_1996}, that allocates the use of network assets to consumers in a locally constrained fashion.
 Combining these two concepts allows for an transparent allocation of all operational (OPEX) and capital expenditures (CAPEX) to the consumers in the network.
@@
 The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however it does not reflect the exact LMP of an optimized network.
-A similar approach was used in \cite{tranberg_flow-based_2018} that allocates electricity prices of a non-optimal power dispatch using flow tracing.
+A similar approach was used in \cite{schafer_allocation_2017} that allocates electricity prices of a non-optimal power dispatch using flow tracing.
 
 In this paper, we bring together the advantages of the studies discussed above.
-It reassures localized cost allocations while fully aligning payments to the nodal pricing scheme based on the LMP.
+Our approach assures localized cost allocations while fully aligning payments to the nodal pricing scheme based on the LMP.
 It serves to facilitate transparency and cost-benefit analysis in network plannings such as the Ten Year Network Development Plan \cite{entso-e_completing_2020} or the German Netzentwicklungsplan (NEP) \cite{bundesnetzagentur_netzentwicklungsplan_2020}.
@@
     \end{subfigure}
-    \caption{Power allocations $\allocatestate$ in the example network \cref{fig:example_network} using Average Participation.
-Consumers at for bus~1, figure (a), are, with 60~MW, totally supplied by the local generator and do not consume any imported power from bus 2.
-In contrast consumer at bus~2, figure (b), retrieve 40~MW from generator 1, induce a flow of 40~MW on the transmission line and consume 50~MW from the local generator.}
+    \caption{Power allocations $A_{i,n}$ for the example network in \cref{fig:example_network} using Average Participation.
+The 60~MW consumption at bus~1 (a) are totally supplied by the local generator.
+In contrast consumers at bus~2 (b) retrieve 50~MW from the local generator, the remaining 40~MW are imported from generator~1 which induces a flow on the transmission line.}
     \label{fig:example_allocation}
@@
 Thus, consumers pay for the local OPEX and CAPEX as well as the corresponding proportion of expenditures at bus~1 and the transmission system.
-As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity cost are assigned to it.
+As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity rent is assigned to it.
 The allocated $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the full investment of generator~2.
@@
     \includegraphics[width=\linewidth]{example_payoff_net_ptpf.png}
-    \caption{Full P2P cost allocation $\allocatecost$ on the optimized example network in \cref{fig:example_network}.
+    \caption{Full P2P cost allocation $\allocatecost$ for the optimized example network in \cref{fig:example_network}.
 Consumers compensate OPEX and CAPEX of the generators they retrieve from (compare with \cref{fig:example_allocation}).
 As bus~1 is totally self-supplying, all it payment is assigned to the local generator.
-As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX at bus~1 and CAPEX for the transmission.}
+As bus~2 imports power from bus~1 and thus induces a flow on line~1, it not only compensates local expenditures but also OPEX and CAPEX of generator~1 and CAPEX of the transmission line.}
     \label{fig:example_payoff}
@@
 % 
-The sum of all values in the payoff matrix in \cref{fig:example_payoff} yield $\totalcost - \scarcitycost$, the total system cost minus the scarcity cost (which in turn is negative).
+The sum of all values in the payoff matrix in \cref{fig:example_payoff} yield $\totalcost - \scarcitycost$, the total system cost minus the scarcity rent (which in turn is negative).
 The sum of a column  yields the total revenue per the asset $i$.
@@
     \end{subfigure}
-    \caption{Comparison of payments of the node with the \textbf{lowest LMP (left)} and the node with the \textbf{highest LMP (right)}.
-The region of the paying bus is colored in dark blue.
-The circles indicate where to which bus and technology combined OPEX and CAPEX payments.
-Further the thickness of the lines indicates the dedicated amount of payments.
-}
+    \caption{Comparison of payments of the region with the \textbf{lowest LMP (left)} and the region with the \textbf{highest LMP (right)}.
+The region is colored in dark blue.
+The circles indicate to which bus and technology OPEX and CAPEX are assigned.
+The thickness of the lines is proportional to dedicated payment.}
     \label{fig:direct-allocation}
@@
 In the real world this money would be spent for augmented land costs or civic participation in the dedicated areas.
-In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity cost manifest in the average cost per consumed MWh.
+In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity rent manifests in the average cost per consumed MWh.
 The scarcity for wind and solar is relatively low and impact the average price at roughly 2~\euro/MWh.
@@
 Note however that the scarcity rent only apportions to a few areas with the highest renewable power potential.
-Remarkably, the scarcity cost per MWh for run-of-river power plants amounts up to 16~\euro.
+Remarkably, the scarcity rent per MWh for run-of-river power plants amounts up to 16~\euro.
 This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
-However, as these power plants in particular are already amortized, the scarcity cost should be reconsidered and removed from a final cost allocation.
+However, as these power plants in particular are already amortized, the scarcity rent should be reconsidered and removed from a final cost allocation.
 
@@
 Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
-However, the AP scheme might not the right choice as the Kirchhoff Voltage Law cannot not be considered subsequently as done here.
-Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it considers by default the two circuit laws.
+However, the AP scheme might not the correct choice as the misalignment from the Kirchhoff Voltage Law cannot not be corrected subsequently as done here.
+Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it naturally respects both circuit laws.
 Allocating on the basis of the full power flow introduces an additional cost term $\remainingcost^\text{Loss}$ accounting for the transmission loss which is compensated by the consumers and mirrors in the LMP.
 
-The used optimization does not consider security constrains of the transmission system, this can however be incorporated by introducing by following security constrained optimal power flow (SCOPF) as done in \cite{nikoukar_transmission_2012}.
+The used optimization does not take security constrains of the transmission system into account. These may be incorporated following the in \cite{nikoukar_transmission_2012}.
 
@@
 
-The optimization assumed a fix demand time series.
+The optimization assumes a fix demand time series.
 As shown in \cref{fig:capex_duration_curve} this leads to high if not unrealistic LMP.
@@
 
-\section{Conclusion and Outlook}
+\section{Conclusion}
 
@@
 These alter the revenues per asset and therefore the cost allocations.
-For lower capacity requirements, the cost have to be subsidized by the overall system operator or the government, as those cannot be allocated.
-Contrary, upper capacity expansion limits lead to an additional charge for consumers which have to compensate for \eg land use restrictions.
-Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent, the cost allocation shows how buses remote from wind farms pay higher prices due to increased reliance in transmission and backup capacity.
+For lower capacity requirements, assets may not recover all the expenses from the revenue.
+In this case the cost difference has to be subsidized or simply be ignored, in case the asset is already amortized.
+Contrary, upper capacity expansion limits lead to an additional charge for consumers which have to compensate for an additional scarcity rent of the assets.
+Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent.
+The cost allocation shows low electricity prices for consumers in a renewable German system are made achieved on a transparent basis.
+The cost allocation shows how buses remote from wind farms pay higher prices due to increased reliance in transmission and backup capacity.
 On the other hand, buses with high renewable installation spend most payments to local assets.
@@
 \subsection*{Funding}
-This research was funded by the by the German Federal Ministry for Economics Affairs and Energy in
-the frame of the NetAllok project \cite{bundesministerium_fur_wirtschaft_und_energie_verbundvorhaben_nodate}.
+This research was funded by the by the German Federal Ministry for Economics Affairs and Energy (BMWi) in
+the frame of the NetAllok project (grant number 03ET4046A) \cite{bundesministerium_fur_wirtschaft_und_energie_verbundvorhaben_nodate}.
 

### 2020-10-23 paper/flow_based.tex
SUBMITTED Version (not the previous)

@@
 
-\newcommand{\kk}{k\euro~}
+\newcommand{\kk}{k~\euro~}
 
@@
 This paper fills this gap.
-In an optimized network, the characteristics each time step are in detailed considered to allocate all system costs.
+In an optimized network, the characteristics of each time step are in detailed considered to allocate all system costs.
 We build on two fundamental concepts: first, the zero profit condition that states that, in the optimum, the revenue of each network asset, i.e., generator, transmission line etc., matches its operational and capital expenditures.
@@
 
-The literature discussed and applied the concept of flow-based cost allocation in a range of papers \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,tranberg_flow-based_2018,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
+The literature discussed and applied the concept of flow-based cost allocation in a range of papers \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,schafer_allocation_2017,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}.
 Shahidehpour et al.
-provide a profound insight into allocating congestion cost and transmission investments to market participants \cite{shahidehpour_market_2002} using different allocation techniques.
+provide a profound insight into allocating congestion cost and transmission investments to market participants using different allocation techniques \cite{shahidehpour_market_2002}.
 Specifically, they set out that Generation Shift Factors, i.e.
@@
 The approach in \cite{meng_investigation_2007} expands this relation for contributions based on the AP scheme, which allows for an accurate estimation of the LMP, however it does not reflect the exact LMP of an optimized network.
-A similar approach was used in \cite{schafer_allocation_2017} that allocates electricity prices of a non-optimal power dispatch using flow tracing.
+A similar approach is used in \cite{schafer_allocation_2017} that allocates electricity prices of a non-optimal power dispatch using flow tracing.
 
@@
     \includegraphics[width=.8\linewidth]{cost_decomposition.png}
-    \caption{Schematic decomposition of the total system cost $\totalcost$ in an long-term investment model.}
+    \caption{Schematic decomposition of the total system cost $\totalcost$ in a long-term investment model.}
     \label{fig:cost_decomposition}
@@
 
-In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms $\cost$.
+In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms $\cost$, as schematically depicted in \cref{fig:cost_decomposition}.
 Typically, these include operational expenditures for generators $\opexgeneration$, expenditures for emissions $\emissioncost$,  capital investments for the transmission system $\capexflow$ and so on, \ie
@@
 \end{align} 
-where an ``asset'' describes any operating components of the network, such as a generator, line, energy storage etc. We refer to the set of assets as $I$. 
+where an ``asset'' describes any operating components of the network, such as a generator, line, energy storage etc. We refer to the set of all assets as $I$. 
 In a long-term equilibrium of a power system with perfect competition, the zero-profit condition states that each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
@@
 Therefore, let $o_i$ denote the operational price per MWh of asset $i$ and $c_i$ the capital price for one MW capacity expansion.
-\Cref{tab:cost_allocation_map} summarizes all derived relations, these can be inserted into \cref{eq:asset_cost,eq:cost_decomposition} for each single cost term.
+\Cref{tab:cost_allocation_map} summarizes all derived relations.
+These can be inserted into \cref{eq:asset_cost,eq:cost_decomposition} for each single cost term.
 
@@
     \end{center}
-    \caption{Mapping of different cost terms to the cost allocation scheme given in \cref{eq:general_scheme}.
-These include OPEX \& CAPEX for production, transmission and storage assets in the network, as well as a cost term for the total Green House Gas (GHG) emissions.}
+    \caption{Representation of different cost terms as a function of the operational state, \ie matching the form in \cref{eq:cost_decomposition}.
+These include OPEX \& CAPEX for production, transmission and storage assets in the network, as well as a cost term for the total Green House Gas (GHG) emissions.
+For storage units, an additional cost term adds to $\capexstorage$ (see \cref{sec:zero_profit_storage_units} for details).}
     \label{tab:cost_allocation_map}
@@
 Here, $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$.
-In contrast to OPEX and emission costs, the CAPEX of $S$ are not a function of the production $\generation$, but of the actual installed capacity $\capacitygeneration$ of generator $s$.
+In contrast to OPEX and emission costs, the CAPEX of $s$ are not a function of the production $\generation$, but of the actual installed capacity $\capacitygeneration$ of generator $s$.
 In the optimization it limits the generation $\generation$ in the form of
@@
 \end{align}
-The constraint yields a shadow-price of $\muuppergeneration$ in literature often denoted as the Quality of Supply \cite{schweppe_spot_1988}.
+The constraint yields a shadow-price of $\muuppergeneration$, given by corresponding the Karush–Kuhn–Tucker (KKT) variable, in literature often denoted as the Quality of Supply \cite{schweppe_spot_1988}.
 It can be interpreted as the price per MW that \cref{eq:upper_generation_capacity_constraint} imposes to the system.
 If $\muuppergeneration$  is bigger than zero, the constraint is binding, which pushes investments in $\capacitygeneration$.
-As shown in \cite{brown_decreasing_2020} and in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$,
+As shown in \cite{brown_decreasing_2020} and in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is recovered by the production $\generation$ times the shadow price $\muuppergeneration$,
 \begin{align}
@@
 At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-Again, we use the relation that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow (for details see \cref{sec:zero_profit_flow})
+Again, we use the relation that over the whole time span, the shadow prices weighted by the flow match the investment in line $\ell$ (for details see \cref{sec:zero_profit_flow})
 \begin{align}
@@
 
-The fact that all asset related costs $\cost_i$ can be represented as a cost-weighted sum of the operational state $\state$, prompts the question how the operational state $\state$, in turn, is allocated to consumers.
-This builds the second crucial ingredient of the cost allocation.
-
-Dispatch and flow in a system can be considered as a superposition of individual contributions of nodes and assets.
+The fact that all asset related costs $\cost_i$ can be represented as a cost-weighted sum of the operation $\state$, prompts the question how $\state$ in turn is allocated to consumers. 
+% The approach we take here builds the second crucial ingredient of the cost allocation.
+
+Dispatch and flow in a system can be considered as a superposition of individual contributions of nodes or assets.
 In order to artificially quantify these contribution, the literature provides various methods, named flow allocation schemes.
-Each of these follow a specific set of assumptions which result in peer-to-peer allocations $A_{m \rightarrow n}$.
+Each of these follow a specific set of assumptions which lead to peer-to-peer allocations $A_{m \rightarrow n}$.
 That is a measure for the power produced at node $m$ and consumed at node $n$.
 
-In the following, we resort to one specific flow allocation scheme Average Participation (AP), also known as Flow Tracing.
+In the following, we resort to one specific flow allocation scheme Average Participation (AP), also known as flow tracing \cite{bialek_tracing_1996}.
 The basic idea is to trace the power injection at bus $m$ through the network while following the real power flow on transmission lines and applying the principal of proportional sharing.
@@
 A mathematical formulation of the AP scheme is documented in \cref{sec:net_ap}.
-For a detailed comparison with other scheme we refer to \cite{hofmann_techno-economic_2020}.
+For a detailed comparison with other schemes we refer to \cite{hofmann_techno-economic_2020}.
 
@@
 Like this nodal generations which do not exceed the nodal demand are completely allocated to local consumers.
-The principal of proportional sharing sets the allocation of a single producing asset $i \in S \cup R$ proportional to the nodal allocation $\allocatepeer$ and weights it by the share $w_{i,t} = \incidenceasset[m] \state / \nodalgeneration[m]$ that asset $i$ contributes to the nodal generation $\nodalgeneration[m]$, leading to
-\begin{align}
-    \allocatestate = w_{i,t} \allocatepeer \Forall{i \in S \cup R, n , t}
+The contribution of a single producing asset $i \in S \cup R$ to the nodal allocation $\allocatepeer$ is in proportion to the share $w_{i,t} = \incidenceasset[m] \state / \nodalgeneration[m]$ that asset $i$ contributes to the nodal generation $\nodalgeneration[m]$, leading to
+\begin{align}
+    \allocatestate = w_{i,t}\, \allocatepeer \Forall{i \in S \cup R, n , t}
     \label{eq:allocate_production}
@@
 Yet, we didn't touch the allocation of transporting assets $i \in L$ to consumers.
-Note that the traced flow based in the AP scheme does obey the Kirchhoff Current Law but not the Kirchhoff Voltage Law, as already pointed out in \cite{galiana_transmission_2003}.
-However, as we show later a consideration of both laws is needed in order to align the allocated costs with the optimized Locational Marginal Prices.
+Note that the traced flow based in the AP scheme obeys the Kirchhoff Current Law but not the Kirchhoff Voltage Law, as already pointed out in \cite{galiana_transmission_2003}.
+However, as we show later a consideration of both laws is needed in order to align the allocated costs with the optimized locational marginal prices (LMP).
 To tackle this, let $\ptdf$ denote a linear mapping between the injection $\left(\nodalgeneration[m] - \demand[m]\right)$ and the flow $\flow$, such that
@@
 Usually, $\ptdf$ is given by the Power Transfer Distribution Factors (PTDF) which indicate the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $m$.
-For transport models or networks with High Voltage Directed Current (HVDC) lines, these can retrospectively be calculated using the formulation presented in \cite{hofmann_flow_2020}.
+For transport models or networks with High Voltage Directed Current (HVDC) lines, these can be set retrospectively using the formulation presented in \cite{hofmann_flow_2020}.
 Inserting \cref{eq:nodalgeneration} into \cref{eq:power_flow_equation} and expanding the sum yields 
@@
 
-With \cref{eq:allocate_production,eq:allocate_flow} a full set of allocations between the operational state $\state$ of asset $i \in I$ is derived.
-Naturally, the sum over all receiving nodes reproduces the operational state $\state$ of asset $i$,
+With \cref{eq:allocate_production,eq:allocate_flow} the allocation from all assets $i \in I$ to consumers in the network is derived.
+Naturally, the sum over all receiving nodes reproduces the operation $\state$ of asset $i$,
 \begin{align}
@@
 \end{align}
-By default, all cost of a cost term $\cost_i$ associated with asset $i$ are allocated, thus $\cost_i = \sum_n \allocatecost$; likewise all system costs are distributed, $\totalcost = \sum_{\circ, i, n, t} \allocatecost$.
+By default, the full costs $\cost_i$ associated with asset $i$ are allocated, \ie $\cost_i = \sum_n \allocatecost$.
+Likewise all system costs are allocated, $\totalcost = \sum_{\circ, i, n, t} \allocatecost$.
 
 The cost allocation entails a further important property.
-In a cost-optimal setup with minimized $\totalcost$, the LMP describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
+In a cost-optimal setup, the LMP describes the change of costs for an incremental increase of electricity demand $\demand$ at node $n$ and time $t$ \cite{schweppe_spot_1988}.
 Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
@@
 which we refer to as the nodal payment or payment.  
-This relation which is in detailed proven in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the optimized nodal pricing scheme, \ie the nodal payment exactly match the payment determined by the LMP $\lmp$.
+This relation which is in detailed proven in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the optimized nodal pricing scheme, \ie the nodal payment exactly matches the payment determined by the LMP $\lmp$.
 
@@
 \end{align}
-
-According to the nature of \cref{eq:design_constraint} and the corresponding $\remainingcost_i$, it is either larger, equal or lower than $\cost$.
-Whereas \cref{eq:payment} still holds true, the total of payments do not return the total system cost $\totalcost$ anymore.
+and result in a mismatch of $\remainingcost^\circ_i$ between the expenses and the revenue, therefore the cost allocation to asset $i$.
+According to the nature of \cref{eq:design_constraint} and $\remainingcost_i$, it is either larger, equal or lower than $\cost$.
+Whereas \cref{eq:payment} still holds true, the total of payments do not return the system cost $\totalcost$ anymore.
 The total mismatch $\remainingcost$ is given by
@@
 \begin{subequations}
-In real-world setups, generators, lines or other assets can often only be built up to a certain limit. This might be due to land use restrictions or social acceptance considerations. 
+In real-world setups the capacity expansion of generators, lines or other assets are often limited. This might be due to land use restrictions or social acceptance considerations. 
 When constraining the capacity $\capacity$ to an upper limit $\capacityupper$, in the form of 
@@
 \begin{subequations}
-In order to take already built infrastructure into account, the capacity $\capacity$ can be constrained to a minimum required capacity $\capacitylower$. This introduces a constraint of the form 
+In order to take already built infrastructure into account, the capacity $\capacity$ may be constrained by a minimum required capacity $\capacitylower_i$. This introduces a constraint of the form 
 \begin{align}
@@
 has to be subsidized by governments or communities or is simply ignored when investments are amortized.
-Note, it is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost = \subsidycost_i $.
+Note, it is rather futile wanting to allocate these cost to consumers as assets may not gain any revenue for their operational state, \ie where $\cost_i = \subsidycost_i $.
 \end{subequations}
@@
 
+After presenting the basics of the cost allocation, we present its behavior by means of a small numerical example.  
 Consider a two bus system, depicted in \cref{fig:example_network}, with one transmission line and one generator per bus.
@@
 With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW.
-The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW.
+The 40~MW excess power, not consumed at bus~1, flow to bus~2 where the generator is built with only 50~MW.
 \\
@@
 
-The left graph \cref{fig:example_allocation_bus1} shows that $d_1$ is with 60~MW totally supplied by the local production.
-Consequently consumers at bus~1 pay 3k~\euro~OPEX, which is the operational price of 50~\euro\, times the 60~MW.
-Further they pay 33k~\euro~ for the CAPEX at generator~1.
-Note that 3k~\euro~ of these account for the scarcity imposed buy the upper expansion limit $\capacitygenerationupper$.
-The rest makes up 60\% of the total CAPEX spent at generator~1, exactly the share of power allocated to $d_1$.
+The left graph \cref{fig:example_allocation_bus1} shows that $d_1$ is totally supplied by the local production.
+Consequently consumers at bus~1 pay 3k~\euro~OPEX to generator~1, which is the operational price of 50~\euro\, times the retrieved power of 60~MW.
+Further they pay 33\kk of its CAPEX.
+Note that 3\kk of these account for the scarcity imposed buy the upper expansion limit $\bar{G}_1$.
+The rest makes up 60\% of the total CAPEX of generator~1, exactly the share of power allocated to $d_1$.
 Consumers at bus~1 don't pay any transmission CAPEX as no flow is assigned.
@@
 We see that 50~MW are self-supplied whereas the remaining 40~MW are imported from generator~1.
-Thus, consumers pay for the local OPEX and CAPEX as well as the corresponding proportion of expenditures at bus~1 and the transmission system.
-As the capacity at generator~2 does not hit the expansion limit $\capacitygenerationupper$, no scarcity rent is assigned to it.
+Thus, consumers pay for the local OPEX and CAPEX as well as the corresponding proportion of to generator~1 and the transmission line.
+As the capacity at generator~2 does not hit the expansion limit $\bar{G}_2$, no scarcity rent is assigned to it.
 The allocated $\allocatecapexgeneration[2\rightarrow 2]$  compensates  the full investment of generator~2.
-In contrast, 2\kk of the 22\kk which are allocated to investment in generator~1 are associated with the scarcity cost for generator~1.
+In contrast, 2\kk of the 22\kk which are allocated to investment in generator~1 are associated with the scarcity rent of generator~1.
 The payed congestion revenue of 4\kk is exactly the CAPEX of the transmission line.
-% 
+
 \begin{figure}[h]
@@
 \end{figure}
-% 
+
 The sum of all values in the payoff matrix in \cref{fig:example_payoff} yield $\totalcost - \scarcitycost$, the total system cost minus the scarcity rent (which in turn is negative).
@@
 This coherence breaks for larger optimization problems with multiple time steps.
-Then CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
+Then CAPEX allocation takes effect only when one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.
 \\
@@
 We impose an carbon price of 120 \euro\, per tonne-CO$_{2}$ which, for OCGT, adds an effective price of 55 \euro/\megawatthour (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%).
-All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in detail in \cref{tab:cost_assumptions}.
+All cost assumptions on operational costs $o_i$ and annualized capital cost $c_i$ are summarized in \cref{tab:cost_assumptions}.
 
 The optimized network is shown in  \cref{fig:network}.
-On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the capacity expansion for generation, storage and transmission.
-The optimization expands solar capacities in the south, onshore and offshore wind in the upper north and most west.
+On the left we find the lower capacity bounds for renewable generators and transmission infrastructure, on the right the optimized capacity expansion for generation, storage and transmission.
+Solar capacities are expanded in the south, onshore and offshore wind in the upper north and most west.
 Open-Cycle Gas Turbines (OCGT) are build within the broad middle of the network.
@@
 In \cref{fig:total_cost} we show the total of all allocated costs.
-The allocated cost to capital expenditures is given by $\cost - \remainingcost^\circ$.
-The difference $\remainingcost$ consists of scarcity rents $\scarcitycost$ and subsidies $\subsidycost$.
+We split the CAPEX allocation into $\cost - \remainingcost^\circ$ and $\remainingcost^\circ$.
+The difference $\remainingcost^\circ$ consists of scarcity rents $\scarcitycost$ and subsidies $\subsidycost$.
 Note that the sum of all contributions in \cref{fig:total_cost} equals the total cost $\totalcost$.
@@
 The largest proportion of the payments is associated with CAPEX for generators, transmission system and storage units in decreasing order. 
-Taking different technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shown in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
+Taking production technologies into account, we observe a fundamental difference between the controllable OCGT  and the variable renewable resources: As shown in detail in \cref{fig:capex_duration_curve}, more than half of all investments in OCGT is determined in one specific hour.
 At this time (morning, end of February), the system hits the highest mismatch between low renewable power potentials and high demands.
-The necessity for backup generators manifests in high allocation of CAPEX for OCGT and consequently high LMP.
+The necessity for backup generators manifests in high allocation of CAPEX to OCGT and consequently high LMP.
 With a few exceptions in the South West and East, all consumers receive power from OCGT at this time, thus all pay high amounts for the needed backup infrastructure (operational state of the system is in detailed shown in \cref{fig:operation_high_expenditure}).
@@
 As for onshore and offshore wind farms, the produced power deeply penetrates the network, see \cref{fig:power_mix}, thus it is not only local, but also remote consumers which cover the CAPEX, see \cref{fig:capex_price}.
-This in turn benefits local consumers, which profit from the cheap operational prices of local wind farms.
+This in turn benefits consumers, which profit from cheap operational prices of local wind farms.
 This explains why these regions end up with a low average LMP.\\
@@
 Since during ordinary demand peaks, it is rather local OCGT generators which serve as backup generators, thus the OPEX allocation of OCGT clearly differs from the CAPEX allocation.
-In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities predominately have high shares of OCGT power.
+In the average power mix per region, see \cref{fig:power_mix}, we observe that regions with strong OCGT capacities also  have high shares of OCGT power consumption.
 The average operational price for renewable generators is extremely low (0.2 \euro/MWh), thus they play an inferior role in the OPEX allocation.
@@
 The circles indicate to which bus and technology OPEX and CAPEX are assigned.
-The thickness of the lines is proportional to dedicated payment.}
+The thickness of the lines is proportional to dedicated payments.}
     \label{fig:direct-allocation}
@@
 The negative segment in \cref{fig:total_cost} is associated with scarcity costs $\scarcitycost$, caused by land use constraints for renewable resources and the transmission expansion limit.
-These sum up approximately to 7.5 bn\euro.
+These sum up to approximately 7.5bn~\euro.
 Note again, that according to \cref{eq:scarcitycost} this term is negative and is part of the allocated CAPEX payments.
@@
 In \cref{fig:scarcity_price} we give a detailed insight of how the scarcity rent manifests in the average cost per consumed MWh.
-The scarcity for wind and solar is relatively low and impact the average price at roughly 2~\euro/MWh.
-It primarily affects, different to the pure CAPEX allocation, regions in close vicinity.
-Note however that the scarcity rent only apportions to a few areas with the highest renewable power potential.
+The scarcity for wind and solar is relatively low and contributes roughly 2~\euro/MWh to the load-weighted average price $\averagelmp$.
+It primarily affects regions in close vicinity to area with the highest renewable power potential.
+\\
 Remarkably, the scarcity rent per MWh for run-of-river power plants amounts up to 16~\euro.
-This high impact is due to the steady power potential from run-off water and the strong limitation of capacity expansion.
+This impact comes from the steady power potential from run-off water and the strong limitation of capacity expansion.
 However, as these power plants in particular are already amortized, the scarcity rent should be reconsidered and removed from a final cost allocation.
-
+\\
 A high influence on the price comes from the scarcity of transmission expansion.
-Right beside regions with high wind infrastructure, it occasions costs higher than MWh 4\euro/MWh (maximal 8\euro/MWh), see \cref{fig:branch_scarcity_price}.
+Right beside regions with high wind infrastructure, it occasions average costs higher than MWh 4~\euro/MWh (maximal 8\euro/MWh), see \cref{fig:branch_scarcity_price}.
 As to expect, the constraint mainly suppresses transmission expansion along the north-south axis.
-
-The last cost term in \cref{fig:total_cost}, is  caused by lower capacity constraints for pre-existing assets.
+\\
+The last cost term $\subsidycost$ in \cref{fig:total_cost}, is caused by lower capacity constraints for pre-existing assets.
 These violate the optimal design and are not recovered by the revenue.
 Most of these ``non-allocatable'' costs account for Pumped Hydro Storage, onshore wind, offshore wind and the transmission system, see \cref{fig:subsidy} for further details.
-Again, as most of the named assets are already amortized (PHS, transmission system), the subsidy should be reconsidered in a final cost allocation.
-
-In \cref{fig:direct-allocation} we compare the P2P cost assignments of the region with the lowest average LMP (left side) against the one with the highest LMP (right side).
-The region with the cheapest electricity is in the Northwest.
-We see that this bus is fairly independent of investments in the transmission system, as it is supplied by steady, cheap offshore power.
-Moreover, the CAPEX of the local offshore wind farms are partly payed by subsidies, see \cref{fig:subsidy}.
-Only a small share of the payments is allocated to remote OCGTs, resulting from the extreme event mentioned earlier.
-
-In contrast the regions with the highest LMP, spends most its money to local OCGT operations, emission cost and the transmission system.
+Again, as most of these generators are amortized (PHS, transmission system), the costs should be reconsidered in a final cost allocation.
+
+\Cref{fig:direct-allocation} compares the P2P cost assignments of the region with the lowest average LMP (left side) against the one with the highest LMP (right side).
+The low-price region in the north-west is fairly independent of investments in the transmission system.
+It profits from local offshore wind farms which are partly payed by subsidies, see \cref{fig:subsidy}.
+Only a small share of the payments is allocated to remote OCGTs.
+In contrast, the high-priced region is highly dependent on local OCGT and the transmission system, which causes high allocated OPEX, emission cost and transmission CAPEX.
 Its payments to  onshore and offshore wind infrastructure are low despite a third of its supply comes from wind power.
-Instead, a lot of its payments are assigned to the transmission infrastructure.
 Hence, the wind power supply at this region is not restricted by exhausted wind power resources but by bottlenecks in the transmission system.
@@
 Offshore wind has the strongest price influence to remote buses.
-Only 10-30\% of its expenses are compensated by local consumers and the rest is assigned to remote buses especially those with high demand (compare \cref{fig:average_demand}).
-Interestingly, the contrast between Hydrogen Storage and Battery sticks out.
+Only 10-30\% of its expenses are compensated by local consumers and the rest is assigned to remote buses, especially those with high demand (compare \cref{fig:average_demand}).
+The contrast between Hydrogen Storage and Battery sticks out.
 Whereas Battery costs are hardly assigned to other buses, almost 50\% of the Hydrogen Storage costs are payed by remote buses.
@@
 Yet, the method is equally applicable to a system with an Optimal Power Flow (OPF), \ie full AC power flows.
-However, the AP scheme might not the correct choice as the misalignment from the Kirchhoff Voltage Law cannot not be corrected subsequently as done here.
-Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might be a better choice as it naturally respects both circuit laws.
+Yet for this case, the AP scheme is not the correct choice as the misalignment from the Kirchhoff Voltage Law cannot not be corrected subsequently.
+Rather, the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} might suit better as it naturally respects both circuit laws.
 Allocating on the basis of the full power flow introduces an additional cost term $\remainingcost^\text{Loss}$ accounting for the transmission loss which is compensated by the consumers and mirrors in the LMP.
 
-The used optimization does not take security constrains of the transmission system into account. These may be incorporated following the in \cite{nikoukar_transmission_2012}.
-
-We restricted the application to long term investment models with perfect foresight.
-However, the cost allocation can as well be applied to short term planning models with fixed capacities.
-The revenue from the capacity limits then builds basis for amortization and future investments.
+The used optimization does not take security constrains of the transmission system into account. These may be incorporated following the approach in \cite{nikoukar_transmission_2012}.
+
+We restricted the application to long-term investment models with perfect foresight.
+However, the cost allocation can as well be applied to short-term planning models with fixed capacities.
+The revenue from the capacity limits then builds the basis for amortization and future investments.
 
@@
 Applied to a optimized German power system with an imposed price of 120 \euro\, per tonne CO$_2$ equivalent.
-The cost allocation shows low electricity prices for consumers in a renewable German system are made achieved on a transparent basis.
+The cost allocation shows low electricity prices for consumers in a renewable German system are achieved on a transparent basis.
 The cost allocation shows how buses remote from wind farms pay higher prices due to increased reliance in transmission and backup capacity.
@@
 
-\subsubsection*{Reproducibility}
+\subsubsection*{Reproducibility and Expansion}
 
 All figures and data points can be reproduced by using the \textit{python} workflow in \cite{hofmann_pypsa-costallocation_2020}.
+The automated workflow allows for higher spatial resolution of the network (scalable up to a the full ENTSO-E Transmission System Map) and optionally taking the total European power system into account.
 
@@
 The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL).
-Alternatively, we can the demand $\demand$ has to be supplied by the attached assets,
+With a given demand $\demand$ this translates to
 \begin{align}
@@
 where $\incidence$ is +1 if line $\ell$ starts at bus $n$, -1 if it ends at $n$, 0 otherwise.
-The nodal generation $\nodalgeneration$ collects all nodal production, see \cref{eq:nodalgeneration}.
+The nodal generation $\nodalgeneration$ collects the production of all nodal assets, see \cref{eq:nodalgeneration}.
 The shadow price of the nodal balance constraint mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step.
-In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\
+In a optimal nodal pricing scheme this is the \euro/\megawatthour price which a consumer has to pay.\\
 
@@
 where $h_j \left(\state, \capacity \right)$ denotes all inequality constraints attached to $\state$ and $\capacity$.
-Note that according to \cref{eq:nodalgeneration} $\nodalgeneration$ is a composition of $\state$ and $\flow = \state[\ell] \forall \ell \in L$.
-
 In order to impose the Kirchhoff Voltage Law (KVL) for the linearized AC flow, the term
@@
 
-The stationarity of the transmission capacity to
+The stationarity of the transmission capacity leads to
 \begin{align}
@@
 \section{Working Example}
-The following figures contain more detailed information about the peer-to-peer cost allocation.
+The following figures contain more detailed information about the peer-to-peer cost allocation discussed in \cref{sec:application_case}.
 The cost or prices payed by consumers are indicated by the region color.

### 2020-10-26 paper/flow_based.tex
follow up

@@
 \title{Tracing prices: A flow-based cost allocation for optimized power systems}
-\author{F. Hofmann}
+\author{Fabian Hofmann\\Frankfurt Institute for Advanced Studies (FIAS), 60438 Frankfurt, Germany\\
+Email: hofmann@fias.uni-frankfurt.de}
 
@@
         Power system models are a valuable and widely used tool to determine cost-minimal future operation and investment under political or ecological boundary conditions.
-Yet they are silent about the allocation of costs of single assets, \ie generators, transmission lines etc., to consumers in the network.
+Yet they are silent about the allocation of costs of single assets, as generators or transmission lines, to consumers in the network.
 Existing cost-allocation methods hardly suit large networks and do not take all relevant costs into account.
 This paper bridges this gap.
-Based on flow tracing, it introduces a peer-to-peer allocation of all costs in an optimized power system to consumers.
+Based on flow tracing, it introduces a peer-to-peer or more precisely an asset-to-consumer allocation of all costs in an optimized power system.
 The resulting cost allocation is both locally constrained and aligned with locational marginal prices in the optimum.
@@
 where an ``asset'' describes any operating components of the network, such as a generator, line, energy storage etc. We refer to the set of all assets as $I$. 
-In a long-term equilibrium of a power system with perfect competition, the zero-profit condition states that each cost term $\cost_{i}$ can be considered as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
+In a long-term equilibrium of a power system with perfect competition and no further constraints, the zero-profit condition states that each cost term $\cost_{i}$ is recovered by the revenue that asset $i$ receives from the market \cite{steiner_peak_1957}.
+As partly shown in \cite{brown_decreasing_2020}, it builds on the fact that $\cost_i$ can be expressed as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
 \begin{align}

### 2020-10-26 submission-joule/paper/main.tex
add submission files

--- 2020-10-26 paper/flow_based.tex
@@
-\documentclass[11pt,twocolumn]{article}
+\documentclass[preprint,5p,12pt]{elsarticle}
 \usepackage{graphicx}
-\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
+% \usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
 \usepackage{amsmath,bm}
 \usepackage[colorlinks]{hyperref}
-\usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8, style=ieee]{biblatex}
+% \usepackage{biblatex}
 \usepackage{eurosym}
@@
 
-\addbibresource{main.bib}
-\graphicspath{{../figures/}{../figures/example/}}
+\hypersetup{
+    colorlinks = true,
+    linkbordercolor = {white},}
+
+% \addbibresource{main.bib}
+\graphicspath{{figures/}{figures/example/}}
 
@@
 
+\begin{frontmatter}
+
 \title{Tracing prices: A flow-based cost allocation for optimized power systems}
-\author{Fabian Hofmann\\Frankfurt Institute for Advanced Studies (FIAS), 60438 Frankfurt, Germany\\
-Email: hofmann@fias.uni-frankfurt.de}
-
-\twocolumn[
-  \begin{@twocolumnfalse}
-
-    \maketitle
+
+\author{Fabian Hofmann\corref{cor1}}
+\ead{hofmann@fias.uni-frankfurt.de}
+\address{Frankfurt Institute for Advanced Studies (FIAS), 60438 Frankfurt, Germany}
 
 \begin{abstract}
-        % Cost minimization of power systems leads to perfectly scheduled operations and investments while securing all power demands. 
-        % System needs reports from regional, national and international agencies increasingly rely on cost minimization techniques.
-However, cost impact to single regions and thefore the question of who profits to what extent from an investment is often neglected.
-The literature provides multiple cost allocation methods dedicated to this problem, yet they hardly suit for large networks and an integration of all costs.
-        % This work introduces a peer-to-peer allocation of all system costs in optimized models.
-Combining basic relations from economic optimization with flow allocation, its cost assignments are locally constraint while still being aligned with the Locational Marginal Prices at the optimum.
-It is showcased by means of an optimized near-future scenario of the German power system.
     Power system models are a valuable and widely used tool to determine cost-minimal future operation and investment under political or ecological boundary conditions.
@@
 The approach is applied and discussed using a future German scenario.
-        \vtop{%
-         \vskip0pt
-          \hbox{%
-          \includegraphics[width=\linewidth]{graphical_abstract.png}
-          }%
-        }
-
 \end{abstract}
 
-\end{@twocolumnfalse}
-]
-
-\subsection*{Highlights}
-\begin{itemize}
-    \item In long-term equilibria a network asset recovers its variable and investment cost from its operation based revenue
-    \item Flow tracing is used as a basis to allocate the operation of assets to consumers. 
-    \item Allocated flows have to be reshaped such that the Kirchhoff Current Law and the Kirchhoff Voltage Law are respected. 
-    \item Using operational and shadow prices from constraints, all costs are assigned in an P2P manner.
-    \item The cost assignments are locally constraint and in alignment with the nodal pricing scheme of the optimum.    
+\begin{graphicalabstract}
+    \includegraphics[width=.9\textwidth]{graphical_abstract.png}
+\end{graphicalabstract}
+
+\begin{highlights}
+    \item Expenditures to a network asset are recovered by  its operation based revenue
+    \item Flow tracing is used to allocate the operation of assets to consumers 
+    \item Allocated flows have to be reshaped to meet the Kirchhoff circuit laws
+    \item Using operational and shadow prices, all system costs are assigned in an P2P manner
+    \item Cost assignments are local while aliging with the nodal pricing scheme of the optimum    
     % \item Upper and lower capacity expansion limits modify the revenue per asset, which however correctly reflects in the cost allocation
-    \item In a low-carbon scenario for Germany, regions with high renewable potentials profit from investments compensated by remote buses.   
-\end{itemize}
+    % \item In a low-carbon scenario for Germany, regions with high renewable potentials profit from investments compensated by remote buses.   
+\end{highlights}
+
+\end{frontmatter}
 
@@
 It is central to identify the drivers of costs and address them appropriately.
-In this respect, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019}.
+In this respect, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019,collins_impacts_2018,kozarcanin_21st_2019}.
 Many studies for countries and regions throughout the world exist that lay out how renewable energy penetration can be expanded at minimum costs.
@@
 
-The first section presents the zero-profit condition for different kind of assets (\cref{sec:zero_profit_rule}),  the AP allocation scheme and derived allocations from asset to consumer (\cref{sec:dispatch_allocations}), the cost allocation and the impact of additional constraints (\cref{sec:general_scheme,sec:design_constraints}), a numerical example (\cref{sec:numerical_example}).
-\Cref{sec:application_case} applies the cost allocation to a optimized German power system with a high share of renewable power generation and evaluates the allocated costs.
-
-\section{Flow-based Cost Allocation}
-
-\subsection{Zero-Profit Rule}
+The first section presents the zero-profit condition for different kind of assets (\ref{sec:zero_profit_rule}),  the AP allocation scheme and derived allocations from asset to consumer (\ref{sec:dispatch_allocations}), the cost allocation and the impact of additional constraints (\ref{sec:general_scheme,sec:design_constraints}), a numerical example (\ref{sec:numerical_example}).
+\ref{sec:application_case} applies the cost allocation to a optimized German power system with a high share of renewable power generation and evaluates the allocated costs.
+
+\section{Results and Discussion}
+
+\subsection{Operation Based Revenue}
 \label{sec:zero_profit_rule}
@@
 In a long-term equilibrium of a power system with perfect competition and no further constraints, the zero-profit condition states that each cost term $\cost_{i}$ is recovered by the revenue that asset $i$ receives from the market \cite{steiner_peak_1957}.
-As partly shown in \cite{brown_decreasing_2020}, it builds on the fact that $\cost_i$ can be expressed as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
+It builds on the fact that $\cost_i$ can be expressed as a cost-weighted sum of the operational state $s_{i,t}$ of asset $i$ and time $t$, \ie
 \begin{align}
@@
 These include OPEX \& CAPEX for production, transmission and storage assets in the network, as well as a cost term for the total Green House Gas (GHG) emissions.
-For storage units, an additional cost term adds to $\capexstorage$ (see \cref{sec:zero_profit_storage_units} for details).}
+For storage units, an additional cost term adds to $\capexstorage$ (see \ref{sec:zero_profit_storage_units} for details).}
     \label{tab:cost_allocation_map}
@@
 If $\muuppergeneration$  is bigger than zero, the constraint is binding, which pushes investments in $\capacitygeneration$.
-As shown in \cite{brown_decreasing_2020} and in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is recovered by the production $\generation$ times the shadow price $\muuppergeneration$,
+As shown in \cite{brown_decreasing_2020} and in detail in \ref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is recovered by the production $\generation$ times the shadow price $\muuppergeneration$,
 \begin{align}
@@
 At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$.
-Again, we use the relation that over the whole time span, the shadow prices weighted by the flow match the investment in line $\ell$ (for details see \cref{sec:zero_profit_flow})
+Again, we use the relation that over the whole time span, the shadow prices weighted by the flow match the investment in line $\ell$ (for details see \ref{sec:zero_profit_flow})
 \begin{align}
@@
 The storage $r$ dispatches power with efficiency $\efficiencydispatch$, charges power with efficiency $\efficiencycharge$ and preserves power from one time step $t$ to the next, $t+1$, with an efficiency of $\efficiencysoc$.
-In \cref{sec:zero_profit_storage_units} we formulate the mathematical details.
+In \ref{sec:zero_profit_storage_units} we formulate the mathematical details.
 The OPEX which adds to $\totalcost$ is given by
@@
 where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint.
-Following the considerations in \cref{sec:zero_profit_storage_units} we restrict to the revenue from dispatched power, \ie the first term in \cref{eq:no_profit_capex_storage}, for the cost allocation.
+Following the considerations in \ref{sec:zero_profit_storage_units} we restrict to the revenue from dispatched power, \ie the first term in \cref{eq:no_profit_capex_storage}, for the cost allocation.
 
@@
 This assumption leads to regionally confined allocation $A_{m \rightarrow n}$ based on a straightforward principal.
-A mathematical formulation of the AP scheme is documented in \cref{sec:net_ap}.
+A mathematical formulation of the AP scheme is documented in \ref{sec:net_ap}.
 For a detailed comparison with other schemes we refer to \cite{hofmann_techno-economic_2020}.
@@
 Note that the traced flow based in the AP scheme obeys the Kirchhoff Current Law but not the Kirchhoff Voltage Law, as already pointed out in \cite{galiana_transmission_2003}.
-However, as we show later a consideration of both laws is needed in order to align the allocated costs with the optimized locational marginal prices (LMP).
+However, as we show later, a consideration of both laws is needed in order to align the allocated costs with the optimized locational marginal prices (LMP).
 To tackle this, let $\ptdf$ denote a linear mapping between the injection $\left(\nodalgeneration[m] - \demand[m]\right)$ and the flow $\flow$, such that
@@
 Mathematically this translates to the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$, $\lmp = \partial \totalcost /\partial \demand$.
-At the optimum this quantity is given by a shadow price of the nodal balance constraint (see \cref{sec:lmp} for details).
+At the optimum this quantity is given by a shadow price of the nodal balance constraint (see \ref{sec:lmp} for details).
 Now, when summing over all assets $i$ and cost terms $\circ$, the cost allocation yields
@@
 which we refer to as the nodal payment or payment.  
-This relation which is in detailed proven in \cref{sec:proof_equivalence} shows that the cost allocation is embedded in the optimized nodal pricing scheme, \ie the nodal payment exactly matches the payment determined by the LMP $\lmp$.
+This relation which is in detailed proven in \ref{sec:proof_equivalence} shows that the cost allocation is embedded in the optimized nodal pricing scheme, \ie the nodal payment exactly matches the payment determined by the LMP $\lmp$.
 
@@
 
-\section{Application Case}
+\subsection{Application Case}
 \label{sec:application_case}
@@
 
-\section{Limitations}
+\subsection{Limitations}
 The presented cost allocation is based on the linear power flow approximation.
@@
 
-\section{Conclusion}
+\subsection{Conclusion}
 
@@
 
-\subsubsection*{Reproducibility and Expansion}
-
-All figures and data points can be reproduced by using the \textit{python} workflow in \cite{hofmann_pypsa-costallocation_2020}.
-The automated workflow allows for higher spatial resolution of the network (scalable up to a the full ENTSO-E Transmission System Map) and optionally taking the total European power system into account.
-
-\subsection*{Funding}
-This research was funded by the by the German Federal Ministry for Economics Affairs and Energy (BMWi) in
-the frame of the NetAllok project (grant number 03ET4046A) \cite{bundesministerium_fur_wirtschaft_und_energie_verbundvorhaben_nodate}.
-
-\subsubsection*{Acknowledgement}
+\section*{Experimental Procedures}
+
+% \subsection*{Resource Availability}
+
+\subsection*{Lead Contact}
+Requests for further information, resources and materials should be directed to Fabian Hofmann (hofmann@fias.uni-frankfurt.de)
+
+\subsection*{Data and Code Availability}
+All figures and data points of this paper can be reproduced by using the \textit{python} workflow published in \cite{hofmann_pypsa-costallocation_2020}. \\
+The automated workflow allows for higher spatial resolution of the network (scalable up to the full ENTSO-E Transmission System Map) and optionally taking the total European power system into account.
+
+\subsection*{Materials Availability}
+This study did not generate any new materials.
+
+\section*{Acknowledgement}
 
@@
 Further, I want to thank Alexander Kies and Markus Schlott who steadily helped with creative thoughts.
+This research was funded by the by the German Federal Ministry for Economics Affairs and Energy (BMWi) in
+the frame of the NetAllok project (grant number 03ET4046A) \cite{bundesministerium_fur_wirtschaft_und_energie_verbundvorhaben_nodate}.
+
+\section*{Author Contributions}
+This work was solely done by Fabian Hofmann.
+
+\section*{Declaration of Interests}
+The author declares no competing interests.
+
+\bibliographystyle{elsarticle-num} 
+\bibliography{main.bib}
 
@@
 \section{Working Example}
-The following figures contain more detailed information about the peer-to-peer cost allocation discussed in \cref{sec:application_case}.
+The following figures contain more detailed information about the peer-to-peer cost allocation discussed in \ref{sec:application_case}.
 The cost or prices payed by consumers are indicated by the region color.
@@
     \centering
-    \input{../tables/de50bf_ptpf_net/prices}    
+    \input{tables/de50bf_ptpf_net/prices}    
     \caption{Operational and capital price assumptions for all type of assets used in the working example.
@@
 \begin{figure*}
-    \includegraphics[width=\linewidth]{de50bf/capex_duration_curve}
+    \includegraphics[width=.9\linewidth]{de50bf/capex_duration_curve}
     \caption{Duration curve of the CAPEX allocation for production and storage technologies.
@@
 
-\clearpage
-\printbibliography
-
 \end{document}

### 2021-05-12 code/example-price-tracing.py
update example section

@@ -29,7 +29,7 @@ n.madd('Generator', [1, 3], bus=[1, 3], p_nom=[50, 50],
 n.madd("Load", [1,2], bus=[1,2], p_set=[30,50])
 n.madd('Line', ['1-2', '3-1', '3-2',],
        bus0=[1, 3, 3,], bus1=[2, 1, 2,],
-       s_nom=[40, 15, 30], x=[0.01, 0.01, 0.01], )
+       s_nom=[60, 30, 30], x=[0.01, 0.01, 0.01], )
 
 n.lopf(pyomo=False, solver_name='gurobi', keep_shadowprices=True,
         keep_references=True)
@@ -113,7 +113,7 @@ fig.savefig(figures / 'network.png', bbox_inches='tight')
 
 C = ntl.allocate_revenue(n, with_cycle_prices=True)
 
-to_index = pd.Series(dict(payer='n', receiver_node='m', receiver_branch='$\ell$'))
+to_index = pd.Series(dict(payer='n', receiver_node='s', receiver_branch='$\ell$'))
 payoff = C.sum(['snapshot', 'receiver_carrier'])\
            .assign_coords(receiver_branch=['1-2', '1-3', '2-3'])\
            .rename(to_index).transpose(..., 'n') \
@@ -132,10 +132,37 @@ for i, v in enumerate(payoff):
                 annot_kws={'horizontalalignment': 'left'})
     if i:
         ax.set_ylabel('')
-    ax.set_title(v.replace("_", " ").title())
+    # ax.set_title(v.replace("_", " ").title())
+    ax.tick_params(left=False, bottom=False)
+fig.tight_layout(w_pad=.4)
+fig.savefig(figures / 'example_payoff-kvl.png', bbox_inches='tight')
+
+# %% revenue allocation wo KVL prices
+
+C = ntl.allocate_revenue(n, with_cycle_prices=False)
+
+to_index = pd.Series(dict(payer='n', receiver_node='s', receiver_branch='$\ell$'))
+payoff = C.sum(['snapshot', 'receiver_carrier'])\
+           .assign_coords(receiver_branch=['1-2', '1-3', '2-3'])\
+           .rename(to_index).transpose(..., 'n') \
+           .drop_sel(n='3')
+vmax = payoff.to_array().max()
+
+fig, axes = plt.subplots(1, 2, figsize=(4,3), sharey=True,
+                         gridspec_kw={'width_ratios':[0.5,.5,]})
[…]

### 2021-08-12 code/allocate_network.py
a lot of review, restructuring

@@ -27,7 +27,7 @@ from config import source_dims
 if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
-        snakemake = mock_snakemake('allocate_network', nname='test-de10')
+        snakemake = mock_snakemake('allocate_network', nname='de50')
 
 
 n = pypsa.Network(snakemake.input.network)

### 2021-08-12 code/example_without_cycles.py

@@ -28,8 +28,8 @@ if 'snakemake' not in globals():
 n = pypsa.Network()
 n.madd("Bus", range(1,4), x=[0,0,0.5], y=[0,0.7, 0.35])
 n.madd('Generator', [1, 3], bus=[1, 3], p_nom=[50, 50],
-       marginal_cost=[6, 4])
-n.madd("Load", [1,2], bus=[1,2], p_set=[30,50])
+       marginal_cost=[6, 4], carrier='gen')
+n.madd("Load", [1,2], bus=[1,2], p_set=[30,50], carrier='load')
 n.madd('Line', ['1-2', '3-1', ],
        bus0=[1, 3,], bus1=[2, 1,],
        s_nom=[60, 30, ], x=[0.01, 0.01,], )

### 2021-08-12 code/solve_and_sanitize_network.py

@@ -14,7 +14,7 @@ if __name__ == "__main__":
     if 'snakemake' not in globals():
         from _helpers import mock_snakemake
         snakemake = mock_snakemake('solve_and_sanitize_german_network',
-                                   clusters=10)
+                                   clusters=50)
 
 
 solver_opts = snakemake.config['solving']['solver']
@@ -28,11 +28,8 @@ for c in ['Generator', 'StorageUnit']:
     n.df(c).p_nom_max.update(n.df(c).query('not p_nom_extendable').p_nom)
     n.df(c)['p_nom_extendable'] = True
 
-n.mremove('Line', n.lines.query('num_parallel == 0').index)
+n.lines.num_parallel = n.lines.num_parallel.clip(lower=1)
 
-for c, attr in pypsa.descriptors.nominal_attrs.items():
-    n.df(c)[attr + '_min'] = n.df(c)[attr]
-    n.df(c)[attr] = 0
 
 n.lopf(pyomo=False, solver_name=solver_name, solver_options=solver_opts,
         keep_shadowprices=True)

### 2021-08-17 code/allocate_network.py
further work

@@ -47,7 +47,7 @@ if 'lv_limit' in n.global_constraints.index:
         n.df(c)['mu_upper_'+nom] += mu_upper
 
 
-ds = ntl.allocate_flow(n, method=method, aggregated=aggregated)  # q=0
+ds = ntl.allocate_flow(n, method=method, aggregated=aggregated, dask=True)  # q=0
 
 if snakemake.config['alloc_to_load_only']:
     ds = expand_by_sink_type(ds.chunk(dict(snapshot=5)), n)

### 2021-08-17 code/check_shadowprices.py

@@ -84,14 +84,14 @@ if not 'test' in snakemake.input[0]:
                  n.buses_t.marginal_price[bus])
 
 
-    gamma = (st.mu_lower_p_dispatch + st.mu_upper_p_dispatch +
-             st.mu_state_of_charge / s.efficiency_dispatch)
-    investment = s.eval('p_nom_opt * capital_cost')
-    # gamma + marginal_cost is the lmp
-    revenue = ((gamma) * st.p_dispatch - (gamma + s.marginal_cost) * st.p_store).sum()
-    subsidy = s.p_nom_opt * s.mu_lower_p_nom
-    scarcity = s.p_nom_opt * s.mu_upper_p_nom
-    assert close(investment, revenue + subsidy + scarcity)
+    # gamma = (st.mu_lower_p_dispatch + st.mu_upper_p_dispatch +
+    #          st.mu_state_of_charge / s.efficiency_dispatch)
+    # investment = s.eval('p_nom_opt * capital_cost')
+    # # gamma + marginal_cost is the lmp
+    # revenue = ((gamma) * st.p_dispatch - (gamma + s.marginal_cost) * st.p_store).sum()
+    # subsidy = s.p_nom_opt * s.mu_lower_p_nom
+    # scarcity = s.p_nom_opt * s.mu_upper_p_nom
+    # assert close(investment, revenue + subsidy + scarcity)
 
 
 

### 2021-08-17 code/plot_network.py

@@ -23,6 +23,8 @@ if __name__ == "__main__":
         snakemake = mock_snakemake('plot_network', nname='de50')
 
 
+plt.rc('axes', titlesize='medium')
+
 n = pypsa.Network(snakemake.input.network)
 regions = gpd.read_file(snakemake.input.regions)
 

### 2021-08-17 code/solve_and_sanitize_network.py

@@ -8,7 +8,7 @@ Created on Fri Jul 10 22:16:58 2020
 
 import pypsa
 import shutil
-
+import numpy as np
 
 if __name__ == "__main__":
     if 'snakemake' not in globals():
@@ -30,6 +30,7 @@ for c in ['Generator', 'StorageUnit']:
 
 n.lines.num_parallel = n.lines.num_parallel.clip(lower=1)
 
+n.generators.p_nom_max.fillna(np.inf, inplace=True)
 
 n.lopf(pyomo=False, solver_name=solver_name, solver_options=solver_opts,
         keep_shadowprices=True)

### 2021-09-23 code/plot_allocated_payment_comparison.py
insert alex comments

@@ -1,0 +1,140 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Aug 10 12:53:08 2020
+
+@author: fabian
+"""
+
+import pypsa
+import geopandas as gpd
+import pandas as pd
+import numpy as np
+import xarray as xr
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+import matplotlib as mpl
+import pypsa
+from config import color, to_symbol, sink_dims, to_explanation
+import netallocation as ntl
+from matplotlib.pyplot import Line2D
+from pypsa.plot import projected_area_factor
+from netallocation.plot_helpers import (make_handler_map_to_scale_circles_as_in,
+                                        make_legend_circles_for)
+
+
+if 'snakemake' not in globals():
+    from _helpers import mock_snakemake
+    snakemake = mock_snakemake('plot_allocated_payment', nname='de50bf',
+                               method='ptpf', power='net', sink='DE0 1')
+
+
+n = pypsa.Network(snakemake.input.network)
+costs = xr.open_dataset(snakemake.input.costs)
+sink = snakemake.wildcards.sink
+regions = gpd.read_file(snakemake.input.regions).set_index('name')
+
+fig, axes = plt.subplots(1, 2, figsize=(8,4.5),
+                         subplot_kw={'projection': ccrs.EqualEarth()})
+
+for sink, ax in zip(['lowest-lmp', 'highest-lmp'], axes):
+    if sink == 'lowest-lmp':
+        sink = n.buses_t.marginal_price.mean().idxmin()
+    elif sink == 'highest-lmp':
+        sink = n.buses_t.marginal_price.mean().idxmax()
+
+
+    payment = (xr.open_dataset(snakemake.input.costs)
+               .sel(sink=sink, drop=True)
+               .fillna(0)
+               .sum('snapshot')
+               .rename(source='bus', source_carrier='carrier')
+               .set_index(branch=['component', 'branch_i']))
+
+
+    branch_widths = payment.branch_investment_cost.to_pandas()
+    branch_colors = (np.sign(payment.branch_investment_cost.to_pandas())
+                     .replace({1: 'cadetblue', -1:'indianred'}))
+
+    emission_cost = payment[['co2_cost']].sum('carrier').to_dataframe().stack()
+    capex = pd.concat([payment.generator_investment_cost.to_series(),
[…]

### 2022-03-16 submission-energy/paper/main.tex
add elsevier preprint template

\documentclass[preprint,12pt]{elsarticle}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath,bm}
\usepackage[colorlinks]{hyperref}
% \usepackage[natbib=true, sorting=none, backend=biber, bibencoding=utf8, style=ieee]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{booktabs} % for table (toprule, bottomrule)
\usepackage{enumitem} % for alphabetical enumeration 
\usepackage{accents}
\usepackage{tikz}
\usepackage{dblfloatfix}
\usepackage[printonlyused]{acronym}
\usepackage{cleveref}

\graphicspath{{figures/}}

\usetikzlibrary{arrows}
\tikzset{
    vertex/.style={circle,draw,minimum size=1.5em},
    edge/.style={->,> = latex',very thick}
    }

    \crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}
\crefname{subequations}{Eqs.}{Eqs.}
\creflabelformat{subequations}{(#2#1#3)}

\setlength\parindent{8pt}

% style operators
\newcommand{\ie}{i.e.}
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}
\newcommand{\resultsin}[1]{\hspace{6pt} \leftrightarrow  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ra}{\rightarrow}

% general symbols
\newcommand{\state}[1][i]{s_{#1,t}}
\newcommand{\capacity}{S_{i}}
\newcommand{\costfactor}{\gamma^\circ_{i,t}}
\newcommand{\capacityupper}{\bar{S}}
\newcommand{\capacitylower}{\ubar{S}}
\newcommand{\muuppernom}{\bar{\mu}^\text{nom}_{i}}
\newcommand{\mulowernom}{\ubar{\mu}^\text{nom}_{i}}

\newcommand{\kk}{k~\euro~}

%generation
\newcommand{\generation}{g_{s,t}}
\newcommand{\generationpotential}{\bar{g}_{s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacitygeneration}{G_{s}}
\newcommand{\capacitygenerationupper}{\bar{G}_{s}}
\newcommand{\capacitygenerationlower}{\ubar{G}_{s}}
\newcommand{\operationalpricegeneration}{o_{s}}
\newcommand{\capitalpricegeneration}{c_{s}}
\newcommand{\mulowergeneration}{\ubar{\mu}_{s,t}}
\newcommand{\muuppergeneration}{\bar{\mu}_{s,t}}
\newcommand{\muuppergenerationnom}{\bar{\mu}^\text{nom}_{s}}
\newcommand{\mulowergenerationnom}{\ubar{\mu}^\text{nom}_{s}}

% flow
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\capacityflow}{F_{\ell}}
\newcommand{\capacityflowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityflowLower}{\ubar{F}_{\ell}}
\newcommand{\operationalpriceflow}{o_\ell}
\newcommand{\capitalpriceflow}{c_{\ell}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\muupperflownom}{\bar{\mu}^\text{nom}_{\ell}}
\newcommand{\mulowerflownom}{\ubar{\mu}^\text{nom}_{\ell}}

% storage
\newcommand{\storage}{g_{r,t}}
\newcommand{\storagedispatch}{\storage^\text{dis}}
\newcommand{\storagecharge}{\storage^\text{sto}}
\newcommand{\storagesoc}{\storage^\text{ene}}
\newcommand{\storageprevioussoc}{g_{r,t-1}^\text{ene}}

\newcommand{\efficiency}{\eta_{r}}
\newcommand{\efficiencydispatch}{\efficiency^\text{dis}}
\newcommand{\efficiencycharge}{\efficiency^\text{sto}}
\newcommand{\efficiencysoc}{\efficiency^\text{ene}}

\newcommand{\operationalpricestorage}{o_r}
\newcommand{\capitalpricestorage}{c_r}
\newcommand{\capacitystorage}{G_r}
\newcommand{\capacitystorageupper}{\bar{G}_r}
\newcommand{\capacitystoragelower}{\ubar{G}_r}
\newcommand{\mulowerstorage}{\ubar{\mu}_{r,t}}
\newcommand{\muupperstorage}{\bar{\mu}_{r,t}}
\newcommand{\mulowerstoragedispatch}{\ubar{\mu}_{r,t}^\text{dis}}
\newcommand{\muupperstoragedispatch}{\bar{\mu}_{r,t}^\text{dis}}
\newcommand{\mulowerstoragecharge}{\ubar{\mu}_{r,t}^\text{sto}}
\newcommand{\muupperstoragecharge}{\bar{\mu}_{r,t}^\text{sto}}
\newcommand{\mulowerstoragesoc}{\ubar{\mu}_{r,t}^\text{ene}}
\newcommand{\muupperstoragesoc}{\bar{\mu}_{r,t}^\text{ene}}
\newcommand{\muupperstoragenom}{\bar{\mu}^\text{nom}_r}
\newcommand{\mulowerstoragenom}{\ubar{\mu}^\text{nom}_r}
\newcommand{\mustateofcharge}{\lambda^\text{ene}_{r,t}}
\newcommand{\munextstateofcharge}{\lambda^\text{ene}_{r,t+1}}

% other
\newcommand{\lagrangian}{\mathcal{L}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\lmpdiff}[1][\ell]{\lmp[#1]^\text{diff}}
\newcommand{\lmpkvl}[1][\ell]{\lmp[#1]^\text{KVL}}
\newcommand{\averagelmp}[1][n]{\bar{\lambda}_{#1}}
\newcommand{\demand}[1][n]{d_{#1,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\injection}[1][n]{p_{#1,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}
\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}

% network
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\incidencegenerator}[1][n]{K_{#1,s}}
\newcommand{\incidencestorage}[1][n]{K_{#1,r}}
\newcommand{\incidenceasset}[1][n]{K_{#1,i}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\reactance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}

% Cost Terms
\newcommand{\emission}{e_{s}}
\newcommand{\emissionprice}{\mu_{\text{CO2}}}
\newcommand{\megawatthour}{MWh$_\text{el}$}
\newcommand{\totalcost}{\mathcal{T}}

\newcommand{\cost}{\mathcal{C}}
\newcommand{\payment}[1][n]{\cost_{#1,t}}

\newcommand{\opex}{\mathcal{O}}
\newcommand{\opexgeneration}{\mathcal{O}^G}
\newcommand{\opexflow}{\mathcal{O}^F}
\newcommand{\opexstorage}{\mathcal{O}^E}
\newcommand{\capex}{\mathcal{I}}
\newcommand{\capexgeneration}{\mathcal{I}^G}
\newcommand{\capexflow}{\mathcal{I}^F} 
\newcommand{\capexstorage}{\mathcal{I}^E}
\newcommand{\emissioncost}{\mathcal{E}}
\newcommand{\scarcitycost}{\mathcal{S}}
\newcommand{\subsidycost}{\mathcal{U}}
\newcommand{\remainingcost}{\mathcal{R}}

% allocation quantities
\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
\newcommand{\allocatestoragedispatch}[1][r, n]{A_{#1,t}}
\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}

\newcommand{\allocatecost}[1][n \rightarrow i]{\cost_{#1,t}}
\newcommand{\allocateassetcost}[1][n \rightarrow i]{\cost_{#1, t}}
\newcommand{\allocategeneratorcost}[1][n \rightarrow s]{\cost_{#1, t}}
\newcommand{\allocatelinecost}[1][n \rightarrow \ell]{\cost_{#1, t}}

\newcommand{\allocatecapex}[1][n \rightarrow s,t]{\mathcal{I}_{#1}}
\newcommand{\allocatecapexgeneration}[1][n \rightarrow s,t]{\capexgeneration_{#1}}
\newcommand{\allocatecapexflow}[1][n \rightarrow \ell,t]{\capexflow_{#1}}
\newcommand{\allocatecapexstorage}[1][n \rightarrow r, t]{\capexstorage_{#1}}
\newcommand{\allocateopex}[1][n \rightarrow s]{\opex_{#1,t}}
\newcommand{\allocateemissioncost}[1][n \rightarrow s]{\emissioncost_{#1,t}}
\newcommand{\allocatescarcitycost}[1][n \rightarrow s]{\scarcitycost_{#1,t}}
\newcommand{\allocatesubsidycost}[1][n \rightarrow s]{\subsidycost_{#1,t}}

\journal{Energy}

\begin{document}

\begin{frontmatter}

    \title{Price Tracing: Linking Nodal Prices in Optimized Power Systems}

    \author[tub,fias]{Fabian Hofmann}
    \ead{hofmann@fias.uni-frankfurt.de}

    \author{Markus Schlott}
    \ead{schlott@fias.uni-frankfurt.de}

    \affiliation[tub]{organization={Technical University Berlin, Department of Energy Systems},
        city={Berlin},
        postcode={10623},
        country={Germany}}

    \affiliation[tub]{organization={Frankfurt Institute for Advanced Studies (FIAS)},
        city={Frankfurt},
        postcode={60438},
        country={Germany}}

    \begin{abstract}
        Optimizing the total cost of power systems is a common tool for network operation and planning. Besides valuable information about how to run and possibly expand a power system, the optimization provides an optimal Locational Marginal Price per node and time step. This price can be seen as the price of electricity paid by consumers and purchased by suppliers, while maximizing social welfare. Naturally, it is a direct result of the optimization problem, and therefore does not give any information about its internal composition. This paper shows that by applying Flow Tracing, an algorithm for tracking flows in complex networks, it is possible to interlink Locational Marginal Prices in a coherent mathematical way. This does not only lead to important insights into the price structure, but also provides an intuitive decomposition and allocation of all system costs. Then individual consumers ``see'' how much they have to pay to individual generators and transmission lines in the power system. This method, introduced as Price Tracing, outperforms similar approaches provided by the literature, since the resulting cost allocations are transparent, plausible, and consistent with the Locational Marginal Prices from the optimization. The Price Tracing method is applied and discussed on behalf of a power system model of Germany with a high share of renewable power. The presented analysis and its implications can help in finding a more efficient market design to promote renewable power supply.
    \end{abstract}

    \begin{highlights}
        \item Flow Tracing is used to interlink optimal Locational Marginal Prices.
        \item This quantifies the payment of individual consumers to individual generators and transmission lines.
        \item Four use-cases are proposed and demonstrated on behalf of a future model of the German power system.
    \end{highlights}

    \begin{keyword}
        keywords here, in the form: keyword \sep keyword

        %% PACS codes here, in the form: \PACS code \sep code

        %% MSC codes here, in the form: \MSC code \sep code
        %% or \MSC[2008] code \sep code (2000 is the default)

    \end{keyword}

\end{frontmatter}

\section{Introduction}

Today's energy systems are undergoing profound and enduring change. The transition from controllable to variable, weather-dependent power generation and the constant improvement and innovation of technology require rigorous system planning and international cooperation. The core of the challenge manifests itself in the total cost of the system. First, these should be as low as possible, while meeting environmental and techno-economic standards. Second, they must be distributed fairly and transparently among all actors in the power system. It is essential to identify the cost drivers and account for them appropriately.
In this regard, power system models are a valuable and widely used tool \cite{pfenninger_energy_2014,bazmi_sustainable_2011,pereira_generation_2017,brown_sectoral_2019}. Many studies looking at different countries and regions around the world show how to expand renewable energy penetration at minimal cost. However, they are largely silent on how and for what reasons these costs are passed on to consumers.

This paper addresses this gap. We introduce a new method called Price Tracing that links electricity prices from different nodes. The method builds on Bialek's \ac{AP} \cite{bialek_tracing_1996}, also known as flow tracing, and weights the resulting power flow allocations with the \ac{LMP}. This provides the basis for a transparent allocation of cost contributions to consumer electricity prices, i.e. it answers not only the question of who supplies electricity to whom, but also how much consumers must pay to particular generators or transmission lines.

In the literature, the concept of flow-based cost allocation has been discussed and applied in a number of works \cite{galiana_transmission_2003,shahidehpour_market_2002,meng_investigation_2007,schafer_allocation_2017,nikoukar_transmission_2012,arabali_pricing_2012,wu_locational_2005}. Shahidehpour et al. provide deep insights into the allocation of congestion costs and transmission investments to market participants using different allocation techniques \cite{shahidehpour_market_2002}. In particular, they suggest that generation shift factors, i.e. the marginal contribution of generators to a flow on a line, allow the \ac{LMP} to be represented as a superposition of (a) the \ac{LMP} on the reference bus, (b) the congestion price, and (c) a price for losses. The approach in \cite{meng_investigation_2007} extends this relationship for contributions based on the \ac{AP} scheme, which provides an accurate, albeit imprecise, estimate of the optimal \ac{LMP}. A similar approach based on \ac{AP} is used in \cite{tranbergFlowbasedNodalCost2018}, which allocates electricity prices from a non-optimal electricity dispatch.

The presented approach combines the advantages of the above studies. It ensures that cost allocations are plausible and spatially constrained, while payments are fully aligned with the \ac{LMP}-based nodal pricing system. It also facilitates transparency and cost-benefit analysis in network planning, such as the ten-year network development plan \cite{entso-e_completing_2020} or the German network development plan \cite{bundesnetzagentur_netzentwicklungsplan_2020}. In addition, it provides a starting point for a usage-based allocation of transmission costs. While the presented method uses the \ac{LOPF} to analyze large network models, it is discussed how it can be extended to nonlinear power flows.

First, we formulate the price-tracing method (\cref{sec:price_tracing}) with the underlying assumptions and a numerical example (\cref{sec:numerical_example}). In \cref{sec:application_case}, possible use cases are presented using an optimized German power system with a high share of renewable resources. \Cref{sec:limitations} shows methodological limitations and \cref{sec:conclusions} draws final conclusions.

\section{Price Tracing}
\label{sec:price_tracing}

\subsection{Mathematical Model}

Consider an electricity network model with N nodes, L lines and T time steps. Using the linearized power flow approximation, the power flow $\flow$ on a passive line $\ell$ at time $t$ relates to the generation $\nodalgeneration$ and demand $\demand$ at node $n$ and time $t$ according to
\begin{align}
    \flow = \sum_n \ptdf \left(\nodalgeneration - \demand \right) \Forall{t}
\end{align}
where $\ptdf$ are the \ac{PTDF}. These translate the nodal injection on the right-hand side to the network flow in compliance with the linearized Kirchhoff Circuit Laws.
% \ac{KCL} and \ac{KVL}.
% \footnote{for transport models or networks with \acl{HVDC} lines, these can be artificially derived from the network flow using the formulation in \cite{hofmann_flow_2020}}

In a nodal pricing scheme, the system cost occasioned by the electrical demand $\demand$ is proportional to the corresponding \ac{LMP} $\lmp$,
\begin{align}
    \text{Demand Cost}: \lmp\, \demand .
    \label{eq:demand_cost}
\end{align}
On the other hand, system assets generate revenue from their operation. Therefore, the revenue of the dispatch $\nodalgeneration[m]$ at bus $m$ is given by
\begin{align}
    \text{Dispatch Revenue}: \lmp[m] \, \nodalgeneration[m]
    \label{eq:dispatch_revenue}
\end{align}
and the congestion revenue of line $\ell$ at time $t$ by
\begin{align}
    \text{Congestion Revenue}: \lmp[\ell]\, \flow .
    \label{eq:congestion_revenue}
\end{align}
where, in the absence of network cycles, the revenue per transported MWh $\lmp[\ell]$ is the price difference $\lmpdiff$ between the end node and beginning node of line $\ell$. In the case of network cycles, the price of the \ac{KVL} $\lmpkvl$ may be added to $\lmp[\ell]$ to match congestion revenues to expenditures in a long-term equilibrium, see \cref{sec:problem_formulation,sec:zero_profit_flow} for details. The \ac{LMP} as well as the prices for the \ac{KVL} are given by the dual values, often called shadow prices, of the corresponding constraints in the underlying cost optimization (see \cref{sec:optimality_conditions} for details).
% Using the Incidence Matrix $\incidence$, which is 1 if $n$ is a starting node of line $\ell$, -1 if $n$ is a ending node of line $\ell$ and zero otherwise, the price differences are given by $\lmpdiff = - \sum_n \incidence \lmp$.
In economic equilibrium, prices naturally arise, which in turn leads to the following equivalency
\begin{align}
    \underbrace{\sum_{n} \lmp\, \demand}_{\text{Total electricity cost at $t$}}  = \underbrace{\sum_{m} \lmp[m]\, \nodalgeneration[m] + \sum_{\ell} \lmp[\ell] \flow}_{\text{Total revenue at $t$}}\, \Forall{t} .
    \label{eq:total-demand-cost}
\end{align}
While this equality relates the totals of costs and revenues, the relationships between individual contributions on the left-hand side and individual contributions on the right-hand side remain undefined. The following shows that by allocating flows in the network, it is possible to relate the power consumption at single nodes to the revenues of single generators and transmission lines:

Dispatch and flow can be considered as a superposition of individual contributions of nodes and assets. The literature provides various methods, called \ac{FA} method, to artificially quantify these. Each method follows a particular set of assumptions that lead to \ac{P2P} allocations $A_{m \rightarrow n}$ which measure the power generated at node $m$ and consumed at node $n$. Flow Tracing, also known as \ac{AP} \cite{bialek_tracing_1996}, is a flow allocation method that traces the power injection at bus $m$ through the network up to its sink $n$ using the principle of proportional sharing. The method is illustrated in \cref{fig:ap-scheme} and mathematically documented in \cref{sec:net_ap}.
\begin{figure}
    \centering
    \begin{tikzpicture}

        % vertices
        \node[vertex, draw=Orange!60, fill=Orange!60] (1) at (0,0) {$+2$};
        \node[vertex, draw=MidnightBlue!60, fill=MidnightBlue!60] (2) at (0,4) {$+6$};
        \node[vertex, draw=Gray!60, fill=Gray!60] (3) at (3,2) {$-2$};
        \node[vertex, draw=Gray!60, fill=Gray!60] (4) at (6,0) {$-2$};
        \node[vertex, draw=Gray!60, fill=Gray!60] (5) at (6,4) {$-4$};

        %edges
        \draw[edge, draw=Orange!60] (1) -- (3) node[midway, above] {\color{Orange}$2$};
        \draw[edge, draw=MidnightBlue!60] (2) -- (3) node[midway, above] {\color{MidnightBlue}$6$};
        \draw[edge, draw=MidnightBlue!60] (3.330) -- (4.140) node[midway, above] {\color{MidnightBlue}$3/2$};
        \draw[edge, draw=Orange!60] (3.310) -- (4.160) node[midway, below] {\color{Orange}$1/2$};
        \draw[edge, draw=MidnightBlue!60] (3.050) -- (5.200) node[midway, above] {\color{MidnightBlue}$3$};
        \draw[edge, draw=Orange!60] (3.030) -- (5.220) node[midway, below] {\color{Orange}$1$};
    \end{tikzpicture}
    \caption{Schematic illustration of the \ac{AP} method. Figuratively, each power source is given one color. As soon as different power flows mix at a node, the mix of ``colors'' in the outgoing flows is in proportion to the ingoing colors. Therefore, all consumed power can be related to a distinct origin.}
    \label{fig:ap-scheme}
\end{figure}
It states that when flows from different sources meet at the same bus, their share determines the mixture of all flows originating from that bus, including nodal withdrawal. The advantage of the \ac{AP} method is the spatial containment of \ac{P2P} allocations $A_{m \rightarrow n}$ \cite{hofmann_techno-economic_2020}. By definition, the sum of all receivers gives the nodal generation at $m$, thus
\begin{align}
    \nodalgeneration[m] = \sum_n \allocatepeer \Forall{m.t}
    \label{eq:nodalgeneration-breakdown}
\end{align}
As explained later in the paper, only the \ac{P2P} mappings of the \ac{AP} method are used. Thus, regardless of the \ac{AP} allocation, $\allocateflow$ denotes the contribution of demand $\demand$ to the flow on line $\ell$ such that the sum of all flow contributions equals flow on line $\ell$,
\begin{align}
    \flow = \sum_n \allocateflow \Forall{\ell,t}.
    \label{eq:flow-breakdown}
\end{align}
Inserting \cref{eq:nodalgeneration-breakdown,eq:flow-breakdown} into \cref{eq:total-demand-cost} and imposing that the equality holds for all summands referring to $n$ separately, leads to
\begin{align}
    \lmp\, \demand & = \sum_m \lmp[m] \allocatepeer + \sum_\ell \lmp[\ell] \allocateflow \Forall{n,t}.
    \label{eq:demand-cost-allocation}
\end{align}
This equation maps the cost of nodal demand on the left-hand side to the contributions of dispatch and congestion revenues on the right-hand side. It thus indicates what consumers at $n$ have to pay to generators at $m$ and the transmission line $\ell$. The equation introduces N new equalities for each time step for which the necessary degree of freedom comes from the yet undefined flow contribution $\allocateflow$. \Cref{sec:proof_equivalence} proves that
\begin{align}
    \allocateflow = \sum_m \ptdf[m] (\allocatepeer - \delta_{nm} \demand)
    \label{eq:allocateflow}
\end{align}
solves \cref{eq:demand-cost-allocation} where $\delta_{mn}$ is 1 for $m=n$ and zero otherwise. Since \cref{eq:allocateflow} is a function of the \ac{P2P}-assignment $\allocatepeer$ only, the cost of electricity in \cref{eq:demand-cost-allocation} is also only a function of the \ac{P2P} allocation $\allocatepeer$ and the \ac{LMP} $\lmp$.
% This holds true because the flow allocations $\allocateflow$ obey the (linearized) \ac{KVL}\footnote{for transport models or networks with \acl{HVDC} lines, the PTDF can be artificially derived from the network flow using the formulation in \cite{hofmann_flow_2020}}. 
Note that it is not possible to take the non-modified flow allocations given by the \ac{AP} as these are not a solution to \cref{eq:demand-cost-allocation}. Instead it is necessary to reassign them through \cref{eq:allocateflow}\footnote{This can be explained by considering that the \ac{AP} flow allocations do not respect the \ac{KVL}, which however have a direct effect on the \ac{LMP}. For the same reason, the cost estimation in \cite{meng_investigation_2007} are also inaccurate.}.

Since generators are often grouped together at network nodes, the \ac{P2P}-assignments $\allocatepeer$ partition into the contributions of generators located at $m$. Following the formulation in \cite{schafer_tracing_2020}, let $\allocategeneration = \allocatepeer \, r_{s,m,t}$ be the power allocation from generator $s$ to demand at $n$ where $r_{s,m,t}$ is the share that generation $\generation$ of generator $s$ at time $t$ contributes to the generation $\nodalgeneration[m]$ at node $m$ and $t$. It fulfills
\begin{align}
    \generation = \sum_n \allocategeneration .
    \label{eq:generation-breakdown}
\end{align}
Further, let $\lmp[s]$ of generator $s$ be set to the \ac{LMP} $\lmp$, where $s$ is located at bus $n$. For consistency with \cref{eq:dispatch_revenue}, the dispatch revenue of generator $s$ is given by $\lmp[s] \, \generation$.\

The \textit{Price Tracing} method for a network with a linearized power flow and a nodal pricing scheme builds on the above equations and defines
\begin{subequations}
    \begin{align}
        \allocategeneratorcost = \lmp[s] \allocategeneration
    \end{align}
    as the payment from consumers at $n$ to generator $s$ as well as
    \begin{align}
        \allocatelinecost = \lmp[\ell] \allocateflow
    \end{align}
\end{subequations}
as the payment to transmission line $\ell$. For consistency, let $\payment = \lmp \demand$ denote the nodal payment of all consumers at node $n$ which according to \cref{eq:demand-cost-allocation} is $\payment = \sum_s \allocategeneratorcost + \sum_\ell \allocatelinecost$.

It should be noted that storage units such as batteries can easily be included in the Price Tracing method: When they discharge electricity, they are treated as generators, and when they charge electricity, they are treated as consumers. For the latter, the $\allocategeneration$ and $\allocateflow$ allocations must be split into allocations to consumers and charging storage units at $n$.

% Price Tracing enables to track the flow of money in a network with optimal or sub-optimal prices. Note that as long as the allocation $\allocatepeer$ fulfills \cref{eq:flow-breakdown,eq:generation-breakdown}, it may be modified. So, there is a degree of freedom to pre-allocate direct \ac{P2P} tradings $\allocatepeer^\text{Trade}$ and to allocate the remaining generation and demand with the \ac{AP} method, thus 
% \begin{align}
%     \allocatepeer = \allocatepeer^\text{Trade} + \allocatepeer^\text{AP} .
% \end{align}
% This optional modification leads to a Price Tracing which includes exchanges between \ac{P2P} traders as well as ordinary market participants.

\subsection{Numerical Example}
\label{sec:numerical_example}

In the following, the Price Tracing method is illustrated by a numerical example, assuming a three-node network with two lines. Considering one time step, the dispatch of the generators is optimized and the costs are allocated. Finally, a third line is added to the system and the cost allocation is repeated in order to show the effects of the network cycle.

\subsubsection*{Network without Cycles}
\Cref{fig:example-network-wo} shows the optimized three-node network with corresponding numerical values.
\begin{figure}[h!]
    \centering
    \includegraphics[width=.5\linewidth]{example-without-cycles/network}
    \caption{Example network with nodal pricing scheme. Dispatch revenue as well as demand cost per MWh are given by the \ac{LMP} $\lambda$. Congestion revenue per MWh is given by the price difference $\lambda^\text{diff}$ of the connecting nodes.}
    \label{fig:example-network-wo}
\end{figure}
Both, bus 1 and 2 have a fixed demand of 30~MW and 50~MW, respectively. Bus~1 has a generator with an operational price of 6~\euro/MWh. Bus~3 has a cheaper generator with an operational price of 4~\euro/MWh. The maximum capacity of both generators is 50~MW. The transmission line from 1 to 2 has a capacity of 60~MW and line from 3 to 1 a capacity of 30 MW. Since the latter limits the use of the cheaper generator, there is a price difference of 2 \euro/MWh between bus 1 and 3. Due to the absence of network cycles, this price difference defines the congestion revenue of line 1--3. The flow on line 1--2 is not bound, hence there is no congestion revenue and bus 2 and 1 have the same electricity price of 6~\euro/MWh.

Since only net injections are allocated to the different buses, all demand at bus~1 is met by the local generator. Bus~2, on the other hand, imports electricity from bus~1 and bus~3. Using the Price Tracing method, the cost allocations $\mathcal{C}_{n\rightarrow s}$ and $\mathcal{C}_{n\rightarrow \ell}$ are calculated and shown in \cref{fig:example-network-payoff-wo}. Consumers at bus~1 only have to compensate the local generator with 180 \euro. Consumers at bus~2 pay for congestion revenues from line 1--3, all electricity generation at bus~3, and the remaining dispatch revenues from the generator at bus~1.

\begin{figure}[h]
    \centering
    \includegraphics[width=.5\linewidth]{example-without-cycles/payoff}
    \caption{Allocation matrices $\mathcal{C}_{n\rightarrow s}$ (left) and $\mathcal{C}_{n\rightarrow \ell}$ (right) for the dispatch and congestion revenue for the network in \cref{fig:example-network-wo}.}
    \label{fig:example-network-payoff-wo}
\end{figure}

The consistency of the cost allocation can be easily checked. The sum of a column gives the payment to a generator or a line. This corresponds exactly to the dispatch and congestion revenue defined in \cref{eq:dispatch_revenue,eq:congestion_revenue}. In turn, the sum of a row gives the total amount consumers have to pay at a bus. This corresponds to the demand cost $\lambda_n \, d_n$, \eg the sum of payments to bus~2 is 300~\euro, equal to the price of 6~\euro/MWh times the 50~MWh consumption. \\

\subsubsection*{Network with Cycle}

Now, we add a line from bus~3 to bus~2 with a capacity of 30~MW. This introduces \ac{KVL} constraints for the three lines that the sum of flows following the cycle is zero\footnote{For simplicity, we assume a uniform impedance in the example.}. As shown in \cref{fig:example-network} the new line's capacity is bound at 30~MW. The new cost optimum results in a higher price at bus~2 of 8~\euro/MWh despite the cheaper generator~2 producing 10~MW more than before. This result may be counterintuitive, but it becomes clear when considering that the low capacity of the new line, combined with the \ac{KVL}, restricts flows on line 1--3 and 1--2.

\begin{figure}[h!]
    \centering
    \includegraphics[width=.5\linewidth]{example-with-cycles/network}
    \caption{Example network with nodal pricing scheme and a network cycle. Now, due to the \ac{KVL} a new constraint adds to the optimization problem. Optionally, the congestion revenue per MWh may include shadow prices of the \ac{KVL} constraint $\lmpkvl$, thus $\lmp[\ell] = \lmpdiff + \lmpkvl$.}
    \label{fig:example-network}
\end{figure}

\begin{figure}[h!]
    \begin{subfigure}[c]{.5\linewidth}
        \includegraphics[width=\linewidth]{example-with-cycles/payoff}
        \subcaption{Without \ac{KVL} shadow prices}
        \label{fig:example-network-payoff}
    \end{subfigure}
    \begin{subfigure}[c]{.5\linewidth}
        \includegraphics[width=\linewidth]{example-with-cycles/payoff-kvl}
        \subcaption{With \ac{KVL} shadow prices}
        \label{fig:example-network-payoff-kvl}
    \end{subfigure}
    \caption{Cost allocations $\allocategeneratorcost$ and $\allocatelinecost$ for the dispatch and congestion revenue in the network in \cref{fig:example-network}.}
    \label{fig:example-network-payoff-all}
\end{figure}

The prices $\lambda_\ell^\text{KVL}$ of line $\ell$ are returned as the shadow price of the \ac{KVL} constraints in the cost-optimization, see \cref{sec:problem_formulation} for details. The Price Tracing method optionally allows these to be included in the congestion revenue without losing consistency with the demand cost.

\Cref{fig:example-network-payoff} shows the cost allocation without considering the \ac{KVL} shadow prices. Here, the congestion revenue is proportional to the price difference of the connected buses only. The new line 2--3 has the highest revenue at 120~\euro/MWh. However, \cref{fig:example-network-payoff-kvl} shows the cost allocation taking into account the \ac{KVL} shadow prices. Here, the congestion revenues are shifted towards 1--2 and 1--3. Note that this shift accounts for the fact that it is the new line 2--3 that limits the flow on line 1--2 and 1--3.

Again, both cost allocation schemes are consistent with the total demand cost, dispatch and congestion revenues. In the following, the \ac{KVL} shadow prices in the Price Tracing method are considered since then, the total congestion revenue of a line equals the sum of all associated line costs in a short and long-term equilibrium, see \cref{sec:zero_profit_flow}.

\begin{figure*}[h!]
    \centering
    \includegraphics[width=\linewidth]{de50/network}
    \caption{Model of the German power system. The left-hand side shows the minimum requirements for renewable and transmission capacities. The right-hand side shows the cost-optimal capacities of all generators as well as transmission expansions. The effective CO$_2$ price is set to 120 \euro\, per ton of CO$_2$ emission.}
    \label{fig:network}
\end{figure*}

\section{Application Cases}
\label{sec:application_case}

Using a cost-optimized model of the German power grid with 50 representative nodes and a time span of one year with hourly resolution, four use cases of the Price Tracing method are demonstrated.

The investment model builds on the PyPSA-EUR workflow \cite{horsch_jonas_pypsa-eur_2020} whose technical details and assumptions are presented in \cite{horsch_pypsa-eur_2018}. Transmission line capacities can be expanded but require a minimum of today's capacities, originally retrieved from the ENTSO-E Transmission System Map \cite{entso-e_entso-e_nodate}. Further, a minimum deployment of renewable capacity is required, derived from matching all entries of the OPSD renewable power plant list \cite{schlechtRenewablePowerPlants2020} to its closest bus. Wind and solar capacity expansion are limited by land-use restrictions considering agriculture, urban, forested and protected areas based on the CORINE and NATURA2000 database \cite{eea_corine_2012,eea_natura_2016}. \ac{PHS} and \ac{ROR} power plants are fixed to today's capacities and may not be expanded. In addition, unlimited expansion of batteries and H$_{2}$-storages and \ac{OCGT} are allowed at each node.
% 
% 
\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{de50/average_price}
    \caption{Load-weighted average electricity price $\averagelmp$ per region in the optimized German power system. Regions in the middle and south of Germany have high prices whereas electricity in the North with a strong wind, transmission and \ac{OCGT} infrastructure is cheaper.}
    \label{fig:average_price}
\end{figure}
We assume a carbon price of 120 \euro\, per ton of CO$_{2}$, which adds an effective price of 55 \euro/\megawatthour\, to the operational price of \acp{OCGT} (assuming a gross emission of 180~kg/MWh and an efficiency of 39\%). All cost assumptions on operational costs and annualized capital costs are summarized in \cref{tab:cost_assumptions}.

The network is shown in  \cref{fig:network}. The left-hand side shows all the minimum capacity requirements, the right-hand side the cost-optimal generation capacities as well as net transmission expansions. Solar capacities are expanded in the south, onshore and offshore wind in the upper north and far west. \ac{OCGT} are built in the middle and north of the country. Transmission lines are amplified along the north-south axis, including a major \ac{DC} link, associated with the German S\"ud-Link, running from the coastal region to the southwest.
The total annualized cost of the power system is roughly 46 billion \euro.
%

\Cref{fig:average_price} displays the load-weighted average electricity price $\averagelmp$ per region, defined by
\begin{align}
    \averagelmp = \dfrac{\sum_{t} \lmp \demand}{\sum_t \demand} = \dfrac{\sum_{t} \payment }{\sum_t \demand}
\end{align}
It reveals a relatively strong gradient from the south (at roughly 98 \euro/MWh) to the north (84 \euro/MWh). Regions with relatively poor renewable resources and therefore little capacity expansion tend to have higher prices.
% The node with the lowest \ac{LMP} in the upper northwest, stands out through high offshore capacity requirements.
\\

\subsection{Usage-based Network Tariff}

The first possible use case of the Price Tracing method discussed here is a usage-based network tariff. As in other countries, electricity consumers in Germany pay a uniform network tariff, the ``Netzentgelte''. These are based on a relatively opaque calculation and valuation of all network costs for each Transmission System Operator, in addition to an individually derived profit margin.

The Price Tracing method allocates all congestion revenues to consumers in the network, given by $\allocatelinecost$. This cost allocation can be used to define a usage-based network tariffs $\averagelmp^\text{grid}$ for each bus-region $n$ in the network:
\begin{align}
    \averagelmp^\text{grid} = \dfrac{\sum_{t,\ell} \allocatelinecost}{\sum_t \demand}.
\end{align}
This is the average price that consumers at $n$ pay for the transmission of electricity. These include compensation for all capital investments taken by the network operators. Consumers with a high degree of nodal autarky, i.e. with a little net electricity import, pay a lower network tariff and vice versa. This gives an incentive to regions with a sub-optimal deployment of renewable technologies to invest in local infrastructure. Note since the cost-optimum represents a long-term equilibrium, the total revenue per network asset equals the total \ac{OPEX} and \ac{CAPEX}. The network tariffs $\averagelmp^\text{grid}$ reflect this equilibrium and therefore do not result in any profits for the grid operator.

A further advantage of this network tariff allocation lies in the possibility of differentiating between network operators. The quantity
\begin{align}
    \averagelmp[n\rightarrow \ell]^\text{grid} = \dfrac{\sum_{t} \allocatelinecost}{\sum_t \demand}
\end{align}
decomposes the network tariff into single lines, naturally fulfilling  $\averagelmp^\text{grid} = \sum_\ell \averagelmp[n\rightarrow \ell]^\text{grid}$. On the one hand this enables operator or country specific network tariffs to be quantified so they are consistent with cross-border flows. On the other hand this facilitates tracing back which congested lines are the strongest drivers to the local network tariff.

\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{de50/maps_price/branch_investment_cost.png}
    \caption{Usage-based network tariff per region. The resulting prices are indicated by the color of the region, the transmission lines are drawn in proportion to their congestion revenues.}
    \label{fig:congestion_revenue}
\end{figure}

\Cref{fig:congestion_revenue} shows the network tariffs $\averagelmp[n]^\text{grid}$ for all regions of the network model as well as the average congestion revenues per transmission line. Regions close to the shore hardly affect the transmission system, hence their network tariff is negligible. In contrast, regions in the middle and south are relying on power transfer from the north. Their demand is the main cause for the transmission system expansion. This translates to a higher network tariff.

\subsection{Emission Cost Allocation}
\label{sec:co2-cost-allocation}

\begin{figure}[h!]
    \centering
    \includegraphics[width=.5\linewidth]{de50/maps_price/co2_cost.png}
    \caption{Average CO$_2$ cost per consumed MWh. The effective prices are indicated by the color of the region, the circles are drawn in proportion to the revenue per regional generators.}
    \label{fig:opex_price}
\end{figure}

The mapping $\allocategeneratorcost$ allocates all dispatch revenues to consumers. Naturally, the revenues of conventional generators include costs accounting for CO$_2$ emissions. Using the mechanism of the Price Tracing method, these CO$_2$ costs may be mapped to consumers by weighting the dispatch allocation $\allocategeneration$ with the effective CO$_2$ price $\emissionprice$ per produced MWh at generator $s$. Applying an emission cost allocation in the form of
\begin{align}
    \allocateemissioncost = \emissionprice \allocategeneration
\end{align}
encourages consumers to reduce their emission-intensive power consumption and leads to a transparent polluter pay principle. The average emission cost per consumed MWh is then given by
\begin{align}
    \averagelmp^{\text{CO}_2} = \dfrac{\sum_{t,\ell} \allocateemissioncost}{\sum_t \demand}.
\end{align}
\Cref{fig:opex_price} depicts the price for the network model. Regions with higher \ac{OCGT} backup capacity tend to have higher allocated emission costs. The size of the circles is proportional to the emission cost occasioned by the local generators.
Note, since the CO$_2$ price is given exogenously, this tracing method may also be applied to networks without a nodal pricing scheme.

\subsection{A Transparent Nodal Pricing Scheme}

\begin{figure*}[h!]
    \includegraphics[width=\linewidth]{de50/allocated_payment_comparison}
    \caption{Comparison of region payments with the lowest \ac{LMP} (left) and the region with the highest \ac{LMP} (right). The region is colored in dark blue. The circles indicate to which bus and technology the payments are assigned. The thickness of the lines is proportional to dedicated payments.}
    \label{fig:direct-allocation}
\end{figure*}

Another application of the Price Tracing method is to make the nodal demand cost in a network more transparent. This answers the question of which parts of the system consumers' electricity payments go to. To demonstrate this, \cref{fig:direct-allocation} shows the total \ac{P2P} cost assignments $\sum_t \allocategeneratorcost$ and $\sum_t  \allocatelinecost$ for the region with the lowest average \ac{LMP} (left side) and for the region with the highest \ac{LMP} (right side). The dedicated region is indicated in blue. The circles and their segments indicate the allocated dispatch revenues which are subdivided by carrier type. The lines indicate the allocated congestion revenue.

The low-cost region in the Northwest has few congestion costs due to its low net imports. It profits from local offshore wind farms and only a small share of the payments is allocated to remote \acp{OCGT}. In contrast, the high-price region is highly dependent on local \ac{OCGT} and the transmission system, which causes high amounts of allocated dispatch and congestion costs. Interestingly, its payments to  onshore and offshore wind infrastructure are low despite a third of its supply coming from wind power, compare \cref{fig:power_mix}. This suggests that wind power supply in this region is constrained not by depleted wind power resources, but by bottlenecks in the transmission grid.

% Finally, \cref{fig:locality} draws the cumulative share of P2P cost assignments of a function of the distance between producer and consumer. The data is shown for all technologies separately. The later the curve reaches 100\% the deeper the price of a technologies penetrates into the network. Offshore wind has the strongest price influence to remote buses. Only 10-30\% of its expenses are compensated by local consumers and the rest is assigned to remote buses, especially those with high demand (compare \cref{fig:average_demand}). The contrast between Hydrogen Storage and Battery sticks out. Whereas Battery costs are hardly assigned to other buses, almost 50\% of the Hydrogen Storage costs are payed by remote buses. It underlines the fundamental functionality of the two technologies. Hydrogen storage are located at buses with high wind generation and balance out their long-term excess and deficit energy. The dispatched power follows a similar way through the network as the wind power. The battery on the other hand pairs with local solar production and its locally constraint dispatch and flow.

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{de50bf/locality_all_ptpf_net}
%     \caption{Average distance between payer and receiver for different technologies and shares of the total production.}
%     \label{fig:locality}
% \end{figure}

\subsection{Revenue Decomposition}

Finally, a more abstract use case of Price Tracing is demonstrated that focuses on the decomposition of congestion and dispatch revenues for each asset. In a long-term equilibrium without any exogenous constraints, a network asset recovers all its \ac{OPEX} and \ac{CAPEX} from the revenue. This relation is known as the zero-profit rule and is an outcome of the \ac{KKT} conditions of the underlying cost-minimization problem, derived in detail in \cref{sec:zero_profit_generation,sec:zero_profit_flow,sec:zero_profit_storage_units}. However, if the cost-optimum is subject to additional constraints, for example by a minimum share of renewable capacity, the zero-profit relation is altered and more cost contributions have to be considered \cite{brown_decreasing_2020} as soon as the new constraint becomes binding. With the Price Tracing method it is possible to decompose the effect of the additional boundary conditions to the cost of individual consumers.

\begin{table}[h!]
    \caption{Exemplary contributions to the revenue of network assets highlighted in this work. Depending on the formulation of the optimization problem, the list changes and may possibly include other terms.}
    \begin{center}
        \begin{tabular}{c|c p{0.3\textwidth}}
            \toprule
            \text{Contribution}  & Symbol          \\
            \midrule
            \text{OPEX}          & $\opex$         \\
            % \midrule
            \text{CAPEX}         & $\capex$        \\
            % \midrule
            \text{Emission Tax}  & $\emissioncost$ \\
            % \midrule 
            \text{Scarcity Rent} & $\scarcitycost$ \\
            \text{Subsidies}     & $\subsidycost$  \\
            \vdots               & \vdots          \\
            \bottomrule
        \end{tabular}
    \end{center}
    \label{tab:contributions}
\end{table}

\noindent
\Cref{tab:contributions} shows a possible set of contributions to the asset revenue. On top of \ac{OPEX} and \ac{CAPEX}, there are possible emission costs which were previously discussed in \cref{sec:co2-cost-allocation}. In presence of capacity expansion limits, additional Scarcity Rent $\scarcitycost$ must be included in the revenues, in case the corresponding constraints are binding, for mathematical details see \cref{sec:upper_capacity_limits}. This rent translates to a compensation for higher investments in a highly competitive market, but can also be ignored in case an asset is fixed to its existent, amortized capacity. Likewise, assets may be constrained to lower capacity requirements which reflect \ie  existing infrastructure, called Brownfield constraints, see \cref{sec:lower_capacity_limits}. If binding, the lower limit reduces the revenue of the corresponding asset and leaves a Subsidy Cost $\subsidycost$ which have to be compensated by external institutions (government, community) or are simply ignored when the asset is already amortized.

Except for the subsidies, all contributions can be expressed in terms of the operation of an asset, that is the generation $\generation$ and the flow $\flow$. Thus, given the set of discussed revenue contributions, the dispatch revenue allocation decomposes into
\begin{align}
    \allocategeneratorcost = \allocateopex + \allocatecapex + \allocateemissioncost + \allocatescarcitycost - \subsidycost_s .
\end{align}
where all expressions are summarized in \cref{tab:cost_allocation_map}. \Cref{sec:application_case_appendix} shows the individual cost allocations for the network model. Note that transmission lines are assumed to have neither \ac{OPEX} nor emission tax, \ie
\begin{align}
    \allocatelinecost = \capex_{n \ra \ell, t} + \scarcitycost_{n \ra \ell, t} - \subsidycost_\ell
\end{align}
This approach allows modelers and system operators to disentangle the combined effect of multiple constraints on the dispatch and congestion revenues. For example, when analyzing these cost allocations for the presented German network model, it becomes clear that buses, which are far from wind farms, pay higher prices due to increased reliance on transmission and backup capacity. On the other hand, buses with high renewable installation spend most payments to local assets, see \cref{sec:application_case_appendix} for details.

\section{Limitations}
\label{sec:limitations}

The method presented is closely linked to the underlying power system model and its optimization. In the usual electricity market mechanism, electricity producers bid on the market and Transmission System Operators make profits by charging grid fees. On the other hand, electricity consumers always pay indirectly for electricity generation costs, grid usage fees, and carbon emission costs.
While this study is less concerned with today's market structures, it is more concerned with showing what other market structures are theoretically possible and what flexibilities result from them. Nodal price structures with optimized prices hold many possibilities to create incentives. The preceding explanation is intended to help illuminate this space of possibilities.

The presented cost allocation is based on the linear power flow approximation. Yet, the method is equally applicable to a system with an optimal power flow, \ie an \ac{AC} power flow. However, in this case, the allocations $\allocategeneration$ and $\allocateflow$ should be computed with the Z-Bus flow allocation presented in \cite{conejo_z-bus_2007} which, by design, respects both Kirchhoff Circuit Laws. Here, an additional cost term accounting for the transmission loss has to be considered in the decomposition of the congestion revenue.

The used optimization does not take security constraints of the transmission system into account. These may be incorporated following the approach in \cite{nikoukar_transmission_2012}.

The method can be equally applied to short and long-term planning network models. However, the paper does not take the detailed structure of today's power markets into account. These vary from country to country and often reveal interlocked mechanisms (energy only markets, reserve markets, redispatch etc.). The aim of the presented work is to reconsider some of the mechanisms and  to reevaluate their market efficiencies.

\section{Conclusion}
\label{sec:conclusions}

A new method, called Price Tracing, was presented. It builds upon the nodal pricing scheme and decomposes the dispatch and congestion revenues of generators and transmission lines into individual payments of consumers. It is based on the Average Participation method which calculates assignments between generators and consumers using the principle of proportional sharing. These assignments are weighted by the market price to determine the allocation of the revenues to consumers. By means of a numerical example, it was shown that there exist two options of how to treat network cycles regarding cost allocation and laid out why it is more sensible to consider shadow prices of the Kirchhoff Voltage constraint.

Using an optimized model of the German power system with a price of 120~\euro\, per ton of CO$_2$, four use cases of the Price Tracing method were presented. First, it was demonstrated how grid tariffs can be adjusted to a transparent, consumption-based tariff that incentivizes investment in local infrastructure. Second, Price Tracing allows the grid to track the costs associated with CO$_2$ emissions. These results can be used to reduce CO$_2$-intensive demand on a grid. The third use case presented uses Price Tracing to derive cost contributions for individual consumers to make nodal payments transparent and explain price differences. This relates to the fourth case where individual dispatch and congestion revenues are decomposed into contributions from operational expenditures, capital expenditures, scarcity rents, and subsidies. It was shown that the allocation can be performed for a single cost contribution to provide insights into electricity system models or to derive policy advice.

\subsubsection*{Reproducibility and Expansion}

All figures and data points can be reproduced by using the \textit{python} workflow in \cite{hofmann_pypsa-costallocation_2020}. The automated workflow allows for higher spatial resolution of the network (scalable up to a the full ENTSO-E Transmission System Map) and optionally taking the total European power system into account.

\subsection*{Funding}
This research was funded by the by the German Federal Ministry for Economics Affairs and Energy (BMWi) in
the frame of the NetAllok project (grant number 03ET4046A) \cite{bundesministeriumfurwirtschaftundenergieVerbundvorhabenNETALLOKMethoden}.

\subsubsection*{Acknowledgement}

In particular, we thank Tom Brown for very fruitful discussions. We are very grateful to Alexander Zerrahn for reviewing. Further, we thank Alexander Kies and Horst Stöcker for their creative and fruitful inputs.

% ----------------------- APPENDIX -------------------------------------

\clearpage
\appendix
\subsection*{Acronyms}
\begin{acronym}[UMLX]
    \acro{AC}{Alternating Current}
    \acro{AP}{Average Participation}
    \acro{CAPEX}{Capital Expenditures}
    \acro{CO2}[CO$_2$]{Carbon Dioxide}
    \acro{DC}{Direct Current}
    \acro{FA}{Flow Allocation}
    \acro{HVAC}{High-Voltage Alternating Current}
    \acro{HVDC}{High-Voltage Direct Current}
    \acro{Hydro}{Hydro-Electric Dams}
    \acro{Hydrogen}[H$_2$-storage]{Hydrogen Storage}
    \acro{KCL}{Kirchhoff's Current Law}
    \acro{KKT}{Karush–Kuhn–Tucker}
    \acro{KVL}{Kirchhoff's Voltage Law}
    \acro{LCOE}{Levelized Cost of Electricity}
    \acro{LMP}{Locational Marginal Price}
    \acro{LOPF}{Linear Optimal Power Flow}
    \acro{OCGT}{Open-Cycle Gas Turbine}
    \acro{OPEX}{Operational Expenditures}
    \acro{P2P}{Peer-to-Peer}
    \acro{PHS}{Pumped-Hydro-Storage}
    \acro{PTDF}{Power Transfer Distribution Factors}
    \acro{PV}{Photovoltaic}
    \acro{PyPSA-EUR}{PyPSA-Europe}
    \acro{PyPSA-Eur}{\acs*{PyPSA} Europe}
    \acro{PyPSA}{Python for Power System Analysis}
    \acro{ROR}{Run-of-River}
    \acro{SOAF}[SO\&AF]{\acs*{ENTSO-E} Scenario Outlook and Adequacy Forecast}
    \acro{TSO}{Transmission System Operator}
    \acrodefplural{TSO}{Transmission System Operators}
\end{acronym}

\renewcommand\theequation{\thesection.\arabic{equation}}
\setcounter{equation}{0}

\renewcommand\thefigure{\thesection.\arabic{figure}}
\setcounter{figure}{0}

\section{Network Optimization}

\subsection{Optimality Conditions}
\label{sec:optimality_conditions}

Consider the following linear minimization problem

\begin{align}
    \min_{x_n} \sum_n c_n \, x_n
\end{align}
subject to inequality constraints
\begin{align}
    g_i(x_n) \le 0 \resultsin{\mu_i}
\end{align}
and equality constraints
\begin{align}
    h_j(x_n) = 0 \resultsin{\lambda_j}
\end{align}
where $\mu_i$ and $\lambda_j$ are the corresponding dual variables, also known as \ac{KKT} variables.
The Lagrangian is given by
\begin{align}
    \lagrangian = \sum_n c_n x_n + \sum_i \mu_i g_i(x_n) + \sum_j \lambda_j h_j(x_n).
\end{align}

At the optimum $x^*_n$ the following \ac{KKT} conditions are satisfied.
\begin{enumerate}
    \item Stationarity \\
          \begin{align}
              \pdv{\lagrangian}{x_n} = c_n + \sum_i \mu_i \pdv{g_i}{x_n} + \sum_j \lambda_j \pdv{h_j}{x_n} & = 0
          \end{align}
    \item Primal Feasibility \\
          \begin{align}
              g_i(x_n) \le 0 \qquad \forall i \\
              h_j(x_n) \le 0 \qquad \forall j
          \end{align}
    \item Dual Feasibility \\
          \begin{align}
              \mu_i \ge 0 \qquad \forall i
          \end{align}
    \item Complementary Slackness \\
          \begin{align}
              \sum_i \mu_i g_i(x_n) = 0
          \end{align}
\end{enumerate}

\subsection{Problem formulation}
\label{sec:problem_formulation}

We follow a cost-minimization approach, which optimizes all operational and capital expenditures of the network. The corresponding objective is given by

\begin{align}
    \min_{g_{s,t}, G_s, F_{\ell, t}} \sum_{s,t} o_s g_{s,t} + \sum_s c_s G_s + \sum_\ell c_\ell F_\ell
\end{align}
where $G_s$ and $F_\ell$ are the nominal capacities per generator $s$ and line $\ell$ respectively. The operational costs of generators are given by $o_s$, the capital costs for generators and lines by $c_s$ and $c_\ell$.

Of particular importance is the nodal balance constraint which ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the \ac{KCL}. With a given demand $\demand$ this translates to
\begin{align}
    \nodalgeneration - \sum_\ell \incidence \flow & =  \demand
    \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
where $\incidence$ is +1 if line $\ell$ starts at bus $n$, -1 if it ends at $n$, 0 otherwise. The nodal generation $\nodalgeneration$ collects the production of all nodal assets. The corresponding dual variable (or shadow price) $\lmp$ is the \ac{LMP} per bus and time step. In an economic equilibrium for a nodal pricing scheme this is the \euro/\megawatthour\, price which a consumer has to pay.\\

In order to impose the \ac{KVL} for the linearized \ac{AC} flow, the constraint
\begin{align}
    \sum_{\ell} \cycle \, \reactance \, \flow = 0 \resultsin{\cycleprice} \Forall{c,t}
\end{align}
is added to the problem. Here, $\reactance$ denotes the line's impedance and $\cycle$ is +1 if $\ell$ lays aligned to the network cycle $c$, --1 if it lays in the opposite direction and zero if it is not part of the network cycle $c$.
For simplicity we define the total price of the \ac{KVL} per line as
\begin{align}
    \lmpkvl = \sum_c \cycleprice\, \cycle\, \reactance \Forall{\ell,t} .
\end{align}
The total price per transmitted unit of power can now be defined by the price differences of the starting and ending bus of the transmission line
\begin{align}
    \lmp[\ell] = \lmpdiff
\end{align}
with
\begin{align}
    \lmpdiff = -\sum_n \incidence \lmp .
\end{align}
or it additionally includes the price for the \ac{KVL}
\begin{align}
    \lmp[\ell] = \lmpkvl + \lmpdiff .
    \label{eq:congestion_price}
\end{align}

\subsection{Zero Profit Generation}
\label{sec:zero_profit_generation}
% Let $S \subseteq I$ be the set of generators in the network, such that $\generation = s_{s,t}$ describes the power production of generator $s \in S$. The OPEX occasioned by generator $s$ is given by a cost-weighted sum of the production, thus 
% \begin{align}
%     \opexgeneration_s = \sum_t \operationalpricegeneration \, \generation 
%     \label{eq:opexgeneration}
% \end{align}
% In case a fix price for emissions $\emissionprice$ in \euro\, per tonne-CO$_2$ equivalents, is assumed, a further the cost term per generator $s$,   
% \begin{align}
%  \emissioncost_s = \emissionprice \, \sum_t  \emission \, \generation
% \end{align}
% adds to $\totalcost$. Here, $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour\, of generator $s$. In contrast to OPEX and emission costs, the CAPEX of $s$ are not a function of the production $\generation$, but of the actual installed capacity $\capacitygeneration$ of generator $s$. In the optimization it limits the generation $\generation$ in the form of 
% \begin{align}
% \generation - \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
% \label[constraint]{eq:upper_generation_capacity_constraint}
% \end{align}
% The constraint yields a shadow-price of $\muuppergeneration$, given by corresponding the Karush–Kuhn–Tucker (KKT) variable, in literature often denoted as the Quality of Supply \cite{schweppe_spot_1988}. It can be interpreted as the price per MW that \cref{eq:upper_generation_capacity_constraint} imposes to the system. If $\muuppergeneration$  is bigger than zero, the constraint is binding, which pushes investments in $\capacitygeneration$. As shown in \cite{brown_decreasing_2020} and in detail in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is recovered by the production $\generation$ times the shadow price $\muuppergeneration$, 
% \begin{align}
%  \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
%  \label{eq:no_profit_capex_generation}
% \end{align}
% This representation connects the CAPEX with the operational state of generator $s$, \ie matches the form in \cref{eq:cost_decomposition}.

For each generator $s$ the optimization defines a lower and upper limit for the power output $\generation$, given by
\begin{align}
    - \generation                     & \le 0 \resultsin{\mulowergeneration} \Forall{s,t}
    \label[constraint]{eq:lower_generation_capacity_constraint}                           \\
    \generation - \capacitygeneration & \le 0 \resultsin{\muuppergeneration} \Forall{s,t}
    \label[constraint]{eq:upper_generation_capacity_constraint}
\end{align}
\Cref{eq:upper_generation_capacity_constraint,eq:lower_generation_capacity_constraint}, which yield the \ac{KKT} variables $\muuppergeneration$ and $\mulowergeneration$, imply the complementary slackness,
\begin{align}
    \muuppergeneration \left( \generation - \generationpotential \, \capacitygeneration \right) & = 0  \Forall{s,t}
    \label{eq:complementary_slackness_upper_generation}                                                             \\
    \mulowergeneration  \, \generation                                                          & = 0 \Forall{s,t}
    \label{eq:complementary_slackness_lower_generation}
\end{align}

The stationarity of the generation capacity variable leads to
\begin{align}
    \pdv{\lagrangian}{\capacitygeneration}  = 0 \,\, \rightarrow \,\,
    \capitalpricegeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{s}
    \label{eq:capex_generation_duality}
\end{align}
and the stationarity of the generation to
\begin{align}
    \pdv{\lagrangian}{\generation} & = 0 \,\, \rightarrow \,\,
    \operationalpricegeneration =  \sum_n \incidencegenerator \, \lmp - \muuppergeneration + \mulowergeneration \Forall{s} \label{eq:opex_duality}
\end{align}
where $\incidencegenerator$ is set to one if generator $s$ is placed at node $n$ and zero otherwise.

Multiplying both sides of \cref{eq:capex_generation_duality} with $\capacitygeneration$ and using \cref{eq:complementary_slackness_upper_generation} leads to
\begin{align}
    \capitalpricegeneration \, \capacitygeneration  = \sum_t \muuppergeneration \, \generation \Forall{s}
    \label{eq:capital_price_generation_sum}
\end{align}
The zero-profit rule for generators is obtained by multiplying \cref{eq:opex_duality} with $\generation$ and using \cref{eq:complementary_slackness_lower_generation,eq:capital_price_generation_sum} which results in
\begin{align}
    \capitalpricegeneration \, \capacitygeneration + \sum_t \operationalpricegeneration \generation = \sum_{n,t} \lmp \incidencegenerator \generation \Forall{s}
\end{align}
It states that over the whole time span, all \ac{OPEX} and \ac{CAPEX} for generator $s$ (left-hand side) are payed back by its revenue (right-hand side).

\subsection{Zero Profit Transmission System}
\label{sec:zero_profit_flow}

% Let $L \subset I$ be the set of transmission lines in the system, these may include Alternating Current (AC) as well as Directed Current (DC) lines. Further let $\flow = s_{\ell,t}$ represent the power flow on line $\ell \in L$.  If the OPEX of the transmission system is taken into account in $\totalcost$ (these are often neglected in power system models), these may be approximated by $\opexflow_\ell = \sum_t o_\ell |\flow|$, that is, a cost weighted sum of the net flow on line $\ell$. Again this stands in contrast to the CAPEX which not a function of $\flow$ but of the transmission capacity $\capacityflow$. It limits the flow $\flow$ in both directions,
% \begin{align}
% \flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
% \label[constraint]{eq:upper_flow_capacity_constraint} \\
% - \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
% \label[constraint]{eq:lower_flow_capacity_constraint}
% \end{align}
% At the cost-optimum, the two constraints yield the shadow prices $\muupperflow$ and $\mulowerflow$. Again, we use the relation that over the whole time span, the shadow prices weighted by the flow match the investment in line $\ell$ (for details see \cref{sec:zero_profit_flow}) 
% \begin{align}
% \capexflow_\ell = \capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
% \label{eq:no_profit_capex_flow}
% \end{align}
% The shadow prices $\muupperflow$ and $\mulowerflow$ can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$. Hence, a non-zero values indicate that \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint} are bound and therefore that the congestion on line $\ell$ at time $t$ is imposing costs to the system.

For transmission lines the flow $\flow$ is limited by the nominal capacity $\capacityflow$ in both directions, mathematically translating to
\begin{align}
    \flow - \capacityflow   & \le 0 \resultsin{\muupperflow} \Forall{\ell,t}
    \label[constraint]{eq:upper_flow_capacity_constraint}                      \\
    - \flow - \capacityflow & \le 0 \resultsin{\mulowerflow} \Forall{\ell,t} .
    \label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
The \ac{KKT} variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the transmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. For flows below the thermal limit, the complementary slackness
\begin{align}
    \muupperflow \left( \flow - \capacityflow \right) & = 0 \Forall{\ell,t}
    \label{eq:complementary_slackness_upper_flow}                            \\
    \mulowerflow \left( \flow - \capacityflow \right) & =  0 \Forall{\ell,t}
    \label{eq:complementary_slackness_lower_flow}
\end{align}
sets the respective \ac{KKT} variables to zero.

The stationarity of the transmission capacity leads to
\begin{align}
    \pdv{\lagrangian}{\capacityflow} = 0 \,\, \rightarrow \,\,
    \capitalpriceflow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
    \label{eq:capacity_flow_duality}
\end{align}
and the stationarity with respect to the flow to
\begin{align}
    0 & = \pdv{\lagrangian}{\flow}                                                                                                                 \\
    0 & = - \sum_n \incidence \lmp  + \sum_c \cycleprice \cycle \reactance  - \muupperflow + \mulowerflow \Forall{\ell, t} \label{eq:flow_duality}
\end{align}

When multiplying \cref{eq:capacity_flow_duality} with $\capacityflow$ and using the complementary slackness \cref{eq:complementary_slackness_upper_flow,eq:complementary_slackness_lower_flow} we obtain
\begin{align}
    \capitalpriceflow \, \capacityflow = \sum_t \left( \muupperflow - \mulowerflow \right)  \, \flow \Forall{\ell}
    \label{eq:capital_price_flow_sum}
\end{align}
Again we can use this to formulate the zero-profit rule for transmission lines. We multiply \cref{eq:flow_duality} with $\flow$, which finally leads us to
\begin{align}
    \capitalpriceflow \, \capacityflow & = - \sum_n \incidence\, \lmp\, \flow + \sum_c \cycleprice\, \cycle\, \reactance\, \flow
    \Forall{\ell}                                                                                                                \\
                                       & = \sum_t \lmp[\ell] \flow \Forall{\ell}
\end{align}
% Watch out Incidence = - ordinary Incidence !!
where $\lmp[\ell]$ is given by \cref{eq:congestion_price}.
It states that the congestion revenue of a line (first term right-hand side) reduced by the cost for cycle constraint exactly matches its \ac{CAPEX}.

\subsection{Zero Profit Storage Units}
\label{sec:zero_profit_storage_units}

% In a simplified storage model, a storage unit $r$ has an associated nominal capacity $\capacitystorage$ that limits the storage dispatch $\storagedispatch$ and charging $\storagecharge$. Further it limits the maximal storage capacity $\storagesoc$ by a fix ratio $h_r$, denoting the maximum hours at full discharge. The storage $r$ dispatches power with efficiency $\efficiencydispatch$, charges power with efficiency $\efficiencycharge$ and preserves power from one time step $t$ to the next, $t+1$, with an efficiency of $\efficiencysoc$. In \cref{sec:zero_profit_storage_units} we formulate the mathematical details. 
% Using the result of \cite{brown_decreasing_2020} the CAPEX can be related to the operation of a storage unit $r$ through
% \begin{align}
%     \notag
%     \capexstorage =& \capitalpricestorage \, \capacitystorage \\
%     \notag
%     =& \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
%     &- \sum_t \lmp \incidencestorage  \storagecharge \Forall{r} 
%     \label{eq:no_profit_capex_storage}
% \end{align}
% where $\muupperstoragedispatch$ and $\mulowerstoragedispatch$ are the shadow prices of the upper and lower dispatch capacity bound and $\mustateofcharge$ is the shadow price of the energy balance constraint. Following the considerations in \cref{sec:zero_profit_storage_units} we restrict to the revenue from dispatched power, \ie the first term in \cref{eq:no_profit_capex_storage}, for the cost allocation.

For an simplified storage model, the upper capacity $\capacitystorage$ limits the discharging dispatch $\storagedispatch$, the storing power $\storagecharge$ and state of charge $\storagesoc$ of a storage unit $r$ by
\begin{align}
    \storagedispatch - \capacitystorage   & \le 0  \resultsin{\muupperstoragedispatch} \Forall{r,t} \\
    \storagecharge - \capacitystorage     & \le 0  \resultsin{\muupperstoragecharge} \Forall{r,t}   \\
    \storagesoc - h_r \, \capacitystorage & \le 0  \resultsin{\muupperstoragesoc} \Forall{r,t}
\end{align}
where we assume a fixed ratio between dispatch and storage capacity of $h_r$.
% From stationarity we obtain 
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\capacitystorage} &= 0 \\
%     \capitalpricestorage &= \sum_t \left( \muupperstoragedispatch + \muupperstoragecharge + h_r \muupperstoragesoc  \right)
% \end{align}
The state of charge must be consistent throughout every time step according to what is dispatched and stored,
\begin{align}
    \notag
    \storagesoc - \efficiencysoc \storageprevioussoc - \efficiencycharge \storagecharge & + (\efficiencydispatch)^{-1} \storagedispatch = 0 \\
                                                                                        & \resultsin{\mustateofcharge} \Forall{r,t}
\end{align}

We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus
\begin{align}
    \notag
    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage & \left(\storagedispatch - \storagecharge \right) \Forall{r,t}
\end{align}
where $\incidencestorage$ is one if storage $r$ is paced at node $n$ and zero otherwise.
The stationarity of the dispatched power leads us to
\begin{align}
    \notag
    \pdv{\lagrangian}{\storagedispatch}                                                                                                                            & = 0                   \\
    \operationalpricestorage - \sum_n \lmp \, \incidencestorage - \mulowerstoragedispatch + \muupperstoragedispatch + (\efficiencydispatch )^{-1} \mustateofcharge & = 0 \;  \forall\; r,t
    \label{eq:stationarity_storagedispatch}
\end{align}
which we  can use to define the revenue which recovers the \ac{CAPEX} at $r$,
\begin{align}
    \notag
    \capitalpricestorage \, \capacitystorage = \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch \\
    - \sum_t \lmp \incidencestorage  \storagecharge \Forall{r}
    \label{eq:no_profit_capex_storage2}
\end{align}

% With this constraint we can derive the following stationarities:
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagedispatch} &= 0 \\
%     \operationalpricestorage + \lmp + \mulowerstoragedispatch - \muupperstoragedispatch - (\efficiencydispatch )^{-1} \mustateofcharge &= 0  
%     \label{eq:stationarity_storagedispatch}
% \end{align}
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagecharge} &= 0 \\
%     - \lmp + \mulowerstoragecharge - \muupperstoragecharge + \efficiencycharge \mustateofcharge &= 0 
%     \label{eq:stationarity_storagecharge}
% \end{align}
% \begin{align}
%     \notag
%     \pdv{\lagrangian}{\storagesoc} &= 0 \\
%     \mulowerstoragesoc - \muupperstoragesoc - \mustateofcharge + \efficiencysoc \munextstateofcharge &= 0 
%     \label{eq:stationarity_storagesoc}
% \end{align}

It stands to reason to assume that when a storage charges power, it does not supply any demand. Thus consumers only pay storage units in times the storage dispatches power. Hence, we restrict the allocatable revenue per storage unit to the first term in \cref{eq:no_profit_capex_storage2}. This allocates the CAPEX of $r$ plus the costs $\remainingcost^E_r$ it needs to buy the charging power,
\begin{align}
    \capexstorage_r + \remainingcost^E_r = \sum_t \left(\muupperstoragedispatch - \mulowerstoragedispatch  + (\efficiencydispatch )^{-1} \mustateofcharge \right) \storagedispatch
\end{align}
In charging times the total of remaining costs $\remainingcost^E_r$ is spent to power from other assets. These costs scale with the amount of installed storage capacity.
% Note that it would be possible to incorporate this redistribution into the cost allocation, by replacing the demand $\demand$ with the power charge $\storagecharge$ in \cref{eq:general_scheme}. Then, the derived payments that a storage unit $r$ has to pay to asset $i$ is given by $\cost_{r \rightarrow i}$. The sum of those payments due to $r$ will the sum up to $\remainingcost^E_r$.

\subsection{Upper Capacity Expansion Limits}
\label{sec:upper_capacity_limits}

\begin{subequations}
    In real-world setups the capacity expansion of generators, lines or other assets are often limited. This might be due to land use restrictions or social acceptance considerations.
    When constraining the capacity $\capacity$ of an arbitrary asset $i$ ($i$ may now be any generator, line etc.) to an upper limit $\capacityupper$, in the form of
    \begin{align}
        \capacity - \capacityupper \le 0 \resultsin{\muuppernom}
        \label{eq:capacityexpansionmaximum},
    \end{align}
    the zero profit condition alters as soon as the constraint becomes binding. Then, asset $i$ is payed an additional scarcity rent
    \begin{align}
        \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
        \label{eq:scarcitycost}
    \end{align}
    This rent may account for different possible realms, as for example the increased market price in higher competed areas or additional costs for social or environmental compensation. The share in $\allocatecost$ which consumers pay for the scarcity rent can be recalculated by a correct weighting of the shadow price $\muuppergenerationnom$ with the capital price $c_i$, leading to
    \begin{align}
        \allocatescarcitycost = \dfrac{\muuppernom}{c_i + \muuppernom} \, \allocatecost \Forall{i}
    \end{align}
\end{subequations}

\subsection{Lower Capacity Expansion Limits}
\label{sec:lower_capacity_limits}

\begin{subequations}
    In order to take already built infrastructure into account, the capacity $\capacity$ of an arbitrary asset $i$ may be constrained by a minimum required capacity $\capacitylower_i$. This introduces a constraint of the form
    \begin{align}
        \capacitylower - \capacity  \le 0 \resultsin{\mulowernom}
        \label{eq:capacityexpansionminimum}
    \end{align}
    Again, such a setup alters the zero profit condition of asset $i$, as soon as the constraint becomes binding.
    In that case, asset $i$ does not collect enough revenue from $\allocatecost$ in order to compensate the \ac{CAPEX}. The difference, given by
    \begin{align}
        \subsidycost_i = \mulowernom \capacity
        \Forall{i}
    \end{align}
    has to be subsidized by governments or communities or is simply ignored when investments are amortized. Allocating the subsidies to consumers is ambiguous as revenue of the assets are not in equilibrium with their \ac{CAPEX} and \ac{OPEX} anymore.
\end{subequations}

\subsection{Proof: Equivalence of local and imported prices}
\label{sec:proof_equivalence}

When inserting
\begin{align}
    \allocateflow = \sum_m \ptdf[m] (\allocatepeer - \delta_{nm} \demand)
    \label{eq:allocateflow2}
\end{align}
into
\begin{align}
    \lmp\, \demand & = \sum_m \lmp[m] \allocatepeer + \sum_\ell \lmp[\ell] \allocateflow \Forall{n,t}
    \label{eq:demand-cost-allocation2}
\end{align}
where
\begin{align}
    \lmp[\ell] = - \sum_m \incidence[m] \lmp[m]  + \sum_c  \cycleprice \cycle \reactance
\end{align}
the second expression on the right-hand side of \cref{eq:demand-cost-allocation2} becomes

\begin{align}
    \notag
      & - \sum_{\ell, m'} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \sum_m \incidence[m] \lmp[m]          \\
    \notag
      & + \sum_{\ell, m'} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \sum_c  \cycleprice \cycle \reactance \\
    \notag
    = & - \sum_{\ell, m', m} \ptdf[m'] \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \incidence[m] \lmp[m]              \\
    \notag
    = & - \sum_{m', m} \delta_{m m'} \left( \allocatepeer[m' \rightarrow n]  - \delta_{n m'} \demand \right) \lmp[m]                              \\
    \notag
    = & - \sum_{m} \left( \allocatepeer - \delta_{n m} \demand \right) \lmp[m]                                                                    \\
    = & - \sum_{m} \allocatepeer \lmp[m] + \demand \lmp
\end{align}
where in the third step we used the relation $\sum_\ell \ptdf \incidence[m] = \delta_{n m}$.
The second term in the first line vanishes as the basis cycles $\cycle$ are the kernel of the PTDF, $\sum_{\ell} \cycle \ptdf  = 0 \; \forall\, c, n$.

\begin{table*}[t]
    \begin{center}
        \begin{tabular}{l|c|c|c|c|c}
                               & Symbol             & Total Cost                                              & Allocation                                                                                                                               \\
            \toprule
            OPEX Production    & $\opexgeneration$  & $\sum_{t} \operationalpricegeneration \, \generation$   & $\operationalpricegeneration \allocategeneration $                                                                                       \\
            OPEX Storage       & $\opexstorage$     & $\sum_{t} \operationalpricestorage \, \storagedispatch$ & $\operationalpricestorage \allocatestoragedispatch $                                                                                     \\
            \midrule
            Emission Cost      & $\emissioncost$    & $ \emissionprice \, \emission \, \generation$           & $\emissionprice \,\emission \allocategeneration$                                                                                         \\
            \midrule
            CAPEX Production   & $\capexgeneration$ & $ \capitalpricegeneration \capacitygeneration$          & $\muuppergeneration \allocategeneration$                                                                                                 \\
            CAPEX Transmission & $\capexflow$       & $ \capitalpriceflow \capacityflow$                      & $\left(\muupperflow  \mulowerflow \right) \allocateflow$                                                                                 \\
            CAPEX Storage      & $\capexstorage$    & $ \capitalpricestorage \capacitystorage$                & $ \left(\muupperstoragedispatch  \mulowerstoragedispatch  + (\efficiencydispatch )^{1} \mustateofcharge\right) \allocatestoragedispatch$ \\
            % \bottomrule   
        \end{tabular}
    \end{center}
    \caption{Different cost terms and cost allocations of different assets, index $s$ refers to generators, $\ell$ to lines and $r$ to storage units. The definition of the cost weightings are defined in \cref{sec:zero_profit_generation,sec:zero_profit_flow,sec:zero_profit_storage_units}.}
    \label{tab:cost_allocation_map}
\end{table*}

\section{Power Allocation}
\label{sec:net_ap}

\newcommand{\incidenceM}{K}
\newcommand{\flowM}{f}
\newcommand{\injectionM}{p}
\newcommand{\slackM}{k}
\newcommand{\DirectedIncidence}{\mathcal{K}}
\newcommand{\InverseAPInjection}{\mathcal{J}}
\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}

Allocating net injections using the AP method is derived from \cite{achayuthakan_electricity_2010}. In a lossless network the downstream and upstream formulations result in the same P2P allocation which is why we restrict ourselves to the downstream formulation only. In a first step we define a time-dependent auxiliary matrix $\InverseAPInjection_t$ which is the inverse of the $N\times N$ with directed power flow $m \rightarrow n$ at entry $(m, n)$ for $m \ne n$ and the total flow passing node $m$ at entry $\left( m, m\right)$ at time step $t$. Mathematically this translates to

\begin{align}
    \InverseAPInjection_t = \left( \diag{\injectionM^+} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1}
\end{align}
where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( f_\ell \right)  \incidence$. Then the P2P allocation for time step $t$ is given by
\begin{align}
    \allocatepeer = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
\end{align}

\newpage
\section{Application Case}
\label{sec:application_case_appendix}

\begin{table*}[h]
    \centering
    \input{tables/de50/prices}
    \caption{Operational and capital price assumptions for all type of assets used in the working example. The capital price for transmission lines are given in [k\,\euro/MW/km]. The cost assumptions are retrieved from the PyPSA-EUR model \cite{horsch_jonas_pypsa-eur_2020}.}
    \label{tab:cost_assumptions}
\end{table*}

The following figures contain more detailed information about the peer-to-peer cost allocation discussed in \cref{sec:application_case}. The cost or prices payed by consumers are indicated by the region color. The dedicated revenue is displayed in proportion to the size of circles (for assets attached to buses) or to the thickness of transmission branches.

\begin{figure}[h]
    \centering
    \includegraphics[width=.5\linewidth]{de50/average_demand}
    \caption{Average demand, $\sum_t \demand/T$ per regions. The regions with high population densities and larger areas reveal a higher demand.}
    \label{fig:average_demand}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{de50/power_mix}
    \caption{Average power mix per region calculated by Average Participation. Coastal regions are mainly supplied by local offshore and onshore wind farms. Their strong power injections additionally penetrate the network up to the southern border. In the middle and South, the supply is dominated by a combination of OCGT and solar power.}
    \label{fig:power_mix}
\end{figure}

\begin{figure*}
    \centering
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50/maps_price/one_port_investment_cost}
        \subcaption{All production and storage technologies}
        \label{fig:total_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50/maps_price/by_carrier/onwind_one_port_investment_cost}
        \subcaption{Onshore Wind}
        \label{fig:onshore_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50/maps_price/by_carrier/solar_one_port_investment_cost}
        \subcaption{Solar}
        \label{fig:solar_capex}
    \end{subfigure}
    \begin{subfigure}[c]{.49\linewidth}
        \includegraphics[width=\linewidth]{de50/maps_price/by_carrier/OCGT_one_port_investment_cost}
        \subcaption{OCGT}
        \label{fig:ocgt_capex}
    \end{subfigure}
    \caption{Average \ac{CAPEX} allocation per MWh, $\sum_t \allocatecapex / \sum_t \demand$  for all production and storage assets (a), onshore wind (b), solar (c) and OCGT (d). Average allocated CAPEX per MWh within the regions are indicated by the color, the revenue per production asset is given by the size of the circles at the corresponding bus.}
    \label{fig:capex_price}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{de50/maps_scarcity_price/branch_scarcity_cost_total}
    \caption{Average allocated transmission scarcity cost per consumed MWh, $\sum_t \scarcitycost_{n \rightarrow \ell, t} / \sum_t \demand $. This scarcity cost results from the upper transmission expansion limit of 25\%. The costs are indicated by the regional color.  The lines are drawn in proportion to revenue dedicated to scarcity cost. }
    \label{fig:branch_scarcity_price}
\end{figure}

\begin{figure}
    \centering
    \vspace{2cm}
    \includegraphics[width=.5\linewidth]{de50/subsidy}
    \caption{Total costs for subsidy $\subsidycost$ resulting from lower capacity expansion bounds (brownfield constraints). The figure shows the built infrastructure that does not gain back its CAPEX from its market revenue, but is only built due to lower capacity limits.}
    \label{fig:subsidy}
\end{figure}

% \begin{figure*}
%     \includegraphics[width=\linewidth]{de50bf/capex_duration_curve}
%     \caption{Duration curve of the CAPEX allocation for production and storage technologies. Hours are sorted by the total amount of allocated expenditure. With 2.7 bn~\euro\, the first value pushes investments extraordinarily high. Due to low renewable potentials, it is dominated by CAPEX for OCGT which receives 92\% of the payments. This hour alone occasions about the half of all OCGT CAPEX. \Cref{fig:capex_duration_curve} gives a detailed picture of the operational state at this time-step. The following 7000 time-steps are dominated by revenues for onshore wind and reveal a rather even distribution. In hours of low CAPEX allocation (after the second drop) spending for OCGT start to increase again. These time-steps however play a minor role.}
%     \label{fig:capex_duration_curve}
% \end{figure*}

% \begin{figure}
%     \includegraphics[width=\linewidth]{de50bf/operation_high_expenditure}
%     \caption{Production, flow and consumption in the system at the hour with the highest allocated expenditures. The size of the circles are proportional to the power production at a node. Size of arrows are proportional to the flow on the transmission line. The depicted hour corresponds to the first value in the duration curve in \cref{fig:capex_duration_curve}.}
%     \label{fig:operation_high_expenditure}
% \end{figure}

% \begin{figure*}
%     \centering
%     \begin{subfigure}[c]{.49\linewidth}
%         \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/offwind-dc_generator_scarcity_cost}
%         \subcaption{Offshore Wind}
%         \label{fig:offwind-dc_generator_scarcity_cost}
%     \end{subfigure}
%     \begin{subfigure}[c]{.49\linewidth}
%         \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/onwind_generator_scarcity_cost}
%         \subcaption{Onshore Wind}
%         \label{fig:onwind_generator_scarcity_cost}
%     \end{subfigure}
%     \begin{subfigure}[c]{.49\linewidth}
%         \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/solar_generator_scarcity_cost}
%         \subcaption{Solar}
%         \label{fig:solar_generator_scarcity_cost}
%     \end{subfigure}
%     \begin{subfigure}[c]{.49\linewidth}
%         \includegraphics[width=\linewidth]{de50bf/maps_scarcity_price_ptpf_net/by_carrier/ror_generator_scarcity_cost}
%         \subcaption{Run-of-River}
%         \label{fig:ror_generator_scarcity_cost}
%     \end{subfigure}
%     \caption{Average \textbf{allocated scarcity cost} per consumed MWh, $\sum_t \allocatescarcitycost / \sum_t \demand$. These cost result from land use restrictions for offshore wind, onshore wind,  solar, run-of-river. The cost per MWh are indicated by the color of a region. The revenue per production asset is given by the size of the circle at the corresponding bus.}
%     \label{fig:scarcity_price}
% \end{figure*}

\clearpage
\bibliographystyle{elsarticle-harv}
\bibliography{main.bib}

\end{document}

### 2022-03-16 submission-energy/paper/main.tex
finish submission to energy

@@
     \begin{highlights}
+        \item Optimal Locational Marginal Prices are a direct result of a power system optimization. 
         \item Flow Tracing is used to interlink optimal Locational Marginal Prices.
         \item This quantifies the payment of individual consumers to individual generators and transmission lines.
-        \item Four use-cases are proposed and demonstrated on behalf of a future model of the German power system.
+        \item Four use-cases for improving transparency and fairness in the electricity market are proposed and demonstrated on behalf of a future model of the German power system.
     \end{highlights}
@@
     \begin{keyword}
-        keywords here, in the form: keyword \sep keyword
-
-        %% PACS codes here, in the form: \PACS code \sep code
-
-        %% MSC codes here, in the form: \MSC code \sep code
-        %% or \MSC[2008] code \sep code (2000 is the default)
+
+        Power System Optimization, Shadow Prices, Cost Allocation, Electricity Market, Power Flow Tracing
 
